### 2021-03-22 papers/icse21-src/main.tex
Initial commit to papers branch

\documentclass[10pt,conference]{IEEEtran}

\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote.
% If that is unneeded, please comment it out.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Packages

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage[dvipsnames]{xcolor}
\usepackage{listings}
\usepackage{xspace}
\usepackage{multirow}
\usepackage{float}
\usepackage{flushend}
%% \usepackage{caption}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Macros and definitions
\linepenalty=1000
\widowpenalty10000
\clubpenalty10000

\lstset{
  frame=none,
  stepnumber=1,
  belowcaptionskip=\bigskipamount,
  captionpos=b,
  escapeinside={*'}{'*},
  language=haskell,
  tabsize=2,
  emphstyle={\bf},
  commentstyle=\it,
  stringstyle=\mdseries\ttfamily,
  showspaces=false,
  keywordstyle=\ttfamily,
  columns=flexible,
  basicstyle=\footnotesize\ttfamily,
  showstringspaces=false,
  morecomment=[l]\%,
}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\setlength{\tabcolsep}{0.5em}

\newcommand{\ourtool}{\textsc{Mutagen}\xspace}

%% \captionsetup[figure]{labelfont={bf},name={Table},labelsep=period}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Document

\begin{document}

\title{\ourtool: Faster Mutation-Based Random Testing}

\author{
  \IEEEauthorblockN{Agust\'in Mista}
  \IEEEauthorblockA{
    %% \textit{Department of Computer Science} \\
    \textit{Chalmers University of Technology}\\
    Gothenburg, Sweden \\
    \texttt{mista@chalmers.se}
  }
}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Abstract

\begin{abstract}
We present \ourtool, a fully automated mutation-oriented framework for
property-based testing.
%
Our tool uses novel heuristics to improve the performance of the testing loop,
and it is capable of finding complex bugs within seconds.
%
We evaluate \ourtool by generating random WebAssembly programs that we use to
find bugs in a faulty validator.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Keywords

\begin{IEEEkeywords}
random testing, mutation, heuristics
\end{IEEEkeywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Sections

\section{Introduction}

% Randomly generated values
%
Using randomly generated inputs is a popular and powerful approach when it comes
to testing software \cite{duran1984evaluation}.
%
While plenty, the tools designed for this purpose can be divided into two main
categories: those that use an existing corpus of inputs, and those that generate
such inputs from scratch.

% Corpus-based approaches
%
On one hand, corpus-based tools create new inputs by combining and mutating
seeds from their input corpora using several heuristics
\cite{zalewski2014american, swiecki2017honggfuzz}.
%
Being mostly black-box, this approach is often limited by the lack of knowledge
about the structure of its inputs, although there exist efforts to improve this
situation, e.g., by using grammars describing the syntactic structure of the
generated data while combining seeds in order to produce syntactically valid
test cases \cite{miller2007analysis, wang2019superion, holler2012fuzzing}.

% Generational approaches
%
On the other hand, generational approaches can circumvent this limitation by
generating valid data from scratch using specialized random generators
\cite{eddington2011peach, dharma, grieco2016quickfuzz, grieco2017quickfuzz}.
%
However, writing good random generators by hand is a demanding task that
involves several iterations of trial and error and can take hundreds of hours
\cite{lampropoulos2019coverage}.
%
While there exist tools that tackle this problem by automatically synthesizing
(with varying degrees of complexity) random generators directly from the static
information present in the codebase \cite{mista2018branching,
  mista2019generating, duregaard2012feat, lampropoulos2017generating}, such
approaches are unable to derive suitable generators when the target data
involves complex invariants like those required to generate random programs,
e.g., well-scopedness and well-typedness.
%
In such cases, generators obtained by automatic tools are extremely unlikely to
produce random data with enough structure to penetrate deep layers of our
systems before being discarded.
%
% early as completely invalid.

%% % Random property-based testing
%% %
%% Random property-based testing is a powerful technique for finding bugs.
%% %
%% The expected behavior of a system under test is expressed in terms of
%% a collection of properties that can be executed when given certain inputs.
%% %
%% Then, automated tools can be used for repeteadly probing these properties using
%% randomly generated inputs until any of them gets falsified.

%% % QuickCheck and random generators
%% %
%% In the Haskell realm, QuickCheck is the dominant tool of this sort
%% \cite{hall1996type}.
%% %
%% As well as most other property-based frameworks, QuickCheck provides default
%% random generators for the base types, along with with a library of combinators
%% for writing random generators for any other used-defined algebraic data type
%% (ADT) -- a task that lies on the developers' shoulders.

%% % Automated derivation tools
%% %
%% For simple ADTs, writing custom random generators manually is often just a
%% cumbersome task.
%% %
%% However, as the complexity of the input data grows, developing random generators
%% capable of generating \emph{good} random values can be extremely tricky.
%% %
%% Fortunately, there exists tools that alleviate this problem by automatically
%% synthesizing (with varying degrees of complexity) random data generators
%% directly from the static information present in the codebase.
%% %
%% Most of these tools are solely focused on the ADT definition of the data to be
%% generated, which is often enough to ensure that the derived generators always
%% produce finite well-typed values that can be use for finding bugs in practice.

%% % Limitation
%% %
%% Despite these automated efforts, most of these tools are unable to synthesize
%% random generators capable of producing data satisfying complex invariants.
%% %
%% Perhaps the most clear example of this issue arises while trying to generate
%% random code.
%% %
%% In most host programming languages, invariants like well-scopedness and
%% well-typedness are quite difficult to express directly at the type level.
%% %
%% Thus, any automated mechanism that synthesizes random generators by only looking
%% at the type definition of the generated data will fail to provide acceptable
%% results in practice.
%% %
%% In this scenario, developers are often forced to spend countless hours of trial
%% and error while developing their random generators by hand.

% Mutation-based approach
%
%% Considering that interesting test cases are very unlikely to be produced from
%% scratch when using automatically derived generators,
%
%% other random testing frameworks like FuzzChick (implemented for the Coq
%% programming language) instead focus their efforts on those that actually are.

Recently, Lampropoulos et al. introduced FuzzChick
\cite{lampropoulos2019coverage}, a property-based testing framework for the Coq
programming language that borrows ideas from the fuzzing community to generate
highly structured values while using automatically derived generators.
%
Instead of continuously generating random invalid test cases from scratch,
FuzzChick keeps a queue of interesting previously executed test cases that can
be mutated using type-preserving transformations in order to produce new ones.
%
The logic behind this is simple, mutating an interesting test case in a
small way (at the data constructor level) has a much higher probability of
producing a new interesting test case than generating a new one from scratch
using a na\"ive generator.
%
FuzzChick is likely to preserve the semantic structure of the mutated data, as
mutations are applied directly at the data type level -- the random AST in case
of generating code.

In order to work in practice, FuzzChick relies on execution traces to
distinguish which executed test cases were interesting and are therefore worth
mutating.
%
%% For this purpose, FuzzChick captures the execution trace of the system under
%% test and compares it with the ones already executed.
%
Mutated test cases are considered interesting for mutation only when they
produce new execution traces -- any other test case is simply discarded.
%
%% %
%% This coverage-based guidance resembles that used by popular software fuzzers
%% like AFL or Honggfuzz, and it helps focusing the mutation mechanism towards a
%% wider variety of tests cases.
%
%% Moreover, it avoids putting effort in undesired cases like continuously mutating
%% irrelevant subexpressions of a test case, or triggering cyclic mutations loops.
%
While powerful, the implementation of FuzzChick is relatively simple, leaving
the door open for future extensions.

In this work we present \ourtool, a random property-based testing framework
implemented in the strongly typed language Haskell that follows the mutational
approach behind FuzzChick, extending it with novel mutation heuristics designed
to converge to counterexamples in fewer tests.
%
We outline these heuristics in the next section and evaluate them in Section
\ref{sec:casestudy}.

%%%%%%%%%%%%%%%%%%%%

\begin{figure*}[t]
  \renewcommand{\figurename}{Table}
  \vspace{-10pt}
  \scriptsize
  \centering
    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|}
      \hline
      \multirow{2}{*}{Bug} &
      \multicolumn{3}{c|}{QuickCheck} &
      \multicolumn{3}{c|}{\ourtool} &
      \multicolumn{3}{c|}{\ourtool (no inheritance)} &
      \multicolumn{3}{c|}{\ourtool (no scheduling)}  \\
      \cline{2-13} &
      Time (s) & Passed & Discarded &
      Time (s) & Passed & Discarded &
      Time (s) & Passed & Discarded &
      Time (s) & Passed & Discarded \\
      \hline
      1 & N/A & 14.8 & 1000000 & 2.16 & 2194 & 11706.7 & 4.06 & 4079 & 15123.3 & 3659.45 & 101536.4 & 134107.1 \\
      \hline
      2 & N/A & 12.7 & 1000000 & 0.07 & 83.7  & 131.6 & 0.23 & 159.5 & 206.5 & 0.33 & 261.9 & 234 \\
      \hline
      3 & N/A & 14.9 & 1000000 & 0.03 & 56.7 & 91.8 & 0.04 & 51 & 71.9 & 0.28 & 217.2 & 107.1 \\
      \hline
      4 & N/A & 13.7 & 1000000 & 0.21 & 157.2 & 314.2 & 0.63 & 427.2 & 903.3 & 10.85 & 6740.6 & 7372.2 \\
      \hline
      5 & N/A & 14.3 & 1000000 & 6.45 & 6464 & 34856 & 0.82 & 560.1  & 1543.6 & 454.26 & 31577.9 & 43837.5 \\
      \hline
    \end{tabular}
    \vspace{-2pt}
    \caption{\label{fig:results} Time to first failure, passed and discarded
      tests accross different bugs for the faulty WebAssembly validator. Mean values computed after 10 executions.}
  \vspace{-15pt}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%

\section{\ourtool's Heuristics}
\label{sec:heuristics}

\noindent This section introduces some of \ourtool's novel heuristics.

\paragraph{Exhaustive Uniform Mutations}

Perhaps the biggest distinction between FuzzChick and \ourtool lies in how
mutations are applied over test cases.
%
Along with random data generators, FuzzChick automatically derives simple
type-preserving mutators.
%
These mutators work in a simple recursive manner: with uniform probability, they
either mutate a node at the top level of the given value, or apply a mutation
recursively to one of its subexpressions.
%
This mechanism favors mutations to happen on the shallower levels of their
inputs, while deeper mutations are unlikely to happen due to the multiplicative
decline of their probability on each recursive call.

To find complex bugs, we believe that mutations should be able to occur deep
inside of the generated values in a reasonable proportion.
%
%% If our target is to generate data with a complex and deep structure, e.g.,
%% program's ASTs, we certainly want every deep subexpression of a test candidate
%% to be likely to be mutated at some point.
%%
To allow this, the mutators derived by our tool follow a different approach.
%
\emph{They first traverse the input, collecting the path to each mutable
  subexpression.}
%
Then, the testing loop schedules mutations targeted over each one of them,
ensuring that all of them are mutated in a timely manner.

Furthermore, for a given target subexpression, the testing loop produces and
tests every possible mutation exhaustively, reducing the reliance on randomness
in order to find bugs.
%
%% ensuring that if a test candidate is one mutation away
%% from discovering a bug, then the bug will be found deterministically.
%
This approach is inspired by exhaustive testing tools like SmallCheck
\cite{runciman2008smallcheck} or Korat \cite{boyapati2002korat}.
%
However, testing mutations exhaustively comes attached to a high testing cost
per test candidate, so the next heuristic is, in part, focused on reducing the
time complexity of the testing loop of our tool.

\paragraph{Mutation Inheritance}

The second mutation heuristic focuses on preserving the semantically important
subexpressions generated on previous steps.
%
For this purpose, child mutants keep a record of the subexpressions that were
already mutated by their ancestors.
%
This allows each mutant to focus on the previously untouched subexpressions, as
well as the ones freshly generated by its parent.

As a consequence, this heuristic greatly reduces the available positions where
mutations can occur on each new mutant.
%
This, in turn, helps to cope with the cost of running mutations exhaustively, as
described above.

\paragraph{Mutants Scheduling}

FuzzChick keeps a queue of mutation candidates obtained by analyzing the
execution trace of every executed mutant.
%
If a candidate executes a completely new branch in the code, it is inserted at
the end of the queue, and the testing loop will first have to process every
candidate ahead of it before it can start mutating this new (and likely more
interesting) test case.

To account for this, our tool uses a preemptive (FIFO) schedule with priority for
scheduling mutant candidates.
%
In this setting, we capture the depth at which each new execution trace differs
from all the previous ones.
%
This depth is later used to give more priority to those candidates that
``discover'' new parts of the code at earlier stages, favoring a wider
traversal of the execution trace space in less time.

%%%%%%%%%%%%%%%%%%%%

\section{Case Study: WebAssembly Validator}
\label{sec:casestudy}

This section evaluates how the heuristics implemented in our \ourtool affect the
testing performance.
%
For this purpose, we use the validator of the WebAssembly programming language \cite{haas2017bringing}
as a case study.
%
%% WebAssembly is recent a stack-based low level programming language that can run
%% natively on most web browsers, and that is proposed as a compilation target for
%% a variety of higher level languages.
%% %
%% In practice, WebAssembly modules need to be validated before they can be used
%% --- a process similar to type checking in other languages.
%
We took an exiting Haskell implementation of the WebAssembly validator
\cite{wasm-lib} and injected several bugs into it.
%
These bugs let invalid WebAssembly modules pass the validation process.
%
Then, we tested for false positives of the faulty validator by comparing it
against the output of the validator from the WebAssembly reference
implementation (written in OCaml).
%
This was expressed using the following property:

\vspace{-3pt}
\begin{lstlisting}
prop_valid :: WasmModule -> Result
prop_valid wasm = validHs wasm ==> validOCaml wasm
\end{lstlisting}
\vspace{-3pt}

Using {\small\texttt{validHs}} as a precondition, we avoid executing the
reference validator on invalid modules.
%
Instead, they get automatically discarded and the testing loop continues.

Table \ref{fig:results} shows the performance of our tool against QuickCheck
\cite{claessen2011quickcheck} (the reference tool for random testing in Haskell)
across five different bugs, using the same automatically derived random
generator.
%
To evaluate the effectiveness of our inheritance and scheduling heuristics, we
included the results obtained when they are disabled.

Unsurprisingly, QuickCheck fails to find any of the bugs we planted after
generating more than a million values.
%
This is because the automatically derived generator is virtually unable to
produce valid modules on its own.
%
In comparison, \ourtool is able to find all the bugs in a couple of seconds.
%
Guided by the execution traces, the type-preserving mutations that our tool
applies allows us to run a much larger number of valid WebAssembly modules through
the reference validator, which in turn helps finding bugs faster.
%
As an example of the level of invariants required to find bugs, the following
WebAssembly module AST is a counterexample for bug \#1:

\vspace{-3pt}
\begin{lstlisting}
Module {
  types=*'\textcolor{red}{[}'*FuncType{params=[], results=*'\textcolor{blue}{[]}'*}*'\textcolor{red}{]}'*,
  functions=[
    Function{resultType=*'\textcolor{red}{0}'*, localTypes=[], body=[
      I32Const 0,
      If{resultType=*'\textcolor{blue}{[]}'*, then=*'\textcolor{blue}{[]}'*, else=[I32Const 0]}]]},
  tables=[], mems=[], globals=[], elems=[],
  datas=[], start=Nothing, imports=[], exports=[],
}
\end{lstlisting}
\vspace{-3pt}

\noindent This bug causes false validations by not checking that the type of an
\textbf{else} branch in an \textbf{if} expression matches that of the
\textbf{then} one.
%
%% when the type of an \textbf{else}
%% branch of an \textbf{if} expression does not match the \textbf{then} one ---
%% it can be observed in the example above.
%
%% Despite the actual cause for this counterexample to trigger the bug is quite
%% simple (\textbf{if} branches with different types),
%
Despite this, several other things need to be in place for this value to be
considered valid: the type of the generated function must be declared
beforehand, and the function must refer to it by its index in the types list
(marked in red).
%
Moreover, the type of its actual outputs must match the one declared by its type
(marked in blue).
%
Generating valid modules satisfying such invariants is extremely rare using
automatically derived generators.\looseness=-1

If we look at the effect of our heuristics, we can observe that mutation
inheritance can provide substantial speedups in most cases.
%
%% (e.g., bugs \#1, \#3, and \#4).
%
In the case of bug \#5, however, disabling mutation inheritance seems to
help as a shortcut to find counterexamples faster.
%
This suggests that perhaps a hybrid approach would work best in the general case.

Furthermore, our preemptive scheduling heuristic greatly improves the testing
performance in most cases, being orders of magnitude faster in the case of
bugs \#1, \#4, and \#5.

%% At the light of these results,
The heuristics we developed for \ourtool offer a substantial improvement over
the simple (though still powerful) approach taken by FuzzChick.
%
Our current work focuses on gathering more empirical evidence to strengthen this
claim.

%%%%%%%%%%%%%%%%%%%%

\section{Future Work}

%% We presented \ourtool, an automated framework for property-based testing in
%% Haskell focused on generating complex test cases by progressively applying
%% type-preserving mutations over previously generated ones.
%% %
%% Our tool follows a testing approach similar to that of FuzzChick, extending it
%% with heuristics to improve the performance in terms of convergence speed to
%% counterexamples and reproductibility.

%% \ourtool is guided by the execution trace of the mutated test cases, and favours
%% mutating the ones that discover new parts of the code on each step.
%% %
%% Compared to just using automatically derived generators in a purely random
%% fashion, our tool is capable of finding bugs triggered only by inputs satisfying
%% complex invariants that are extremely unlikey to be produced by such generators
%% alone.

%% In the future, we are focused on extending \ourtool with the ability of producing
%% semantic-preserving paired mutations.
%% %
%% This would entail finding related subexpressions within mutated test cases that
%% need to be handled in tandem in order to preserve their semantic correctness.
%% %
%% With this information at hand, we foresee that it could be possible to train our
%% mutation mechanism so these hard-to-find semantic relations are preserved
%% throughout the testing process.

As for future work, we are focused on extending \ourtool with dependent mutations,
where the existence of certain subexpressions would allow or disallow certain
mutations to occur.
%
Concretely, this could be useful to improve the performance of our tool when
used for testing systems expecting programs as inputs.
%
There, for instance, a mutation that introduces a new identifier would allow
subsequent mutations to refer to it, preserving the well-scopedness of the
program.
%% %
%% We foresee that this extension would require external knowdledge from the
%% developer in order to recognize which kinds of subexpression define identifiers,
%% and which other kinds can use them.
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Bibliography

\bibliography{references}{}
\bibliographystyle{IEEEtran}

\end{document}

### 2021-03-22 papers/oopsla-21/main.tex
Initial commit to papers branch

%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[acmsmall, anonymous]{acmart}

% ----------------------------------------

%% Packages

\usepackage{xcolor}
\usepackage{color}
\usepackage{multicol}
\usepackage[ruled,vlined,noend]{algorithm2e}
\usepackage{xspace}
\usepackage{tikz}
\usepackage{tikzscale}
\usepackage{pgfplots}
\usepackage{paralist}
\pgfplotsset{compat=1.3}

\makeatletter
\patchcmd{\@verbatim}
  {\verbatim@font}
  {\verbatim@font\footnotesize}
  {}{}
\makeatother

\let\oldv\verbatim
\let\oldendv\endverbatim

\def\verbatim{\par\setbox0\vbox\bgroup\oldv}
\def\endverbatim{\oldendv\egroup\fboxsep0pt \noindent\colorbox[gray]{0.95}{\usebox0}\par}

% ----------------------------------------

%% Rights management information. This information is sent to you when you
%% complete the rights form. These commands have SAMPLE values in them; it is
%% your responsibility as an author to replace the commands and values with
%% those provided to you when you complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2021}
\acmYear{2021}
%% \acmDOI{10.1145/1122445.1122456}

%% These commands are for a JOURNAL article.
\acmConference[OOPSLA '21]{OOPSLA '21: ACM SIGPLAN conference on Systems,
  Programming, Languages, and Applications}{October 17--22 2021}{Chicago, IL}

%% \acmJournal{JACM}
%% \acmVolume{37}
%% \acmNumber{4}
%% \acmArticle{111}
%% \acmMonth{8}

%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
\citestyle{acmauthoryear}

% ----------------------------------------
%% Macros

\newcommand{\quickcheck}{\textit{QuickCheck}\xspace}
\newcommand{\quickchick}{\textit{QuickChick}\xspace}
\newcommand{\fuzzchick}{\textit{FuzzChick}\xspace}
\newcommand{\mutagen}{\textsc{Mutagen}\xspace}

% ----------------------------------------

%% end of the preamble, start of the body of the document source.

\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{\mutagen: Working title}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.

\author{Agust\'in Mista}
\email{mista@chalmers.se}
\affiliation{%
  \institution{Chalmers University of Technology}
  \city{Gothenburg}
  \country{Sweden}
}

\author{Alejandro Russo}
\email{russo@chalmers.se}
\affiliation{%
  \institution{Chalmers University of Technology}
  \city{Gothenburg}
  \country{Sweden}
}

%% By default, the full list of authors will be used in the page headers. Often,
%% this list is too long, and will overlap other information printed in the page
%% headers. This command allows the author to define a more concise list of
%% authors' names for this purpose.

%% \renewcommand{\shortauthors}{Mista and Russo}

% ----------------------------------------
%% The abstract is a short summary of the work to be presented in the article.

\begin{abstract}
  Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
  facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc
  porta vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam
  posuere. Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada
  massa, quis varius mi purus non odio. Pellentesque condimentum, magna ut
  suscipit hendrerit, ipsum augue ornare nulla, non luctus diam neque sit amet
  urna. Curabitur vulputate vestibulum lorem. Fusce sagittis, libero non
  molestie mollis, magna orci ultrices dolor, at vulputate neque nulla lacinia
  eros. Sed id ligula quis est convallis tempor. Curabitur lacinia pulvinar
  nibh. Nam a sapien.
\end{abstract}

% ----------------------------------------

%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm. Please
%% copy and paste the code instead of the example below.
\begin{CCSXML}
% CCS XML here
\end{CCSXML}

%% \ccsdesc[500]{Computer systems organization~Embedded systems}
%% \ccsdesc[300]{Computer systems organization~Redundancy}
%% \ccsdesc{Computer systems organization~Robotics}
%% \ccsdesc[100]{Networks~Network reliability}

% ----------------------------------------

%% Keywords. The author(s) should pick words that accurately describe the work
%% being presented. Separate the keywords with commas.
\keywords{random testing, mutations, heuristics}

% ----------------------------------------

%% This command processes the author and affiliation and title information and
%% builds the first part of the formatted document.
\maketitle

% ----------------------------------------

\section{Introduction}
\label{sec:intro}

Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.

The contributions of this work are:

\begin{itemize}
\item Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec
  facilisis facilisis, est dui fermentum leo, quis tempor ligula erat quis odio.
  Nunc porta vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum.
  Aliquam posuere. Nunc aliquet, augue nec adipiscing interdum, lacus tellus
  malesuada massa, quis varius mi purus non odio.
\item Pellentesque condimentum,
  magna ut suscipit hendrerit, ipsum augue ornare nulla, non luctus diam neque
  sit amet urna. Curabitur vulputate vestibulum lorem. Fusce sagittis, libero
  non molestie mollis, magna orci ultrices dolor, at vulputate neque nulla
  lacinia eros. Sed id ligula quis est convallis tempor. Curabitur lacinia
  pulvinar nibh. Nam a sapien.
\item Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec
  facilisis facilisis, est dui fermentum leo, quis tempor ligula erat quis odio.
  Nunc porta vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum.
  Aliquam posuere. Nunc aliquet, augue nec adipiscing interdum, lacus tellus
  malesuada massa, quis varius mi purus non odio.
\item Pellentesque condimentum, magna ut suscipit hendrerit, ipsum augue ornare
  nulla, non luctus diam neque sit amet urna. Curabitur vulputate vestibulum
  lorem. Fusce sagittis, libero non molestie mollis, magna orci ultrices dolor,
  at vulputate neque nulla lacinia eros. Sed id ligula quis est convallis
  tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
\end{itemize}

\section{Background}
\label{sec:background}

In this section we briefly introduce the concept of Property-based Testing along
with \quickcheck, the most popular tool of this sort and often considered as the
baseline when comparing PBT algoritms.
%
Moreover, we describe the ideas and limitations behind \fuzzchick, which
enhances the PBT approach with execution information and serves as the
foundation for \mutagen.
%
\fuzzchick is implemented as an extension of \quickchick, Coq's own
reimplementation of \quickcheck.

\subsection{Property-Based Testing and \quickcheck}

Property-based testing is a powerful technique for finding bugs without having
to write test cases by hand.
%
Originally introduced by \citet{quickcheck} in tandem with the first version of
\quickcheck, this technique focuses on aiming the developer's efforts into
testing systems via executable specifications using randomly generated inputs.
%
Moreover, tools like \quickchick and Isabelle's \quickcheck demonstrate
that PBT can also be used in the formal verification realm.
%
There, one can quickly spot bugs in system specifications before directing the
efforts into pointlessly trying to prove bogus propositions.

In the simplest form, there are four main elements the user needs to provide in
order to perform property-based testing on their systems:
%
\begin{itemize}
\item one or more \emph{executable properties}, often implemented simply as a
  boolean predicates,
\item \emph{random data generators}, used to repeatedly instantiate the testing
  properties,
\item \emph{printers}, used to show the user the random inputs that falsify some
  testing property (the counterexample) whenever a bug is found, and
\item \emph{shrinkers}, to minimize counterexamples making them easier to
  understand by humans.
\end{itemize}

\noindent In this work we focus solely on the first two elements introduced
above, namely the testing properties and the random data generators used to feed
them.
%
Printers and shrinkers, for the most part, can be obtained automatically using
generic programming capabilities present in the compiler, and although being
crucial for the testing process as a whole, their role becomes irrelevant when
it comes to \emph{finding} bugs.

\begin{algorithm}[ b]
  \SetKw{KwNot}{not}
  \SetAlgoLined
  \DontPrintSemicolon
  \SetKwFunction{QC}{Loop}
  \SetKwProg{Fn}{Function}{:}{}
  \Fn{\QC{P, N, gen}}{
    i $\gets$ 0\;
    \While{i $<$ N}{
      x $\gets$ Sample(gen)\;
      \lIf{\KwNot P(x)}{\KwRet Bug(x)}
      i $\gets$ i+1 \;
    }
    \KwRet Ok\;
  }
\caption{\label{algo:quickcheck}\quickcheck Testing Loop}
\end{algorithm}

Perhaps the simplest PBT technique is to repeatedly generate random inputs and
instantiate the testing properties until they either get falsified by a
counterexample, or we ran a sufficiently large amount of tests --- suggesting
that the properties holds.
%
\quickcheck implements a testing loop that closely follows this simple idea,
which is outlined in Algorithm \ref{algo:quickcheck}, where $P$ is the testing
property, $N$ is the maximum number of tests to perform, and $gen$ is the random
generator to be used to instantiate $P$.

To illustrate this technique, let us focus on the same motivating example used
by \citeauthor{lampropoulos2019coverage}, who propose a simple property defined over
binary trees.
%
Such data structure can be easily defined in Haskell using a custom data type
with two data constructors for leaves and branches respectively:

\begin{verbatim}
data Tree a = Leaf a | Branch (Tree a) a (Tree a)
\end{verbatim}

\noindent The type parameter \texttt{a} indicates that trees can be instantiated
using any other type as payload, so the value \texttt{Leaf Bool} has type
\texttt{Tree Bool}, whereas the value \texttt{Branch (Leaf 1) 2 (Leaf 3)} has
type \texttt{Tree Int}.
%
Then, we can define tree reflections using a simple recursive function that
pattern matches against the two possible constructors, inverting the order of
the subtrees whenever it encounters a branch:

\begin{verbatim}
mirror :: Tree a -> Tree a
mirror (Leaf x)       = Leaf x
mirror (Branch l x r) = Branch (mirror r) x (mirror l)
\end{verbatim}

Later, a reasonable requirement to assert for is that \texttt{mirror} must be
\emph{involutive}, i.e., reflecting a tree twice always yields the original
tree.
%
We can simply capture this property using a boolean predicate written as a
normal function:

\begin{verbatim}
prop_mirror :: Tree Int -> Bool
prop_mirror t = mirror (mirror t) == t
\end{verbatim}

\noindent For simplicity, here we instantiate the tree payload with integers,
altough this predicate should clearly hold for any other type with a properly
defined notion of equality as well.

With our simple specification in place, the last missing piece is a random
generator of trees.
%
In \quickcheck, this is usually done via the type class mechanism, instantiating
the \texttt{Arbitrary} type class, providing a random generator as the
implementation of the overloaded \texttt{arbitrary} operation:

\begin{verbatim}
instance Arbitrary a => Arbitrary (Tree a) where
  arbitrary = sized gen
    where
      gen 0 = do { x <- arbitrary; return (Leaf x) }
      gen n = oneof [ do {x <- arbitrary; return (Leaf x) }
                    , do {l <- gen (n-1); x <- arbitrary; r <- gen (n-1); return (Branch l x r)} ]
\end{verbatim}

\noindent Let us break this definition into parts.
%
The first line states that we will provide an \texttt{Arbitrary} instance for
trees with payload of type \texttt{a}, provided that values of type \texttt{a}
can also be randomly generated.
%
This allows us to use \texttt{arbitrary} to generate \texttt{a}'s inside the
definition of our tree generator.

Moreover, \quickcheck internally keeps track of the \emph{maximum generation
  size}, a parameter that can be tuned by the user in order to limit the size of
the randomly generated values.
%
Our definition exposes this internal parameter via \quickcheck's \texttt{sized}
combinator, allowing us to parameterize the maximum size of the randomly
generated trees.
%
If the generation size is zero (\texttt{gen 0}), our generator is limited to
produce just leaves with randomly generated payloads.
%
In turn, when the generation size is strictly positive (\texttt{gen n}), the
generator is able to perform a random uniform choice between generating either a
single leaf or a branch.
%
When generating branches, the generator calls itself recursively in order to
produce random subtrees (\texttt{gen (n-1)}).
%
Notice the importance of reducing the generation size on each recursive call.
%
This way we ensure that randomly generated trees using a generation size
\texttt{n} are always finite and have at most \texttt{n} levels.

Finally, we are ready to let \quickcheck test \texttt{prop\_mirror} against a
large number of inputs (100 by default) produced by our brand new random tree
generator:

\begin{verbatim}
quickCheck prop_mirror
+++ Ok ####
\end{verbatim}

\noindent Should we mistakenly implement \texttt{mirror}, e.g., by dropping the
right subtree altogether:

\begin{verbatim}
mirror (Branch l x r) = Branch (mirror l) x (mirror l)
\end{verbatim}

\noindent then \quickcheck will quickly falsify \texttt{prop\_mirror}, reporting
a minimized counterexample that we can use to find the root of the issue:

\begin{verbatim}
quickCheck prop_mirror
*** Failed ####
\end{verbatim}

At this point, it is clear that the \emph{quality} of our random generators is
paramount to the performance of the overall PBT process.
%
Random generators that rarely produce interesting values will fail to trigger
bugs in our code, potentially leaving entire parts of the codebase virtually
untested.

Recalling our tree generator, the reader (far from mistaken) might already have
imagined better ways for implementing it.
%
For most practical purposes, this generator is in fact quite bad.
%
However, it follows a simple type-directed fashion, and it is a good example of
what to expect from a random generator synthesized automatically using a process
that knows very little about the values to be generated apart from their
(syntactic) data type structure.

As introduced earlier, there exist multiple tools that can automatically derive
better random generators solely from the static information present in the
codebase.
%
%% Such tools vary on the degree of invariants and requirements
%% %
%% However, such approaches are unable to derive useful generators whenever the
%% target data involves complex invariants that cannot be easily extracted from the
%% codebase or that cannot be easily expressed in terms of inductive relations.
%
Sadly, these tools lack the domain knowledge required to generate random data
with complex invariants --- especially those present in programming languages
like well-scopedness and well-typedeness.

In particular, automatically derived generators are remarkably uneffective when
used to test properties with sparse preconditions.
%
Let us continue with the example by \citeauthor{lampropoulos2019coverage} to
illustrate this problem in more detail.
%
For this, consider that we want to use our \texttt{Tree} data type to encode
binary-search trees (BST).
%
Then, given a predicate \texttt{isBST} that asserts if a tree satisfies the BST
invariants, we might want to use it as pre- and post-condition to assert that
BST operations like \texttt{insert} preserve them:

\begin{verbatim}
prop_bst_insert :: Tree a -> a -> Bool
prop_bst_insert t a =
  isBST t ==> isBST (insert a t)
\end{verbatim}

\noindent Attempting to test this property using \quickcheck does not work well:

\begin{verbatim}
quickCheck prop_bst_insert
*** Gave up! ####
\end{verbatim}

\noindent Here, \quickcheck discards random inputs as soon as it finds they do
not pass the precondition (\texttt{isBST t}).
%
Sadly, most of the inputs generated by our na\"ive generator suffer from this
problem, and the interesting part of the property (\texttt{isBST (insert a t)})
is tested very sporadically as a result.

At this point it is reasonable to think that, to obtain the best results when
using PBT over complex systems, one is forced to put a large amount of time on
developing manually-written generators.
%
In practice, that is most often the case, no automatic effort can beat a
well-thought manually-written generator that produces interesting complex values
and finds bugs in very few tests.
%
Not all is lost, however.
%
It is still possible to obtain acceptable results automatically by incorporating
dynamic information from the system under test into the testing loop.
%
The next subsection introduces the clever technique used by \fuzzchick to find
bugs in complex systems while using simple automatically derived random
generators.

%% Far from trying to compete against manual-efforts, our work focuses on improving
%% automatic approaches so they can be used paliatively in earlier development
%% stages.

\subsection{Coverage-Guided Property-Based Testing with \fuzzchick}

To alleviate the problem of testing properties with sparse preconditions while
using simple automatically derived random generators, \fuzzchick introduces
\emph{coverage-guided property-based testing} (CGPT) by enhancing the testing
process with two key characteristics:
%
\begin{inparaenum}
\item \emph{target code instrumentation}, to capture execution information from
  each test case; and
\item \emph{high-level, well-typed mutations}, to produce syntactically valid
  test cases by altering existing ones at the datatype level.
\end{inparaenum}

Using code instrumentation in tandem with mutations is a well-known technique in
the fuzzing community.
%
Generic fuzzing tools like AFL, libFuzzer or HonggFuzz, as well as
language-specific ones like Crowbar use execution traces to recognize
interesting test cases, e.g, those that exercise previously undiscovered parts
of the target code.
%
Later, such tools use generic mutators to combine and produce new test cases
from previously executed interesting ones.
%
\fuzzchick, however, does this in a clever way.
%
Instead of mutating any previously executed test case that discovers a new part
of the code, \fuzzchick integrates these fuzzing techniques into the PBT testing
loop itself.

Since it is possible to distinguish semantically valid test cases from invalid
ones, i.e., those passing the sparse preconditions of our testing properties as
opossed to those that are discarded early, \fuzzchick exploits this information
in order to focus the testing efforts into mutating valid test cases with a
higher priority than those that were discarded.

In addition, high-level mutators are better suited for producing syntactically
valid mutants, avoiding the time wasted by using generic low-level mutators that
act at the ``serialized'' level and know very little about the structure of the
generated data, thus producing syntactically broken mutants most of the time.
%
This grammar-aware mutation technique has shown to be quite useful when fuzzing
systems accepting structurally complex inputs.
%
Tools like Criterion, XSmith and LangFuzz use existing grammars to tailor the
generic mutators to the specific input structure used by the system under test.
%
In \fuzzchick, external grammars are not required.
%
The datatypes used by the inputs of the testing properties already describe the
structure of the random data we want to mutate in a concrete manner, and
specialized mutators acting at the data constructor level can be automatically
derived directly from their definition.

The next subsections describe \fuzzchick's testing loop and well-typed mutations
in detail.

\subsubsection{Testing loop}

Outlined in Algorithm \ref{algo:fuzzchick:loop}, the process starts by creating
two queues, \textit{QSucc} and \textit{QDisc} for valid and discarded previously
executed test cases, respectively.
%
Enqueued values are stored along with a given mutation energy, that controls how
many times a given test case can be mutated before being finally discared.

Once inside of the main loop, \fuzzchick picks the next test case using a simple
criterion: if there are valid values enqueued for mutation, it picks the first
one, mutates it and returns it, decreasing its energy by one.
%
If \textit{QSucc} is empty, then the same is attempted using \textit{QDisc}.
%
If none of the mutation queues contain any candidates, \fuzzchick generates a
new value from scratch.
%
This selection process is illustrated in detail in Algorithm
\ref{algo:fuzzchick:pick}.

Having selected the next test case, the main loop proceeds to execute it,
capturing both the result (passed, discarded, or failed) and its execution trace
over the system under test.
%
If the test case fails, it is immediately reported as a bug.
%
If not, \fuzzchick evaluates whether it was interesting (i.e., it exercises a
new path) based on its trace information and the one from previously executed
test cases (represented by \textit{TLog}).
%
If the test case does in fact discover a new path, it is enqueued at the end of
its corresponding queue, depending on whether it passed or was discarded.
%
This process alternates between generation and mutation until a bug is found or
we reach the test limit.

The energy assigned to each test case follows that of AFL's power schedule: more
energy to test cases that lead to shorter executions, or that discover more
parts of the code.
%
Moreover, to favour mutating interesting valid test cases, they get more energy
than those that were discarded.

\begin{figure}[t]
\begin{multicols}{2}

\begin{algorithm}[H]
  \SetArgSty{textnormal}
  \SetInd{0em}{0.75em}
  \SetKw{KwNot}{not}
  \SetAlgoLined
  \DontPrintSemicolon
  \SetKwFunction{FC}{Loop}
  \SetKwProg{Fn}{Function}{:}{}
  \Fn{\FC{P, N, gen, mut}}{
    i $\gets$ 0\;
    TLog, QSucc, QDisc $\gets$ $\varnothing$;\;
    \While{i $<$ N}{
      x $\gets$ Pick(QSucc, QDisc, gen, mut)\;
      (result, trace) $\gets$ WithTrace(P(x))\;
      \lIf{\KwNot result}{\KwRet Bug(x)}
      \If{Interesting(TLog, trace)}{
        e $\gets$ Energy(TLog, x, trace)\;
        \eIf{\KwNot Discarded(result)}{
          Enqueue(QSucc, (x, e))\;
        }{
          Enqueue(QDisc, (x, e))\;
        }
      }
      i $\gets$ i+1\;
    }
    \KwRet Ok\;
  }
\caption{\label{algo:fuzzchick:loop}\fuzzchick Testing Loop}
\end{algorithm}

\columnbreak

\begin{algorithm}[H]
  \SetArgSty{textnormal}
  \SetInd{0em}{0.75em}
  \SetKw{KwNot}{not}
  \SetAlgoLined
  \DontPrintSemicolon
  \SetKwFunction{Pick}{Pick}
  \SetKwProg{Fn}{Function}{:}{}
  \Fn{\Pick{QSucc, QDisc, gen, mut}}{
    \If{\KwNot Empty(QSucc)}{
      (x,e) $\gets$ Deque(QSucc)\;
      \If{e $>$ 0}{
        PushFront(QSucc, (x, e-1))
      }
      \KwRet Sample(mut(x))\;
    }
    \uElseIf{\KwNot Empty(QDisc)}{
      (x,e) $\gets$ Deque(QDisc)\;
      \If{e $>$ 0}{
        PushFront(QDisc, (x, e-1))
      }
      \KwRet Sample(mut(x))\;
    }
    \lElse{\KwRet Sample(gen)}
  }
\caption{\label{algo:fuzzchick:pick}\fuzzchick Seed Selection}
\end{algorithm}

\end{multicols}
\end{figure}

\subsubsection{Well-typed mutations}

Mutators in \fuzzchick are no more than specialized random generators,
parameterized by the original input to be mutated.
%
They use a simple set of mutation operations that are randomly applied at the
datatype level.
%
In simple terms, these operations encompass:
%
\begin{itemize}
\item \emph{shrinking the value}, replacing its top-level data constructor with
  one that contains a subset of its fields, reusing existing subexpressions;
\item \emph{growing the value}, replacing its top-level data constructor with
  one that contains a superset of its fields, reusing existing subexpressions
  and generating random ones when needed;
\item \emph{returning a subexpression of the same type};
\item \emph{mutating recursively}, applying a mutation operation over an
  immediate subexpression.
\end{itemize}

\begin{figure}
\begin{verbatim}
mutate_tree :: (a -> Gen a) -> Tree a -> Gen (Tree a)
mutate_tree mutate_a (Leaf x) =
  oneof [ do { x' <- mutate_a x; return (Leaf x')}                     -- Mutate recursively
        , do { l <- arbitrary; r <- arbitrary; return (Node l x r) } ] -- Grow constructor
mutate_tree (Node l x r) =
  oneof [ return l                                                     -- Return subexpression
        , return r                                                     -- Return subexpression
        , return (Leaf x)                                              -- Shrink constructor
        , do { l' <- mutate_tree l; return (Node l' x r) }             -- Mutate recursively
        , do { x' <- mutate_a    x; return (Node l x' r) }             -- Mutate recursively
        , do { r' <- mutate_tree r; return (Node l x r') } ]           -- Mutate recursively
\end{verbatim}
\caption{\label{fig:fuzzchick:mutator}\fuzzchick mutator for the \texttt{Tree} data type. }
\end{figure}

Fig. \ref{fig:fuzzchick:mutator} illustrates the idea behind a random \fuzzchick
mutator for our previously used \texttt{Tree} data type example.
%
Since trees are parametric, for clarity this definition is also parameterized by
a mutator for the payload (\texttt{mutate\_a}), although in practice this can be
abstracted away by the means of the type class system (where
\texttt{mutate\_tree} and \texttt{mutate\_a} can be transparently coerced into
an overloaded \texttt{mutate} function.)

In this mutator, branches can be shrinked into leaves by dropping the subtrees,
whereas leaves can grow into branches, by reusing the payload and generating two
random subtrees.
%
Moreover, branches can be replaced with one of their subtrees.
%
Finally, mutations can be recursively applied over both the payload and the
subtrees.
%
At the top level, all these operations are put together using the \texttt{oneof}
combinator that randomly picks one of them with uniform probability.

%
% FuzzChick Limitations
%
\subsubsection{Limitations of \fuzzchick}

\citeauthor{lampropoulos2019coverage} demonstrated empirically that \fuzzchick
lies comfortably in the middle ground between using na\"ive automatically
derived random generators and complex manually-written ones.
%
Their results suggest that CGPT is an appealing technique for finding bugs while
still offering a mostly automated workflow.

However, the authors acknowledge that certain parts of their implementation have
room for improvement, especially when it comes to the mutators design.
%
Moreover, when we recreated the evaluation of the IFC stack machine case study
(described in detail in Section \ref{sec:casestudies}), we found that after 30
runs (as opposed to the 5 runs used original by
\citeauthor{lampropoulos2019coverage}), \emph{\fuzzchick was only able to find 5
  out of the 20 injected bugs with a failure rate of 1}, the hardest one being
found only around 13\% of the time after an hour of testing.
%
These results are presented in detail in Section \ref{sec:evaluation}.

At the light of these observations, we identified several aspects of \fuzzchick
that can be improved upon --- and that constitute the main goal of this work.
%
In no particular order:
%
\begin{itemize}
\item \emph{Mutators distribution:} if we inspect the mutator defined in Fig.
  \ref{fig:fuzzchick:mutator}, there are two compromises that the authors of
  \fuzzchick adopted for the sake of simplicity.
  %
  On one hand, deep recursive mutations are very unlikely, since their
  probability decreases multiplicatively with each recursive call.
  %
  For instance, mutating a subexpression that lies on the third level of a
  \texttt{Tree} happens with a probability smaller than $(1/6)^3 = \sim 0.0046$,
  and this only worsens as the type of the mutated value becomes more complex.
  %
  Hence, \fuzzchick mutators can only be effectively used to transform to
  shallow data structures, potentially excluding interesting applications that
  might require producing deeper valid values, e.g., programming languages,
  network protocols interactions, etc.
  %
  Ideally, mutations should be able to happen on every subexpression of the
  input seed in a reasonable basis.

  On the other hand, using random generators to produce neeeded subexpressions
  when growing data constructors can be dangerous, as we are introducing the
  very same ``uncontrolled'' randomness that we wanted to mitigate in the first
  place!
  %
  If the random generator produces an invalid subexpression (something quite
  likely), this might just invalidate the whole mutated test case.
  %
  We believe that growing data constructors needs to be done carefully.
  %
  For instance, by using just a minimal piece of data to make the overall
  mutated test case type correct.
  %
  If that mutated test case to be interesting, that subexpression can always be
  mutated later.

\item \emph{Enqueuing mutation candidates:} \fuzzchick uses two single queues for
  keeping valid and discarded mutation candidates.
  %
  Whenever a new test case is found interesting, it is placed \emph{at the end
    of its corresponding queue}.
  %
  If this test case happens to have discovered a whole new portion of the target
  code, it will not be further mutated until the rest of the queue ahead of it
  gets processed.
  %
  This can limit the effectiveness of the testing loop if the queues tend to
  grow more often than they tend to shrink, as interesting mutation candidates
  can get buried at the end of a long queue that only exercises the same portion
  of the target code.
  %
  In the extreme case, they might not processed at all within the testing budget.
  %
  Ideally, one would like a mechanism that prioritizes mutating test cases
  that discovers new portions of the code right away, and that is capable of
  jumping back and forth from mutation candidates whenever this happens.
  %
  We show in Section \ref{sec:heuristics} how this can be achieved by analyzing
  the execution information in order to prioritize test cases with novel
  execution traces.

\item \emph{Power schedule:} It is not clear how the power schedule used to
  assign energy to each mutable test case in \fuzzchick works in the context of
  high-level well-typed mutations.
  %
  If it assigns too much energy to certain not-so-interesting seeds, some bugs
  might not be discovered in a timely basis.
  %
  Conversely, assigning too little energy to interesting test cases might cause
  that some bugs cannot be discovered at all unless the right mutation happens
  within the small available energy window --- randomly generating the same test
  case later on does not help, as it becomes uninteresting based on historic
  trace information.

  To keep the comparison fair, the authors replicated the same power schedule
  configuration used in AFL.
  %
  However, AFL uses a different mutation approach that works at the bit level.
  %
  This raises the question about what is the best power schedule configuration
  when using a high-level mutation approach --- something quite challenging to
  characterize in general given the expressivity of the data types used to
  drive the mutators.
\end{itemize}

The next section introduces \mutagen, our CGPT tool written in Haskell that aims
to tackle the main limitations of \fuzzchick using an exhaustive mutation
approach that requires very little randomness and no power schedule.

\section{\mutagen: Testing Mutants Exhaustively}
\label{sec:mutagen}

Aliquam erat volutpat \citep{devai2013edsl}. Nunc eleifend leo vitae magna. In
id erat non orci commodo lobortis. Proin neque massa, cursus ut, gravida ut,
lobortis eget, lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus.
Mauris ac felis vel velit tristique imperdiet.
nec dui dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a,
aliquet quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet
tortor. Nam euismod tellus id erat.

\begin{verbatim}
type Mutation a = a -> [Mutant a]
\end{verbatim}

\begin{verbatim}
data Mutant a = PURE a | RAND (Gen a)
\end{verbatim}

\begin{verbatim}
mutate (Leaf x)     = [ PURE (Node def x def) ]   -- Swap constructor
mutate (Node l x r) = [ PURE l, PURE r            -- Return subexpression
                      , PURE (Leaf x)             -- Swap constructor
                      , PURE (Node l x l)         -- Rearrange subexpressions
                      , PURE (Node r x r)         -- Rearrange subexpressions
                      , PURE (Node r x l) ]       -- Rearrange subexpressions
\end{verbatim}

\begin{verbatim}
positions (Leaf x)       = node [ (0, positions x) ]
positions (Branch l x r) = node [ (0, positions l), (1, positions x), (2, positions r) ]
\end{verbatim}

\begin{verbatim}
inside []        mut x              = mut x
inside (0 : pos) mut (Leaf x)       = [ Leaf x'       | x' <- inside pos mut x ]
inside (0 : pos) mut (Branch l x r) = [ Branch l' x r | l' <- inside pos mut l ]
inside (1 : pos) mut (Branch l x r) = [ Branch l x' r | x' <- inside pos mut x ]
inside (2 : pos) mut (Branch l x r) = [ Branch l x r' | r' <- inside pos mut r ]
\end{verbatim}

Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
Nam euismod tellus id erat.

\begin{figure}
\begin{multicols}{2}

\begin{algorithm}[H]
  \SetArgSty{textnormal}
  \SetInd{0em}{0.75em}
  \SetKw{KwNot}{not}
  \SetAlgoLined
  \DontPrintSemicolon
  \SetKwFunction{MG}{Loop}
  \SetKwProg{Fn}{Function}{:}{}
  \Fn{\MG{P, N, R, gen, mut}}{
    i $\gets$ 0\;
    TLog, QSucc, QDisc $\gets$ $\varnothing$;\;
    \While{i $<$ N}{
      x $\gets$ Pick(QSucc, QDisc, gen, mut)\;
      (result, trace) $\gets$ WithTrace(P(x))\;
      \lIf{\KwNot result}{\KwRet Bug(x)}
      \If{Interesting(TLog, trace)}{
        \If{\KwNot Discarded(result)}{
          muts $\gets$ Mutants(x, mut, R)\;
          Enqueue(QSucc, muts)\;
        }
        \uElseIf{Passed(Parent(x))}{
          muts $\gets$ Mutants(x, mut, R)\;
          Enqueue(QDisc, muts)\;
        }
      }
      i $\gets$ i+1\;
    }
    \KwRet Ok\;
  }
\caption{\mutagen Testing Loop}
\end{algorithm}

\columnbreak

\begin{algorithm}[H]
  \SetArgSty{textnormal}
  \SetInd{0em}{0.75em}
  \SetKw{KwNot}{not}
  \SetAlgoLined
  \DontPrintSemicolon
  \SetKwFunction{Pick}{Pick}
  \SetKwProg{Fn}{Function}{:}{}
  \Fn{\Pick{QSucc, QDisc, gen}}{
    \If{\KwNot Empty(QSucc)}{
      muts $\gets$ Deque(QSucc)\;
      \If{Empty(muts)}{
        Pick(QSucc, QDisc, gen)\;
      } \uElse{
        PushFront(QSucc, Rest(muts))\;
        \KwRet First(muts)\;
      }
    }
    \If{\KwNot Empty(QSucc)}{
      muts $\gets$ Deque(QSucc)\;
      \If{Empty(muts)}{
        Pick(QSucc, QDisc, gen)\;
      } \uElse{
        PushFront(QSucc, Rest(muts))\;
        \KwRet First(muts)\;
      }
    }
    \lElse{\KwRet Sample(gen)}
  }
\caption{\mutagen Seed Selection}
\end{algorithm}
\end{multicols}
\end{figure}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
Nam vestibulum accumsan nisl.

\begin{figure}

\begin{algorithm}[H]
  \SetArgSty{textnormal}
  \SetInd{0em}{0.75em}
  \SetKw{KwNot}{not}
  \SetKw{KwIn}{in}
  \SetKwFor{KwRepeat}{repeat}{times}
  \SetAlgoLined
  \DontPrintSemicolon
  \SetKwFunction{Muts}{Mutants}
  \SetKwProg{Fn}{Function}{:}{}
  \Fn{\Muts{x, mut, R}}{
    muts $\gets$ $\varnothing$\;
    \For{pos \KwIn Flatten(Positions(x))}{
      \For{mutant \KwIn Inside(pos, mut, x)}{
        \Switch{mutant}{
          \Case{PURE $\hat{x}$}{
            Enqueue($\hat{x}$, muts)\;
          }
          \Case{RAND gen}{
            \KwRepeat{R}{
              $\hat{x}$ $\gets$ Sample(gen)\;
              Enqueue($\hat{x}$, muts)\;
            }
          }
        }
      }
    }
    \KwRet muts\;
  }
\caption{Mutants Initialization}
\end{algorithm}

\end{figure}

\section{\mutagen Heuristics}
\label{sec:heuristics}

Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
commodo.

\subsection{Priority FIFO Scheduling}

Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
lacinia purus, et dictum nunc justo sit amet elit.

\begin{figure}
\begin{multicols}{2}

\begin{algorithm}[H]
  \SetArgSty{textnormal}
  \SetInd{0em}{0.75em}
  \SetKw{KwNot}{not}
  \SetAlgoLined
  \DontPrintSemicolon
  \SetKwFunction{MG}{Loop}
  \SetKwProg{Fn}{Function}{:}{}
  \Fn{\MG{P, N, R, gen, mut}}{
    $\cdots$\;
    (result, trace) $\gets$ WithTrace(P(x))\;
    $\cdots$\;
    \If{Interesting(TLog, trace)}{
      \If{\KwNot Discarded(result)}{
        muts $\gets$ Mutants(x, mut, R)\;
        {\color{red} prio $\gets$ BranchDepth(TLog, trace)}\; %% NEW
        {\color{red} PushFront(QSucc, prio, muts)}\;          %% NEW
      }
      $\cdots$\;
    }
  }
  \vspace{10pt}
  \SetKwFunction{Pick}{Pick}
  \SetKwProg{Fn}{Function}{:}{}
  \Fn{\Pick{QSucc, QDisc, gen}}{
    \If{\KwNot Empty(QSucc)}{
      {\color{red} (muts, prio) $\gets$ DequeMin(QSucc)}\; %%NEW
      \If{Empty(muts)}{
        Pick(QSucc, QDisc, gen)\;
      } \uElse{
        {\color{red} PushFront(QSucc, prio, Rest(muts))}\; %% NEW
        \KwRet First(muts)\;
      }
    }
    $\cdots$\;
  }

\caption{Priority FIFO Heuristic}
\end{algorithm}

\columnbreak

\begin{algorithm}[H]
  \SetArgSty{textnormal}
  \SetInd{0em}{0.75em}
  \SetKw{KwNot}{not}
  \SetAlgoLined
  \DontPrintSemicolon
  \SetKwFunction{MG}{Loop}
  \SetKwProg{Fn}{Function}{:}{}
  \Fn{\MG{P, N, gen, mut}}{
    {\color{red}
    boring  $\gets$ 0\;
    reset   $\gets$ 1000\;
    R       $\gets$ 1\;
    }
    $\cdots$\;
    \While{i $<$ N}{
      {\color{red}
      \If{boring $>$ reset}{
        TLog   $\gets$ $\varnothing$\;
        reset $\gets$ reset * 2\;
        R     $\gets$ R * 2\;
      }
      }
      $\cdots$\;
      \lIf{\KwNot result}{\KwRet Bug(x)}
      \If{Interesting(TLog, trace)}{
        $\cdots$\;
      }
      \lElse {
        {\color{red} boring $\gets$ boring + 1}
      }
      $\cdots$\;
    }
  }

\caption{Trace Saturation Heuristic}
\end{algorithm}

\end{multicols}
\caption{\label{fig:heuristics} The two main heuristics implemented in \mutagen.
  Statements in {\color{red} red} indicate important changes to the base
  algorithm. Ellipses denote parts of the code that are not relevant for the
  point being made. }
\end{figure}

\subsection{Detecting Trace Space Saturation And Tuning Random Mutations}

Vestibulum convallis, lorem a tempus semper, dui dui euismod elit, vitae
placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis, varius
et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat tellus ut
neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus lacinia purus,
et dictum nunc justo sit amet elit.

\subsection{Mutation Inheritance}

Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
lacinia purus, et dictum nunc justo sit amet elit.

\section{Case studies}
\label{sec:casestudies}

Aliquam erat volutpat \citep{devai2013edsl}. Nunc eleifend leo vitae magna. In id
erat non orci commodo lobortis. Proin neque massa, cursus ut, gravida ut,
lobortis eget, lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus.
Mauris ac felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque
nec dui dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a,
aliquet quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet
tortor. Nam euismod tellus id erat.

\subsection{IFC Stack Machine}

Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
lacinia purus, et dictum nunc justo sit amet elit.

\subsection{WebAssembly Engine}

Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.

\subsubsection{Testing the WebAssembly Validator}

Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.

\begin{verbatim}
prop_validator m =
  isValidHaskell m ==> isValidSpec m
\end{verbatim}

Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
Nam euismod tellus id erat.

\begin{verbatim}
isValidHaskell m =
  case validate m of
    Left  validationError -> return False
    Right validModule     -> return True
\end{verbatim}

Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
Nam euismod tellus id erat.

\begin{verbatim}
isValidSpec m = do
  writeFile "testcase.wasm" (dumpModule m)
  res <- shell "./wasm" ["-d", "-i", "testcase.wasm"]
  return (res == "")
\end{verbatim}

Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
Nam euismod tellus id erat.

\subsubsection{Testing the WebAssembly Interpreter}

Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
Nam euismod tellus id erat.

\begin{verbatim}
mk_module ty name fun =
  emptyModule { types     = [ ty ]
              , functions = [ fun ]
              , exports   = [ Export name (ExportFunc 0) ]
              , mems      = [ Memory (Limit 1 Nothing) ]
              }
\end{verbatim}

\begin{verbatim}
prop_interpreter ty (args, fun) =
  discardAfter 20
  (do let m = mk_module ty "f" fun;
      resHs   <- invokeHaskell m "f" args;
      resSpec <- invokeSpec    m "f" args;
      return (equiv resHs resSpec))
\end{verbatim}

\begin{verbatim}
prop_interpreter_i32 (i, f, fun) =
  prop_interpreter
    (FuncType { params = [I32, F32], result = [I32]})
    ([VI32 i, VF32 f], fun)
\end{verbatim}

\subsubsection{Real Bugs and Discrepancies found by \mutagen in \textit{haskell-wasm}}

\paragraph{Bug \#1: Invalid Memory Alignment Validation}
Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
Nam vestibulum accumsan nisl.

\paragraph{Bug \#2: Lax invocation checks}
Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
Nam euismod tellus id erat.

\paragraph{Bug \#3: Allowed Out-of-bounds Memory Access}
Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
lacinia purus, et dictum nunc justo sit amet elit.

\paragraph{Discrepancy \#1: Different Reinterpretation Semantics of NaN Values}
Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
Nam vestibulum accumsan nisl.

\paragraph{Discrepancy \#2: Allowed Blocks with Multiple Return Types}
Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
Nam vestibulum accumsan nisl.

\section{Evaluation}
\label{sec:evaluation}

\subsection{IFC Stack Machine}

Aliquam erat volutpat \citep{devai2013edsl}. Nunc eleifend leo vitae magna. In
id erat non orci commodo lobortis. Proin neque massa, cursus ut, gravida ut,
lobortis eget, lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus.
Mauris ac felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque
nec dui dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a,
aliquet quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet
tortor. Nam euismod tellus id erat.

\begin{center}
  \includegraphics[width=\linewidth]{tikz/ifc-mttf.tikz}
  \includegraphics[width=\linewidth]{tikz/ifc-fr.tikz}
\end{center}

\subsection{Webassembly Engine}

Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
Nam euismod tellus id erat.

\begin{center}
  \includegraphics[width=\linewidth]{tikz/wasm-mttf.tikz}
  \includegraphics[width=\linewidth]{tikz/wasm-fr.tikz}\vspace{-5pt}
  \includegraphics[width=\linewidth]{tikz/wasm-legend.tikz}
\end{center}

\section{Implementation}
\label{sec:implementation}

\subsection{Deriving Mutation Machinery}

Integer placerat tristique nisl. Praesent augue. Fusce commodo. Vestibulum
convallis, lorem a tempus semper, dui dui euismod elit, vitae placerat urna
tortor vitae lacus. Nullam libero mauris, consequat quis, varius et, dictum id,
arcu. Mauris mollis tincidunt felis. Aliquam feugiat tellus ut neque. Nulla
facilisis, risus a rhoncus fermentum, tellus tellus lacinia purus, et dictum
nunc justo sit amet elit.

\subsection{Tracing Haskell Programs}

\begin{verbatim}
  sorted []       = True
  sorted [x]      = True
  sorted (x:y:xs) = if x <= y then sorted (y:xs) else False
\end{verbatim}

\begin{verbatim}
  sorted []       = _trace_ 1 (True)
  sorted [x]      = _trace_ 2 (True)
  sorted (x:y:xs) = _trace_ 3 (if x <= y then _trace_ 4 (sorted (y:xs)) else _trace_ 5 (False))
\end{verbatim}

Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
mollis ultricies.

\section{Related work}
\label{sec:related}

\subsection{Grammar-based Fuzzing}

Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
lacinia purus, et dictum nunc justo sit amet elit.

\subsection{Random Data Generation}

%% Don't forget to cite "Deriving good generators ..."

Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
lacinia purus, et dictum nunc justo sit amet elit.

\subsection{Exhaustive Bounded Testing}

Aliquam erat volutpat \citep{devai2013edsl}. Nunc eleifend leo vitae magna. In id
erat non orci commodo lobortis. Proin neque massa, cursus ut, gravida ut,
lobortis eget, lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus.
Mauris ac felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque
nec dui dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a,
aliquet quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet
tortor. Nam euismod tellus id erat.

\section{Conclusions}
\label{sec:conclusions}

Aliquam erat volutpat \citep{devai2013edsl}. Nunc eleifend leo vitae magna. In id
erat non orci commodo lobortis. Proin neque massa, cursus ut, gravida ut,
lobortis eget, lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus.
Mauris ac felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque
nec dui dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a,
aliquet quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet
tortor. Nam euismod tellus id erat.

% ----------------------------------------

%% The acknowledgments section is defined using the "acks" environment (and NOT
%% an unnumbered section). This ensures the proper identification of the section
%% in the article metadata, and the consistent spelling of the heading.

%% \begin{acks}
%% To Robert, for the bagels and explaining CMYK and color spaces.
%% \end{acks}

% ----------------------------------------

%% The next two lines define the bibliography style to be used, and the
%% bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

% ----------------------------------------

%% If your work has an appendix, this is the place to put it.
\appendix

\section{Detailed Empirical Results}

Fusce suscipit, wisi nec facilisis facilisis, est dui fermentum leo, quis tempor
ligula erat quis odio. Nunc porta vulputate tellus. Nunc rutrum turpis sed pede.
Sed bibendum. Aliquam posuere. Nunc aliquet, augue nec adipiscing interdum,
lacus tellus malesuada massa, quis varius mi purus non odio. Pellentesque
condimentum, magna ut suscipit hendrerit, ipsum augue ornare nulla, non luctus
diam neque sit amet urna. Curabitur vulputate vestibulum lorem. Fusce sagittis,
libero non molestie mollis, magna orci ultrices dolor, at vulputate neque nulla
lacinia eros. Sed id ligula quis est convallis tempor. Curabitur lacinia
pulvinar nibh. Nam a sapien.

\end{document}
\endinput
%%
%% End of `main.tex'.

### 2021-03-23 papers/oopsla-21/_TZ_34837-obiwan.tex
WIP

--- 2021-03-22 papers/oopsla-21/main.tex
@@
 
-%%
-%% The "title" command has an optional parameter,
-%% allowing the author to define a "short title" to be used in page headers.
-\title{\mutagen: Working title}
-
-%%
-%% The "author" command and its associated commands are used to define
-%% the authors and their affiliations.
-%% Of note is the shared affiliation of the first two authors, and the
-%% "authornote" and "authornotemark" commands
-%% used to denote shared contribution to the research.
-
-\author{Agust\'in Mista}
-\email{mista@chalmers.se}
-\affiliation{%
-  \institution{Chalmers University of Technology}
-  \city{Gothenburg}
-  \country{Sweden}
-}
-
-\author{Alejandro Russo}
-\email{russo@chalmers.se}
-\affiliation{%
-  \institution{Chalmers University of Technology}
-  \city{Gothenburg}
-  \country{Sweden}
-}
-
-%% By default, the full list of authors will be used in the page headers. Often,
-%% this list is too long, and will overlap other information printed in the page
-%% headers. This command allows the author to define a more concise list of
-%% authors' names for this purpose.
-
-%% \renewcommand{\shortauthors}{Mista and Russo}
-
-% ----------------------------------------
-%% The abstract is a short summary of the work to be presented in the article.
-
-\begin{abstract}
-  Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-  facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc
-  porta vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam
-  posuere. Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada
-  massa, quis varius mi purus non odio. Pellentesque condimentum, magna ut
-  suscipit hendrerit, ipsum augue ornare nulla, non luctus diam neque sit amet
-  urna. Curabitur vulputate vestibulum lorem. Fusce sagittis, libero non
-  molestie mollis, magna orci ultrices dolor, at vulputate neque nulla lacinia
-  eros. Sed id ligula quis est convallis tempor. Curabitur lacinia pulvinar
-  nibh. Nam a sapien.
-\end{abstract}
-
-% ----------------------------------------
-
-%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm. Please
-%% copy and paste the code instead of the example below.
-\begin{CCSXML}
-% CCS XML here
-\end{CCSXML}
-
-%% \ccsdesc[500]{Computer systems organization~Embedded systems}
-%% \ccsdesc[300]{Computer systems organization~Redundancy}
-%% \ccsdesc{Computer systems organization~Robotics}
-%% \ccsdesc[100]{Networks~Network reliability}
-
-% ----------------------------------------
-
-%% Keywords. The author(s) should pick words that accurately describe the work
-%% being presented. Separate the keywords with commas.
-\keywords{random testing, mutations, heuristics}
-
-% ----------------------------------------
-
-%% This command processes the author and affiliation and title information and
-%% builds the first part of the formatted document.
-\maketitle
-
-% ----------------------------------------
-
-\section{Introduction}
-\label{sec:intro}
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
-The contributions of this work are:
-
-\begin{itemize}
-\item Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec
-  facilisis facilisis, est dui fermentum leo, quis tempor ligula erat quis odio.
-  Nunc porta vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum.
-  Aliquam posuere. Nunc aliquet, augue nec adipiscing interdum, lacus tellus
-  malesuada massa, quis varius mi purus non odio.
-\item Pellentesque condimentum,
-  magna ut suscipit hendrerit, ipsum augue ornare nulla, non luctus diam neque
-  sit amet urna. Curabitur vulputate vestibulum lorem. Fusce sagittis, libero
-  non molestie mollis, magna orci ultrices dolor, at vulputate neque nulla
-  lacinia eros. Sed id ligula quis est convallis tempor. Curabitur lacinia
-  pulvinar nibh. Nam a sapien.
-\item Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec
-  facilisis facilisis, est dui fermentum leo, quis tempor ligula erat quis odio.
-  Nunc porta vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum.
-  Aliquam posuere. Nunc aliquet, augue nec adipiscing interdum, lacus tellus
-  malesuada massa, quis varius mi purus non odio.
-\item Pellentesque condimentum, magna ut suscipit hendrerit, ipsum augue ornare
-  nulla, non luctus diam neque sit amet urna. Curabitur vulputate vestibulum
-  lorem. Fusce sagittis, libero non molestie mollis, magna orci ultrices dolor,
-  at vulputate neque nulla lacinia eros. Sed id ligula quis est convallis
-  tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-\end{itemize}
-
-\section{Background}
-\label{sec:background}
-
-In this section we briefly introduce the concept of Property-based Testing along
-with \quickcheck, the most popular tool of this sort and often considered as the
-baseline when comparing PBT algoritms.
-%
-Moreover, we describe the ideas and limitations behind \fuzzchick, which
-enhances the PBT approach with execution information and serves as the
-foundation for \mutagen.
-%
-\fuzzchick is implemented as an extension of \quickchick, Coq's own
-reimplementation of \quickcheck.
-
-\subsection{Property-Based Testing and \quickcheck}
-
-Property-based testing is a powerful technique for finding bugs without having
-to write test cases by hand.
-%
-Originally introduced by \citet{quickcheck} in tandem with the first version of
-\quickcheck, this technique focuses on aiming the developer's efforts into
-testing systems via executable specifications using randomly generated inputs.
-%
-Moreover, tools like \quickchick and Isabelle's \quickcheck demonstrate
-that PBT can also be used in the formal verification realm.
-%
-There, one can quickly spot bugs in system specifications before directing the
-efforts into pointlessly trying to prove bogus propositions.
-
-In the simplest form, there are four main elements the user needs to provide in
-order to perform property-based testing on their systems:
-%
-\begin{itemize}
-\item one or more \emph{executable properties}, often implemented simply as a
-  boolean predicates,
-\item \emph{random data generators}, used to repeatedly instantiate the testing
-  properties,
-\item \emph{printers}, used to show the user the random inputs that falsify some
-  testing property (the counterexample) whenever a bug is found, and
-\item \emph{shrinkers}, to minimize counterexamples making them easier to
-  understand by humans.
-\end{itemize}
-
-\noindent In this work we focus solely on the first two elements introduced
-above, namely the testing properties and the random data generators used to feed
-them.
-%
-Printers and shrinkers, for the most part, can be obtained automatically using
-generic programming capabilities present in the compiler, and although being
-crucial for the testing process as a whole, their role becomes irrelevant when
-it comes to \emph{finding} bugs.
-
-\begin{algorithm}[ b]
-  \SetKw{KwNot}{not}
-  \SetAlgoLined
-  \DontPrintSemicolon
-  \SetKwFunction{QC}{Loop}
-  \SetKwProg{Fn}{Function}{:}{}
-  \Fn{\QC{P, N, gen}}{
-    i $\gets$ 0\;
-    \While{i $<$ N}{
-      x $\gets$ Sample(gen)\;
-      \lIf{\KwNot P(x)}{\KwRet Bug(x)}
-      i $\gets$ i+1 \;
-    }
-    \KwRet Ok\;
-  }
-\caption{\label{algo:quickcheck}\quickcheck Testing Loop}
-\end{algorithm}
-
-Perhaps the simplest PBT technique is to repeatedly generate random inputs and
-instantiate the testing properties until they either get falsified by a
-counterexample, or we ran a sufficiently large amount of tests --- suggesting
-that the properties holds.
-%
-\quickcheck implements a testing loop that closely follows this simple idea,
-which is outlined in Algorithm \ref{algo:quickcheck}, where $P$ is the testing
-property, $N$ is the maximum number of tests to perform, and $gen$ is the random
-generator to be used to instantiate $P$.
-
-To illustrate this technique, let us focus on the same motivating example used
-by \citeauthor{lampropoulos2019coverage}, who propose a simple property defined over
-binary trees.
-%
-Such data structure can be easily defined in Haskell using a custom data type
-with two data constructors for leaves and branches respectively:
-
-\begin{verbatim}
-data Tree a = Leaf a | Branch (Tree a) a (Tree a)
-\end{verbatim}
-
-\noindent The type parameter \texttt{a} indicates that trees can be instantiated
-using any other type as payload, so the value \texttt{Leaf Bool} has type
-\texttt{Tree Bool}, whereas the value \texttt{Branch (Leaf 1) 2 (Leaf 3)} has
-type \texttt{Tree Int}.
-%
-Then, we can define tree reflections using a simple recursive function that
-pattern matches against the two possible constructors, inverting the order of
-the subtrees whenever it encounters a branch:
-
-\begin{verbatim}
-mirror :: Tree a -> Tree a
-mirror (Leaf x)       = Leaf x
-mirror (Branch l x r) = Branch (mirror r) x (mirror l)
-\end{verbatim}
-
-Later, a reasonable requirement to assert for is that \texttt{mirror} must be
-\emph{involutive}, i.e., reflecting a tree twice always yields the original
-tree.
-%
-We can simply capture this property using a boolean predicate written as a
-normal function:
-
-\begin{verbatim}
-prop_mirror :: Tree Int -> Bool
-prop_mirror t = mirror (mirror t) == t
-\end{verbatim}
-
-\noindent For simplicity, here we instantiate the tree payload with integers,
-altough this predicate should clearly hold for any other type with a properly
-defined notion of equality as well.
-
-With our simple specification in place, the last missing piece is a random
-generator of trees.
-%
-In \quickcheck, this is usually done via the type class mechanism, instantiating
-the \texttt{Arbitrary} type class, providing a random generator as the
-implementation of the overloaded \texttt{arbitrary} operation:
-
-\begin{verbatim}
-instance Arbitrary a => Arbitrary (Tree a) where
-  arbitrary = sized gen
-    where
-      gen 0 = do { x <- arbitrary; return (Leaf x) }
-      gen n = oneof [ do {x <- arbitrary; return (Leaf x) }
-                    , do {l <- gen (n-1); x <- arbitrary; r <- gen (n-1); return (Branch l x r)} ]
-\end{verbatim}
-
-\noindent Let us break this definition into parts.
-%
-The first line states that we will provide an \texttt{Arbitrary} instance for
-trees with payload of type \texttt{a}, provided that values of type \texttt{a}
-can also be randomly generated.
-%
-This allows us to use \texttt{arbitrary} to generate \texttt{a}'s inside the
-definition of our tree generator.
-
-Moreover, \quickcheck internally keeps track of the \emph{maximum generation
-  size}, a parameter that can be tuned by the user in order to limit the size of
-the randomly generated values.
-%
-Our definition exposes this internal parameter via \quickcheck's \texttt{sized}
-combinator, allowing us to parameterize the maximum size of the randomly
-generated trees.
-%
-If the generation size is zero (\texttt{gen 0}), our generator is limited to
-produce just leaves with randomly generated payloads.
-%
-In turn, when the generation size is strictly positive (\texttt{gen n}), the
-generator is able to perform a random uniform choice between generating either a
-single leaf or a branch.
-%
-When generating branches, the generator calls itself recursively in order to
-produce random subtrees (\texttt{gen (n-1)}).
-%
-Notice the importance of reducing the generation size on each recursive call.
-%
-This way we ensure that randomly generated trees using a generation size
-\texttt{n} are always finite and have at most \texttt{n} levels.
-
-Finally, we are ready to let \quickcheck test \texttt{prop\_mirror} against a
-large number of inputs (100 by default) produced by our brand new random tree
-generator:
-
-\begin{verbatim}
-quickCheck prop_mirror
-+++ Ok ####
-\end{verbatim}
-
-\noindent Should we mistakenly implement \texttt{mirror}, e.g., by dropping the
-right subtree altogether:
-
-\begin{verbatim}
-mirror (Branch l x r) = Branch (mirror l) x (mirror l)
-\end{verbatim}
-
-\noindent then \quickcheck will quickly falsify \texttt{prop\_mirror}, reporting
-a minimized counterexample that we can use to find the root of the issue:
-
-\begin{verbatim}
-quickCheck prop_mirror
-*** Failed ####
-\end{verbatim}
-
-At this point, it is clear that the \emph{quality} of our random generators is
-paramount to the performance of the overall PBT process.
-%
-Random generators that rarely produce interesting values will fail to trigger
-bugs in our code, potentially leaving entire parts of the codebase virtually
-untested.
-
-Recalling our tree generator, the reader (far from mistaken) might already have
-imagined better ways for implementing it.
-%
-For most practical purposes, this generator is in fact quite bad.
-%
-However, it follows a simple type-directed fashion, and it is a good example of
-what to expect from a random generator synthesized automatically using a process
-that knows very little about the values to be generated apart from their
-(syntactic) data type structure.
-
-As introduced earlier, there exist multiple tools that can automatically derive
-better random generators solely from the static information present in the
-codebase.
-%
-%% Such tools vary on the degree of invariants and requirements
-%% %
-%% However, such approaches are unable to derive useful generators whenever the
-%% target data involves complex invariants that cannot be easily extracted from the
-%% codebase or that cannot be easily expressed in terms of inductive relations.
-%
-Sadly, these tools lack the domain knowledge required to generate random data
-with complex invariants --- especially those present in programming languages
-like well-scopedness and well-typedeness.
-
-In particular, automatically derived generators are remarkably uneffective when
-used to test properties with sparse preconditions.
-%
-Let us continue with the example by \citeauthor{lampropoulos2019coverage} to
-illustrate this problem in more detail.
-%
-For this, consider that we want to use our \texttt{Tree} data type to encode
-binary-search trees (BST).
-%
-Then, given a predicate \texttt{isBST} that asserts if a tree satisfies the BST
-invariants, we might want to use it as pre- and post-condition to assert that
-BST operations like \texttt{insert} preserve them:
-
-\begin{verbatim}
-prop_bst_insert :: Tree a -> a -> Bool
-prop_bst_insert t a =
-  isBST t ==> isBST (insert a t)
-\end{verbatim}
-
-\noindent Attempting to test this property using \quickcheck does not work well:
-
-\begin{verbatim}
-quickCheck prop_bst_insert
-*** Gave up! ####
-\end{verbatim}
-
-\noindent Here, \quickcheck discards random inputs as soon as it finds they do
-not pass the precondition (\texttt{isBST t}).
-%
-Sadly, most of the inputs generated by our na\"ive generator suffer from this
-problem, and the interesting part of the property (\texttt{isBST (insert a t)})
-is tested very sporadically as a result.
-
-At this point it is reasonable to think that, to obtain the best results when
-using PBT over complex systems, one is forced to put a large amount of time on
-developing manually-written generators.
-%
-In practice, that is most often the case, no automatic effort can beat a
-well-thought manually-written generator that produces interesting complex values
-and finds bugs in very few tests.
-%
-Not all is lost, however.
-%
-It is still possible to obtain acceptable results automatically by incorporating
-dynamic information from the system under test into the testing loop.
-%
-The next subsection introduces the clever technique used by \fuzzchick to find
-bugs in complex systems while using simple automatically derived random
-generators.
-
-%% Far from trying to compete against manual-efforts, our work focuses on improving
-%% automatic approaches so they can be used paliatively in earlier development
-%% stages.
-
-\subsection{Coverage-Guided Property-Based Testing with \fuzzchick}
-
-To alleviate the problem of testing properties with sparse preconditions while
-using simple automatically derived random generators, \fuzzchick introduces
-\emph{coverage-guided property-based testing} (CGPT) by enhancing the testing
-process with two key characteristics:
-%
-\begin{inparaenum}
-\item \emph{target code instrumentation}, to capture execution information from
-  each test case; and
-\item \emph{high-level, well-typed mutations}, to produce syntactically valid
-  test cases by altering existing ones at the datatype level.
-\end{inparaenum}
-
-Using code instrumentation in tandem with mutations is a well-known technique in
-the fuzzing community.
-%
-Generic fuzzing tools like AFL, libFuzzer or HonggFuzz, as well as
-language-specific ones like Crowbar use execution traces to recognize
-interesting test cases, e.g, those that exercise previously undiscovered parts
-of the target code.
-%
-Later, such tools use generic mutators to combine and produce new test cases
-from previously executed interesting ones.
-%
-\fuzzchick, however, does this in a clever way.
-%
-Instead of mutating any previously executed test case that discovers a new part
-of the code, \fuzzchick integrates these fuzzing techniques into the PBT testing
-loop itself.
-
-Since it is possible to distinguish semantically valid test cases from invalid
-ones, i.e., those passing the sparse preconditions of our testing properties as
-opossed to those that are discarded early, \fuzzchick exploits this information
-in order to focus the testing efforts into mutating valid test cases with a
-higher priority than those that were discarded.
-
-In addition, high-level mutators are better suited for producing syntactically
-valid mutants, avoiding the time wasted by using generic low-level mutators that
-act at the ``serialized'' level and know very little about the structure of the
-generated data, thus producing syntactically broken mutants most of the time.
-%
-This grammar-aware mutation technique has shown to be quite useful when fuzzing
-systems accepting structurally complex inputs.
-%
-Tools like Criterion, XSmith and LangFuzz use existing grammars to tailor the
-generic mutators to the specific input structure used by the system under test.
-%
-In \fuzzchick, external grammars are not required.
-%
-The datatypes used by the inputs of the testing properties already describe the
-structure of the random data we want to mutate in a concrete manner, and
-specialized mutators acting at the data constructor level can be automatically
-derived directly from their definition.
-
-The next subsections describe \fuzzchick's testing loop and well-typed mutations
-in detail.
-
-\subsubsection{Testing loop}
-
-Outlined in Algorithm \ref{algo:fuzzchick:loop}, the process starts by creating
-two queues, \textit{QSucc} and \textit{QDisc} for valid and discarded previously
-executed test cases, respectively.
-%
-Enqueued values are stored along with a given mutation energy, that controls how
-many times a given test case can be mutated before being finally discared.
-
-Once inside of the main loop, \fuzzchick picks the next test case using a simple
-criterion: if there are valid values enqueued for mutation, it picks the first
-one, mutates it and returns it, decreasing its energy by one.
-%
-If \textit{QSucc} is empty, then the same is attempted using \textit{QDisc}.
-%
-If none of the mutation queues contain any candidates, \fuzzchick generates a
-new value from scratch.
-%
-This selection process is illustrated in detail in Algorithm
-\ref{algo:fuzzchick:pick}.
-
-Having selected the next test case, the main loop proceeds to execute it,
-capturing both the result (passed, discarded, or failed) and its execution trace
-over the system under test.
-%
-If the test case fails, it is immediately reported as a bug.
-%
-If not, \fuzzchick evaluates whether it was interesting (i.e., it exercises a
-new path) based on its trace information and the one from previously executed
-test cases (represented by \textit{TLog}).
-%
-If the test case does in fact discover a new path, it is enqueued at the end of
-its corresponding queue, depending on whether it passed or was discarded.
-%
-This process alternates between generation and mutation until a bug is found or
-we reach the test limit.
-
-The energy assigned to each test case follows that of AFL's power schedule: more
-energy to test cases that lead to shorter executions, or that discover more
-parts of the code.
-%
-Moreover, to favour mutating interesting valid test cases, they get more energy
-than those that were discarded.
-
-\begin{figure}[t]
-\begin{multicols}{2}
-
-\begin{algorithm}[H]
-  \SetArgSty{textnormal}
-  \SetInd{0em}{0.75em}
-  \SetKw{KwNot}{not}
-  \SetAlgoLined
-  \DontPrintSemicolon
-  \SetKwFunction{FC}{Loop}
-  \SetKwProg{Fn}{Function}{:}{}
-  \Fn{\FC{P, N, gen, mut}}{
-    i $\gets$ 0\;
-    TLog, QSucc, QDisc $\gets$ $\varnothing$;\;
-    \While{i $<$ N}{
-      x $\gets$ Pick(QSucc, QDisc, gen, mut)\;
-      (result, trace) $\gets$ WithTrace(P(x))\;
-      \lIf{\KwNot result}{\KwRet Bug(x)}
-      \If{Interesting(TLog, trace)}{
-        e $\gets$ Energy(TLog, x, trace)\;
-        \eIf{\KwNot Discarded(result)}{
-          Enqueue(QSucc, (x, e))\;
-        }{
-          Enqueue(QDisc, (x, e))\;
-        }
-      }
-      i $\gets$ i+1\;
-    }
-    \KwRet Ok\;
-  }
-\caption{\label{algo:fuzzchick:loop}\fuzzchick Testing Loop}
-\end{algorithm}
-
-\columnbreak
-
-\begin{algorithm}[H]
-  \SetArgSty{textnormal}
-  \SetInd{0em}{0.75em}
-  \SetKw{KwNot}{not}
-  \SetAlgoLined
-  \DontPrintSemicolon
-  \SetKwFunction{Pick}{Pick}
-  \SetKwProg{Fn}{Function}{:}{}
-  \Fn{\Pick{QSucc, QDisc, gen, mut}}{
-    \If{\KwNot Empty(QSucc)}{
-      (x,e) $\gets$ Deque(QSucc)\;
-      \If{e $>$ 0}{
-        PushFront(QSucc, (x, e-1))
-      }
-      \KwRet Sample(mut(x))\;
-    }
-    \uElseIf{\KwNot Empty(QDisc)}{
-      (x,e) $\gets$ Deque(QDisc)\;
-      \If{e $>$ 0}{
-        PushFront(QDisc, (x, e-1))
-      }
-      \KwRet Sample(mut(x))\;
-    }
-    \lElse{\KwRet Sample(gen)}
-  }
-\caption{\label{algo:fuzzchick:pick}\fuzzchick Seed Selection}
-\end{algorithm}
-
-\end{multicols}
-\end{figure}
-
-\subsubsection{Well-typed mutations}
-
-Mutators in \fuzzchick are no more than specialized random generators,
-parameterized by the original input to be mutated.
-%
-They use a simple set of mutation operations that are randomly applied at the
-datatype level.
-%
-In simple terms, these operations encompass:
-%
-\begin{itemize}
-\item \emph{shrinking the value}, replacing its top-level data constructor with
-  one that contains a subset of its fields, reusing existing subexpressions;
-\item \emph{growing the value}, replacing its top-level data constructor with
-  one that contains a superset of its fields, reusing existing subexpressions
-  and generating random ones when needed;
-\item \emph{returning a subexpression of the same type};
-\item \emph{mutating recursively}, applying a mutation operation over an
-  immediate subexpression.
-\end{itemize}
-
-\begin{figure}
-\begin{verbatim}
-mutate_tree :: (a -> Gen a) -> Tree a -> Gen (Tree a)
-mutate_tree mutate_a (Leaf x) =
-  oneof [ do { x' <- mutate_a x; return (Leaf x')}                     -- Mutate recursively
-        , do { l <- arbitrary; r <- arbitrary; return (Node l x r) } ] -- Grow constructor
-mutate_tree (Node l x r) =
-  oneof [ return l                                                     -- Return subexpression
-        , return r                                                     -- Return subexpression
-        , return (Leaf x)                                              -- Shrink constructor
-        , do { l' <- mutate_tree l; return (Node l' x r) }             -- Mutate recursively
-        , do { x' <- mutate_a    x; return (Node l x' r) }             -- Mutate recursively
-        , do { r' <- mutate_tree r; return (Node l x r') } ]           -- Mutate recursively
-\end{verbatim}
-\caption{\label{fig:fuzzchick:mutator}\fuzzchick mutator for the \texttt{Tree} data type. }
-\end{figure}
-
-Fig. \ref{fig:fuzzchick:mutator} illustrates the idea behind a random \fuzzchick
-mutator for our previously used \texttt{Tree} data type example.
-%
-Since trees are parametric, for clarity this definition is also parameterized by
-a mutator for the payload (\texttt{mutate\_a}), although in practice this can be
-abstracted away by the means of the type class system (where
-\texttt{mutate\_tree} and \texttt{mutate\_a} can be transparently coerced into
-an overloaded \texttt{mutate} function.)
-
-In this mutator, branches can be shrinked into leaves by dropping the subtrees,
-whereas leaves can grow into branches, by reusing the payload and generating two
-random subtrees.
-%
-Moreover, branches can be replaced with one of their subtrees.
-%
-Finally, mutations can be recursively applied over both the payload and the
-subtrees.
-%
-At the top level, all these operations are put together using the \texttt{oneof}
-combinator that randomly picks one of them with uniform probability.
-
-%
-% FuzzChick Limitations
-%
-\subsubsection{Limitations of \fuzzchick}
-
-\citeauthor{lampropoulos2019coverage} demonstrated empirically that \fuzzchick
-lies comfortably in the middle ground between using na\"ive automatically
-derived random generators and complex manually-written ones.
-%
-Their results suggest that CGPT is an appealing technique for finding bugs while
-still offering a mostly automated workflow.
-
-However, the authors acknowledge that certain parts of their implementation have
-room for improvement, especially when it comes to the mutators design.
-%
-Moreover, when we recreated the evaluation of the IFC stack machine case study
-(described in detail in Section \ref{sec:casestudies}), we found that after 30
-runs (as opposed to the 5 runs used original by
-\citeauthor{lampropoulos2019coverage}), \emph{\fuzzchick was only able to find 5
-  out of the 20 injected bugs with a failure rate of 1}, the hardest one being
-found only around 13\% of the time after an hour of testing.
-%
-These results are presented in detail in Section \ref{sec:evaluation}.
-
-At the light of these observations, we identified several aspects of \fuzzchick
-that can be improved upon --- and that constitute the main goal of this work.
-%
-In no particular order:
-%
-\begin{itemize}
-\item \emph{Mutators distribution:} if we inspect the mutator defined in Fig.
-  \ref{fig:fuzzchick:mutator}, there are two compromises that the authors of
-  \fuzzchick adopted for the sake of simplicity.
-  %
-  On one hand, deep recursive mutations are very unlikely, since their
-  probability decreases multiplicatively with each recursive call.
-  %
-  For instance, mutating a subexpression that lies on the third level of a
-  \texttt{Tree} happens with a probability smaller than $(1/6)^3 = \sim 0.0046$,
-  and this only worsens as the type of the mutated value becomes more complex.
-  %
-  Hence, \fuzzchick mutators can only be effectively used to transform to
-  shallow data structures, potentially excluding interesting applications that
-  might require producing deeper valid values, e.g., programming languages,
-  network protocols interactions, etc.
-  %
-  Ideally, mutations should be able to happen on every subexpression of the
-  input seed in a reasonable basis.
-
-  On the other hand, using random generators to produce neeeded subexpressions
-  when growing data constructors can be dangerous, as we are introducing the
-  very same ``uncontrolled'' randomness that we wanted to mitigate in the first
-  place!
-  %
-  If the random generator produces an invalid subexpression (something quite
-  likely), this might just invalidate the whole mutated test case.
-  %
-  We believe that growing data constructors needs to be done carefully.
-  %
-  For instance, by using just a minimal piece of data to make the overall
-  mutated test case type correct.
-  %
-  If that mutated test case to be interesting, that subexpression can always be
-  mutated later.
-
-\item \emph{Enqueuing mutation candidates:} \fuzzchick uses two single queues for
-  keeping valid and discarded mutation candidates.
-  %
-  Whenever a new test case is found interesting, it is placed \emph{at the end
-    of its corresponding queue}.
-  %
-  If this test case happens to have discovered a whole new portion of the target
-  code, it will not be further mutated until the rest of the queue ahead of it
-  gets processed.
-  %
-  This can limit the effectiveness of the testing loop if the queues tend to
-  grow more often than they tend to shrink, as interesting mutation candidates
-  can get buried at the end of a long queue that only exercises the same portion
-  of the target code.
-  %
-  In the extreme case, they might not processed at all within the testing budget.
-  %
-  Ideally, one would like a mechanism that prioritizes mutating test cases
-  that discovers new portions of the code right away, and that is capable of
-  jumping back and forth from mutation candidates whenever this happens.
-  %
-  We show in Section \ref{sec:heuristics} how this can be achieved by analyzing
-  the execution information in order to prioritize test cases with novel
-  execution traces.
-
-\item \emph{Power schedule:} It is not clear how the power schedule used to
-  assign energy to each mutable test case in \fuzzchick works in the context of
-  high-level well-typed mutations.
-  %
-  If it assigns too much energy to certain not-so-interesting seeds, some bugs
-  might not be discovered in a timely basis.
-  %
-  Conversely, assigning too little energy to interesting test cases might cause
-  that some bugs cannot be discovered at all unless the right mutation happens
-  within the small available energy window --- randomly generating the same test
-  case later on does not help, as it becomes uninteresting based on historic
-  trace information.
-
-  To keep the comparison fair, the authors replicated the same power schedule
-  configuration used in AFL.
-  %
-  However, AFL uses a different mutation approach that works at the bit level.
-  %
-  This raises the question about what is the best power schedule configuration
-  when using a high-level mutation approach --- something quite challenging to
-  characterize in general given the expressivity of the data types used to
-  drive the mutators.
-\end{itemize}
-
-The next section introduces \mutagen, our CGPT tool written in Haskell that aims
-to tackle the main limitations of \fuzzchick using an exhaustive mutation
-approach that requires very little randomness and no power schedule.
-
 \section{\mutagen: Testing Mutants Exhaustively}
@@
 
-Aliquam erat volutpat \citep{devai2013edsl}. Nunc eleifend leo vitae magna. In
-id erat non orci commodo lobortis. Proin neque massa, cursus ut, gravida ut,
-lobortis eget, lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus.
-Mauris ac felis vel velit tristique imperdiet.
-nec dui dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a,
-aliquet quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet
-tortor. Nam euismod tellus id erat.
+In this section we describe the base ideas behind \mutagen, our CGPT tool
+written in Haskell.
+%
+On top of them, Section \ref{sec:heuristics} introduces heuristics that help
+finding bugs faster in certain testing scenarios.
+
+In constrast with \fuzzchick, \mutagen does not employ a power schedule to
+assign energy to mutable candidates.
+%
+In turn, it resorts to mutate them in an exhaustive and precise basis, where
+%
+\begin{inparaenum}
+  \item each subexpression of a mutation candidate is associated with a set of
+    deterministic mutations, and
+  \item for every mutable subexpression, each of these mutations is evaluated
+    \emph{exactly once}.
+\end{inparaenum}
+%
+There is a small exception to this rules that we will introduce soon.
+
+This idea is inspired by exhaustive bounded testing tools like SmallCheck (in
+Haskell) or Korat (in Java), that produce test cases exhaustively.
+%
+In simple words, such tools work by enumerating all possible values of the
+datatypes used in the testing properties, and then executing them from smaller
+to larger until a a bug is found, or a certain size bound is reached.
+%
+The main problem with this approach is that the space of all possible test cases
+often grows exponentially as we increment the size bound, and the user
+experiences what it looks like ``hitting a wall'', where no larger test cases
+can be evaluated until we exhausted all the immediately smaller ones \cite{feat}.
+%
+In consequence, such tools can only be applied to relatively simple systems,
+where the space of inputs does not grow extremely fast.
+
+Not to be confused by these tools, in \mutagen we do not enumerate all possible
+test cases exhaustively.
+%
+\mutagen uses random generators to find interesting initial seeds, and only then
+proceeds to execute all possible mutations.
+%
+Moreover, these mutants will be mutated further only if they discover new paths
+in the target code, so the testing loop automatically prunes the space of test
+cases that are worth mutating.
+
+\subsection{Exhaustive Mutations}
+
+To describe \mutagen's testing loop, we first need to introduce the mechanism
+used for testing mutations exhaustively.
+%
+In contrast to \fuzzchick, where mutators are parameterized random generators,
+in \mutagen we define mutations as the set of mutants that can be obtained by
+altering the input value at the top-level (the root).
+%
+In Haskell, we define mutations as:
 
@@
 
+Where \texttt{Mutant}s come in two flavours, pure and random:
+
 \begin{verbatim}
@@
 
-\begin{verbatim}
-mutate (Leaf x)     = [ PURE (Node def x def) ]   -- Swap constructor
-mutate (Node l x r) = [ PURE l, PURE r            -- Return subexpression
-                      , PURE (Leaf x)             -- Swap constructor
-                      , PURE (Node l x l)         -- Rearrange subexpressions
-                      , PURE (Node r x r)         -- Rearrange subexpressions
-                      , PURE (Node r x l) ]       -- Rearrange subexpressions
-\end{verbatim}
-
-\begin{verbatim}
-positions (Leaf x)       = node [ (0, positions x) ]
-positions (Branch l x r) = node [ (0, positions l), (1, positions x), (2, positions r) ]
-\end{verbatim}
-
-\begin{verbatim}
-inside []        mut x              = mut x
-inside (0 : pos) mut (Leaf x)       = [ Leaf x'       | x' <- inside pos mut x ]
-inside (0 : pos) mut (Branch l x r) = [ Branch l' x r | l' <- inside pos mut l ]
-inside (1 : pos) mut (Branch l x r) = [ Branch l x' r | x' <- inside pos mut x ]
-inside (2 : pos) mut (Branch l x r) = [ Branch l x r' | r' <- inside pos mut r ]
-\end{verbatim}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-\begin{figure}
-\begin{multicols}{2}
-
-\begin{algorithm}[H]
-  \SetArgSty{textnormal}
-  \SetInd{0em}{0.75em}
-  \SetKw{KwNot}{not}
-  \SetAlgoLined
-  \DontPrintSemicolon
-  \SetKwFunction{MG}{Loop}
-  \SetKwProg{Fn}{Function}{:}{}
-  \Fn{\MG{P, N, R, gen, mut}}{
-    i $\gets$ 0\;
-    TLog, QSucc, QDisc $\gets$ $\varnothing$;\;
-    \While{i $<$ N}{
-      x $\gets$ Pick(QSucc, QDisc, gen, mut)\;
-      (result, trace) $\gets$ WithTrace(P(x))\;
-      \lIf{\KwNot result}{\KwRet Bug(x)}
-      \If{Interesting(TLog, trace)}{
-        \If{\KwNot Discarded(result)}{
-          muts $\gets$ Mutants(x, mut, R)\;
-          Enqueue(QSucc, muts)\;
-        }
-        \uElseIf{Passed(Parent(x))}{
-          muts $\gets$ Mutants(x, mut, R)\;
-          Enqueue(QDisc, muts)\;
-        }
-      }
-      i $\gets$ i+1\;
-    }
-    \KwRet Ok\;
-  }
-\caption{\mutagen Testing Loop}
-\end{algorithm}
-
-\columnbreak
-
-\begin{algorithm}[H]
-  \SetArgSty{textnormal}
-  \SetInd{0em}{0.75em}
-  \SetKw{KwNot}{not}
-  \SetAlgoLined
-  \DontPrintSemicolon
-  \SetKwFunction{Pick}{Pick}
-  \SetKwProg{Fn}{Function}{:}{}
-  \Fn{\Pick{QSucc, QDisc, gen}}{
-    \If{\KwNot Empty(QSucc)}{
-      muts $\gets$ Deque(QSucc)\;
-      \If{Empty(muts)}{
-        Pick(QSucc, QDisc, gen)\;
-      } \uElse{
-        PushFront(QSucc, Rest(muts))\;
-        \KwRet First(muts)\;
-      }
-    }
-    \If{\KwNot Empty(QSucc)}{
-      muts $\gets$ Deque(QSucc)\;
-      \If{Empty(muts)}{
-        Pick(QSucc, QDisc, gen)\;
-      } \uElse{
-        PushFront(QSucc, Rest(muts))\;
-        \KwRet First(muts)\;
-      }
-    }
-    \lElse{\KwRet Sample(gen)}
-  }
-\caption{\mutagen Seed Selection}
-\end{algorithm}
-\end{multicols}
-\end{figure}
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-
-\begin{figure}
-
-\begin{algorithm}[H]
-  \SetArgSty{textnormal}
-  \SetInd{0em}{0.75em}
-  \SetKw{KwNot}{not}
-  \SetKw{KwIn}{in}
-  \SetKwFor{KwRepeat}{repeat}{times}
-  \SetAlgoLined
-  \DontPrintSemicolon
-  \SetKwFunction{Muts}{Mutants}
-  \SetKwProg{Fn}{Function}{:}{}
-  \Fn{\Muts{x, mut, R}}{
-    muts $\gets$ $\varnothing$\;
-    \For{pos \KwIn Flatten(Positions(x))}{
-      \For{mutant \KwIn Inside(pos, mut, x)}{
-        \Switch{mutant}{
-          \Case{PURE $\hat{x}$}{
-            Enqueue($\hat{x}$, muts)\;
-          }
-          \Case{RAND gen}{
-            \KwRepeat{R}{
-              $\hat{x}$ $\gets$ Sample(gen)\;
-              Enqueue($\hat{x}$, muts)\;
-            }
-          }
-        }
-      }
-    }
-    \KwRet muts\;
-  }
-\caption{Mutants Initialization}
-\end{algorithm}
-
-\end{figure}
-
-\section{\mutagen Heuristics}
-\label{sec:heuristics}
-
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo.
-
-\subsection{Priority FIFO Scheduling}
-
-Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
-
-\begin{figure}
-\begin{multicols}{2}
-
-\begin{algorithm}[H]
-  \SetArgSty{textnormal}
-  \SetInd{0em}{0.75em}
-  \SetKw{KwNot}{not}
-  \SetAlgoLined
-  \DontPrintSemicolon
-  \SetKwFunction{MG}{Loop}
-  \SetKwProg{Fn}{Function}{:}{}
-  \Fn{\MG{P, N, R, gen, mut}}{
-    $\cdots$\;
-    (result, trace) $\gets$ WithTrace(P(x))\;
-    $\cdots$\;
-    \If{Interesting(TLog, trace)}{
-      \If{\KwNot Discarded(result)}{
-        muts $\gets$ Mutants(x, mut, R)\;
-        {\color{red} prio $\gets$ BranchDepth(TLog, trace)}\; %% NEW
-        {\color{red} PushFront(QSucc, prio, muts)}\;          %% NEW
-      }
-      $\cdots$\;
-    }
-  }
-  \vspace{10pt}
-  \SetKwFunction{Pick}{Pick}
-  \SetKwProg{Fn}{Function}{:}{}
-  \Fn{\Pick{QSucc, QDisc, gen}}{
-    \If{\KwNot Empty(QSucc)}{
-      {\color{red} (muts, prio) $\gets$ DequeMin(QSucc)}\; %%NEW
-      \If{Empty(muts)}{
-        Pick(QSucc, QDisc, gen)\;
-      } \uElse{
-        {\color{red} PushFront(QSucc, prio, Rest(muts))}\; %% NEW
-        \KwRet First(muts)\;
-      }
-    }
-    $\cdots$\;
-  }
-
-\caption{Priority FIFO Heuristic}
-\end{algorithm}
-
-\columnbreak
-
-\begin{algorithm}[H]
-  \SetArgSty{textnormal}
-  \SetInd{0em}{0.75em}
-  \SetKw{KwNot}{not}
-  \SetAlgoLined
-  \DontPrintSemicolon
-  \SetKwFunction{MG}{Loop}
-  \SetKwProg{Fn}{Function}{:}{}
-  \Fn{\MG{P, N, gen, mut}}{
-    {\color{red}
-    boring  $\gets$ 0\;
-    reset   $\gets$ 1000\;
-    R       $\gets$ 1\;
-    }
-    $\cdots$\;
-    \While{i $<$ N}{
-      {\color{red}
-      \If{boring $>$ reset}{
-        TLog   $\gets$ $\varnothing$\;
-        reset $\gets$ reset * 2\;
-        R     $\gets$ R * 2\;
-      }
-      }
-      $\cdots$\;
-      \lIf{\KwNot result}{\KwRet Bug(x)}
-      \If{Interesting(TLog, trace)}{
-        $\cdots$\;
-      }
-      \lElse {
-        {\color{red} boring $\gets$ boring + 1}
-      }
-      $\cdots$\;
-    }
-  }
-
-\caption{Trace Saturation Heuristic}
-\end{algorithm}
-
-\end{multicols}
-\caption{\label{fig:heuristics} The two main heuristics implemented in \mutagen.
-  Statements in {\color{red} red} indicate important changes to the base
-  algorithm. Ellipses denote parts of the code that are not relevant for the
-  point being made. }
-\end{figure}
-
-\subsection{Detecting Trace Space Saturation And Tuning Random Mutations}
-
-Vestibulum convallis, lorem a tempus semper, dui dui euismod elit, vitae
-placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis, varius
-et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat tellus ut
-neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus lacinia purus,
-et dictum nunc justo sit amet elit.
-
-\subsection{Mutation Inheritance}
-
-Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
-
-\section{Case studies}
-\label{sec:casestudies}
-
-Aliquam erat volutpat \citep{devai2013edsl}. Nunc eleifend leo vitae magna. In id
-erat non orci commodo lobortis. Proin neque massa, cursus ut, gravida ut,
-lobortis eget, lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus.
-Mauris ac felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque
-nec dui dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a,
-aliquet quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet
-tortor. Nam euismod tellus id erat.
-
-\subsection{IFC Stack Machine}
-
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
-
-\subsection{WebAssembly Engine}
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
-\subsubsection{Testing the WebAssembly Validator}
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
-\begin{verbatim}
-prop_validator m =
-  isValidHaskell m ==> isValidSpec m
-\end{verbatim}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-\begin{verbatim}
-isValidHaskell m =
-  case validate m of
-    Left  validationError -> return False
-    Right validModule     -> return True
-\end{verbatim}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-\begin{verbatim}
-isValidSpec m = do
-  writeFile "testcase.wasm" (dumpModule m)
-  res <- shell "./wasm" ["-d", "-i", "testcase.wasm"]
-  return (res == "")
-\end{verbatim}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-\subsubsection{Testing the WebAssembly Interpreter}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-\begin{verbatim}
-mk_module ty name fun =
-  emptyModule { types     = [ ty ]
-              , functions = [ fun ]
-              , exports   = [ Export name (ExportFunc 0) ]
-              , mems      = [ Memory (Limit 1 Nothing) ]
-              }
-\end{verbatim}
-
-\begin{verbatim}
-prop_interpreter ty (args, fun) =
-  discardAfter 20
-  (do let m = mk_module ty "f" fun;
-      resHs   <- invokeHaskell m "f" args;
-      resSpec <- invokeSpec    m "f" args;
-      return (equiv resHs resSpec))
-\end{verbatim}
-
-\begin{verbatim}
-prop_interpreter_i32 (i, f, fun) =
-  prop_interpreter
-    (FuncType { params = [I32, F32], result = [I32]})
-    ([VI32 i, VF32 f], fun)
-\end{verbatim}
-
-\subsubsection{Real Bugs and Discrepancies found by \mutagen in \textit{haskell-wasm}}
-
-\paragraph{Bug \#1: Invalid Memory Alignment Validation}
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-
-\paragraph{Bug \#2: Lax invocation checks}
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-\paragraph{Bug \#3: Allowed Out-of-bounds Memory Access}
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
-
-\paragraph{Discrepancy \#1: Different Reinterpretation Semantics of NaN Values}
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-
-\paragraph{Discrepancy \#2: Allowed Blocks with Multiple Return Types}
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-
-\section{Evaluation}
-\label{sec:evaluation}
-
-\subsection{IFC Stack Machine}
-
-Aliquam erat volutpat \citep{devai2013edsl}. Nunc eleifend leo vitae magna. In
-id erat non orci commodo lobortis. Proin neque massa, cursus ut, gravida ut,
-lobortis eget, lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus.
-Mauris ac felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque
-nec dui dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a,
-aliquet quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet
-tortor. Nam euismod tellus id erat.
-
-\begin{center}
-  \includegraphics[width=\linewidth]{tikz/ifc-mttf.tikz}
-  \includegraphics[width=\linewidth]{tikz/ifc-fr.tikz}
-\end{center}
-
-\subsection{Webassembly Engine}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-\begin{center}
-  \includegraphics[width=\linewidth]{tikz/wasm-mttf.tikz}
-  \includegraphics[width=\linewidth]{tikz/wasm-fr.tikz}\vspace{-5pt}
-  \includegraphics[width=\linewidth]{tikz/wasm-legend.tikz}
-\end{center}
-
-\section{Implementation}
-\label{sec:implementation}
-
-\subsection{Deriving Mutation Machinery}
-
-Integer placerat tristique nisl. Praesent augue. Fusce commodo. Vestibulum
-convallis, lorem a tempus semper, dui dui euismod elit, vitae placerat urna
-tortor vitae lacus. Nullam libero mauris, consequat quis, varius et, dictum id,
-arcu. Mauris mollis tincidunt felis. Aliquam feugiat tellus ut neque. Nulla
-facilisis, risus a rhoncus fermentum, tellus tellus lacinia purus, et dictum
-nunc justo sit amet elit.
-
-\subsection{Tracing Haskell Programs}
-
-\begin{verbatim}
-  sorted []       = True
-  sorted [x]      = True
-  sorted (x:y:xs) = if x <= y then sorted (y:xs) else False
-\end{verbatim}
-
-\begin{verbatim}
-  sorted []       = _trace_ 1 (True)
-  sorted [x]      = _trace_ 2 (True)
-  sorted (x:y:xs) = _trace_ 3 (if x <= y then _trace_ 4 (sorted (y:xs)) else _trace_ 5 (False))
-\end{verbatim}
-
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies.
-
-\section{Related work}
-\label{sec:related}
-
-\subsection{Grammar-based Fuzzing}
-
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
-
-\subsection{Random Data Generation}
-
-%% Don't forget to cite "Deriving good generators ..."
-
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
-
-\subsection{Exhaustive Bounded Testing}
-
-Aliquam erat volutpat \citep{devai2013edsl}. Nunc eleifend leo vitae magna. In id
-erat non orci commodo lobortis. Proin neque massa, cursus ut, gravida ut,
-lobortis eget, lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus.
-Mauris ac felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque
-nec dui dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a,
-aliquet quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet
-tortor. Nam euismod tellus id erat.
-
-\section{Conclusions}
-\label{sec:conclusions}
-
-Aliquam erat volutpat \citep{devai2013edsl}. Nunc eleifend leo vitae magna. In id
-erat non orci commodo lobortis. Proin neque massa, cursus ut, gravida ut,
-lobortis eget, lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus.
-Mauris ac felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque
-nec dui dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a,
-aliquet quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet
-tortor. Nam euismod tellus id erat.
-
-% ----------------------------------------
-
-%% The acknowledgments section is defined using the "acks" environment (and NOT
-%% an unnumbered section). This ensures the proper identification of the section
-%% in the article metadata, and the consistent spelling of the heading.
-
-%% \begin{acks}
-%% To Robert, for the bagels and explaining CMYK and color spaces.
-%% \end{acks}
-
-% ----------------------------------------
-
-%% The next two lines define the bibliography style to be used, and the
-%% bibliography file.
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{references}
-
-% ----------------------------------------
-
-%% If your work has an appendix, this is the place to put it.
-\appendix
-
-\section{Detailed Empirical Results}
-
-Fusce suscipit, wisi nec facilisis facilisis, est dui fermentum leo, quis tempor
-ligula erat quis odio. Nunc porta vulputate tellus. Nunc rutrum turpis sed pede.
-Sed bibendum. Aliquam posuere. Nunc aliquet, augue nec adipiscing interdum,
-lacus tellus malesuada massa, quis varius mi purus non odio. Pellentesque
-condimentum, magna ut suscipit hendrerit, ipsum augue ornare nulla, non luctus
-diam neque sit amet urna. Curabitur vulputate vestibulum lorem. Fusce sagittis,
-libero non molestie mollis, magna orci ultrices dolor, at vulputate neque nulla
-lacinia eros. Sed id ligula quis est convallis tempor. Curabitur lacinia
-pulvinar nibh. Nam a sapien.
+Pure mutants are used most of the time, and encode simple deterministic
+transformations over the top-level of the input.
+%
+Like in \fuzzchick, these tranformation can:
 
 \end{document}
-\endinput
-%%
-%% End of `main.tex'.

### 2021-03-23 papers/oopsla-21/main.tex
WIP

@@
 \pgfplotsset{compat=1.3}
+
+\usetikzlibrary{trees}
 
@@
 %
-In simple terms, these operations encompass:
-%
-\begin{itemize}
+In simple terms, these operations encompass
+%
+\begin{inparaenum}
 \item \emph{shrinking the value}, replacing its top-level data constructor with
@@
   and generating random ones when needed;
-\item \emph{returning a subexpression of the same type};
+\item \emph{returning a subexpression of the same type}; and
 \item \emph{mutating recursively}, applying a mutation operation over an
   immediate subexpression.
-\end{itemize}
+\end{inparaenum}
 
@@
 
-Fig. \ref{fig:fuzzchick:mutator} illustrates the idea behind a random \fuzzchick
-mutator for our previously used \texttt{Tree} data type example.
+Fig. \ref{fig:fuzzchick:mutator} illustrates a \fuzzchick mutator for our
+previously used \texttt{Tree} data type example.
 %
 Since trees are parametric, for clarity this definition is also parameterized by
-a mutator for the payload (\texttt{mutate\_a}), although in practice this can be
-abstracted away by the means of the type class system (where
-\texttt{mutate\_tree} and \texttt{mutate\_a} can be transparently coerced into
-an overloaded \texttt{mutate} function.)
+a mutator for the payload (\texttt{mutate\_a}), although this can be abstracted
+away using the type class system.
+
+%% (where \texttt{mutate\_tree} and
+%% \texttt{mutate\_a} can be transparently coerced into an overloaded
+%% \texttt{mutate} function.)
 
@@
 \citeauthor{lampropoulos2019coverage} demonstrated empirically that \fuzzchick
-lies comfortably in the middle ground between using na\"ive automatically
-derived random generators and complex manually-written ones.
+lies comfortably in the middle ground between using pure random testing with
+na\"ive automatically derived random generators and complex manually-written
+ones.
 %
 Their results suggest that CGPT is an appealing technique for finding bugs while
-still offering a mostly automated workflow.
+still using a mostly automated workflow.
 
@@
 
+\noindent Where \texttt{Mutant}s come in two flavours, pure and random:
+
 \begin{verbatim}
@@
 
+Pure mutants are used most of the time, and encode simple deterministic
+transformations over the top-level data constructor of the input --- recursive
+mutations will be introduced soon.
+%
+These tranformation can:
+%
+\begin{inparaenum}
+\item return an immediate subexpression of the same type as the input;
+\item swap the top-level data constructor with any other constructor of the same
+  type, reusing existing subexpressions; and
+\item rearrange and replace fields using existing ones of the same type.
+\end{inparaenum}
+%
+To illustrate this, Fig. \ref{fig:mutagen:mutator} outlines a mutator for the
+\texttt{Tree} data type.
+%
+Notice how this definition simply enumerates mutants that transform the
+top-level data constructor, hence no recursion is needed here.
+%
+Moreover, notice how a default value (\texttt{def}) used to fill the subtrees
+when transforming a leaf into a branch.
+%
+This value corresponds to the smallest value we can construct in order for the
+mutant to be type-correct.
+%
+In practice (\texttt{def = Leaf def}), where the inner \texttt{def} is the
+smallest value of the payload --- we use the type class system to abstract this
+complexity away in our implementation.
+%
+Using a default smallest value is again inspired by exhaustive bounded testing
+tools, and avoids introducing unnecesary randomness when growing data
+constructors.
+
+\begin{figure}
 \begin{verbatim}
@@
 \end{verbatim}
+\caption{\label{fig:mutagen:mutator}\mutagen mutator for the \texttt{Tree} data
+  type.}
+\end{figure}
+
+At this point, the reader might consider:
+%
+what about large enumeration types like integers or characters?
+%
+Does \mutagen transform them exaustively too?
+%
+\emph{Certainly not.}
+%
+Random mutants serve as a way to break exhaustiveness when mutating values of
+large enumeration types.
+%
+Instead of trying every possible integer or character, we resolve in using a
+random generator and sample a small amount of values from it --- the precise
+amount is a tunable parameter of \mutagen.
+%
+This way, a mutator for integers simply becomes:
+
+\begin{verbatim}
+mutate n = [ RAND arbitrary ]
+\end{verbatim}
+
+\subsubsection{Mapping top-level mutations everywhere}
+
+So far we defined mutations that transform the root of the input.
+%
+Now it is time to apply these mutations to every subexpression as well.
+%
+To do so, we will use two functions that can also be derived from the data type
+definition.
+
+In first place, a function \texttt{positions} traverses the mutable candidate
+and builds a multi-way tree of mutable positions.
+%
+These positions are essentially a list of indices encoding the path from the
+root to every mutable subexpression.
+%
+For instance, the positions of a \texttt{Tree} are computed as follows:
 
@@
 
-\begin{verbatim}
-inside []        mut x              = mut x
-inside (0 : pos) mut (Leaf x)       = [ Leaf x'       | x' <- inside pos mut x ]
-inside (0 : pos) mut (Branch l x r) = [ Branch l' x r | l' <- inside pos mut l ]
-inside (1 : pos) mut (Branch l x r) = [ Branch l x' r | x' <- inside pos mut x ]
-inside (2 : pos) mut (Branch l x r) = [ Branch l x r' | r' <- inside pos mut r ]
-\end{verbatim}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
+\noindent In this light, the mutable positions of the value \texttt{Branch (Leaf 1) 2 (Leaf 3)} are:
+%
+\begin{equation}
+  \nonumber
+  {\small \texttt{positions}}\quad
+  \left(
+  \tikz[
+    baseline=-7mm,
+    level distance=0.6cm,
+    level 1/.style={sibling distance=0.75cm},
+    level 2/.style={sibling distance=0.75cm}
+  ]{
+    \node {\small \texttt{Branch}}
+      child {node {\small \texttt{Leaf}}
+        child {node {\small \texttt{1}}}
+      }
+      child {node {\small \texttt{2}}}
+      child {node {\small \texttt{Leaf}}
+        child {node {\small \texttt{3}}}
+      };
+  }
+  \right)
+  \quad
+  =
+  \quad
+  \tikz[
+    baseline=-7mm,
+    level distance=0.6cm,
+    level 1/.style={sibling distance=0.75cm},
+    level 2/.style={sibling distance=0.75cm}
+  ]{
+    \node {\texttt{\small []}}
+      child {node {\small \texttt{[0]}}
+        child {node {\small \texttt{[0,0]}}}
+      }
+      child {node {\small \texttt{[1]}}}
+      child {node {\small \texttt{[2]}}
+        child {node {\small \texttt{[2,0]}}}
+      };
+  }
+\end{equation}
+
+\noindent Later, given a desired position to mutate within a candidate, we
+define another function \texttt{inside} that precisely finds the subexpression
+corresponding to it and applies the mutation.
+%
+A slightly simplified version of this function for the \texttt{Tree} datatype is
+as follows:
+
+\begin{verbatim}
+inside []        x              = mutate x
+inside (0 : pos) (Leaf x)       = [ Leaf x'       | x' <- inside pos x ]
+inside (0 : pos) (Branch l x r) = [ Branch l' x r | l' <- inside pos l ]
+inside (1 : pos) (Branch l x r) = [ Branch l x' r | x' <- inside pos x ]
+inside (2 : pos) (Branch l x r) = [ Branch l x r' | r' <- inside pos r ]
+\end{verbatim}
+
+\noindent This function simply traverses the desired position, calling itself
+recursively until it reaches the desired subexpression, where the mutation can
+be applied locally at the top-level (case \texttt{inside [] x}).
+%
+The rest of the function takes care of unwrapping and rewrapping the
+intermediate subexpressions and is not partically relevant for the point being
+made.
+
+\subsection{Testing loop}
+
+Having the mutation mechanism in place, we can finally sketch the base testing
+loop used by \mutagen.
+%
 
@@
         \uElseIf{Passed(Parent(x))}{
-          muts $\gets$ Mutants(x, mut, R)\;
+          muts $\gets$ Mutate(x, mut, R)\;
           Enqueue(QDisc, muts)\;
@@
   \DontPrintSemicolon
-  \SetKwFunction{Muts}{Mutants}
+  \SetKwFunction{Muts}{Mutate}
   \SetKwProg{Fn}{Function}{:}{}
@@
         \Switch{mutant}{
-          \Case{PURE $\hat{x}$}{
+          \Case{\texttt{PURE} $\hat{x}$}{
             Enqueue($\hat{x}$, muts)\;
           }
-          \Case{RAND gen}{
+          \Case{\texttt{RAND} gen}{
             \KwRepeat{R}{
@@
 \end{figure}
+
+The next section introduces three heurisitics we added to the base testing loop
+of \mutagen and can that offer a substantial improvement in practice.
 

### 2021-03-26 papers/oopsla-21/main.tex
WIP

@@
 
+\usepackage{wrapfig}
 \usepackage{xcolor}
@@
 
-Having the mutation mechanism in place, we can finally sketch the base testing
-loop used by \mutagen.
-%
-
-\begin{figure}
+Having the exhaustive mutation mechanism in place, we can finally introduce the
+base testing loop used by \mutagen.
+%
+This is outlined in Algorithm \ref{algo:mutagen:loop}.
+%
+As it can be observed, we closely follow \fuzzchick's testing loop, using qtwo
+queues to keep mutant candidates, and enqueueing and retrieving interesting test
+cases from them before falling back to random generation.
+
+\begin{wrapfigure}{r}{0.5\textwidth}
+\vspace{-10pt}
+\begin{algorithm}[H]
+  \SetArgSty{textnormal}
+  \SetInd{0em}{0.75em}
+  \SetKw{KwNot}{not}
+  \SetKw{KwIn}{in}
+  \SetKwFor{KwRepeat}{repeat}{times}
+  \SetAlgoLined
+  \DontPrintSemicolon
+  \SetKwFunction{Muts}{Mutate}
+  \SetKwProg{Fn}{Function}{:}{}
+  \Fn{\Muts{x, mut, R}}{
+    muts $\gets$ $\varnothing$\;
+    \For{pos \KwIn Flatten(Positions(x))}{
+      \For{mutant \KwIn Inside(pos, mut, x)}{
+        \Switch{mutant}{
+          \Case{\texttt{PURE} $\hat{x}$}{
+            Enqueue($\hat{x}$, muts)\;
+          }
+          \Case{\texttt{RAND} gen}{
+            \KwRepeat{R}{
+              $\hat{x}$ $\gets$ Sample(gen)\;
+              Enqueue($\hat{x}$, muts)\;
+            }
+          }
+        }
+      }
+    }
+    \KwRet muts\;
+  }
+\caption{\label{algo:mutagen:init}Mutants Initialization}
+\end{algorithm}
+\vspace{-10pt}
+\end{wrapfigure}
+
+The main difference lies in that we precompute all the mutations of a given
+mutation candidate before enqueueing them.
+%
+These mutations are put together into lists that we call \emph{mutation batches}
+--- one for each mutation candidate.
+%
+To initialize a mutation batch (outlined in Algorithm \ref{algo:mutagen:init}),
+we first flatten all the mutable positions of the input value in level order
+(recall that positions are stored as a multi-way tree).
+%
+Then, we iterate over all of them and retrieve all the mutants defined for each
+subexpression.
+%
+For each one of these, there are two posible cases:
+%
+\begin{inparaenum}
+\item if it is a pure mutant carrying a concrete mutated value, we enqueue it
+  into the mutation batch directly; otherwise
+\item it is a random mutant that carries a random generator with it (e.g., a
+  numeric subexpression), in which case we sample and enqueue \textit{R} random
+  values using this generator, where \textit{R} is a parameter set by the user.
+\end{inparaenum}
+%
+At the end, we simply return the accummulated batch.
+
+Finally, the seed selection algorithm (Algorithm \ref{algo:mutagen:pick}) simply
+selects the next test case using the same criteria as \fuzzchick, prioritizing
+valid candidates over discarded ones, falling back to random generation when
+both queues are empty.
+%
+Since mutations are precomputed, this function only needs to pick the next test
+case from the current batch, until it becomes empty and can switch to the next
+precomputed one in line.
+
+\begin{figure}[b]
 \begin{multicols}{2}
@@
   }
-\caption{\mutagen Testing Loop}
+\caption{\label{algo:mutagen:loop}\mutagen Testing Loop}
 \end{algorithm}
@@
   }
-\caption{\mutagen Seed Selection}
+\caption{\label{algo:mutagen:pick}\mutagen Seed Selection}
+\vspace{1pt}
 \end{algorithm}
@@
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-
-\begin{figure}
-
-\begin{algorithm}[H]
-  \SetArgSty{textnormal}
-  \SetInd{0em}{0.75em}
-  \SetKw{KwNot}{not}
-  \SetKw{KwIn}{in}
-  \SetKwFor{KwRepeat}{repeat}{times}
-  \SetAlgoLined
-  \DontPrintSemicolon
-  \SetKwFunction{Muts}{Mutate}
-  \SetKwProg{Fn}{Function}{:}{}
-  \Fn{\Muts{x, mut, R}}{
-    muts $\gets$ $\varnothing$\;
-    \For{pos \KwIn Flatten(Positions(x))}{
-      \For{mutant \KwIn Inside(pos, mut, x)}{
-        \Switch{mutant}{
-          \Case{\texttt{PURE} $\hat{x}$}{
-            Enqueue($\hat{x}$, muts)\;
-          }
-          \Case{\texttt{RAND} gen}{
-            \KwRepeat{R}{
-              $\hat{x}$ $\gets$ Sample(gen)\;
-              Enqueue($\hat{x}$, muts)\;
-            }
-          }
-        }
-      }
-    }
-    \KwRet muts\;
-  }
-\caption{Mutants Initialization}
-\end{algorithm}
-
-\end{figure}
+Another small difference between \mutagen and \fuzzchick is the criteria for
+enqueuing discarded tests.
+%
+We found that, especially for large data types, the queue of discarded
+candidates tends to grow disproportionately during testing, making them hardly
+usable and consuming large amounts of memory.
+%
+To improve this, we resort to mutate discarded tests cases only when we have
+some evidence that they are ``almost valid.''
+%
+For this, each mutated test case remembers whether its parent (the original test
+case they derive from after being mutated) was valid.
+%
+Then, we enqueue discarded test cases only if they meet this condition.
+%
+As a result, we fill the discarded queue with lesser but much more interesting
+mutation candidates.
+%
+Moreover, this can potentially help introducing \emph{2-step mutations}, where
+an initial mutation breaks a valid test case in a small way, it gets enqueued as
+discarded, and later a subsequent mutation fixes it by changing a different
+subexpression.
 
 The next section introduces three heurisitics we added to the base testing loop
-of \mutagen and can that offer a substantial improvement in practice.
+of \mutagen based on the limitations we found in \fuzzchick.
+%
+These heuristics can that offer a substantial improvement in practice.
 

### 2021-03-29 papers/oopsla-21/main.tex
WIP

@@
 
+% ----------------------------------------
+
 \section{Background}
@@
 
+% ----------------------------------------
+
 \section{\mutagen: Testing Mutants Exhaustively}
@@
 
-The next section introduces three heurisitics we added to the base testing loop
-of \mutagen based on the limitations we found in \fuzzchick.
-%
-These heuristics can that offer a substantial improvement in practice.
+The next section introduces two heurisitics we added to the base testing loop of
+\mutagen based on the limitations we found in \fuzzchick.
+%
+%% These heuristics can that offer a substantial improvement in practice.
+
+% ----------------------------------------
 
@@
 
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo.
+In this section we introduce two heuristics implemented on top of the base
+testing loop of our tool.
+%
+\mutagen enables them all by default, although they can be individually disabled
+by the user if deemed appropriate.
 
@@
 
-Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
+This heuristic tackles the issue of enqueuing novel mutation candidates at the
+end of (possibly) long queues of not-so-interesting previously executed ones.
+
+\fuzzchick uses AFL instrumentation under the hood, which in turn uses an
+\emph{edge coverage} criteria to distinguish novel executions and to assign each
+mutation candidate a given energy.
+%
+In contrast, execution traces \mutagen represent the \emph{path} in the code
+taken by the program, as opposed to just the set of edges traversed in the
+control-flow graph (CFG).
+%
+Using this criteria lets us gather precise information from the each new
+execution.
+%
+%% To keep track of the execution path of each test, our tool keeps an internal
+%% prefix tree that records every new entry by traversing this data structure and
+%% inserting new nodes after a new path prefix gets discovered.
+%% %
+In particular, we are interested in the \emph{depth} where each new execution
+branches from already seen ones.
+%
+Our assumption here is that test cases that differ (branch) at shallower depths
+from the ones already executed are more likely to discover completely new
+portions of the code under test, and hence we want to assign them a higher
+priority.
+
+In this light, every time we insert a new execution path into the internal trace
+log, we calculate the number of new nodes that were executed, as well as the
+\emph{branching depth} where they got inserted.
+%
+The former is used to distinguish interesting test cases (whether or not new
+nodes were inserted), whereas the latter is used by this heuristic to schedule
+mutation candidates.
+%
+Fig. \ref{fig:mutagen:tracelog} illustrates this idea, inserting two execution
+traces (one after another) into a trace log that initially contains a single
+execution path.
+%
+The second insertion (with trace $1 \rightarrow 2 \rightarrow 6 \rightarrow 7$)
+branches at a shallower depth than the first one (2 vs. 3), hence its
+corresponding test case should be given a higher priority.
+
+\begin{figure}[b]
+\newcommand\stacked[2]{\underset{\textstyle #2\mathstrut}{#1}}
+\begin{equation}
+  \nonumber
+  \left[
+  \tikz[
+    baseline=-10mm,
+    level distance=0.6cm,
+    level 1/.style={sibling distance=0.75cm},
+    level 2/.style={sibling distance=0.75cm}
+  ]{
+    \node {\small \texttt{1}}
+      child {node {\small \texttt{2}}
+        child {node {\small \texttt{3}}
+          child {node {\small \texttt{4}}}
+        }
+      }
+  }
+  \right]
+  \;
+  +
+  \;
+  \tikz[
+    baseline=-10mm,
+    level distance=0.6cm,
+    level 1/.style={sibling distance=0.75cm},
+    level 2/.style={sibling distance=0.75cm}
+  ]{
+    \node {\small \texttt{1}}
+      child {node {\small \texttt{2}}
+        child {node {\small \texttt{3}}
+          child {node {\small \texttt{5}}}
+        }
+      }
+  }
+  \;
+  =
+  \;
+  \left[
+  \tikz[
+    baseline=-10mm,
+    level distance=0.6cm,
+    level 1/.style={sibling distance=0.75cm},
+    level 2/.style={sibling distance=0.75cm}
+  ]{
+    \node {\color{red} \small \texttt{1}}
+      child {node {\color{red} \small \texttt{2}}
+        child {node {\color{red} \small \texttt{3}}
+          child {node {\small \texttt{4}}}
+          child {node {\color{blue} \small \texttt{5}}}
+        }
+      }
+  }
+  \right]
+  ,
+  \stacked{{\color{blue} new=1\phantom{x}}}{{\color{red} depth=3}}
+  \qquad
+  \qquad
+  \left[
+  \tikz[
+    baseline=-10mm,
+    level distance=0.6cm,
+    level 1/.style={sibling distance=0.75cm},
+    level 2/.style={sibling distance=0.75cm}
+  ]{
+    \node {\small \texttt{1}}
+      child {node {\small \texttt{2}}
+        child {node {\small \texttt{3}}
+          child {node {\small \texttt{4}}}
+          child {node {\small \texttt{5}}}
+        }
+      }
+  }
+  \right]
+  \;
+  +
+  \;
+  \tikz[
+    baseline=-10mm,
+    level distance=0.6cm,
+    level 1/.style={sibling distance=0.75cm},
+    level 2/.style={sibling distance=0.75cm}
+  ]{
+    \node {\small \texttt{1}}
+      child {node {\small \texttt{2}}
+        child {node {\small \texttt{6}}
+          child {node {\small \texttt{7}}
+          }
+        }
+      }
+  }
+  \;
+  =
+  \;
+  \left[
+  \tikz[
+    baseline=-10mm,
+    level distance=0.6cm,
+    level 1/.style={sibling distance=0.75cm},
+    level 2/.style={sibling distance=1cm}
+  ]{
+    \node {\color{red} \small \texttt{1}}
+      child {node {\color{red} \small \texttt{2}}
+        child {node {\small \texttt{3}}
+          child {node {\small \texttt{4}}}
+          child {node {\small \texttt{5}}}
+        }
+        child {node {\color{blue}\small \texttt{6}}
+          child {node {\color{blue}\small \texttt{7}}}
+        }
+      }
+  }
+  \right]
+  ,
+  \stacked{{\color{blue} new=2\phantom{x}}}{{\color{red} depth=2}}
+\end{equation}
+\caption{\label{fig:mutagen:tracelog}Inserting two new execution traces into the
+  internal trace log (represented using brackets).}
+\end{figure}
+
+With this mechanism in place, we can modify \mutagen's base testing loop by
+replacing each mutation queue with a priority queue indexed by the branching
+depth of each new execution.
+%
+These changes are illustrated in Algorithm \ref{algo:mutagen:fifo}.
+%
+Statements in {\color{red} red} indicate important changes to the base
+algorithm, whereas ellipses denote parts of the code that are not relevant for
+the point being made.
+
+To pick the next test case, we simply retrieve the one with highest priority
+(the smallest branching depth).
+%
+Then, whenever we find a new interesting test case, we enqueue it at the
+beginning of the queue of its corresponding priority.
+%
+This allows the testing loop to jump immediately onto processing new interesting
+candidates as soon as they are found (even at the same priority), and to jump
+back to previous candidates as soon as their mutants become progressively less
+novel.
 
@@
     $\cdots$\;
+    x $\gets$ Pick(QSucc, QDisc, gen, mut)\;
     (result, trace) $\gets$ WithTrace(P(x))\;
@@
   }
-  \vspace{10pt}
+  \vspace{5pt}
   \SetKwFunction{Pick}{Pick}
@@
 
-\caption{Priority FIFO Heuristic}
+\caption{\label{algo:mutagen:fifo}Priority FIFO Heuristic}
 \end{algorithm}
@@
       \If{Interesting(TLog, trace)}{
+        {\color{red} boring  $\gets$ 0}\;
         $\cdots$\;
@@
 
-\caption{Trace Saturation Heuristic}
+\caption{\label{algo:mutagen:reset}Trace Saturation Heuristic}
 \end{algorithm}
@@
 \end{multicols}
-\caption{\label{fig:heuristics} The two main heuristics implemented in \mutagen.
-  Statements in {\color{red} red} indicate important changes to the base
-  algorithm. Ellipses denote parts of the code that are not relevant for the
-  point being made. }
+%% \caption{\label{fig:heuristics} The two main heuristics implemented in \mutagen.
+%%   Statements in {\color{red} red} indicate important changes to the base
+%%   algorithm. Ellipses denote parts of the code that are not relevant for the
+%%   point being made. }
 \end{figure}
@@
 
-Vestibulum convallis, lorem a tempus semper, dui dui euismod elit, vitae
-placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis, varius
-et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat tellus ut
-neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus lacinia purus,
-et dictum nunc justo sit amet elit.
-
-\subsection{Mutation Inheritance}
-
-Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
+As introduced in Section \ref{sec:mutagen}, our tool is parameterized by the
+amount of random mutations to be generated over each mutable subexpression
+defined using a random mutant, e.g., numeric values, characters, etc.
+%
+But, how many random mutations should we use? A single one? A few tenths? A few
+hundreds?
+%
+Clearly, using too little, can put finding bugs at risk.
+%
+For instance, when the system under test branches on numeric values, we should
+make sure that we set enough random mutations to test each case in a reasonable
+basis.
+%
+Using too many, the other hand, can degrade the performance of the testing
+loop, as it will spend too much time producing uninteresting mutations.
+%
+This can happen for instance if the subexpressions defined using random mutants
+are only used as payloads, and their value does not affect the execution in any
+way.
+
+Answering this question precisely is not an easy task, and the second heuristic
+we introduce in this work aims to tackle this issue.
+%
+We found that, the smaller the number of random mutations we set, the easier it
+is for the trace log that records executions to start getting saturated, i.e.,
+when interesting test cases stop getting discovered or are discovered very
+sporadically.
+%
+Our realization is that we can use this information to automatically optimize
+the number of random mutations used by our tool.
+%
+This idea is described in Algorithm \ref{algo:mutagen:reset}.
+%
+The process is simple:
+%
+\begin{inparaenum}
+  \item we start the testing loop with the amount of random mutations set to one,
+  %
+  \item each time we find that a test is not interesting (i.e. boring), we
+    increment a counter,
+  \item if we have not produced any interesting test case after a certain
+    threshold (1000 tests seems to be a reasonable value in practice), we
+    increment the amount of random mutations and the threshold by twice the
+    current amount.
+    %
+    Additionally, each time this happens we also reset the trace log, so
+    interesting test cases found on a previous iteration can be found and
+    enqueued for mutation again --- this time with a higher effort dedicated to
+    producing random mutations.
+\end{inparaenum}
+
+Then, if the execution of the system under test depends heavily on the values
+stored at randomly mutable subexpressions, starting with a single random mutation
+will quickly saturate the trace space, and this heuristic will continuosly
+increase the random mutations parameter until that stops happening.
+
+%% \subsection{Mutation Inheritance}
+
+%% The last heuristic we discuss in this work attempts to reduce the amount of
+%% mutations required when testing properties that lazily consume foldable inputs
+%% like lists or trees.
+%% %
+%% To illustrate this, suppose we define a property that aims to test a list
+%% sorting algorithm:
+
+% ----------------------------------------
 
@@
 
-\section{Implementation}
-\label{sec:implementation}
-
-\subsection{Deriving Mutation Machinery}
-
-Integer placerat tristique nisl. Praesent augue. Fusce commodo. Vestibulum
-convallis, lorem a tempus semper, dui dui euismod elit, vitae placerat urna
-tortor vitae lacus. Nullam libero mauris, consequat quis, varius et, dictum id,
-arcu. Mauris mollis tincidunt felis. Aliquam feugiat tellus ut neque. Nulla
-facilisis, risus a rhoncus fermentum, tellus tellus lacinia purus, et dictum
-nunc justo sit amet elit.
-
-\subsection{Tracing Haskell Programs}
-
-\begin{verbatim}
-  sorted []       = True
-  sorted [x]      = True
-  sorted (x:y:xs) = if x <= y then sorted (y:xs) else False
-\end{verbatim}
-
-\begin{verbatim}
-  sorted []       = _trace_ 1 (True)
-  sorted [x]      = _trace_ 2 (True)
-  sorted (x:y:xs) = _trace_ 3 (if x <= y then _trace_ 4 (sorted (y:xs)) else _trace_ 5 (False))
-\end{verbatim}
-
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies.
+% ----------------------------------------
+
+%% \section{Implementation}
+%% \label{sec:implementation}
+
+%% \subsection{Deriving Mutation Machinery}
+
+%% Integer placerat tristique nisl. Praesent augue. Fusce commodo. Vestibulum
+%% convallis, lorem a tempus semper, dui dui euismod elit, vitae placerat urna
+%% tortor vitae lacus. Nullam libero mauris, consequat quis, varius et, dictum id,
+%% arcu. Mauris mollis tincidunt felis. Aliquam feugiat tellus ut neque. Nulla
+%% facilisis, risus a rhoncus fermentum, tellus tellus lacinia purus, et dictum
+%% nunc justo sit amet elit.
+
+%% \subsection{Tracing Haskell Programs}
+
+%% \begin{verbatim}
+%%   sorted []       = True
+%%   sorted [x]      = True
+%%   sorted (x:y:xs) = if x <= y then sorted (y:xs) else False
+%% \end{verbatim}
+
+%% \begin{verbatim}
+%%   sorted []       = _trace_ 1 (True)
+%%   sorted [x]      = _trace_ 2 (True)
+%%   sorted (x:y:xs) = _trace_ 3 (if x <= y then _trace_ 4 (sorted (y:xs)) else _trace_ 5 (False))
+%% \end{verbatim}
+
+%% Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
+%% tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
+%% leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
+%% wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
+%% mollis ultricies.
+
+% ----------------------------------------
 
@@
 
+% ----------------------------------------
+
 \section{Conclusions}

### 2021-03-30 papers/oopsla-21/main.tex
WIP

@@
 
-\begin{algorithm}[ b]
+\begin{wrapfigure}{r}{0.5\textwidth}
+\vspace{-10pt}
+\begin{algorithm}[H]
   \SetKw{KwNot}{not}
@@
 \end{algorithm}
+\vspace{-10pt}
+\end{wrapfigure}
 
@@
 %
-Such data structure can be easily defined in Haskell using a custom data type
-with two data constructors for leaves and branches respectively:
+Such data structure can be defined in Haskell using a custom data type with two
+data constructors for leaves and branches respectively:
 
@@
 
-\noindent Should we mistakenly implement \texttt{mirror}, e.g., by dropping the
-right subtree altogether:
+\noindent Should we mistakenly introduce a bug in \texttt{mirror}, e.g., by
+dropping the right subtree altogether:
 
@@
 
-Aliquam erat volutpat \citep{devai2013edsl}. Nunc eleifend leo vitae magna. In id
-erat non orci commodo lobortis. Proin neque massa, cursus ut, gravida ut,
+We evaluated the performance of \mutagen using two main case studies.
+%
+In both of them, we show how the heuristics described in Section
+\ref{sec:heuristics} affect the testing performance.
+
+The first one is a simple abstract stack machine that enforces
+\emph{noninterference} using runtime checks.
+%
+The implementation of this case study was originally proven correct by \cite{}
+in Coq, and subsequently degraded by systematically introducing 20 bugs on its
+enforcing mechanism.
+%
+\citeauthor{lampropoulos2019coverage} used this same case study to compare
+\fuzzchick against random testing approaches using na\"ive and manually-written
+smart random generators.
+%
+In this work we replicate their results and compare them against our tool.
+
+The second case study aims to evaluate \mutagen in a more realistic scenario,
+and focuses on testing \textit{haskell-wasm} an existing WebAssembly engine
+written in Haskell of industrial strength.
+%
+We manually injected 10 bugs in the validator as well as 5 bugs in the
+interpreter of this engine, and used the reference implementation to find them
+using differential testing.
+%
+During this process, we quickly discovered 3 bugs and 2 discrepancies on this
+engine with respect to the reference implementation.
+%
+All of them were reported and confirmed by the authors of \textit{haskell-wasm}.
+
+\subsection{IFC Stack Machine}
+
+Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
+facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
+vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
+Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
+varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
+ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
+vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
+orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
+est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
+
+\subsection{WebAssembly Engine}
+
+WebAssembly is a popular assembly-like language designed to be an open standard
+for executing low-level code in web pages, although it has become increasingly
+popular in standalone, non-web contexts as well.
+%
+WebAssembly programs are first validated and later executed in a sandboxed
+environment (a virtual machine), making this language an attractive target for
+virtualization in \emph{functions-as-a-service} platforms.
+%
+The language is relatively simple, in esence
+%
+\begin{inparaenum}
+  \item it contains only four numerical types, representing both integers and
+    IEEE754 floating point numbers of either 32 or 64 bits;
+  \item values of these types are manipulated by writing functions using
+    sequences of stack instructions;
+  \item functions are organized in modules and must be explicitly imported and
+    exported;
+  \item memory blocks can be imported, exported and grown dinamically;
+\end{inparaenum}
+%
+among others.
+%
+To give the reader a taste, Fig. \ref{fig:wasm:factorial} defines a WebAssembly
+module that exports a recursive implementation of the factorial function.
+%
+This language is a perfect match for our tool, and the space of WebAssembly
+programs contains mostly invalid ones, and automatically derived random
+generators cannot satisfy all the invariants required to produce interesting
+ones.
+
+\begin{wrapfigure}{r}{0.35\textwidth}
+\vspace{-10pt}
+\begin{verbatim}
+(module
+(func $f (param f64) (result f64)
+  get_local 0
+  f64.const 1
+  f64.lt
+  if (result f64)
+    f64.const 1
+  else
+    get_local 0
+    get_local 0
+    f64.const 1
+    f64.sub
+    call $fact
+    f64.mul
+  end)
+(export "fact" (func $f)))
+\end{verbatim}
+\caption{\label{fig:wasm:factorial}Factorial in WebAssembly}
+\vspace{-10pt}
+\end{wrapfigure}
+
+Unlike most other programming languages, its behavior is fully specified, and
+WebAssembly programs are expected to be consistently interpreted across engines
+--- despite some subtle details that we will address soon.
+%
+For this purpose, the WebAssembly standard provides a reference engine
+implementation with all the basic functionality expected from a compliant
+WebAssembly engine.
+
+In this work, we are interested on using our tool to test the two most complex
+subsystems of \textit{haskell-wasm}: the \emph{validator} and the
+\emph{interpreter} --- both being previously tested using a unit test suite.
+%
+For this purpose, we avoided spending countless hours writing an extensive
+property-based specification to mimic the reference WebAssembly specification.
+%
+Instead, we take advantage of the readily available reference implementation via
+differential testing.
+%
+In this light, our testing properties assert that any result produced by
+\textit{haskell-wasm} matches that of the reference implementation.
+
+Unsurprisingly, this engine had several subtle latent bugs that were not caught
+by the existing unit tests and that we discovered using \mutagen while
+developing the test suite used in this work.
+%
+Moreover, \mutagen exposed two discrepancies between \textit{haskell-wasm} and
+the reference implementation.
+%
+Not severe enough to be classified as bugs, these discrepancies trigger parts of
+the WebAssembly specification that are either not yet suported by the reference
+implementation (multi-value blocks), or that produce a well-known
+non-deterministic undefined behavior allowed by the specification (NaN
+reinterpretation).
+%
+These findings are briefly outlined in Table \ref{table:wasm:bugs}.
+
+The rest of this section introduces the test suite we used to test the different
+parts of \textit{haskell-wasm}.
+
+\begin{table}[b]
+\footnotesize
+\begin{tabular}{|c|c|c|l|}
+\hline
+\textbf{Id}
+& \textbf{Subsystem}
+& \textbf{Category}
+& \textbf{Description} \\
+\hline
+1
+& Validator
+& Bug
+& Invalid memory alignment validation \\
+\hline
+2
+& Validator
+& Discrepancy
+& Validator accepts blocks returning multiple values  \\
+\hline
+3
+& Interpreter
+& Bug
+& Instance function invoker silently proceeds after arity mismatch \\
+\hline
+4
+& Interpreter
+& Bug
+& Allowed out-of-bounds memory access \\
+\hline
+5
+& Interpreter
+& Discrepancy
+& NaN reinterpretation does not follow reference implementation\\
+\hline
+\end{tabular}
+\caption{\label{table:wasm:bugs}Bugs and discrepancies found by \mutagen in
+  \textit{haskell-wasm}.}
+\end{table}
+
+\subsubsection{Testing the WebAssembly Validator}
+
+We begin by designing a property to test the WebAssembly validator implemented
+in \textit{haskell-wasm}.
+%
+To keep things simple, we can simply assert that, whenever a randomly generated
+(or mutated for that matter) WebAssembly module is valid according to
+\textit{haskell-wasm}, then the reference implementation agrees upon it.
+%
+In other words, we are testing for false positives.
+%
+In Haskell, we write the following predicate:
+
+\begin{verbatim}
+prop_validator m =
+  isValidHaskell m ==> isValidSpec m
+\end{verbatim}
+
+Where the precondition (\texttt{isValidHaskell m}) runs the input WebAssembly
+module \texttt{m} against the \textit{haskell-wasm} validator, whereas the
+postcondition (\texttt{isValidSpec m}) serializes \texttt{m} to a file, runs it
+against the reference implementation validator and checks that no errors are
+produced.
+
+We want to remark that, although here for simplicity we only focus on finding
+false positives, in a realistic test suite, one would also want to test for
+false negatives, i.e., when a module is valid and \textit{haskell-wasm} rejects
+it.
+%
+This can be easily done by inverting the direction of the implication
+(\texttt{<==}) in the property above.
+%
+However, the resulting property is much slower, as every tested module will
+first be serialized and run against the reference implementation.
+
+%% \begin{verbatim}
+%% isValidHaskell m =
+%%   case validate m of
+%%     Left  validationError -> return False
+%%     Right validModule     -> return True
+%% \end{verbatim}
+
+%% \begin{verbatim}
+%% isValidSpec m = do
+%%   writeFile "testcase.wasm" (dumpModule m)
+%%   res <- shell "./wasm" ["-d", "-i", "testcase.wasm"]
+%%   return (res == "")
+%% \end{verbatim}
+
+\subsubsection{Testing the WebAssembly Interpreter}
+
+Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
+commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
+lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
+felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
+dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
+quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
+Nam euismod tellus id erat.
+
+\begin{verbatim}
+mk_module ty name fun =
+  emptyModule { types     = [ ty ]
+              , functions = [ fun ]
+              , exports   = [ Export name (ExportFunc 0) ]
+              , mems      = [ Memory (Limit 1 Nothing) ]
+              }
+\end{verbatim}
+
+\begin{verbatim}
+prop_interpreter ty (args, fun) =
+  discardAfter 20
+  (do let m = mk_module ty "f" fun;
+      resHs   <- invokeHaskell m "f" args
+      resSpec <- invokeSpec    m "f" args
+      return (equivalent resHs resSpec))
+\end{verbatim}
+
+\begin{verbatim}
+prop_interpreter_i32 (i, f, fun) =
+  prop_interpreter (FuncType { params = [I32, F32], result = [I32]}) ([VI32 i, VF32 f], fun)
+\end{verbatim}
+
+%% \subsubsection{Real Bugs and Discrepancies found by \mutagen in \textit{haskell-wasm}}
+
+%% \paragraph{Bug \#1: Invalid Memory Alignment Validation}
+%% \paragraph{Bug \#2: Lax invocation checks}
+%% \paragraph{Bug \#3: Allowed Out-of-bounds Memory Access}
+%% \paragraph{Discrepancy \#1: Different Reinterpretation Semantics of NaN Values}
+%% \paragraph{Discrepancy \#2: Allowed Blocks with Multiple Return Types}
+
+% ----------------------------------------
+
+\section{Evaluation}
+\label{sec:evaluation}
+
+All the experiments were done in a dedicated workstation with an Intel Core
+i7-8700 CPU running at 3.20GHz, and equipped with 32Gb of RAM.
+%
+We ran each experiment 30 times except for the ones involving the bugs on the
+WebAssembly interpreter, which were run 10 times.
+%
+From there, we followed the same approach taken by
+\citeauthor{lampropoulos2019coverage} and collected the Mean-Time-To-Failure
+(MTTF) of each bug, i.e., how quickly a bug can be found in wall time.
+%
+In all cases, we used a one-hour timeout to stop the execution of both tools if
+they have not yet found a counterexample.
+
+Additionally, we collected the Failure Rate (FR) observed for each bug, i.e. the
+proportion of times each tool finds each bug within the one-hour testing bugdet.
+%
+We found this metric crucial to be analyzed when replicating \fuzzchick's
+results, as opposed to just paying attention at the MTTF.
+
+\subsection{IFC Stack Machine}
+
+Aliquam erat volutpat \citep{devai2013edsl}. Nunc eleifend leo vitae magna. In
+id erat non orci commodo lobortis. Proin neque massa, cursus ut, gravida ut,
 lobortis eget, lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus.
@@
 
-\subsection{IFC Stack Machine}
+\begin{center}
+  \includegraphics[width=\linewidth]{tikz/ifc-mttf.tikz}
+  \includegraphics[width=\linewidth]{tikz/ifc-fr.tikz}
+\end{center}
+
+\subsection{WebAssembly Engine}
+
+Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
+commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
+lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
+felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
+dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
+quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
+Nam euismod tellus id erat.
+
+\begin{center}
+  \includegraphics[width=\linewidth]{tikz/wasm-mttf.tikz}
+  \includegraphics[width=\linewidth]{tikz/wasm-fr.tikz}\vspace{-5pt}
+  \includegraphics[width=\linewidth]{tikz/wasm-legend.tikz}
+\end{center}
+
+% ----------------------------------------
+
+%% \section{Implementation}
+%% \label{sec:implementation}
+
+%% \subsection{Deriving Mutation Machinery}
+
+%% Integer placerat tristique nisl. Praesent augue. Fusce commodo. Vestibulum
+%% convallis, lorem a tempus semper, dui dui euismod elit, vitae placerat urna
+%% tortor vitae lacus. Nullam libero mauris, consequat quis, varius et, dictum id,
+%% arcu. Mauris mollis tincidunt felis. Aliquam feugiat tellus ut neque. Nulla
+%% facilisis, risus a rhoncus fermentum, tellus tellus lacinia purus, et dictum
+%% nunc justo sit amet elit.
+
+%% \subsection{Tracing Haskell Programs}
+
+%% \begin{verbatim}
+%%   sorted []       = True
+%%   sorted [x]      = True
+%%   sorted (x:y:xs) = if x <= y then sorted (y:xs) else False
+%% \end{verbatim}
+
+%% \begin{verbatim}
+%%   sorted []       = _trace_ 1 (True)
+%%   sorted [x]      = _trace_ 2 (True)
+%%   sorted (x:y:xs) = _trace_ 3 (if x <= y then _trace_ 4 (sorted (y:xs)) else _trace_ 5 (False))
+%% \end{verbatim}
+
+%% Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
+%% tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
+%% leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
+%% wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
+%% mollis ultricies.
+
+% ----------------------------------------
+
+\section{Related work}
+\label{sec:related}
+
+\subsection{Grammar-based Fuzzing}
 
@@
 
-\subsection{WebAssembly Engine}
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
-\subsubsection{Testing the WebAssembly Validator}
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
-\begin{verbatim}
-prop_validator m =
-  isValidHaskell m ==> isValidSpec m
-\end{verbatim}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-\begin{verbatim}
-isValidHaskell m =
-  case validate m of
-    Left  validationError -> return False
-    Right validModule     -> return True
-\end{verbatim}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-\begin{verbatim}
-isValidSpec m = do
-  writeFile "testcase.wasm" (dumpModule m)
-  res <- shell "./wasm" ["-d", "-i", "testcase.wasm"]
-  return (res == "")
-\end{verbatim}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-\subsubsection{Testing the WebAssembly Interpreter}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-\begin{verbatim}
-mk_module ty name fun =
-  emptyModule { types     = [ ty ]
-              , functions = [ fun ]
-              , exports   = [ Export name (ExportFunc 0) ]
-              , mems      = [ Memory (Limit 1 Nothing) ]
-              }
-\end{verbatim}
-
-\begin{verbatim}
-prop_interpreter ty (args, fun) =
-  discardAfter 20
-  (do let m = mk_module ty "f" fun;
-      resHs   <- invokeHaskell m "f" args;
-      resSpec <- invokeSpec    m "f" args;
-      return (equiv resHs resSpec))
-\end{verbatim}
-
-\begin{verbatim}
-prop_interpreter_i32 (i, f, fun) =
-  prop_interpreter
-    (FuncType { params = [I32, F32], result = [I32]})
-    ([VI32 i, VF32 f], fun)
-\end{verbatim}
-
-\subsubsection{Real Bugs and Discrepancies found by \mutagen in \textit{haskell-wasm}}
-
-\paragraph{Bug \#1: Invalid Memory Alignment Validation}
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-
-\paragraph{Bug \#2: Lax invocation checks}
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-\paragraph{Bug \#3: Allowed Out-of-bounds Memory Access}
+\subsection{Random Data Generation}
+
+%% Don't forget to cite "Deriving good generators ..."
+
 Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
@@
 
-\paragraph{Discrepancy \#1: Different Reinterpretation Semantics of NaN Values}
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-
-\paragraph{Discrepancy \#2: Allowed Blocks with Multiple Return Types}
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-
-\section{Evaluation}
-\label{sec:evaluation}
-
-\subsection{IFC Stack Machine}
-
-Aliquam erat volutpat \citep{devai2013edsl}. Nunc eleifend leo vitae magna. In
-id erat non orci commodo lobortis. Proin neque massa, cursus ut, gravida ut,
-lobortis eget, lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus.
-Mauris ac felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque
-nec dui dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a,
-aliquet quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet
-tortor. Nam euismod tellus id erat.
-
-\begin{center}
-  \includegraphics[width=\linewidth]{tikz/ifc-mttf.tikz}
-  \includegraphics[width=\linewidth]{tikz/ifc-fr.tikz}
-\end{center}
-
-\subsection{Webassembly Engine}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-\begin{center}
-  \includegraphics[width=\linewidth]{tikz/wasm-mttf.tikz}
-  \includegraphics[width=\linewidth]{tikz/wasm-fr.tikz}\vspace{-5pt}
-  \includegraphics[width=\linewidth]{tikz/wasm-legend.tikz}
-\end{center}
-
-% ----------------------------------------
-
-%% \section{Implementation}
-%% \label{sec:implementation}
-
-%% \subsection{Deriving Mutation Machinery}
-
-%% Integer placerat tristique nisl. Praesent augue. Fusce commodo. Vestibulum
-%% convallis, lorem a tempus semper, dui dui euismod elit, vitae placerat urna
-%% tortor vitae lacus. Nullam libero mauris, consequat quis, varius et, dictum id,
-%% arcu. Mauris mollis tincidunt felis. Aliquam feugiat tellus ut neque. Nulla
-%% facilisis, risus a rhoncus fermentum, tellus tellus lacinia purus, et dictum
-%% nunc justo sit amet elit.
-
-%% \subsection{Tracing Haskell Programs}
-
-%% \begin{verbatim}
-%%   sorted []       = True
-%%   sorted [x]      = True
-%%   sorted (x:y:xs) = if x <= y then sorted (y:xs) else False
-%% \end{verbatim}
-
-%% \begin{verbatim}
-%%   sorted []       = _trace_ 1 (True)
-%%   sorted [x]      = _trace_ 2 (True)
-%%   sorted (x:y:xs) = _trace_ 3 (if x <= y then _trace_ 4 (sorted (y:xs)) else _trace_ 5 (False))
-%% \end{verbatim}
-
-%% Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-%% tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-%% leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-%% wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-%% mollis ultricies.
-
-% ----------------------------------------
-
-\section{Related work}
-\label{sec:related}
-
-\subsection{Grammar-based Fuzzing}
-
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
-
-\subsection{Random Data Generation}
-
-%% Don't forget to cite "Deriving good generators ..."
-
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
-
 \subsection{Exhaustive Bounded Testing}
@@
 %% If your work has an appendix, this is the place to put it.
-\appendix
-
-\section{Detailed Empirical Results}
-
-Fusce suscipit, wisi nec facilisis facilisis, est dui fermentum leo, quis tempor
-ligula erat quis odio. Nunc porta vulputate tellus. Nunc rutrum turpis sed pede.
-Sed bibendum. Aliquam posuere. Nunc aliquet, augue nec adipiscing interdum,
-lacus tellus malesuada massa, quis varius mi purus non odio. Pellentesque
-condimentum, magna ut suscipit hendrerit, ipsum augue ornare nulla, non luctus
-diam neque sit amet urna. Curabitur vulputate vestibulum lorem. Fusce sagittis,
-libero non molestie mollis, magna orci ultrices dolor, at vulputate neque nulla
-lacinia eros. Sed id ligula quis est convallis tempor. Curabitur lacinia
-pulvinar nibh. Nam a sapien.
+%% \appendix
 

### 2021-03-30 papers/oopsla-21/main.tex
WIP

@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{\mutagen: Working title}
+\title[\mutagen]{\mutagen: Reliable Coverage-Guided, Property-Based Testing using
+  Exhaustive Well-Typed Mutations}
 
@@
 \begin{algorithm}[H]
+  \SetInd{0em}{0.75em}
   \SetKw{KwNot}{not}
@@
 quickCheck prop_mirror
-+++ Ok ####
++++ OK, passed 100 tests.
 \end{verbatim}
@@
 quickCheck prop_mirror
-*** Failed ####
+*** Failed! Falsified (after 2 tests and 2 shrinks):
+Branch (Leaf 0) 0 (Leaf 1)
 \end{verbatim}
@@
 For this, consider that we want to use our \texttt{Tree} data type to encode
-binary-search trees (BST).
+binary-search trees (BST) --- this requires some minor tweaks in practice.
 %
@@
 quickCheck prop_bst_insert
-*** Gave up! ####
+*** Gave up! Passed only 44 tests; 1000 discarded tests.
 \end{verbatim}
 
-\noindent Here, \quickcheck discards random inputs as soon as it finds they do
-not pass the precondition (\texttt{isBST t}).
+\noindent \quickcheck discards random inputs as soon as it finds they do not
+pass the precondition (\texttt{isBST t}).
 %
@@
 The second case study aims to evaluate \mutagen in a more realistic scenario,
-and focuses on testing \textit{haskell-wasm} an existing WebAssembly engine
+and focuses on testing \textit{haskell-wasm}, an existing WebAssembly engine
 written in Haskell of industrial strength.
@@
 
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
+The abstract stack machine used in this case study consists of four main
+elements: a program counter, a stack, and data- and instruction memories.
+%
+Moreover, every runtime value is labeled with its security level, i.e., L (for
+``low'' or public) or H (for ``high'' or secret).
+%
+Labels form a trivial 2-lattice where information can either stay at the same
+level, or public information can flow to secret one but not the opposite.
+%
+Security labels are propagated throughout the execution of the program every
+time the machine executes an instruction.
+%
+There are eight different instructions defined as:
+%
+\begin{verbatim}
+data Instr = Nop | Push Int | Call Int | Ret | Add | Load | Store | Halt
+\end{verbatim}
+
+\noindent Control flow is achieved using the \texttt{Call} and \texttt{Ret}
+instructions, that let the program jump back and forth within the instruction
+memory using specially labelled values in the stack respresenting memory
+addresses.
+%
+The argument in the \texttt{Push} instruction represents a value to be
+inserted in the stack, whereas the argument of the \texttt{Call} instruction
+encodes the number of elements of the stack to be treated as arguments.
+%
+Then, programs are simply modeled as sequences of instructions.
+%
+To preserve space, we encourage the reader to refer to the work of
+\citeauthor{catalin} as well as the original \fuzzchick paper for more details
+about the implementation and semantics of this case study.
+
+\subsubsection{Noninterference}
+
+\subsection{Single-step Noninterference}
 
@@
 WebAssembly is a popular assembly-like language designed to be an open standard
-for executing low-level code in web pages, although it has become increasingly
+for executing low-level code in the web, although it has become increasingly
 popular in standalone, non-web contexts as well.
@@
 %
-To give the reader a taste, Fig. \ref{fig:wasm:factorial} defines a WebAssembly
-module that exports a recursive implementation of the factorial function.
-%
-This language is a perfect match for our tool, and the space of WebAssembly
-programs contains mostly invalid ones, and automatically derived random
-generators cannot satisfy all the invariants required to produce interesting
-ones.
+To give the reader a taste, Fig. \ref{fig:wasm:factorial} the WebAssembly text
+representation of module that exports a recursive implementation of the
+factorial function.
 
@@
 %
-For this purpose, the WebAssembly standard provides a reference engine
-implementation with all the basic functionality expected from a compliant
-WebAssembly engine.
-
-In this work, we are interested on using our tool to test the two most complex
+For this purpose, the WebAssembly standard provides a reference implementation
+with all the basic functionality expected from a compliant WebAssembly engine.
+
+In this work, we are interested on using \mutagen to test the two most complex
 subsystems of \textit{haskell-wasm}: the \emph{validator} and the
@@
 %
-For this purpose, we avoided spending countless hours writing an extensive
-property-based specification to mimic the reference WebAssembly specification.
+Our tool is an attractive match for testing \textit{haskell-wasm}, as the space
+of WebAssembly programs contains mostly invalid ones, and automatically derived
+random generators cannot satisfy all the invariants required to produce
+interesting test cases.
+%
+Here, we avoided spending countless hours writing an extensive property-based
+specification to mimic the reference WebAssembly specification.
 %
@@
 %
-In Haskell, we write the following predicate:
+In Haskell, we write the following testing property:
 
@@
 %
-However, the resulting property is much slower, as every tested module will
-first be serialized and run against the reference implementation.
+However, the resulting property is much slower, as every tested module will be
+serialized and run against the reference implementation.
 
@@
 \subsubsection{Testing the WebAssembly Interpreter}
+
+Testing the WebAssembly interpreter is substantially more complicated than
+testing the validator, as it entails running actual programs.
+%
+To achieve this, we need the generated test cases to comply a with stable
+interface that can be invoked both by \textit{haskell-wasm} and the reference
+WebAssembly implementation.
+
+To keep things simple here as well, we use a stub definition that provides a
+module that initializes a memory block and exports a single function.
+%
+This module can be instantiated by providing the definition of the single
+function, along with a name and a type signature.
+%
+In Haskell:
+
+\begin{verbatim}
+mk_module ty name fun =
+  emptyModule { types     = [ ty ]
+              , functions = [ fun ]
+              , exports   = [ Export name (ExportFunc 0) ]
+              , mems      = [ Memory (Limit 1 Nothing) ]
+              }
+\end{verbatim}
+
+Using this stub module, we can define a testing property parameterized by the
+function type signature, the function implementation and a list of invocation
+arguments:
+
+\begin{verbatim}
+prop_interpreter ty fun args =
+  (discardAfter 20)
+  (do let m = mk_module ty "f" fun;
+      resHs   <- invokeHaskell m "f" args
+      resSpec <- invokeSpec    m "f" args
+      return (equivalent resHs resSpec))
+\end{verbatim}
+
+This property instantiates the module stub using the input function and its type
+signature, and uses it to invoke both the \textit{haskell-wasm} and reference
+implementation interepreters with the provided arguments.
+%
+Then, the property asserts whether their results are equivalent.
+%
+Interestingly, equivalence in this context does not imply equality,
+non-deterministic operations in WebAssembly like NaN reinterpretations can
+produce different equivalent results (as exposed by the discrepancy \#5 in Table
+\ref{table:wasm:bugs}), and our equivalence relation needs to take that into
+acount.
+%
+Notice that we additionally set a 20ms timeout to discard potentially diverging
+programs with infinite loops.
+
+Using this testing property directly might not sound like a great idea, as
+randomly generated lists of inputs will be very unlikely to match the type
+signature of randomly generated functions.
+%
+However, it lets us test what happens when programs are not properly invoked,
+and it quickly discovered the bug \#3 in \textit{haskell-wasm} mentioned above.
+%
+Having solved this issue in \textit{haskell-wasm}, we proceed to define a more
+useful specialized version of \texttt{prop\_interpreter} that fixes the type of
+the generated function to take two arguments (of type \texttt{I32} and {F32})
+and return an \texttt{I32} as a result:
+
+\begin{verbatim}
+prop_interpreter_i32 fun i f =
+  prop_interpreter (FuncType { params = [I32, F32], result = [I32]}) fun [VI32 i, VF32 f]
+\end{verbatim}
+
+This specilized property let us generate functions using this fixed type and
+invoke them with the exact number and type of arguments required.
+%
+In our experiments (presented in next section), we use this property when
+finding all the injected bugs into the \textit{haskell-wasm} interpreter.
+%
+Worth mentioning again, a realistic test suite should at least include different
+variants of this property testing functions of several different types.
+
+%% \subsubsection{Real Bugs and Discrepancies found by \mutagen in \textit{haskell-wasm}}
+
+%% \paragraph{Bug \#1: Invalid Memory Alignment Validation}
+%% \paragraph{Bug \#2: Lax invocation checks}
+%% \paragraph{Bug \#3: Allowed Out-of-bounds Memory Access}
+%% \paragraph{Discrepancy \#1: Different Reinterpretation Semantics of NaN Values}
+%% \paragraph{Discrepancy \#2: Allowed Blocks with Multiple Return Types}
+
+% ----------------------------------------
+
+\section{Evaluation}
+\label{sec:evaluation}
+
+All the experiments were done in a dedicated workstation with an Intel Core
+i7-8700 CPU running at 3.20GHz, and equipped with 32Gb of RAM.
+%
+We ran each experiment 30 times except for the ones involving the bugs on the
+WebAssembly interpreter, which were run 10 times.
+%
+From there, we followed the same approach taken by
+\citeauthor{lampropoulos2019coverage} and collected the Mean-Time-To-Failure
+(MTTF) of each bug, i.e., how quickly a bug can be found in wall time.
+%
+In all cases, we used a one-hour timeout to stop the execution of both tools if
+they have not yet found a counterexample.
+
+Additionally, we collected the Failure Rate (FR) observed for each bug, i.e. the
+proportion of times each tool finds each bug within the one-hour testing bugdet.
+%
+We found this metric crucial to be analyzed when replicating \fuzzchick's
+results, as opposed to just paying attention at the MTTF.
+
+\subsection{IFC Stack Machine}
+
+Aliquam erat volutpat \citep{devai2013edsl}. Nunc eleifend leo vitae magna. In
+id erat non orci commodo lobortis. Proin neque massa, cursus ut, gravida ut,
+lobortis eget, lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus.
+Mauris ac felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque
+nec dui dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a,
+aliquet quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet
+tortor. Nam euismod tellus id erat.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{tikz/ifc-mttf.tikz}
+  \includegraphics[width=\linewidth]{tikz/ifc-fr.tikz}
+  \caption{\label{fig:results:ifc} Comparison between \fuzzchick and \mutagen
+    across 20 different bugs for the IFC stack machine. }
+\end{figure}
+
+\subsection{WebAssembly Engine}
 
@@
 
-\begin{verbatim}
-mk_module ty name fun =
-  emptyModule { types     = [ ty ]
-              , functions = [ fun ]
-              , exports   = [ Export name (ExportFunc 0) ]
-              , mems      = [ Memory (Limit 1 Nothing) ]
-              }
-\end{verbatim}
-
-\begin{verbatim}
-prop_interpreter ty (args, fun) =
-  discardAfter 20
-  (do let m = mk_module ty "f" fun;
-      resHs   <- invokeHaskell m "f" args
-      resSpec <- invokeSpec    m "f" args
-      return (equivalent resHs resSpec))
-\end{verbatim}
-
-\begin{verbatim}
-prop_interpreter_i32 (i, f, fun) =
-  prop_interpreter (FuncType { params = [I32, F32], result = [I32]}) ([VI32 i, VF32 f], fun)
-\end{verbatim}
-
-%% \subsubsection{Real Bugs and Discrepancies found by \mutagen in \textit{haskell-wasm}}
-
-%% \paragraph{Bug \#1: Invalid Memory Alignment Validation}
-%% \paragraph{Bug \#2: Lax invocation checks}
-%% \paragraph{Bug \#3: Allowed Out-of-bounds Memory Access}
-%% \paragraph{Discrepancy \#1: Different Reinterpretation Semantics of NaN Values}
-%% \paragraph{Discrepancy \#2: Allowed Blocks with Multiple Return Types}
-
-% ----------------------------------------
-
-\section{Evaluation}
-\label{sec:evaluation}
-
-All the experiments were done in a dedicated workstation with an Intel Core
-i7-8700 CPU running at 3.20GHz, and equipped with 32Gb of RAM.
-%
-We ran each experiment 30 times except for the ones involving the bugs on the
-WebAssembly interpreter, which were run 10 times.
-%
-From there, we followed the same approach taken by
-\citeauthor{lampropoulos2019coverage} and collected the Mean-Time-To-Failure
-(MTTF) of each bug, i.e., how quickly a bug can be found in wall time.
-%
-In all cases, we used a one-hour timeout to stop the execution of both tools if
-they have not yet found a counterexample.
-
-Additionally, we collected the Failure Rate (FR) observed for each bug, i.e. the
-proportion of times each tool finds each bug within the one-hour testing bugdet.
-%
-We found this metric crucial to be analyzed when replicating \fuzzchick's
-results, as opposed to just paying attention at the MTTF.
-
-\subsection{IFC Stack Machine}
-
-Aliquam erat volutpat \citep{devai2013edsl}. Nunc eleifend leo vitae magna. In
-id erat non orci commodo lobortis. Proin neque massa, cursus ut, gravida ut,
-lobortis eget, lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus.
-Mauris ac felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque
-nec dui dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a,
-aliquet quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet
-tortor. Nam euismod tellus id erat.
-
-\begin{center}
-  \includegraphics[width=\linewidth]{tikz/ifc-mttf.tikz}
-  \includegraphics[width=\linewidth]{tikz/ifc-fr.tikz}
-\end{center}
-
-\subsection{WebAssembly Engine}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-\begin{center}
+\begin{figure}
+  \centering
   \includegraphics[width=\linewidth]{tikz/wasm-mttf.tikz}
@@
   \includegraphics[width=\linewidth]{tikz/wasm-legend.tikz}
-\end{center}
+  \caption{\label{fig:results:wasm} Comparison between \quickcheck and \mutagen
+    across 15 different bugs for the WebAssembly case study. }
+\end{figure}
 
@@
 
-Aliquam erat volutpat \citep{devai2013edsl}. Nunc eleifend leo vitae magna. In id
-erat non orci commodo lobortis. Proin neque massa, cursus ut, gravida ut,
-lobortis eget, lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus.
-Mauris ac felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque
-nec dui dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a,
-aliquet quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet
-tortor. Nam euismod tellus id erat.
+We presented \mutagen, a coverage-guided, property-based testing framework
+written in Haskell.
+%
+Inspired by \fuzzchick, our tool uses coverage information to guide
+automatically derived mutators producing high-level, well-typed mutations.
+%
+However, instead of relying heavily on randomness and power schedules to find
+bugs, our tool uses an exhaustive mutation approach that generates every
+possible mutant for each interesting input candidate, and schedules it to be
+tested exactly once.
+%
+This is in turn inspired by exhaustive bounded testing tools that focus on
+testing every possible input value of the system under test --- a more generic
+technique of limited applicability.
+
+Our experimental results indicate that \mutagen outperforms the simpler approach
+taken by \fuzzchick in terms of failure rate.
+%
+Moreover, we show how our tool can be applied in a real-world testing scenario,
+where it quickly discovers several planted and existent previously unknown bugs.
+
+In the future, we will investigate how to redefine our automatically synthesized
+mutators in a stateful manner.
+%
+This way, it would be possible to apply mutations that preserve complex
+properties of the generated data like well-scopedness and well-typedness simply
+by construction, e.g., mutations that \emph{always} produce well-typed
+subexpressions that refer only to identifiers in the current scope.
+%
+The main challenge will be to achieve this while keeping the testing process as
+automatable as possible.
 

### 2021-03-30 papers/oopsla-21/main.tex
WIP

@@
 %
-In both of them, we show how the heuristics described in Section
-\ref{sec:heuristics} affect the testing performance.
-
 The first one is a simple abstract stack machine that enforces
@@
 %
-In this work we replicate their results and compare them against our tool.
+In this work, we replicate their results and compare them against our tool.
+%
+Worth mentioning, since this case study was originally implemented in Coq, we
+first translated it to Haskell in order to run \mutagen on its test suite.
 
@@
 interpreter of this engine, and used the reference implementation to find them
-using differential testing.
+via differential testing.
 %
@@
 %
-All of them were reported and confirmed by the authors of \textit{haskell-wasm}.
+All of them were reported and later confirmed by the authors of
+\textit{haskell-wasm}.
 
@@
 %
-Moreover, every runtime value is labeled with its security level, i.e., L (for
+Moreover, every runtime value is labeled with a security level, i.e., L (for
 ``low'' or public) or H (for ``high'' or secret).
@@
 Labels form a trivial 2-lattice where information can either stay at the same
-level, or public information can flow to secret one but not the opposite.
+level, or public information can flow to secret one but not the opposite
+(represented using the familiar $\sqsubseteq$ binary operator).
 %
@@
 %
+Finally, machine states can be modeled using a 4-tuple \textit{(pc, stk, m, i)}
+that represents a particular configuration of the program counter, stack and
+data- and instruction memories, respectively.
+%
 To preserve space, we encourage the reader to refer to the work of
@@
 
-\subsubsection{Noninterference}
-
-\subsection{Single-step Noninterference}
+\subsubsection{Single-Step Noninterference (SSNI)}
+
+This abstract machine is designed to enforce noninterference, which is based on
+the notion of \emph{indistinguishability}.
+%
+Intuitively, two machine states are indistinguishable from each other if they
+only differ on secret data.
+%
+Using this notion, the particular variant of noninterference we are interested
+in this work is called \emph{single-step noninterference}.
+%
+In simple terms, this property asserts that, given two indistinguishable
+machines states, running a single instruction on both machines brings them to
+resulting states that are again indistinguishable.
+
+The tricky part about this property is, of course, satisfying its sparse
+precondition: we need to generate two valid indistinguishable machine states in
+order to even proceed to execute the next instruction.
+%
+As demonstrated by \citeauthor{lampropoulos2019coverage}, generating two
+independent machine states using \quickcheck has virtually no chances of
+producing valid indistinguishable ones.
+%
+However, having the mutation mechanism available, we can use a clever trick: we
+can obtain a pair of valid indistinguishable machine states by generating a
+single valid machine state (something still hard but much easier than before),
+and then producing a similar mutated copy.
+%
+By doing this, we have a much higher chance of producing two almost identical
+states that pass the sparse precondition.
+
+\subsubsection{IFC Rules}
+
+In this abstract machine, the enforced IFC rules are implemented using a rule
+table indexed by the operation that the machine is about to perform.
+%
+Such table contains: the dynamic check that the abstract machine needs to
+perform in order to enforce the IFC policy, along with the corresponding
+security label of the program counter and the operation result after the
+operation is executed.
+%
+For instance, to execute the \texttt{Store} operation (which stores a value in a
+memory pointer) the machine needs to check that both the label of the program
+counter and the label of the pointer together ``can flow'' to the label of the
+destination memory cell.
+%
+If this condition is not met, the machine then halts as indication of a
+violation in the IFC policy.
+%
+After this check, this operation overwrites the value at the destination cell
+and updates its label with the maximum sensibility of the involved labels.
+%
+In the rule table, this looks as follows:
+
+\begin{center}
+\footnotesize
+\begin{tabular}{|c|c|c|c|}
+\hline
+\textbf{Operation}
+& \textbf{Precondition Check}
+& \textbf{Final PC Label}
+& \textbf{Final Result Label} \\
+\hline
+\texttt{Store}
+& $l_{pc} \vee l_{p} \sqsubseteq l_{v}$
+& $l_{pc}$
+& $l_{v'} \vee l_{pc} \vee l_{p}$ \\
+\hline
+\end{tabular}
+\end{center}
+
+Where $l_{pc}$, $l_{p}$, $l_{v}$, and $l_{v'}$ represent the labels of: the
+program counter, the memory pointer, and the old a new values stored in that
+memory cell.
+%
+The symbol $\vee$ simply denotes the join of two labels, i.e., the
+\emph{maximum} of their sensibilities.
+
+In this case study, we systematically introduce several bugs on the IFC
+enforcing mechanism by removing or weaking the checks stored in this rule table.
 
@@
 %
-To give the reader a taste, Fig. \ref{fig:wasm:factorial} the WebAssembly text
-representation of module that exports a recursive implementation of the
-factorial function.
-
-\begin{wrapfigure}{r}{0.35\textwidth}
-\vspace{-10pt}
-\begin{verbatim}
-(module
-(func $f (param f64) (result f64)
-  get_local 0
-  f64.const 1
-  f64.lt
-  if (result f64)
-    f64.const 1
-  else
-    get_local 0
-    get_local 0
-    f64.const 1
-    f64.sub
-    call $fact
-    f64.mul
-  end)
-(export "fact" (func $f)))
-\end{verbatim}
-\caption{\label{fig:wasm:factorial}Factorial in WebAssembly}
-\vspace{-10pt}
-\end{wrapfigure}
+%% To give the reader a taste, Fig. \ref{fig:wasm:factorial} the WebAssembly text
+%% representation of module that exports a recursive implementation of the
+%% factorial function.
 
@@
 
+%% \begin{wrapfigure}{r}{0.35\textwidth}
+%% \vspace{-10pt}
+%% \begin{verbatim}
+%% (module
+%% (func $f (param f64) (result f64)
+%%   get_local 0
+%%   f64.const 1
+%%   f64.lt
+%%   if (result f64)
+%%     f64.const 1
+%%   else
+%%     get_local 0
+%%     get_local 0
+%%     f64.const 1
+%%     f64.sub
+%%     call $fact
+%%     f64.mul
+%%   end)
+%% (export "fact" (func $f)))
+%% \end{verbatim}
+%% \caption{\label{fig:wasm:factorial}Factorial in WebAssembly}
+%% \vspace{-10pt}
+%% \end{wrapfigure}
+
 In this work, we are interested on using \mutagen to test the two most complex
@@
 Our tool is an attractive match for testing \textit{haskell-wasm}, as the space
-of WebAssembly programs contains mostly invalid ones, and automatically derived
-random generators cannot satisfy all the invariants required to produce
-interesting test cases.
+of WebAssembly programs that can be respresented using its AST contains mostly
+invalid ones, and automatically derived random generators cannot satisfy all the
+invariants required to produce interesting test cases.
 %
@@
 %
-However, the resulting property is much slower, as every tested module will be
-serialized and run against the reference implementation.
+However, the resulting property is much slower, as every tested module will
+always be serialized and run against the reference implementation.
 
@@
 Testing the WebAssembly interpreter is substantially more complicated than
-testing the validator, as it entails running actual programs.
+testing the validator, since it requires running actual programs.
 %
@@
 %
-Interestingly, equivalence in this context does not imply equality,
-non-deterministic operations in WebAssembly like NaN reinterpretations can
+Interestingly, equivalence in this context does not imply equality.
+%
+Non-deterministic operations in WebAssembly like NaN reinterpretations can
 produce different equivalent results (as exposed by the discrepancy \#5 in Table
@@
 results, as opposed to just paying attention at the MTTF.
+
+In both case studies, we additionally show how the heuristics described in
+Section \ref{sec:heuristics} affect the testing performance by disabling them
+when using our tool.
 
@@
 
-\subsection{Grammar-based Fuzzing}
+\paragraph{Grammar-based Fuzzing}
 
@@
 
-\subsection{Random Data Generation}
+\paragraph{Random Data Generation}
 
@@
 
-\subsection{Exhaustive Bounded Testing}
+\paragraph{Exhaustive Bounded Testing}
 

### 2021-03-31 papers/oopsla-21/main.tex
WIP

@@
 
-All the experiments were done in a dedicated workstation with an Intel Core
+All the experiments were performed in a dedicated workstation with an Intel Core
 i7-8700 CPU running at 3.20GHz, and equipped with 32Gb of RAM.
@@
 
-\paragraph{Grammar-based Fuzzing}
+There exists a vast literature on fuzzing and property-based testing.
+%
+In this section we focus on three main topics:
+%
+automated random data generation using static information,
+%
+fuzzing using coverage information to guide mutations, and
+%
+exhaustive bounded testing tools, which inspired the design of \mutagen.
+
+\paragraph{Automated Random Data Generation}
+
+Obtaining good random generators automatically from static information (e.g.,
+grammars or datatype definitions) has been tackled from different angles.
+
+% DRAGEN and DRAGEN2
+In Haskell, DRAGEN \cite{dragen} is a meta-programming tool that synthezises
+random generators from datatypes definitions, using stochastic models to predict
+and optimize their distribution based on a target distribution set by the user.
+%
+DRAGEN2 \cite{dragen2} extends this idea adding support for generating richer
+random data by extracting additional static information like library APIs and
+function input patterns from the codebase.
+%
+% Bolztmann-brain
+Similarly, \cite{bendkowski} developed a polinomial tuning mechanism based on
+Boltzmann samplers \cite{boltzmann} that synthesizes generators with
+approximate-size distributions for combinatorial structures like lists or trees.
+%
+
+Automatically deriving random generators is substantially more complicated when
+the generated data must satisfy by (often sparse) preconditions.
+%
+% Generating Constrained Random Data with Uniform Distribution
+%
+\cite{claessen2014} developed an algorithm for generating inputs constrained by
+boolean preconditions with almost-uniform distribution.
+%
+% Luck
+%
+Later, \cite{luck} extends this approach by adding a limited form of constraint
+solving controllable by the user in a domain-specific language called
+\emph{Luck}.
+%
+% Deriving good generators
+%
+Recently, \cite{generatinggood} proposed a derivation mechanism to obtain
+constrained random generators directly by defining their structure
+using inductively defined relations in Coq.
+
+We consider all these generational approaches to be orthogonal to the ideas
+behind \mutagen.
+%
+In principle, our tool is tailored to improve the performance of poor
+automatically derived generators.
+%
+However, more specialized generators can by used directly by \mutagen, and
+combining them with well-typed mutations using perhaps a hybrid technique is an
+appealing idea that we aim to address in the future.
+
+\paragraph{Coverage-guided Fuzzing}
+
+% AFL
+
+% AFLfast
+
+% CollAFL
+
+% Criterion
+
+% CSmith, XSmith
+
+% QuickFuzz
 
@@
 
-\paragraph{Random Data Generation}
-
-%% Don't forget to cite "Deriving good generators ..."
-
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
-
 \paragraph{Exhaustive Bounded Testing}
 
-Aliquam erat volutpat \citep{devai2013edsl}. Nunc eleifend leo vitae magna. In id
-erat non orci commodo lobortis. Proin neque massa, cursus ut, gravida ut,
-lobortis eget, lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus.
-Mauris ac felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque
-nec dui dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a,
-aliquet quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet
-tortor. Nam euismod tellus id erat.
+A popular category of property-based testing tools does not rely on randomness.
+%
+Instead, all possible inputs can be enumerated and tested from smaller to larger
+up to a certain size bound.
+
+%
+% Feat
+%
+\emph{Feat} \cite{feat} formalizes the notion of \emph{functional enumerations},
+For any algebraic type, it synthesizes a bijection between a finite prefix of
+the natural numbers and set of increasingly growing values of the input type.
+%
+Later, this bijection can be traversed exhaustively or, more interestingly,
+randomly accesed.
+%
+This allows the user to easily generate random values in a uniform basis simply
+by sampling natural numbers.
+%
+However, values are enumerated based only on their type definition, so this
+technique is not suitable for testing properties with sparse preconditions
+expressed elsewhere.
+
+% SmallCheck and LazySmallCheck
+\emph{SmallCheck} \cite{smallcheck} is a Haskell tool that follows this idea.
+%
+\emph{SmallCheck} is an exhaustive testing tool for Haskell that progressively
+executes the testing properties against all possible inputs values of within
+a certain size bound.
+%
+Later, \emph{LazySmallCheck} is a lazy variant of \emph{SmallCheck} that uses
+lazy evaluation to automatically prune the search space by detecting input
+subexpressions that are not evaluated by the testing properties, and hence do
+not need to be populated exhaustively with concrete values.
+
+% Korat
+On the object-oriented realm, \emph{Korat} \cite{korat} is a Java tool that uses
+method specification predicates to automatically generate all (nonisomorphic)
+test cases up to a given small size.
+%
+Similar to \emph{LazySmallCheck}, \emph{Korat} uses method preconditions to
+prune the search space.
+%
+However, in \emph{Korat}, pruning is done by analyzing the execution trace of
+the precondition predicate instead of by keeping track of unused subexpressions.
+
+Being exhaustive, these approaches can only be used to enumerate and test a
+relatively little space of small inputs.
+%
+In this work we use exhaustiveness as a way to reliably schedule different
+mutants.
+%
+In contrast to fully-exhaustive testing tools, \mutagen relies initially on
+randomness to find interesting mutable candidates.
+%
+Later, enumerating mutations for a given interesting candidate is much cheaper
+than enumerating \emph{all posible values}.
+%
+Hence, our tool can be used to produce much larger test cases before their size
+becomes computationally too restrictive.
 

### 2021-03-31 papers/oopsla-21/main.tex
WIP

@@
 %
-In this section we focus on three main topics:
+To preserve space, in this section we only focus on three main topics:
 %
@@
 %
-fuzzing using coverage information to guide mutations, and
-%
-exhaustive bounded testing tools, which inspired the design of \mutagen.
+fuzzing using coverage information, and
+%
+exhaustive bounded testing tools, all of which inspired the design of \mutagen.
 
@@
 random generators from datatypes definitions, using stochastic models to predict
-and optimize their distribution based on a target distribution set by the user.
+and optimize their distribution based on a target one set by the user.
 %
@@
 approximate-size distributions for combinatorial structures like lists or trees.
-%
+
+%
+% QuickFuzz
+%
+Exploiting some of the ideas described above, \emph{QuickFuzz} \cite{quickfuzz,
+  quickfuzzjournal} is a generational grammar-based fuzzer that leverages
+existing Haskell libraries describing common file formats to synthesize random
+data generators that are used in tandem with off-the-shelf low-level fuzzers to
+find bugs in massively used programs.
 
@@
 However, more specialized generators can by used directly by \mutagen, and
-combining them with well-typed mutations using perhaps a hybrid technique is an
+combining them with well-typed mutations using a hybrid technique is an
 appealing idea that we aim to address in the future.
@@
 % AFL
-
+\emph{AFL} \cite{afl} is the reference tool when it comes to coverage-guided
+fuzzing.
+%
 % AFLfast
-
+\emph{AFLFast} is an extension to AFL that uses Markov chain models to tune the
+power scheduler towards testing low-frequency paths.
+%
+In \mutagen, the scheduler does not account for path frequency.
+%
+Instead, it favours a breadth-first traversal of the execution path space.
+%
 % CollAFL
-
-% Criterion
-
-% CSmith, XSmith
-
-% QuickFuzz
-
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
+Similar to our tool, \emph{CollAFL} \cite{collAFL} is a variant of AFL that uses
+path- instead of edge-based coverage, which helps distinguishing executions more
+precisely by reducing path collisions.
+
+%AFLGo
+When the source code history is available, \emph{AFLGo} \cite{aflgo} is a fuzzer
+that can be targeted to exercise the specific parts of the the code affected by
+recent commits in order to find potential new bugs.
+%
+Back to PBT, a related idea called \emph{targeted property-based testing} (TPBT)
+is to use fitness functions to guide the testing efforts towards user defined
+goals.
+%
+\emph{Target} \cite{targetedpbt} is a tool that implements TPBT using
+optimization techniques like hill climbing and simulated annealing.
+
+% Coverage-guided fuzzing + symbolic execution
+%
+Moreover, a popular technique is to combine coverage-guided fuzzing with ideas
+borrowed from symbolic execution tools like \emph{KLEE} \cite{klee} to avoid
+getting stuck in superficial paths \cite{driller, qsym}.
+%
+When using off-the-shelf symbolic executors, this technique is often limited by
+the path explosion problem, altough recent tools like \emph{QSYM} \cite{qsym}
+demonstrate that using tailored concolic executors can help overcoming this
+limitation.
+
+Most of these ideas can potentially be incorporated in our tool, and we keep
+them as a challenge for future work.
 
@@
 %
-\emph{Feat} \cite{feat} formalizes the notion of \emph{functional enumerations},
+\emph{Feat} \cite{feat} formalizes the notion of \emph{functional enumerations}.
+%
 For any algebraic type, it synthesizes a bijection between a finite prefix of
@@
 Later, this bijection can be traversed exhaustively or, more interestingly,
-randomly accesed.
+randomly accessed.
 %
@@
 % SmallCheck and LazySmallCheck
-\emph{SmallCheck} \cite{smallcheck} is a Haskell tool that follows this idea.
-%
-\emph{SmallCheck} is an exhaustive testing tool for Haskell that progressively
-executes the testing properties against all possible inputs values of within
-a certain size bound.
-%
-Later, \emph{LazySmallCheck} is a lazy variant of \emph{SmallCheck} that uses
-lazy evaluation to automatically prune the search space by detecting input
-subexpressions that are not evaluated by the testing properties, and hence do
-not need to be populated exhaustively with concrete values.
+\emph{SmallCheck} \cite{smallcheck} is a Haskell tool that also follows this
+idea.
+%
+It progressively executes the testing properties against all possible inputs
+values up to a certain size bound.
+%
+Later, \emph{LazySmallCheck} is a variant of \emph{SmallCheck} that uses lazy
+evaluation to automatically prune the search space by detecting which
+subexpressions of the generated values are not evaluated by the testing
+properties, and hence do not need to be populated exhaustively with concrete
+values.
 
@@
 %
-In this work we use exhaustiveness as a way to reliably schedule different
-mutants.
+In this work we use exhaustiveness as a way to reliably enforce that all
+possible mutants of an interesting seed are executed.
 %
 In contrast to fully-exhaustive testing tools, \mutagen relies initially on
-randomness to find interesting mutable candidates.
+randomness to find initial interesting mutable candidates.
 %

### 2021-04-01 papers/oopsla-21/main.tex
WIP

@@
 \usepackage{multicol}
+\usepackage{multirow}
+\usepackage{caption}
 \usepackage[ruled,vlined,noend]{algorithm2e}
@@
 
-In both case studies, we additionally show how the heuristics described in
-Section \ref{sec:heuristics} affect the testing performance by disabling them
-when using our tool.
+In both case studies, we additionally show how the FIFO scheduling and trace
+reset heuristics described in Section \ref{sec:heuristics} affect the testing
+performance by disabling them when using our tool.
+%
+We call these variants \textit{no-FIFO} and \textit{no-reset}, respectively.
+%
+In the case of \textit{no-reset}, the amount of random mutations is no longer
+controlled by this heuristic, so we fixed it to 25 random mutations throughout
+the execution of the tool.
 
@@
 
-Aliquam erat volutpat \citep{devai2013edsl}. Nunc eleifend leo vitae magna. In
-id erat non orci commodo lobortis. Proin neque massa, cursus ut, gravida ut,
-lobortis eget, lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus.
-Mauris ac felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque
-nec dui dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a,
-aliquet quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet
-tortor. Nam euismod tellus id erat.
+The results of this case study are shown in Fig. \ref{fig:results:ifc}, ordered
+by the failure rate achieved by \fuzzchick in decreasing order --- notice the
+logarithmic scale used on the MTTF.
+%
+As introduced earlier, \fuzzchick can only find 5 out of the 20 bugs injected
+with a \%100 success rate within the one hour testing budget.
+%
+In turn, our tool manages to find every bug on all runs, and taking less than a
+minute in the worst absolute case.
+%
+These results are somewhat pesimistic towards \fuzzchick if compared to the ones
+presented originally by \citeauthor{lampropoulos2019coverage}.
+%
+However, we want to remark that our experiments encompass 30 independent runs
+instead of the 5 originally used when presenting \fuzzchick.
+%
+Moreover, since the MTTF simply aggregates all runs, regardless of if they found
+a bug or timed out, this metric is quite sensitive to the particular timeout
+used on each experiment, and will tend to inflate the results as soon as the
+failure rate goes below 1.
+%
+For this reason, additionally comparing the failure rate between tools gives us
+a better estimation of the reliability of both tools, where \mutagen shows a
+clear improvement.
+
+\begin{wrapfigure}{r}{0.5\textwidth}
+\vspace{-10pt}
+\footnotesize
+\begin{tabular}{ccccc}
+\hline
+\multicolumn{1}{c}{
+  \multirow{2}{*}{\textbf{Bug}}} &
+  \multicolumn{1}{c}{\multirow{2}{*}{\textbf{\fuzzchick}}} &
+  \multicolumn{1}{c}{\multirow{2}{*}{\textbf{\mutagen}}} &
+  \multicolumn{2}{c}{\textbf{VD.A}} \\
+\cline{4-5}
+\multicolumn{1}{c}{} &
+\multicolumn{1}{c}{} &
+\multicolumn{1}{c}{} &
+\multicolumn{1}{c}{\textbf{Value}} &
+\multicolumn{1}{c}{\textbf{Estimate}} \\
+\hline
+1  & 13974.9 & 3632.7 & 0.89 & {\color{red} large} \\ \hline
+2  & 23962.9 & 7122.3 & 0.90 & {\color{red} large} \\ \hline
+3  & 19678.5 & 4633.9 & 0.89 & {\color{red} large} \\ \hline
+4  & 20398.4 & 16831.2 & 0.72 & {\color{red} medium} \\ \hline
+5  & 17348.1 & 2326.6 & 0.94 & {\color{red} large} \\ \hline
+6  & 10727.1 & 2312.9 & 0.92 & {\color{red} large} \\ \hline
+7  & 5070.7 & 332.9 & 0.91 & {\color{red} large} \\ \hline
+8  & 5596.4 & 298.7 & 0.90 & {\color{red} large} \\ \hline
+9  & 16402.4 & 26342.5 & 0.51 & negligible \\ \hline
+10 & 11553.0 & 25044.5 & 0.55 & negligible \\ \hline
+11 & 11304.3 & 2536.8 & 0.82 & {\color{red} large} \\ \hline
+12 & 18507.3 & 14482.7 & 0.65 & {\color{red} small} \\ \hline
+13 & 17961.5 & 13454.9 & 0.65 & {\color{red} small} \\ \hline
+14 & 10621.9 & 3928.3 & 0.78 & {\color{red} large} \\ \hline
+15 & 23866.3 & 43678.2 & 0.23 & {\color{blue} large} \\ \hline
+16 & 25321.7  & 27602.0 & 0.54 & negligible \\ \hline
+17 & 28515.6 & 52344.9 & 0.47 & negligible \\ \hline
+18 & 24218.6 & 123401.3 & 0.14 & {\color{blue} large} \\ \hline
+19 & 18638.9 & 30682.3 & 0.45 & negligible \\ \hline
+20 & 18883.1 & 15841.9 & 0.60 & {\color{red} small} \\ \hline
+& & \textbf{Mean} & 0.67 & {\color{red} medium} \\
+\hline
+\end{tabular}
+\captionof{table}{\label{table:ifc:tests} Mean tests until first failure and effect size
+  comparison between \fuzzchick and \mutagen when considering only succesfull
+  runs.
+  %
+  Estimates in {\color{blue} blue} and {\color{red} red} indicate results
+  favouring \fuzzchick and \mutagen, respectively. }
+\vspace{-15pt}
+\end{wrapfigure}
+
+If we only consider the succesfull runs where \fuzzchick does not time out, we
+observed a peculiarity.
+%
+This tool either finds bugs relatively quickly (after a few thouthands tests) or
+does not find them at all within the time budget, which suggest that its power
+scheduler assigns some mutation candidates too little energy before they become
+uniteresting and get ultimately discarded.
+%
+Under this consideration, it is fair to think that \fuzzchick might be a better
+choice if we set a shorter timeout and run it several times.
+%
+Thus, to be an improvement over \fuzzchick, our tool must be able to find bugs
+not only more reliably, but also relatively fast!
+
+Table \ref{table:ifc:tests} shows a comparison between the mean number of tests
+required by both tools to find the first failure when we only consider
+successful runs.
+%
+Additionally, we compute the non-parametric Vargha-Delaney A measure \cite{vda,
+  arcuri}, that encodes the probability of \mutagen to yield better results than
+\fuzzchick.
+%
+As it can be observed, using an exhaustive mutation approach can be not only
+more reliable, but also faster, where \mutagen is likely to find bugs faster
+that \fuzzchick in 14 of the 20 bugs injected into the abstract machine.
 
@@
 \end{figure}
+
+In terms of the \mutagen heuristics, we can also arrive to some conclusions.
+%
+In first place, this case study does not seem to be affected by using a FIFO
+scheduling.
+%
+Upon inspection, we found that the reason behind this is that the mutation
+candidate queues remain empty most of the time (generation mode), and when a new
+interesting candidate gets inserted, all its mutants (and their descendents) are
+processed before the next one is enqueued.
+
+On the other hand, disabling the trace saturation heuristic (\textit{no-reset})
+we observed two interesting effects:
+%
+havig fixed the amount of random mutations to 25 adds a seemingly constant
+overhead when finding most of the (easier) bugs, suggesting that we are spending
+worthless time mutating numeric subexpressions inside the generated machine
+states and that a smaller number of random mutations could be equally effective
+to find such bugs.
+%
+However, some of the hardest to find bugs cannot be reliably exposed using this
+amount of random mutations (\#4, \#9, \#12, \#18 and \#10), suggesting that one
+should increase this number even further to be able to discover all bugs within
+the time budget.
+%
+In consequence, we believe this heuristic is effective at automatically tunning
+this internal parameter of our tool, especially when the user is unsure about
+what the best value for it might be.
 

### 2021-04-05 papers/oopsla-21/main.tex
WIO
WIP

@@
     i $\gets$ 0\;
-    TLog, QSucc, QDisc $\gets$ $\varnothing$;\;
+    TLog, QSucc, QDisc $\gets$ $\varnothing$\;
     \While{i $<$ N}{
@@
         , do { l <- arbitrary; r <- arbitrary; return (Node l x r) } ] -- Grow constructor
-mutate_tree (Node l x r) =
+mutate_tree mutate_a (Node l x r) =
   oneof [ return l                                                     -- Return subexpression
@@
     i $\gets$ 0\;
-    TLog, QSucc, QDisc $\gets$ $\varnothing$;\;
+    TLog, QSucc, QDisc $\gets$ $\varnothing$\;
     \While{i $<$ N}{
-      x $\gets$ Pick(QSucc, QDisc, gen, mut)\;
+      x $\gets$ Pick(QSucc, QDisc, gen)\;
       (result, trace) $\gets$ WithTrace(P(x))\;
@@
         \If{\KwNot Discarded(result)}{
-          muts $\gets$ Mutants(x, mut, R)\;
+          muts $\gets$ Mutate(x, mut, R)\;
           Enqueue(QSucc, muts)\;

### 2021-04-06 papers/oopsla-21/main.tex
WIP

@@
   oneof [ do { x' <- mutate_a x; return (Leaf x')}                       -- Mutate recursively
-        , do { l <- arbitrary; r <- arbitrary; return (Node l x r) } ] -- Grow constructor
-mutate_tree mutate_a (Node l x r) =
+        , do { l <- arbitrary; r <- arbitrary; return (Branch l x r) } ] -- Grow constructor
+mutate_tree mutate_a (Branch l x r) =
   oneof [ return l                                                       -- Return subexpression
@@
         , return (Leaf x)                                                -- Shrink constructor
-        , do { l' <- mutate_tree l; return (Node l' x r) }             -- Mutate recursively
-        , do { x' <- mutate_a    x; return (Node l x' r) }             -- Mutate recursively
-        , do { r' <- mutate_tree r; return (Node l x r') } ]           -- Mutate recursively
+        , do { l' <- mutate_tree l; return (Branch l' x r) }             -- Mutate recursively
+        , do { x' <- mutate_a    x; return (Branch l x' r) }             -- Mutate recursively
+        , do { r' <- mutate_tree r; return (Branch l x r') } ]           -- Mutate recursively
 \end{verbatim}
@@
 \begin{verbatim}
-mutate (Leaf x)     = [ PURE (Node def x def) ]   -- Swap constructor
-mutate (Node l x r) = [ PURE l, PURE r            -- Return subexpression
+mutate (Leaf x)       = [ PURE (Branch def x def) ]   -- Swap constructor
+mutate (Branch l x r) = [ PURE l, PURE r              -- Return subexpression
                         , PURE (Leaf x)               -- Swap constructor
-                      , PURE (Node l x l)         -- Rearrange subexpressions
-                      , PURE (Node r x r)         -- Rearrange subexpressions
-                      , PURE (Node r x l) ]       -- Rearrange subexpressions
+                        , PURE (Branch l x l)         -- Rearrange subexpressions
+                        , PURE (Branch r x r)         -- Rearrange subexpressions
+                        , PURE (Branch r x l) ]       -- Rearrange subexpressions
 \end{verbatim}
@@
       \If{\KwNot Discarded(result)}{
-        muts $\gets$ Mutants(x, mut, R)\;
+        muts $\gets$ Mutate(x, mut, R)\;
         {\color{red} prio $\gets$ BranchDepth(TLog, trace)}\; %% NEW
@@
 %
-In other words, we are testing for false positives.
+In other words, we are testing for false negatives.
 %

### 2021-04-07 papers/oopsla-21/main.tex
Wrote introduction

@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title[\mutagen]{\mutagen: Reliable Coverage-Guided, Property-Based Testing using
-  Exhaustive Well-Typed Mutations}
+\title[\mutagen]{\mutagen: Coverage-Guided, Property-Based Testing using
+  Exhaustive Type-Preserving Mutations}
 
@@
 
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
-The contributions of this work are:
-
-\begin{itemize}
-\item Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec
-  facilisis facilisis, est dui fermentum leo, quis tempor ligula erat quis odio.
-  Nunc porta vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum.
-  Aliquam posuere. Nunc aliquet, augue nec adipiscing interdum, lacus tellus
-  malesuada massa, quis varius mi purus non odio.
-\item Pellentesque condimentum,
-  magna ut suscipit hendrerit, ipsum augue ornare nulla, non luctus diam neque
-  sit amet urna. Curabitur vulputate vestibulum lorem. Fusce sagittis, libero
-  non molestie mollis, magna orci ultrices dolor, at vulputate neque nulla
-  lacinia eros. Sed id ligula quis est convallis tempor. Curabitur lacinia
-  pulvinar nibh. Nam a sapien.
-\item Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec
-  facilisis facilisis, est dui fermentum leo, quis tempor ligula erat quis odio.
-  Nunc porta vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum.
-  Aliquam posuere. Nunc aliquet, augue nec adipiscing interdum, lacus tellus
-  malesuada massa, quis varius mi purus non odio.
-\item Pellentesque condimentum, magna ut suscipit hendrerit, ipsum augue ornare
-  nulla, non luctus diam neque sit amet urna. Curabitur vulputate vestibulum
-  lorem. Fusce sagittis, libero non molestie mollis, magna orci ultrices dolor,
-  at vulputate neque nulla lacinia eros. Sed id ligula quis est convallis
-  tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-\end{itemize}
+%% PBT and generation of interesting data for free
+%
+Random Property-Based Testing (RPBT) is a popular technique for finding bugs
+using executable specifications (properties) encoding the expected behavior of
+our system.
+%
+Exemplified by frameworks Haskell's \quickcheck \cite{quickcheck} and a pletora
+of ports of this tool on other programming languages \cite{qcports}, the most
+common approach is to instantiate the testing properties using independent,
+randomly generated inputs.
+
+Users of \quickcheck (and RPBT in general) are well aware that one of the
+biggest challenges while using these frameworks is to provide suitable random
+data generators that are used to instantiate the testing properties.
+%
+In some extreme cases, writing highly-tuned random generators capable of
+exercising every part of a complex system in a reasonable basis can take several
+thouthands person-hours of trial and error \cite{fuzzchick}.
+
+To alleviate this, there exist several techniques to automatically synthesize
+random data generators by reyfing the static information present in the
+codebase, e.g., data type definitions, application public interfaces (APIs),
+etc.
+%
+These techniques, however, are unable to synthesize random generators capable of
+producing data satisfying complex invariants not easily derivable from the
+codebase, being randomly generated programs are common example of this problem.
+%
+Moreover, automatically synthesized random generators offer particularly poor
+results when the testing properties are constrained by sparse preconditions.
+%
+In this scenario, most of the generated tests are simply discarded before they
+can be used to test the main components of our system.
+
+%% FuzzChick: PBT + code instrumentation + type-preserving mutations
+\fuzzchick \cite{lampropoulos2019coverage} is a property-based testing framework
+for the Coq programming language that borrows ideas from the fuzzing community
+to generate highly structured values while using automatically derived
+generators.
+%
+Notably, this tool is implemented as an extension of \quickchick, Coq's own
+reimplementation of \quickcheck.
+
+Instead of continuously generating random invalid test cases from scratch,
+\fuzzchick keeps a queue of interesting previously executed test cases that can
+be mutated using type-preserving transformations in order to produce new ones.
+%
+Intuitively, mutating an interesting test case in a small way (at the data
+constructor level) has a much higher probability of producing a new interesting
+test case than generating a new one from scratch using a na\"ive generator.
+%
+\fuzzchick is likely to preserve the semantic structure of the mutated data, as
+mutations are applied directly at the data type level -- the random AST in case
+of generating code.
+
+This tool relies on execution traces to distinguish which test cases resulted
+interesting and are therefore worth mutating.
+%
+Mutated test cases are considered interesting for mutation only when they
+produce new execution traces -- any other test case is simply thrown away.
+
+Of course, this hybrid technique is not meant to replace smart manually-written
+generators.
+%
+In turn, it provides an acceptable solution that can be used in the early stages
+of development, and while a better test suite using manually-written generators
+is still under construction.
+
+%% Problems of FuzzChick: mutations are superficial, long queues, power
+%% scheduler
+%
+In this work, we take a close look at \fuzzchick for opportunities to improve.
+%
+Notably, we establish several aspects of its implementation that can be revised
+--- some of these already being identified by the authors of \fuzzchick on their
+original work.
+%
+In particular:
+%
+\begin{inparaenum}
+\item \emph{the type-preserving mutations produced by this tool are superficial
+  and rely heavily on randomness to produce diverse mutants},
+\item \emph{using normal queues can cause large delays when interesting values
+  are enqueued frequently}, and
+\item \emph{using a power schedule to assing a certain mutation energy to each
+  interesting test case requires fine tuning and can be hard to generalize to
+  work well under different testing scenarios.}
+\end{inparaenum}
+
+Moreover, we replicated the IFC stack machine case study bundled with \fuzzchick
+and observed a noticeable limitation:
+%
+\emph{after repeating each experiment 30 times, \fuzzchick was only able to find
+  5 out of the 20 injected bugs with a 100\% efficacy, the hardest one being
+  found only around 13\% of the time after an hour of testing.}
+%
+Despite this limitation, \citeauthor{lampropoulos2019coverage} show that
+\fuzzchick is far better than using normal random testing coupled with na\"ive
+random generators by comparing the mean-time-to-failure (MTTF) against
+\emph{QuickChick}.
+%
+However, we believe that a proper evaluation ought to use the failure rate as an
+important metric when comparing RPBT tools.
+
+%% Our solution
+In this work we introduce \mutagen, a RPBT tool implemented in Haskell that
+build upon the basic idea behind \fuzzchick, i.e., code instrumentation and
+type-preserving mutations.
+
+%% Exhaustive mutations
+Unlike \fuzzchick, mutations in \mutagen are applied \emph{exhaustively}.
+%
+This is, given an interesting, previously executed test case, our tool
+precomputes and schedules every possible type-presserving mutation that can be
+applied to it.
+%
+This approach has two inherent advantages when compared to \fuzzchick.
+%
+On one hand, mutations do not rely on randomness to be generated.
+%
+Instead, every subexpression of the input test case is mutated on the same
+basis, which guarantees that no interesting mutation is left behind due to
+randomness.
+%
+On the other hand, scheduling mutations exaustively eliminates the need of using
+a power schedule to assign a given mutation energy to each input test case.
+
+Notably, \emph{scheduling mutations exhaustively within an input test case does
+  not mean that mutations are exhaustive by themselves.}
+%
+Our tool distinguishes two kinds of mutations, those that can be computed
+purely, yielding a single mutated value; and those that can be obtained
+non-deterministically.
+%
+Pure mutations encode transformations that can be done by swapping data
+constructor variants or rearranging subexpressions, whereas non-deterministic
+(random) mutations are useful to represent transformations over large
+enumeration types, e.g., numbers, characters, etc.
+%
+In this setting, random mutations let us selectively escape the exhaustiveness
+of our tool by randomly sampling a small random generator a reduced number of
+times.
+%
+This way, our tool avoids, for instance, mutating an given integer into every
+other integer in the 32 bits range.
+
+%% Heuristics to improve testing loop
+Additionally, the testing loop of our tool includes implements two heuristics
+that help finding bugs faster.
+%
+In first place, \mutagen uses first-in first-out (FIFO) scheduling with priority
+to enqueue interesting test cases for mutation.
+%
+This scheduling algorithm is indexed by the novelty of the test case with
+respect to the ones already seen.
+%
+In this light, test cases that discover new parts of the code earlier during
+execution are given a higher priority while picking the next mutated test case.
+%
+Additionally, our scheduling lets the testing loop jump back and forth between
+scheduled test cases as soon as a new more interesting one is enqueued for mutation,
+eliminating potential delays when the mutation queues grow large.
+
+The second heuristic is used to tune the number of times our tool samples random
+mutants, i.e., those corresponding to large enumeration types as described above.
+%
+For this, we keep track of how often we generate interesting test cases.
+%
+If this frequency suddenly stalls, \mutagen automatically resets the testing
+loop increasing the amount of times we sample such mutants.
+%
+This automatically finds a suitable value for this external on the fly.
+
+%% Validation
+To validate our ideas, we use two main case studies (described in detail in
+Section \ref{sec:casestudies}): the IFC stack machine used by
+\citeauthor{lampropoulos2019coverage} on their original work, and a WebAssembly
+engine written in Haskell.
+%
+In both case studies, we additionally compare the effect of the heuristics
+implemented on top of the base testing loop of \mutagen.
+%
+In the IFC stack machine case study, we compare \fuzzchick against different
+variants of our tool, where our results (Section \ref{sec:evaluation}) indicate
+that \mutagen outperforms \fuzzchick both in terms of time to failure and
+failure rate.
+%
+On the other hand, our WebAssembly case study shows the performance of our tool
+in a more realistic scenario.
+%
+There, \mutagen is capable of reliaby finding 15 planted bugs in the validator
+and interpreter, as well as 3 previously unknown bugs that flew under the radar
+of the existing unit test suite of this engine.
+%
+All in all, our evaluation encompass ?? hours of CPU time and suggest that
+testing mutants exhaustively is an appealing technique for finding bugs fast
+without sacrificing efficacy.
+
+%% Last two sections
+
+Finally, we discuss related work on Section \ref{sec:related} and conclude
+on Section \ref{sec:conclusions}.
 
@@
 
-In this section we briefly introduce the concept of Property-based Testing along
-with \quickcheck, the most popular tool of this sort and often considered as the
-baseline when comparing PBT algoritms.
-%
-Moreover, we describe the ideas and limitations behind \fuzzchick, which
-enhances the PBT approach with execution information and serves as the
-foundation for \mutagen.
-%
-\fuzzchick is implemented as an extension of \quickchick, Coq's own
-reimplementation of \quickcheck.
+In this section we briefly introduce the concept of Property-Based Testing along
+with \quickcheck, one of the most popular tools of this sort and often used as
+the baseline when comparing PBT algoritms.
+%
+Moreover, we describe in detail the ideas and limitations behind \fuzzchick,
+which served as the foundation while designing \mutagen.
 
@@
   each test case; and
-\item \emph{high-level, well-typed mutations}, to produce syntactically valid
-  test cases by altering existing ones at the datatype level.
+\item \emph{high-level, type-preserving mutations}, to produce syntactically
+  valid test cases by altering existing ones at the datatype level.
 \end{inparaenum}
@@
 
-The next subsections describe \fuzzchick's testing loop and well-typed mutations
-in detail.
+The next subsections describe \fuzzchick's testing loop and type-preserving
+mutations in detail.
 
@@
 
-\subsubsection{Well-typed mutations}
+\subsubsection{Type-preserving mutations}
 
@@
 
-However, the authors acknowledge that certain parts of their implementation have
+However, the authors acknowledge that certain parts of its implementation have
 room for improvement, especially when it comes to the mutators design.
 %
-Moreover, when we recreated the evaluation of the IFC stack machine case study
-(described in detail in Section \ref{sec:casestudies}), we found that after 30
-runs (as opposed to the 5 runs used original by
-\citeauthor{lampropoulos2019coverage}), \emph{\fuzzchick was only able to find 5
-  out of the 20 injected bugs with a failure rate of 1}, the hardest one being
-found only around 13\% of the time after an hour of testing.
-%
-These results are presented in detail in Section \ref{sec:evaluation}.
-
-At the light of these observations, we identified several aspects of \fuzzchick
-that can be improved upon --- and that constitute the main goal of this work.
+In tandem with the lack of efficacy observed when replicating the evaluation of
+the IFC stack machine case study, these observations led us to consider three
+main aspects of \fuzzchick that can be improved upon --- and that constitute the
+main goal of this work.
+%
+%% As mentioned earlier, when we recreated the evaluation of the IFC stack machine
+%% case study (described in detail in Section \ref{sec:casestudies}), we found that
+%% after 30 runs (as opposed to the 5 runs used original by
+%% \citeauthor{lampropoulos2019coverage}), \emph{\fuzzchick was only able to find 5
+%%   out of the 20 injected bugs with a failure rate of 1}, the hardest one being
+%% found only around 13\% of the time after an hour of testing.
+%% %
+%% These results are presented in detail in Section \ref{sec:evaluation}.
+%
+%% At the light of these observations, we identified three main aspects of
+%% \fuzzchick that can be improved upon --- and that constitute the main goal of
+%% this work.
 %
@@
   assign energy to each mutable test case in \fuzzchick works in the context of
-  high-level well-typed mutations.
+  high-level type-preserving mutations.
   %
@@
 \begin{wrapfigure}{r}{0.5\textwidth}
+\centering
 \vspace{-10pt}
@@
 \multicolumn{1}{c}{
-  \multirow{2}{*}{\textbf{Bug}}} &
-  \multicolumn{1}{c}{\multirow{2}{*}{\textbf{\fuzzchick}}} &
-  \multicolumn{1}{c}{\multirow{2}{*}{\textbf{\mutagen}}} &
-  \multicolumn{2}{c}{\textbf{VD.A}} \\
+  \multirow{2}{*}{Bug}} &
+  \multicolumn{1}{c}{\multirow{2}{*}{\fuzzchick}} &
+  \multicolumn{1}{c}{\multirow{2}{*}{\mutagen}} &
+  \multicolumn{2}{c}{$A_{12}$ measure} \\
 \cline{4-5}
@@
 \multicolumn{1}{c}{} &
-\multicolumn{1}{c}{\textbf{Value}} &
-\multicolumn{1}{c}{\textbf{Estimate}} \\
+\multicolumn{1}{c}{Value} &
+\multicolumn{1}{c}{Estimate} \\
 \hline
@@
 20 & 18883.1 & 15841.9 & 0.60 & {\color{red} small} \\ \hline
-& & \textbf{Mean} & 0.67 & {\color{red} medium} \\
+& & Mean & 0.67 & {\color{red} medium} \\
 \hline
@@
 %
-Additionally, we compute the non-parametric Vargha-Delaney A measure \cite{vda,
-  arcuri}, that encodes the probability of \mutagen to yield better results than
-\fuzzchick.
+Additionally, as adviced by \cite{arcuri} when comparing random testing tools,
+we computed the non-parametric Vargha-Delaney $A\_{12}$ measure \cite{vda}.
+%
+Intuitively, this measure encodes the probability of \mutagen to yield better
+results than \fuzzchick.
 %
@@
   \caption{\label{fig:results:wasm} Comparison between \quickcheck and \mutagen
-    across 15 different bugs for the WebAssembly case study. }
+    across 15 different bugs injected into \textit{haskell-wasm}. }
 \end{figure}
@@
 However, more specialized generators can by used directly by \mutagen, and
-combining them with well-typed mutations using a hybrid technique is an
+combining them with type-preserving mutations using a hybrid technique is an
 appealing idea that we aim to address in the future.
@@
 values up to a certain size bound.
-%
-Later, \emph{LazySmallCheck} is a variant of \emph{SmallCheck} that uses lazy
-evaluation to automatically prune the search space by detecting which
-subexpressions of the generated values are not evaluated by the testing
-properties, and hence do not need to be populated exhaustively with concrete
-values.
 
@@
 test cases up to a given small size.
-%
-Similar to \emph{LazySmallCheck}, \emph{Korat} uses method preconditions to
-prune the search space.
-%
-However, in \emph{Korat}, pruning is done by analyzing the execution trace of
-the precondition predicate instead of by keeping track of unused subexpressions.
-
-Being exhaustive, these approaches can only be used to enumerate and test a
-relatively little space of small inputs.
-%
+
+Being exhaustive, these approaches rely on prunning mechanisms to avoid
+populating unevaluated subexpresions exhaustively before the computational cost
+becomes too restrictive.
+%
+\emph{LazySmallCheck} is a variant of \emph{SmallCheck} that uses lazy
+evaluation to automatically prune the search space by detecting unevaluated
+subexpressions using lazy evaluation.
+%
+In the case of \emph{Korat}, pruning is done by instrumenting method
+precondition predicates and analyzing which parts of the execution trace
+correspond to each evaluated subexpression.
+
 In this work we use exhaustiveness as a way to reliably enforce that all
@@
 %
-In contrast to fully-exhaustive testing tools, \mutagen relies initially on
+In contrast to fully-exhaustive testing tools, \mutagen initially relies on
 randomness to find initial interesting mutable candidates.
 %
-Later, enumerating mutations for a given interesting candidate is much cheaper
-than enumerating \emph{all posible values}.
-%
-Hence, our tool can be used to produce much larger test cases before their size
-becomes computationally too restrictive.
+In terms of pruning, it is possible to instruct \mutagen to detect unused
+subexpressions (using a technique similar to that of \emph{LazySmallCheck}) to
+avoid scheduling mutations over their corresponding positions.
+%
+This could, in principle, improve the overall performance when testing
+properties that tend to short-circuit, leaving parts of the input unevaluated.
+%
+In our case studies, however, we observed that their preconditions tend to be
+quite strict when mutating from passed test cases, fully evaluating their input
+before executing the postcondition.
+%
+Thus, we do not expect to see considerable improvements by applying this idea in
+this particular scenario.
+%
+Gathering empirical evidence about using prunning via lazy evaluation in
+\mutagen is an effort that we keep as future work.
 
@@
 Inspired by \fuzzchick, our tool uses coverage information to guide
-automatically derived mutators producing high-level, well-typed mutations.
+automatically derived mutators producing high-level, type-preserving mutations.
 %
@@
 Our experimental results indicate that \mutagen outperforms the simpler approach
-taken by \fuzzchick in terms of failure rate.
+taken by \fuzzchick in terms of both failure rate and tests until first failure.
 %

### 2021-04-07 papers/oopsla-21/main.tex
Added WebAssembly evaluation

@@
 In no particular order:
-%
-\begin{itemize}
-\item \emph{Mutators distribution:} if we inspect the mutator defined in Fig.
-  \ref{fig:fuzzchick:mutator}, there are two compromises that the authors of
-  \fuzzchick adopted for the sake of simplicity.
-  %
-  On one hand, deep recursive mutations are very unlikely, since their
-  probability decreases multiplicatively with each recursive call.
+
+\paragraph{Mutators distribution:}
+%
+if we inspect the mutator defined in Fig. \ref{fig:fuzzchick:mutator}, there are
+two compromises that the authors of \fuzzchick adopted for the sake of
+simplicity.
+%
+On one hand, deep recursive mutations are very unlikely, since their probability
+decreases multiplicatively with each recursive call.
 %
@@
 %
-  Hence, \fuzzchick mutators can only be effectively used to transform to
-  shallow data structures, potentially excluding interesting applications that
-  might require producing deeper valid values, e.g., programming languages,
-  network protocols interactions, etc.
-  %
-  Ideally, mutations should be able to happen on every subexpression of the
-  input seed in a reasonable basis.
+Hence, \fuzzchick mutators can only be effectively used to transform to shallow
+data structures, potentially excluding interesting applications that might
+require producing deeper valid values, e.g., programming languages, network
+protocols interactions, etc.
+%
+Ideally, mutations should be able to happen on every subexpression of the input
+seed in a reasonable basis.
 
 On the other hand, using random generators to produce neeeded subexpressions
-  when growing data constructors can be dangerous, as we are introducing the
-  very same ``uncontrolled'' randomness that we wanted to mitigate in the first
-  place!
+when growing data constructors can be dangerous, as we are introducing the very
+same ``uncontrolled'' randomness that we wanted to mitigate in the first place!
 %
@@
 %
-  For instance, by using just a minimal piece of data to make the overall
-  mutated test case type correct.
+For instance, by using just a minimal piece of data to make the overall mutated
+test case type correct.
 %
@@
 
-\item \emph{Enqueuing mutation candidates:} \fuzzchick uses two single queues for
-  keeping valid and discarded mutation candidates.
-  %
-  Whenever a new test case is found interesting, it is placed \emph{at the end
-    of its corresponding queue}.
+\paragraph{Enqueuing mutation candidates:}
+%
+\fuzzchick uses two single queues for keeping valid and discarded mutation
+candidates.
+%
+Whenever a new test case is found interesting, it is placed \emph{at the end of
+  its corresponding queue}.
 %
@@
 %
-  This can limit the effectiveness of the testing loop if the queues tend to
-  grow more often than they tend to shrink, as interesting mutation candidates
-  can get buried at the end of a long queue that only exercises the same portion
-  of the target code.
+This can limit the effectiveness of the testing loop if the queues tend to grow
+more often than they tend to shrink, as interesting mutation candidates can get
+buried at the end of a long queue that only exercises the same portion of the
+target code.
 %
@@
 %
-  Ideally, one would like a mechanism that prioritizes mutating test cases
-  that discovers new portions of the code right away, and that is capable of
-  jumping back and forth from mutation candidates whenever this happens.
+Ideally, one would like a mechanism that prioritizes mutating test cases that
+discovers new portions of the code right away, and that is capable of jumping
+back and forth from mutation candidates whenever this happens.
 %
 We show in Section \ref{sec:heuristics} how this can be achieved by analyzing
-  the execution information in order to prioritize test cases with novel
-  execution traces.
-
-\item \emph{Power schedule:} It is not clear how the power schedule used to
-  assign energy to each mutable test case in \fuzzchick works in the context of
-  high-level type-preserving mutations.
+the execution information in order to prioritize test cases with novel execution
+traces.
+
+\paragraph{Power schedule:}
+%
+It is not clear how the power schedule used to assign energy to each mutable
+test case in \fuzzchick works in the context of high-level type-preserving
+mutations.
 %
@@
 within the small available energy window --- randomly generating the same test
-  case later on does not help, as it becomes uninteresting based on historic
-  trace information.
+case later on does not help, as it becomes uninteresting based on historic trace
+information.
 
@@
 when using a high-level mutation approach --- something quite challenging to
-  characterize in general given the expressivity of the data types used to
-  drive the mutators.
-\end{itemize}
+characterize in general given the expressivity of the data types used to drive
+the mutators.\\
 
@@
 
-\begin{figure}[b]
+\begin{figure}[t]
 \begin{multicols}{2}
@@
 \textit{haskell-wasm}.
+
+\begin{table}[t]
+\footnotesize
+\begin{tabular}{|c|c|l|}
+\hline
+\textbf{Id}
+& \textbf{Subsystem}
+& \textbf{Description} \\
+\hline
+1
+& Validator
+& Wrong if-then-else type validation on else branch \\
+\hline
+2
+& Validator
+& Wrong stack type validation \\
+\hline
+3
+& Validator
+& Removed function type mismatch assertion \\
+\hline
+4
+& Validator
+& Removed max memory instances assertion \\
+\hline
+5
+& Validator
+& Removed function index out-of-range assertion \\
+\hline
+6
+& Validator
+& Wrong type validation on \texttt{i64.eqz} instruction \\
+\hline
+7
+& Validator
+& Wrong type validation on \texttt{i32} binary opertations \\
+\hline
+8
+& Validator
+& Removed memory index out-of-range assertion \\
+\hline
+9
+& Validator
+& Wrong type validation on \texttt{i64} constants \\
+\hline
+10
+& Validator
+& Removed memory alignment validation on \texttt{i32.load} instruction \\
+\hline
+11
+& Interpreter
+& Wrong intepretation of \texttt{i32.sub} instruction \\
+\hline
+12
+& Interpreter
+& Wrong intepretation of \texttt{i32.lt\_u} instruction when operands are equal \\
+\hline
+13
+& Interpreter
+& Wrong intepretation of \texttt{i32.shr\_u} instruction \\
+\hline
+14
+& Interpreter
+& Wrong local variable initialization \\
+\hline
+15
+& Interpreter
+& Wrong memory address casting on \texttt{i32.load8\_s} instruction \\
+\hline
+\end{tabular}
+\caption{\label{table:wasm:injectedbugs}Bugs injected in \textit{haskell-wasm}.}
+\end{table}
 
@@
 %
-We call these variants \textit{no-FIFO} and \textit{no-reset}, respectively.
-%
-In the case of \textit{no-reset}, the amount of random mutations is no longer
+We call these variants \textit{no FIFO} and \textit{no reset}, respectively.
+%
+In the case of \textit{no reset}, the amount of random mutations is no longer
 controlled by this heuristic, so we fixed it to 25 random mutations throughout
@@
 \subsection{IFC Stack Machine}
+
+\begin{figure}[b]
+  \centering
+  \includegraphics[width=\linewidth]{tikz/ifc-mttf.tikz}
+  \includegraphics[width=\linewidth]{tikz/ifc-fr.tikz}
+  \caption{\label{fig:results:ifc} Comparison between \fuzzchick and \mutagen
+    across 20 different bugs for the IFC stack machine. }
+\end{figure}
 
@@
 \centering
-\vspace{-10pt}
+\vspace{-20pt}
 \footnotesize
@@
   favouring \fuzzchick and \mutagen, respectively. }
-\vspace{-15pt}
+\vspace{-25pt}
 \end{wrapfigure}
@@
 Additionally, as adviced by \cite{arcuri} when comparing random testing tools,
-we computed the non-parametric Vargha-Delaney $A\_{12}$ measure \cite{vda}.
+we computed the non-parametric Vargha-Delaney $A_{12}$ measure \cite{vda}.
 %
@@
 that \fuzzchick in 14 of the 20 bugs injected into the abstract machine.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{tikz/ifc-mttf.tikz}
-  \includegraphics[width=\linewidth]{tikz/ifc-fr.tikz}
-  \caption{\label{fig:results:ifc} Comparison between \fuzzchick and \mutagen
-    across 20 different bugs for the IFC stack machine. }
-\end{figure}
 
@@
 
-On the other hand, disabling the trace saturation heuristic (\textit{no-reset})
+On the other hand, disabling the trace saturation heuristic (\textit{no reset})
 we observed two interesting effects:
@@
 
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
+The results of this case study are shown in Fig. \ref{fig:results:wasm}, ordered
+by the mean-time-to-failure achieved by \mutagen.
+%
+Given that we used different properties to test the WebAssembly validator and
+interpreter, we will focus on these results independently.
+
+Firstly, we focus on the bugs injected in the validator (Fig.
+\ref{fig:results:wasm} left).
+%
+There, we can quickly conclude that \quickcheck is well suited to find most of
+the bugs --- it merely finds a counterexample for the easier bugs \#4 and \#5 in
+just 1 out of 30 runs!
+%
+The reason behind this simple: WebAssembly modules comprise several different
+interconnected components, e.g., types and functions definitions, memory blocks,
+imports, exports, etc.
+%
+All these components need to be valid and agree with each other in order for the
+module to be correctly validated (modulo the injected bugs).
+%
+Using an automatically derived generator that generates each one of these
+components independently is virtually unable to produce valid WebAssembly
+modules.
+%
+Using the same random generator, however, \mutagen is able to consistently find
+every bug in less than 20 seconds in the absolute worst case.
+
+In terms of the heuristics implemented in our tool, we can observe two clear
+phenomena.
+%
+On one hand, disabling the FIFO schedulng (case \textit{no FIFO}), the
+hardest-to-find bugs (\#7, \#8 and \#1) are no longer found on every run.
+%
+Moreover, although bugs \#2, \#9, \#8 and \#6 are found with \%100 success rate
+across runs, it takes several times longer for \mutagen to find them.
+%
+On the other hand, disabling the the trace saturation heuristic (case \textit{no
+  reset}) and using a fixed number of 25 random mutation does not affect the
+effectiveness of our tool, apart from adding again a mostly constant overhead on
+the time required to find every bug --- suggesting that the value we arbitrarily
+chose for this parameter is too large and should be left to be handled
+automatically by our tool.
+
+If we now pay attention to the bugs injected into the interpreter (Fig.
+\ref{fig:results:wasm} right), we notice that finding bugs requires
+substantially more time (minutes instead of seconds), since both interpreters
+need to validate and run the inputs before producing a result to compare.
+%
+Perhaps more interestingly, we can observe a significant improvement on the
+performance of \quickcheck in terms of failure rate.
+%
+This is of no surprise: we deliberately reduced the search space by using a
+module stub when defining \texttt{prop\_interpreter} in our test suite.
+%
+Nonetheless, \mutagen still outperforms \quickcheck both in terms of failure
+rate and mean time to failure.
+%
+Moreover, by disabling the FIFO scheduling we observe that the performance of
+our tool gets more unstable, where the bugs \#12 and \#11 take substantially
+less time to be found, whereas the bugs \#15 and \#14 cannot be found on every
+run.
+%
+Finding where the best tradeoff lies in cases where the heuristics cause mixed
+results like this is an interesting direction to aim future work.
 
@@
 
+Finally, this case study gives us the opportunity to analyze the overhead
+introduced by the code instrumentation and internal processing used in \mutagen
+versus the stateless black-box approach of \quickcheck.
+%
+Table \ref{table:wasm:overhead} compares the total number of executed and passed
+test per second that we obtained using each tool.
+%
+It is easy to conclude that \mutagen executes tests several times slower than
+\quickcheck -- roughly 16x and 70x slower when testing \textit{prop\_validator}
+and \textit{prop\_interpreter\_i32}, respectively.
+%
+Part of the reason behind this slowdown is that the code instrumentation our
+tool currently uses is built upon a superficial tranformation that logs traces
+at the user level.
+%
+We anticipate that a large portion of the overhead could be reduced considerably
+by integrating the instrumentation directly into the Haskell runtime --- a
+substantial challenge to tackle in the future.
+
+More importantly, in the same time, our tool is capable of running many more
+tests that pass the sparse preconditions than \quickcheck, and which ultimately
+lead us to finding bugs.
+
+\begin{table}[H]
+\begin{tabular}{|l|c|c|c|c|}
+\hline
+\multicolumn{1}{|c|}{
+  \multirow{2}{*}{Property}}
+  & \multicolumn{2}{c|}{\quickcheck}
+  & \multicolumn{2}{c|}{\mutagen} \\
+  \cline{2-5}
+  \multicolumn{1}{|c|}{}
+  & \multicolumn{1}{c|}{total}
+  & passed
+  & \multicolumn{1}{c|}{total}
+  & passed \\
+\hline
+\textit{prop\_validator}
+& 42136
+& 0.00028
+& 2583
+& 539.19 \\
+\hline
+\textit{prop\_interpreter\_i32}
+& 134000
+& 23.16
+& 1890
+& 372.75 \\
+\hline
+\end{tabular}
+\caption{\label{table:wasm:overhead}Mean number of tests per second executed by
+  each tool on the WebAssembly case study.}
+\end{table}
+
 % ----------------------------------------
@@
 This is in turn inspired by exhaustive bounded testing tools that focus on
-testing every possible input value of the system under test --- a more generic
-technique of limited applicability.
+testing every possible input value of the system under test.
+
+%% --- a more generic technique of limited applicability.
 
@@
 
-In the future, we will investigate how to redefine our automatically synthesized
-mutators in a stateful manner.
+As already indicated throughout this work, there are several direction for
+future work we aim to investigate:
+%
+combining \mutagen with specialized generators; improving the code
+instrumentation in order to effectively target mutations towards specific parts
+of the code; and evaluating the effect of lazy prunning; among others.
+%
+In addition, we want to investigate how to redefine our automatically
+synthesized mutators in a stateful manner.
 %
 This way, it would be possible to apply mutations that preserve complex
-properties of the generated data like well-scopedness and well-typedness simply
-by construction, e.g., mutations that \emph{always} produce well-typed
-subexpressions that refer only to identifiers in the current scope.
+properties of the generated data simply by construction, e.g., mutations that
+\emph{always} produce well-typed subexpressions that refer only to identifiers
+in scope.
 %

### 2021-04-08 papers/oopsla-21/main.tex
WIP

@@
 
+\linepenalty=1000
+\widowpenalty=10000
+\clubpenalty=10000
+
 % ----------------------------------------
@@
 Random Property-Based Testing (RPBT) is a popular technique for finding bugs
-using executable specifications (properties) encoding the expected behavior of
-our system.
-%
-Exemplified by frameworks Haskell's \quickcheck \cite{quickcheck} and a pletora
-of ports of this tool on other programming languages \cite{qcports}, the most
-common approach is to instantiate the testing properties using independent,
-randomly generated inputs.
+using executable properties encoding the expected behavior of our system.
+%
+Exemplified by frameworks Haskell's \quickcheck \cite{ClaessenH00} and the
+abundant ports of this tool in other programming languages
+\cite{hughes2003erlang, papadakis2011proper, bulwahn2012new,
+  denes2014quickchick}, the most common approach is to instantiate the testing
+properties using randomly generated inputs.
 
@@
 exercising every part of a complex system in a reasonable basis can take several
-thouthands person-hours of trial and error \cite{fuzzchick}.
+thouthands person-hours of trial and error \cite{lampropoulos2019coverage}.
 
@@
 codebase, e.g., data type definitions, application public interfaces (APIs),
-etc.
+etc. \cite{GriecoCB16, DBLP:conf/haskell/MistaRH18, Mista2019GeneratingRS,
+  DuregardJW12, Lampropoulos2017, Bendkowski2017}.
 %
@@
 producing data satisfying complex invariants not easily derivable from the
-codebase, being randomly generated programs are common example of this problem.
+codebase.
+%
+Randomly generating valid programs is common example of this problem, where
+developers are forced to put a substantial effort writing complex generators by
+hand \cite{Palka11, perenyi2020stack, yang2011finding}.
 %
@@
 %
-Notably, this tool is implemented as an extension of \quickchick, Coq's own
-reimplementation of \quickcheck.
+Notably, this tool is implemented as an extension of \quickchick
+\cite{denes2014quickchick}, Coq's own reimplementation of \quickcheck.
 
@@
 This tool relies on execution traces to distinguish which test cases resulted
-interesting and are therefore worth mutating.
+interesting and are therefore worth mutating --- a prolific technique known
+\emph{coverage-guided fuzzing}, popularized by tools like \emph{AFL} \cite{afl}.
 %
@@
 type-preserving mutations.
+%
+\mutagen is introduced in detail in Section \ref{sec:mutagen}.
 
@@
 %
-On the other hand, scheduling mutations exaustively eliminates the need of using
-a power schedule to assign a given mutation energy to each input test case.
+On the other hand, scheduling mutations exaustively eliminates the need of a
+heuristic power schedule to a given mutation energy to each input test
+case.
 
@@
 %
-This way, our tool avoids, for instance, mutating an given integer into every
-other integer in the 32 bits range.
+This way, our tool avoids, for instance, mutating a given number into every
+other number in the 32 or 64 bits range.
 
 %% Heuristics to improve testing loop
-Additionally, the testing loop of our tool includes implements two heuristics
-that help finding bugs faster.
+Additionally, the testing loop of our tool incorporates two heuristics that help
+finding bugs faster and more reliably (Section \ref{sec:heuristics}).
 %
@@
 %
-This scheduling algorithm is indexed by the novelty of the test case with
+This scheduling algorithm is indexed by the novelty of each new test case with
 respect to the ones already seen.
@@
 %
-This automatically finds a suitable value for this external on the fly.
+This automatically finds a suitable value for this external parametere on the
+fly.
 
@@
 %
-All in all, our evaluation encompass ?? hours of CPU time and suggest that
-testing mutants exhaustively is an appealing technique for finding bugs fast
-without sacrificing efficacy.
+All in all, our evaluation encompass more than 600 hours of cpu time and suggest
+that testing mutants exhaustively can be an appealing technique for finding bugs
+reliably without sacrificing speed.
 
 %% Last two sections
-
 Finally, we discuss related work on Section \ref{sec:related} and conclude
@@
 
-In this section we briefly introduce the concept of Property-Based Testing along
-with \quickcheck, one of the most popular tools of this sort and often used as
-the baseline when comparing PBT algoritms.
+In this section, we briefly introduce the concept of Property-Based Testing
+along with \quickcheck, one of the most popular tools of this sort and often
+used as the baseline when comparing PBT algoritms.
 %
@@
 \quickcheck, this technique focuses on aiming the developer's efforts into
-testing systems via executable specifications using randomly generated inputs.
+testing systems via executable properties using randomly generated inputs.
 %
@@
 \begin{itemize}
-\item one or more \emph{executable properties}, often implemented simply as a
+\item one or more \emph{executable properties}, often implemented simply as
   boolean predicates,
@@
 Perhaps the simplest PBT technique is to repeatedly generate random inputs and
-instantiate the testing properties until they either get falsified by a
+use them instantiate the testing properties until they either get falsified by a
 counterexample, or we ran a sufficiently large amount of tests --- suggesting
-that the properties holds.
+that the properties hold.
 %
@@
 
-In particular, automatically derived generators are remarkably uneffective when
-used to test properties with sparse preconditions.
+In particular, \emph{automatically derived generators are remarkably uneffective
+  when used to test properties with sparse preconditions.}
 %
@@
 %
-Generic fuzzing tools like AFL, libFuzzer or HonggFuzz, as well as
-language-specific ones like Crowbar use execution traces to recognize
-interesting test cases, e.g, those that exercise previously undiscovered parts
-of the target code.
+Generic fuzzing tools like \emph{AFL}, \emph{libFuzzer} \citeyearpar{libfuzzer}
+or \emph{HonggFuzz} \cite{honggfuzz}, as well as language-specific ones like
+Crowbar \cite{dolan2017testing} or \emph{Kelinci} \cite{kersten2017poster} use
+execution traces to recognize interesting test cases, e.g, those that exercise
+previously undiscovered parts of the target code.
 %
@@
 Enqueued values are stored along with a given mutation energy, that controls how
-many times a given test case can be mutated before being finally discared.
+many times a given test case can be mutated before being finally thrown away.
 
@@
 If the test case does in fact discover a new path, it is enqueued at the end of
-its corresponding queue, depending on whether it passed or was discarded.
+its corresponding queue, depending on whether it passed or was discarded (i.e.,
+it failed the precondition).
 %
@@
 \end{multicols}
+\vspace{-15pt}
 \end{figure}
@@
 \begin{inparaenum}
-\item \emph{shrinking the value}, replacing its top-level data constructor with
+\item \emph{shrinking the value}, replacing its outtermost data constructor with
   one that contains a subset of its fields, reusing existing subexpressions;
-\item \emph{growing the value}, replacing its top-level data constructor with
+\item \emph{growing the value}, replacing its outtermost data constructor with
   one that contains a superset of its fields, reusing existing subexpressions
@@
 \texttt{Tree} happens with a probability smaller than $(1/6)^3 = \sim 0.0046$,
-and this only worsens as the type of the mutated value becomes more complex.
+i.e., when applying a recursive mutation (with probability $1/6$) three
+consecutive times.
+%
+This only worsens as the type of the mutated value becomes more complex.
 %
@@
 data structures, potentially excluding interesting applications that might
-require producing deeper valid values, e.g., programming languages, network
-protocols interactions, etc.
+require producing deeper valid values, e.g., programs, network protocols
+interactions, etc.
 %
@@
 %
-On top of them, Section \ref{sec:heuristics} introduces heuristics that help
-finding bugs faster in certain testing scenarios.
+Despite the choice of implementation language, we remark that the ideas
+presented in this work extend beyond Haskell and functional programming
+languages in general.
+%
+%% Later, Section \ref{sec:heuristics} introduces heuristics that help finding bugs
+%% faster in certain testing scenarios.
 
@@
     deterministic mutations, and
-  \item for every mutable subexpression, each of these mutations is evaluated
-    \emph{exactly once}.
+  \item for every mutable subexpression, each one of these mutations is
+    evaluated \emph{exactly once}.
 \end{inparaenum}
 %
-There is a small exception to this rules that we will introduce soon.
+%% There is a small exception to this rules that we will introduce soon.
 
@@
 %
-In consequence, such tools can only be applied to relatively simple systems,
-where the space of inputs does not grow extremely fast.
+To alleviate this problem, such tools often rely on heuristics to prune the
+search space --- this is discussed in Section \ref{sec:related}.
+
+%% In consequence, such tools can only be applied to relatively simple systems,
+%% where the space of inputs does not grow extremely fast.
 
@@
 %
-Moreover, these mutants will be mutated further only if they discover new paths
-in the target code, so the testing loop automatically prunes the space of test
-cases that are worth mutating.
+Moreover, like in \fuzzchick, the testing loop of our tool automatically filters
+the test cases that are worth mutating by keeping only those that discover new
+execution paths in the code under test.
 
@@
 Pure mutants are used most of the time, and encode simple deterministic
-transformations over the top-level data constructor of the input --- recursive
+transformations over the outtermost data constructor of the input --- recursive
 mutations will be introduced soon.
@@
 \item return an immediate subexpression of the same type as the input;
-\item swap the top-level data constructor with any other constructor of the same
-  type, reusing existing subexpressions; and
+\item swap the outtermost data constructor with any other constructor of the
+  same type, reusing existing subexpressions; and
 \item rearrange and replace fields using existing ones of the same type.
@@
 Notice how this definition simply enumerates mutants that transform the
-top-level data constructor, hence no recursion is needed here.
+outtermost data constructor, hence no recursion is needed here.
 %
@@
 In first place, a function \texttt{positions} traverses the mutable candidate
-and builds a multi-way tree of mutable positions.
+and builds a rose tree of mutable positions.
 %
@@
 
-\noindent In this light, the mutable positions of the value \texttt{Branch (Leaf 1) 2 (Leaf 3)} are:
+\noindent where the function \texttt{node} simply builds a node of the positions
+rose tree and propagates the accummulated prefix (the path from the root) to its
+children.
+%
+In this light, the mutable positions of the value \texttt{Branch (Leaf 1) 2
+  (Leaf 3)} are:
 %
@@
   \Fn{\Muts{x, mut, R}}{
-    muts $\gets$ $\varnothing$\;
+    batch $\gets$ $\varnothing$\;
     \For{pos \KwIn Flatten(Positions(x))}{
@@
           \Case{\texttt{PURE} $\hat{x}$}{
-            Enqueue($\hat{x}$, muts)\;
+            Enqueue($\hat{x}$, batch)\;
           }
@@
               $\hat{x}$ $\gets$ Sample(gen)\;
-              Enqueue($\hat{x}$, muts)\;
-            }
-          }
-        }
-      }
-    }
-    \KwRet muts\;
+              Enqueue($\hat{x}$, batch)\;
+            }
+          }
+        }
+      }
+    }
+    \KwRet batch\;
   }
@@
 we first flatten all the mutable positions of the input value in level order
-(recall that positions are stored as a multi-way tree).
+(recall that positions are stored as a rose tree).
 %
@@
         \If{\KwNot Discarded(result)}{
-          muts $\gets$ Mutate(x, mut, R)\;
-          Enqueue(QSucc, muts)\;
+          batch $\gets$ Mutate(x, mut, R)\;
+          Enqueue(QSucc, batch)\;
         }
         \uElseIf{Passed(Parent(x))}{
-          muts $\gets$ Mutate(x, mut, R)\;
-          Enqueue(QDisc, muts)\;
+          batch $\gets$ Mutate(x, mut, R)\;
+          Enqueue(QDisc, batch)\;
         }
@@
     \If{\KwNot Empty(QSucc)}{
-      muts $\gets$ Deque(QSucc)\;
-      \If{Empty(muts)}{
+      batch $\gets$ Deque(QSucc)\;
+      \If{Empty(batch)}{
         Pick(QSucc, QDisc, gen)\;
       } \uElse{
-        PushFront(QSucc, Rest(muts))\;
-        \KwRet First(muts)\;
+        PushFront(QSucc, Rest(batch))\;
+        \KwRet First(batch)\;
       }
@@
 
-This heuristic tackles the issue of enqueuing novel mutation candidates at the
-end of (possibly) long queues of not-so-interesting previously executed ones.
+This heuristic tackles the issue of enqueuing new interesting mutation
+candidates at the end of (possibly) long queues of not-so-interesting previously
+executed ones.
 
@@
 %
-In contrast, execution traces \mutagen represent the \emph{path} in the code
-taken by the program, as opposed to just the set of edges traversed in the
+In contrast, execution traces \mutagen represent the specific \emph{path} in the
+code taken by the program, as opposed to just the edges traversed in the
 control-flow graph (CFG).
@@
       \If{\KwNot Discarded(result)}{
-        muts $\gets$ Mutate(x, mut, R)\;
+        batch $\gets$ Mutate(x, mut, R)\;
         {\color{red} prio $\gets$ BranchDepth(TLog, trace)}\; %% NEW
-        {\color{red} PushFront(QSucc, prio, muts)}\;          %% NEW
+        {\color{red} PushFront(QSucc, prio, batch)}\;          %% NEW
       }
@@
     \If{\KwNot Empty(QSucc)}{
-      {\color{red} (muts, prio) $\gets$ DequeMin(QSucc)}\; %%NEW
-      \If{Empty(muts)}{
+      {\color{red} (batch, prio) $\gets$ DequeMin(QSucc)}\; %%NEW
+      \If{Empty(batch)}{
         Pick(QSucc, QDisc, gen)\;
       } \uElse{
-        {\color{red} PushFront(QSucc, prio, Rest(muts))}\; %% NEW
-        \KwRet First(muts)\;
+        {\color{red} PushFront(QSucc, prio, Rest(batch))}\; %% NEW
+        \KwRet First(batch)\;
       }
@@
 
-\subsection{Detecting Trace Space Saturation And Tuning Random Mutations}
+\subsection{Detecting Trace Space Saturation And Tuning Random Mutations Parameter}
 
@@
 The first one is a simple abstract stack machine that enforces
-\emph{noninterference} using runtime checks.
-%
-The implementation of this case study was originally proven correct by \cite{}
-in Coq, and subsequently degraded by systematically introducing 20 bugs on its
+\emph{noninterference} \cite{goguen1982security} using runtime checks.
+%
+The implementation of this case study was originally proven correct by
+\citeauthor{10.1145/2578855.2535839} \citeyearpar{10.1145/2578855.2535839} in
+Coq, and subsequently degraded by systematically introducing 20 bugs on its
 enforcing mechanism.
@@
 The second case study aims to evaluate \mutagen in a more realistic scenario,
-and focuses on testing \textit{haskell-wasm}, an existing WebAssembly engine
-written in Haskell of industrial strength.
+and focuses on testing \textit{haskell-wasm} \cite{haskellwasm}, an existing
+WebAssembly engine written in Haskell of industrial strength.
 %
@@
 interpreter of this engine, and used the reference implementation to find them
-via differential testing.
+via differential testing \cite{mckeeman1998differential}.
 %
@@
 \end{tabular}
-\caption{\label{table:wasm:injectedbugs}Bugs injected in \textit{haskell-wasm}.}
+\caption{\label{table:wasm:injectedbugs}Bugs injected into
+  \textit{haskell-wasm}.}
+\vspace{-20pt}
 \end{table}
@@
 To preserve space, we encourage the reader to refer to the work of
-\citeauthor{catalin} as well as the original \fuzzchick paper for more details
+\citeauthor{hritcu2013testing} \citeyearpar{hritcu2013testing,
+  hrictcu2016testing} as well as the original \fuzzchick paper for more details
 about the implementation and semantics of this case study.
@@
 Using this notion, the particular variant of noninterference we are interested
-in this work is called \emph{single-step noninterference}.
+in this work is called \emph{single-step noninterference}
+\cite{hritcu2013testing}.
 %
@@
 
-WebAssembly is a popular assembly-like language designed to be an open standard
-for executing low-level code in the web, although it has become increasingly
-popular in standalone, non-web contexts as well.
+WebAssembly \cite{haas2017bringing} is a popular assembly-like language designed
+to be an open standard for executing low-level code in the web, although it has
+become increasingly popular in standalone, non-web contexts as well.
 %
@@
 non-deterministic undefined behavior allowed by the specification (NaN
-reinterpretation).
+reinterpretation) \cite{perenyi2020stack}.
 %
@@
 prop_validator m =
-  isValidHaskell m ==> isValidSpec m
+  isValidHaskellWasm m ==> isValidRefImpl m
 \end{verbatim}
 
-Where the precondition (\texttt{isValidHaskell m}) runs the input WebAssembly
+Where the precondition (\texttt{isValidHaskellWasm m}) runs the input WebAssembly
 module \texttt{m} against the \textit{haskell-wasm} validator, whereas the
-postcondition (\texttt{isValidSpec m}) serializes \texttt{m} to a file, runs it
+postcondition (\texttt{isValidRefImpl m}) serializes \texttt{m} to a file, runs it
 against the reference implementation validator and checks that no errors are
@@
 %% \begin{verbatim}
-%% isValidHaskell m =
+%% isValidHaskellWasm m =
 %%   case validate m of
@@
 %% \begin{verbatim}
-%% isValidSpec m = do
+%% isValidRefImpl m = do
 %%   writeFile "testcase.wasm" (dumpModule m)
@@
 All the experiments were performed in a dedicated workstation with an Intel Core
-i7-8700 CPU running at 3.20GHz, and equipped with 32Gb of RAM.
+i7-8700 CPU running at 3.20GHz, and equipped with 32GB of RAM.
 %
@@
 \citeauthor{lampropoulos2019coverage} and collected the Mean-Time-To-Failure
-(MTTF) of each bug, i.e., how quickly a bug can be found in wall time.
+(MTTF) of each bug, i.e., how quickly a bug can be found in wall clock time.
 %
@@
 
-\begin{figure}[b]
+\begin{figure}[t]
   \centering
@@
 
+If we only consider the succesfull runs where \fuzzchick does not time out, we
+observed a peculiarity.
+%
+This tool either finds bugs relatively quickly (after a few thouthands tests) or
+does not find them at all within the time budget, which suggest that its power
+scheduler assigns some mutation candidates too little energy before they become
+uniteresting and get ultimately thrown away.
+%
+Under this consideration, it is fair to think that \fuzzchick might be a better
+choice if we set a shorter timeout and run it several times.
+%
+Thus, to be an improvement over \fuzzchick, our tool must be able to find bugs
+not only more reliably, but also relatively fast!
+
+Table \ref{table:ifc:tests} shows a comparison between the mean number of tests
+required by both tools to find the first failure when we only consider
+successful runs.
+%
+Additionally, as adviced by \cite{arcuri2014hitchhiker} when comparing random
+testing tools, we computed the non-parametric Vargha-Delaney $A_{12}$ measure
+\cite{vargha2000critique}.
+%
+Intuitively, this measure encodes the probability of \mutagen to yield better
+results than \fuzzchick.
+
 \begin{wrapfigure}{r}{0.5\textwidth}
 \centering
-\vspace{-20pt}
+%% \vspace{-1}
 \footnotesize
@@
   favouring \fuzzchick and \mutagen, respectively. }
-\vspace{-25pt}
+\vspace{-20pt}
 \end{wrapfigure}
-
-If we only consider the succesfull runs where \fuzzchick does not time out, we
-observed a peculiarity.
-%
-This tool either finds bugs relatively quickly (after a few thouthands tests) or
-does not find them at all within the time budget, which suggest that its power
-scheduler assigns some mutation candidates too little energy before they become
-uniteresting and get ultimately discarded.
-%
-Under this consideration, it is fair to think that \fuzzchick might be a better
-choice if we set a shorter timeout and run it several times.
-%
-Thus, to be an improvement over \fuzzchick, our tool must be able to find bugs
-not only more reliably, but also relatively fast!
-
-Table \ref{table:ifc:tests} shows a comparison between the mean number of tests
-required by both tools to find the first failure when we only consider
-successful runs.
-%
-Additionally, as adviced by \cite{arcuri} when comparing random testing tools,
-we computed the non-parametric Vargha-Delaney $A_{12}$ measure \cite{vda}.
-%
-Intuitively, this measure encodes the probability of \mutagen to yield better
-results than \fuzzchick.
 %
 As it can be observed, using an exhaustive mutation approach can be not only
-more reliable, but also faster, where \mutagen is likely to find bugs faster
-that \fuzzchick in 14 of the 20 bugs injected into the abstract machine.
+more reliable, but also faster.
+%
+\mutagen is likely to find bugs faster than \fuzzchick in 14 of the 20 bugs
+injected into the abstract machine, while being substantially slower in only 2
+cases.
 
@@
 components independently is virtually unable to produce valid WebAssembly
-modules.
+modules apart from the trivial empty one.
 %
@@
 %
-Nonetheless, \mutagen still outperforms \quickcheck both in terms of failure
-rate and mean time to failure.
+Notably, \quickcheck find counterexamples for the bug \%14 almost instantly.
+%
+The reason behind this is that this bug can be found using a rather small
+counterexample, and \quickcheck prefers sampling small test cases at the
+beginning of the testing loop.
+%
+Our tool follows this same approach when in generation mode.
+%
+However, when a test case is found interesting, the scheduler does not take its
+size into account while computing its priority --- future work should
+inverstigate this posibility.
+%
+Nonetheless, \mutagen still outperforms \quickcheck on the remaining bugs both
+in terms of failure rate and mean time to failure.
 %
@@
 Finding where the best tradeoff lies in cases where the heuristics cause mixed
-results like this is an interesting direction to aim future work.
+results like this is an interesting direction to aim our future work.
 
@@
 Table \ref{table:wasm:overhead} compares the total number of executed and passed
-test per second that we obtained using each tool.
+tests per second that we observed using each tool.
 %
 It is easy to conclude that \mutagen executes tests several times slower than
-\quickcheck -- roughly 16x and 70x slower when testing \textit{prop\_validator}
+\quickcheck --- roughly 16x and 70x slower when testing \textit{prop\_validator}
 and \textit{prop\_interpreter\_i32}, respectively.
@@
 \textit{prop\_validator}
-& 42136
-& 0.00028
-& 2583
-& 539.19 \\
+& 41374.92
+& 0.00029
+& 2243.26
+& 519.83\\
 \hline
@@
 & 23.16
-& 1890
-& 372.75 \\
+& 1815.81
+& 351.16 \\
 \hline
@@
 % DRAGEN and DRAGEN2
-In Haskell, DRAGEN \cite{dragen} is a meta-programming tool that synthezises
-random generators from datatypes definitions, using stochastic models to predict
-and optimize their distribution based on a target one set by the user.
-%
-DRAGEN2 \cite{dragen2} extends this idea adding support for generating richer
-random data by extracting additional static information like library APIs and
-function input patterns from the codebase.
+In Haskell, DRAGEN \cite{DBLP:conf/haskell/MistaRH18} is a meta-programming tool
+that synthezises random generators from datatypes definitions, using stochastic
+models to predict and optimize their distribution based on a target one set by
+the user.
+%
+DRAGEN2 \cite{Mista2019GeneratingRS} extends this idea adding support for
+generating richer random data by extracting additional static information like
+library APIs and function input patterns from the codebase.
 %
 % Bolztmann-brain
-Similarly, \cite{bendkowski} developed a polinomial tuning mechanism based on
-Boltzmann samplers \cite{boltzmann} that synthesizes generators with
+Similarly, \cite{Bendkowski2017} developed a polinomial tuning mechanism based
+on Boltzmann samplers \cite{Duchon2004} that synthesizes generators with
 approximate-size distributions for combinatorial structures like lists or trees.
@@
 %
-Exploiting some of the ideas described above, \emph{QuickFuzz} \cite{quickfuzz,
-  quickfuzzjournal} is a generational grammar-based fuzzer that leverages
-existing Haskell libraries describing common file formats to synthesize random
-data generators that are used in tandem with off-the-shelf low-level fuzzers to
-find bugs in massively used programs.
+Exploiting some of the ideas described above, \emph{QuickFuzz} \cite{GriecoCB16,
+  grieco2017} is a generational fuzzer that leverages existing Haskell libraries
+describing common file formats to synthesize random data generators that are
+used in tandem with existing off-the-shelf low-level fuzzers to find bugs in
+massively used programs.
 
@@
 %
-\cite{claessen2014} developed an algorithm for generating inputs constrained by
+\cite{ClaessenDP14} developed an algorithm for generating inputs constrained by
 boolean preconditions with almost-uniform distribution.
@@
 %
-Later, \cite{luck} extends this approach by adding a limited form of constraint
-solving controllable by the user in a domain-specific language called
-\emph{Luck}.
+Later, \cite{LampropoulosGHH17} extends this approach by adding a limited form
+of constraint solving controllable by the user in a domain-specific language
+called \emph{Luck}.
 %
@@
 %
-Recently, \cite{generatinggood} proposed a derivation mechanism to obtain
-constrained random generators directly by defining their structure
-using inductively defined relations in Coq.
+Recently, \cite{Lampropoulos2017} proposed a derivation mechanism to obtain
+constrained random generators directly by defining their structure using
+inductively defined relations in Coq.
 
@@
 % AFLfast
-\emph{AFLFast} is an extension to AFL that uses Markov chain models to tune the
-power scheduler towards testing low-frequency paths.
+\emph{AFLFast} \cite{bohme2017coverage} is an extension to AFL that uses Markov
+chain models to tune the power scheduler towards testing low-frequency paths.
 %
@@
 % CollAFL
-Similar to our tool, \emph{CollAFL} \cite{collAFL} is a variant of AFL that uses
-path- instead of edge-based coverage, which helps distinguishing executions more
-precisely by reducing path collisions.
+Similar to our tool, \emph{CollAFL} \cite{gan2018collafl} is a variant of AFL
+that uses path- instead of edge-based coverage, which helps distinguishing
+executions more precisely by reducing path collisions.
 
 %AFLGo
-When the source code history is available, \emph{AFLGo} \cite{aflgo} is a fuzzer
-that can be targeted to exercise the specific parts of the the code affected by
-recent commits in order to find potential new bugs.
+When the source code history is available, \emph{AFLGo} \cite{bohme2017directed}
+is a fuzzer that can be targeted to exercise the specific parts of the the code
+affected by recent commits in order to find potential new bugs.
 %
@@
 is to use fitness functions to guide the testing efforts towards user defined
-goals.
-%
-\emph{Target} \cite{targetedpbt} is a tool that implements TPBT using
-optimization techniques like hill climbing and simulated annealing.
+goals \cite{loscher2017targeted, loscher2018automating}.
 
@@
 Moreover, a popular technique is to combine coverage-guided fuzzing with ideas
-borrowed from symbolic execution tools like \emph{KLEE} \cite{klee} to avoid
-getting stuck in superficial paths \cite{driller, qsym}.
+borrowed from symbolic execution tools like \emph{KLEE} \cite{cadar2008klee} to
+avoid getting stuck in superficial paths \cite{stephens2016driller}.
 %
 When using off-the-shelf symbolic executors, this technique is often limited by
-the path explosion problem, altough recent tools like \emph{QSYM} \cite{qsym}
-demonstrate that using tailored concolic executors can help overcoming this
-limitation.
+the path explosion problem, altough recent tools like \emph{QSYM}
+\cite{yun2018qsym} demonstrate that using tailored concolic executors can help
+overcoming this limitation.
 
@@
 %
-\emph{Feat} \cite{feat} formalizes the notion of \emph{functional enumerations}.
+\emph{Feat} \cite{DuregardJW12} formalizes the notion of \emph{functional
+  enumerations}.
 %
@@
 % SmallCheck and LazySmallCheck
-\emph{SmallCheck} \cite{smallcheck} is a Haskell tool that also follows this
-idea.
+\emph{SmallCheck} \cite{runciman2008smallcheck} is a Haskell tool that also
+follows this idea.
 %
@@
 % Korat
-On the object-oriented realm, \emph{Korat} \cite{korat} is a Java tool that uses
-method specification predicates to automatically generate all (nonisomorphic)
-test cases up to a given small size.
+On the object-oriented realm, \emph{Korat} \cite{boyapati2002korat} is a Java
+tool that uses method specification predicates to automatically generate all
+(nonisomorphic) test cases up to a given small size.
 
@@
 
-As already indicated throughout this work, there are several direction for
+As already indicated throughout this work, there are several directions for
 future work we aim to investigate:

### 2021-04-08 papers/oopsla-21/main.tex
Added references, general pass

@@
 %
-Random Property-Based Testing (RPBT) is a popular technique for finding bugs
-using executable properties encoding the expected behavior of our system.
-%
-Exemplified by frameworks Haskell's \quickcheck \cite{ClaessenH00} and the
+Property-Based Testing (PBT) is a popular technique for finding bugs using
+executable properties that encode the expected behavior of our systems.
+%
+Exemplified by frameworks like Haskell's \quickcheck \cite{ClaessenH00} and the
 abundant ports of this tool in other programming languages
 \cite{hughes2003erlang, papadakis2011proper, bulwahn2012new,
-  denes2014quickchick}, the most common approach is to instantiate the testing
-properties using randomly generated inputs.
-
-Users of \quickcheck (and RPBT in general) are well aware that one of the
-biggest challenges while using these frameworks is to provide suitable random
-data generators that are used to instantiate the testing properties.
+  denes2014quickchick}, the most common PBT approach is to instantiate the
+testing properties using randomly generated inputs.
+
+Users of \quickcheck (and PBT in general) are well aware that one of the biggest
+challenges while using these frameworks is to provide suitable random data
+generators needed to instantiate the testing properties.
 %
@@
 exercising every part of a complex system in a reasonable basis can take several
-thouthands person-hours of trial and error \cite{lampropoulos2019coverage}.
+thouthand person-hours of trial and error \cite{lampropoulos2019coverage}.
 
@@
 Randomly generating valid programs is common example of this problem, where
-developers are forced to put a substantial effort writing complex generators by
-hand \cite{Palka11, perenyi2020stack, yang2011finding}.
+developers are forced to put a substantial efforts in writing specialized random
+generators by hand \cite{Palka11, perenyi2020stack, yang2011finding}.
 %
@@
 %
-\fuzzchick is likely to preserve the semantic structure of the mutated data, as
-mutations are applied directly at the data type level -- the random AST in case
-of generating code.
+In this light, \fuzzchick is likely to preserve the semantic structure of the
+mutated data, as mutations are applied directly at the data type level, e.g.,
+the program's AST in case of generating code.
 
 This tool relies on execution traces to distinguish which test cases resulted
-interesting and are therefore worth mutating --- a prolific technique known
-\emph{coverage-guided fuzzing}, popularized by tools like \emph{AFL} \cite{afl}.
+interesting and are therefore worth mutating --- a prolific technique known as
+\emph{coverage-guided fuzzing} and popularized by tools like \emph{AFL}
+\cite{afl}.
 %
 Mutated test cases are considered interesting for mutation only when they
-produce new execution traces -- any other test case is simply thrown away.
+produce new execution traces --- any other test case is simply thrown away.
 
@@
 Notably, we establish several aspects of its implementation that can be revised
---- some of these already being identified by the authors of \fuzzchick on their
+--- some of these already being identified by the authors of \fuzzchick in their
 original work.
@@
   are enqueued frequently}, and
-\item \emph{using a power schedule to assing a certain mutation energy to each
+\item \emph{the mechanism used to assing a certain mutation energy to each
   interesting test case requires fine tuning and can be hard to generalize to
@@
 However, we believe that a proper evaluation ought to use the failure rate as an
-important metric when comparing RPBT tools.
+important metric when comparing PBT tools.
 
 %% Our solution
-In this work we introduce \mutagen, a RPBT tool implemented in Haskell that
-build upon the basic idea behind \fuzzchick, i.e., code instrumentation and
+In this work we introduce \mutagen, a PBT tool implemented in Haskell that
+builds upon the basic ideas behind \fuzzchick, i.e., code instrumentation and
 type-preserving mutations.
@@
 This is, given an interesting, previously executed test case, our tool
-precomputes and schedules every possible type-presserving mutation that can be
-applied to it.
+precomputes and schedules every type-presserving mutation that can be applied to
+it.
 %
@@
 On the other hand, scheduling mutations exaustively eliminates the need of a
-heuristic power schedule to a given mutation energy to each input test
-case.
+heuristic that assigns a given mutation energy to each input test case.
 
@@
 Our tool distinguishes two kinds of mutations, those that can be computed
-purely, yielding a single mutated value; and those that can be obtained
-non-deterministically.
-%
-Pure mutations encode transformations that can be done by swapping data
-constructor variants or rearranging subexpressions, whereas non-deterministic
-(random) mutations are useful to represent transformations over large
-enumeration types, e.g., numbers, characters, etc.
-%
-In this setting, random mutations let us selectively escape the exhaustiveness
-of our tool by randomly sampling a small random generator a reduced number of
-times.
+purely, yielding a single deterministic mutated value; and those that can be
+obtained non-deterministically.
+%
+On one hand, pure mutations encode transformations that swap data constructors
+around, as well as return or rearrange subexpressions.
+%
+Non-deterministic (random) mutations, on the other hand, are useful to represent
+transformations over large enumeration types, e.g., numbers, characters, etc.
+%
+This mechanism let us selectively escape the exhaustiveness of our tool by
+randomly sampling a small random generator a reduced number of times.
 %
@@
 %
-In this light, test cases that discover new parts of the code earlier during
-execution are given a higher priority while picking the next mutated test case.
+In this light, interesting test cases that discover new parts of the code
+earlier during execution are given a higher priority.
 %
 Additionally, our scheduling lets the testing loop jump back and forth between
-scheduled test cases as soon as a new more interesting one is enqueued for mutation,
-eliminating potential delays when the mutation queues grow large.
+mutable test cases as soon as a new more interesting one is enqueued for
+mutation, eliminating potential delays when the mutation queues grow large.
 
@@
 %
-This automatically finds a suitable value for this external parametere on the
+This automatically finds a suitable value for this external parameter on the
 fly.
@@
 %
-There, \mutagen is capable of reliaby finding 15 planted bugs in the validator
+There, \mutagen is capable of reliably finding 15 planted bugs in the validator
 and interpreter, as well as 3 previously unknown bugs that flew under the radar
@@
 %
-Originally introduced by \citet{quickcheck} in tandem with the first version of
-\quickcheck, this technique focuses on aiming the developer's efforts into
-testing systems via executable properties using randomly generated inputs.
-%
-Moreover, tools like \quickchick and Isabelle's \quickcheck demonstrate
-that PBT can also be used in the formal verification realm.
-%
-There, one can quickly spot bugs in system specifications before directing the
-efforts into pointlessly trying to prove bogus propositions.
+Popularized by Haskell's \quickcheck, this technique focuses on aiming the
+developer's efforts into testing systems via executable properties using
+randomly generated inputs.
+%
+%% Moreover, tools like \quickchick and Isabelle's \quickcheck demonstrate
+%% that PBT can also be used in the formal verification realm.
+%% %
+%% There, one can quickly spot bugs in system specifications before directing the
+%% efforts into pointlessly trying to prove bogus propositions.
 
@@
 %
-Printers and shrinkers, for the most part, can be obtained automatically using
-generic programming capabilities present in the compiler, and although being
-crucial for the testing process as a whole, their role becomes irrelevant when
-it comes to \emph{finding} bugs.
+Printers and shrinkers, for the most part, can often be obtained automatically
+using generic programming capabilities present in the compiler, and although
+being crucial for the testing process as a whole, their role becomes irrelevant
+when it comes to \emph{finding} bugs.
 
@@
 \noindent The type parameter \texttt{a} indicates that trees can be instantiated
-using any other type as payload, so the value \texttt{Leaf Bool} has type
+using any other type as payload, so the value \texttt{Leaf True} has type
 \texttt{Tree Bool}, whereas the value \texttt{Branch (Leaf 1) 2 (Leaf 3)} has
@@
 
-With our simple specification in place, the last missing piece is a random
-generator of trees.
-%
-In \quickcheck, this is usually done via the type class mechanism, instantiating
-the \texttt{Arbitrary} type class, providing a random generator as the
-implementation of the overloaded \texttt{arbitrary} operation:
+The last missing piece is a random generator of trees.
+%
+In \quickcheck, this is commonly done via the type class mechanism
+\cite{jones1997type}, instantiating the \texttt{Arbitrary} type class with a
+random generator as the implementation of the overloaded \texttt{arbitrary}
+operation:
 
@@
 To alleviate the problem of testing properties with sparse preconditions while
-using simple automatically derived random generators, \fuzzchick introduces
-\emph{coverage-guided property-based testing} (CGPT) by enhancing the testing
-process with two key characteristics:
+using automatically derived random generators, \fuzzchick introduces
+\emph{coverage-guided, property-based testing} (CGPT), a technique that enhances
+the testing process with two key characteristics:
 %
@@
 \item \emph{high-level, type-preserving mutations}, to produce syntactically
-  valid test cases by altering existing ones at the datatype level.
+  valid test cases by altering existing ones at the data type level.
 \end{inparaenum}
@@
 %
-Tools like Criterion, XSmith and LangFuzz use existing grammars to tailor the
-generic mutators to the specific input structure used by the system under test.
+Tools like LangFuzz \cite{holler2012fuzzing}, Superion \cite{wang2019superion},
+XSmith \cite{xsmith} use existing grammars to tailor the mutators to the
+specific input structure used by the system under test.
 %
@@
 %
-The datatypes used by the inputs of the testing properties already describe the
+The data types used by the inputs of the testing properties already describe the
 structure of the random data we want to mutate in a concrete manner, and
@@
 They use a simple set of mutation operations that are randomly applied at the
-datatype level.
+data type level.
 %
@@
 
-Fig. \ref{fig:fuzzchick:mutator} illustrates a \fuzzchick mutator for our
+Fig. \ref{fig:fuzzchick:mutator} illustrates a \fuzzchick-like mutator for our
 previously used \texttt{Tree} data type example.
@@
 %
-For instance, mutating a subexpression that lies on the third level of a
+For instance, mutating a subexpression that lies on the fourth level of a
 \texttt{Tree} happens with a probability smaller than $(1/6)^3 = \sim 0.0046$,
@@
 %
-If that mutated test case to be interesting, that subexpression can always be
-mutated later.
+If that mutated test case turns to be interesting, that subexpression can always
+be mutated later.
 
@@
 %
-We show in Section \ref{sec:heuristics} how this can be achieved by analyzing
-the execution information in order to prioritize test cases with novel execution
-traces.
+%% We show in Section \ref{sec:heuristics} how this can be achieved by analyzing
+%% the execution information in order to prioritize test cases with novel execution
+%% traces.
 
@@
 %
-Despite the choice of implementation language, we remark that the ideas
-presented in this work extend beyond Haskell and functional programming
-languages in general.
+Despite the choice of implementation language, the ideas presented in this work
+extend beyond Haskell and functional programming languages in general.
 %
@@
   \item each subexpression of a mutation candidate is associated with a set of
-    deterministic mutations, and
+    type-preserving mutations, and
   \item for every mutable subexpression, each one of these mutations is
@@
 
-This idea is inspired by exhaustive bounded testing tools like SmallCheck (in
-Haskell) or Korat (in Java), that produce test cases exhaustively.
+This idea is inspired by exhaustive bounded testing tools like SmallCheck
+\cite{runciman2008smallcheck} or Korat \cite{boyapati2002korat}, that produce
+test cases exhaustively.
 %
 In simple words, such tools work by enumerating all possible values of the
-datatypes used in the testing properties, and then executing them from smaller
+data types used in the testing properties, and then executing them from smaller
 to larger until a a bug is found, or a certain size bound is reached.
@@
 experiences what it looks like ``hitting a wall'', where no larger test cases
-can be evaluated until we exhausted all the immediately smaller ones \cite{feat}.
+can be evaluated until we exhausted all the immediately smaller ones
+\cite{DuregardJW12}.
 %
 To alleviate this problem, such tools often rely on heuristics to prune the
-search space --- this is discussed in Section \ref{sec:related}.
+search space based on detecting unevaluated subexpressions --- this is discussed
+in Section \ref{sec:related}.
 
@@
 %
-\mutagen uses random generators to find interesting initial seeds, and only then
-proceeds to execute all possible mutations.
+Our tool uses random generators to find interesting initial seeds, and when it
+finds them, only then proceeds to schedule all possible mutations.
 %
@@
 In contrast to \fuzzchick, where mutators are parameterized random generators,
-in \mutagen we define mutations as the set of mutants that can be obtained by
+in our tool we define mutations as the set of mutants that can be obtained by
 altering the input value at the top-level (the root).
@@
 %
-These tranformation can:
+These tranformations can:
 %
@@
 %
-This value corresponds to the smallest value we can construct in order for the
+This value corresponds to the simplest value we can construct in order for the
 mutant to be type-correct.
@@
 In practice (\texttt{def = Leaf def}), where the inner \texttt{def} is the
-smallest value of the payload --- we use the type class system to abstract this
+simplest value of the payload --- we use the type class system to abstract this
 complexity away in our implementation.
 %
-Using a default smallest value is again inspired by exhaustive bounded testing
-tools, and avoids introducing unnecesary randomness when growing data
-constructors.
+Using a small value is again inspired by exhaustive bounded testing tools, and
+avoids introducing unnecesary randomness when growing data constructors.
 
@@
 
-At this point, the reader might consider:
-%
-what about large enumeration types like integers or characters?
-%
-Does \mutagen transform them exaustively too?
-%
-\emph{Certainly not.}
-%
-Random mutants serve as a way to break exhaustiveness when mutating values of
-large enumeration types.
-%
-Instead of trying every possible integer or character, we resolve in using a
-random generator and sample a small amount of values from it --- the precise
-amount is a tunable parameter of \mutagen.
+%% At this point, the reader might consider:
+%% %
+%% what about large enumeration types like integers or characters?
+%% %
+%% Does \mutagen transform them exaustively too?
+%% %
+%% \emph{Certainly not.}
+%
+Random mutants, on the other hand, serve as a way to break exhaustiveness when
+mutating values of large enumeration types --- or any other type the user might
+want to introduce random mutations to.
+%
+Instead of mutating into every other possible number or character, we resolve in
+using a random generator and sample a small amount of values from it --- the
+precise amount is a tunable parameter of \mutagen.
 %
@@
 
-So far we defined mutations that transform the root of the input.
+So far we have defined mutations that transform only the root of the input.
 %
@@
 %
-To do so, we will use two functions that can also be derived from the data type
+To do so, we will use two functions that can be derived from the data type
 definition.
 
-In first place, a function \texttt{positions} traverses the mutable candidate
-and builds a rose tree of mutable positions.
+In first place, a function \texttt{positions} traverses the input and builds a
+rose tree of mutable positions.
 %
@@
 
-\noindent Later, given a desired position to mutate within a candidate, we
-define another function \texttt{inside} that precisely finds the subexpression
+\noindent Later, given a desired position to mutate within a test case, we
+define another function \texttt{inside} that finds the subexpression
 corresponding to it and applies the mutation.
 %
-A slightly simplified version of this function for the \texttt{Tree} datatype is
+A slightly simplified version of this function for the \texttt{Tree} data type is
 as follows:
@@
 %
-As it can be observed, we closely follow \fuzzchick's testing loop, using qtwo
-queues to keep mutant candidates, and enqueueing and retrieving interesting test
-cases from them before falling back to random generation.
+As it can be observed, we closely follow \fuzzchick's testing loop, using two
+queues to keep mutable candidates, and enqueueing and retrieving interesting
+test cases from them before falling back to random generation.
 
@@
 These mutations are put together into lists that we call \emph{mutation batches}
---- one for each mutation candidate.
+--- one for each mutated test case.
 %
@@
   into the mutation batch directly; otherwise
-\item it is a random mutant that carries a random generator with it (e.g., a
-  numeric subexpression), in which case we sample and enqueue \textit{R} random
-  values using this generator, where \textit{R} is a parameter set by the user.
+\item it is a random mutant that carries a random generator with it (e.g.,
+  corresponding to a numeric subexpression), in which case we sample and enqueue
+  \textit{R} random values using this generator, where \textit{R} is a parameter
+  set by the user.
 \end{inparaenum}
@@
 We found that, especially for large data types, the queue of discarded
-candidates tends to grow disproportionately during testing, making them hardly
-usable and consuming large amounts of memory.
+candidates tends to grow disproportionately large during testing, making them
+hardly usable and consuming large amounts of memory.
 %
@@
 %
-Moreover, this can potentially help introducing \emph{2-step mutations}, where
-an initial mutation breaks a valid test case in a small way, it gets enqueued as
-discarded, and later a subsequent mutation fixes it by changing a different
-subexpression.
+Moreover, this can potentially lead to \emph{2-step mutations}, where an initial
+mutation breaks a valid test case in a small way, it gets enqueued as discarded,
+and later a subsequent mutation fixes it by changing a different subexpression.
 
@@
 \mutagen enables them all by default, although they can be individually disabled
-by the user if deemed appropriate.
+by the user if deems appropriate.
 
@@
 In contrast, execution traces \mutagen represent the specific \emph{path} in the
-code taken by the program, as opposed to just the edges traversed in the
-control-flow graph (CFG).
+code taken by the program, as opposed to just the (unordered) set of edges
+traversed in the control-flow graph (CFG).
 %
@@
 %
-For instance, when the system under test branches on numeric values, we should
-make sure that we set enough random mutations to test each case in a reasonable
-basis.
+For instance, when the system under test branches based on numeric values, we
+should make sure that we set enough random mutations to test each branch in a
+reasonable basis.
 %
@@
 and focuses on testing \textit{haskell-wasm} \cite{haskellwasm}, an existing
-WebAssembly engine written in Haskell of industrial strength.
+WebAssembly engine of industrial strength written in Haskell.
 %
 We manually injected 10 bugs in the validator as well as 5 bugs in the
-interpreter of this engine, and used the reference implementation to find them
-via differential testing \cite{mckeeman1998differential}.
+interpreter of this engine, and used the reference WebAssembly implementation to
+find them via differential testing \cite{mckeeman1998differential}.
 %
@@
 Labels form a trivial 2-lattice where information can either stay at the same
-level, or public information can flow to secret one but not the opposite
+level, or public information ``can flow'' to secret one but not the opposite
 (represented using the familiar $\sqsubseteq$ binary operator).
@@
 \noindent Control flow is achieved using the \texttt{Call} and \texttt{Ret}
-instructions, that let the program jump back and forth within the instruction
+instructions, which let the program jump back and forth within the instruction
 memory using specially labelled values in the stack respresenting memory
@@
 memory pointer) the machine needs to check that both the label of the program
-counter and the label of the pointer together ``can flow'' to the label of the
+counter and the label of the pointer together can flow to the label of the
 destination memory cell.
@@
 %
-WebAssembly programs are first validated and later executed in a sandboxed
-environment (a virtual machine), making this language an attractive target for
-virtualization in \emph{functions-as-a-service} platforms.
+WebAssembly programs are first validated (using a simple form of type-checking)
+and later executed in a sandboxed environment (a virtual machine).
+%
+%% , making this language an attractive target for virtualization in
+%% \emph{functions-as-a-service} platforms.
 %
@@
     IEEE754 floating point numbers of either 32 or 64 bits;
-  \item values of these types are manipulated by writing functions using
-    sequences of stack instructions;
+  \item values of these types are by writing functions written using sequences
+    of stack instructions;
   \item functions are organized in modules and must be explicitly imported and
@@
 We want to remark that, although here for simplicity we only focus on finding
-false positives, in a realistic test suite, one would also want to test for
-false negatives, i.e., when a module is valid and \textit{haskell-wasm} rejects
+false negatives, in a realistic test suite, one would also want to test for
+false positives, i.e., when a module is valid and \textit{haskell-wasm} rejects
 it.
@@
 %
-To achieve this, we need the generated test cases to comply a with stable
+To achieve this, we need the generated test cases to comply a with common
 interface that can be invoked both by \textit{haskell-wasm} and the reference
@@
 Worth mentioning again, a realistic test suite should at least include different
-variants of this property testing functions of several different types.
+variants of this property testing functions of several different types, as well
+as properties testing multiple functions simultaneously.
 
@@
 
-Additionally, we collected the Failure Rate (FR) observed for each bug, i.e. the
+Additionally, we collected the Failure Rate observed for each bug, i.e. the
 proportion of times each tool finds each bug within the one-hour testing bugdet.
@@
 
-The results of this case study are shown in Fig. \ref{fig:results:ifc}, ordered
-by the failure rate achieved by \fuzzchick in decreasing order --- notice the
-logarithmic scale used on the MTTF.
-%
-As introduced earlier, \fuzzchick can only find 5 out of the 20 bugs injected
-with a \%100 success rate within the one hour testing budget.
-%
-In turn, our tool manages to find every bug on all runs, and taking less than a
-minute in the worst absolute case.
-%
-These results are somewhat pesimistic towards \fuzzchick if compared to the ones
-presented originally by \citeauthor{lampropoulos2019coverage}.
-%
-However, we want to remark that our experiments encompass 30 independent runs
-instead of the 5 originally used when presenting \fuzzchick.
-%
-Moreover, since the MTTF simply aggregates all runs, regardless of if they found
-a bug or timed out, this metric is quite sensitive to the particular timeout
-used on each experiment, and will tend to inflate the results as soon as the
-failure rate goes below 1.
-%
-For this reason, additionally comparing the failure rate between tools gives us
-a better estimation of the reliability of both tools, where \mutagen shows a
-clear improvement.
-
-If we only consider the succesfull runs where \fuzzchick does not time out, we
-observed a peculiarity.
-%
-This tool either finds bugs relatively quickly (after a few thouthands tests) or
-does not find them at all within the time budget, which suggest that its power
-scheduler assigns some mutation candidates too little energy before they become
-uniteresting and get ultimately thrown away.
-%
-Under this consideration, it is fair to think that \fuzzchick might be a better
-choice if we set a shorter timeout and run it several times.
-%
-Thus, to be an improvement over \fuzzchick, our tool must be able to find bugs
-not only more reliably, but also relatively fast!
-
-Table \ref{table:ifc:tests} shows a comparison between the mean number of tests
-required by both tools to find the first failure when we only consider
-successful runs.
-%
-Additionally, as adviced by \cite{arcuri2014hitchhiker} when comparing random
-testing tools, we computed the non-parametric Vargha-Delaney $A_{12}$ measure
-\cite{vargha2000critique}.
-%
-Intuitively, this measure encodes the probability of \mutagen to yield better
-results than \fuzzchick.
-
 \begin{wrapfigure}{r}{0.5\textwidth}
 \centering
-%% \vspace{-1}
+\vspace{-5pt}
 \footnotesize
@@
 \end{wrapfigure}
+
+The results of this case study are shown in Fig. \ref{fig:results:ifc}, ordered
+by the failure rate achieved by \fuzzchick in decreasing order --- notice the
+logarithmic scale used on the MTTF.
+%
+As introduced earlier, \fuzzchick can only find 5 out of the 20 bugs injected
+with a \%100 success rate within the one hour testing budget.
+%
+In turn, our tool manages to find every bug on all runs, and taking less than a
+minute in the worst absolute case.
+%
+These results are somewhat pesimistic towards \fuzzchick if compared to the ones
+presented originally by \citeauthor{lampropoulos2019coverage}.
+%
+However, we want to remark that our experiments encompass 30 independent runs
+instead of the 5 originally used when presenting \fuzzchick.
+%
+Moreover, since the MTTF simply aggregates all runs, regardless of if they found
+a bug or timed out, this metric is quite sensitive to the particular timeout
+used on each experiment, and will tend to inflate the results as soon as the
+failure rate goes below 1.
+%
+For this reason, additionally comparing the failure rate between tools gives us
+a better estimation of the reliability of both tools, where \mutagen shows a
+clear improvement.
+
+%% If we only consider the succesfull runs where \fuzzchick does not time out, we
+After analysing the results obtained using \fuzzchick, we observed a
+peculiarity.
+%
+This tool either finds bugs relatively quickly (after a few thouthands tests) or
+does not find them at all within the time budget, which suggest that its power
+scheduler assigns some mutation candidates too little energy before they become
+uniteresting and get ultimately thrown away.
+%
+Under this consideration, it is fair to think that \fuzzchick might be a better
+choice if we set a shorter timeout and run it several times.
+%
+Thus, to be an improvement over \fuzzchick, our tool must be able to find bugs
+not only more reliably, but also relatively fast!
+
+Table \ref{table:ifc:tests} shows a comparison between the mean number of tests
+required by both tools to find the first failure \emph{when we only consider
+  successful runs.}
+%
+Additionally, as adviced by \cite{arcuri2014hitchhiker} when comparing random
+testing tools, we computed the non-parametric Vargha-Delaney $A_{12}$ measure
+\cite{vargha2000critique}.
+%
+Intuitively, this measure encodes the probability of \mutagen to yield better
+results than \fuzzchick.
 %
@@
 %
-In first place, this case study does not seem to be affected by using a FIFO
-scheduling.
+In first place, this case study does not seem to be strongly affected by using a
+FIFO scheduling, where the same benchmark takes rougly only \%20 longer to
+complete when disabling this heuristic.
 %
@@
 processed before the next one is enqueued.
+%
+So, the proportion of time this heuristic is actually active tends to be minor
+under this case study.
 
@@
 %
-havig fixed the amount of random mutations to 25 adds a seemingly constant
+having fixed the amount of random mutations to 25 adds a seemingly constant
 overhead when finding most of the (easier) bugs, suggesting that we are spending
@@
 However, some of the hardest to find bugs cannot be reliably exposed using this
-amount of random mutations (\#4, \#9, \#12, \#18 and \#10), suggesting that one
-should increase this number even further to be able to discover all bugs within
-the time budget.
+fixed amount of random mutations (\#4, \#9, \#12, \#18 and \#10), suggesting
+that one should increase this number even further to be able to discover all
+bugs within the time budget.
 %
@@
 The results of this case study are shown in Fig. \ref{fig:results:wasm}, ordered
-by the mean-time-to-failure achieved by \mutagen.
+by the MTTF achieved by \mutagen.
 %
@@
 %
-There, we can quickly conclude that \quickcheck is well suited to find most of
-the bugs --- it merely finds a counterexample for the easier bugs \#4 and \#5 in
-just 1 out of 30 runs!
+There, we can quickly conclude that \quickcheck is not well suited to find most
+of the bugs --- it merely finds a counterexample for the easier bugs \#4 and \#5
+in just 1 out of 30 runs!
 %
@@
 %
-Notably, \quickcheck find counterexamples for the bug \%14 almost instantly.
+Notably, \quickcheck finds counterexamples for the bug \#14 almost instantly.
 %
@@
 Finding where the best tradeoff lies in cases where the heuristics cause mixed
-results like this is an interesting direction to aim our future work.
+results like this is another interesting direction to aim our future work.
 
@@
 It is easy to conclude that \mutagen executes tests several times slower than
-\quickcheck --- roughly 16x and 70x slower when testing \textit{prop\_validator}
+\quickcheck --- roughly 20x and 75x slower when testing \textit{prop\_validator}
 and \textit{prop\_interpreter\_i32}, respectively.
 %
-Part of the reason behind this slowdown is that the code instrumentation our
-tool currently uses is built upon a superficial tranformation that logs traces
-at the user level.
-%
-We anticipate that a large portion of the overhead could be reduced considerably
-by integrating the instrumentation directly into the Haskell runtime --- a
-substantial challenge to tackle in the future.
-
-More importantly, in the same time, our tool is capable of running many more
-tests that pass the sparse preconditions than \quickcheck, and which ultimately
-lead us to finding bugs.
+This slowdown is partly caused because the code instrumentation our tool
+currently uses is built upon a superficial tranformation that logs traces at the
+user level.
+%
+We anticipate that this overhead could be reduced considerably by integrating
+the instrumentation directly into the Haskell runtime --- a substantial
+challenge to tackle in the future.
+
+Despite this, our tool is still capable of running substantially more tests that
+pass the sparse preconditions than \quickcheck in the same amount of time, and
+which ultimately lead us to finding bugs.
 
@@
 Obtaining good random generators automatically from static information (e.g.,
-grammars or datatype definitions) has been tackled from different angles.
+grammars or data type definitions) has been tackled from different angles.
 
@@
 In Haskell, DRAGEN \cite{DBLP:conf/haskell/MistaRH18} is a meta-programming tool
-that synthezises random generators from datatypes definitions, using stochastic
+that synthezises random generators from data types definitions, using stochastic
 models to predict and optimize their distribution based on a target one set by
@@
 properties of the generated data simply by construction, e.g., mutations that
-\emph{always} produce well-typed subexpressions that refer only to identifiers
-in scope.
+\emph{always} produce well-typed programs where variables are always in scope.
 %

### 2021-04-09 papers/oopsla-21/main.tex
Added abstract. Grammar and syntax check.

@@
 \begin{abstract}
-  Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-  facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc
-  porta vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam
-  posuere. Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada
-  massa, quis varius mi purus non odio. Pellentesque condimentum, magna ut
-  suscipit hendrerit, ipsum augue ornare nulla, non luctus diam neque sit amet
-  urna. Curabitur vulputate vestibulum lorem. Fusce sagittis, libero non
-  molestie mollis, magna orci ultrices dolor, at vulputate neque nulla lacinia
-  eros. Sed id ligula quis est convallis tempor. Curabitur lacinia pulvinar
-  nibh. Nam a sapien.
+Automatically synthesized random data generators are an appealing option when
+using property-based testing to validate our systems.
+%
+Such random generators can be obtained using a variety of techniques that work
+by extracting static information from the codebase in order to produce random
+test cases.
+%
+For the most part, however, such techniques are not suitable for deriving
+generators producing random values satisfying complex invariants, which in turn
+complicates testing properties with sparse preconditions.
+
+\fuzzchick is a recent tool that borrows ideas from the fuzzing community to
+alleviate this limitation.
+%
+It enhances automatically synthesized generators with type-preserving mutations
+guided by execution traces.
+%
+Despite being enlighthening, we observed several limitations in the
+implementation of \fuzzchick that reduce its applicability.
+%
+These limitations include a large reliance on randomness when mutating test
+cases, as well as poor scheduling in certain situations, among others.
+
+In this work we present \mutagen, a coverage-guided, property-based testing
+framework that builds upon the ideas behind \fuzzchick.
+%
+Instead of relying heavily on randomness, \mutagen uses an exhaustive approach
+to generate new mutants.
+%
+This is coupled with heuristics that help to schedule mutants based on their
+novelty and to minimize mutating subexpressions of large enumeration types more
+than necessary.
+%
+Our results indicate that our approach is capable of outperforming the simpler
+one used by \fuzzchick, as well as finding subtle bugs in real-world scenarios.
 \end{abstract}
@@
 In some extreme cases, writing highly-tuned random generators capable of
-exercising every part of a complex system in a reasonable basis can take several
-thouthand person-hours of trial and error \cite{lampropoulos2019coverage}.
+exercising every part of a complex system on a reasonable basis can take several
+thousand person-hours of trial and error \cite{lampropoulos2019coverage}.
 
 To alleviate this, there exist several techniques to automatically synthesize
-random data generators by reyfing the static information present in the
+random data generators by reifying the static information present in the
 codebase, e.g., data type definitions, application public interfaces (APIs),
@@
 %
-Randomly generating valid programs is common example of this problem, where
-developers are forced to put a substantial efforts in writing specialized random
+Randomly generating valid programs is a common example of this problem, where
+developers are forced to put substantial efforts in writing specialized random
 generators by hand \cite{Palka11, perenyi2020stack, yang2011finding}.
@@
 \fuzzchick keeps a queue of interesting previously executed test cases that can
-be mutated using type-preserving transformations in order to produce new ones.
+be mutated using type-preserving transformations to produce new ones.
 %
@@
   are enqueued frequently}, and
-\item \emph{the mechanism used to assing a certain mutation energy to each
+\item \emph{the mechanism used to assign a certain mutation energy to each
   interesting test case requires fine tuning and can be hard to generalize to
@@
 This is, given an interesting, previously executed test case, our tool
-precomputes and schedules every type-presserving mutation that can be applied to
+precomputes and schedules every type-preserving mutation that can be applied to
 it.
@@
 %
-On the other hand, scheduling mutations exaustively eliminates the need of a
+On the other hand, scheduling mutations exhaustively eliminates the need of a
 heuristic that assigns a given mutation energy to each input test case.
@@
 along with \quickcheck, one of the most popular tools of this sort and often
-used as the baseline when comparing PBT algoritms.
+used as the baseline when comparing PBT algorithms.
 %
@@
 \noindent For simplicity, here we instantiate the tree payload with integers,
-altough this predicate should clearly hold for any other type with a properly
-defined notion of equality as well.
+although this predicate should hold for any other type with a properly defined
+notion of equality as well.
 
@@
 The first line states that we will provide an \texttt{Arbitrary} instance for
-trees with payload of type \texttt{a}, provided that values of type \texttt{a}
+trees with a payload of type \texttt{a}, provided that values of type \texttt{a}
 can also be randomly generated.
@@
 Moreover, \quickcheck internally keeps track of the \emph{maximum generation
-  size}, a parameter that can be tuned by the user in order to limit the size of
-the randomly generated values.
+  size}, a parameter that can be tuned by the user to limit the size of the
+randomly generated values.
 %
@@
 In turn, when the generation size is strictly positive (\texttt{gen n}), the
-generator is able to perform a random uniform choice between generating either a
-single leaf or a branch.
+generator can perform a random uniform choice between generating either a single
+leaf or a branch.
 %
@@
 
-In particular, \emph{automatically derived generators are remarkably uneffective
+In particular, \emph{automatically derived generators are remarkably ineffective
   when used to test properties with sparse preconditions.}
@@
 At this point it is reasonable to think that, to obtain the best results when
-using PBT over complex systems, one is forced to put a large amount of time on
+using PBT over complex systems, one is forced to put a large amount of time into
 developing manually-written generators.
@@
 ones, i.e., those passing the sparse preconditions of our testing properties as
-opossed to those that are discarded early, \fuzzchick exploits this information
+opposed to those that are discarded early, \fuzzchick exploits this information
 in order to focus the testing efforts into mutating valid test cases with a
@@
 %
-Moreover, to favour mutating interesting valid test cases, they get more energy
+Moreover, to favor mutating interesting valid test cases, they get more energy
 than those that were discarded.
@@
 \begin{inparaenum}
-\item \emph{shrinking the value}, replacing its outtermost data constructor with
+\item \emph{shrinking the value}, replacing its outermost data constructor with
   one that contains a subset of its fields, reusing existing subexpressions;
-\item \emph{growing the value}, replacing its outtermost data constructor with
+\item \emph{growing the value}, replacing its outermost data constructor with
   one that contains a superset of its fields, reusing existing subexpressions
@@
 \end{verbatim}
+\vspace{-5pt}
 \caption{\label{fig:fuzzchick:mutator}\fuzzchick mutator for the \texttt{Tree} data type. }
+\vspace{-10pt}
 \end{figure}
@@
 
-In this mutator, branches can be shrinked into leaves by dropping the subtrees,
+In this mutator, branches can be shrunk into leaves by dropping the subtrees,
 whereas leaves can grow into branches, by reusing the payload and generating two
@@
 However, the authors acknowledge that certain parts of its implementation have
-room for improvement, especially when it comes to the mutators design.
+room for improvement, especially when it comes to the mutator's design.
 %
@@
 data structures, potentially excluding interesting applications that might
-require producing deeper valid values, e.g., programs, network protocols
+require producing deeper valid values, e.g., programs, network protocol
 interactions, etc.
@@
 Ideally, mutations should be able to happen on every subexpression of the input
-seed in a reasonable basis.
-
-On the other hand, using random generators to produce neeeded subexpressions
+seed on a reasonable basis.
+
+On the other hand, using random generators to produce needed subexpressions
 when growing data constructors can be dangerous, as we are introducing the very
@@
 %
-In the extreme case, they might not processed at all within the testing budget.
+In an extreme case, they might not be processed at all within the testing
+budget.
 %
 Ideally, one would like a mechanism that prioritizes mutating test cases that
-discovers new portions of the code right away, and that is capable of jumping
+discover new portions of the code right away, and that is capable of jumping
 back and forth from mutation candidates whenever this happens.
@@
 If it assigns too much energy to certain not-so-interesting seeds, some bugs
-might not be discovered in a timely basis.
+might not be discovered on a timely basis.
 %
@@
 within the small available energy window --- randomly generating the same test
-case later on does not help, as it becomes uninteresting based on historic trace
-information.
+case again later does not help, as it becomes uninteresting based on historic
+trace information.
 
@@
 
-In constrast with \fuzzchick, \mutagen does not employ a power schedule to
-assign energy to mutable candidates.
-%
-In turn, it resorts to mutate them in an exhaustive and precise basis, where
+In contrast with \fuzzchick, \mutagen does not employ a power schedule to assign
+energy to mutable candidates.
+%
+In turn, it resorts to mutate them on an exhaustive and precise manner, where
 %
@@
 %
-In simple words, such tools work by enumerating all possible values of the
-data types used in the testing properties, and then executing them from smaller
-to larger until a a bug is found, or a certain size bound is reached.
+In simple words, such tools work by enumerating all possible values of the data
+types used in the testing properties, and then executing them from smaller to
+larger until a bug is found, or a certain size bound is reached.
 %
@@
 Pure mutants are used most of the time, and encode simple deterministic
-transformations over the outtermost data constructor of the input --- recursive
+transformations over the outermost data constructor of the input --- recursive
 mutations will be introduced soon.
 %
-These tranformations can:
+These transformations can:
 %
@@
 \item return an immediate subexpression of the same type as the input;
-\item swap the outtermost data constructor with any other constructor of the
+\item swap the outermost data constructor with any other constructor of the
   same type, reusing existing subexpressions; and
@@
 Notice how this definition simply enumerates mutants that transform the
-outtermost data constructor, hence no recursion is needed here.
+outermost data constructor, hence no recursion is needed here.
 %
@@
 %
-This value corresponds to the simplest value we can construct in order for the
+This value corresponds to the simplest expression we can construct for the
 mutant to be type-correct.
@@
 Using a small value is again inspired by exhaustive bounded testing tools, and
-avoids introducing unnecesary randomness when growing data constructors.
+avoids introducing unnecessary randomness when growing data constructors.
 
@@
 \end{verbatim}
+\vspace{-5pt}
 \caption{\label{fig:mutagen:mutator}\mutagen mutator for the \texttt{Tree} data
   type.}
+\vspace{-10pt}
 \end{figure}
@@
 mutating values of large enumeration types --- or any other type the user might
-want to introduce random mutations to.
+want to use random mutations with.
 %
 Instead of mutating into every other possible number or character, we resolve in
-using a random generator and sample a small amount of values from it --- the
+using a random generator and sample a small number of values from it --- the
 precise amount is a tunable parameter of \mutagen.
@@
 \noindent where the function \texttt{node} simply builds a node of the positions
-rose tree and propagates the accummulated prefix (the path from the root) to its
+rose tree and propagates the accumulated prefix (the path from the root) to its
 children.
@@
 
-\noindent Later, given a desired position to mutate within a test case, we
+\noindent Later, given the desired position to mutate within a test case, we
 define another function \texttt{inside} that finds the subexpression
@@
 The rest of the function takes care of unwrapping and rewrapping the
-intermediate subexpressions and is not partically relevant for the point being
+intermediate subexpressions and is not particularly relevant for the point being
 made.
@@
 %
-For each one of these, there are two posible cases:
+For each one of these, there are two possible cases:
 %
@@
 %
-At the end, we simply return the accummulated batch.
+At the end, we simply return the accumulated batch.
 
@@
 \end{multicols}
+\vspace{-10pt}
 \end{figure}
 
-Another small difference between \mutagen and \fuzzchick is the criteria for
+Another small difference between \mutagen and \fuzzchick is the criterion for
 enqueuing discarded tests.
@@
 
-The next section introduces two heurisitics we added to the base testing loop of
+The next section introduces two heuristics we added to the base testing loop of
 \mutagen based on the limitations we found in \fuzzchick.
@@
 \fuzzchick uses AFL instrumentation under the hood, which in turn uses an
-\emph{edge coverage} criteria to distinguish novel executions and to assign each
-mutation candidate a given energy.
+\emph{edge coverage} criterion to distinguish novel executions and to assign
+each mutation candidate a given energy.
 %
@@
 %
-Using this criteria lets us gather precise information from the each new
-execution.
+Using this criterion lets us gather precise information from each new execution.
 %
@@
 \begin{figure}[b]
+\vspace{-15pt}
 \newcommand\stacked[2]{\underset{\textstyle #2\mathstrut}{#1}}
@@
 \end{equation}
+\vspace{-10pt}
 \caption{\label{fig:mutagen:tracelog}Inserting two new execution traces into the
@@
 
-To pick the next test case, we simply retrieve the one with highest priority
+To pick the next test case, we simply retrieve the one with the highest priority
 (the smallest branching depth).
@@
 %%   point being made. }
+\vspace{-20pt}
 \end{figure}
@@
 As introduced in Section \ref{sec:mutagen}, our tool is parameterized by the
-amount of random mutations to be generated over each mutable subexpression
+number of random mutations to be generated over each mutable subexpression
 defined using a random mutant, e.g., numeric values, characters, etc.
@@
 But, how many random mutations should we use? A single one? A few tenths? A few
-hundreds?
-%
-Clearly, using too little, can put finding bugs at risk.
+hundred?
+%
+Using too little can put finding bugs at risk.
 %
 For instance, when the system under test branches based on numeric values, we
-should make sure that we set enough random mutations to test each branch in a
+should make sure that we set enough random mutations to test each branch on a
 reasonable basis.
 %
-Using too many, the other hand, can degrade the performance of the testing
+Using too many, on the other hand, can degrade the performance of the testing
 loop, as it will spend too much time producing uninteresting mutations.
@@
 %
-Our realization is that we can use this information to automatically optimize
-the number of random mutations used by our tool.
+We realized that we can use this information to automatically optimize the
+number of random mutations used by our tool.
 %
@@
 \begin{inparaenum}
-  \item we start the testing loop with the amount of random mutations set to one,
+  \item we start the testing loop with the number of random mutations set to
+    one,
   %
@@
     threshold (1000 tests seems to be a reasonable value in practice), we
-    increment the amount of random mutations and the threshold by twice the
+    increment the number of random mutations and the threshold by twice the
     current amount.
@@
 Then, if the execution of the system under test depends heavily on the values
-stored at randomly mutable subexpressions, starting with a single random mutation
-will quickly saturate the trace space, and this heuristic will continuosly
-increase the random mutations parameter until that stops happening.
+stored at randomly mutable subexpressions, starting with a single random
+mutation will quickly saturate the trace space, and this heuristic will
+continuously increase the random mutations parameter until that stops happening.
 
@@
   \textit{haskell-wasm}.}
-\vspace{-20pt}
+\vspace{-25pt}
 \end{table}
@@
 instructions, which let the program jump back and forth within the instruction
-memory using specially labelled values in the stack respresenting memory
+memory using specially labeled values in the stack representing memory
 addresses.
@@
 Finally, machine states can be modeled using a 4-tuple \textit{(pc, stk, m, i)}
-that represents a particular configuration of the program counter, stack and
-data- and instruction memories, respectively.
+that represents a particular configuration of the program counter, the stack and
+the data- and instruction memories, respectively.
 %
@@
 %
-In simple terms, this property asserts that, given two indistinguishable
-machines states, running a single instruction on both machines brings them to
-resulting states that are again indistinguishable.
+In simple terms, this property asserts that, given two indistinguishable machine
+states, running a single instruction on both machines brings them to resulting
+states that are again indistinguishable.
 
@@
 As demonstrated by \citeauthor{lampropoulos2019coverage}, generating two
-independent machine states using \quickcheck has virtually no chances of
+independent machine states using \quickcheck has virtually no chance of
 producing valid indistinguishable ones.
@@
 %
-Such table contains: the dynamic check that the abstract machine needs to
+This table contains: the dynamic check that the abstract machine needs to
 perform in order to enforce the IFC policy, along with the corresponding
@@
 In this case study, we systematically introduce several bugs on the IFC
-enforcing mechanism by removing or weaking the checks stored in this rule table.
+enforcing mechanism by removing or weakening the checks stored in this rule
+table.
 
@@
   \item it contains only four numerical types, representing both integers and
-    IEEE754 floating point numbers of either 32 or 64 bits;
-  \item values of these types are by writing functions written using sequences
-    of stack instructions;
+    IEEE754 floating-point numbers of either 32 or 64 bits;
+  \item values of these types are by manipulated by functions written using
+    sequences of stack instructions;
   \item functions are organized in modules and must be explicitly imported and
     exported;
-  \item memory blocks can be imported, exported and grown dinamically;
+  \item memory blocks can be imported, exported and grown dynamically;
 \end{inparaenum}
@@
 
-In this work, we are interested on using \mutagen to test the two most complex
+In this work, we are interested in using \mutagen to test the two most complex
 subsystems of \textit{haskell-wasm}: the \emph{validator} and the
@@
 Our tool is an attractive match for testing \textit{haskell-wasm}, as the space
-of WebAssembly programs that can be respresented using its AST contains mostly
+of WebAssembly programs that can be represented using its AST contains mostly
 invalid ones, and automatically derived random generators cannot satisfy all the
@@
 
-Unsurprisingly, this engine had several subtle latent bugs that were not caught
-by the existing unit tests and that we discovered using \mutagen while
-developing the test suite used in this work.
-%
-Moreover, \mutagen exposed two discrepancies between \textit{haskell-wasm} and
-the reference implementation.
-%
-Not severe enough to be classified as bugs, these discrepancies trigger parts of
-the WebAssembly specification that are either not yet suported by the reference
-implementation (multi-value blocks), or that produce a well-known
-non-deterministic undefined behavior allowed by the specification (NaN
-reinterpretation) \cite{perenyi2020stack}.
-%
-These findings are briefly outlined in Table \ref{table:wasm:bugs}.
-
-The rest of this section introduces the test suite we used to test the different
-parts of \textit{haskell-wasm}.
-
-\begin{table}[b]
+\begin{table}[t]
 \footnotesize
@@
   \textit{haskell-wasm}.}
+\vspace{-30pt}
 \end{table}
+
+Unsurprisingly, this engine had several subtle latent bugs that were not caught
+by the existing unit tests and that we discovered using \mutagen while
+developing the test suite used in this work.
+%
+Moreover, \mutagen exposed two discrepancies between \textit{haskell-wasm} and
+the reference implementation.
+%
+Not severe enough to be classified as bugs, these discrepancies trigger parts of
+the WebAssembly specification that are either not yet supported by the reference
+implementation (multi-value blocks), or that produce a well-known
+non-deterministic undefined behavior allowed by the specification (NaN
+reinterpretation) \cite{perenyi2020stack}.
+%
+These findings are briefly outlined in Table \ref{table:wasm:bugs}.
+
+%% The rest of this section introduces the test suite we used to test the different
+%% parts of \textit{haskell-wasm}.
 
@@
 %
-To achieve this, we need the generated test cases to comply a with common
+To achieve this, we need the generated test cases to comply a with a common
 interface that can be invoked both by \textit{haskell-wasm} and the reference
@@
               , exports   = [ Export name (ExportFunc 0) ]
-              , mems      = [ Memory (Limit 1 Nothing) ]
-              }
+              , mems      = [ Memory (Limit 1 Nothing) ] }
 \end{verbatim}
@@
 Using this stub module, we can define a testing property parameterized by the
-function type signature, the function implementation and a list of invocation
+function type signature, the function implementation, and a list of invocation
 arguments:
@@
 prop_interpreter ty fun args =
-  (discardAfter 20)
-  (do let m = mk_module ty "f" fun;
+  do let m = mk_module ty "f" fun;
      resHs   <- invokeHaskell m "f" args
      resSpec <- invokeSpec    m "f" args
-      return (equivalent resHs resSpec))
+     return (equivalent resHs resSpec)
 \end{verbatim}
@@
 signature, and uses it to invoke both the \textit{haskell-wasm} and reference
-implementation interepreters with the provided arguments.
+implementation interpreter with the provided arguments.
 %
@@
 \ref{table:wasm:bugs}), and our equivalence relation needs to take that into
-acount.
-%
-Notice that we additionally set a 20ms timeout to discard potentially diverging
+account.
+%
+Moreover, we additionally set a 20ms timeout to discard potentially diverging
 programs with infinite loops.
@@
 
-This specilized property let us generate functions using this fixed type and
+This specialized property lets us generate functions using this fixed type and
 invoke them with the exact number and type of arguments required.
 %
-In our experiments (presented in next section), we use this property when
+In our experiments (presented in the next section), we use this property when
 finding all the injected bugs into the \textit{haskell-wasm} interpreter.
@@
 Additionally, we collected the Failure Rate observed for each bug, i.e. the
-proportion of times each tool finds each bug within the one-hour testing bugdet.
+proportion of times each tool finds each bug within the one-hour testing budget.
 %
@@
   \includegraphics[width=\linewidth]{tikz/ifc-fr.tikz}
+  \vspace{-20pt}
   \caption{\label{fig:results:ifc} Comparison between \fuzzchick and \mutagen
     across 20 different bugs for the IFC stack machine. }
+\vspace{-10pt}
 \end{figure}
+
+The results of this case study are shown in Fig. \ref{fig:results:ifc}, ordered
+by the failure rate achieved by \fuzzchick in decreasing order --- notice the
+logarithmic scale used on the MTTF.
+%
+As introduced earlier, \fuzzchick can only find 5 out of the 20 bugs injected
+with a \%100 success rate within the one hour testing budget.
+%
+In turn, our tool manages to find every bug on all runs, and taking less than a
+minute in the worst absolute case.
 
@@
   Estimates in {\color{blue} blue} and {\color{red} red} indicate results
-  favouring \fuzzchick and \mutagen, respectively. }
+  favoring \fuzzchick and \mutagen, respectively. }
 \vspace{-20pt}
@@
 
-The results of this case study are shown in Fig. \ref{fig:results:ifc}, ordered
-by the failure rate achieved by \fuzzchick in decreasing order --- notice the
-logarithmic scale used on the MTTF.
-%
-As introduced earlier, \fuzzchick can only find 5 out of the 20 bugs injected
-with a \%100 success rate within the one hour testing budget.
-%
-In turn, our tool manages to find every bug on all runs, and taking less than a
-minute in the worst absolute case.
-%
-These results are somewhat pesimistic towards \fuzzchick if compared to the ones
-presented originally by \citeauthor{lampropoulos2019coverage}.
+These results are somewhat pessimistic towards \fuzzchick if compared to the
+ones presented originally by \citeauthor{lampropoulos2019coverage}.
 %
@@
 %
-For this reason, additionally comparing the failure rate between tools gives us
-a better estimation of the reliability of both tools, where \mutagen shows a
-clear improvement.
+For this reason, additionally comparing the failure rate between tools give us a
+better estimation of the reliability of both tools, where \mutagen shows a clear
+improvement.
 
 %% If we only consider the succesfull runs where \fuzzchick does not time out, we
-After analysing the results obtained using \fuzzchick, we observed a
+After analyzing the results obtained using \fuzzchick, we observed a
 peculiarity.
 %
-This tool either finds bugs relatively quickly (after a few thouthands tests) or
-does not find them at all within the time budget, which suggest that its power
+This tool either finds bugs relatively quickly (after a few thousands tests) or
+does not find them at all within the time budget, which suggests that its power
 scheduler assigns some mutation candidates too little energy before they become
-uniteresting and get ultimately thrown away.
+uninteresting and get ultimately thrown away.
 %
@@
 %
-Additionally, as adviced by \cite{arcuri2014hitchhiker} when comparing random
+Additionally, as advised by \cite{arcuri2014hitchhiker} when comparing random
 testing tools, we computed the non-parametric Vargha-Delaney $A_{12}$ measure
@@
 %
-In first place, this case study does not seem to be strongly affected by using a
-FIFO scheduling, where the same benchmark takes rougly only \%20 longer to
-complete when disabling this heuristic.
+Firstly, this case study does not seem to be strongly affected by using a FIFO
+scheduling, where the same benchmark takes only roughly \%20 longer to complete
+when disabling this heuristic.
 %
@@
 candidate queues remain empty most of the time (generation mode), and when a new
-interesting candidate gets inserted, all its mutants (and their descendents) are
+interesting candidate gets inserted, all its mutants (and their descendants) are
 processed before the next one is enqueued.
@@
 %
-having fixed the amount of random mutations to 25 adds a seemingly constant
+having fixed the number of random mutations to 25 adds a seemingly constant
 overhead when finding most of the (easier) bugs, suggesting that we are spending
@@
 However, some of the hardest to find bugs cannot be reliably exposed using this
-fixed amount of random mutations (\#4, \#9, \#12, \#18 and \#10), suggesting
+fixed amount of random mutations (\#4, \#9, \#12, \#18, and \#10), suggesting
 that one should increase this number even further to be able to discover all
@@
 %
-On one hand, disabling the FIFO schedulng (case \textit{no FIFO}), the
-hardest-to-find bugs (\#7, \#8 and \#1) are no longer found on every run.
-%
-Moreover, although bugs \#2, \#9, \#8 and \#6 are found with \%100 success rate
+On one hand, disabling the FIFO scheduling (case \textit{no FIFO}), the
+hardest-to-find bugs (\#7, \#8, and \#1) are no longer found on every run.
+%
+Moreover, although bugs \#2, \#9, \#8, and \#6 are found with \%100 success rate
 across runs, it takes several times longer for \mutagen to find them.
 %
-On the other hand, disabling the the trace saturation heuristic (case \textit{no
+On the other hand, disabling the trace saturation heuristic (case \textit{no
   reset}) and using a fixed number of 25 random mutation does not affect the
@@
 %
-Perhaps more interestingly, we can observe a significant improvement on the
+Perhaps more interestingly, we can observe a significant improvement in the
 performance of \quickcheck in terms of failure rate.
@@
 size into account while computing its priority --- future work should
-inverstigate this posibility.
+investigate this possibility.
 %
@@
   \includegraphics[width=\linewidth]{tikz/wasm-legend.tikz}
+  \vspace{-20pt}
   \caption{\label{fig:results:wasm} Comparison between \quickcheck and \mutagen
     across 15 different bugs injected into \textit{haskell-wasm}. }
+\vspace{-15pt}
 \end{figure}
 
-Finally, this case study gives us the opportunity to analyze the overhead
-introduced by the code instrumentation and internal processing used in \mutagen
-versus the stateless black-box approach of \quickcheck.
+Finally, this case study allows us to analyze the overhead introduced by the
+code instrumentation and internal processing used in \mutagen versus the
+stateless black-box approach of \quickcheck.
 %
@@
 This slowdown is partly caused because the code instrumentation our tool
-currently uses is built upon a superficial tranformation that logs traces at the
-user level.
+currently uses is built upon a superficial transformation that logs traces at
+the user level.
 %
@@
 challenge to tackle in the future.
-
+%
 Despite this, our tool is still capable of running substantially more tests that
 pass the sparse preconditions than \quickcheck in the same amount of time, and
-which ultimately lead us to finding bugs.
+which ultimately leads us to find bugs.
 
@@
   each tool on the WebAssembly case study.}
+\vspace{-20pt}
 \end{table}
@@
 In Haskell, DRAGEN \cite{DBLP:conf/haskell/MistaRH18} is a meta-programming tool
-that synthezises random generators from data types definitions, using stochastic
+that synthesises random generators from data types definitions, using stochastic
 models to predict and optimize their distribution based on a target one set by
@@
 DRAGEN2 \cite{Mista2019GeneratingRS} extends this idea adding support for
-generating richer random data by extracting additional static information like
-library APIs and function input patterns from the codebase.
+generating richer random data by extracting static information like library APIs
+and function input patterns from the codebase.
 %
 % Bolztmann-brain
-Similarly, \cite{Bendkowski2017} developed a polinomial tuning mechanism based
+Similarly, \cite{Bendkowski2017} developed a polynomial tuning mechanism based
 on Boltzmann samplers \cite{Duchon2004} that synthesizes generators with
@@
 %
-However, more specialized generators can by used directly by \mutagen, and
+However, more specialized generators can be used directly by \mutagen, and
 combining them with type-preserving mutations using a hybrid technique is an
-appealing idea that we aim to address in the future.
+idea that we aim to address in the future.
 
@@
 %
-Instead, it favours a breadth-first traversal of the execution path space.
+Instead, it favors a breadth-first traversal of the execution path space.
 %
@@
 When the source code history is available, \emph{AFLGo} \cite{bohme2017directed}
-is a fuzzer that can be targeted to exercise the specific parts of the the code
+is a fuzzer that can be targeted to exercise the specific parts of the code
 affected by recent commits in order to find potential new bugs.
@@
 When using off-the-shelf symbolic executors, this technique is often limited by
-the path explosion problem, altough recent tools like \emph{QSYM}
+the path explosion problem, although recent tools like \emph{QSYM}
 \cite{yun2018qsym} demonstrate that using tailored concolic executors can help
-overcoming this limitation.
+to overcome this limitation.
 
@@
 For any algebraic type, it synthesizes a bijection between a finite prefix of
-the natural numbers and set of increasingly growing values of the input type.
+the natural numbers and a set of increasingly larger values of the input type.
 %
@@
 %
-This allows the user to easily generate random values in a uniform basis simply
-by sampling natural numbers.
+This allows the user to easily generate random values uniformly simply by
+sampling natural numbers.
 %
@@
 %
-It progressively executes the testing properties against all possible inputs
+It progressively executes the testing properties against all possible input
 values up to a certain size bound.
-
+%
 % Korat
@@
 tool that uses method specification predicates to automatically generate all
-(nonisomorphic) test cases up to a given small size.
-
-Being exhaustive, these approaches rely on prunning mechanisms to avoid
+nonisomorphic test cases up to a given small size.
+
+Being exhaustive, these approaches rely on pruning mechanisms to avoid
 populating unevaluated subexpresions exhaustively before the computational cost
@@
 %
-In contrast to fully-exhaustive testing tools, \mutagen initially relies on
+In contrast to fully exhaustive testing tools, \mutagen initially relies on
 randomness to find initial interesting mutable candidates.
@@
 %
-Gathering empirical evidence about using prunning via lazy evaluation in
-\mutagen is an effort that we keep as future work.
+Gathering empirical evidence about using pruning via lazy evaluation in \mutagen
+is an effort that we keep as future work.
 

### 2021-04-11 papers/oopsla-21/main.tex
Another pass

@@
 \usepackage{paralist}
+
 \pgfplotsset{compat=1.3}
@@
 
-\fuzzchick is a recent tool that borrows ideas from the fuzzing community to
-alleviate this limitation.
-%
-It enhances automatically synthesized generators with type-preserving mutations
-guided by execution traces.
+%% \fuzzchick is a recent tool that borrows ideas from the fuzzing community to
+%% alleviate this limitation.
+%
+\fuzzchick is a tool coverage-guided, property-based testing (CGPBT) tool that
+alleviates this limitation by enhancing automatically synthesized generators
+with type-preserving mutations guided by execution traces.
 %
@@
 
-In this work we present \mutagen, a coverage-guided, property-based testing
-framework that builds upon the ideas behind \fuzzchick.
+In this work we present \mutagen, a CGPBT framework that builds upon the ideas
+behind \fuzzchick.
 %
@@
 producing data satisfying complex invariants not easily derivable from the
-codebase.
+codebase, e.g., a data type definition.
 %
@@
 %
-In this scenario, most of the generated tests are simply discarded before they
+In this scenario, most of the generated tests are simply thrown away before they
 can be used to test the main components of our system.
@@
 
-This tool relies on execution traces to distinguish which test cases resulted
+\fuzzchick relies on execution traces to distinguish which test cases resulted
 interesting and are therefore worth mutating --- a prolific technique known as
@@
 
-Of course, this hybrid technique is not meant to replace smart manually-written
+Of course, this technique is not meant to replace smart manually-written
 generators.
@@
 is still under construction.
-
+%
 %% Problems of FuzzChick: mutations are superficial, long queues, power
@@
 %
-In this work, we take a close look at \fuzzchick for opportunities to improve.
-%
-Notably, we establish several aspects of its implementation that can be revised
---- some of these already being identified by the authors of \fuzzchick in their
-original work.
+%% In this work, we take a close look at \fuzzchick for opportunities to improve.
+%
+Notably, we establish several aspects of its implementation that can be
+revised. \footnote{Some of these aspects were already identified by the authors
+  of \fuzzchick in their original work.}
 %
@@
 \end{inparaenum}
-
+%
 Moreover, we replicated the IFC stack machine case study bundled with \fuzzchick
@@
 builds upon the basic ideas behind \fuzzchick, i.e., code instrumentation and
-type-preserving mutations.
-%
-\mutagen is introduced in detail in Section \ref{sec:mutagen}.
+type-preserving mutations (see Section \ref{sec:mutagen}).
+%
+%% \mutagen is introduced in detail in Section \ref{sec:mutagen}.
 
@@
 
-Notably, \emph{scheduling mutations exhaustively within an input test case does
-  not mean that mutations are exhaustive by themselves.}
+Notably, \emph{scheduling mutations exhaustively does not mean that mutations
+  are always exhaustive.}
 %
@@
 %
-This mechanism let us selectively escape the exhaustiveness of our tool by
-randomly sampling a small random generator a reduced number of times.
+This mechanism let us selectively escape the scalability issues of
+exhaustiveness by randomly sampling a small random generator a reduced number of
+times.
 %
@@
 %
-Moreover, we describe in detail the ideas and limitations behind \fuzzchick,
-which served as the foundation while designing \mutagen.
+We also describe in detail the ideas and limitations behind \fuzzchick.
+
+%% which served as the foundation while designing \mutagen.
 
@@
 %
-Our definition exposes this internal parameter via \quickcheck's \texttt{sized}
-combinator, allowing us to parameterize the maximum size of the randomly
-generated trees.
+Our definition exposes this size via \quickcheck's \texttt{sized} combinator,
+allowing us to parameterize the maximum size of the randomly generated trees.
 %
@@
 In turn, when the generation size is strictly positive (\texttt{gen n}), the
-generator can perform a random uniform choice between generating either a single
-leaf or a branch.
+generator can perform a random uniform choice (\texttt{oneof}) between
+generating either a single leaf or a branch.
 %
@@
 
-\noindent \quickcheck discards random inputs as soon as it finds they do not
-pass the precondition (\texttt{isBST t}).
+\noindent \quickcheck \emph{discards} random inputs as soon as it finds they do
+not pass the precondition (\texttt{isBST t}).
 %
@@
 valid mutants, avoiding the time wasted by using generic low-level mutators that
-act at the ``serialized'' level and know very little about the structure of the
-generated data, thus producing syntactically broken mutants most of the time.
+act over the ``serialized'' of the test cases and know very little about the
+structure of the generated data, thus producing syntactically broken mutants
+most of the time.
 %
@@
 Having selected the next test case, the main loop proceeds to execute it,
-capturing both the result (passed, discarded, or failed) and its execution trace
-over the system under test.
+capturing both the result (passed, discarded due to preconditions, or failed)
+and its execution trace over the system under test.
 %
@@
 %
-Despite the choice of implementation language, the ideas presented in this work
-extend beyond Haskell and functional programming languages in general.
+Notably, the ideas presented in this work extend beyond Haskell and functional
+programming languages in general.
 %
@@
 In first place, a function \texttt{positions} traverses the input and builds a
-rose tree of mutable positions.
+Rose tree of mutable positions.
 %
@@
 \noindent where the function \texttt{node} simply builds a node of the positions
-rose tree and propagates the accumulated prefix (the path from the root) to its
+Rose tree and propagates the accumulated prefix (the path from the root) to its
 children.
@@
 we first flatten all the mutable positions of the input value in level order
-(recall that positions are stored as a rose tree).
+(recall that positions are stored as a Rose tree).
 %
@@
 %
-The first one is a simple abstract stack machine that enforces
-\emph{noninterference} \cite{goguen1982security} using runtime checks.
+The first one is a simple abstract stack machine that enforces the
+hyper-property \emph{noninterference} \cite{goguen1982security} using runtime
+checks.
 %
@@
 
-To keep things simple here as well, we use a stub definition that provides a
-module that initializes a memory block and exports a single function.
-%
-This module can be instantiated by providing the definition of the single
-function, along with a name and a type signature.
-%
-In Haskell:
-
-\begin{verbatim}
-mk_module ty name fun =
-  emptyModule { types     = [ ty ]
-              , functions = [ fun ]
-              , exports   = [ Export name (ExportFunc 0) ]
-              , mems      = [ Memory (Limit 1 Nothing) ] }
-\end{verbatim}
-
-Using this stub module, we can define a testing property parameterized by the
-function type signature, the function implementation, and a list of invocation
-arguments:
+To keep things simple here as well, we use a function \texttt{mk\_module} to
+creates a stub module which initializes one memory block and exports a single
+function.
+%
+This function is parameterized by the definition of the single function, its
+name and type signature.
+%
+%% In Haskell:
+%
+%% \begin{verbatim}
+%% mk_module ty name fun =
+%%   emptyModule { types     = [ ty ]
+%%               , functions = [ fun ]
+%%               , exports   = [ Export name (ExportFunc 0) ]
+%%               , mems      = [ Memory (Limit 1 Nothing) ] }
+%% \end{verbatim}
+%
+This way, we can define a testing property that can be instantiated with
+randomly generated combinations of function types, function definitions and
+lists of invocation arguments:
+
+%% function type signature, the function implementation, and a list of invocation
+%% arguments:
 
@@
   do let m = mk_module ty "f" fun;
-     resHs   <- invokeHaskell m "f" args
-     resSpec <- invokeSpec    m "f" args
+     resHs   <- invokeHaskellWasm m "f" args
+     resSpec <- invokeRefImpl     m "f" args
      return (equivalent resHs resSpec)
@@
 Then, the property asserts whether their results are equivalent.
+%
+\footnote{In our implementation, we additionally set a 20ms timeout to discard
+  potentially diverging programs with infinite loops.}
 %
@@
 %
-Moreover, we additionally set a 20ms timeout to discard potentially diverging
-programs with infinite loops.
 
@@
 
+These results are somewhat pessimistic towards \fuzzchick if compared to the
+ones presented originally by \citeauthor{lampropoulos2019coverage}.
+%
+However, we want to remark that our experiments encompass 30 independent runs
+instead of the 5 originally used when presenting \fuzzchick.
+%
+Moreover, since the MTTF simply aggregates all runs, regardless of if they found
+a bug or timed out, this metric is quite sensitive to the particular timeout
+used on each experiment, and will tend to inflate the results as soon as the
+failure rate goes below 1.
+%
+For this reason, additionally comparing the failure rate between tools give us a
+better estimation of the reliability of both tools, where \mutagen shows a clear
+improvement.
+
 \begin{wrapfigure}{r}{0.5\textwidth}
 \centering
-\vspace{-5pt}
+\vspace{-10pt}
 \footnotesize
@@
 
-These results are somewhat pessimistic towards \fuzzchick if compared to the
-ones presented originally by \citeauthor{lampropoulos2019coverage}.
-%
-However, we want to remark that our experiments encompass 30 independent runs
-instead of the 5 originally used when presenting \fuzzchick.
-%
-Moreover, since the MTTF simply aggregates all runs, regardless of if they found
-a bug or timed out, this metric is quite sensitive to the particular timeout
-used on each experiment, and will tend to inflate the results as soon as the
-failure rate goes below 1.
-%
-For this reason, additionally comparing the failure rate between tools give us a
-better estimation of the reliability of both tools, where \mutagen shows a clear
-improvement.
-
 %% If we only consider the succesfull runs where \fuzzchick does not time out, we
@@
 \begin{table}[H]
+\vspace{-10pt}
 \begin{tabular}{|l|c|c|c|c|}
@@
 In our case studies, however, we observed that their preconditions tend to be
-quite strict when mutating from passed test cases, fully evaluating their input
-before executing the postcondition.
+quite strict when executing test cases obtained by mutating passed existing
+ones, fully evaluating their input before executing the postcondition.
 %

### 2021-04-12 papers/oopsla-21/main.tex
Made it fit withing page limit

@@
 \usepackage{paralist}
+\usepackage{enumitem}
+
+\setitemize{noitemsep,nosep,wide,topsep=0pt,parsep=0pt,partopsep=0pt}
 
@@
 %
-\fuzzchick is a tool coverage-guided, property-based testing (CGPBT) tool that
+\fuzzchick is a tool coverage-guided, property-based testing (CGPT) tool that
 alleviates this limitation by enhancing automatically synthesized generators
@@
 
-In this work we present \mutagen, a CGPBT framework that builds upon the ideas
+In this work we present \mutagen, a CGPT framework that builds upon the ideas
 behind \fuzzchick.
@@
 Our results indicate that our approach is capable of outperforming the simpler
-one used by \fuzzchick, as well as finding subtle bugs in real-world scenarios.
+one used by \fuzzchick, as well as finding bugs in real-world scenarios.
 \end{abstract}
@@
 % ----------------------------------------
+
+%% \vspace{-5pt}
 
@@
 
-Notably, \emph{scheduling mutations exhaustively does not mean that mutations
-  are always exhaustive.}
+%% Notably, \emph{scheduling mutations exhaustively does not mean that mutations
+%%   are always exhaustive.}
 %
@@
 %
-Additionally, our scheduling lets the testing loop jump back and forth between
+Moreover, our scheduling lets the testing loop jump back and forth between
 mutable test cases as soon as a new more interesting one is enqueued for
@@
 To validate our ideas, we use two main case studies (described in detail in
-Section \ref{sec:casestudies}): the IFC stack machine used by
-\citeauthor{lampropoulos2019coverage} on their original work, and a WebAssembly
-engine written in Haskell.
+Section \ref{sec:casestudies}): the Information-Flow Control (IFC) stack machine
+used by \citeauthor{lampropoulos2019coverage} on their original work, and a
+WebAssembly engine written in Haskell.
 %
@@
 %
-All in all, our evaluation encompass more than 600 hours of cpu time and suggest
-that testing mutants exhaustively can be an appealing technique for finding bugs
-reliably without sacrificing speed.
+This case study also help us comparing the performance our tool against a more
+traditional PBT approach that does not rely on code instrumentation.
+%
+All in all, our evaluation encompass more than 600 hours of computing time and
+suggest that testing mutants exhaustively can be an appealing technique for
+finding bugs reliably without sacrificing speed.
 
@@
 property, $N$ is the maximum number of tests to perform, and $gen$ is the random
-generator to be used to instantiate $P$.
+generator used to instantiate $P$.
 
@@
 
-In addition, high-level mutators are better suited for producing syntactically
-valid mutants, avoiding the time wasted by using generic low-level mutators that
-act over the ``serialized'' of the test cases and know very little about the
-structure of the generated data, thus producing syntactically broken mutants
-most of the time.
+In addition, high-level, type-preserving mutators are better suited for
+producing syntactically valid mutants, avoiding the time wasted by using generic
+low-level mutators that act over the ``serialized'' test cases and know very
+little about their structure, thus producing syntactically broken mutants most
+of the time.
 %
@@
 \caption{\label{fig:fuzzchick:mutator}\fuzzchick mutator for the \texttt{Tree} data type. }
-\vspace{-10pt}
+\vspace{-15pt}
 \end{figure}
@@
 
-\paragraph{Mutators distribution:}
+\begin{itemize}
+\item \emph{Mutators distribution:}
 %
@@
 
-\paragraph{Enqueuing mutation candidates:}
+\item \emph{Enqueuing mutation candidates:}
 %
@@
 
-\paragraph{Power schedule:}
+\item \emph{Power schedule:}
 %
@@
 the mutators.\\
+
+\end{itemize}
 
@@
 
-This idea is inspired by exhaustive bounded testing tools like SmallCheck
-\cite{runciman2008smallcheck} or Korat \cite{boyapati2002korat}, that produce
-test cases exhaustively.
+This idea is inspired by exhaustive bounded testing tools like \emph{SmallCheck}
+\cite{runciman2008smallcheck} or \emph{Korat} \cite{boyapati2002korat}, that
+produce test cases exhaustively.
 %
@@
   type.}
-\vspace{-10pt}
+\vspace{-25pt}
 \end{figure}
@@
 In this light, the mutable positions of the value \texttt{Branch (Leaf 1) 2
-  (Leaf 3)} are:
-%
+  (Leaf 3)} are as follows:
+
 \begin{equation}
@@
 \begin{figure}[t]
+\vspace{-15pt}
 \begin{multicols}{2}
@@
 \end{multicols}
-\vspace{-10pt}
+\vspace{-20pt}
 \end{figure}
@@
 
-\begin{figure}
+\begin{figure}[t]
+\vspace{-15pt}
 \begin{multicols}{2}
@@
 \textit{haskell-wasm}.
+%
+Sadly, comparing \mutagen with \fuzzchick has unfortunately been out of the
+scope of this work.
+%
+The main reason behind this is that porting \textit{haskell-wasm} into Coq, and
+later adapting our test suite to work with \fuzzchick requires a subtantial
+effort.
+%
+Instead, we focus on comparing \mutagen against a more traditional PBT approach
+using \quickcheck, to evaluate the relative overhead of the code instrumentation
+used in our tool.
 
@@
   \textit{haskell-wasm}.}
-\vspace{-30pt}
+\vspace{-20pt}
 \end{table}
@@
 To keep things simple here as well, we use a function \texttt{mk\_module} to
-creates a stub module which initializes one memory block and exports a single
-function.
+create a stub module which initializes one memory block and exports a single
+WebAssembly function.
 %
@@
 prop_interpreter ty fun args =
-  do let m = mk_module ty "f" fun;
-     resHs   <- invokeHaskellWasm m "f" args
-     resSpec <- invokeRefImpl     m "f" args
+  do let m = mk_module ty "foo" fun;
+     resHs   <- invokeHaskellWasm m "foo" args
+     resSpec <- invokeRefImpl     m "foo" args
      return (equivalent resHs resSpec)
@@
 This property instantiates the module stub using the input function and its type
-signature, and uses it to invoke both the \textit{haskell-wasm} and reference
-implementation interpreter with the provided arguments.
-%
-Then, the property asserts whether their results are equivalent.
-%
-\footnote{In our implementation, we additionally set a 20ms timeout to discard
+signature, and uses it to invoke the function \texttt{foo} both on the
+\textit{haskell-wasm} and reference implementation interpreter with the provided
+arguments.
+%
+Then, the property asserts whether their results are equivalent.%
+%
+\footnote{In our implementation, we additionally set a short timeout to discard
   potentially diverging programs with infinite loops.}
@@
 We ran each experiment 30 times except for the ones involving the bugs on the
-WebAssembly interpreter, which were run 10 times.
+WebAssembly interpreter, which were run 10 times due to time constraints.
 %
@@
 In the case of \textit{no reset}, the amount of random mutations is no longer
-controlled by this heuristic, so we fixed it to 25 random mutations throughout
-the execution of the tool.
+controlled by this heuristic, so we arbitrarily fixed it to 25 random mutations
+throughout the execution of the tool.
 
@@
 
-The results of this case study are shown in Fig. \ref{fig:results:ifc}, ordered
-by the failure rate achieved by \fuzzchick in decreasing order --- notice the
-logarithmic scale used on the MTTF.
-%
+The results of this case study are shown in Fig. \ref{fig:results:ifc} --- we
+encourage the reader to obtain a color-printed version of this paper.
+%
+The injected bugs are ordered by the failure rate achieved by \fuzzchick in
+decreasing order.
+%
+Moreover, notice the logarithmic scale used on the MTTF.
+
 As introduced earlier, \fuzzchick can only find 5 out of the 20 bugs injected
@@
 minute in the worst absolute case.
-
+%
 These results are somewhat pessimistic towards \fuzzchick if compared to the
@@
 However, we want to remark that our experiments encompass 30 independent runs
-instead of the 5 originally used when presenting \fuzzchick.
+instead of the 5 originally used when presenting \fuzzchick
+\citeyearpar{lampropoulos2019coverage}.
 %
@@
 
+%% If we only consider the succesfull runs where \fuzzchick does not time out, we
+After analyzing the results obtained using \fuzzchick, we observed a
+peculiarity.
+%
+This tool either finds bugs relatively quickly (after a few thousands tests) or
+does not find them at all within the time budget, which suggests that its power
+scheduler assigns some mutation candidates too little energy before they become
+uninteresting and get ultimately thrown away.
+%
+Under this consideration, it is fair to think that \fuzzchick might be a better
+choice if we set a shorter timeout and run it several times.
+%
+Thus, to be an improvement over \fuzzchick, our tool must be able to find bugs
+not only more reliably, but also relatively fast!
+
 \begin{wrapfigure}{r}{0.5\textwidth}
 \centering
-\vspace{-10pt}
+\vspace{-20pt}
 \footnotesize
@@
   favoring \fuzzchick and \mutagen, respectively. }
-\vspace{-20pt}
+\vspace{-25pt}
 \end{wrapfigure}
-
-%% If we only consider the succesfull runs where \fuzzchick does not time out, we
-After analyzing the results obtained using \fuzzchick, we observed a
-peculiarity.
-%
-This tool either finds bugs relatively quickly (after a few thousands tests) or
-does not find them at all within the time budget, which suggests that its power
-scheduler assigns some mutation candidates too little energy before they become
-uninteresting and get ultimately thrown away.
-%
-Under this consideration, it is fair to think that \fuzzchick might be a better
-choice if we set a shorter timeout and run it several times.
-%
-Thus, to be an improvement over \fuzzchick, our tool must be able to find bugs
-not only more reliably, but also relatively fast!
 
@@
 On one hand, disabling the FIFO scheduling (case \textit{no FIFO}), the
-hardest-to-find bugs (\#7, \#8, and \#1) are no longer found on every run.
+hardest-to-find bugs (\#7, \#10, and \#1) are no longer found on every run.
 %
@@
     across 15 different bugs injected into \textit{haskell-wasm}. }
-\vspace{-15pt}
+\vspace{-10pt}
 \end{figure}
@@
 \begin{table}[H]
-\vspace{-10pt}
+\footnotesize
+\vspace{-5pt}
 \begin{tabular}{|l|c|c|c|c|}
@@
 %
-To preserve space, in this section we only focus on three main topics:
+In this section we only focus on three main topics:
 %
@@
 %
-exhaustive bounded testing tools, all of which inspired the design of \mutagen.
+exhaustive bounded testing, all of which inspired the design of \mutagen.
 
@@
 %
-\cite{ClaessenDP14} developed an algorithm for generating inputs constrained by
-boolean preconditions with almost-uniform distribution.
+\citeauthor{ClaessenDP14} \citeyearpar{ClaessenDP14} developed an algorithm for
+generating inputs constrained by boolean preconditions with almost-uniform
+distribution.
 %
@@
 %
-Later, \cite{LampropoulosGHH17} extends this approach by adding a limited form
-of constraint solving controllable by the user in a domain-specific language
-called \emph{Luck}.
+Later, \citeauthor{LampropoulosGHH17} \citeyearpar{LampropoulosGHH17} extends
+this approach by adding a limited form of constraint solving controllable by the
+user in a domain-specific language called \emph{Luck}.
 %
@@
 %
-Recently, \cite{Lampropoulos2017} proposed a derivation mechanism to obtain
-constrained random generators directly by defining their structure using
-inductively defined relations in Coq.
+Recently, \citeauthor{Lampropoulos2017} \citeyearpar{Lampropoulos2017} proposed
+a derivation mechanism to obtain constrained random generators directly by
+defining their structure using inductively defined relations in Coq.
 
@@
 expressed elsewhere.
-
+%
 % SmallCheck and LazySmallCheck
@@
 We presented \mutagen, a coverage-guided, property-based testing framework
-written in Haskell.
+written in Haskell.%
+\footnote{As meantioned earlier, we foresee no problems into translating our
+  tool in other programming languages beyond Haskell.}
 %
@@
 This is in turn inspired by exhaustive bounded testing tools that focus on
-testing every possible input value of the system under test.
+testing every possible input value of the system under test up to a certain size
+bound.
+%
 
@@
 instrumentation in order to effectively target mutations towards specific parts
-of the code; and evaluating the effect of lazy prunning; among others.
+of the code; and evaluating the effect of lazy pruning; among others.
 %

### 2021-04-12 papers/oopsla-21/main.tex
NIT

@@
 thousand person-hours of trial and error \cite{lampropoulos2019coverage}.
-
+%
 To alleviate this, there exist several techniques to automatically synthesize
@@
 producing data satisfying complex invariants not easily derivable from the
-codebase, e.g., a data type definition.
+codebase.
 %
@@
 \fuzzchick keeps a queue of interesting previously executed test cases that can
-be mutated using type-preserving transformations to produce new ones.
+be mutated using high-level, type-preserving transformations to produce new
+ones.
 %
@@
 %
-Notably, we establish several aspects of its implementation that can be
-revised. \footnote{Some of these aspects were already identified by the authors
-  of \fuzzchick in their original work.}
+Notably, we establish several aspects of \fuzzchick that can be
+revised.%
+\footnote{Some of these aspects were already identified by the authors of
+  \fuzzchick in their original work.}
 %
@@
   are enqueued frequently}, and
-\item \emph{the mechanism used to assign a certain mutation energy to each
-  interesting test case requires fine tuning and can be hard to generalize to
-  work well under different testing scenarios.}
+\item \emph{the power schedule (i.e., the heuristic used to assign a certain
+  mutation energy to each interesting test case) requires fine tuning and can be
+  hard to generalize to work well under different testing scenarios.}
 \end{inparaenum}
 %
-Moreover, we replicated the IFC stack machine case study bundled with \fuzzchick
-and observed a noticeable limitation:
+Moreover, we replicated the Information-Flow Control (IFC) stack machine case
+study bundled with \fuzzchick and observed a noticeable limitation:
 %
@@
 %% \mutagen is introduced in detail in Section \ref{sec:mutagen}.
-
+%
 %% Exhaustive mutations
@@
 On the other hand, scheduling mutations exhaustively eliminates the need of a
-heuristic that assigns a given mutation energy to each input test case.
+power schedule to assign a given mutation energy to each input test case.
 
@@
 To validate our ideas, we use two main case studies (described in detail in
-Section \ref{sec:casestudies}): the Information-Flow Control (IFC) stack machine
-used by \citeauthor{lampropoulos2019coverage} on their original work, and a
-WebAssembly engine written in Haskell.
+Section \ref{sec:casestudies}): the IFC stack machine used by
+\citeauthor{lampropoulos2019coverage} on their original work, and a WebAssembly
+engine written in Haskell.
 %
@@
 % ----------------------------------------
+\vspace{-10pt}
 
@@
 %
-Generic fuzzing tools like \emph{AFL}, \emph{libFuzzer} \citeyearpar{libfuzzer}
-or \emph{HonggFuzz} \cite{honggfuzz}, as well as language-specific ones like
-Crowbar \cite{dolan2017testing} or \emph{Kelinci} \cite{kersten2017poster} use
-execution traces to recognize interesting test cases, e.g, those that exercise
-previously undiscovered parts of the target code.
+Generic fuzzing tools like \emph{AFL} \cite{afl}, \emph{HonggFuzz}
+\cite{honggfuzz}, or \emph{libFuzzer} \citeyearpar{libfuzzer} as well as
+language-specific ones like Crowbar \cite{dolan2017testing} or \emph{Kelinci}
+\cite{kersten2017poster} use execution traces to recognize interesting test
+cases, e.g, those that exercise previously undiscovered parts of the target
+code.
 %
@@
 If the test case does in fact discover a new path, it is enqueued at the end of
-its corresponding queue, depending on whether it passed or was discarded (i.e.,
-it failed the precondition).
+its corresponding queue, depending on whether it passed or was discarded.
+%
+%% (i.e., it failed the precondition).
 %
@@
 %
-Hence, \fuzzchick mutators can only be effectively used to transform to shallow
+Hence, \fuzzchick mutators can only be effectively used to transform shallow
 data structures, potentially excluding interesting applications that might
@@
 %
-If that mutated test case turns to be interesting, that subexpression can always
-be mutated later.
+If that mutated test case turns to be interesting, that new subexpression can
+always be mutated later.
 
@@
   \item for every mutable subexpression, each one of these mutations is
-    evaluated \emph{exactly once}.
+    scheduled \emph{exactly once}.
 \end{inparaenum}
@@
 %
-Instead of mutating into every other possible number or character, we resolve in
-using a random generator and sample a small number of values from it --- the
-precise amount is a tunable parameter of \mutagen.
+Instead of mutating every numerical or character subexpression exhaustively, we
+resolve in using a random generator and sample a small number of values from it
+--- the precise amount is a tunable parameter of \mutagen.
 %
@@
 We manually injected 10 bugs in the validator as well as 5 bugs in the
-interpreter of this engine, and used the reference WebAssembly implementation to
-find them via differential testing \cite{mckeeman1998differential}.
+interpreter of this engine (see Table \ref{table:wasm:injectedbugs}), and used
+the reference WebAssembly implementation to find them via differential testing
+\cite{mckeeman1998differential}.
 %
@@
 In this abstract machine, the enforced IFC rules are implemented using a rule
-table indexed by the operation that the machine is about to perform.
+table indexed by the instruction that the machine is about to perform.
 %
@@
 perform in order to enforce the IFC policy, along with the corresponding
-security label of the program counter and the operation result after the
+security label of the program counter and the instruction result after the
 operation is executed.
 %
-For instance, to execute the \texttt{Store} operation (which stores a value in a
-memory pointer) the machine needs to check that both the label of the program
+For instance, to execute the \texttt{Store} instruction (which stores a value in
+a memory pointer) the machine needs to check that both the label of the program
 counter and the label of the pointer together can flow to the label of the
@@
 %
-After this check, this operation overwrites the value at the destination cell
+After this check, this instruction overwrites the value at the destination cell
 and updates its label with the maximum sensibility of the involved labels.
@@
 \hline
-\textbf{Operation}
+\textbf{Instruction}
 & \textbf{Precondition Check}
@@
 
-\begin{table}[t]
+\begin{table}[b]
 \footnotesize
@@
 
-To keep things simple here as well, we use a function \texttt{mk\_module} to
-create a stub module which initializes one memory block and exports a single
+To keep things simple here as well, we write a function \texttt{mk\_module} to
+build a stub module which initializes one memory block and exports a single
 WebAssembly function.
 %
-This function is parameterized by the definition of the single function, its
-name and type signature.
+\texttt{mk\_module} is parameterized by the definition of the single WebAssembly
+function, along with its name and type signature.
 %
@@
 prop_interpreter ty fun args =
-  do let m = mk_module ty "foo" fun;
+  do let m = mk_module ty "foo" fun
      resHs   <- invokeHaskellWasm m "foo" args
@@
 Using this testing property directly might not sound like a great idea, as
-randomly generated lists of inputs will be very unlikely to match the type
-signature of randomly generated functions.
+randomly generated lists of input arguments will be very unlikely to match the
+type signature of randomly generated functions.
 %
 However, it lets us test what happens when programs are not properly invoked,
-and it quickly discovered the bug \#3 in \textit{haskell-wasm} mentioned above.
+and it quickly discovered the previously unknown bug \#3 in
+\textit{haskell-wasm} mentioned above.
 %
@@
 useful specialized version of \texttt{prop\_interpreter} that fixes the type of
-the generated function to take two arguments (of type \texttt{I32} and {F32})
-and return an \texttt{I32} as a result:
+the generated function to take two arguments (of type \texttt{I32} and
+\texttt{F32}) and return an \texttt{I32} as a result:
 
@@
 From there, we followed the same approach taken by
-\citeauthor{lampropoulos2019coverage} and collected the Mean-Time-To-Failure
+\citeauthor{lampropoulos2019coverage} and collected the mean-time-to-failure
 (MTTF) of each bug, i.e., how quickly a bug can be found in wall clock time.
@@
 
-Additionally, we collected the Failure Rate observed for each bug, i.e. the
+Additionally, we collected the failure rate observed for each bug, i.e. the
 proportion of times each tool finds each bug within the one-hour testing budget.
@@
 The results of this case study are shown in Fig. \ref{fig:results:ifc} --- we
-encourage the reader to obtain a color-printed version of this paper.
+encourage the reader to obtain a color copy of this work.
 %
@@
 
-Firstly, we focus on the bugs injected in the validator (Fig.
+We first focus on the bugs injected in the validator (Fig.
 \ref{fig:results:wasm} left).
@@
 Automatically deriving random generators is substantially more complicated when
-the generated data must satisfy by (often sparse) preconditions.
+the generated data must satisfy (often sparse) preconditions.
 %
@@
 up to a certain size bound.
-
+%
 %
@@
 We presented \mutagen, a coverage-guided, property-based testing framework
-written in Haskell.%
-\footnote{As meantioned earlier, we foresee no problems into translating our
-  tool in other programming languages beyond Haskell.}
+written in Haskell.
 %
@@
 %
+Moreover, we foresee no problems into translating our tool in other programming
+languages beyond Haskell.
 

### 2021-04-16 papers/oopsla-21/main.tex
Fixed John's comments

@@
 
+\setlength{\skip\footins}{5pt}
+
 % ----------------------------------------
@@
 \setcopyright{acmcopyright}
-\copyrightyear{2021}
-\acmYear{2021}
+%% \copyrightyear{2021}
+%% \acmYear{2021}
 %% \acmDOI{10.1145/1122445.1122456}
@@
 %
-\fuzzchick is a tool coverage-guided, property-based testing (CGPT) tool that
+\fuzzchick is a coverage-guided, property-based testing (CGPT) tool that
 alleviates this limitation by enhancing automatically synthesized generators
@@
 %
-Despite being enlighthening, we observed several limitations in the
+Despite being enlightening, we observed several limitations in the
 implementation of \fuzzchick that reduce its applicability.
@@
 
-\fuzzchick relies on execution traces to distinguish which test cases resulted
-interesting and are therefore worth mutating --- a prolific technique known as
-\emph{coverage-guided fuzzing} and popularized by tools like \emph{AFL}
+\fuzzchick relies on execution traces to distinguish which executed test cases
+were interesting and are therefore worth mutating --- a popular technique known
+as \emph{coverage-guided fuzzing} and popularized by tools like \emph{AFL}
 \cite{afl}.
@@
 
+%% Of course, this technique is not meant to replace smart manually-written
+%% generators.
+%% %
+%% In turn, it provides an acceptable solution that can be used in the early stages
+%% of development, and while a better test suite using manually-written generators
+%% is still under construction.
+%% %
+%% %% Problems of FuzzChick: mutations are superficial, long queues, power
+%% %% scheduler
+%% %
+%% %% In this work, we take a close look at \fuzzchick for opportunities to improve.
+%% %
+%% %% Notably, we establish several aspects of \fuzzchick that can be
+%% %% revised.%
+%% %
+%% But \fuzzchick suffers several disadvantages.%
+%% \footnote{Some of these were already identified by the authors of \fuzzchick in
+%%   their original work.}
+%% %
+%% In particular:
+%% %
+%% \begin{inparaenum}
+%% \item the type-preserving mutations produced by this tool are superficial
+%%   and rely heavily on randomness to produce diverse mutants,
+%% \item the queueing mechanism can cause large delays when interesting
+%%   values are enqueued frequently, and
+%% \item the heuristic used to assign a certain ``mutation energy'' to each
+%%   interesting test case (often referred to as the \emph{power schedule})
+%%   requires fine tuning and can be hard to generalize to work well under
+%%   different testing scenarios.
+%% \end{inparaenum}
+%% %
+%% Moreover, we replicated the Information-Flow Control (IFC) stack machine case
+%% study bundled with \fuzzchick and observed a surprising limitation:
+%% %
+%% \emph{after repeating each experiment 30 times, \fuzzchick was only able to find
+%%   5 out of the 20 injected bugs with a 100\% efficacy, the hardest one being
+%%   found only around 13\% of the time after an hour of testing.}
+%% %
+%% Despite this limitation, \citeauthor{lampropoulos2019coverage} show that
+%% \fuzzchick is far better than using normal random testing coupled with na\"ive
+%% random generators by comparing the mean-time-to-failure (MTTF) against
+%% \emph{QuickChick}.
+%% %
+%% However, we believe that a proper evaluation ought to
+%% %
+%% take failure to find a counterexample as an
+%% %
+%% important metric when comparing PBT tools, not just mean time to failure ---
+%% when one is found.
+
 Of course, this technique is not meant to replace smart manually-written
@@
 %
-In turn, it provides an acceptable solution that can be used in the early stages
-of development, and while a better test suite using manually-written generators
-is still under construction.
+In turn, it provides an adequate solution to be used in early stages of
+development, and while a better test suite using manually-written generators is
+still under construction.
 %
@@
 %
-Notably, we establish several aspects of \fuzzchick that can be
-revised.%
-\footnote{Some of these aspects were already identified by the authors of
-  \fuzzchick in their original work.}
+%% Notably, we establish several aspects of \fuzzchick that can be
+%% revised.%
+%
+But \fuzzchick suffers several disadvantages.%
+\footnote{Some of these were already identified by the authors of \fuzzchick in
+  their original work.}
 %
@@
 \begin{inparaenum}
-\item \emph{the type-preserving mutations produced by this tool are superficial
-  and rely heavily on randomness to produce diverse mutants},
-\item \emph{using normal queues can cause large delays when interesting values
-  are enqueued frequently}, and
-\item \emph{the power schedule (i.e., the heuristic used to assign a certain
-  mutation energy to each interesting test case) requires fine tuning and can be
-  hard to generalize to work well under different testing scenarios.}
+\item type-preserving mutations, if not done carefully, can become superficial
+  and rely heavily on randomness to produce diverse mutants,
+\item the queuing mechanism can cause large delays if interesting values are
+  enqueued frequently and there is no way to prioritize them, and
+\item using an heuristic to assign a certain ``mutation energy'' to each
+  interesting test case (often referred to as a \emph{power schedule}) requires
+  fine tuning and can be hard to generalize to work well under different testing
+  scenarios.
 \end{inparaenum}
@@
 Moreover, we replicated the Information-Flow Control (IFC) stack machine case
-study bundled with \fuzzchick and observed a noticeable limitation:
-%
-\emph{after repeating each experiment 30 times, \fuzzchick was only able to find
-  5 out of the 20 injected bugs with a 100\% efficacy, the hardest one being
-  found only around 13\% of the time after an hour of testing.}
-%
-Despite this limitation, \citeauthor{lampropoulos2019coverage} show that
-\fuzzchick is far better than using normal random testing coupled with na\"ive
-random generators by comparing the mean-time-to-failure (MTTF) against
-\emph{QuickChick}.
-%
-However, we believe that a proper evaluation ought to use the failure rate as an
-important metric when comparing PBT tools.
+study bundled with \fuzzchick and observed a surprising limitation:
+%
+after repeating each experiment 30 times, \fuzzchick could only find 5 out of
+the 20 injected bugs with a 100\% efficacy.
+%
+%% the hardest one being found only around 13\% of the time after an hour of
+%% testing.
+%
+Despite this, \citeauthor{lampropoulos2019coverage} show that \fuzzchick is far
+better than using normal random testing coupled with na\"ive random generators
+by comparing the mean-time-to-failure (MTTF) against \emph{QuickChick}.
+%
+However, we believe that a more meticulous evaluation ought to take failure to
+find a counterexample as an important metric when comparing PBT tools, not just
+mean time to failure --- when one is found.
 
@@
 %
-This is, given an interesting, previously executed test case, our tool
+That is, given an interesting, previously executed test case, our tool
 precomputes and schedules every type-preserving mutation that can be applied to
@@
 Instead, every subexpression of the input test case is mutated on the same
-basis, which guarantees that no interesting mutation is left behind due to
+basis, which guarantees that no interesting mutation is omitted due to
 randomness.
 %
-On the other hand, scheduling mutations exhaustively eliminates the need of a
+%% On the other hand, scheduling mutations exhaustively eliminates the need for a
+%% power schedule to assign a given mutation energy to each input test case.
+%
+On the other hand, scheduling mutations exhaustively eliminates the need for a
 power schedule to assign a given mutation energy to each input test case.
@@
 %
-This way, our tool avoids, for instance, mutating a given number into every
+In this way, our tool avoids, for instance, mutating a given number into every
 other number in the 32 or 64 bits range.
@@
 %
-In first place, \mutagen uses first-in first-out (FIFO) scheduling with priority
-to enqueue interesting test cases for mutation.
+In the first place, \mutagen uses first-in first-out (FIFO) scheduling with
+priority to enqueue interesting test cases for mutation.
 %
@@
 Section \ref{sec:casestudies}): the IFC stack machine used by
-\citeauthor{lampropoulos2019coverage} on their original work, and a WebAssembly
+\citeauthor{lampropoulos2019coverage} in their original work, and a WebAssembly
 engine written in Haskell.
@@
 %
-This case study also help us comparing the performance our tool against a more
-traditional PBT approach that does not rely on code instrumentation.
-%
-All in all, our evaluation encompass more than 600 hours of computing time and
-suggest that testing mutants exhaustively can be an appealing technique for
+This case study also help us to compare the performance of our tool against a
+more traditional PBT approach that does not rely on code instrumentation.
+%
+All in all, our evaluation encompasses more than 600 hours of computing time and
+suggests that testing mutants exhaustively can be an appealing technique for
 finding bugs reliably without sacrificing speed.
@@
 %% Last two sections
-Finally, we discuss related work on Section \ref{sec:related} and conclude
-on Section \ref{sec:conclusions}.
+Finally, we discuss related work in Section \ref{sec:related} and conclude
+in Section \ref{sec:conclusions}.
 
 % ----------------------------------------
-\vspace{-10pt}
+\vspace{-5pt}
 
@@
 
+\vspace{-5pt}
 \subsection{Property-Based Testing and \quickcheck}
@@
 %
-Conversely, assigning too little energy to interesting test cases might cause
+Conversely, assigning too little energy to interesting test cases might mean
 that some bugs cannot be discovered at all unless the right mutation happens
@@
 To alleviate this problem, such tools often rely on heuristics to prune the
-search space based on detecting unevaluated subexpressions --- this is discussed
-in Section \ref{sec:related}.
+search space based on detecting unused subexpressions --- this is discussed in
+Section \ref{sec:related}.
 
@@
 %
-Moreover, like in \fuzzchick, the testing loop of our tool automatically filters
+Moreover, as in \fuzzchick, the testing loop of our tool automatically filters
 the test cases that are worth mutating by keeping only those that discover new
@@
 & Validator
-& Wrong type validation on \texttt{i32} binary opertations \\
+& Wrong type validation on \texttt{i32} binary operations \\
 \hline
@@
 & Interpreter
-& Wrong intepretation of \texttt{i32.sub} instruction \\
+& Wrong interpretation of \texttt{i32.sub} instruction \\
 \hline
@@
 & Interpreter
-& Wrong intepretation of \texttt{i32.lt\_u} instruction when operands are equal \\
+& Wrong interpretation of \texttt{i32.lt\_u} instruction when operands are equal \\
 \hline
@@
 & Interpreter
-& Wrong intepretation of \texttt{i32.shr\_u} instruction \\
+& Wrong interpretation of \texttt{i32.shr\_u} instruction \\
 \hline
@@
 %
-The language is relatively simple, in esence
+The language is relatively simple, in essence
 %
@@
 \captionof{table}{\label{table:ifc:tests} Mean tests until first failure and effect size
-  comparison between \fuzzchick and \mutagen when considering only succesfull
+  comparison between \fuzzchick and \mutagen when considering only successful
   runs.
@@
 %
-In consequence, we believe this heuristic is effective at automatically tunning
+In consequence, we believe this heuristic is effective at automatically tuning
 this internal parameter of our tool, especially when the user is unsure about
@@
 %
-Finding where the best tradeoff lies in cases where the heuristics cause mixed
+Finding where the best trade-off lies in cases where the heuristics cause mixed
 results like this is another interesting direction to aim our future work.
@@
     across 15 different bugs injected into \textit{haskell-wasm}. }
-\vspace{-10pt}
+\vspace{-5pt}
 \end{figure}
@@
 It is easy to conclude that \mutagen executes tests several times slower than
-\quickcheck --- roughly 20x and 75x slower when testing \textit{prop\_validator}
-and \textit{prop\_interpreter\_i32}, respectively.
+\quickcheck --- roughly 20x and 75x slower when testing \texttt{prop\_validator}
+and \texttt{prop\_interpreter\_i32}, respectively.
 %
@@
 used in tandem with existing off-the-shelf low-level fuzzers to find bugs in
-massively used programs.
+heavily used programs.
 
@@
 tool that uses method specification predicates to automatically generate all
-nonisomorphic test cases up to a given small size.
+non-isomorphic test cases up to a given small size.
 
@@
 Moreover, we show how our tool can be applied in a real-world testing scenario,
-where it quickly discovers several planted and existent previously unknown bugs.
+where it quickly discovers several planted and some previously unknown bugs.
 

### 2021-04-16 papers/oopsla-21/main.tex
NIT

@@
 evaluation to automatically prune the search space by detecting unevaluated
-subexpressions using lazy evaluation.
+subexpressions.
 %

### 2021-04-16 papers/oopsla-21/main.tex
Massaging the intro

@@
 
-Instead of continuously generating random invalid test cases from scratch,
+%Instead of continuously generating random invalid test cases from scratch,
 \fuzzchick keeps a queue of interesting previously executed test cases that can
@@
 mutated data, as mutations are applied directly at the data type level, e.g.,
-the program's AST in case of generating code.
-
-\fuzzchick relies on execution traces to distinguish which executed test cases
-were interesting and are therefore worth mutating --- a popular technique known
-as \emph{coverage-guided fuzzing} and popularized by tools like \emph{AFL}
+the program's AST in case of generating code. Internally, \fuzzchick relies on
+execution traces to distinguish which executed test cases were interesting and
+are therefore worth mutating --- a popular technique known as
+\emph{coverage-guided fuzzing} and popularized by tools like \emph{AFL}
 \cite{afl}.
@@
 produce new execution traces --- any other test case is simply thrown away.
+%
+The techniques behind \fuzzchick are not meant to replace sophisticated,
+manually written random generators, but rather providing an adequate testing
+solution for early stages of development.
+%, and while a better test suite using manually-written generators is
+%still under construction.
 
@@
 
-Of course, this technique is not meant to replace smart manually-written
-generators.
-%
-In turn, it provides an adequate solution to be used in early stages of
-development, and while a better test suite using manually-written generators is
-still under construction.
 %
@@
 %
-But \fuzzchick suffers several disadvantages.%
-\footnote{Some of these were already identified by the authors of \fuzzchick in
-  their original work.}
+The approach taken by \fuzzchick leaves considerable room for improvements in
+many aspects\footnote{Some of these were already identified by the authors of
+  \fuzzchick in their original work.}:
+%But \fuzzchick suffers several disadvantages.%
 %
@@
 %% Our solution
-In this work we introduce \mutagen, a PBT tool implemented in Haskell that
-builds upon the basic ideas behind \fuzzchick, i.e., code instrumentation and
-type-preserving mutations (see Section \ref{sec:mutagen}).
+In this work, we introduce \mutagen, a PBT tool implemented %in Haskell
+that also builds upon code instrumentation and type-preserving mutations (see
+Section \ref{sec:mutagen}).
+%
+% the basic ideas behind \fuzzchick, i.e.,
 %
@@
 %
-This approach has two inherent advantages when compared to \fuzzchick.
+This approach has two inherent advantages. %when compared to \fuzzchick.
 %
@@
 Section \ref{sec:casestudies}): the IFC stack machine used by
-\citeauthor{lampropoulos2019coverage} in their original work, and a WebAssembly
+\citeauthor{lampropoulos2019coverage} in their original work of \fuzzchick, and a WebAssembly
 engine written in Haskell.

### 2021-04-16 papers/oopsla-21/main.tex
NIT

@@
 mutated data, as mutations are applied directly at the data type level, e.g.,
-the program's AST in case of generating code. Internally, \fuzzchick relies on
-execution traces to distinguish which executed test cases were interesting and
-are therefore worth mutating --- a popular technique known as
-\emph{coverage-guided fuzzing} and popularized by tools like \emph{AFL}
-\cite{afl}.
+the program's AST in case of generating code.
+%
+Internally, \fuzzchick relies on execution traces to distinguish which executed
+test cases were interesting and are therefore worth mutating --- a popular
+technique known as \emph{coverage-guided fuzzing} and popularized by tools like
+\emph{AFL} \cite{afl}.
 %
@@
 The techniques behind \fuzzchick are not meant to replace sophisticated,
-manually written random generators, but rather providing an adequate testing
+manually written random generators, but rather to provide an adequate testing
 solution for early stages of development.
@@
 %
-The approach taken by \fuzzchick leaves considerable room for improvements in
-many aspects\footnote{Some of these were already identified by the authors of
-  \fuzzchick in their original work.}:
+The approach taken by \fuzzchick leaves considerable room for
+improvements in several aspects.%
+\footnote{Some of these were already identified by the authors of \fuzzchick in
+  their original work.}
 %But \fuzzchick suffers several disadvantages.%
@@
 %% Our solution
-In this work, we introduce \mutagen, a PBT tool implemented %in Haskell
+In this work, we introduce \mutagen, a PBT tool %implemented in Haskell
 that also builds upon code instrumentation and type-preserving mutations (see

### 2021-04-16 papers/oopsla-21/main.tex
Improved sentence

@@
 %
-The techniques behind \fuzzchick are not meant to replace sophisticated,
-manually written random generators, but rather to provide an adequate testing
-solution for early stages of development.
+Worth mentioning, the techniques behind \fuzzchick are not meant to replace
+sophisticated, manually written random generators, but rather to provide an
+adequate testing solution for early stages of development.
 %, and while a better test suite using manually-written generators is

### 2021-04-17 papers/oopsla-21/main.tex
Moved paragraph to section 2. Fixed figures.

@@
 %
-Moreover, we replicated the Information-Flow Control (IFC) stack machine case
-study bundled with \fuzzchick and observed a surprising limitation:
-%
-after repeating each experiment 30 times, \fuzzchick could only find 5 out of
-the 20 injected bugs with a 100\% efficacy.
-%
-%% the hardest one being found only around 13\% of the time after an hour of
-%% testing.
-%
-Despite this, \citeauthor{lampropoulos2019coverage} show that \fuzzchick is far
-better than using normal random testing coupled with na\"ive random generators
-by comparing the mean-time-to-failure (MTTF) against \emph{QuickChick}.
-%
-However, we believe that a more meticulous evaluation ought to take failure to
-find a counterexample as an important metric when comparing PBT tools, not just
-mean time to failure --- when one is found.
+%% Moreover, we replicated the Information-Flow Control (IFC) stack machine case
+%% study bundled with \fuzzchick and observed a surprising limitation:
+%% %
+%% after repeating each experiment 30 times, \fuzzchick could only find 5 out of
+%% the 20 injected bugs with a 100\% efficacy.
+%% %
+%% %% the hardest one being found only around 13\% of the time after an hour of
+%% %% testing.
+%% %
+%% Despite this, \citeauthor{lampropoulos2019coverage} show that \fuzzchick is far
+%% better than using normal random testing coupled with na\"ive random generators
+%% by comparing the mean-time-to-failure (MTTF) against \emph{QuickChick}.
+%% %
+%% However, we believe that a more meticulous evaluation ought to take failure to
+%% find a counterexample as an important metric when comparing PBT tools, not just
+%% mean time to failure --- when one is found.
 
@@
 still using a mostly automated workflow.
-
+%
 However, the authors acknowledge that certain parts of its implementation have
@@
 %
-In tandem with the lack of efficacy observed when replicating the evaluation of
-the IFC stack machine case study, these observations led us to consider three
-main aspects of \fuzzchick that can be improved upon --- and that constitute the
-main goal of this work.
+Moreover, we replicated the Information-Flow Control (IFC) stack machine case
+study bundled with \fuzzchick and observed a surprising limitation:
+%
+\emph{after repeating each experiment 30 times, \fuzzchick could only find 5 out
+  of the 20 injected bugs with a 100\% efficacy, the hardest one being found
+  only around 13\% of the time after an hour of testing.}
+%
+Despite this, the authors show that \fuzzchick outperforms normal random testing
+coupled with na\"ive random generators by comparing the mean-time-to-failure
+(MTTF) against \emph{QuickChick}.
+%
+In turn, we believe that a more meticulous evaluation ought to take failure to find a
+counterexample as an important metric when comparing PBT tools, not just mean
+time to failure --- when one is found.
+%
+%% In tandem with the lack of efficacy observed when replicating the evaluation of
+%% the IFC stack machine case study,
+
+All these observations led us to consider three main aspects of \fuzzchick that
+can be improved upon --- and that constitute the main goal of this work.
 %
@@
 \end{multicols}
-\vspace{-20pt}
+\vspace{-10pt}
 \end{figure}
@@
 
-\begin{figure}[b]
+\begin{figure}[t]
 \vspace{-15pt}
@@
 
-\begin{figure}[t]
+\begin{figure}[b]
 \vspace{-15pt}
@@
   \textit{haskell-wasm}.}
-\vspace{-25pt}
+\vspace{-15pt}
 \end{table}
@@
 
-\begin{table}[b]
+\begin{table}[t]
 \footnotesize
@@
     across 15 different bugs injected into \textit{haskell-wasm}. }
-\vspace{-5pt}
+%% \vspace{-5pt}
 \end{figure}
@@
 \footnotesize
-\vspace{-5pt}
+%% \vspace{-5pt}
 \begin{tabular}{|l|c|c|c|c|}
@@
   each tool on the WebAssembly case study.}
-\vspace{-20pt}
+\vspace{-10pt}
 \end{table}

### 2021-04-17 papers/oopsla-21/main.tex
Introduce IFC earlier

@@
 To validate our ideas, we use two main case studies (described in detail in
-Section \ref{sec:casestudies}): the IFC stack machine used by
-\citeauthor{lampropoulos2019coverage} in their original work of \fuzzchick, and a WebAssembly
-engine written in Haskell.
+Section \ref{sec:casestudies}): the Information-Flow Control (IFC) stack machine
+used by \citeauthor{lampropoulos2019coverage} in their original work of
+\fuzzchick, and a WebAssembly engine written in Haskell.
 %
@@
 %
-Moreover, we replicated the Information-Flow Control (IFC) stack machine case
-study bundled with \fuzzchick and observed a surprising limitation:
+%% Moreover, we replicated the Information-Flow Control (IFC) stack machine case
+Moreover, we replicated the IFC stack machine case study bundled with \fuzzchick
+and observed a surprising limitation:
 %

### 2021-04-17 papers/oopsla-21/main.tex
NIT

@@
 %% Our solution
-In this work, we introduce \mutagen, a PBT tool %implemented in Haskell
+In this work, we introduce \mutagen, a PBT tool implemented in Haskell
 that also builds upon code instrumentation and type-preserving mutations (see
@@
 used by \citeauthor{lampropoulos2019coverage} in their original work of
-\fuzzchick, and a WebAssembly engine written in Haskell.
+\fuzzchick, and an existing WebAssembly engine of industrial strenght written in
+Haskell.
 %

### 2021-06-14 papers/oopsla-21/rebuttal.md
Create rebuttal.md

Overview
========

Changes
=======

Detailed comments
=================

Reviewer #1
-----------

How did you decide on what bugs to inject? How concerned are you about their realism? For the IFC example, there was a systematic strategy; what about for WASM?

Replied below under "Reviewer #2".

Reviewer #2
-----------

given the absence of a direct comparison against FuzzChick, how do you make the case that your technique is an improvement over the state of the art?

The IFC stack machine case study shows a direct comparison between FuzzChick and MUTAGEN, where our tool shows a noticeable improvement in terms of mean-time-to-failure, failure rate. We acknowledge that the introduction makes it unclear that we actually evaluate MUTAGEN against FuzzChick in one of the case studies but not in the other. This should be rephrased.

on a related note, Table 3 suggests that Mutagen needs to generate more tests until failure than FuzzChick in at least 7 of the 20 cases. Based on these results, and given the synthetic nature of the bugs, how can you argue that the Mutagen technique is really better?

Table 3 compares results only for the runs in which FuzzChick actually finds the injected bug --- a distinction that clearly benefits FuzzChick, since it shows a low failure rate in this case study, and was only able to find 5 out of 20 bugs with 100% failure rate, whereas MUTAGEN achieves 100% failure rate across all bugs.

Given this observation, comparing only the MTTF across all runs (bug-found or timeout) can skew the representation of the performance of both tools since timed-out executions inflate the overall calculation. This is clear in Fig. 4, where FuzzChick's MTTF clearly jumps over 10^5ms as soon as the failure rate goes below 100%.

Moreover, we observed that FuzzChick tends to find bugs either relatively fast, or not at all within the 1-hour testing budget. In this light, the unstable behavior of FuzzChick could be alleviated by running it several times in a row using much shorter timeout (e.g. 30 seconds). Under this consideration, for the comparison against FuzzChick to be fair, our tool needs to be reliable (always find the bug) without being (possibly orders of magniture) slower on individual runs. Table 3 shows that this in fact the case, where the A12 measure indicates that MUTAGEN is not prone to need many more tests to find a bug that FuzzChick in most of the cases (except only for bugs #15 and #18, which FuzzChick is not capable of finding with 100% success rate).

Extra clarification: we included mean-tests-to-failure in Table 3 merely as an indication of the order of magniture. This metric, hovewer, is quite sensitive to outliers, which are to be expected due to both tools having to sometimes fall back to pure random generation, which normally follows a geometric distribution to find counterexamples. For this reason, we included the A12 measure between the two sets of measures, which is much more robust against outliers and should be considered over simply mean-tests-to-failure when comparing both approaches.

how were bugs introduced in the benchmarks? If you introduced these manually, what steps were taken to mitigate potential bias?

The IFC stack machine case study uses a systematic approach for introducing bugs, removing a single IFC policy check over a single machine instruction at a time. This was originally introduced by Lampropoulos et al. in their original FuzzChick article. We simply translated this full case study (along with the mechanism introducing the bugs) directly into Haskell.

The WebAssembly case study uses manually-inserted bugs. Such bugs are of two different kind:

* Removing an existing integrity check 
  (to weaken the WebAssembly type-system/validator)
 * Replacing the implementation of an operation with a similar compatible one, e.g., (+) -> (-) 
    (to simulate copy-and-paste induced bugs, inspired by one of the real bugs we found on haskell-wasm while developing this case study)

how much time do the techniques require to generate tests?

While we did not gather specific numbers, we don't expect MUTAGEN to be substantially slower than FuzzChick at generating (or mutating) test cases. In both cases, the complexity is at most linear in the size of the generated (or mutated) test case.

Moreover, Haskell's lazy evaluation poses a particular challenge: test cases will not be fully generated/mutated unless the testing property actually fully evaluates them --- test cases are constructed "on the fly" in tandem with the evaluation of the testing property. This is the case with most random testing tools in Haskell, including QuickCheck.

Reviewer #3
-----------

It seems to be the case that Mutagen is effectively a new driver for QuickCheck. Is that right? How does one invoke Mutagen? Can you drop it in to replace QuickCheck? How is the tracing/coverage achieved?

Yes, MUTAGEN aims to be a drop-in replacement for QuickCheck as much as possible. Our tool is invoked in the same way as QuickCheck, using an IO function as a driver that accepts a testing property.

The major difference lies in that MUTAGEN requires some extra constraints to be satisfied, e.g., property inputs should be Mutable (automatically derivable using Template Haskell).

Tracing/coverage is achieved by instrumenting the user code using a source-to-source plugin bundled with MUTAGEN. The plugin can be enabled in a per-module basis to add or remove tracing granularity if needed.

How does Mutagen's memory profile compare to QuickCheck's and FuzzChick's?

 Memory consumption is QuickCheck is almost negligible, as it is completely black-box and "almost" stateless.
FuzzChick needs extra memory for two purposes: the mutation queues (size/length depends on each particularcase study), and the tracing mechanism (a constant bitmap used to keep track of evaluated edges).
Like FuzzChick, MUTAGEN requires extra space for the mutation queues. In contrast, the "path" tracing mechanism used in MUTAGEN is expected to grow continuously as new executions are discovered. In our experiments, we never observed this to be a problem, and multi-day executions never exhausted the 32GB of memory available in our workstation.

If memory consumption becomes a problem for a certain case study, MUTAGEN can also fall to edge-based tracing, using a constant bitmap like AFL/FuzzChick, at the cost of less trace granularity and a less precise FIFO prioritization of test cases (the only metric available to prioritize them becomes the number of discovered edges.)

How much retesting do you do? Do you ever run the exact same test twice?

To the best of our knowledge, there are two situations when retesting can occur:

Randomly generating the same test case twice while on generation mode. This depends on the quality of the random generator used. 
Mutating a test case back to its immediate ancestor. This can happen when a mutation reverts a previous one. In principle, this could only happen once per mutation batch. We initially introduced a rather complex mechanism to avoid this by keeping track of previously mutated subexpressions and making mutants "inherit" this history from their ancestors, however, the added complexity did not show a big improvement, so we decided to keep the mutation mechanism simple.

In WASM, are you only testing single-module programs (p17)?

Yes. Testing multi-module interactions (a.k.a. WebAssembly scripts) using MUTAGEN would be feasible, although we foresee that a more statefull approach for mutating test cases would be helpful for facilitating interaction between mutated WebAssembly modules --- we are currently investigating this posibility.

You fix 25 as the number of mutations in 'no reset' mode. Is that higher or lower than typical?

In our experiments, this depended heavily on the case study. The WebAssembly case study requires almost no resets (since programs with new executions are found continously), whereas in the IFC stack machine case study we typically observed about 6/7 resets before finding a counterexample. Setting the number of random mutations to 25 was an arbitrary decision to evaluate what happened when this parameter is smaller than the one optimized by MUTAGEN.

Table 3 is just about the IFC example, right? What do these bug numbers correspond to? What's special about bugs 15 and 18 that makes FuzzChick so much better than Mutagen?

Yes. As mentioned above, FuzzChick is better at finding bugs #15 and #18 when it actually finds them. In particular, bug #18 is the second hardest one to find, where FuzzChick manages to find a counterexample in 1 out of 4 executions. In this light, our tool trades speed for effectiveness, finding the bugs reliably by using exhaustive mutations at the cost of (potentially) needing longer to find some pathological bugs.

### 2021-06-14 papers/oopsla-21/rebuttal.md
Pass rebuttal

@@
-Overview
-========
-
-Changes
-=======
-
-Detailed comments
-=================
-
-Reviewer #1
------------
-
-How did you decide on what bugs to inject? How concerned are you about their realism? For the IFC example, there was a systematic strategy; what about for WASM?
-
-Replied below under "Reviewer #2".
+We thank all the reviewers for their helpful and detailed feedback. We answer
+reviewers' questions below -- review #1, please refer to the answer given to
+reviewer #2.
 
@@
 
-given the absence of a direct comparison against FuzzChick, how do you make the case that your technique is an improvement over the state of the art?
+> given the absence of a direct comparison against FuzzChick, how do you make
+> the case that your technique is an improvement over the state of the art?
 
-The IFC stack machine case study shows a direct comparison between FuzzChick and MUTAGEN, where our tool shows a noticeable improvement in terms of mean-time-to-failure, failure rate.
-We acknowledge that the introduction makes it unclear that we actually evaluate MUTAGEN against FuzzChick in one of the case studies but not in the other.
-This should be rephrased.
+We clarify that IFC stack machine case study shows a direct comparison between
+FuzzChick and MUTAGEN, where our tool shows a noticeable improvement in terms of
+mean-time-to-failure, failure rate. We will update the introduction to reflect
+that explicitly.
 
-on a related note, Table 3 suggests that Mutagen needs to generate more tests until failure than FuzzChick in at least 7 of the 20 cases.
-Based on these results, and given the synthetic nature of the bugs, how can you argue that the Mutagen technique is really better?
+> on a related note, Table 3 suggests that Mutagen needs to generate more tests
+> until failure than FuzzChick in at least 7 of the 20 cases. Based on these
+> results, and given the synthetic nature of the bugs, how can you argue that
+> the Mutagen technique is really better?
 
-Table 3 compares results only for the runs in which FuzzChick actually finds the injected bug --- a distinction that clearly benefits FuzzChick, since it shows a low failure rate in this case study, and was only able to find 5 out of 20 bugs with 100% failure rate, whereas MUTAGEN achieves 100% failure rate across all bugs.
+We argue that we are fair when comparing against FuzzChick for the following
+reasons:
 
-Given this observation, comparing only the MTTF across all runs (bug-found or timeout) can skew the representation of the performance of both tools since timed-out executions inflate the overall calculation.
-This is clear in Fig.
-4, where FuzzChick's MTTF clearly jumps over 10^5ms as soon as the failure rate goes below 100%.
+- Table 3 compares results only for the runs in which *FuzzChick actually finds
+the injected bug* --- a distinction that benefits FuzzChick, since it shows a
+low failure rate in this case study, and was only able to find 5 out of 20 bugs
+with 100% failure rate, whereas MUTAGEN achieves 100% failure rate across all
+bugs.
 
-Moreover, we observed that FuzzChick tends to find bugs either relatively fast, or not at all within the 1-hour testing budget.
-In this light, the unstable behavior of FuzzChick could be alleviated by running it several times in a row using much shorter timeout (e.g.
-30 seconds).
-Under this consideration, for the comparison against FuzzChick to be fair, our tool needs to be reliable (always find the bug) without being (possibly orders of magniture) slower on individual runs.
-Table 3 shows that this in fact the case, where the A12 measure indicates that MUTAGEN is not prone to need many more tests to find a bug that FuzzChick in most of the cases (except only for bugs #15 and #18, which FuzzChick is not capable of finding with 100% success rate).
+- Given the above observation, comparing only the MTTF across all runs
+(bug-found or timeout) can skew the representation of the performance of both
+tools since timed-out executions inflate the overall calculation. This is clear
+in Fig. 4, where FuzzChick's MTTF clearly jumps over 10^5ms as soon as the
+failure rate goes below 100%.
 
-Extra clarification: we included mean-tests-to-failure in Table 3 merely as an indication of the order of magniture.
-This metric, hovewer, is quite sensitive to outliers, which are to be expected due to both tools having to sometimes fall back to pure random generation, which normally follows a geometric distribution to find counterexamples.
-For this reason, we included the A12 measure between the two sets of measures, which is much more robust against outliers and should be considered over simply mean-tests-to-failure when comparing both approaches.
+- We observed that FuzzChick tends to find bugs either relatively fast, or not
+at all within the 1-hour testing budget. The unstable behavior of FuzzChick
+could be alleviated by running it several times in a row using much shorter
+timeout (e.g. 30 seconds). Under this consideration, for the comparison against
+FuzzChick to be fair, our tool needs to be reliable (always find the bug)
+without being (possibly orders of magniture) slower on individual runs. Table 3
+shows that this in the case, where the *A12* measure indicates that MUTAGEN is
+not prone to need many more tests to find a bug that FuzzChick in most of the
+cases (except only for bugs #15 and #18, which FuzzChick is not capable of
+finding with 100% success rate).
 
-how were bugs introduced in the benchmarks? If you introduced these manually, what steps were taken to mitigate potential bias?
+- We included MTTF in Table 3 merely as an indication of the order of magniture.
+This metric, however, is quite sensitive to outliers. For this reason, we
+included the A12 measure between the two sets of measures, which is much more
+robust against outliers and should be considered over simply
+mean-tests-to-failure when comparing both approaches.
 
-The IFC stack machine case study uses a systematic approach for introducing bugs, removing a single IFC policy check over a single machine instruction at a time.
-This was originally introduced by Lampropoulos et al.
-in their original FuzzChick article.
-We simply translated this full case study (along with the mechanism introducing the bugs) directly into Haskell.
+> how were bugs introduced in the benchmarks? If you introduced these manually,
+> what steps were taken to mitigate potential bias?
+
+In the FuzzChick article, the IFC stack machine case study uses a systematic
+approach for introducing bugs. We simply translate (implement) this *full case
+study* directly into Haskell along with the mechanism introducing the bugs.
 
@@
 
-how much time do the techniques require to generate tests?
+> how much time do the techniques require to generate tests?
 
-While we did not gather specific numbers, we don't expect MUTAGEN to be substantially slower than FuzzChick at generating (or mutating) test cases.
-In both cases, the complexity is at most linear in the size of the generated (or mutated) test case.
-
-Moreover, Haskell's lazy evaluation poses a particular challenge: test cases will not be fully generated/mutated unless the testing property actually fully evaluates them --- test cases are constructed "on the fly" in tandem with the evaluation of the testing property.
-This is the case with most random testing tools in Haskell, including QuickCheck.
+We expect MUTAGEN to be not substantially slower than FuzzChick at generating
+(or mutating) test cases. In both cases, the complexity is at most linear in the
+size of the generated (or mutated) test case. Futhermore, test cases will not be
+fully generated/mutated unless the testing property actually fully evaluates
+them --- test cases are constructed "on the fly" in tandem with the evaluation
+of the testing property. This is the case with most random testing tools in
+Haskell, including QuickCheck.
 
@@
 
-It seems to be the case that Mutagen is effectively a new driver for QuickCheck.
-Is that right?
-How does one invoke Mutagen?
-Can you drop it in to replace QuickCheck?
-How is the tracing/coverage achieved?
+> It seems to be the case that Mutagen is effectively a new driver for
+> QuickCheck. Is that right? How does one invoke Mutagen? Can you drop it in to
+> replace QuickCheck? How is the tracing/coverage achieved?
 
-Yes, MUTAGEN aims to be a drop-in replacement for QuickCheck as much as possible. Our tool is invoked in the same way as QuickCheck, using an IO function as a driver that accepts a testing property.
+Yes, MUTAGEN aims to be a drop-in replacement for QuickCheck as much as
+possible. Our tool is invoked in the same way as QuickCheck, using an IO
+function as a driver that accepts a testing property.
 
-The major difference lies in that MUTAGEN requires some extra constraints to be satisfied, e.g., property inputs should be Mutable (automatically derivable using Template Haskell).
+The major difference lies in that MUTAGEN requires some extra constraints to be
+satisfied, e.g., property inputs should be Mutable (automatically derivable
+using Template Haskell).
 
-Tracing/coverage is achieved by instrumenting the user code using a source-to-source plugin bundled with MUTAGEN.
-The plugin can be enabled in a per-module basis to add or remove tracing granularity if needed.
+Tracing/coverage is achieved by instrumenting the user code using our own
+source-to-source plugin to the Haskell compiler. The plugin can be enabled in a
+per-module basis to add or remove tracing granularity if needed.
 
-How does Mutagen's memory profile compare to QuickCheck's and FuzzChick's?
+> How does Mutagen's memory profile compare to QuickCheck's and FuzzChick's?
 
- Memory consumption is QuickCheck is almost negligible, as it is completely black-box and "almost" stateless.
-FuzzChick needs extra memory for two purposes: the mutation queues (size/length depends on each particularcase study), and the tracing mechanism (a constant bitmap used to keep track of evaluated edges).
-Like FuzzChick, MUTAGEN requires extra space for the mutation queues.
-In contrast, the "path" tracing mechanism used in MUTAGEN is expected to grow continuously as new executions are discovered.
-In our experiments, we never observed this to be a problem, and multi-day executions never exhausted the 32GB of memory available in our workstation.
+Memory consumption is QuickCheck is almost negligible, as it is completely
+black-box and "almost" stateless. FuzzChick, however, needs extra memory for two
+purposes, e.g., the mutation queues. Like FuzzChick, MUTAGEN requires extra
+space for the mutation queues. The "path" tracing mechanism used in MUTAGEN is
+expected to grow continuously as new executions are discovered. In our
+experiments, we never observed this to be a problem, and multi-day executions
+never exhausted the 32GB of memory available in our workstation.
 
-If memory consumption becomes a problem for a certain case study, MUTAGEN can also fall to edge-based tracing, using a constant bitmap like AFL/FuzzChick, at the cost of less trace granularity and a less precise FIFO prioritization of test cases (the only metric available to prioritize them becomes the number of discovered edges.)
+If memory consumption becomes a problem, MUTAGEN can fall to edge-based tracing,
+i.e., using a constant bitmap like AFL/FuzzChick, at the cost of less trace
+granularity and a less precise FIFO prioritization of test cases (the only
+metric available to prioritize them becomes the number of discovered edges.)
 
-How much retesting do you do? Do you ever run the exact same test twice?
+> How much retesting do you do? Do you ever run the exact same test twice?
 
@@
 
-Randomly generating the same test case twice while on generation mode. This depends on the quality of the random generator used. 
-Mutating a test case back to its immediate ancestor.
-This can happen when a mutation reverts a previous one.
-In principle, this could only happen once per mutation batch.
-We initially introduced a rather complex mechanism to avoid this by keeping track of previously mutated subexpressions and making mutants "inherit" this history from their ancestors, however, the added complexity did not show a big improvement, so we decided to keep the mutation mechanism simple.
+- Randomly generating the same test case twice while on generation mode. This
+depends on the quality of the random generator used.
 
-In WASM, are you only testing single-module programs (p17)?
+- Mutating a test case back to its immediate ancestor. This can happen when a
+mutation reverts a previous one. We initially introduced a complex mechanism to
+avoid this, however, it did not show a big improvement, so we decided to keep
+the mutation mechanism simple.
 
-Yes.
-Testing multi-module interactions (a.k.a.
-WebAssembly scripts) using MUTAGEN would be feasible, although we foresee that a more statefull approach for mutating test cases would be helpful for facilitating interaction between mutated WebAssembly modules --- we are currently investigating this posibility.
+> In WASM, are you only testing single-module programs (p17)?
 
-You fix 25 as the number of mutations in 'no reset' mode. Is that higher or lower than typical?
+Yes. Testing multi-module interactions (a.k.a. WebAssembly scripts) using
+MUTAGEN would be feasible, although we foresee that a more statefull approach
+for mutating test cases would be helpful for facilitating interaction between
+mutated WebAssembly modules --- we are currently investigating this
+possibility.
 
-In our experiments, this depended heavily on the case study.
-The WebAssembly case study requires almost no resets (since programs with new executions are found continously), whereas in the IFC stack machine case study we typically observed about 6/7 resets before finding a counterexample.
-Setting the number of random mutations to 25 was an arbitrary decision to evaluate what happened when this parameter is smaller than the one optimized by MUTAGEN.
+> You fix 25 as the number of mutations in 'no reset' mode. Is that higher or
+> lower than typical?
 
-Table 3 is just about the IFC example, right? What do these bug numbers correspond to? What's special about bugs 15 and 18 that makes FuzzChick so much better than Mutagen?
+In our experiments, this depended heavily on the case study. The WebAssembly
+case study requires almost no resets (since programs with new executions are
+found continously), whereas in the IFC stack machine case study we typically
+observed about 6/7 resets before finding a counterexample. Setting the number of
+random mutations to 25 was an arbitrary decision to evaluate what happened when
+this parameter is smaller than the one optimized by MUTAGEN.
 
-Yes.
-As mentioned above, FuzzChick is better at finding bugs #15 and #18 when it actually finds them.
-In particular, bug #18 is the second hardest one to find, where FuzzChick manages to find a counterexample in 1 out of 4 executions.
-In this light, our tool trades speed for effectiveness, finding the bugs reliably by using exhaustive mutations at the cost of (potentially) needing longer to find some pathological bugs.
+> Table 3 is just about the IFC example, right? What do these bug numbers
+> correspond to? What's special about bugs 15 and 18 that makes FuzzChick so
+> much better than Mutagen?
+
+Yes. As mentioned above, FuzzChick is better at finding bugs #15 and #18 when
+*it actually finds them*. In particular, bug #18 is the second hardest one to
+find, where FuzzChick manages to find a counterexample in 1 out of 4 executions.
+In this light, our tool trades speed for effectiveness, finding the bugs
+reliably by using exhaustive mutations at the cost of (potentially) needing
+longer to find some pathological bugs.

### 2021-06-15 papers/oopsla-21/rebuttal.md
Pass over rebutal. 165 words left.

@@
 We thank all the reviewers for their helpful and detailed feedback. We answer
-reviewers' questions below -- review #1, please refer to the answer given to
+reviewers' questions below -- reviewer #1, please refer to the answer given to
 reviewer #2.
+
+For the final version, we propose to use the extra available space to clarify
+the questions and concerns of the reviewers to the furthest extent. Moreover, we
+will expand on implementation details and usage examples of MUTAGEN.
 
@@
 low failure rate in this case study, and was only able to find 5 out of 20 bugs
-with 100% failure rate, whereas MUTAGEN achieves 100% failure rate across all
+with a 100% failure rate, whereas MUTAGEN achieves a 100% failure rate across all
 bugs.
@@
 FuzzChick to be fair, our tool needs to be reliable (always find the bug)
-without being (possibly orders of magniture) slower on individual runs. Table 3
+without being (possibly orders of magnitude) slower on individual runs. Table 3
 shows that this in the case, where the *A12* measure indicates that MUTAGEN is
@@
 cases (except only for bugs #15 and #18, which FuzzChick is not capable of
-finding with 100% success rate).
+finding with a 100% success rate).
 
-- We included MTTF in Table 3 merely as an indication of the order of magniture.
-This metric, however, is quite sensitive to outliers. For this reason, we
-included the A12 measure between the two sets of measures, which is much more
-robust against outliers and should be considered over simply
+- We included mean-tests-to-failure in Table 3 merely as an indication of the
+order of magnitude. This metric, however, is quite sensitive to outliers. For
+this reason, we included the A12 measure between the two sets of measures, which
+is much more robust against outliers and should be considered over simply
 mean-tests-to-failure when comparing both approaches.
@@
 
-The WebAssembly case study uses manually-inserted bugs. Such bugs are of two different kind:
+The WebAssembly case study uses manually inserted bugs. Such bugs are of two 
+different kinds:
 
-* Removing an existing integrity check 
-  (to weaken the WebAssembly type-system/validator)
- * Replacing the implementation of an operation with a similar compatible one, e.g., (+) -> (-) 
-    (to simulate copy-and-paste induced bugs, inspired by one of the real bugs we found on haskell-wasm while developing this case study)
+* Removing an existing integrity check (to weaken the WebAssembly 
+   type-system/validator)
+ * Replacing the implementation of an operation with a similar compatible one,
+   e.g., (+) -> (-) (to simulate copy-and-paste induced bugs, inspired by one
+   of the real bugs we found on `haskell-wasm` while developing this case study)
 
@@
 (or mutating) test cases. In both cases, the complexity is at most linear in the
-size of the generated (or mutated) test case. Futhermore, test cases will not be
-fully generated/mutated unless the testing property actually fully evaluates
-them --- test cases are constructed "on the fly" in tandem with the evaluation
-of the testing property. This is the case with most random testing tools in
-Haskell, including QuickCheck.
+size of the generated (or mutated) test case.
+
+However, test cases will not be fully generated/mutated unless the testing 
+property actually fully evaluates them --- test cases are constructed "on the fly"
+in tandem with the evaluation of the testing property. This is the case with most
+random testing tools in Haskell, including QuickCheck. This lazy behavior greatly
+difficults to measure generation/mutation overhead in a precise manner.
 
@@
 
-Memory consumption is QuickCheck is almost negligible, as it is completely
+Memory consumption in QuickCheck is almost negligible, as it is completely
 black-box and "almost" stateless. FuzzChick, however, needs extra memory for two
-purposes, e.g., the mutation queues. Like FuzzChick, MUTAGEN requires extra
-space for the mutation queues. The "path" tracing mechanism used in MUTAGEN is
-expected to grow continuously as new executions are discovered. In our
-experiments, we never observed this to be a problem, and multi-day executions
+purposes: the mutation queues and the edge-based tracing mechanism (a constant bitmap).
+Like FuzzChick, MUTAGEN requires extra space for the mutation queues, whereas the
+more precise path-based tracing mechanism implemented in MUTAGEN uses a trie data
+structure that is expected to grow continuously as new executions are discovered.
+In our experiments, we never observed this to be a problem, and multi-day executions
 never exhausted the 32GB of memory available in our workstation.
 
-If memory consumption becomes a problem, MUTAGEN can fall to edge-based tracing,
-i.e., using a constant bitmap like AFL/FuzzChick, at the cost of less trace
-granularity and a less precise FIFO prioritization of test cases (the only
+If memory consumption becomes a problem, MUTAGEN can be used with to edge-based
+tracing, i.e., using a constant bitmap like AFL/FuzzChick, at the cost of less
+trace granularity and a less precise FIFO prioritization of test cases (the only
 metric available to prioritize them becomes the number of discovered edges.)
@@
 Yes. Testing multi-module interactions (a.k.a. WebAssembly scripts) using
-MUTAGEN would be feasible, although we foresee that a more statefull approach
-for mutating test cases would be helpful for facilitating interaction between
-mutated WebAssembly modules --- we are currently investigating this
-possibility.
+MUTAGEN would be feasible, although we foresee that a more stateful approach
+for mutating test cases would help to facilitate interaction between mutated 
+WebAssembly modules --- we are currently investigating this possibility.
 
@@
 case study requires almost no resets (since programs with new executions are
-found continously), whereas in the IFC stack machine case study we typically
+found continuously), whereas in the IFC stack machine case study we typically
 observed about 6/7 resets before finding a counterexample. Setting the number of
 random mutations to 25 was an arbitrary decision to evaluate what happened when
-this parameter is smaller than the one optimized by MUTAGEN.
+this parameter is *smaller* than the one optimized by MUTAGEN.
 
@@
 
-Yes. As mentioned above, FuzzChick is better at finding bugs #15 and #18 when
-*it actually finds them*. In particular, bug #18 is the second hardest one to
-find, where FuzzChick manages to find a counterexample in 1 out of 4 executions.
-In this light, our tool trades speed for effectiveness, finding the bugs
-reliably by using exhaustive mutations at the cost of (potentially) needing
-longer to find some pathological bugs.
+Yes. Bug numbers are systematically assigned. As mentioned above, FuzzChick is
+better at finding bugs #15 and #18 when *it actually finds them*. In particular,
+bug #18 is the second hardest one to find, where FuzzChick manages to find a
+counterexample in only about 1 out of 4 executions. In this light, our tool trades
+speed for effectiveness, finding the bugs reliably by using exhaustive mutations
+at the cost of (potentially) needing longer to find some pathological bugs.

### 2021-06-15 papers/oopsla-21/rebuttal.md
Pass over rebuttal

@@
 For the final version, we propose to use the extra available space to clarify
-the questions and concerns of the reviewers to the furthest extent. Moreover, we
-will expand on implementation details and usage examples of MUTAGEN.
+the questions and concerns of the reviewers to the furthest extent. We will also
+expand on implementation details and usage examples of MUTAGEN.
 
@@
 not prone to need many more tests to find a bug that FuzzChick in most of the
-cases (except only for bugs #15 and #18, which FuzzChick is not capable of
-finding with a 100% success rate).
+cases (except only for bugs #15 and #18, where FuzzChick finds it more quickly
+but not in all runs).
 
@@
 
-If memory consumption becomes a problem, MUTAGEN can be used with to edge-based
+If memory consumption becomes a problem, MUTAGEN can be used with edge-based
 tracing, i.e., using a constant bitmap like AFL/FuzzChick, at the cost of less
@@
 - Randomly generating the same test case twice while on generation mode. This
-depends on the quality of the random generator used.
+depends on the quality of the used random generators.
 

### 2021-08-10 papers/icse22/results/wasm/prop_interpreter/mutagen-base/prop_interpreter_i32.BUG_INTERPRETER_5.MUTAGEN_NO_INHERITANCE.csv.1.hs
Ported paper to ICSE22 format

@@ -1,0 +1,1 @@
+(2,1.4241114,Function {funcType = 0, localTypes = [I64,F32,F32], body = [GetLocal 0,GetLocal 0,I32Store16 (MemArg {offset = 0, align = 0}),I32Const 2,I32Load8S (MemArg {offset = 0, align = 0})]})

### 2021-08-10 papers/icse22/results/wasm/prop_interpreter/mutagen-no-fifo/prop_interpreter_i32.BUG_INTERPRETER_1.MUTAGEN_NO_FIFO.csv.10.hs

@@ -1,0 +1,1 @@
+(1,0.55339277,Function {funcType = 0, localTypes = [F32,I64,I64], body = [I32Const 2,I32Const 2,IBinOp BS32 ISub]})

### 2021-08-10 papers/icse22/results/wasm/prop_interpreter/mutagen-no-fifo/prop_interpreter_i32.BUG_INTERPRETER_2.MUTAGEN_NO_FIFO.csv.7.hs

@@ -1,0 +1,1 @@
+(2,-0.49816766,Function {funcType = 0, localTypes = [], body = [I32Const 1,I32Const 1,IRelOp BS32 ILtU]})

### 2021-08-10 papers/icse22/results/wasm/prop_interpreter/mutagen-no-fifo/prop_interpreter_i32.BUG_INTERPRETER_3.MUTAGEN_NO_FIFO.csv.8.hs

@@ -1,0 +1,1 @@
+(0,-2.5966632,Function {funcType = 0, localTypes = [I64], body = [CurrentMemory,GetLocal 0,IBinOp BS32 IShrU]})

### 2021-08-10 papers/icse22/results/wasm/prop_interpreter/mutagen-no-reset/prop_interpreter_i32.BUG_INTERPRETER_1.MUTAGEN_NO_RESET.csv.10.hs

@@ -1,0 +1,1 @@
+(0,-2.3528986,Function {funcType = 0, localTypes = [], body = [GetLocal 0,I32Const 0,GrowMemory,IBinOp BS32 ISub]})

### 2021-08-10 papers/icse22/results/wasm/prop_interpreter/mutagen-no-reset/prop_interpreter_i32.BUG_INTERPRETER_4.MUTAGEN_NO_RESET.csv.6.hs

@@ -1,0 +1,1 @@
+(0,-0.7305002,Function {funcType = 0, localTypes = [I32], body = [GetLocal 2,Return,TeeLocal 2,F32Const (-1.9906845),I32Const 1,I64Const 2,CurrentMemory,GetLocal 0,I64Load32S (MemArg {offset = 0, align = 0}),CurrentMemory,I32Const 0,Return]})

### 2021-08-10 papers/icse22/results/wasm/prop_interpreter/mutagen-no-reset/prop_interpreter_i32.BUG_INTERPRETER_5.MUTAGEN_NO_RESET.csv.7.hs

@@ -1,0 +1,1 @@
+(0,-1.0923313,Function {funcType = 0, localTypes = [F32], body = [GetLocal 1,ITruncFS BS32 BS32,I64Const 1,I64Store32 (MemArg {offset = 3, align = 0}),I32Const 2,I32Eqz,GrowMemory,I32Load (MemArg {offset = 2, align = 1}),F64Const (-0.7314185108772099),GetLocal 2,F32Const 0.55392754,GetLocal 0,Br 0]})

### 2021-08-10 papers/icse22/results/wasm/prop_validator/mutagen-base/prop_validator.BUG_VALIDATOR_1.MUTAGEN_NO_INHERITANCE.csv.27.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [I32Const 0,If {resultType = [], true = [Call 0], false = [F32Const (-8.636211),I32Const 28]}]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2021-08-10 papers/icse22/results/wasm/prop_validator/mutagen-base/prop_validator.BUG_VALIDATOR_1.MUTAGEN_NO_INHERITANCE.csv.29.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [I32Const 0,If {resultType = [], true = [], false = [I64Const 79]}]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2021-08-10 papers/icse22/results/wasm/prop_validator/mutagen-base/prop_validator.BUG_VALIDATOR_10.MUTAGEN_NO_INHERITANCE.csv.10.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [I32Const 0,I64ExtendSI32,I64Eqz,I32Load (MemArg {offset = 0, align = 0}),F32Const (-2.7480369),Loop {resultType = [], body = []},Return]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2021-08-10 papers/icse22/results/wasm/prop_validator/mutagen-base/prop_validator.BUG_VALIDATOR_7.MUTAGEN_NO_INHERITANCE.csv.12.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [I32Const 0,IBinOp BS32 IAdd,I32Const 0,Return]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2021-08-10 papers/icse22/results/wasm/prop_validator/mutagen-no-fifo/prop_validator.MUTAGEN_NO_FIFO.BUG_VALIDATOR_4.csv.22.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [], tables = [], mems = [Memory (Limit 1 Nothing),Memory (Limit 1 Nothing)], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2021-08-10 papers/icse22/results/wasm/prop_validator/mutagen-no-fifo/prop_validator.MUTAGEN_NO_FIFO.BUG_VALIDATOR_6.csv.22.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [I64Eqz,Return]},Function {funcType = 0, localTypes = [], body = []}], tables = [Table (TableType (Limit 0 Nothing) AnyFunc)], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2021-08-10 papers/icse22/results/wasm/prop_validator/mutagen-no-fifo/prop_validator.MUTAGEN_NO_FIFO.BUG_VALIDATOR_9.csv.28.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [Nop,I64Const 211,CallIndirect 0]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [Import {sourceModule = "SxPW\DEL\NUL\995545", name = "m&\153938\132752%vcb", desc = ImportTable (TableType (Limit 6 Nothing) AnyFunc)}], exports = []}

### 2021-08-10 papers/icse22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_1.MUTAGEN_NO_RESET.csv.12.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [I32Const 0,If {resultType = [], true = [Nop], false = [I32Const 0,I64Const 117,I32Const 28]}]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2021-08-10 papers/icse22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_1.MUTAGEN_NO_RESET.csv.15.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [I64Const 0,Loop {resultType = [I64,I64], body = [Return]},FConvertIS BS64 BS64,I32Const 9,Call 0,Loop {resultType = [], body = []},Block {resultType = [], body = []},Br 0]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2021-08-10 papers/icse22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_1.MUTAGEN_NO_RESET.csv.2.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [I32Const 0,If {resultType = [], true = [], false = [I64Const 121]}]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2021-08-10 papers/icse22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_1.MUTAGEN_NO_RESET.csv.23.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [F64Const 0.0,Block {resultType = [], body = []},Nop,Block {resultType = [I64,I64,I64,F64], body = [Return]},Return,IReinterpretF BS32,BrTable [] 0]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2021-08-10 papers/icse22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_10.MUTAGEN_NO_RESET.csv.28.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [I32Const 0,I32Load (MemArg {offset = 0, align = 0}),Nop,Block {resultType = [], body = []},Nop,Br 0]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2021-08-10 papers/icse22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_3.MUTAGEN_NO_RESET.csv.28.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [I64,I32,F64,F32,F64,I32,I32,I32,F32], results = [F64]},FuncType {params = [], results = [I64]}], functions = [Function {funcType = 0, localTypes = [], body = []}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2021-08-10 papers/icse22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_4.MUTAGEN_NO_RESET.csv.17.hs

@@ -1,0 +1,1 @@
+Module {types = [], functions = [], tables = [], mems = [Memory (Limit 8 Nothing),Memory (Limit 2 Nothing)], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2021-08-10 papers/icse22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_6.MUTAGEN_NO_RESET.csv.10.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [F32Const (-1.9060678),I64Eqz,Nop,Br 0]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2021-08-10 papers/icse22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_6.MUTAGEN_NO_RESET.csv.16.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [Nop,Nop,I64Eqz,Drop]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2021-08-10 papers/icse22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_7.MUTAGEN_NO_RESET.csv.4.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [F32Const 0.0,ITruncFU BS32 BS32,IBinOp BS32 IAdd,Drop]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2021-08-10 papers/icse22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_8.MUTAGEN_NO_RESET.csv.30.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [Br 0,F64Const 8.21381393881019,Drop,CurrentMemory,Return]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2021-08-18 papers/icse22/main.tex
Ported paper to ICSE22 format
Fixed figures and tables to somewhat fit now

--- 2021-04-17 papers/oopsla-21/main.tex
@@
 %% The first command in your LaTeX source must be the \documentclass command.
-\documentclass[acmsmall, anonymous]{acmart}
+\documentclass[sigconf, anonymous, review]{acmart}
 
@@
 % ----------------------------------------
-\vspace{-5pt}
+%% \vspace{-5pt}
 
@@
 
-\vspace{-5pt}
+%% \vspace{-5pt}
 \subsection{Property-Based Testing and \quickcheck}
@@
 
-\begin{wrapfigure}{r}{0.5\textwidth}
-\vspace{-10pt}
-\begin{algorithm}[H]
+\begin{algorithm}
   \SetInd{0em}{0.75em}
@@
 \end{algorithm}
-\vspace{-10pt}
-\end{wrapfigure}
 
@@
 instance Arbitrary a => Arbitrary (Tree a) where
-  arbitrary = sized gen
-    where
-      gen 0 = do { x <- arbitrary; return (Leaf x) }
-      gen n = oneof [ do {x <- arbitrary; return (Leaf x) }
-                    , do {l <- gen (n-1); x <- arbitrary; r <- gen (n-1); return (Branch l x r)} ]
+  arbitrary = sized gen where
+  gen 0 = Leaf <$> arbitrary
+  gen n = oneof [ Leaf <$> arbitrary
+                , Branch <$> gen (n-1) <*> arbitrary <*> gen (n-1) ]
 \end{verbatim}
@@
 
-\begin{figure}[t]
-\begin{multicols}{2}
-
-\begin{algorithm}[H]
+\begin{algorithm}
   \SetArgSty{textnormal}
@@
 
-\columnbreak
-
-\begin{algorithm}[H]
+\begin{algorithm}
   \SetArgSty{textnormal}
@@
 
-\end{multicols}
-\vspace{-15pt}
-\end{figure}
-
 \subsubsection{Type-preserving mutations}
@@
 
+%% \begin{figure}
+%% \vspace{-5pt}
+%% \begin{verbatim}
+%% mutate_tree :: (a -> Gen a) -> Tree a -> Gen (Tree a)
+%% mutate_tree mutate_a (Leaf x) =
+%%   oneof [ do { x' <- mutate_a x; return (Leaf x')}                       -- Mutate recursively
+%%         , do { l <- arbitrary; r <- arbitrary; return (Branch l x r) } ] -- Grow constructor
+%% mutate_tree mutate_a (Branch l x r) =
+%%   oneof [ return l                                                       -- Return subexpression
+%%         , return r                                                       -- Return subexpression
+%%         , return (Leaf x)                                                -- Shrink constructor
+%%         , do { l' <- mutate_tree l; return (Branch l' x r) }             -- Mutate recursively
+%%         , do { x' <- mutate_a    x; return (Branch l x' r) }             -- Mutate recursively
+%%         , do { r' <- mutate_tree r; return (Branch l x r') } ]           -- Mutate recursively
+%% \end{verbatim}
+%% \caption{\label{fig:fuzzchick:mutator}\fuzzchick mutator for the \texttt{Tree} data type. }
+%% \end{figure}
+
 \begin{figure}
+%% \vspace{-5pt}
 \begin{verbatim}
@@
 mutate_tree mutate_a (Leaf x) =
-  oneof [ do { x' <- mutate_a x; return (Leaf x')}                       -- Mutate recursively
-        , do { l <- arbitrary; r <- arbitrary; return (Branch l x r) } ] -- Grow constructor
+  oneof [ Leaf <$> mutate_a x
+        , Branch <$> arbitrary <*> return x <*> arbitrary ]
 mutate_tree mutate_a (Branch l x r) =
-  oneof [ return l                                                       -- Return subexpression
-        , return r                                                       -- Return subexpression
-        , return (Leaf x)                                                -- Shrink constructor
-        , do { l' <- mutate_tree l; return (Branch l' x r) }             -- Mutate recursively
-        , do { x' <- mutate_a    x; return (Branch l x' r) }             -- Mutate recursively
-        , do { r' <- mutate_tree r; return (Branch l x r') } ]           -- Mutate recursively
+  oneof [ return l
+        , return r
+        , Leaf   <$> return x
+        , Branch <$> mutate_tree l <*> return x   <*> return r
+        , Branch <$> return l      <*> mutate_a x <*> return r
+        , Branch <$> return l      <*> return x   <*> mutate_tree r
 \end{verbatim}
-\vspace{-5pt}
 \caption{\label{fig:fuzzchick:mutator}\fuzzchick mutator for the \texttt{Tree} data type. }
-\vspace{-15pt}
 \end{figure}
@@
 
+%% \begin{figure}
+%% \begin{verbatim}
+%% mutate (Leaf x)       = [ PURE (Branch def x def) ]   -- Swap constructor
+%% mutate (Branch l x r) = [ PURE l, PURE r              -- Return subexpression
+%%                         , PURE (Leaf x)               -- Swap constructor
+%%                         , PURE (Branch l x l)         -- Rearrange subexpressions
+%%                         , PURE (Branch r x r)         -- Rearrange subexpressions
+%%                         , PURE (Branch r x l) ]       -- Rearrange subexpressions
+%% \end{verbatim}
+%% \vspace{-5pt}
+%% \caption{\label{fig:mutagen:mutator}\mutagen mutator for the \texttt{Tree} data
+%%   type.}
+%% \vspace{-10pt}
+%% \end{figure}
+
 \begin{figure}
 \begin{verbatim}
-mutate (Leaf x)       = [ PURE (Branch def x def) ]   -- Swap constructor
-mutate (Branch l x r) = [ PURE l, PURE r              -- Return subexpression
-                        , PURE (Leaf x)               -- Swap constructor
-                        , PURE (Branch l x l)         -- Rearrange subexpressions
-                        , PURE (Branch r x r)         -- Rearrange subexpressions
-                        , PURE (Branch r x l) ]       -- Rearrange subexpressions
+mutate (Leaf x)       = [ PURE (Branch def x def) ]
+mutate (Branch l x r) = [ PURE l, PURE r
+                        , PURE (Leaf x)
+                        , PURE (Branch l x l)
+                        , PURE (Branch r x r)
+                        , PURE (Branch r x l) ]
 \end{verbatim}
-\vspace{-5pt}
+%% \vspace{-5pt}
 \caption{\label{fig:mutagen:mutator}\mutagen mutator for the \texttt{Tree} data
   type.}
-\vspace{-25pt}
+%% \vspace{-10pt}
 \end{figure}
@@
 positions (Leaf x)       = node [ (0, positions x) ]
-positions (Branch l x r) = node [ (0, positions l), (1, positions x), (2, positions r) ]
+positions (Branch l x r) = node [ (0, positions l)
+                                , (1, positions x)
+                                , (2, positions r) ]
 \end{verbatim}
@@
 inside []      x              = mutate x
-inside (0 : pos) (Leaf x)       = [ Leaf x'       | x' <- inside pos x ]
-inside (0 : pos) (Branch l x r) = [ Branch l' x r | l' <- inside pos l ]
-inside (1 : pos) (Branch l x r) = [ Branch l x' r | x' <- inside pos x ]
-inside (2 : pos) (Branch l x r) = [ Branch l x r' | r' <- inside pos r ]
+inside (0:pos) (Leaf x)       = [ Leaf x'       | x' <- inside pos x ]
+inside (0:pos) (Branch l x r) = [ Branch l' x r | l' <- inside pos l ]
+inside (1:pos) (Branch l x r) = [ Branch l x' r | x' <- inside pos x ]
+inside (2:pos) (Branch l x r) = [ Branch l x r' | r' <- inside pos r ]
 \end{verbatim}
@@
 
-\begin{wrapfigure}{r}{0.5\textwidth}
-\vspace{-10pt}
-\begin{algorithm}[H]
+\begin{algorithm}
   \SetArgSty{textnormal}
@@
 
-\begin{figure}[t]
-\vspace{-15pt}
-\begin{multicols}{2}
-
-\begin{algorithm}[H]
+\begin{algorithm}
   \SetArgSty{textnormal}
@@
 \caption{\label{algo:mutagen:pick}\mutagen Seed Selection}
-\vspace{1pt}
 \end{algorithm}
-\end{multicols}
-\vspace{-10pt}
-\end{figure}
 
@@
 \begin{figure}[t]
-\vspace{-15pt}
+%% \vspace{-15pt}
 \newcommand\stacked[2]{\underset{\textstyle #2\mathstrut}{#1}}
@@
   \stacked{{\color{blue} new=1\phantom{x}}}{{\color{red} depth=3}}
-  \qquad
-  \qquad
+  \end{equation}
+  \begin{equation}
+  \nonumber
+  %% \\
+  %% \qquad
+  %% \qquad
   \left[
@@
 \end{equation}
-\vspace{-10pt}
+%% \vspace{-10pt}
 \caption{\label{fig:mutagen:tracelog}Inserting two new execution traces into the
@@
 
-\begin{figure}[b]
-\vspace{-15pt}
-\begin{multicols}{2}
-
-\begin{algorithm}[H]
+\begin{algorithm}
   \SetArgSty{textnormal}
@@
   }
-  \vspace{5pt}
+  %% \vspace{5pt}
   \SetKwFunction{Pick}{Pick}
@@
 
-\end{multicols}
 %% \caption{\label{fig:heuristics} The two main heuristics implemented in \mutagen.
@@
 %%   point being made. }
-\vspace{-20pt}
-\end{figure}
 
@@
 & Interpreter
-& Wrong interpretation of \texttt{i32.lt\_u} instruction when operands are equal \\
+& Wrong interpretation of \texttt{i32.lt\_u} instruction \\
 \hline
@@
   \textit{haskell-wasm}.}
-\vspace{-15pt}
+%% \vspace{-15pt}
 \end{table}
@@
 \begin{table}[t]
-\footnotesize
+\scriptsize
 \begin{tabular}{|c|c|c|l|}
@@
 & Bug
-& Instance function invoker silently proceeds after arity mismatch \\
+& Instance function invoker silently ignores arity mismatch \\
 \hline
@@
 & Discrepancy
-& NaN reinterpretation does not follow reference implementation\\
+& NaN reinterpretation does not follow reference \\
 \hline
@@
   \textit{haskell-wasm}.}
-\vspace{-20pt}
+%% \vspace{-20pt}
 \end{table}
@@
 prop_interpreter_i32 fun i f =
-  prop_interpreter (FuncType { params = [I32, F32], result = [I32]}) fun [VI32 i, VF32 f]
+  prop_interpreter
+    (FuncType { params = [I32, F32], result = [I32] })
+    fun
+    [VI32 i, VF32 f]
 \end{verbatim}
@@
   \includegraphics[width=\linewidth]{tikz/ifc-fr.tikz}
-  \vspace{-20pt}
+  %% \vspace{-20pt}
   \caption{\label{fig:results:ifc} Comparison between \fuzzchick and \mutagen
     across 20 different bugs for the IFC stack machine. }
-\vspace{-10pt}
+%% \vspace{-10pt}
 \end{figure}
@@
 
-\begin{wrapfigure}{r}{0.5\textwidth}
-\centering
-\vspace{-20pt}
-\footnotesize
+Table \ref{table:ifc:tests} shows a comparison between the mean number of tests
+required by both tools to find the first failure \emph{when we only consider
+  successful runs.}
+%
+Additionally, as advised by \cite{arcuri2014hitchhiker} when comparing random
+testing tools, we computed the non-parametric Vargha-Delaney $A_{12}$ measure
+\cite{vargha2000critique}.
+%
+Intuitively, this measure encodes the probability of \mutagen to yield better
+results than \fuzzchick.
+%
+As it can be observed, using an exhaustive mutation approach can be not only
+more reliable, but also faster.
+%
+\mutagen is likely to find bugs faster than \fuzzchick in 14 of the 20 bugs
+injected into the abstract machine, while being substantially slower in only 2
+cases.
+
+\begin{figure}
+%% \begin{wrapfigure}{r}{0.5\textwidth}
+%% \centering
+%% \vspace{-20pt}
+\small
 \begin{tabular}{ccccc}
@@
   favoring \fuzzchick and \mutagen, respectively. }
-\vspace{-25pt}
-\end{wrapfigure}
-
-Table \ref{table:ifc:tests} shows a comparison between the mean number of tests
-required by both tools to find the first failure \emph{when we only consider
-  successful runs.}
-%
-Additionally, as advised by \cite{arcuri2014hitchhiker} when comparing random
-testing tools, we computed the non-parametric Vargha-Delaney $A_{12}$ measure
-\cite{vargha2000critique}.
-%
-Intuitively, this measure encodes the probability of \mutagen to yield better
-results than \fuzzchick.
-%
-As it can be observed, using an exhaustive mutation approach can be not only
-more reliable, but also faster.
-%
-\mutagen is likely to find bugs faster than \fuzzchick in 14 of the 20 bugs
-injected into the abstract machine, while being substantially slower in only 2
-cases.
+%% \vspace{-25pt}
+%% \end{wrapfigure}
+\end{figure}
 
@@
   \includegraphics[width=\linewidth]{tikz/wasm-legend.tikz}
-  \vspace{-20pt}
+  %% \vspace{-20pt}
   \caption{\label{fig:results:wasm} Comparison between \quickcheck and \mutagen
@@
 
-\begin{table}[H]
+\begin{table}
 \footnotesize
@@
   each tool on the WebAssembly case study.}
-\vspace{-10pt}
+%% \vspace{-10pt}
 \end{table}

### 2021-08-19 papers/icse22/main.tex
Changed Abstract

@@
 Automatically synthesized random data generators are an appealing option when
-using property-based testing to validate our systems.
-%
-Such random generators can be obtained using a variety of techniques that work
-by extracting static information from the codebase in order to produce random
-test cases.
-%
-For the most part, however, such techniques are not suitable for deriving
-generators producing random values satisfying complex invariants, which in turn
-complicates testing properties with sparse preconditions.
-
-%% \fuzzchick is a recent tool that borrows ideas from the fuzzing community to
-%% alleviate this limitation.
-%
-\fuzzchick is a coverage-guided, property-based testing (CGPT) tool that
+using property-based testing.
+%
+Such generators can be obtained using a variety of techniques that extract
+static information from the codebase in order to produce random test cases.
+%
+However, such techniques are not suitable for deriving generators producing
+random values satisfying complex invariants, which in turn complicates testing
+properties with sparse preconditions.
+
+Coverage-guided, property-based testing (CGPT) is a recent technique that
 alleviates this limitation by enhancing automatically synthesized generators
@@
 %
-Despite being enlightening, we observed several limitations in the
-implementation of \fuzzchick that reduce its applicability.
-%
-These limitations include a large reliance on randomness when mutating test
-cases, as well as poor scheduling in certain situations, among others.
-
-In this work we present \mutagen, a CGPT framework that builds upon the ideas
-behind \fuzzchick.
-%
-Instead of relying heavily on randomness, \mutagen uses an exhaustive approach
-to generate new mutants.
+Albeit effective, the initial CGPT approach is limited by factors like large
+reliance on randomness and poor mutants scheduling, reducing its applicability
+and bug-finding capacity.
+
+In this work we present \mutagen, a CGPT framework that overcomes these
+limitations by generating mutants in an \emph{exhaustive} manner.
 %
@@
 %
-Our results indicate that our approach is capable of outperforming the simpler
-one used by \fuzzchick, as well as finding bugs in real-world scenarios.
+Our results indicate that our approach is capable of outperforming existing CGPT
+tools, as well as finding bugs in real-world scenarios.
 \end{abstract}

### 2021-08-19 papers/icse22/main.tex
Pass over introduction

@@
 %% These commands are for a JOURNAL article.
-\acmConference[OOPSLA '21]{OOPSLA '21: ACM SIGPLAN conference on Systems,
-  Programming, Languages, and Applications}{October 17--22 2021}{Chicago, IL}
+\acmConference[ICSE 2022]{The 44th International Conference on Software
+  Engineering}{May 21–29, 2022}{Pittsburgh, PA, USA}
 
@@
 %
-Property-Based Testing (PBT) is a popular technique for finding bugs using
-executable properties that encode the expected behavior of our systems.
-%
-Exemplified by frameworks like Haskell's \quickcheck \cite{ClaessenH00} and the
-abundant ports of this tool in other programming languages
-\cite{hughes2003erlang, papadakis2011proper, bulwahn2012new,
-  denes2014quickchick}, the most common PBT approach is to instantiate the
-testing properties using randomly generated inputs.
-
-Users of \quickcheck (and PBT in general) are well aware that one of the biggest
-challenges while using these frameworks is to provide suitable random data
-generators needed to instantiate the testing properties.
+Popularized by \quickcheck \cite{ClaessenH00}, Random Property-Based Testing
+(RPBT) is a popular technique for finding bugs
+\cite{ClaessenH00,hughes2003erlang, papadakis2011proper, bulwahn2012new,
+  denes2014quickchick}.
+%
+This is achieved by using randomly generated inputs to instantiate executable
+properties encoding the expected behavior of the system under test.
+%
+%% Exemplified by frameworks like Haskell's \quickcheck \cite{ClaessenH00} and the
+%% Abundant ports of this tool in other programming languages
+%% \cite{hughes2003erlang, papadakis2011proper, bulwahn2012new,
+%%   denes2014quickchick}
+%
+%% The most common PBT approach is to instantiate the testing properties using
+%% randomly generated inputs.
+%
+However, users of \quickcheck (and RPBT in general) are well aware that one of
+the biggest challenges while using this technique is to provide suitable random
+data generators needed to instantiate the testing properties.
 %
@@
 In this scenario, most of the generated tests are simply thrown away before they
-can be used to test the main components of our system.
+can be used to test the main components of the system under test.
 
 %% FuzzChick: PBT + code instrumentation + type-preserving mutations
-\fuzzchick \cite{lampropoulos2019coverage} is a property-based testing framework
-for the Coq programming language that borrows ideas from the fuzzing community
-to generate highly structured values while using automatically derived
-generators.
-%
-Notably, this tool is implemented as an extension of \quickchick
-\cite{denes2014quickchick}, Coq's own reimplementation of \quickcheck.
-
-%Instead of continuously generating random invalid test cases from scratch,
-\fuzzchick keeps a queue of interesting previously executed test cases that can
-be mutated using high-level, type-preserving transformations to produce new
-ones.
+%% \fuzzchick \cite{lampropoulos2019coverage} is a property-based testing framework
+%% for the Coq programming language that
+%
+Coverage-Guided, Property-Based Testing (CGPT) \cite{lampropoulos2019coverage}
+is a technique that borrows ideas from the fuzzing community to generate highly
+structured values while still using automatically derived generators.
+%
+%% Notably, this tool is implemented as an extension of \quickchick
+%% \cite{denes2014quickchick}, Coq's own reimplementation of \quickcheck.
+%
+Instead of continuously generating random invalid test cases from scratch, CGPT
+works by keeping a queue of \emph{interesting} previously executed test cases
+that can be mutated to produce new ones.
+%
+Mutations are achieved using high-level, structure-preserving transformations
+that can be derived directly from the codebase.
 %
 Intuitively, mutating an interesting test case in a small way (at the data
-constructor level) has a much higher probability of producing a new interesting
-test case than generating a new one from scratch using a na\"ive generator.
-%
-In this light, \fuzzchick is likely to preserve the semantic structure of the
-mutated data, as mutations are applied directly at the data type level, e.g.,
-the program's AST in case of generating code.
-%
-Internally, \fuzzchick relies on execution traces to distinguish which executed
-test cases were interesting and are therefore worth mutating --- a popular
-technique known as \emph{coverage-guided fuzzing} and popularized by tools like
-\emph{AFL} \cite{afl}.
+constructor level) has a higher probability of producing a new interesting test
+case than generating a new one using a na\"ive generator.
+%
+In this light, CGPT is likely to preserve the semantic structure of the mutated
+data, as mutations are applied directly at the data type level, e.g., the
+program's AST in case of generating code.
+%
+Moreover, CGPT relies on execution traces to distinguish which executed test
+cases were interesting and are therefore worth mutating --- a popular technique
+known as \emph{coverage-guided fuzzing} and popularized by tools like \emph{AFL}
+\cite{afl}.
 %
@@
 %
-Worth mentioning, the techniques behind \fuzzchick are not meant to replace
-sophisticated, manually written random generators, but rather to provide an
-adequate testing solution for early stages of development.
+Worth mentioning, this technique is not meant to replace sophisticated, manually
+written random generators, but rather to provide an adequate testing solution
+for early stages of development.
 %, and while a better test suite using manually-written generators is
@@
 %
-The approach taken by \fuzzchick leaves considerable room for
-improvements in several aspects.%
-\footnote{Some of these were already identified by the authors of \fuzzchick in
-  their original work.}
-%But \fuzzchick suffers several disadvantages.%
+%% The approach taken by \fuzzchick leaves considerable room for
+%% improvements in several aspects.%
+%
+
+In this work, we establish several aspects of the original CGPT approach by
+\citeauthor{lampropoulos2019coverage} that leave considerable room for
+improvement.%
+%
+\footnote{Some of these were already identified by
+  \citeauthor{lampropoulos2019coverage} in their original work.}
 %
@@
 \begin{inparaenum}
-\item type-preserving mutations, if not done carefully, can become superficial
-  and rely heavily on randomness to produce diverse mutants,
+\item structure-preserving mutations, if not done carefully, can become
+  superficial and rely heavily on randomness to produce diverse mutants,
 \item the queuing mechanism can cause large delays if interesting values are
   enqueued frequently and there is no way to prioritize them, and
-\item using an heuristic to assign a certain ``mutation energy'' to each
+\item using an heuristic to assign a certain ``mutation budget'' to each
   interesting test case (often referred to as a \emph{power schedule}) requires
@@
 %% Our solution
-In this work, we introduce \mutagen, a PBT tool implemented in Haskell
-that also builds upon code instrumentation and type-preserving mutations (see
+In this work, we introduce \mutagen, a CGPT framework that tackles the
+aforementioned limitations by applying mutations \emph{exhaustively} (see
 Section \ref{sec:mutagen}).
 %
+%% implemented in Haskell
+%% that also builds upon code instrumentation and type-preserving mutations
+%
+%
 % the basic ideas behind \fuzzchick, i.e.,
@@
 %% Exhaustive mutations
-Unlike \fuzzchick, mutations in \mutagen are applied \emph{exhaustively}.
+%% Unlike \fuzzchick, mutations in \mutagen are applied \emph{exhaustively}.
 %
 That is, given an interesting, previously executed test case, our tool
-precomputes and schedules every type-preserving mutation that can be applied to
-it.
+precomputes and schedules every structure-preserving mutation that can be
+applied to it.
 %
@@
 Our tool distinguishes two kinds of mutations, those that can be computed
-purely, yielding a single deterministic mutated value; and those that can be
+deterministically, yielding a single mutated value; and those that can be
 obtained non-deterministically.
 %
-On one hand, pure mutations encode transformations that swap data constructors
-around, as well as return or rearrange subexpressions.
+On one hand, deterministic mutations encode transformations that swap data
+constructors around, as well as return or rearrange subexpressions.
 %
@@
 
-The second heuristic is used to tune the number of times our tool samples random
-mutants, i.e., those corresponding to large enumeration types as described above.
+The second heuristic adjusts the number of times our tool samples random
+mutants, i.e., those corresponding to large enumeration types as described
+above.
 %
@@
 %
-If this frequency suddenly stalls, \mutagen automatically resets the testing
-loop increasing the amount of times we sample such mutants.
+If this frequency suddenly stalls, \mutagen resets the testing loop increasing
+the amount of times we sample such mutants.
 %
@@
 Section \ref{sec:casestudies}): the Information-Flow Control (IFC) stack machine
-used by \citeauthor{lampropoulos2019coverage} in their original work of
-\fuzzchick, and an existing WebAssembly engine of industrial strenght written in
-Haskell.
+used by \citeauthor{lampropoulos2019coverage} in their original work, and an
+existing WebAssembly engine of industrial strenght written in Haskell.
 %
@@
 %
-In the IFC stack machine case study, we compare \fuzzchick against different
-variants of our tool, where our results (Section \ref{sec:evaluation}) indicate
-that \mutagen outperforms \fuzzchick both in terms of time to failure and
-failure rate.
+In the IFC stack machine case study, we compare \mutagen against \fuzzchick, the
+reference CGPT implementation by \citeauthor{lampropoulos2019coverage}
+implemented in Coq.
+%
+Our results (Section \ref{sec:evaluation}) indicate that \mutagen outperforms
+\fuzzchick both in terms of time to failure and failure rate.
 %
 On the other hand, our WebAssembly case study shows the performance of our tool
-in a more realistic scenario.
+in a realistic scenario.
 %

### 2021-08-19 papers/icse22/main.tex
Pass over background

@@
 On the other hand, scheduling mutations exhaustively eliminates the need for a
-power schedule to assign a given mutation energy to each input test case.
+power schedule to assign a given mutation budget to each input test case.
 
@@
 
-In this section, we briefly introduce the concept of Property-Based Testing
-along with \quickcheck, one of the most popular tools of this sort and often
-used as the baseline when comparing PBT algorithms.
-%
-We also describe in detail the ideas and limitations behind \fuzzchick.
-
-%% which served as the foundation while designing \mutagen.
-
-%% \vspace{-5pt}
-\subsection{Property-Based Testing and \quickcheck}
-
-Property-based testing is a powerful technique for finding bugs without having
-to write test cases by hand.
-%
-Popularized by Haskell's \quickcheck, this technique focuses on aiming the
-developer's efforts into testing systems via executable properties using
-randomly generated inputs.
-%
-%% Moreover, tools like \quickchick and Isabelle's \quickcheck demonstrate
-%% that PBT can also be used in the formal verification realm.
-%% %
-%% There, one can quickly spot bugs in system specifications before directing the
-%% efforts into pointlessly trying to prove bogus propositions.
-
-In the simplest form, there are four main elements the user needs to provide in
-order to perform property-based testing on their systems:
-%
-\begin{itemize}
-\item one or more \emph{executable properties}, often implemented simply as
-  boolean predicates,
-\item \emph{random data generators}, used to repeatedly instantiate the testing
-  properties,
-\item \emph{printers}, used to show the user the random inputs that falsify some
-  testing property (the counterexample) whenever a bug is found, and
-\item \emph{shrinkers}, to minimize counterexamples making them easier to
-  understand by humans.
-\end{itemize}
-
-\noindent In this work we focus solely on the first two elements introduced
-above, namely the testing properties and the random data generators used to feed
-them.
-%
-Printers and shrinkers, for the most part, can often be obtained automatically
-using generic programming capabilities present in the compiler, and although
-being crucial for the testing process as a whole, their role becomes irrelevant
-when it comes to \emph{finding} bugs.
-
-\begin{algorithm}
-  \SetInd{0em}{0.75em}
-  \SetKw{KwNot}{not}
-  \SetAlgoLined
-  \DontPrintSemicolon
-  \SetKwFunction{QC}{Loop}
-  \SetKwProg{Fn}{Function}{:}{}
-  \Fn{\QC{P, N, gen}}{
-    i $\gets$ 0\;
-    \While{i $<$ N}{
-      x $\gets$ Sample(gen)\;
-      \lIf{\KwNot P(x)}{\KwRet Bug(x)}
-      i $\gets$ i+1 \;
-    }
-    \KwRet Ok\;
-  }
-\caption{\label{algo:quickcheck}\quickcheck Testing Loop}
-\end{algorithm}
-
-Perhaps the simplest PBT technique is to repeatedly generate random inputs and
-use them instantiate the testing properties until they either get falsified by a
-counterexample, or we ran a sufficiently large amount of tests --- suggesting
-that the properties hold.
-%
-\quickcheck implements a testing loop that closely follows this simple idea,
-which is outlined in Algorithm \ref{algo:quickcheck}, where $P$ is the testing
-property, $N$ is the maximum number of tests to perform, and $gen$ is the random
-generator used to instantiate $P$.
-
-To illustrate this technique, let us focus on the same motivating example used
-by \citeauthor{lampropoulos2019coverage}, who propose a simple property defined over
+In this section, we describe in detail the motivation, ideas and limitations
+behind coverage-guided, property-based testing as depicted by
+\citeauthor{lampropoulos2019coverage}.
+%
+To illustrate this technique, let us focus on a simple property defined over
 binary trees.
 %
-Such data structure can be defined in Haskell using a custom data type with two
-data constructors for leaves and branches respectively:
+Such a data structure can be defined in Haskell using a custom algebraic data
+type with two data constructors for leaves and branches respectively:
 
@@
 %
-Then, we can define tree reflections using a simple recursive function that
-pattern matches against the two possible constructors, inverting the order of
-the subtrees whenever it encounters a branch:
+%% Then, we can define tree reflections using a simple recursive function that
+%% pattern matches against the two possible constructors, inverting the order of
+%% the subtrees whenever it encounters a branch:
+%
+%% \begin{verbatim}
+%% mirror :: Tree a -> Tree a
+%% mirror (Leaf x)       = Leaf x
+%% mirror (Branch l x r) = Branch (mirror r) x (mirror l)
+%% \end{verbatim}
+%
+%% Later, a reasonable requirement to assert for is that \texttt{mirror} must be
+%% \emph{involutive}, i.e., reflecting a tree twice always yields the original
+%% tree.
+%% %
+%
+Then, if we assume existence of a predicate \texttt{balanced} of type
+\texttt{Tree a -> Bool} that asserts that a tree satisfies a certain notion of
+balancedness, we can write testing properties to assert that the operations
+defined over the \texttt{Tree} data type preserve this invariant.
+%
+For instance, the following testing property can be used to find bugs in the
+implementation of the \texttt{insert} function:
 
 \begin{verbatim}
-mirror :: Tree a -> Tree a
-mirror (Leaf x)       = Leaf x
-mirror (Branch l x r) = Branch (mirror r) x (mirror l)
+prop_insert_inv :: a -> Tree a -> Property
+prop_insert_inv x t =
+  balanced t ==> balanced (insert x t)
 \end{verbatim}
 
-Later, a reasonable requirement to assert for is that \texttt{mirror} must be
-\emph{involutive}, i.e., reflecting a tree twice always yields the original
-tree.
-%
-We can simply capture this property using a boolean predicate written as a
-normal function:
-
-\begin{verbatim}
-prop_mirror :: Tree Int -> Bool
-prop_mirror t = mirror (mirror t) == t
-\end{verbatim}
-
-\noindent For simplicity, here we instantiate the tree payload with integers,
-although this predicate should hold for any other type with a properly defined
-notion of equality as well.
+\noindent This property asserts that, given a balanced tree \texttt{t} and an
+element \texttt{x} as input, inserting \texttt{x} into \texttt{t} will produce a
+balanced tree as output.
+%
+(The implementation details of \texttt{balanced} and \texttt{insert} are not
+relevant for the point being made here.)
+%
+Moreover, the arrow operator (\texttt{==>}) indicates that the predicate
+\texttt{balanced t} is a precondition of this property, and so attempting to
+test it using an unbalanced tree as an input will result in the test case
+getting \emph{discarded}.
 
@@
 
-\noindent Let us break this definition into parts.
-%
-The first line states that we will provide an \texttt{Arbitrary} instance for
-trees with a payload of type \texttt{a}, provided that values of type \texttt{a}
-can also be randomly generated.
-%
-This allows us to use \texttt{arbitrary} to generate \texttt{a}'s inside the
-definition of our tree generator.
-
-Moreover, \quickcheck internally keeps track of the \emph{maximum generation
-  size}, a parameter that can be tuned by the user to limit the size of the
-randomly generated values.
-%
-Our definition exposes this size via \quickcheck's \texttt{sized} combinator,
-allowing us to parameterize the maximum size of the randomly generated trees.
-%
-If the generation size is zero (\texttt{gen 0}), our generator is limited to
-produce just leaves with randomly generated payloads.
-%
-In turn, when the generation size is strictly positive (\texttt{gen n}), the
-generator can perform a random uniform choice (\texttt{oneof}) between
-generating either a single leaf or a branch.
-%
-When generating branches, the generator calls itself recursively in order to
-produce random subtrees (\texttt{gen (n-1)}).
-%
-Notice the importance of reducing the generation size on each recursive call.
-%
-This way we ensure that randomly generated trees using a generation size
-\texttt{n} are always finite and have at most \texttt{n} levels.
-
-Finally, we are ready to let \quickcheck test \texttt{prop\_mirror} against a
-large number of inputs (100 by default) produced by our brand new random tree
-generator:
-
-\begin{verbatim}
-quickCheck prop_mirror
-+++ OK, passed 100 tests.
-\end{verbatim}
-
-\noindent Should we mistakenly introduce a bug in \texttt{mirror}, e.g., by
-dropping the right subtree altogether:
-
-\begin{verbatim}
-mirror (Branch l x r) = Branch (mirror l) x (mirror l)
-\end{verbatim}
-
-\noindent then \quickcheck will quickly falsify \texttt{prop\_mirror}, reporting
-a minimized counterexample that we can use to find the root of the issue:
-
-\begin{verbatim}
-quickCheck prop_mirror
-*** Failed! Falsified (after 2 tests and 2 shrinks):
-Branch (Leaf 0) 0 (Leaf 1)
-\end{verbatim}
-
-At this point, it is clear that the \emph{quality} of our random generators is
-paramount to the performance of the overall PBT process.
-%
-Random generators that rarely produce interesting values will fail to trigger
-bugs in our code, potentially leaving entire parts of the codebase virtually
-untested.
-
-Recalling our tree generator, the reader (far from mistaken) might already have
-imagined better ways for implementing it.
-%
-For most practical purposes, this generator is in fact quite bad.
-%
-However, it follows a simple type-directed fashion, and it is a good example of
-what to expect from a random generator synthesized automatically using a process
-that knows very little about the values to be generated apart from their
-(syntactic) data type structure.
-
-As introduced earlier, there exist multiple tools that can automatically derive
-better random generators solely from the static information present in the
-codebase.
-%
-%% Such tools vary on the degree of invariants and requirements
-%% %
-%% However, such approaches are unable to derive useful generators whenever the
-%% target data involves complex invariants that cannot be easily extracted from the
-%% codebase or that cannot be easily expressed in terms of inductive relations.
-%
-Sadly, these tools lack the domain knowledge required to generate random data
-with complex invariants --- especially those present in programming languages
-like well-scopedness and well-typedeness.
-
-In particular, \emph{automatically derived generators are remarkably ineffective
-  when used to test properties with sparse preconditions.}
-%
-Let us continue with the example by \citeauthor{lampropoulos2019coverage} to
-illustrate this problem in more detail.
-%
-For this, consider that we want to use our \texttt{Tree} data type to encode
-binary-search trees (BST) --- this requires some minor tweaks in practice.
-%
-Then, given a predicate \texttt{isBST} that asserts if a tree satisfies the BST
-invariants, we might want to use it as pre- and post-condition to assert that
-BST operations like \texttt{insert} preserve them:
-
-\begin{verbatim}
-prop_bst_insert :: Tree a -> a -> Bool
-prop_bst_insert t a =
-  isBST t ==> isBST (insert a t)
-\end{verbatim}
-
-\noindent Attempting to test this property using \quickcheck does not work well:
-
-\begin{verbatim}
-quickCheck prop_bst_insert
-*** Gave up! Passed only 44 tests; 1000 discarded tests.
-\end{verbatim}
-
-\noindent \quickcheck \emph{discards} random inputs as soon as it finds they do
-not pass the precondition (\texttt{isBST t}).
+\noindent This definition follows a simple type-directed generation approach,
+and it is a good example of what to expect from a random generator synthesized
+automatically from the codebase.
+%
+In esence, this generator randomly picks one data constructor from its
+corresponding data type definition with uniform probability (i.e, either a
+\texttt{Leaf} or a \texttt{Branch}), and then calls itself for every recursive
+subexpression needed to produce a well-typed value, carefully reducing the depth
+limit (controlled by \quickcheck) by a unit on each step, ensuring termination.
+%
+Subexpressions of any other type are generated simply by delegating the task to
+their corresponding \texttt{Arbitrary} instance --- as noted by the use of
+\texttt{arbitrary} when generating random payloads inside of the \texttt{Leaf}
+and \texttt{Branch} constructors.
+
+Readers familiar with RBPT will recognize that attempting to test this property
+using \quickcheck directly does not work well.
+%
+In practice, \quickcheck will give up quite early if most of the generated test
+cases get discarded because they do not satisfy the balanced precondition.
 %
 Sadly, most of the inputs generated by our na\"ive generator suffer from this
-problem, and the interesting part of the property (\texttt{isBST (insert a t)})
+problem, and the interesting part of the property (\texttt{balanced (insert x t)})
 is tested very sporadically as a result.
-
-At this point it is reasonable to think that, to obtain the best results when
-using PBT over complex systems, one is forced to put a large amount of time into
-developing manually-written generators.
-%
-In practice, that is most often the case, no automatic effort can beat a
-well-thought manually-written generator that produces interesting complex values
-and finds bugs in very few tests.
-%
-Not all is lost, however.
-%
-It is still possible to obtain acceptable results automatically by incorporating
-dynamic information from the system under test into the testing loop.
-%
-The next subsection introduces the clever technique used by \fuzzchick to find
-bugs in complex systems while using simple automatically derived random
-generators.
+%
+In other words, \emph{the precondition of our testing property becomes very
+  sparse when we use simple, automatically derived generators.}
+
+The next subsection show how coverage-guided, property-based testing can help
+overcoming this limitation without having to write specialized random generators
+by hand.
+
+%% At this point it is reasonable to think that, to obtain the best results when
+%% using PBT over complex systems, one is forced to put a large amount of time into
+%% developing manually-written generators.
+%% %
+%% In practice, that is most often the case, no automatic effort can beat a
+%% well-thought manually-written generator that produces interesting complex values
+%% and finds bugs in very few tests.
+%% %
+%% Not all is lost, however.
+%% %
+%% It is still possible to obtain acceptable results automatically by incorporating
+%% dynamic information from the system under test into the testing loop.
+%% %
+%% The next subsection introduces the clever technique used by \fuzzchick to find
+%% bugs in complex systems while using simple automatically derived random
+%% generators.
 
@@
 
-\subsection{Coverage-Guided Property-Based Testing with \fuzzchick}
-
-To alleviate the problem of testing properties with sparse preconditions while
-using automatically derived random generators, \fuzzchick introduces
-\emph{coverage-guided, property-based testing} (CGPT), a technique that enhances
-the testing process with two key characteristics:
+\subsection{Coverage-Guided Property-Based Testing}
+
+To alleviate the problem of testing properties with non-trivial preconditions
+while using automatically derived random generators,
+\citeauthor{lampropoulos2019coverage} introduced \emph{coverage-guided,
+  property-based testing} (CGPT), a technique that enhances the testing process
+with two key characteristics:
 %
@@
   each test case; and
-\item \emph{high-level, type-preserving mutations}, to produce syntactically
-  valid test cases by altering existing ones at the data type level.
+\item \emph{high-level, structure-preserving mutations}, to produce
+  syntactically valid test cases by altering existing ones at the data type
+  level.
 \end{inparaenum}
@@
 %
-\fuzzchick, however, does this in a clever way.
+CGPT, however, does this in a clever way.
 %
 Instead of mutating any previously executed test case that discovers a new part
-of the code, \fuzzchick integrates these fuzzing techniques into the PBT testing
-loop itself.
+of the code, this technique integrates these fuzzing mechanisms into the PBT
+testing loop itself.
 
 Since it is possible to distinguish semantically valid test cases from invalid
-ones, i.e., those passing the sparse preconditions of our testing properties as
-opposed to those that are discarded early, \fuzzchick exploits this information
-in order to focus the testing efforts into mutating valid test cases with a
-higher priority than those that were discarded.
-
-In addition, high-level, type-preserving mutators are better suited for
+ones, i.e., those passing the testing property preconditions as opposed to those
+that are discarded early, CGPT exploits this information to prefer mutating
+valid test cases than discarded ones.
+
+In addition, high-level, structure-preserving mutators are better suited for
 producing syntactically valid mutants, avoiding the time wasted by using generic
@@
 %
-In \fuzzchick, external grammars are not required.
+In CGPT, external grammars are not required.
 %
@@
 
-The next subsections describe \fuzzchick's testing loop and type-preserving
-mutations in detail.
-
-\subsubsection{Testing loop}
-
-Outlined in Algorithm \ref{algo:fuzzchick:loop}, the process starts by creating
-two queues, \textit{QSucc} and \textit{QDisc} for valid and discarded previously
+%% The next subsections describe the CGPT testing loop and structure-preserving
+%% mutations in detail.
+
+%% \subsubsection{Testing loop}
+
+\paragraph{Testing Loop}
+Outlined in Algorithm \ref{algo:fuzzchick:loop}, CGPT starts by creating two
+queues, \textit{QSucc} and \textit{QDisc} for valid and discarded previously
 executed test cases, respectively.
 %
-Enqueued values are stored along with a given mutation energy, that controls how
+Enqueued values are stored along with a given mutation budget, that controls how
 many times a given test case can be mutated before being finally thrown away.
 
-Once inside of the main loop, \fuzzchick picks the next test case using a simple
-criterion: if there are valid values enqueued for mutation, it picks the first
-one, mutates it and returns it, decreasing its energy by one.
+Once inside of the main loop, the algorithm picks the next test case using a
+simple criterion: if there are valid values enqueued for mutation, it picks the
+first one, mutates it and returns it, decreasing its energy by one.
 %
@@
 %
-If none of the mutation queues contain any candidates, \fuzzchick generates a
-new value from scratch.
+If none of the mutation queues contain any candidates, CGPT generates a new
+value from scratch.
 %
@@
 %
-If not, \fuzzchick evaluates whether it was interesting (i.e., it exercises a
+If not, the algorithm evaluates whether it was interesting (i.e., it exercises a
 new path) based on its trace information and the one from previously executed
@@
 
-The energy assigned to each test case follows that of AFL's power schedule: more
-energy to test cases that lead to shorter executions, or that discover more
-parts of the code.
-%
-Moreover, to favor mutating interesting valid test cases, they get more energy
-than those that were discarded.
+%% The energy assigned to each test case follows that of AFL's power schedule: more
+%% energy to test cases that lead to shorter executions, or that discover more
+%% parts of the code.
+%% %
+%% Moreover, to favor mutating interesting valid test cases, they get more energy
+%% than those that were discarded.
 
@@
   \SetArgSty{textnormal}
-  \SetInd{0em}{0.75em}
+  %% \SetInd{0em}{0.75em}
   \SetKw{KwNot}{not}
@@
   }
-\caption{\label{algo:fuzzchick:loop}\fuzzchick Testing Loop}
+\caption{\label{algo:fuzzchick:loop}CGPT Testing Loop}
 \end{algorithm}
@@
   }
-\caption{\label{algo:fuzzchick:pick}\fuzzchick Seed Selection}
+\caption{\label{algo:fuzzchick:pick}CGPT Seed Selection}
 \end{algorithm}
 
-\subsubsection{Type-preserving mutations}
-
-Mutators in \fuzzchick are no more than specialized random generators,
-parameterized by the original input to be mutated.
-%
-They use a simple set of mutation operations that are randomly applied at the
-data type level.
+%% \subsubsection{Type-preserving mutations}
+
+\paragraph{Structure-preserving Mutators}
+
+%% are no more than specialized random generators, parameterized
+%% by the original input to be mutated.
+%% %
+%% They
+
+Mutators in GCPT use a simple set of mutation operations that are randomly
+applied to their inputs at the data type level.
 %
@@
 \begin{inparaenum}
-\item \emph{shrinking the value}, replacing its outermost data constructor with
-  one that contains a subset of its fields, reusing existing subexpressions;
-\item \emph{growing the value}, replacing its outermost data constructor with
-  one that contains a superset of its fields, reusing existing subexpressions
-  and generating random ones when needed;
-\item \emph{returning a subexpression of the same type}; and
+\item \emph{swapping a data constructor}, replacing its outermost data
+  constructor by another one, e.g., a \texttt{Leaf} by a \texttt{Branch} and
+  vice-versa, either dropping unused subexpressions or generating new random
+  ones when the new data constructor contains extra fields that cannot be filled
+  using existing ones;
+%% \item \emph{shrinking the value}, replacing its outermost data constructor with
+%%   one that contains a subset of its fields, reusing existing subexpressions;
+%% \item \emph{growing the value}, replacing its outermost data constructor with
+%%   one that contains a superset of its fields, reusing existing subexpressions
+%%   and generating random ones when needed;
+\item \emph{returning a subexpression of the same type}, e.g, returning a left
+  or right subtree when applied to a \texttt{Branch}; and
 \item \emph{mutating recursively}, applying a mutation operation over an
@@
 \end{inparaenum}
+%
+We will provide more details about structure-preserving mutations in Section
+\ref{sec:mutagen}.
 
@@
 
-\begin{figure}
-%% \vspace{-5pt}
-\begin{verbatim}
-mutate_tree :: (a -> Gen a) -> Tree a -> Gen (Tree a)
-mutate_tree mutate_a (Leaf x) =
-  oneof [ Leaf <$> mutate_a x
-        , Branch <$> arbitrary <*> return x <*> arbitrary ]
-mutate_tree mutate_a (Branch l x r) =
-  oneof [ return l
-        , return r
-        , Leaf   <$> return x
-        , Branch <$> mutate_tree l <*> return x   <*> return r
-        , Branch <$> return l      <*> mutate_a x <*> return r
-        , Branch <$> return l      <*> return x   <*> mutate_tree r
-\end{verbatim}
-\caption{\label{fig:fuzzchick:mutator}\fuzzchick mutator for the \texttt{Tree} data type. }
-\end{figure}
-
-Fig. \ref{fig:fuzzchick:mutator} illustrates a \fuzzchick-like mutator for our
-previously used \texttt{Tree} data type example.
-%
-Since trees are parametric, for clarity this definition is also parameterized by
-a mutator for the payload (\texttt{mutate\_a}), although this can be abstracted
-away using the type class system.
-
-%% (where \texttt{mutate\_tree} and
-%% \texttt{mutate\_a} can be transparently coerced into an overloaded
-%% \texttt{mutate} function.)
-
-In this mutator, branches can be shrunk into leaves by dropping the subtrees,
-whereas leaves can grow into branches, by reusing the payload and generating two
-random subtrees.
-%
-Moreover, branches can be replaced with one of their subtrees.
-%
-Finally, mutations can be recursively applied over both the payload and the
-subtrees.
-%
-At the top level, all these operations are put together using the \texttt{oneof}
-combinator that randomly picks one of them with uniform probability.
+%% \begin{figure}
+%% %% \vspace{-5pt}
+%% \begin{verbatim}
+%% mutate_tree :: (a -> Gen a) -> Tree a -> Gen (Tree a)
+%% mutate_tree mutate_a (Leaf x) =
+%%   oneof [ Leaf <$> mutate_a x
+%%         , Branch <$> arbitrary <*> return x <*> arbitrary ]
+%% mutate_tree mutate_a (Branch l x r) =
+%%   oneof [ return l
+%%         , return r
+%%         , Leaf   <$> return x
+%%         , Branch <$> mutate_tree l <*> return x   <*> return r
+%%         , Branch <$> return l      <*> mutate_a x <*> return r
+%%         , Branch <$> return l      <*> return x   <*> mutate_tree r
+%% \end{verbatim}
+%% \caption{\label{fig:fuzzchick:mutator}\fuzzchick mutator for the \texttt{Tree} data type. }
+%% \end{figure}
+
+%% Fig. \ref{fig:fuzzchick:mutator} illustrates a \fuzzchick-like mutator for our
+%% previously used \texttt{Tree} data type example.
+%% %
+%% Since trees are parametric, for clarity this definition is also parameterized by
+%% a mutator for the payload (\texttt{mutate\_a}), although this can be abstracted
+%% away using the type class system.
+
+%% %% (where \texttt{mutate\_tree} and
+%% %% \texttt{mutate\_a} can be transparently coerced into an overloaded
+%% %% \texttt{mutate} function.)
+
+%% In this mutator, branches can be shrunk into leaves by dropping the subtrees,
+%% whereas leaves can grow into branches, by reusing the payload and generating two
+%% random subtrees.
+%% %
+%% Moreover, branches can be replaced with one of their subtrees.
+%% %
+%% Finally, mutations can be recursively applied over both the payload and the
+%% subtrees.
+%% %
+%% At the top level, all these operations are put together using the \texttt{oneof}
+%% combinator that randomly picks one of them with uniform probability.
 
@@
 %
-\subsubsection{Limitations of \fuzzchick}
-
-\citeauthor{lampropoulos2019coverage} demonstrated empirically that \fuzzchick
-lies comfortably in the middle ground between using pure random testing with
-na\"ive automatically derived random generators and complex manually-written
-ones.
-%
-Their results suggest that CGPT is an appealing technique for finding bugs while
-still using a mostly automated workflow.
+\subsubsection{Limitations of CGPT}
+
+\citeauthor{lampropoulos2019coverage} demonstrated empirically that CGPT lies
+comfortably in the middle ground between using pure random testing with na\"ive
+automatically derived random generators and complex manually-written ones.
+%
+%% Their results suggest that CGPT is an appealing technique for finding bugs while
+%% still using a mostly automated workflow.
 %
@@
 %% Moreover, we replicated the Information-Flow Control (IFC) stack machine case
+%
 Moreover, we replicated the IFC stack machine case study bundled with \fuzzchick
-and observed a surprising limitation:
+(their reference implementation) and observed a surprising limitation:
 %
@@
 %
-Despite this, the authors show that \fuzzchick outperforms normal random testing
+Despite this, the authors show that CGPT outperforms normal random testing
 coupled with na\"ive random generators by comparing the mean-time-to-failure
-(MTTF) against \emph{QuickChick}.
+(MTTF) against normal random property-based testing.
 %
@@
 
-All these observations led us to consider three main aspects of \fuzzchick that
-can be improved upon --- and that constitute the main goal of this work.
+All these observations led us to consider three main aspects of CGPT that can be
+improved upon --- and that constitute the main goal of this work.
 %
@@
 %
-if we inspect the mutator defined in Fig. \ref{fig:fuzzchick:mutator}, there are
-two compromises that the authors of \fuzzchick adopted for the sake of
-simplicity.
-%
-On one hand, deep recursive mutations are very unlikely, since their probability
+for the sake of simplicity, the mutators proposed by
+\citeauthor{lampropoulos2019coverage} are derived to follow a top-down approach,
+similar to that used by the automatically derived random generators.
+%
+In this light, a mutation can happen at top-level or be recursively applied to
+an immediate subexpression with approximately the same probability.
+%
+%%   if we inspect the mutator defined in Fig. \ref{fig:fuzzchick:mutator}, there are
+%% two compromises that the authors of \fuzzchick adopted for the sake of
+%% simplicity.
+%
+This makes deep recursive mutations very unlikely, since their probability
 decreases multiplicatively with each recursive call.
 %
-For instance, mutating a subexpression that lies on the fourth level of a
-\texttt{Tree} happens with a probability smaller than $(1/6)^3 = \sim 0.0046$,
-i.e., when applying a recursive mutation (with probability $1/6$) three
-consecutive times.
-%
-This only worsens as the type of the mutated value becomes more complex.
-%
-Hence, \fuzzchick mutators can only be effectively used to transform shallow
+%% For instance, mutating a subexpression that lies on the fourth level of a
+%% \texttt{Tree} happens with a probability smaller than $(1/6)^3 = \sim 0.0046$,
+%% i.e., when applying a recursive mutation (with probability $1/6$) three
+%% consecutive times.
+%
+%% This only worsens as the type of the mutated value becomes more complex.
+%
+Hence, these simple mutators can only be effectively used to transform shallow
 data structures, potentially excluding interesting applications that might
@@
 %
-Ideally, mutations should be able to happen on every subexpression of the input
-seed on a reasonable basis.
+Ideally, \emph{mutations should be able to happen on every subexpression of the
+  input test case on a reasonable basis}.
 
@@
 We believe that growing data constructors needs to be done carefully.
-%
-For instance, by using just a minimal piece of data to make the overall mutated
-test case type correct.
-%
-If that mutated test case turns to be interesting, that new subexpression can
-always be mutated later.
+%% %
+%% For instance, by using just a minimal piece of data to make the overall mutated
+%% test case type correct.
+%% %
+%% If that mutated test case turns to be interesting, that new subexpression can
+%% always be mutated later.
 
@@
 %
-\fuzzchick uses two single queues for keeping valid and discarded mutation
-candidates.
+the CGPT testing loop uses two single queues for keeping valid and discarded
+mutation candidates.
 %
@@
 more often than they tend to shrink, as interesting mutation candidates can get
-buried at the end of a long queue that only exercises the same portion of the
-target code.
+``buried'' at the end of a long queue that only exercises the same portion of
+the target code.
 %
@@
 %
-Ideally, one would like a mechanism that prioritizes mutating test cases that
-discover new portions of the code right away, and that is capable of jumping
-back and forth from mutation candidates whenever this happens.
+Ideally, \emph{one would like a mechanism that prioritizes mutating test cases
+  that discover new portions of the code right away}, and that is capable of
+jumping back and forth from mutation candidates whenever this happens.
 %
@@
 %
-It is not clear how the power schedule used to assign energy to each mutable
-test case in \fuzzchick works in the context of high-level type-preserving
-mutations.
+it is not clear how the power schedule used to assign a given budget to each
+mutable test case in CGPT works in the context of high-level
+structure-preserving mutations.
 %
@@
 characterize in general given the expressivity of the data types used to drive
-the mutators.\\
-
+the mutators.
 \end{itemize}
@@
 The next section introduces \mutagen, our CGPT tool written in Haskell that aims
-to tackle the main limitations of \fuzzchick using an exhaustive mutation
-approach that requires very little randomness and no power schedule.
+to tackle the main limitations of CGPT using an exhaustive mutation approach
+that requires very little randomness and no power schedule.
 
@@
 
-\section{\mutagen: Testing Mutants Exhaustively}
+\section{\mutagen: No Mutant Left Behind}
 \label{sec:mutagen}
@@
   \SetArgSty{textnormal}
-  \SetInd{0em}{0.75em}
+  %% \SetInd{1em}{em}
   \SetKw{KwNot}{not}

### 2021-08-20 papers/icse22/main.tex
Work in section 3

@@
 
-The next section introduces \mutagen, our CGPT tool written in Haskell that aims
-to tackle the main limitations of CGPT using an exhaustive mutation approach
-that requires very little randomness and no power schedule.
+The next section introduces \mutagen, our CGPT tool that aims to tackle the main
+limitations of CGPT.
 
@@
 
-In this section we describe the base ideas behind \mutagen, our CGPT tool
-written in Haskell.
-%
-Notably, the ideas presented in this work extend beyond Haskell and functional
-programming languages in general.
+In this section we describe the main ideas behind \mutagen, our enhanced CGPT
+tool written in Haskell.
+%
+Notably, these ideas extend beyond Haskell and functional programming languages
+in general.
 %
@@
 
-In contrast with \fuzzchick, \mutagen does not employ a power schedule to assign
-energy to mutable candidates.
-%
-In turn, it resorts to mutate them on an exhaustive and precise manner, where
+In contrast to CGPT, and many other popular fuzzing tools for that matter,
+\mutagen works by to mutating test cases on an exhaustive and precise manner,
+where
 %
 \begin{inparaenum}
-  \item each subexpression of a mutation candidate is associated with a set of
-    type-preserving mutations, and
-  \item for every mutable subexpression, each one of these mutations is
-    scheduled \emph{exactly once}.
+  \item each subexpression of a mutable test case is associated with a set of
+    structure-preserving mutations, and
+  \item each one of these mutations is scheduled \emph{exactly once}.
 \end{inparaenum}
@@
 %% There is a small exception to this rules that we will introduce soon.
-
-This idea is inspired by exhaustive bounded testing tools like \emph{SmallCheck}
+%
+Our realization is that, by following an exhaustive mutation approach, we can
+avoid the need for a power schedule, i.e., a heuristic to assign a certain
+mutation budget to each mutated test case.
+%
+Moreover, producing mutants exhaustively ensures that no interesting mutation is
+omitted due to randomness, as well a no mutation is evaluated more than once.
+%
+This is inspired by exhaustive bounded testing tools like \emph{SmallCheck}
 \cite{runciman2008smallcheck} or \emph{Korat} \cite{boyapati2002korat}, that
-produce test cases exhaustively.
-%
-In simple words, such tools work by enumerating all possible values of the data
-types used in the testing properties, and then executing them from smaller to
-larger until a bug is found, or a certain size bound is reached.
-%
-The main problem with this approach is that the space of all possible test cases
-often grows exponentially as we increment the size bound, and the user
-experiences what it looks like ``hitting a wall'', where no larger test cases
-can be evaluated until we exhausted all the immediately smaller ones
-\cite{DuregardJW12}.
-%
-To alleviate this problem, such tools often rely on heuristics to prune the
-search space based on detecting unused subexpressions --- this is discussed in
-Section \ref{sec:related}.
+produce test cases exhaustively --- please refer to Section \ref{sec:related}
+for a detailed discussion.
+
+The rest of this section describes our exhaustive mutation mechanism, as well as
+the adapted testing loop used in \mutagen.
+
+%% In simple words, such tools work by enumerating all possible values of the data
+%% types used in the testing properties, and then executing them from smaller to
+%% larger until a bug is found, or a certain size bound is reached.
+%% %
+%% The main problem with this approach is that the space of all possible test cases
+%% often grows exponentially as we increment the size bound, and the user
+%% experiences what it looks like ``hitting a wall'', where no larger test cases
+%% can be evaluated until we exhausted all the immediately smaller ones
+%% \cite{DuregardJW12}.
+%% %
+%% To alleviate this problem, such tools often rely on heuristics to prune the
+%% search space based on detecting unused subexpressions --- this is discussed in
+%% Section \ref{sec:related}.
 
@@
 
-Not to be confused by these tools, in \mutagen we do not enumerate all possible
-test cases exhaustively.
-%
-Our tool uses random generators to find interesting initial seeds, and when it
-finds them, only then proceeds to schedule all possible mutations.
-%
-Moreover, as in \fuzzchick, the testing loop of our tool automatically filters
-the test cases that are worth mutating by keeping only those that discover new
-execution paths in the code under test.
+%% Not to be confused by these tools, in \mutagen we do not enumerate all possible
+%% test cases exhaustively.
+%% %
+%% Our tool uses random generators to find interesting initial seeds, and when it
+%% finds them, only then proceeds to schedule all possible mutations.
+%% %
+%% Moreover, as in \fuzzchick, the testing loop of our tool automatically filters
+%% the test cases that are worth mutating by keeping only those that discover new
+%% execution paths in the code under test.
 
@@
 
-To describe \mutagen's testing loop, we first need to introduce the mechanism
-used for testing mutations exhaustively.
-%
-In contrast to \fuzzchick, where mutators are parameterized random generators,
-in our tool we define mutations as the set of mutants that can be obtained by
-altering the input value at the top-level (the root).
+%% Before describing \mutagen's testing loop, we first need to introduce the
+%% mechanism used for testing mutations exhaustively.
+%% %
+%% In contrast to \fuzzchick, where mutators are parameterized random generators,
+%% in our tool
+
+Mutations in \mutagen are defined as the set of mutants that can be obtained by
+transforming the input test case \emph{at the top-level} (the root).
 %
@@
 %
-Notice how this definition simply enumerates mutants that transform the
-outermost data constructor, hence no recursion is needed here.
+Notably, this definition simply enumerates mutants that transform the outermost
+data constructor, hence no recursion is needed here.
 %
@@
 
-\begin{figure}
+\begin{figure}[t]
 \begin{verbatim}
@@
 %
-Random mutants, on the other hand, serve as a way to break exhaustiveness when
-mutating values of large enumeration types --- or any other type the user might
-want to use random mutations with.
+Random mutants, on the other hand, serve as a way to \emph{selectively avoid
+  exhaustiveness} when mutating values of large enumeration types --- or any
+other type the user might want to use random mutations with.
 %
@@
 %
-This way, a mutator for integers simply becomes:
+Then, for instance, a mutator for integers becomes:
 
@@
 %
-To do so, we will use two functions that can be derived from the data type
-definition.
+To do so, we will use two functions that can be derived from the code.
 
 In first place, a function \texttt{positions} traverses the input and builds a
-Rose tree of mutable positions.
-%
-These positions are essentially a list of indices encoding the path from the
-root to every mutable subexpression.
+Rose tree of \emph{mutable positions}, where positions are lists of indices
+encoding the path from the root to every mutable subexpression.
 %
@@
 inside []     x              = mutate x
-inside (0:pos) (Leaf x)       = [ Leaf x'       | x' <- inside pos x ]
-inside (0:pos) (Branch l x r) = [ Branch l' x r | l' <- inside pos l ]
-inside (1:pos) (Branch l x r) = [ Branch l x' r | x' <- inside pos x ]
-inside (2:pos) (Branch l x r) = [ Branch l x r' | r' <- inside pos r ]
+inside (0:ps) (Leaf x)       = [ Leaf x'       | x' <- inside ps x ]
+inside (0:ps) (Branch l x r) = [ Branch l' x r | l' <- inside ps l ]
+inside (1:ps) (Branch l x r) = [ Branch l x' r | x' <- inside ps x ]
+inside (2:ps) (Branch l x r) = [ Branch l x r' | r' <- inside ps r ]
 \end{verbatim}
@@
 %
-As it can be observed, we closely follow \fuzzchick's testing loop, using two
-queues to keep mutable candidates, and enqueueing and retrieving interesting
-test cases from them before falling back to random generation.
-
-\begin{algorithm}
-  \SetArgSty{textnormal}
-  %% \SetInd{1em}{em}
-  \SetKw{KwNot}{not}
-  \SetKw{KwIn}{in}
-  \SetKwFor{KwRepeat}{repeat}{times}
-  \SetAlgoLined
-  \DontPrintSemicolon
-  \SetKwFunction{Muts}{Mutate}
-  \SetKwProg{Fn}{Function}{:}{}
-  \Fn{\Muts{x, mut, R}}{
-    batch $\gets$ $\varnothing$\;
-    \For{pos \KwIn Flatten(Positions(x))}{
-      \For{mutant \KwIn Inside(pos, mut, x)}{
-        \Switch{mutant}{
-          \Case{\texttt{PURE} $\hat{x}$}{
-            Enqueue($\hat{x}$, batch)\;
-          }
-          \Case{\texttt{RAND} gen}{
-            \KwRepeat{R}{
-              $\hat{x}$ $\gets$ Sample(gen)\;
-              Enqueue($\hat{x}$, batch)\;
-            }
-          }
-        }
-      }
-    }
-    \KwRet batch\;
-  }
-\caption{\label{algo:mutagen:init}Mutants Initialization}
-\end{algorithm}
-
-The main difference lies in that we precompute all the mutations of a given
-mutation candidate before enqueueing them.
-%
-These mutations are put together into lists that we call \emph{mutation batches}
---- one for each mutated test case.
-%
-To initialize a mutation batch (outlined in Algorithm \ref{algo:mutagen:init}),
-we first flatten all the mutable positions of the input value in level order
-(recall that positions are stored as a Rose tree).
-%
-Then, we iterate over all of them and retrieve all the mutants defined for each
-subexpression.
-%
-For each one of these, there are two possible cases:
-%
-\begin{inparaenum}
-\item if it is a pure mutant carrying a concrete mutated value, we enqueue it
-  into the mutation batch directly; otherwise
-\item it is a random mutant that carries a random generator with it (e.g.,
-  corresponding to a numeric subexpression), in which case we sample and enqueue
-  \textit{R} random values using this generator, where \textit{R} is a parameter
-  set by the user.
-\end{inparaenum}
-%
-At the end, we simply return the accumulated batch.
-
-Finally, the seed selection algorithm (Algorithm \ref{algo:mutagen:pick}) simply
-selects the next test case using the same criteria as \fuzzchick, prioritizing
-valid candidates over discarded ones, falling back to random generation when
-both queues are empty.
-%
-Since mutations are precomputed, this function only needs to pick the next test
-case from the current batch, until it becomes empty and can switch to the next
-precomputed one in line.
+As it can be observed, we follow the CGPT testing loop, using two queues to keep
+mutable test cases, and enqueueing and retrieving interesting test cases from
+them before falling back to random generation.
 
@@
   \SetArgSty{textnormal}
+  %% \SetInd{1em}{em}
+  \SetKw{KwNot}{not}
+  \SetKw{KwIn}{in}
+  \SetKwFor{KwRepeat}{repeat}{times}
+  \SetAlgoLined
+  \DontPrintSemicolon
+  \SetKwFunction{Muts}{Mutate}
+  \SetKwProg{Fn}{Function}{:}{}
+  \Fn{\Muts{x, mut, R}}{
+    batch $\gets$ $\varnothing$\;
+    \For{pos \KwIn Flatten(Positions(x))}{
+      \For{mutant \KwIn Inside(pos, mut, x)}{
+        \Switch{mutant}{
+          \Case{\texttt{PURE} $\hat{x}$}{
+            Enqueue($\hat{x}$, batch)\;
+          }
+          \Case{\texttt{RAND} gen}{
+            \KwRepeat{R}{
+              $\hat{x}$ $\gets$ Sample(gen)\;
+              Enqueue($\hat{x}$, batch)\;
+            }
+          }
+        }
+      }
+    }
+    \KwRet batch\;
+  }
+\caption{\label{algo:mutagen:init}Mutants Initialization}
+\end{algorithm}
+
+The difference with CGPT lies in that we precompute all the mutations of a given
+mutable test case before enqueueing them.
+%
+These mutations are put together into lists that we call \emph{mutation batches}
+--- one for each mutated test case.
+%
+To initialize a mutation batch (outlined in Algorithm \ref{algo:mutagen:init}),
+we first flatten all the mutable positions of the input test case in level order
+(recall that positions are stored as a Rose tree).
+%
+Then, we iterate over all of them and retrieve all the mutants defined for each
+subexpression.
+%
+For each one of these, there are two possible cases:
+%
+\begin{inparaenum}
+\item if it is a pure mutant carrying a concrete mutated value, we enqueue it
+  into the mutation batch directly; otherwise
+\item it is a random mutant that carries a random generator with it (e.g.,
+  corresponding to a numeric subexpression), in which case we sample and enqueue
+  \textit{R} random values using this generator, where \textit{R} is a parameter
+  set by the user.
+\end{inparaenum}
+%
+At the end, we simply return the accumulated batch.
+
+Finally, the seed selection algorithm (Algorithm \ref{algo:mutagen:pick}) picks
+the next test case using the same criteria as CGPT, prioritizing valid test
+cases over discarded ones, falling back to random generation when both queues
+are empty.
+%
+Since mutations are precomputed, this function simply picks the next test case
+from the current batch, until it becomes empty and can switch to the next
+precomputed one in line.
+
+Another difference is the criterion for enqueuing discarded tests.
+%
+We found that, especially for large data types, the queue of discarded
+candidates tends to grow disproportionately large during testing, making them
+hardly usable and consuming large amounts of memory.
+%
+To improve this, we resort to mutate discarded tests cases only when we have
+some evidence that they are ``almost valid.''
+%
+For this, each mutated test case remembers whether its parent (the original test
+case they derive from after being mutated) was valid.
+%
+Then, we enqueue discarded test cases only if they meet this condition.
+%
+As a result, we fill the discarded queue with lesser but much more interesting
+mutable test cases.
+%% %
+%% Moreover, this can potentially lead to \emph{2-step mutations}, where an initial
+%% mutation breaks a valid test case in a small way, it gets enqueued as discarded,
+%% and later a subsequent mutation fixes it by changing a different subexpression.
+
+\begin{algorithm}
+  \SetArgSty{textnormal}
   %% \SetInd{0em}{0.75em}
@@
 
-Another small difference between \mutagen and \fuzzchick is the criterion for
-enqueuing discarded tests.
-%
-We found that, especially for large data types, the queue of discarded
-candidates tends to grow disproportionately large during testing, making them
-hardly usable and consuming large amounts of memory.
-%
-To improve this, we resort to mutate discarded tests cases only when we have
-some evidence that they are ``almost valid.''
-%
-For this, each mutated test case remembers whether its parent (the original test
-case they derive from after being mutated) was valid.
-%
-Then, we enqueue discarded test cases only if they meet this condition.
-%
-As a result, we fill the discarded queue with lesser but much more interesting
-mutation candidates.
-%
-Moreover, this can potentially lead to \emph{2-step mutations}, where an initial
-mutation breaks a valid test case in a small way, it gets enqueued as discarded,
-and later a subsequent mutation fixes it by changing a different subexpression.
-
-The next section introduces two heuristics we added to the base testing loop of
-\mutagen based on the limitations we found in \fuzzchick.
+The next section describes two heuristics we added to \mutagen's testing loop
+based on the limitations we found in CGPT.
 %
@@
 %
-\mutagen enables them all by default, although they can be individually disabled
-by the user if deems appropriate.
+\mutagen enables them all by default, and can be individually disabled by the
+user if needed.
 
@@
 
-\fuzzchick uses AFL instrumentation under the hood, which in turn uses an
-\emph{edge coverage} criterion to distinguish novel executions and to assign
-each mutation candidate a given energy.
-%
-In contrast, execution traces \mutagen represent the specific \emph{path} in the
-code taken by the program, as opposed to just the (unordered) set of edges
-traversed in the control-flow graph (CFG).
+%% \fuzzchick uses AFL instrumentation under the hood, which in turn uses an
+%% \emph{edge coverage} criterion to distinguish novel executions and to assign
+%% each mutation candidate a given energy.
+%% %
+Execution traces in \mutagen represent the specific \emph{path} in the code
+taken by the program, as opposed to just the (unordered) set of edges traversed
+in the control-flow graph (CFG) used by default by most coverage-guided fuzzers.
 %
@@
 from the ones already executed are more likely to discover completely new
-portions of the code under test, and hence we want to assign them a higher
-priority.
+portions of the code under test, and hence we want to prioritize them.
 
 In this light, every time we insert a new execution path into the internal trace
-log, we calculate the number of new nodes that were executed, as well as the
-\emph{branching depth} where they got inserted.
+log, we calculate the number of \emph{new nodes} that were executed, as well as
+the \emph{branching depth} where they got inserted.
 %
@@
 hundred?
-%
-Using too little can put finding bugs at risk.
-%
-For instance, when the system under test branches based on numeric values, we
-should make sure that we set enough random mutations to test each branch on a
-reasonable basis.
-%
-Using too many, on the other hand, can degrade the performance of the testing
-loop, as it will spend too much time producing uninteresting mutations.
-%
-This can happen for instance if the subexpressions defined using random mutants
-are only used as payloads, and their value does not affect the execution in any
-way.
-
+%% %
+%% Using too little can put finding bugs at risk.
+%% %
+%% For instance, when the system under test branches based on numeric values, we
+%% should make sure that we set enough random mutations to test each branch on a
+%% reasonable basis.
+%% %
+%% Using too many, on the other hand, can degrade the performance of the testing
+%% loop, as it will spend too much time producing uninteresting mutations.
+%% %
+%% This can happen for instance if the subexpressions defined using random mutants
+%% are only used as payloads, and their value does not affect the execution in any
+%% way.
+%
 Answering this question precisely is not an easy task, and the second heuristic

### 2021-08-20 papers/icse22/main.tex
Re-structured intro and background sections

@@
 \usepackage{enumitem}
+%% \usepackage[square,numbers]{natbib}
+
+\settopmatter{printacmref=false}
 
@@
 %% the next command will enable that style.
-\citestyle{acmauthoryear}
+%% \citestyle{authoryear}
 
@@
 works by keeping a queue of \emph{interesting} previously executed test cases
-that can be mutated to produce new ones.
-%
-Mutations are achieved using high-level, structure-preserving transformations
-that can be derived directly from the codebase.
+that can be mutated using high-level, structure-preserving transformations to
+produce new ones.
+%
+%% Mutations are achieved using high-level, structure-preserving transformations
+%% that can be derived directly from the codebase.
 %
@@
 %
-In this light, CGPT is likely to preserve the semantic structure of the mutated
-data, as mutations are applied directly at the data type level, e.g., the
-program's AST in case of generating code.
+In addition, high-level, structure-preserving mutators are better suited for
+producing syntactically valid mutants, avoiding the time wasted by using generic
+low-level mutators that act over the ``serialized'' test cases and know very
+little about their structure, thus producing syntactically broken mutants most
+of the time.
+%
+This grammar-aware mutation technique has shown to be quite useful when fuzzing
+systems accepting structurally complex inputs \cite{holler2012fuzzing,
+  wang2019superion, xsmith}.
+%
+While most of the tools following this itechnique use existing grammars to
+tailor the mutators to the specific input structure used by the system under
+test, external grammars are not required in CGPT.
+%
+%% Tools like LangFuzz \cite{holler2012fuzzing}, Superion \cite{wang2019superion},
+%% XSmith \cite{xsmith} use existing grammars to tailor the mutators to the
+%% specific input structure used by the system under test.
+%
+%% In CGPT, external grammars are not required.
+%
+The data types used by the inputs of the testing properties already describe the
+structure of the random data we want to mutate in a concrete manner, and
+specialized mutators acting at the data constructor level can be automatically
+derived directly from their definition --- making strongly-typed programming
+languages with expressive type systems an ideal match for CGPT.
+
+%% In this light, CGPT is likely to preserve the semantic structure of the mutated
+%% data, as mutations are applied directly at the data type level, e.g., the
+%% program's AST in case of generating code.
 %
@@
 In the IFC stack machine case study, we compare \mutagen against \fuzzchick, the
-reference CGPT implementation by \citeauthor{lampropoulos2019coverage}
-implemented in Coq.
+reference CGPT implementation developed in Coq by
+\citeauthor{lampropoulos2019coverage}.
 %
@@
 
-In this section, we describe in detail the motivation, ideas and limitations
-behind coverage-guided, property-based testing as depicted by
-\citeauthor{lampropoulos2019coverage}.
+In this section, we briefly describe the motivation, ideas and limitations
+behind CGPT as depicted by \citeauthor{lampropoulos2019coverage}.
 %
@@
 %
-In \quickcheck, this is commonly done via the type class mechanism
-\cite{jones1997type}, instantiating the \texttt{Arbitrary} type class with a
-random generator as the implementation of the overloaded \texttt{arbitrary}
-operation:
+%% In \quickcheck, this is commonly done via the type class mechanism
+%% \cite{jones1997type}, instantiating the \texttt{Arbitrary} type class with a
+%% random generator as the implementation of the overloaded \texttt{arbitrary}
+%% operation:
+A simple implementation of such a generator (simplified to make it more
+accessible%
+\footnote{In practice, this code will most likely be implemented using Haskell's
+  type-class mechanism, hiding the complexity behind the size limit and the payload generator.}%
+) can be written as follows:
+
+%% \begin{verbatim}
+%% instance Arbitrary a => Arbitrary (Tree a) where
+%%   arbitrary = sized gen where
+%%   gen 0 = Leaf <$> arbitrary
+%%   gen n = oneof [ Leaf <$> arbitrary
+%%                 , Branch <$> gen (n-1) <*> arbitrary <*> gen (n-1) ]
+%% \end{verbatim}
 
 \begin{verbatim}
-instance Arbitrary a => Arbitrary (Tree a) where
-  arbitrary = sized gen where
-  gen 0 = Leaf <$> arbitrary
-  gen n = oneof [ Leaf <$> arbitrary
-                , Branch <$> gen (n-1) <*> arbitrary <*> gen (n-1) ]
+gen_tree :: Size -> Gen a -> Gen (Tree a)
+gen_tree gen_a 0 = Leaf <$> gen_a
+gen_tree gen_a n = oneof
+  [ Leaf <$> gen_a
+  , Branch <$> gen_tree (n-1) <*> gen_a <*> gen_tree (n-1) ]
 \end{verbatim}
 
-\noindent This definition follows a simple type-directed generation approach,
+\noindent This definition follows a common type-directed generation approach,
 and it is a good example of what to expect from a random generator synthesized
@@
 \texttt{Leaf} or a \texttt{Branch}), and then calls itself for every recursive
-subexpression needed to produce a well-typed value, carefully reducing the depth
-limit (controlled by \quickcheck) by a unit on each step, ensuring termination.
-%
-Subexpressions of any other type are generated simply by delegating the task to
-their corresponding \texttt{Arbitrary} instance --- as noted by the use of
-\texttt{arbitrary} when generating random payloads inside of the \texttt{Leaf}
-and \texttt{Branch} constructors.
-
-Readers familiar with RBPT will recognize that attempting to test this property
-using \quickcheck directly does not work well.
+subexpression needed to produce a well-typed value, carefully reducing the size
+limit by a unit on each step, ensuring termination by restricting itself to
+generating leaves when it reaches zero.
+%
+Payloads are generated simply by delegating the task to an external random
+generator (\texttt{gen\_a}) defined elsewhere.
+%
+
+%% Subexpressions of any other type are generated
+%% simply by delegating the task to their corresponding \texttt{Arbitrary} instance
+%% --- as noted by the use of \texttt{arbitrary} when generating random payloads
+%% inside of the \texttt{Leaf} and \texttt{Branch} constructors.
+
+Readers familiar with RBPT will recognize that attempting to test
+\texttt{prop\_insert\_inv} using this generator in \quickcheck directly does not
+work well.
 %
@@
 
-The next subsection show how coverage-guided, property-based testing can help
-overcoming this limitation without having to write specialized random generators
-by hand.
+The next subsection show how GCPT helps overcoming this limitation without
+having to write specialized random generators by hand.
 
@@
 To alleviate the problem of testing properties with non-trivial preconditions
-while using automatically derived random generators,
-\citeauthor{lampropoulos2019coverage} introduced \emph{coverage-guided,
-  property-based testing} (CGPT), a technique that enhances the testing process
-with two key characteristics:
+while using automatically derived random generators, CGPT
+%
+%% \citeauthor{lampropoulos2019coverage} introduced \emph{coverage-guided,
+%%   property-based testing} (CGPT), a technique that
+%
+enhances the testing process with two key characteristics:
 %
@@
 \item \emph{high-level, structure-preserving mutations}, to produce
-  syntactically valid test cases by altering existing ones at the data type
-  level.
+  syntactically valid test cases by altering existing ones at the data
+  constructor level.
 \end{inparaenum}
@@
 %
-CGPT, however, does this in a clever way.
-%
-Instead of mutating any previously executed test case that discovers a new part
-of the code, this technique integrates these fuzzing mechanisms into the PBT
-testing loop itself.
-
-Since it is possible to distinguish semantically valid test cases from invalid
-ones, i.e., those passing the testing property preconditions as opposed to those
-that are discarded early, CGPT exploits this information to prefer mutating
-valid test cases than discarded ones.
-
-In addition, high-level, structure-preserving mutators are better suited for
-producing syntactically valid mutants, avoiding the time wasted by using generic
-low-level mutators that act over the ``serialized'' test cases and know very
-little about their structure, thus producing syntactically broken mutants most
-of the time.
-%
-This grammar-aware mutation technique has shown to be quite useful when fuzzing
-systems accepting structurally complex inputs.
-%
-Tools like LangFuzz \cite{holler2012fuzzing}, Superion \cite{wang2019superion},
-XSmith \cite{xsmith} use existing grammars to tailor the mutators to the
-specific input structure used by the system under test.
-%
-In CGPT, external grammars are not required.
-%
-The data types used by the inputs of the testing properties already describe the
-structure of the random data we want to mutate in a concrete manner, and
-specialized mutators acting at the data constructor level can be automatically
-derived directly from their definition.
-
-%% The next subsections describe the CGPT testing loop and structure-preserving
-%% mutations in detail.
-
-%% \subsubsection{Testing loop}
-
-\paragraph{Testing Loop}
-Outlined in Algorithm \ref{algo:fuzzchick:loop}, CGPT starts by creating two
-queues, \textit{QSucc} and \textit{QDisc} for valid and discarded previously
-executed test cases, respectively.
-%
-Enqueued values are stored along with a given mutation budget, that controls how
-many times a given test case can be mutated before being finally thrown away.
-
-Once inside of the main loop, the algorithm picks the next test case using a
-simple criterion: if there are valid values enqueued for mutation, it picks the
-first one, mutates it and returns it, decreasing its energy by one.
-%
-If \textit{QSucc} is empty, then the same is attempted using \textit{QDisc}.
-%
-If none of the mutation queues contain any candidates, CGPT generates a new
-value from scratch.
-%
-This selection process is illustrated in detail in Algorithm
-\ref{algo:fuzzchick:pick}.
-
-Having selected the next test case, the main loop proceeds to execute it,
-capturing both the result (passed, discarded due to preconditions, or failed)
-and its execution trace over the system under test.
-%
-If the test case fails, it is immediately reported as a bug.
-%
-If not, the algorithm evaluates whether it was interesting (i.e., it exercises a
-new path) based on its trace information and the one from previously executed
-test cases (represented by \textit{TLog}).
+%% Instead of mutating any previously executed test case that discovers a new part
+%% of the code, this technique integrates these fuzzing mechanisms into the PBT
+%% testing loop itself.
+%
+In contrast with traditional fuzzing approaches, CGPT can distinguish
+semantically valid test cases from invalid ones, i.e., those passing the testing
+property preconditions as opposed to those that are discarded early.
+%
+This information is used to favour mutating valid test cases over discarded
+ones.
+
+%% Since it is possible to distinguish semantically valid test cases from invalid
+%% ones, i.e., those passing the testing property preconditions as opposed to those
+%% that are discarded early, CGPT exploits this information to prefer mutating
+%% valid test cases than discarded ones.
+
+The testing loop in CGPT uses two queues to store valid and discarded previously
+executed test cases.
+%
+Enqueued test cases are stored along with a given mutation budget, that controls
+how many times a given test case can be mutated before being finally thrown
+away.
+%
+This budget is calculated using a heuristic derived from AFL's power schedule,
+i.e., more budget to test cases that lead to shorter executions, or that
+discover more parts of the code.
+%
+On each iteration, the algorithm selects the next test case by mutating the
+first value on the queue of valid test cases or, alternatively, from the queue
+of discarded test cases if the former one is empty.
+%
+If none of the queues contain test cases to mutate, CGPT generates a new value
+from scratch.
+%
+Then, the algorithm run this test case and evaluates whether it was interesting
+(i.e., it exercises a new path) based on its trace information.
 %
 If the test case does in fact discover a new path, it is enqueued at the end of
-its corresponding queue, depending on whether it passed or was discarded.
+its corresponding queue, depending on whether it passed the property
+precondition or was discarded.
 %
@@
 This process alternates between generation and mutation until a bug is found or
-we reach the test limit.
-
-%% The energy assigned to each test case follows that of AFL's power schedule: more
-%% energy to test cases that lead to shorter executions, or that discover more
-%% parts of the code.
-%% %
-%% Moreover, to favor mutating interesting valid test cases, they get more energy
-%% than those that were discarded.
-
-\begin{algorithm}
-  \SetArgSty{textnormal}
-  %% \SetInd{0em}{0.75em}
-  \SetKw{KwNot}{not}
-  \SetAlgoLined
-  \DontPrintSemicolon
-  \SetKwFunction{FC}{Loop}
-  \SetKwProg{Fn}{Function}{:}{}
-  \Fn{\FC{P, N, gen, mut}}{
-    i $\gets$ 0\;
-    TLog, QSucc, QDisc $\gets$ $\varnothing$\;
-    \While{i $<$ N}{
-      x $\gets$ Pick(QSucc, QDisc, gen, mut)\;
-      (result, trace) $\gets$ WithTrace(P(x))\;
-      \lIf{\KwNot result}{\KwRet Bug(x)}
-      \If{Interesting(TLog, trace)}{
-        e $\gets$ Energy(TLog, x, trace)\;
-        \eIf{\KwNot Discarded(result)}{
-          Enqueue(QSucc, (x, e))\;
-        }{
-          Enqueue(QDisc, (x, e))\;
-        }
-      }
-      i $\gets$ i+1\;
-    }
-    \KwRet Ok\;
-  }
-\caption{\label{algo:fuzzchick:loop}CGPT Testing Loop}
-\end{algorithm}
-
-\begin{algorithm}
-  \SetArgSty{textnormal}
-  %% \SetInd{0em}{0.75em}
-  \SetKw{KwNot}{not}
-  \SetAlgoLined
-  \DontPrintSemicolon
-  \SetKwFunction{Pick}{Pick}
-  \SetKwProg{Fn}{Function}{:}{}
-  \Fn{\Pick{QSucc, QDisc, gen, mut}}{
-    \If{\KwNot Empty(QSucc)}{
-      (x,e) $\gets$ Deque(QSucc)\;
-      \If{e $>$ 0}{
-        PushFront(QSucc, (x, e-1))
-      }
-      \KwRet Sample(mut(x))\;
-    }
-    \uElseIf{\KwNot Empty(QDisc)}{
-      (x,e) $\gets$ Deque(QDisc)\;
-      \If{e $>$ 0}{
-        PushFront(QDisc, (x, e-1))
-      }
-      \KwRet Sample(mut(x))\;
-    }
-    \lElse{\KwRet Sample(gen)}
-  }
-\caption{\label{algo:fuzzchick:pick}CGPT Seed Selection}
-\end{algorithm}
-
-%% \subsubsection{Type-preserving mutations}
-
-\paragraph{Structure-preserving Mutators}
-
-%% are no more than specialized random generators, parameterized
-%% by the original input to be mutated.
-%% %
-%% They
-
-Mutators in GCPT use a simple set of mutation operations that are randomly
-applied to their inputs at the data type level.
-%
-In simple terms, these operations encompass
-%
-\begin{inparaenum}
-\item \emph{swapping a data constructor}, replacing its outermost data
-  constructor by another one, e.g., a \texttt{Leaf} by a \texttt{Branch} and
-  vice-versa, either dropping unused subexpressions or generating new random
-  ones when the new data constructor contains extra fields that cannot be filled
-  using existing ones;
-%% \item \emph{shrinking the value}, replacing its outermost data constructor with
-%%   one that contains a subset of its fields, reusing existing subexpressions;
-%% \item \emph{growing the value}, replacing its outermost data constructor with
-%%   one that contains a superset of its fields, reusing existing subexpressions
-%%   and generating random ones when needed;
-\item \emph{returning a subexpression of the same type}, e.g, returning a left
-  or right subtree when applied to a \texttt{Branch}; and
-\item \emph{mutating recursively}, applying a mutation operation over an
-  immediate subexpression.
-\end{inparaenum}
-%
-We will provide more details about structure-preserving mutations in Section
-\ref{sec:mutagen}.
-
-%% \begin{figure}
-%% \vspace{-5pt}
-%% \begin{verbatim}
-%% mutate_tree :: (a -> Gen a) -> Tree a -> Gen (Tree a)
-%% mutate_tree mutate_a (Leaf x) =
-%%   oneof [ do { x' <- mutate_a x; return (Leaf x')}                       -- Mutate recursively
-%%         , do { l <- arbitrary; r <- arbitrary; return (Branch l x r) } ] -- Grow constructor
-%% mutate_tree mutate_a (Branch l x r) =
-%%   oneof [ return l                                                       -- Return subexpression
-%%         , return r                                                       -- Return subexpression
-%%         , return (Leaf x)                                                -- Shrink constructor
-%%         , do { l' <- mutate_tree l; return (Branch l' x r) }             -- Mutate recursively
-%%         , do { x' <- mutate_a    x; return (Branch l x' r) }             -- Mutate recursively
-%%         , do { r' <- mutate_tree r; return (Branch l x r') } ]           -- Mutate recursively
-%% \end{verbatim}
-%% \caption{\label{fig:fuzzchick:mutator}\fuzzchick mutator for the \texttt{Tree} data type. }
-%% \end{figure}
-
-%% \begin{figure}
-%% %% \vspace{-5pt}
-%% \begin{verbatim}
-%% mutate_tree :: (a -> Gen a) -> Tree a -> Gen (Tree a)
-%% mutate_tree mutate_a (Leaf x) =
-%%   oneof [ Leaf <$> mutate_a x
-%%         , Branch <$> arbitrary <*> return x <*> arbitrary ]
-%% mutate_tree mutate_a (Branch l x r) =
-%%   oneof [ return l
-%%         , return r
-%%         , Leaf   <$> return x
-%%         , Branch <$> mutate_tree l <*> return x   <*> return r
-%%         , Branch <$> return l      <*> mutate_a x <*> return r
-%%         , Branch <$> return l      <*> return x   <*> mutate_tree r
-%% \end{verbatim}
-%% \caption{\label{fig:fuzzchick:mutator}\fuzzchick mutator for the \texttt{Tree} data type. }
-%% \end{figure}
-
-%% Fig. \ref{fig:fuzzchick:mutator} illustrates a \fuzzchick-like mutator for our
-%% previously used \texttt{Tree} data type example.
-%% %
-%% Since trees are parametric, for clarity this definition is also parameterized by
-%% a mutator for the payload (\texttt{mutate\_a}), although this can be abstracted
-%% away using the type class system.
-
-%% %% (where \texttt{mutate\_tree} and
-%% %% \texttt{mutate\_a} can be transparently coerced into an overloaded
-%% %% \texttt{mutate} function.)
-
-%% In this mutator, branches can be shrunk into leaves by dropping the subtrees,
-%% whereas leaves can grow into branches, by reusing the payload and generating two
-%% random subtrees.
-%% %
-%% Moreover, branches can be replaced with one of their subtrees.
-%% %
-%% Finally, mutations can be recursively applied over both the payload and the
-%% subtrees.
-%% %
-%% At the top level, all these operations are put together using the \texttt{oneof}
-%% combinator that randomly picks one of them with uniform probability.
+it reaches the test limit.
+
+%% %
+%% Enqueued values are stored along with a given mutation budget, that controls how
+%% many times a given test case can be mutated before being finally thrown away.
+
+%% Once inside of the main loop, the algorithm picks the next test case using a
+%% simple criterion: if there are valid values enqueued for mutation, it picks the
+%% first one, mutates it and returns it, decreasing its energy by one.
+%% %
+%% If \textit{QSucc} is empty, then the same is attempted using \textit{QDisc}.
+%% %
+%% If none of the mutation queues contain any candidates, CGPT generates a new
+%% value from scratch.
+%% %
+%% This selection process is illustrated in detail in Algorithm
+%% \ref{algo:fuzzchick:pick}.
+
+%% Having selected the next test case, the main loop proceeds to execute it,
+%% capturing both the result (passed, discarded due to preconditions, or failed)
+%% and its execution trace over the system under test.
+%% %
+%% If the test case fails, it is immediately reported as a bug.
+%% %
+%% If not, the algorithm evaluates whether it was interesting (i.e., it exercises a
+%% new path) based on its trace information and the one from previously executed
+%% test cases (represented by \textit{TLog}).
+%% %
+%% If the test case does in fact discover a new path, it is enqueued at the end of
+%% its corresponding queue, depending on whether it passed or was discarded.
+%% %
+%% %% (i.e., it failed the precondition).
+%% %
+%% This process alternates between generation and mutation until a bug is found or
+%% we reach the test limit.
+
+%% In addition, high-level, structure-preserving mutators are better suited for
+%% producing syntactically valid mutants, avoiding the time wasted by using generic
+%% low-level mutators that act over the ``serialized'' test cases and know very
+%% little about their structure, thus producing syntactically broken mutants most
+%% of the time.
+%% %
+%% This grammar-aware mutation technique has shown to be quite useful when fuzzing
+%% systems accepting structurally complex inputs.
+%% %
+%% Tools like LangFuzz \cite{holler2012fuzzing}, Superion \cite{wang2019superion},
+%% XSmith \cite{xsmith} use existing grammars to tailor the mutators to the
+%% specific input structure used by the system under test.
+%% %
+%% In CGPT, external grammars are not required.
+%% %
+%% The data types used by the inputs of the testing properties already describe the
+%% structure of the random data we want to mutate in a concrete manner, and
+%% specialized mutators acting at the data constructor level can be automatically
+%% derived directly from their definition.
+
+%% \paragraph{Testing Loop}
+%% Outlined in Algorithm \ref{algo:fuzzchick:loop}, CGPT starts by creating two
+%% queues, \textit{QSucc} and \textit{QDisc} for valid and discarded previously
+%% executed test cases, respectively.
+%% %
+%% Enqueued values are stored along with a given mutation budget, that controls how
+%% many times a given test case can be mutated before being finally thrown away.
+
+%% Once inside of the main loop, the algorithm picks the next test case using a
+%% simple criterion: if there are valid values enqueued for mutation, it picks the
+%% first one, mutates it and returns it, decreasing its energy by one.
+%% %
+%% If \textit{QSucc} is empty, then the same is attempted using \textit{QDisc}.
+%% %
+%% If none of the mutation queues contain any candidates, CGPT generates a new
+%% value from scratch.
+%% %
+%% This selection process is illustrated in detail in Algorithm
+%% \ref{algo:fuzzchick:pick}.
+
+%% Having selected the next test case, the main loop proceeds to execute it,
+%% capturing both the result (passed, discarded due to preconditions, or failed)
+%% and its execution trace over the system under test.
+%% %
+%% If the test case fails, it is immediately reported as a bug.
+%% %
+%% If not, the algorithm evaluates whether it was interesting (i.e., it exercises a
+%% new path) based on its trace information and the one from previously executed
+%% test cases (represented by \textit{TLog}).
+%% %
+%% If the test case does in fact discover a new path, it is enqueued at the end of
+%% its corresponding queue, depending on whether it passed or was discarded.
+%% %
+%% %% (i.e., it failed the precondition).
+%% %
+%% This process alternates between generation and mutation until a bug is found or
+%% we reach the test limit.
+
+%% \begin{algorithm}
+%%   \SetArgSty{textnormal}
+%%   %% \SetInd{0em}{0.75em}
+%%   \SetKw{KwNot}{not}
+%%   \SetAlgoLined
+%%   \DontPrintSemicolon
+%%   \SetKwFunction{FC}{Loop}
+%%   \SetKwProg{Fn}{Function}{:}{}
+%%   \Fn{\FC{P, N, gen, mut}}{
+%%     i $\gets$ 0\;
+%%     TLog, QSucc, QDisc $\gets$ $\varnothing$\;
+%%     \While{i $<$ N}{
+%%       x $\gets$ Pick(QSucc, QDisc, gen, mut)\;
+%%       (result, trace) $\gets$ WithTrace(P(x))\;
+%%       \lIf{\KwNot result}{\KwRet Bug(x)}
+%%       \If{Interesting(TLog, trace)}{
+%%         e $\gets$ Energy(TLog, x, trace)\;
+%%         \eIf{\KwNot Discarded(result)}{
+%%           Enqueue(QSucc, (x, e))\;
+%%         }{
+%%           Enqueue(QDisc, (x, e))\;
+%%         }
+%%       }
+%%       i $\gets$ i+1\;
+%%     }
+%%     \KwRet Ok\;
+%%   }
+%% \caption{\label{algo:fuzzchick:loop}CGPT Testing Loop}
+%% \end{algorithm}
+
+%% \begin{algorithm}
+%%   \SetArgSty{textnormal}
+%%   %% \SetInd{0em}{0.75em}
+%%   \SetKw{KwNot}{not}
+%%   \SetAlgoLined
+%%   \DontPrintSemicolon
+%%   \SetKwFunction{Pick}{Pick}
+%%   \SetKwProg{Fn}{Function}{:}{}
+%%   \Fn{\Pick{QSucc, QDisc, gen, mut}}{
+%%     \If{\KwNot Empty(QSucc)}{
+%%       (x,e) $\gets$ Deque(QSucc)\;
+%%       \If{e $>$ 0}{
+%%         PushFront(QSucc, (x, e-1))
+%%       }
+%%       \KwRet Sample(mut(x))\;
+%%     }
+%%     \uElseIf{\KwNot Empty(QDisc)}{
+%%       (x,e) $\gets$ Deque(QDisc)\;
+%%       \If{e $>$ 0}{
+%%         PushFront(QDisc, (x, e-1))
+%%       }
+%%       \KwRet Sample(mut(x))\;
+%%     }
+%%     \lElse{\KwRet Sample(gen)}
+%%   }
+%% \caption{\label{algo:fuzzchick:pick}CGPT Seed Selection}
+%% \end{algorithm}
+
+%% %% \subsubsection{Type-preserving mutations}
+
+%% \paragraph{Structure-preserving Mutators}
+
+%% %% are no more than specialized random generators, parameterized
+%% %% by the original input to be mutated.
+%% %% %
+%% %% They
+
+%% Mutators in GCPT use a simple set of mutation operations that are randomly
+%% applied to their inputs at the data type level.
+%% %
+%% In simple terms, these operations encompass
+%% %
+%% \begin{inparaenum}
+%% \item \emph{swapping a data constructor}, replacing its outermost data
+%%   constructor by another one, e.g., a \texttt{Leaf} by a \texttt{Branch} and
+%%   vice-versa, either dropping unused subexpressions or generating new random
+%%   ones when the new data constructor contains extra fields that cannot be filled
+%%   using existing ones;
+%% %% \item \emph{shrinking the value}, replacing its outermost data constructor with
+%% %%   one that contains a subset of its fields, reusing existing subexpressions;
+%% %% \item \emph{growing the value}, replacing its outermost data constructor with
+%% %%   one that contains a superset of its fields, reusing existing subexpressions
+%% %%   and generating random ones when needed;
+%% \item \emph{returning a subexpression of the same type}, e.g, returning a left
+%%   or right subtree when applied to a \texttt{Branch}; and
+%% \item \emph{mutating recursively}, applying a mutation operation over an
+%%   immediate subexpression.
+%% \end{inparaenum}
+%% %
+%% We will provide more details about structure-preserving mutations in Section
+%% \ref{sec:mutagen}.
 
@@
 %
-If it assigns too much energy to certain not-so-interesting seeds, some bugs
+If it assigns too much budget to certain not-so-interesting seeds, some bugs
 might not be discovered on a timely basis.
 %
-Conversely, assigning too little energy to interesting test cases might mean
-that some bugs cannot be discovered at all unless the right mutation happens
-within the small available energy window --- randomly generating the same test
+Conversely, assigning too little budget to interesting test cases might mean
+that some bugs cannot be discovered at all --- randomly generating the same test
 case again later does not help, as it becomes uninteresting based on historic
@@
 \bibliographystyle{ACM-Reference-Format}
+%% \bibliographystyle{plainnat}
 \bibliography{references}

### 2021-08-23 papers/icse22/main.tex
Pass over heuristics

@@
 %% %
-%% If \textit{QSucc} is empty, then the same is attempted using \textit{QDisc}.
+%% If \textit{QValid} is empty, then the same is attempted using \textit{QDisc}.
 %% %
@@
 %% Outlined in Algorithm \ref{algo:fuzzchick:loop}, CGPT starts by creating two
-%% queues, \textit{QSucc} and \textit{QDisc} for valid and discarded previously
+%% queues, \textit{QValid} and \textit{QDisc} for valid and discarded previously
 %% executed test cases, respectively.
@@
 %%     i $\gets$ 0\;
-%%     TLog, QSucc, QDisc $\gets$ $\varnothing$\;
+%%     TLog, QValid, QDisc $\gets$ $\varnothing$\;
 %%     \While{i $<$ N}{
-%%       x $\gets$ Pick(QSucc, QDisc, gen, mut)\;
+%%       x $\gets$ Pick(QValid, QDisc, gen, mut)\;
 %%       (result, trace) $\gets$ WithTrace(P(x))\;
@@
 %%         \eIf{\KwNot Discarded(result)}{
-%%           Enqueue(QSucc, (x, e))\;
+%%           Enqueue(QValid, (x, e))\;
 %%         }{
@@
 %%   \SetKwProg{Fn}{Function}{:}{}
-%%   \Fn{\Pick{QSucc, QDisc, gen, mut}}{
-%%     \If{\KwNot Empty(QSucc)}{
-%%       (x,e) $\gets$ Deque(QSucc)\;
+%%   \Fn{\Pick{QValid, QDisc, gen, mut}}{
+%%     \If{\KwNot Empty(QValid)}{
+%%       (x,e) $\gets$ Deque(QValid)\;
 %%       \If{e $>$ 0}{
-%%         PushFront(QSucc, (x, e-1))
+%%         PushFront(QValid, (x, e-1))
 %%       }
@@
 
-The next section introduces \mutagen, our CGPT tool that aims to tackle the main
-limitations of CGPT.
+The next section introduces \mutagen, our revised CGPT that aims to tackle the
+main limitations of CGPT.
 
@@
 
-In this section we describe the main ideas behind \mutagen, our enhanced CGPT
-tool written in Haskell.
+In this section we describe the main ideas behind \mutagen, our CGPT tool
+written in Haskell.
 %
@@
 %
-In practice (\texttt{def = Leaf def}), where the inner \texttt{def} is the
-simplest value of the payload --- we use the type class system to abstract this
-complexity away in our implementation.
-%
-Using a small value is again inspired by exhaustive bounded testing tools, and
-avoids introducing unnecessary randomness when growing data constructors.
+In our example, the default value of a \texttt{Tree} is \texttt{Leaf def}, where
+the inner \texttt{def} is the default value of the payload --- we use the type
+class system to abstract this complexity away in our implementation.
+%
+Using a small default value is again inspired by exhaustive bounded testing
+tools, and avoids introducing unnecessary randomness when growing mutated
+expressions.
 
@@
 \begin{verbatim}
-mutate n = [ RAND arbitrary ]
+mutate n = [ RAND gen_int ]
 \end{verbatim}
@@
 %
-As it can be observed, we follow the CGPT testing loop, using two queues to keep
-mutable test cases, and enqueueing and retrieving interesting test cases from
-them before falling back to random generation.
-
-\begin{algorithm}
+Like in CGPT, we use two queues, QValid and QDisc to keep valid and discarded mutable test cases,
+respectively.
+%
+%% The testing loop enqueues and retrieves interesting test cases from them before
+%% falling back to random generation.
+
+\begin{algorithm}[b]
   \SetArgSty{textnormal}
@@
     i $\gets$ 0\;
-    TLog, QSucc, QDisc $\gets$ $\varnothing$\;
+    TLog, QValid, QDisc $\gets$ $\varnothing$\;
     \While{i $<$ N}{
-      x $\gets$ Pick(QSucc, QDisc, gen)\;
+      x $\gets$ Pick(QValid, QDisc, gen)\;
       (result, trace) $\gets$ WithTrace(P(x))\;
@@
           batch $\gets$ Mutate(x, mut, R)\;
-          Enqueue(QSucc, batch)\;
-        }
-        \uElseIf{Passed(Parent(x))}{
+          Enqueue(QValid, batch)\;
+        }
+        \uElseIf{\KwNot Discarded(Parent(x))}{
           batch $\gets$ Mutate(x, mut, R)\;
@@
 
-\begin{algorithm}
+\begin{algorithm}[t]
   \SetArgSty{textnormal}
@@
 
-The difference with CGPT lies in that we precompute all the mutations of a given
-mutable test case before enqueueing them.
+In \mutagen, we precompute all the mutations of a given mutable test case before
+enqueueing them.
 %
@@
 
-Another difference is the criterion for enqueuing discarded tests.
+Having selected the next test case, the main loop proceeds to execute it,
+capturing both the result (valid, discarded due to preconditions, or failed)
+and its execution trace over the system under test.
+%
+If the test case fails, it is immediately reported as a bug.
+%
+If not, the algorithm evaluates whether it was interesting (i.e., it exercises a
+new path) based on its trace information and the one from previously executed
+test cases (represented by \textit{TLog}).
+%
+If the test case does in fact discover a new path, it is enqueued on its
+corresponding queue, depending on whether it was valid or got discarded.
+%
+%% (i.e., it failed the precondition).
+%
+This process alternates between generation and mutation until a bug is found or
+we reach the test limit.
+
+Another difference with CGPT's testing loop is the criterion for enqueuing
+discarded tests.
 %
@@
 
-\begin{algorithm}
+\begin{algorithm}[b]
   \SetArgSty{textnormal}
@@
   \SetKwProg{Fn}{Function}{:}{}
-  \Fn{\Pick{QSucc, QDisc, gen}}{
-    \If{\KwNot Empty(QSucc)}{
-      batch $\gets$ Deque(QSucc)\;
+  \Fn{\Pick{QValid, QDisc, gen}}{
+    \If{\KwNot Empty(QValid)}{
+      batch $\gets$ Deque(QValid)\;
       \If{Empty(batch)}{
-        Pick(QSucc, QDisc, gen)\;
+        Pick(QValid, QDisc, gen)\;
       } \uElse{
-        PushFront(QSucc, Rest(batch))\;
+        PushFront(QValid, Rest(batch))\;
         \KwRet First(batch)\;
@@
     }
-    \If{\KwNot Empty(QSucc)}{
-      batch $\gets$ Deque(QSucc)\;
+    \If{\KwNot Empty(QValid)}{
+      batch $\gets$ Deque(QValid)\;
       \If{Empty(batch)}{
-        Pick(QSucc, QDisc, gen)\;
+        Pick(QValid, QDisc, gen)\;
       } \uElse{
-        PushFront(QSucc, Rest(batch))\;
+        PushFront(QValid, Rest(batch))\;
         \KwRet First(batch)\;
@@
 
-\section{\mutagen Heuristics}
-\label{sec:heuristics}
-
-In this section we introduce two heuristics implemented on top of the base
-testing loop of our tool.
-%
-\mutagen enables them all by default, and can be individually disabled by the
-user if needed.
-
-\subsection{Priority FIFO Scheduling}
-
-This heuristic tackles the issue of enqueuing new interesting mutation
-candidates at the end of (possibly) long queues of not-so-interesting previously
-executed ones.
-
-%% \fuzzchick uses AFL instrumentation under the hood, which in turn uses an
-%% \emph{edge coverage} criterion to distinguish novel executions and to assign
-%% each mutation candidate a given energy.
-%% %
-Execution traces in \mutagen represent the specific \emph{path} in the code
-taken by the program, as opposed to just the (unordered) set of edges traversed
-in the control-flow graph (CFG) used by default by most coverage-guided fuzzers.
-%
-Using this criterion lets us gather precise information from each new execution.
-%
-%% To keep track of the execution path of each test, our tool keeps an internal
-%% prefix tree that records every new entry by traversing this data structure and
-%% inserting new nodes after a new path prefix gets discovered.
-%% %
-In particular, we are interested in the \emph{depth} where each new execution
-branches from already seen ones.
-%
-Our assumption here is that test cases that differ (branch) at shallower depths
-from the ones already executed are more likely to discover completely new
-portions of the code under test, and hence we want to prioritize them.
-
-In this light, every time we insert a new execution path into the internal trace
-log, we calculate the number of \emph{new nodes} that were executed, as well as
-the \emph{branching depth} where they got inserted.
-%
-The former is used to distinguish interesting test cases (whether or not new
-nodes were inserted), whereas the latter is used by this heuristic to schedule
-mutation candidates.
-%
-Fig. \ref{fig:mutagen:tracelog} illustrates this idea, inserting two execution
-traces (one after another) into a trace log that initially contains a single
-execution path.
-%
-The second insertion (with trace $1 \rightarrow 2 \rightarrow 6 \rightarrow 7$)
-branches at a shallower depth than the first one (2 vs. 3), hence its
-corresponding test case should be given a higher priority.
-
-\begin{figure}[t]
+\begin{figure*}[t]
 %% \vspace{-15pt}
 \newcommand\stacked[2]{\underset{\textstyle #2\mathstrut}{#1}}
-\begin{equation}
+\begin{align*}
   \nonumber
+  \qquad\qquad\qquad\qquad
   \left[
@@
   \stacked{{\color{blue} new=1\phantom{x}}}{{\color{red} depth=3}}
-  \end{equation}
-  \begin{equation}
-  \nonumber
+  %% \end{equation}
+  %% \begin{equation}
+  %% \nonumber
   %% \\
-  %% \qquad
-  %% \qquad
+  \qquad
+  \qquad
   \left[
@@
   \stacked{{\color{blue} new=2\phantom{x}}}{{\color{red} depth=2}}
-\end{equation}
+\end{align*}
 %% \vspace{-10pt}
@@
   internal trace log (represented using brackets).}
-\end{figure}
-
-With this mechanism in place, we can modify \mutagen's base testing loop by
-replacing each mutation queue with a priority queue indexed by the branching
-depth of each new execution.
-%
-These changes are illustrated in Algorithm \ref{algo:mutagen:fifo}.
-%
-Statements in {\color{red} red} indicate important changes to the base
-algorithm, whereas ellipses denote parts of the code that are not relevant for
-the point being made.
-
-To pick the next test case, we simply retrieve the one with the highest priority
-(the smallest branching depth).
-%
-Then, whenever we find a new interesting test case, we enqueue it at the
-beginning of the queue of its corresponding priority.
-%
-This allows the testing loop to jump immediately onto processing new interesting
-candidates as soon as they are found (even at the same priority), and to jump
-back to previous candidates as soon as their mutants become progressively less
-novel.
-
-\begin{algorithm}
+\end{figure*}
+
+\section{\mutagen Heuristics}
+\label{sec:heuristics}
+
+In this section we introduce two heuristics implemented on top of the base
+testing loop of our tool.
+%% %
+%% \mutagen enables them by default, and can be individually disabled by the user
+%% if needed.
+
+\subsection{Priority FIFO Scheduling}
+
+This heuristic tackles the issue of enqueuing new interesting mutation
+candidates at the end of (possibly) long queues of not-so-interesting previously
+executed ones.
+
+%% \fuzzchick uses AFL instrumentation under the hood, which in turn uses an
+%% \emph{edge coverage} criterion to distinguish novel executions and to assign
+%% each mutation candidate a given energy.
+%% %
+Execution traces in \mutagen represent the specific \emph{path} in the code
+taken by the program, as opposed to just the (unordered) set of edges traversed
+in the control-flow graph (CFG) used by default by most coverage-guided fuzzers.
+%
+Using this criterion lets us gather precise information from each new execution.
+%
+%% To keep track of the execution path of each test, our tool keeps an internal
+%% prefix tree that records every new entry by traversing this data structure and
+%% inserting new nodes after a new path prefix gets discovered.
+%% %
+In particular, we are interested in the \emph{depth} where each new execution
+branches from already seen ones.
+%
+Our assumption is that test cases that differ (branch) at shallower depths from
+the ones already executed are more likely to discover new portions of the code
+under test, and hence we want to prioritize them.
+
+\begin{algorithm}[b]
   \SetArgSty{textnormal}
@@
     $\cdots$\;
-    x $\gets$ Pick(QSucc, QDisc, gen, mut)\;
+    x $\gets$ Pick(QValid, QDisc, gen, mut)\;
     (result, trace) $\gets$ WithTrace(P(x))\;
@@
         {\color{red} prio $\gets$ BranchDepth(TLog, trace)}\; %% NEW
-        {\color{red} PushFront(QSucc, prio, batch)}\;          %% NEW
+        {\color{red} PushFront(QValid, prio, batch)}\;          %% NEW
       }
@@
   \SetKwProg{Fn}{Function}{:}{}
-  \Fn{\Pick{QSucc, QDisc, gen}}{
-    \If{\KwNot Empty(QSucc)}{
-      {\color{red} (batch, prio) $\gets$ DequeMin(QSucc)}\; %%NEW
+  \Fn{\Pick{QValid, QDisc, gen}}{
+    \If{\KwNot Empty(QValid)}{
+      {\color{red} (batch, prio) $\gets$ DequeMin(QValid)}\; %%NEW
       \If{Empty(batch)}{
-        Pick(QSucc, QDisc, gen)\;
+        Pick(QValid, QDisc, gen)\;
       } \uElse{
-        {\color{red} PushFront(QSucc, prio, Rest(batch))}\; %% NEW
+        {\color{red} PushFront(QValid, prio, Rest(batch))}\; %% NEW
         \KwRet First(batch)\;
@@
 \end{algorithm}
+
+In this light, every time we insert a new execution path into the internal trace
+log, we calculate the number of \emph{new nodes} that were executed, as well as
+the \emph{branching depth} where they got inserted.
+%
+The former is used to distinguish interesting test cases (whether or not new
+nodes were inserted), whereas the latter is used by this heuristic to schedule
+mutation candidates.
+%
+Fig. \ref{fig:mutagen:tracelog} illustrates this idea, inserting two execution
+traces (one after another) into a trace log that initially contains a single
+execution path.
+%
+The second insertion (with trace $1 \rightarrow 2 \rightarrow 6 \rightarrow 7$)
+branches at a shallower depth than the first one (2 vs. 3), and so its
+corresponding test case should be given a higher priority.
+
+Using this mechanism, we can modify \mutagen's base testing loop by replacing
+each mutation queue with a priority queue indexed by the branching depth of each
+new execution.
+%
+These changes are illustrated in Algorithm \ref{algo:mutagen:fifo}.
+%
+Statements in {\color{red} red} indicate important changes to the base
+algorithm, whereas ellipses denote parts of the code that are not relevant for
+the point being made.
+
+To pick the next test case, we retrieve the one with the highest priority (the
+smallest branching depth).
+%
+Then, when we find a new interesting test case, it gets enqueued at the
+\emph{beginning} of the queue at its corresponding priority.
+%
+This allows the testing loop to jump immediately onto mutating new interesting
+test cases as soon as they are found (even at the same priority), and to jump
+back to previous test cases as soon as their mutants become less novel.
+
+%% \caption{\label{fig:heuristics} The two main heuristics implemented in \mutagen.
+%%   Statements in {\color{red} red} indicate important changes to the base
+%%   algorithm. Ellipses denote parts of the code that are not relevant for the
+%%   point being made. }
+
+\subsection{Detecting Trace Space Saturation And Tuning Random Mutations Parameter}
+
+As introduced in Section \ref{sec:mutagen}, our tool is parameterized by the
+number of random mutations to be generated over each mutable subexpression
+defined using a random mutant, e.g., numeric values.
+%
+But, how many random mutations should we use? A single one? A few tenths? A few
+hundred?
+%% %
+%% Using too little can put finding bugs at risk.
+%% %
+%% For instance, when the system under test branches based on numeric values, we
+%% should make sure that we set enough random mutations to test each branch on a
+%% reasonable basis.
+%% %
+%% Using too many, on the other hand, can degrade the performance of the testing
+%% loop, as it will spend too much time producing uninteresting mutations.
+%% %
+%% This can happen for instance if the subexpressions defined using random mutants
+%% are only used as payloads, and their value does not affect the execution in any
+%% way.
+%
+Answering this question precisely is not an easy task, and this second heuristic
+aims to tackle this issue.
+
+We found that, the smaller the number of random mutations we set, the easier it
+is for the trace log that records executions to start getting saturated, i.e.,
+when interesting test cases stop getting discovered or are discovered very
+sporadically.
+%
+We realized that we can use this information to automatically optimize the
+number of random mutations used by our tool.
+%
+This idea is described in Algorithm \ref{algo:mutagen:reset}.
+%
+The process is simple:
+%
+\begin{inparaenum}
+  \item we start the testing loop with the number of random mutations set to
+    one,
+  %
+  \item each time we find that a test is not interesting (i.e. boring), we
+    increment a counter,
+  \item if we have not produced any interesting test case after a certain
+    threshold (1000 tests seems to be a reasonable value in practice), we
+    increment the number of random mutations and the threshold by twice the
+    current amount.
+    %
+    Additionally, each time this happens we also reset the trace log, so
+    interesting test cases found on a previous iteration can be found and
+    enqueued for mutation again --- this time with a higher effort dedicated to
+    producing random mutations.
+\end{inparaenum}
+
+Then, if the execution of the system under test depends heavily on the values
+stored at randomly mutable subexpressions, starting with a single random
+mutation will quickly saturate the trace space, and this heuristic will
+continuously increase the random mutations parameter until that stops happening.
 
@@
 \end{algorithm}
-
-%% \caption{\label{fig:heuristics} The two main heuristics implemented in \mutagen.
-%%   Statements in {\color{red} red} indicate important changes to the base
-%%   algorithm. Ellipses denote parts of the code that are not relevant for the
-%%   point being made. }
-
-\subsection{Detecting Trace Space Saturation And Tuning Random Mutations Parameter}
-
-As introduced in Section \ref{sec:mutagen}, our tool is parameterized by the
-number of random mutations to be generated over each mutable subexpression
-defined using a random mutant, e.g., numeric values, characters, etc.
-%
-But, how many random mutations should we use? A single one? A few tenths? A few
-hundred?
-%% %
-%% Using too little can put finding bugs at risk.
-%% %
-%% For instance, when the system under test branches based on numeric values, we
-%% should make sure that we set enough random mutations to test each branch on a
-%% reasonable basis.
-%% %
-%% Using too many, on the other hand, can degrade the performance of the testing
-%% loop, as it will spend too much time producing uninteresting mutations.
-%% %
-%% This can happen for instance if the subexpressions defined using random mutants
-%% are only used as payloads, and their value does not affect the execution in any
-%% way.
-%
-Answering this question precisely is not an easy task, and the second heuristic
-we introduce in this work aims to tackle this issue.
-%
-We found that, the smaller the number of random mutations we set, the easier it
-is for the trace log that records executions to start getting saturated, i.e.,
-when interesting test cases stop getting discovered or are discovered very
-sporadically.
-%
-We realized that we can use this information to automatically optimize the
-number of random mutations used by our tool.
-%
-This idea is described in Algorithm \ref{algo:mutagen:reset}.
-%
-The process is simple:
-%
-\begin{inparaenum}
-  \item we start the testing loop with the number of random mutations set to
-    one,
-  %
-  \item each time we find that a test is not interesting (i.e. boring), we
-    increment a counter,
-  \item if we have not produced any interesting test case after a certain
-    threshold (1000 tests seems to be a reasonable value in practice), we
-    increment the number of random mutations and the threshold by twice the
-    current amount.
-    %
-    Additionally, each time this happens we also reset the trace log, so
-    interesting test cases found on a previous iteration can be found and
-    enqueued for mutation again --- this time with a higher effort dedicated to
-    producing random mutations.
-\end{inparaenum}
-
-Then, if the execution of the system under test depends heavily on the values
-stored at randomly mutable subexpressions, starting with a single random
-mutation will quickly saturate the trace space, and this heuristic will
-continuously increase the random mutations parameter until that stops happening.
 

### 2021-08-23 papers/icse22/main.tex
Pass over case studies (Section 5)

@@
 
-We evaluated the performance of \mutagen using two main case studies.
+We evaluated the performance of \mutagen using two case studies.
 %
@@
 %
-The implementation of this case study was originally proven correct by
+Its implementation was originally proven correct by
 \citeauthor{10.1145/2578855.2535839} \citeyearpar{10.1145/2578855.2535839} in
@@
 %
-\citeauthor{lampropoulos2019coverage} used this same case study to compare
-\fuzzchick against random testing approaches using na\"ive and manually-written
-smart random generators.
+\citeauthor{lampropoulos2019coverage} used this case study to compare \fuzzchick
+against random testing approaches using na\"ive and manually-written smart
+random generators.
 %
@@
 
-The second case study aims to evaluate \mutagen in a more realistic scenario,
+The second case study aims to evaluate \mutagen in a realistic scenario,
 and focuses on testing \textit{haskell-wasm} \cite{haskellwasm}, an existing
@@
 %
-We manually injected 10 bugs in the validator as well as 5 bugs in the
-interpreter of this engine (see Table \ref{table:wasm:injectedbugs}), and used
-the reference WebAssembly implementation to find them via differential testing
-\cite{mckeeman1998differential}.
-%
-During this process, we quickly discovered 3 bugs and 2 discrepancies on this
-engine with respect to the reference implementation.
-%
-All of them were reported and later confirmed by the authors of
-\textit{haskell-wasm}.
-%
-Sadly, comparing \mutagen with \fuzzchick has unfortunately been out of the
+%
+Sadly, comparing \mutagen with \fuzzchick on this case study has been out of the
 scope of this work.
 %
-The main reason behind this is that porting \textit{haskell-wasm} into Coq, and
-later adapting our test suite to work with \fuzzchick requires a subtantial
-effort.
-%
-Instead, we focus on comparing \mutagen against a more traditional PBT approach
-using \quickcheck, to evaluate the relative overhead of the code instrumentation
-used in our tool.
-
-\begin{table}[t]
-\footnotesize
-\begin{tabular}{|c|c|l|}
-\hline
-\textbf{Id}
-& \textbf{Subsystem}
-& \textbf{Description} \\
-\hline
-1
-& Validator
-& Wrong if-then-else type validation on else branch \\
-\hline
-2
-& Validator
-& Wrong stack type validation \\
-\hline
-3
-& Validator
-& Removed function type mismatch assertion \\
-\hline
-4
-& Validator
-& Removed max memory instances assertion \\
-\hline
-5
-& Validator
-& Removed function index out-of-range assertion \\
-\hline
-6
-& Validator
-& Wrong type validation on \texttt{i64.eqz} instruction \\
-\hline
-7
-& Validator
-& Wrong type validation on \texttt{i32} binary operations \\
-\hline
-8
-& Validator
-& Removed memory index out-of-range assertion \\
-\hline
-9
-& Validator
-& Wrong type validation on \texttt{i64} constants \\
-\hline
-10
-& Validator
-& Removed memory alignment validation on \texttt{i32.load} instruction \\
-\hline
-11
-& Interpreter
-& Wrong interpretation of \texttt{i32.sub} instruction \\
-\hline
-12
-& Interpreter
-& Wrong interpretation of \texttt{i32.lt\_u} instruction \\
-\hline
-13
-& Interpreter
-& Wrong interpretation of \texttt{i32.shr\_u} instruction \\
-\hline
-14
-& Interpreter
-& Wrong local variable initialization \\
-\hline
-15
-& Interpreter
-& Wrong memory address casting on \texttt{i32.load8\_s} instruction \\
-\hline
-\end{tabular}
-\caption{\label{table:wasm:injectedbugs}Bugs injected into
-  \textit{haskell-wasm}.}
-%% \vspace{-15pt}
-\end{table}
+The reason for this is that porting \textit{haskell-wasm} into Coq, and later
+adapting our test suite to work with \fuzzchick requires a subtantial effort.
+%
+Instead, we focus on comparing \mutagen against a more traditional RPBT approach
+using \quickcheck, to evaluate its effectiveness versus the relative overhead of
+the code instrumentation used in our tool.
 
@@
 
-The abstract stack machine used in this case study consists of four main
-elements: a program counter, a stack, and data- and instruction memories.
+This abstract stack machine consists of four main elements: a program counter, a
+stack, and data- and instruction memories.
 %
@@
 %
-Security labels are propagated throughout the execution of the program every
-time the machine executes an instruction.
-%
-There are eight different instructions defined as:
+Security labels are propagated throughout the execution of the program whenever
+the machine executes an instruction.
+%
+Such instructions are defined as:
 %
@@
 \citeauthor{hritcu2013testing} \citeyearpar{hritcu2013testing,
-  hrictcu2016testing} as well as the original \fuzzchick paper for more details
-about the implementation and semantics of this case study.
+  hrictcu2016testing} as well as the original CGPT paper
+\cite{lampropoulos2019coverage} for more details about the implementation and
+semantics of this case study.
 
@@
 in this work is called \emph{single-step noninterference}
-\cite{hritcu2013testing}.
-%
-In simple terms, this property asserts that, given two indistinguishable machine
-states, running a single instruction on both machines brings them to resulting
-states that are again indistinguishable.
-
-The tricky part about this property is, of course, satisfying its sparse
+\cite{hritcu2013testing} (SSNI).
+%
+SSNI asserts that, given two indistinguishable machine states, running a single
+instruction on both machines brings them to resulting states that are also
+indistinguishable.
+
+The tricky part about testing this property is to satisfy its sparse
 precondition: we need to generate two valid indistinguishable machine states in
@@
 %
-As demonstrated by \citeauthor{lampropoulos2019coverage}, generating two
+\citeauthor{lampropoulos2019coverage} demonstrated that generating two
 independent machine states using \quickcheck has virtually no chance of
@@
 %
-However, having the mutation mechanism available, we can use a clever trick: we
-can obtain a pair of valid indistinguishable machine states by generating a
-single valid machine state (something still hard but much easier than before),
-and then producing a similar mutated copy.
-%
-By doing this, we have a much higher chance of producing two almost identical
-states that pass the sparse precondition.
+However, using the mutation mechanism, we can obtain a pair of valid
+indistinguishable machine states by generating a single valid machine state
+(something still hard but much easier than before), and then producing a similar
+mutated copy.
+%
+This way, we have a higher chance of producing two almost identical states that
+pass the sparse precondition.
 
@@
 
-In this abstract machine, the enforced IFC rules are implemented using a rule
+The IFC rules enforced by this abstract machine are implemented using a rule
 table indexed by the instruction that the machine is about to perform.
@@
 For instance, to execute the \texttt{Store} instruction (which stores a value in
-a memory pointer) the machine needs to check that both the label of the program
-counter and the label of the pointer together can flow to the label of the
-destination memory cell.
+a memory pointer) the machine checks that both the labels of the program counter
+and the pointer together can flow to the label of the destination memory cell.
 %
@@
 %
-After this check, this instruction overwrites the value at the destination cell
+After this check, this machine overwrites the value at the destination cell
 and updates its label with the maximum sensibility of the involved labels.
@@
 
-Where $l_{pc}$, $l_{p}$, $l_{v}$, and $l_{v'}$ represent the labels of: the
-program counter, the memory pointer, and the old a new values stored in that
-memory cell.
+\noindent Where $l_{pc}$, $l_{p}$, $l_{v}$, and $l_{v'}$ represent the labels
+of: the program counter, the memory pointer, and the old a new values stored in
+that memory cell.
 %
@@
 
-In this case study, we systematically introduce several bugs on the IFC
-enforcing mechanism by removing or weakening the checks stored in this rule
-table.
+Bugs can be systematically introduced in the IFC enforcing mechanism by removing
+or weakening the checks stored in this rule table.
 
@@
 WebAssembly \cite{haas2017bringing} is a popular assembly-like language designed
-to be an open standard for executing low-level code in the web, although it has
-become increasingly popular in standalone, non-web contexts as well.
+to be an open standard for executing low-level code in the web.%
+%
+\footnote{Although it has become increasingly popular in standalone, non-web
+  contexts as well.}
 %
@@
 
-Unlike most other programming languages, its behavior is fully specified, and
-WebAssembly programs are expected to be consistently interpreted across engines
---- despite some subtle details that we will address soon.
-%
-For this purpose, the WebAssembly standard provides a reference implementation
-with all the basic functionality expected from a compliant WebAssembly engine.
-
+WebAssembly semantics are fully specified, and programs are expected to be
+consistently interpreted across engines --- despite some subtle details that we
+will address soon.
+%
+Moreover, the WebAssembly standard provides a reference implementation with all
+the functionality expected from a compliant engine.
+%
 %% \begin{wrapfigure}{r}{0.35\textwidth}
@@
 
+Our tool is an attractive match for testing WebAssembly engines, and
+\textit{haskell-wasm} in particular, as the space of WebAssembly programs that
+can be represented using its AST contains mostly invalid ones, and automatically
+derived random generators cannot satisfy all the invariants required to produce
+interesting test cases.
+
 In this work, we are interested in using \mutagen to test the two most complex
@@
 %
-Our tool is an attractive match for testing \textit{haskell-wasm}, as the space
-of WebAssembly programs that can be represented using its AST contains mostly
-invalid ones, and automatically derived random generators cannot satisfy all the
-invariants required to produce interesting test cases.
-%
-Here, we avoided spending countless hours writing an extensive property-based
+We avoided spending countless hours writing an extensive property-based
 specification to mimic the reference WebAssembly specification.
 %
-Instead, we take advantage of the readily available reference implementation via
-differential testing.
+Instead, we used the reference WebAssembly implementation to find discrepancies
+(that could potentially lead to bugs) via differential testing
+\cite{mckeeman1998differential}.
 %
@@
 \textit{haskell-wasm} matches that of the reference implementation.
-
-\begin{table}[t]
-\scriptsize
-\begin{tabular}{|c|c|c|l|}
-\hline
-\textbf{Id}
-& \textbf{Subsystem}
-& \textbf{Category}
-& \textbf{Description} \\
-\hline
-1
-& Validator
-& Bug
-& Invalid memory alignment validation \\
-\hline
-2
-& Validator
-& Discrepancy
-& Validator accepts blocks returning multiple values  \\
-\hline
-3
-& Interpreter
-& Bug
-& Instance function invoker silently ignores arity mismatch \\
-\hline
-4
-& Interpreter
-& Bug
-& Allowed out-of-bounds memory access \\
-\hline
-5
-& Interpreter
-& Discrepancy
-& NaN reinterpretation does not follow reference \\
-\hline
-\end{tabular}
-\caption{\label{table:wasm:bugs}Bugs and discrepancies found by \mutagen in
-  \textit{haskell-wasm}.}
-%% \vspace{-20pt}
-\end{table}
-
-Unsurprisingly, this engine had several subtle latent bugs that were not caught
-by the existing unit tests and that we discovered using \mutagen while
-developing the test suite used in this work.
-%
-Moreover, \mutagen exposed two discrepancies between \textit{haskell-wasm} and
-the reference implementation.
+%
+Unsurprisingly, this engine had several latent bugs that were not caught by the
+existing unit tests and that we discovered by \mutagen while developing our test
+suite.
+%
+Moreover, \mutagen exposed two other discrepancies between \textit{haskell-wasm}
+and the reference implementation.
 %
@@
 These findings are briefly outlined in Table \ref{table:wasm:bugs}.
+%
+All of them were reported and later confirmed by the authors of
+\textit{haskell-wasm}.
+%
+Having sorted these issues out, we injected 10 new bugs in the validator as well
+as 5 new bugs in the interpreter of this engine (see Table
+\ref{table:wasm:injectedbugs}).
+%
+These bugs were mechanically injected by either
+%
+\begin{inparaenum}
+\item removing an existing integrity check (to weaken the WebAssembly
+  type-system/validator); or by
+\item simulating a copy-and-paste induced bug.%
+  \footnote{This kind of bugs was inspired by the real bug \#1 we found prior
+    this step.}
+  %
+  For instance, by replacing the implementation of an operation with a similar
+  compatible one (e.g., \texttt{i32.add} by \texttt{i32.sub}).
+\end{inparaenum}
+
+%% During this process, we quickly discovered 3 bugs and 2 discrepancies on this
+%% engine with respect to the reference implementation.
+%% %
+
+\begin{table}[t]
+\scriptsize
+\begin{tabular}{|c|c|c|l|}
+\hline
+\textbf{Id}
+& \textbf{Subsystem}
+& \textbf{Category}
+& \textbf{Description} \\
+\hline
+1
+& Validator
+& Bug
+& Invalid memory alignment validation \\
+\hline
+2
+& Validator
+& Discrepancy
+& Validator accepts blocks returning multiple values  \\
+\hline
+3
+& Interpreter
+& Bug
+& Instance function invoker silently ignores arity mismatch \\
+\hline
+4
+& Interpreter
+& Bug
+& Allowed out-of-bounds memory access \\
+\hline
+5
+& Interpreter
+& Discrepancy
+& NaN reinterpretation does not follow reference \\
+\hline
+\end{tabular}
+\caption{\label{table:wasm:bugs}Issues found by \mutagen in
+  \textit{haskell-wasm}.}
+%% \vspace{-20pt}
+\end{table}
+
+\begin{table}[t]
+\footnotesize
+\begin{tabular}{|c|c|l|}
+\hline
+\textbf{Id}
+& \textbf{Subsystem}
+& \textbf{Description} \\
+\hline
+1
+& Validator
+& Wrong if-then-else type validation on else branch \\
+\hline
+2
+& Validator
+& Wrong stack type validation \\
+\hline
+3
+& Validator
+& Removed function type mismatch assertion \\
+\hline
+4
+& Validator
+& Removed max memory instances assertion \\
+\hline
+5
+& Validator
+& Removed function index out-of-range assertion \\
+\hline
+6
+& Validator
+& Wrong type validation on \texttt{i64.eqz} instruction \\
+\hline
+7
+& Validator
+& Wrong type validation on \texttt{i32} binary operations \\
+\hline
+8
+& Validator
+& Removed memory index out-of-range assertion \\
+\hline
+9
+& Validator
+& Wrong type validation on \texttt{i64} constants \\
+\hline
+10
+& Validator
+& Removed alignment validation on \texttt{i32.load} instruction \\
+\hline
+11
+& Interpreter
+& Wrong interpretation of \texttt{i32.sub} instruction \\
+\hline
+12
+& Interpreter
+& Wrong interpretation of \texttt{i32.lt\_u} instruction \\
+\hline
+13
+& Interpreter
+& Wrong interpretation of \texttt{i32.shr\_u} instruction \\
+\hline
+14
+& Interpreter
+& Wrong local variable initialization \\
+\hline
+15
+& Interpreter
+& Wrong memory address casting on \texttt{i32.load8\_s} instruction \\
+\hline
+\end{tabular}
+\caption{\label{table:wasm:injectedbugs}Bugs injected into
+  \textit{haskell-wasm}.}
+%% \vspace{-15pt}
+\end{table}
+
+%% %
+%% Here, we avoided spending countless hours writing an extensive property-based
+%% specification to mimic the reference WebAssembly specification.
+%% %
+%% Instead, we take advantage of the readily available reference implementation via
+%% differential testing.
+%% %
+%% In this light, our testing properties assert that any result produced by
+%% \textit{haskell-wasm} matches that of the reference implementation.
 
@@
 
-We begin by designing a property to test the WebAssembly validator implemented
-in \textit{haskell-wasm}.
-%
-To keep things simple, we can simply assert that, whenever a randomly generated
-(or mutated for that matter) WebAssembly module is valid according to
-\textit{haskell-wasm}, then the reference implementation agrees upon it.
+%% We begin by designing a property to test the WebAssembly validator implemented
+%% in \textit{haskell-wasm}.
+%% %
+To keep things simple, we assert that, whenever a randomly generated (or
+mutated) WebAssembly module is valid according to \textit{haskell-wasm}, then
+the reference implementation agrees upon it.
 %
@@
 %
-In Haskell, we write the following testing property:
+In Haskell, we write the following property:
 
@@
 
-We want to remark that, although here for simplicity we only focus on finding
-false negatives, in a realistic test suite, one would also want to test for
-false positives, i.e., when a module is valid and \textit{haskell-wasm} rejects
-it.
-%
-This can be easily done by inverting the direction of the implication
-(\texttt{<==}) in the property above.
-%
-However, the resulting property is much slower, as every tested module will
-always be serialized and run against the reference implementation.
+We note that, although we only focus on finding false negatives, in a complete
+test suite, one would also want to test for false positives, i.e., when a module
+is valid and \textit{haskell-wasm} rejects it.
+%
+%% This can be easily done by inverting the direction of the implication
+%% (\texttt{<==}) in the property above.
+%
+However, the resulting property is much slower, as every test case will always
+be serialized and run against the reference implementation.
 
@@
 Testing the WebAssembly interpreter is substantially more complicated than
-testing the validator, since it requires running actual programs.
+testing the validator, since it requires running and comparing the output of
+actual programs.
 %
@@
 
-To keep things simple here as well, we write a function \texttt{mk\_module} to
+To simplify things here as well, we write a function \texttt{mk\_module} to
 build a stub module which initializes one memory block and exports a single
@@
 %
-Having solved this issue in \textit{haskell-wasm}, we proceed to define a more
-useful specialized version of \texttt{prop\_interpreter} that fixes the type of
-the generated function to take two arguments (of type \texttt{I32} and
-\texttt{F32}) and return an \texttt{I32} as a result:
+Having fixed this issue, we define a more useful specialized version of
+\texttt{prop\_interpreter} that fixes the type of the generated function to take
+two arguments (of type \texttt{I32} and \texttt{F32}) and return an \texttt{I32}
+as a result:
 
@@
 prop_interpreter_i32 fun i f =
-  prop_interpreter
-    (FuncType { params = [I32, F32], result = [I32] })
+  prop_interpreter (FuncType { params = [I32, F32], result = [I32] })
                    fun
@@
 %
-Worth mentioning again, a realistic test suite should at least include different
-variants of this property testing functions of several different types, as well
-as properties testing multiple functions simultaneously.
+As before, we remark that a complete test suite should at least include
+different variants of this property testing functions of several different
+types, as well as properties testing multiple functions simultaneously.
 

### 2021-08-23 papers/icse22/main.tex
Pass over evaluation (Section 6). Needs some compression.

@@
 
-All the experiments were performed in a dedicated workstation with an Intel Core
+We performed all experiments in a dedicated workstation with an Intel Core
 i7-8700 CPU running at 3.20GHz, and equipped with 32GB of RAM.
 %
-We ran each experiment 30 times except for the ones involving the bugs on the
+Each experiment was repeated 30 times except for the ones testing the
 WebAssembly interpreter, which were run 10 times due to time constraints.
 %
-From there, we followed the same approach taken by
+From there, we followed the approach taken by
 \citeauthor{lampropoulos2019coverage} and collected the mean-time-to-failure
@@
 %
-In all cases, we used a one-hour timeout to stop the execution of both tools if
-they have not yet found a counterexample.
-
-Additionally, we collected the failure rate observed for each bug, i.e. the
+We used a one-hour timeout to stop the execution of both tools if they have not
+yet found a counterexample.
+%
+In addition, we collected the failure rate observed for each bug, i.e. the
 proportion of times each tool finds each bug within the one-hour testing budget.
 %
-We found this metric crucial to be analyzed when replicating \fuzzchick's
-results, as opposed to just paying attention at the MTTF.
-
-In both case studies, we additionally show how the FIFO scheduling and trace
-reset heuristics described in Section \ref{sec:heuristics} affect the testing
-performance by disabling them when using our tool.
+%% We found this metric crucial to be analyzed when replicating \fuzzchick's
+%% results, as opposed to just paying attention at the MTTF.
+
+In both case studies, we show how the FIFO scheduling and trace reset heuristics
+described in Section \ref{sec:heuristics} affect the testing performance of
+\mutagen by individually disabling them.
 %
@@
 As introduced earlier, \fuzzchick can only find 5 out of the 20 bugs injected
-with a \%100 success rate within the one hour testing budget.
+with a 100\% success rate within the one hour testing budget.
 %
@@
 %
-However, we want to remark that our experiments encompass 30 independent runs
-instead of the 5 originally used when presenting \fuzzchick
+However, we remark that our experiments encompass 30 independent runs instead of
+the 5 originally used when presenting \fuzzchick
 \citeyearpar{lampropoulos2019coverage}.
@@
 Moreover, since the MTTF simply aggregates all runs, regardless of if they found
-a bug or timed out, this metric is quite sensitive to the particular timeout
-used on each experiment, and will tend to inflate the results as soon as the
-failure rate goes below 1.
-%
-For this reason, additionally comparing the failure rate between tools give us a
-better estimation of the reliability of both tools, where \mutagen shows a clear
+a bug or timed out, this metric is quite sensitive to the timeout used on each
+experiment, and will inflate the results as soon as the failure rate goes below
+1.
+%
+For this reason, comparing the failure rate between tools give us a better
+estimation of the reliability of both tools, where \mutagen shows a clear
 improvement.
@@
 This tool either finds bugs relatively quickly (after a few thousands tests) or
-does not find them at all within the time budget, which suggests that its power
-scheduler assigns some mutation candidates too little energy before they become
+does not find them at all within the one-hour budget, suggesting that its power
+scheduler assigns some enqueued test cases too little energy before they become
 uninteresting and get ultimately thrown away.
@@
 %
-Additionally, as advised by \cite{arcuri2014hitchhiker} when comparing random
-testing tools, we computed the non-parametric Vargha-Delaney $A_{12}$ measure
-\cite{vargha2000critique}.
+Additionally, as advised by \citeauthor{arcuri2014hitchhiker}
+\cite{arcuri2014hitchhiker} when comparing random testing tools, we computed the
+non-parametric Vargha-Delaney $A_{12}$ measure \cite{vargha2000critique}.
 %
@@
 %
-As it can be observed, using an exhaustive mutation approach can be not only
-more reliable, but also faster.
+The \textit{Estimate} of this measure simply denotes the statistical difference
+between the compared distributions in a categorical manner.
+%
+Estimates in {\color{blue} blue} and {\color{red} red} indicate results
+favoring \fuzzchick and \mutagen, respectively.
+%
+As it can be observed,
+%
+%% using an exhaustive mutation approach can be not only
+%% more reliable, but also faster.
 %
@@
   comparison between \fuzzchick and \mutagen when considering only successful
-  runs.
-  %
-  Estimates in {\color{blue} blue} and {\color{red} red} indicate results
-  favoring \fuzzchick and \mutagen, respectively. }
+  runs.}
 %% \vspace{-25pt}
@@
 Firstly, this case study does not seem to be strongly affected by using a FIFO
-scheduling, where the same benchmark takes only roughly \%20 longer to complete
+scheduling, where the same benchmark takes only roughly 20\% longer to complete
 when disabling this heuristic.
 %
-Upon inspection, we found that the reason behind this is that the mutation
-candidate queues remain empty most of the time (generation mode), and when a new
-interesting candidate gets inserted, all its mutants (and their descendants) are
-processed before the next one is enqueued.
-%
-So, the proportion of time this heuristic is actually active tends to be minor
-under this case study.
+Upon inspection, the reason behind this is that the mutation queues remain empty
+most of the time (generation mode), and when a new interesting candidate gets
+inserted, all its mutants (and their descendants) are processed before the next
+one is enqueued.
+%
+So, the proportion of time this heuristic is active tends to be minor.
+%% under this case study.
 
@@
 %
-However, some of the hardest to find bugs cannot be reliably exposed using this
+However, some of the hardest-to-find bugs cannot be reliably exposed using this
 fixed amount of random mutations (\#4, \#9, \#12, \#18, and \#10), suggesting
@@
 %
-Moreover, although bugs \#2, \#9, \#8, and \#6 are found with \%100 success rate
+Moreover, although bugs \#2, \#9, \#8, and \#6 are found with 100\% success rate
 across runs, it takes several times longer for \mutagen to find them.

### 2021-08-23 papers/icse22/main.tex
Pass over last two sections and compression. One page to go.

@@
 
+\setlength{\textfloatsep}{5pt}
+\setlength{\floatsep}{5pt}
+
 % ----------------------------------------
@@
 aforementioned limitations by applying mutations \emph{exhaustively} (see
-Section \ref{sec:mutagen}).
+\S \ref{sec:mutagen}).
 %
@@
 %
-%% \mutagen is introduced in detail in Section \ref{sec:mutagen}.
+%% \mutagen is introduced in detail in \S \ref{sec:mutagen}.
 %
@@
 Additionally, the testing loop of our tool incorporates two heuristics that help
-finding bugs faster and more reliably (Section \ref{sec:heuristics}).
+finding bugs faster and more reliably (\S \ref{sec:heuristics}).
 %
@@
 To validate our ideas, we use two main case studies (described in detail in
-Section \ref{sec:casestudies}): the Information-Flow Control (IFC) stack machine
+\S \ref{sec:casestudies}): the Information-Flow Control (IFC) stack machine
 used by \citeauthor{lampropoulos2019coverage} in their original work, and an
@@
 %
-Our results (Section \ref{sec:evaluation}) indicate that \mutagen outperforms
+Our results (\S \ref{sec:evaluation}) indicate that \mutagen outperforms
 \fuzzchick both in terms of time to failure and failure rate.
@@
 %% Last two sections
-Finally, we discuss related work in Section \ref{sec:related} and conclude
-in Section \ref{sec:conclusions}.
+Finally, we discuss related work in \S \ref{sec:related} and conclude
+in \S \ref{sec:conclusions}.
 
@@
 
-The next subsection show how GCPT helps overcoming this limitation without
-having to write specialized random generators by hand.
+The next subsection shows how GCPT helps overcoming this limitation without
+having to write specialized generators by hand.
 
@@
 %% %
-%% We will provide more details about structure-preserving mutations in Section
+%% We will provide more details about structure-preserving mutations in \S
 %% \ref{sec:mutagen}.
@@
 %% As mentioned earlier, when we recreated the evaluation of the IFC stack machine
-%% case study (described in detail in Section \ref{sec:casestudies}), we found that
+%% case study (described in detail in \S \ref{sec:casestudies}), we found that
 %% after 30 runs (as opposed to the 5 runs used original by
@@
 %% %
-%% These results are presented in detail in Section \ref{sec:evaluation}.
+%% These results are presented in detail in \S \ref{sec:evaluation}.
 %
@@
 %
-%% We show in Section \ref{sec:heuristics} how this can be achieved by analyzing
+%% We show in \S \ref{sec:heuristics} how this can be achieved by analyzing
 %% the execution information in order to prioritize test cases with novel execution
@@
 %
-%% Later, Section \ref{sec:heuristics} introduces heuristics that help finding bugs
+%% Later, \S \ref{sec:heuristics} introduces heuristics that help finding bugs
 %% faster in certain testing scenarios.
@@
 \cite{runciman2008smallcheck} or \emph{Korat} \cite{boyapati2002korat}, that
-produce test cases exhaustively --- please refer to Section \ref{sec:related}
+produce test cases exhaustively --- please refer to \S \ref{sec:related}
 for a detailed discussion.
@@
 %% search space based on detecting unused subexpressions --- this is discussed in
-%% Section \ref{sec:related}.
+%% \S \ref{sec:related}.
 
@@
 \end{verbatim}
-%% \vspace{-5pt}
+\vspace{-5pt}
 \caption{\label{fig:mutagen:mutator}\mutagen mutator for the \texttt{Tree} data
   type.}
-%% \vspace{-10pt}
+%% \vspace{-15pt}
 \end{figure}
@@
   \nonumber
-  {\small \texttt{positions}}\quad
+  \footnotesize
+  \texttt{positions}\quad
   \left(
   \tikz[
-    baseline=-7mm,
-    level distance=0.6cm,
-    level 1/.style={sibling distance=0.75cm},
-    level 2/.style={sibling distance=0.75cm}
+    baseline=-6mm,
+    level distance=0.5cm,
+    level 1/.style={sibling distance=0.5cm},
+    level 2/.style={sibling distance=0.5cm}
   ]{
-    \node {\small \texttt{Branch}}
-      child {node {\small \texttt{Leaf}}
-        child {node {\small \texttt{1}}}
-      }
-      child {node {\small \texttt{2}}}
-      child {node {\small \texttt{Leaf}}
-        child {node {\small \texttt{3}}}
+    \node {\texttt{Branch}}
+      child {node {\texttt{Leaf}}
+        child {node {\texttt{1}}}
+      }
+      child {node {\texttt{2}}}
+      child {node {\texttt{Leaf}}
+        child {node {\texttt{3}}}
       };
@@
   \tikz[
-    baseline=-7mm,
-    level distance=0.6cm,
-    level 1/.style={sibling distance=0.75cm},
-    level 2/.style={sibling distance=0.75cm}
+    baseline=-6mm,
+    level distance=0.5cm,
+    level 1/.style={sibling distance=0.5cm},
+    level 2/.style={sibling distance=0.5cm}
   ]{
-    \node {\texttt{\small []}}
-      child {node {\small \texttt{[0]}}
-        child {node {\small \texttt{[0,0]}}}
-      }
-      child {node {\small \texttt{[1]}}}
-      child {node {\small \texttt{[2]}}
-        child {node {\small \texttt{[2,0]}}}
+    \node {\texttt{[]}}
+      child {node {\texttt{[0]}}
+        child {node {\texttt{[0,0]}}}
+      }
+      child {node {\texttt{[1]}}}
+      child {node {\texttt{[2]}}
+        child {node {\texttt{[2,0]}}}
       };
@@
 
-\begin{figure*}[t]
-%% \vspace{-15pt}
-\newcommand\stacked[2]{\underset{\textstyle #2\mathstrut}{#1}}
-\begin{align*}
-  \nonumber
-  \qquad\qquad\qquad\qquad
-  \left[
-  \tikz[
-    baseline=-10mm,
-    level distance=0.6cm,
-    level 1/.style={sibling distance=0.75cm},
-    level 2/.style={sibling distance=0.75cm}
-  ]{
-    \node {\small \texttt{1}}
-      child {node {\small \texttt{2}}
-        child {node {\small \texttt{3}}
-          child {node {\small \texttt{4}}}
-        }
-      }
-  }
-  \right]
-  \;
-  +
-  \;
-  \tikz[
-    baseline=-10mm,
-    level distance=0.6cm,
-    level 1/.style={sibling distance=0.75cm},
-    level 2/.style={sibling distance=0.75cm}
-  ]{
-    \node {\small \texttt{1}}
-      child {node {\small \texttt{2}}
-        child {node {\small \texttt{3}}
-          child {node {\small \texttt{5}}}
-        }
-      }
-  }
-  \;
-  =
-  \;
-  \left[
-  \tikz[
-    baseline=-10mm,
-    level distance=0.6cm,
-    level 1/.style={sibling distance=0.75cm},
-    level 2/.style={sibling distance=0.75cm}
-  ]{
-    \node {\color{red} \small \texttt{1}}
-      child {node {\color{red} \small \texttt{2}}
-        child {node {\color{red} \small \texttt{3}}
-          child {node {\small \texttt{4}}}
-          child {node {\color{blue} \small \texttt{5}}}
-        }
-      }
-  }
-  \right]
-  ,
-  \stacked{{\color{blue} new=1\phantom{x}}}{{\color{red} depth=3}}
-  %% \end{equation}
-  %% \begin{equation}
-  %% \nonumber
-  %% \\
-  \qquad
-  \qquad
-  \left[
-  \tikz[
-    baseline=-10mm,
-    level distance=0.6cm,
-    level 1/.style={sibling distance=0.75cm},
-    level 2/.style={sibling distance=0.75cm}
-  ]{
-    \node {\small \texttt{1}}
-      child {node {\small \texttt{2}}
-        child {node {\small \texttt{3}}
-          child {node {\small \texttt{4}}}
-          child {node {\small \texttt{5}}}
-        }
-      }
-  }
-  \right]
-  \;
-  +
-  \;
-  \tikz[
-    baseline=-10mm,
-    level distance=0.6cm,
-    level 1/.style={sibling distance=0.75cm},
-    level 2/.style={sibling distance=0.75cm}
-  ]{
-    \node {\small \texttt{1}}
-      child {node {\small \texttt{2}}
-        child {node {\small \texttt{6}}
-          child {node {\small \texttt{7}}
-          }
-        }
-      }
-  }
-  \;
-  =
-  \;
-  \left[
-  \tikz[
-    baseline=-10mm,
-    level distance=0.6cm,
-    level 1/.style={sibling distance=0.75cm},
-    level 2/.style={sibling distance=1cm}
-  ]{
-    \node {\color{red} \small \texttt{1}}
-      child {node {\color{red} \small \texttt{2}}
-        child {node {\small \texttt{3}}
-          child {node {\small \texttt{4}}}
-          child {node {\small \texttt{5}}}
-        }
-        child {node {\color{blue}\small \texttt{6}}
-          child {node {\color{blue}\small \texttt{7}}}
-        }
-      }
-  }
-  \right]
-  ,
-  \stacked{{\color{blue} new=2\phantom{x}}}{{\color{red} depth=2}}
-\end{align*}
-%% \vspace{-10pt}
-\caption{\label{fig:mutagen:tracelog}Inserting two new execution traces into the
-  internal trace log (represented using brackets).}
-\end{figure*}
-
 \section{\mutagen Heuristics}
@@
 
-\begin{algorithm}[b]
+\begin{algorithm}[t]
   \SetArgSty{textnormal}
@@
 
+\begin{figure}[t]
+%% \vspace{-15pt}
+\scriptsize
+\newcommand\stacked[2]{\underset{\textstyle #2\mathstrut}{#1}}
+\begin{align*}
+  \nonumber
+  %% \qquad\qquad\qquad\qquad
+  \left[
+  \tikz[
+    baseline=-8mm,
+    level distance=0.5cm,
+    level 1/.style={sibling distance=0.5cm},
+    level 2/.style={sibling distance=0.5cm}
+  ]{
+    \node {\small \texttt{1}}
+      child {node {\small \texttt{2}}
+        child {node {\small \texttt{3}}
+          child {node {\small \texttt{4}}}
+        }
+      }
+  }
+  \right]
+  \;
+  +
+  \;
+  \tikz[
+    baseline=-8mm,
+    level distance=0.5cm,
+    level 1/.style={sibling distance=0.5cm},
+    level 2/.style={sibling distance=0.5cm}
+  ]{
+    \node {\small \texttt{1}}
+      child {node {\small \texttt{2}}
+        child {node {\small \texttt{3}}
+          child {node {\small \texttt{5}}}
+        }
+      }
+  }
+  \;
+  =
+  \;
+  \left[
+  \tikz[
+    baseline=-8mm,
+    level distance=0.5cm,
+    level 1/.style={sibling distance=0.5cm},
+    level 2/.style={sibling distance=0.5cm}
+  ]{
+    \node {\color{red} \small \texttt{1}}
+      child {node {\color{red} \small \texttt{2}}
+        child {node {\color{red} \small \texttt{3}}
+          child {node {\small \texttt{4}}}
+          child {node {\color{blue} \small \texttt{5}}}
+        }
+      }
+  }
+  \right]
+  ,
+  \stacked{{\color{blue} new=1\phantom{x}}}{{\color{red} depth=3}}
+  %% \end{equation}
+  %% \begin{equation}
+  %% \nonumber
+  %% \\
+  %% \qquad
+  \quad
+  \left[
+  \tikz[
+    baseline=-8mm,
+    level distance=0.5cm,
+    level 1/.style={sibling distance=0.5cm},
+    level 2/.style={sibling distance=0.5cm}
+  ]{
+    \node {\small \texttt{1}}
+      child {node {\small \texttt{2}}
+        child {node {\small \texttt{3}}
+          child {node {\small \texttt{4}}}
+          child {node {\small \texttt{5}}}
+        }
+      }
+  }
+  \right]
+  \;
+  +
+  \;
+  \tikz[
+    baseline=-8mm,
+    level distance=0.5cm,
+    level 1/.style={sibling distance=0.5cm},
+    level 2/.style={sibling distance=0.5cm}
+  ]{
+    \node {\small \texttt{1}}
+      child {node {\small \texttt{2}}
+        child {node {\small \texttt{6}}
+          child {node {\small \texttt{7}}
+          }
+        }
+      }
+  }
+  \;
+  =
+  \;
+  \left[
+  \tikz[
+    baseline=-8mm,
+    level distance=0.5cm,
+    level 1/.style={sibling distance=0.5cm},
+    level 2/.style={sibling distance=0.75cm}
+  ]{
+    \node {\color{red} \small \texttt{1}}
+      child {node {\color{red} \small \texttt{2}}
+        child {node {\small \texttt{3}}
+          child {node {\small \texttt{4}}}
+          child {node {\small \texttt{5}}}
+        }
+        child {node {\color{blue}\small \texttt{6}}
+          child {node {\color{blue}\small \texttt{7}}}
+        }
+      }
+  }
+  \right]
+  ,
+  \stacked{{\color{blue} new=2\phantom{x}}}{{\color{red} depth=2}}
+\end{align*}
+\vspace{-10pt}
+\caption{\label{fig:mutagen:tracelog}Inserting two new execution traces into the
+  internal trace log (represented using brackets).}
+\end{figure}
+
 \subsection{Detecting Trace Space Saturation And Tuning Random Mutations Parameter}
 
-As introduced in Section \ref{sec:mutagen}, our tool is parameterized by the
+As introduced in \S \ref{sec:mutagen}, our tool is parameterized by the
 number of random mutations to be generated over each mutable subexpression
@@
 \begin{center}
-\footnotesize
+\scriptsize
 \begin{tabular}{|c|c|c|c|}
@@
 \begin{table}[t]
-\footnotesize
+\scriptsize
 \begin{tabular}{|c|c|l|}
@@
 In both case studies, we show how the FIFO scheduling and trace reset heuristics
-described in Section \ref{sec:heuristics} affect the testing performance of
+described in \S \ref{sec:heuristics} affect the testing performance of
 \mutagen by individually disabling them.
@@
 
-\begin{figure}
+\setlength{\intextsep}{10pt}%
+\setlength{\columnsep}{10pt}%
+\begin{wrapfigure}{r}{0.5\linewidth}
 %% \begin{wrapfigure}{r}{0.5\textwidth}
 %% \centering
-%% \vspace{-20pt}
-\small
+\vspace{-15pt}
+\scriptsize
+
+\setlength{\tabcolsep}{2pt}
+\renewcommand{\arraystretch}{1}
+
+%% \begin{minipage}{0.5\linewidth}
 \begin{tabular}{ccccc}
@@
 \end{tabular}
+%% \end{minipage}
+
 \captionof{table}{\label{table:ifc:tests} Mean tests until first failure and effect size
@@
   runs.}
-%% \vspace{-25pt}
+\vspace{-20pt}
 %% \end{wrapfigure}
-\end{figure}
+\end{wrapfigure}
 
@@
 %
-So, the proportion of time this heuristic is active tends to be minor.
+Hence, the small proportion of time this heuristic is active tends to be
+inconsequential.
 %% under this case study.
@@
 Despite this, our tool is still capable of running substantially more tests that
-pass the sparse preconditions than \quickcheck in the same amount of time, and
-which ultimately leads us to find bugs.
+pass the sparse preconditions than \quickcheck in the same amount of time,
+\emph{and which ultimately leads us to find bugs}.
 
@@
 %
-In this section we only focus on three main topics:
+In this section we will focus on three main topics:
 %
@@
 %
-fuzzing using coverage information, and
+coverage-guided fuzzing, and
 %
@@
 grammars or data type definitions) has been tackled from different angles.
-
+%
 % DRAGEN and DRAGEN2
@@
 %
-DRAGEN2 \cite{Mista2019GeneratingRS} extends this idea adding support for
-generating richer random data by extracting static information like library APIs
-and function input patterns from the codebase.
+DRAGEN2 \cite{Mista2019GeneratingRS} extends this idea with support for
+generating richer random data by extracting library APIs and function input
+patterns from the codebase.
 %
@@
 %
-Exploiting some of the ideas described above, \emph{QuickFuzz} \cite{GriecoCB16,
-  grieco2017} is a generational fuzzer that leverages existing Haskell libraries
-describing common file formats to synthesize random data generators that are
-used in tandem with existing off-the-shelf low-level fuzzers to find bugs in
-heavily used programs.
+\emph{QuickFuzz} \cite{GriecoCB16, grieco2017} is a fuzzer that exploits these
+ideas and uses existing Haskell libraries describing common file formats to
+synthesize random data generators that are used in tandem with existing
+off-the-shelf low-level fuzzers to find bugs in heavily used programs.
 
@@
 %
-Later, \citeauthor{LampropoulosGHH17} \citeyearpar{LampropoulosGHH17} extends
+Later, \citeauthor{LampropoulosGHH17} \citeyearpar{LampropoulosGHH17} extended
 this approach by adding a limited form of constraint solving controllable by the
@@
 a derivation mechanism to obtain constrained random generators directly by
-defining their structure using inductively defined relations in Coq.
+defining the target structure using inductively defined relations in Coq.
 
@@
 However, more specialized generators can be used directly by \mutagen, and
-combining them with type-preserving mutations using a hybrid technique is an
-idea that we aim to address in the future.
+combining them with type-preserving mutations is an idea we aim to address in
+the future.
 
@@
 % AFLfast
-\emph{AFLFast} \cite{bohme2017coverage} is an extension to AFL that uses Markov
-chain models to tune the power scheduler towards testing low-frequency paths.
+\emph{AFLFast} \cite{bohme2017coverage} extends AFL with Markov chain models to
+tune the power scheduler towards testing low-frequency paths.
 %
@@
 Similar to our tool, \emph{CollAFL} \cite{gan2018collafl} is a variant of AFL
-that uses path- instead of edge-based coverage, which helps distinguishing
-executions more precisely by reducing path collisions.
+that uses path- instead of edge-based coverage to distinguish executions more
+precisely by reducing path collisions.
 
 %AFLGo
-When the source code history is available, \emph{AFLGo} \cite{bohme2017directed}
+If the source code history is available, \emph{AFLGo} \cite{bohme2017directed}
 is a fuzzer that can be targeted to exercise the specific parts of the code
@@
 %
-Back to PBT, a related idea called \emph{targeted property-based testing} (TPBT)
-is to use fitness functions to guide the testing efforts towards user defined
-goals \cite{loscher2017targeted, loscher2018automating}.
-
-% Coverage-guided fuzzing + symbolic execution
-%
-Moreover, a popular technique is to combine coverage-guided fuzzing with ideas
-borrowed from symbolic execution tools like \emph{KLEE} \cite{cadar2008klee} to
-avoid getting stuck in superficial paths \cite{stephens2016driller}.
-%
-When using off-the-shelf symbolic executors, this technique is often limited by
-the path explosion problem, although recent tools like \emph{QSYM}
-\cite{yun2018qsym} demonstrate that using tailored concolic executors can help
-to overcome this limitation.
-
-Most of these ideas can potentially be incorporated in our tool, and we keep
-them as a challenge for future work.
+In PBT, a related idea called \emph{targeted property-based testing} (TPBT) is
+to use fitness functions to guide testing efforts towards user defined goals
+\cite{loscher2017targeted, loscher2018automating}.
+
+%% % Coverage-guided fuzzing + symbolic execution
+%% %
+%% Moreover, a popular technique is to combine coverage-guided fuzzing with ideas
+%% borrowed from symbolic execution tools like \emph{KLEE} \cite{cadar2008klee} to
+%% avoid getting stuck in superficial paths \cite{stephens2016driller}.
+%% %
+%% When using off-the-shelf symbolic executors, this technique is often limited by
+%% the path explosion problem, although recent tools like \emph{QSYM}
+%% \cite{yun2018qsym} demonstrate that using tailored concolic executors can help
+%% to overcome this limitation.
+
+All these ideas can potentially be incorporated in our tool, and we keep them as
+a challenge for future work.
 
@@
 % SmallCheck and LazySmallCheck
-\emph{SmallCheck} \cite{runciman2008smallcheck} is a Haskell tool that also
-follows this idea.
+\emph{SmallCheck} \cite{runciman2008smallcheck} is a Haskell tool that follows
+this idea.
 %
@@
 %
-In contrast to fully exhaustive testing tools, \mutagen initially relies on
-randomness to find initial interesting mutable candidates.
+In contrast to fully exhaustive testing tools, \mutagen relies on randomness to
+find initial interesting mutable test cases.
 %
@@
 In our case studies, however, we observed that their preconditions tend to be
-quite strict when executing test cases obtained by mutating passed existing
-ones, fully evaluating their input before executing the postcondition.
-%
-Thus, we do not expect to see considerable improvements by applying this idea in
-this particular scenario.
+quite strict when executing test cases obtained by mutating valid existing ones,
+fully evaluating their input before executing the postcondition.
+%
+Thus, we do not forsee considerable improvements by applying this idea in this
+particular scenario.
 %
@@
 %
-Inspired by \fuzzchick, our tool uses coverage information to guide
-automatically derived mutators producing high-level, type-preserving mutations.
-%
-However, instead of relying heavily on randomness and power schedules to find
-bugs, our tool uses an exhaustive mutation approach that generates every
-possible mutant for each interesting input candidate, and schedules it to be
-tested exactly once.
-%
-This is in turn inspired by exhaustive bounded testing tools that focus on
-testing every possible input value of the system under test up to a certain size
-bound.
-%
-Moreover, we foresee no problems into translating our tool in other programming
-languages beyond Haskell.
+Inspired by exhaustive bounded testing tools, our tool extends the CGPT approach
+introduced by \citeauthor{lampropoulos2019coverage} with an exhaustive mutation
+mechanism that generates every possible mutant for each interesting test case,
+and schedules it to be tested exactly once.
+
+%% , our tool uses coverage information
+%% to guide automatically derived mutators producing high-level, type-preserving
+%% mutations.
+%% %
+%% However, instead of relying heavily on randomness and power schedules to find
+%% bugs, our tool uses an exhaustive mutation approach that generates every
+%% possible mutant for each interesting input candidate, and schedules it to be
+%% tested exactly once.
+%% %
+%% This is in turn inspired by exhaustive bounded testing tools that focus on
+%% testing every possible input value of the system under test up to a certain size
+%% bound.
+%% %
+%% Moreover, we foresee no problems into translating our tool in other programming
+%% languages beyond Haskell.
 
@@
 
-As already indicated throughout this work, there are several directions for
-future work we aim to investigate:
+As indicated throughout this work, there are several directions for future work:
 %
 combining \mutagen with specialized generators; improving the code
-instrumentation in order to effectively target mutations towards specific parts
-of the code; and evaluating the effect of lazy pruning; among others.
+instrumentation in order to target mutations towards specific parts of the code;
+and evaluating the effect of lazy pruning; among others.
 %

### 2021-08-24 papers/icse22/main.tex
Compressed sections 1 and 2

@@
 alleviates this limitation by enhancing automatically synthesized generators
-with type-preserving mutations guided by execution traces.
+with structure-preserving mutations guided by execution traces.
 %
@@
 However, users of \quickcheck (and RPBT in general) are well aware that one of
-the biggest challenges while using this technique is to provide suitable random
-data generators needed to instantiate the testing properties.
-%
-In some extreme cases, writing highly-tuned random generators capable of
-exercising every part of a complex system on a reasonable basis can take several
-thousand person-hours of trial and error \cite{lampropoulos2019coverage}.
+the biggest challenges while using RPBT is to provide suitable random data
+generators needed to instantiate the testing properties.
+%
+In extreme cases, writing highly-tuned random generators capable of exercising
+every part of a complex system on a reasonable basis can take several thousand
+person-hours of trial and error \cite{lampropoulos2019coverage}.
 %
@@
 Instead of continuously generating random invalid test cases from scratch, CGPT
-works by keeping a queue of \emph{interesting} previously executed test cases
+works by keeping queues of \emph{interesting} previously executed test cases
 that can be mutated using high-level, structure-preserving transformations to
@@
 structure of the random data we want to mutate in a concrete manner, and
-specialized mutators acting at the data constructor level can be automatically
-derived directly from their definition --- making strongly-typed programming
-languages with expressive type systems an ideal match for CGPT.
+specialized mutators can be automatically derived directly from their definition
+--- making strongly-typed programming languages an ideal match for CGPT.
 
@@
 produce new execution traces --- any other test case is simply thrown away.
-%
+
 Worth mentioning, this technique is not meant to replace sophisticated, manually
@@
 
-In this work, we establish several aspects of the original CGPT approach by
-\citeauthor{lampropoulos2019coverage} that leave considerable room for
-improvement.%
-%
+In this work, we establish several aspects%
 \footnote{Some of these were already identified by
   \citeauthor{lampropoulos2019coverage} in their original work.}
+%
+of the original CGPT approach by \citeauthor{lampropoulos2019coverage} that
+leave considerable room for improvement (see \S \ref{sec:background}).
 %
@@
   superficial and rely heavily on randomness to produce diverse mutants,
-\item the queuing mechanism can cause large delays if interesting values are
-  enqueued frequently and there is no way to prioritize them, and
+\item the queuing mechanism can cause delays if interesting values are enqueued
+  frequently and there is no way to prioritize them, and
 \item using an heuristic to assign a certain ``mutation budget'' to each
   interesting test case (often referred to as a \emph{power schedule}) requires
-  fine tuning and can be hard to generalize to work well under different testing
-  scenarios.
+  fine tuning and can be hard to generalize to different kinds of testing
+  properties.
 \end{inparaenum}
@@
 %% Heuristics to improve testing loop
-Additionally, the testing loop of our tool incorporates two heuristics that help
+Additionally, the testing loop of \mutagen incorporates two heuristics that help
 finding bugs faster and more reliably (\S \ref{sec:heuristics}).
 %
-In the first place, \mutagen uses first-in first-out (FIFO) scheduling with
+In the first place, our tool uses first-in first-out (FIFO) scheduling with
 priority to enqueue interesting test cases for mutation.
@@
 %% Validation
-To validate our ideas, we use two main case studies (described in detail in
-\S \ref{sec:casestudies}): the Information-Flow Control (IFC) stack machine
-used by \citeauthor{lampropoulos2019coverage} in their original work, and an
-existing WebAssembly engine of industrial strenght written in Haskell.
+To validate our ideas, we use two case studies (described in detail in \S
+\ref{sec:casestudies}): the Information-Flow Control (IFC) stack machine used by
+\citeauthor{lampropoulos2019coverage} in their original work, and an existing
+{Web-Assembly} engine of industrial strenght written in Haskell.
 %
@@
 gen_tree :: Size -> Gen a -> Gen (Tree a)
-gen_tree gen_a 0 = Leaf <$> gen_a
-gen_tree gen_a n = oneof
+gen_tree 0 gen_a = Leaf <$> gen_a
+gen_tree n gen_a = oneof
   [ Leaf <$> gen_a
@@
 %
-In practice, \quickcheck will give up quite early if most of the generated test
-cases get discarded because they do not satisfy the balanced precondition.
-%
-Sadly, most of the inputs generated by our na\"ive generator suffer from this
-problem, and the interesting part of the property (\texttt{balanced (insert x t)})
-is tested very sporadically as a result.
+In practice, \quickcheck will give up quite if most of the generated test cases
+get discarded because they do not satisfy the balanced precondition.
+%
+Sadly, most of the inputs produced by our na\"ive generator suffer from this
+problem, and the interesting part of the property (\texttt{balanced (insert x
+  t)}) is tested very sporadically as a result.
 %
@@
 Moreover, we replicated the IFC stack machine case study bundled with \fuzzchick
-(their reference implementation) and observed a surprising limitation:
+and observed a surprising limitation:
 %
@@
 All these observations led us to consider three main aspects of CGPT that can be
-improved upon --- and that constitute the main goal of this work.
+improved upon:
+%
+%% --- and that constitute the main goal of this work.
 %
@@
 %
-In no particular order:
+%% In no particular order:
 
@@
 for the sake of simplicity, the mutators proposed by
-\citeauthor{lampropoulos2019coverage} are derived to follow a top-down approach,
-similar to that used by the automatically derived random generators.
+\citeauthor{lampropoulos2019coverage} are derived to follow a top-down approach.
+%% similar to that used by the automatically derived random generators.
 %
@@
 data structures, potentially excluding interesting applications that might
-require producing deeper valid values, e.g., programs, network protocol
-interactions, etc.
+require producing deeper valid values.
+%
+%e.g., programs, network protocol interactions, etc.
 %
@@
 
-On the other hand, using random generators to produce needed subexpressions
-when growing data constructors can be dangerous, as we are introducing the very
-same ``uncontrolled'' randomness that we wanted to mitigate in the first place!
-%
-If the random generator produces an invalid subexpression (something quite
-likely), this might just invalidate the whole mutated test case.
-%
-We believe that growing data constructors needs to be done carefully.
+%% On the other hand, using random generators inside mutators to produce new
+%% subexpressions when ``growing'' a data constructor can be dangerous, as this
+%% introduces the same ``uncontrolled'' randomness we want to mitigate in the first
+%% place!
+%% %
+%% If such generator produces an invalid subexpression, it might invalidate the
+%% whole mutated test case.
+%
+%% Hence, growing data constructors must be done carefully.
+
 %% %
@@
 %
-the CGPT testing loop uses two single queues for keeping valid and discarded
-mutation candidates.
+the CGPT testing loop uses two single-ended queues for keeping valid and
+discarded interesting test cases.
 %
@@
 %
-This can limit the effectiveness of the testing loop if the queues tend to grow
-more often than they tend to shrink, as interesting mutation candidates can get
-``buried'' at the end of a long queue that only exercises the same portion of
-the target code.
-%
-In an extreme case, they might not be processed at all within the testing
-budget.
-%
-Ideally, \emph{one would like a mechanism that prioritizes mutating test cases
-  that discover new portions of the code right away}, and that is capable of
-jumping back and forth from mutation candidates whenever this happens.
+This can limit the effectiveness of the testing loop if the queues grow more
+often than they shrink, as interesting test cases can get ``buried'' at the end
+of a long queue and,
+%
+%% that only exercises the same portion of the target code.
+%
+in an extreme case, not get processed at all within the testing budget.
+%
+Ideally, \emph{we need a mechanism that prioritizes mutating test cases that
+  discover new portions of the code right away}.
+%
+%% , and that is capable of jumping
+%% back and forth from  whenever this happens.
 %
@@
 %
-it is not clear how the power schedule used to assign a given budget to each
-mutable test case in CGPT works in the context of high-level
-structure-preserving mutations.
+it is not clear how the heuristic used to assign a given budget to each mutable
+test case in CGPT works in the context of high-level structure-preserving
+mutations.
 %
@@
 %
-However, AFL uses a different mutation approach that works at the bit level.
-%
-This raises the question about what is the best power schedule configuration
-when using a high-level mutation approach --- something quite challenging to
-characterize in general given the expressivity of the data types used to drive
-the mutators.
+However, AFL uses a bit-level mutation approach.
+%
+It is unclear what the best heuristic configuration is when using a high-level
+mutation approach --- something challenging to characterize in general given the
+expressivity of the data types used to drive the mutators.
 \end{itemize}
@@
 
-In contrast to CGPT, and many other popular fuzzing tools for that matter,
-\mutagen works by to mutating test cases on an exhaustive and precise manner,
-where
+In contrast to CGPT and many other popular fuzzing tools, \mutagen works by to
+mutating test cases on an exhaustive and precise manner, where
 %
@@
 %
-Our realization is that, by following an exhaustive mutation approach, we can
-avoid the need for a power schedule, i.e., a heuristic to assign a certain
-mutation budget to each mutated test case.
+Our realization is that, by using an exhaustive mutation approach, we avoid
+needing a heuristic to assign a mutation budget to each mutated test case.
 %
@@
 %
-Using a small default value is again inspired by exhaustive bounded testing
-tools, and avoids introducing unnecessary randomness when growing mutated
-expressions.
+Using a small default value, as oppossed to a randomly generated one (as done by
+the original GCPT), is again inspired by exhaustive bounded testing tools, and
+avoids introducing unnecessary randomness when growing mutated expressions.
 
@@
 
-\begin{figure}[t]
+\begin{figure}[b]
 \begin{verbatim}

### 2021-08-24 papers/icse22/main.tex
More general compression
Compressing. Compressing.

@@
 \usepackage{enumitem}
-%% \usepackage[square,numbers]{natbib}
+\usepackage{titlesec}
+
+\titlespacing*{\section}{0pt}{5pt}{5pt}
+\titlespacing*{\subsection}{0pt}{5pt}{5pt}
+\titlespacing*{\subsubsection}{0pt}{5pt}{5pt}
+\titlespacing*{\paragraph}{10pt}{0pt}{10pt}
 
@@
 \setlength{\floatsep}{5pt}
+
+\SetAlFnt{\small}
+\SetAlCapFnt{\small}
+\SetAlCapNameFnt{\small}
 
@@
 prop_insert_inv :: a -> Tree a -> Property
-prop_insert_inv x t =
-  balanced t ==> balanced (insert x t)
+prop_insert_inv x t = balanced t ==> balanced (insert x t)
 \end{verbatim}
@@
 
+% gen_tree :: Size -> Gen a -> Gen (Tree a)
+
 \begin{verbatim}
-gen_tree :: Size -> Gen a -> Gen (Tree a)
-gen_tree 0 gen_a = Leaf <$> gen_a
-gen_tree n gen_a = oneof
-  [ Leaf <$> gen_a
-  , Branch <$> gen_tree (n-1) <*> gen_a <*> gen_tree (n-1) ]
+genTree 0 = Leaf <$> genA
+genTree n = oneof
+  [Leaf <$> genA, Branch <$> genTree(n-1) <*> genA <*> genTree(n-1)]
 \end{verbatim}
@@
 Payloads are generated simply by delegating the task to an external random
-generator (\texttt{gen\_a}) defined elsewhere.
+generator (\texttt{genA}) defined elsewhere.
 %
@@
 
-Readers familiar with RBPT will recognize that attempting to test
-\texttt{prop\_insert\_inv} using this generator in \quickcheck directly does not
-work well.
-%
-In practice, \quickcheck will give up quite if most of the generated test cases
-get discarded because they do not satisfy the balanced precondition.
+Readers familiar with RBPT will recognize that using \texttt{genTree} to test
+\texttt{prop\_insert\_inv} with \quickcheck directly does not work well.
+%
+In practice, \quickcheck gives up if most of the generated test cases get
+discarded due to failed preconditions.
+%
+%because they do not satisfy the balanced precondition.
 %
@@
   t)}) is tested very sporadically as a result.
-%
-In other words, \emph{the precondition of our testing property becomes very
-  sparse when we use simple, automatically derived generators.}
+%% %
+%% In other words, \emph{the precondition of our testing property becomes very
+%%   sparse when we use simple, automatically derived generators.}
 
@@
 %% %
-%% If \textit{QValid} is empty, then the same is attempted using \textit{QDisc}.
+%% If \textit{QValid} is empty, then the same is attempted using \textit{QDiscarded}.
 %% %
@@
 %% Outlined in Algorithm \ref{algo:fuzzchick:loop}, CGPT starts by creating two
-%% queues, \textit{QValid} and \textit{QDisc} for valid and discarded previously
+%% queues, \textit{QValid} and \textit{QDiscarded} for valid and discarded previously
 %% executed test cases, respectively.
@@
 %%     i $\gets$ 0\;
-%%     TLog, QValid, QDisc $\gets$ $\varnothing$\;
+%%     TLog, QValid, QDiscarded $\gets$ $\varnothing$\;
 %%     \While{i $<$ N}{
-%%       x $\gets$ Pick(QValid, QDisc, gen, mut)\;
+%%       x $\gets$ Pick(QValid, QDiscarded, gen, mut)\;
 %%       (result, trace) $\gets$ WithTrace(P(x))\;
@@
 %%         }{
-%%           Enqueue(QDisc, (x, e))\;
+%%           Enqueue(QDiscarded, (x, e))\;
 %%         }
@@
 %%   \SetKwProg{Fn}{Function}{:}{}
-%%   \Fn{\Pick{QValid, QDisc, gen, mut}}{
+%%   \Fn{\Pick{QValid, QDiscarded, gen, mut}}{
 %%     \If{\KwNot Empty(QValid)}{
@@
 %%     }
-%%     \uElseIf{\KwNot Empty(QDisc)}{
-%%       (x,e) $\gets$ Deque(QDisc)\;
+%%     \uElseIf{\KwNot Empty(QDiscarded)}{
+%%       (x,e) $\gets$ Deque(QDiscarded)\;
 %%       \If{e $>$ 0}{
-%%         PushFront(QDisc, (x, e-1))
+%%         PushFront(QDiscarded, (x, e-1))
 %%       }
@@
 %
-If it assigns too much budget to certain not-so-interesting seeds, some bugs
-might not be discovered on a timely basis.
+If it assigns too much budget to certain not-so-interesting test cases, some
+bugs might not be discovered on a timely basis.
 %
@@
 
-\begin{figure}[b]
+\begin{figure}[t]
 \begin{verbatim}
-mutate (Leaf x)       = [ PURE (Branch def x def) ]
-mutate (Branch l x r) = [ PURE l, PURE r
+mutate (Leaf x) =
+  [ PURE (Branch def x def) ]
+mutate (Branch l x r) =
+  [ PURE l, PURE r
   , PURE (Leaf x)
-                        , PURE (Branch l x l)
-                        , PURE (Branch r x r)
-                        , PURE (Branch r x l) ]
+  , PURE (Branch l x l), PURE (Branch r x r), PURE (Branch r x l) ]
 \end{verbatim}
@@
   type.}
-%% \vspace{-15pt}
+\vspace{-5pt}
 \end{figure}
@@
 \begin{verbatim}
-positions (Leaf x)       = node [ (0, positions x) ]
-positions (Branch l x r) = node [ (0, positions l)
-                                , (1, positions x)
-                                , (2, positions r) ]
+positions (Leaf x) =
+  node [ (0, positions x) ]
+positions (Branch l x r) =
+  node [ (0, positions l), (1, positions x), (2, positions r) ]
 \end{verbatim}
@@
 
+\vspace{-5pt}
 \begin{equation}
@@
 %
-Like in CGPT, we use two queues, QValid and QDisc to keep valid and discarded mutable test cases,
+Like in CGPT, we use two queues, QValid and QDiscarded to keep valid and discarded mutable test cases,
 respectively.
@@
     i $\gets$ 0\;
-    TLog, QValid, QDisc $\gets$ $\varnothing$\;
+    TLog, QValid, QDiscarded $\gets$ $\varnothing$\;
     \While{i $<$ N}{
-      x $\gets$ Pick(QValid, QDisc, gen)\;
+      x $\gets$ Pick(QValid, QDiscarded, gen)\;
       (result, trace) $\gets$ WithTrace(P(x))\;
@@
           batch $\gets$ Mutate(x, mut, R)\;
-          Enqueue(QDisc, batch)\;
+          Enqueue(QDiscarded, batch)\;
         }
@@
   \SetKwProg{Fn}{Function}{:}{}
-  \Fn{\Pick{QValid, QDisc, gen}}{
+  \Fn{\Pick{QValid, QDiscarded, gen}}{
     \If{\KwNot Empty(QValid)}{
@@
       \If{Empty(batch)}{
-        Pick(QValid, QDisc, gen)\;
+        Pick(QValid, QDiscarded, gen)\;
       } \uElse{
@@
     $\cdots$\;
-    x $\gets$ Pick(QValid, QDisc, gen, mut)\;
+    x $\gets$ Pick(QValid, QDiscarded, gen, mut)\;
     (result, trace) $\gets$ WithTrace(P(x))\;
@@
 
+%% \vspace{-10pt}
 \section{Case studies}
@@
   \textit{haskell-wasm}.}
-%% \vspace{-20pt}
+\vspace{-10pt}
 \end{table}
@@
 \begin{verbatim}
-prop_validator m =
-  isValidHaskellWasm m ==> isValidRefImpl m
+prop_validator m = isValidHaskellWasm m ==> isValidRefImpl m
 \end{verbatim}
@@
 \begin{verbatim}
-prop_interpreter_i32 fun i f =
-  prop_interpreter (FuncType { params = [I32, F32], result = [I32] })
-                   fun
-                   [VI32 i, VF32 f]
+prop_interpreter_i32 fun i f = prop_interpreter
+  (FuncType { params = [I32, F32], result = [I32] }) fun [VI32 i, VF32 f]
 \end{verbatim}
@@
 
-In both case studies, we show how the FIFO scheduling and trace reset heuristics
-described in \S \ref{sec:heuristics} affect the testing performance of
-\mutagen by individually disabling them.
+In both case studies, we show how the two \mutagen heuristics described in \S
+\ref{sec:heuristics} affect the testing performance by individually disabling
+them.
 %
@@
 controlled by this heuristic, so we arbitrarily fixed it to 25 random mutations
-throughout the execution of the tool.
+throughout the testing loop.
 
@@
   each tool on the WebAssembly case study.}
-%% \vspace{-10pt}
+\vspace{-10pt}
 \end{table}
@@
 In this work we use exhaustiveness as a way to reliably enforce that all
-possible mutants of an interesting seed are executed.
+possible mutants of an interesting test case are executed.
 %
@@
 and evaluating the effect of lazy pruning; among others.
-%
-In addition, we want to investigate how to redefine our automatically
-synthesized mutators in a stateful manner.
-%
-This way, it would be possible to apply mutations that preserve complex
-properties of the generated data simply by construction, e.g., mutations that
-\emph{always} produce well-typed programs where variables are always in scope.
-%
-The main challenge will be to achieve this while keeping the testing process as
-automatable as possible.
+%% %
+%% In addition, we want to investigate how to redefine our automatically
+%% synthesized mutators in a stateful manner.
+%% %
+%% This way, it would be possible to apply mutations that preserve complex
+%% properties of the generated data simply by construction, e.g., mutations that
+%% \emph{always} produce well-typed programs where variables are always in scope.
+%% %
+%% The main challenge will be to achieve this while keeping the testing process as
+%% automatable as possible.
 

### 2021-08-26 papers/icse22/main.tex
Heavy compression. Paper now fits!

@@
 %
-\subsubsection{Limitations of CGPT}
+\paragraph{Limitations of CGPT}
 
@@
 Instead of mutating every numerical or character subexpression exhaustively, we
-resolve in using a random generator and sample a small number of values from it
---- the precise amount is a tunable parameter of \mutagen.
+sample a small number of values from a given random generator --- this amount is
+a tunable parameter of \mutagen.
 %
@@
 
-\subsubsection{Mapping top-level mutations everywhere}
+\paragraph{Mapping top-level mutations everywhere}
 
@@
 Having selected the next test case, the main loop proceeds to execute it,
-capturing both the result (valid, discarded due to preconditions, or failed)
+capturing both the result (valid, discarded, or failed)
 and its execution trace over the system under test.
 %
-If the test case fails, it is immediately reported as a bug.
+If the test case fails, it is reported as a bug.
 %
@@
 %
-If the test case does in fact discover a new path, it is enqueued on its
-corresponding queue, depending on whether it was valid or got discarded.
+If the test case discovers a new path, it is enqueued on its corresponding
+queue.
+% depending on whether it was valid or got discarded.
 %
@@
 %
-This process alternates between generation and mutation until a bug is found or
-we reach the test limit.
+This process alternates between generation and mutation until it finds a bug or
+reaches the test limit N.
 
@@
   \Fn{\MG{P, N, gen, mut}}{
-    {\color{red}
-    boring  $\gets$ 0\;
-    reset   $\gets$ 1000\;
-    R       $\gets$ 1\;
-    }
+    {\color{red} boring $\gets$ 0}; {\color{red} reset $\gets$ 1000}; {\color{red}R $\gets$ 1}\;
     $\cdots$\;
@@
 The first one is a simple abstract stack machine that enforces the
-hyper-property \emph{noninterference} \cite{goguen1982security} using runtime
-checks.
-%
-Its implementation was originally proven correct by
-\citeauthor{10.1145/2578855.2535839} \citeyearpar{10.1145/2578855.2535839} in
-Coq, and subsequently degraded by systematically introducing 20 bugs on its
-enforcing mechanism.
+\emph{noninterference} \cite{goguen1982security} using runtime checks.
+%
+Its implementation was proven correct by \citeauthor{10.1145/2578855.2535839}
+\citeyearpar{10.1145/2578855.2535839} in Coq, and subsequently degraded by
+systematically introducing 20 bugs on its enforcing mechanism.
 %
 \citeauthor{lampropoulos2019coverage} used this case study to compare \fuzzchick
-against random testing approaches using na\"ive and manually-written smart
-random generators.
-%
-In this work, we replicate their results and compare them against our tool.
-%
-Worth mentioning, since this case study was originally implemented in Coq, we
-first translated it to Haskell in order to run \mutagen on its test suite.
-
-The second case study aims to evaluate \mutagen in a realistic scenario,
-and focuses on testing \textit{haskell-wasm} \cite{haskellwasm}, an existing
-WebAssembly engine of industrial strength written in Haskell.
+against RPBT using na\"ive- and hand-tuned random generators.
+%
+Here, we replicate their results and compare them against our tool.
+%
+Worth mentioning, we translated this case study into Haskell in order to run
+\mutagen on its test suite.
+
+The second case study aims to evaluate \mutagen in a realistic scenario, and
+targets \textit{haskell-wasm} \cite{haskellwasm}, an existing WebAssembly engine
+of industrial strength written in Haskell.
 %
@@
 Sadly, comparing \mutagen with \fuzzchick on this case study has been out of the
-scope of this work.
-%
-The reason for this is that porting \textit{haskell-wasm} into Coq, and later
+scope of this work, since porting \textit{haskell-wasm} into Coq, and later
 adapting our test suite to work with \fuzzchick requires a subtantial effort.
 %
-Instead, we focus on comparing \mutagen against a more traditional RPBT approach
-using \quickcheck, to evaluate its effectiveness versus the relative overhead of
-the code instrumentation used in our tool.
+Instead, we compare \mutagen against a traditional RPBT approach, evaluating its
+effectiveness versus the relative overhead of the code instrumentation used by
+our tool.
 
@@
 %
-Finally, machine states can be modeled using a 4-tuple \textit{(pc, stk, m, i)}
-that represents a particular configuration of the program counter, the stack and
+Finally, machine states are modeled using a 4-tuple \textit{(pc, stk, m, i)}
+representing a particular configuration of the program counter, the stack, and
 the data- and instruction memories, respectively.
@@
 \citeauthor{hritcu2013testing} \citeyearpar{hritcu2013testing,
-  hrictcu2016testing} as well as the original CGPT paper
-\cite{lampropoulos2019coverage} for more details about the implementation and
-semantics of this case study.
-
-\subsubsection{Single-Step Noninterference (SSNI)}
+  hrictcu2016testing} and \citeauthor{lampropoulos2019coverage} for more details
+about the implementation and semantics of this case study.
+
+\paragraph{Single-Step Noninterference (SSNI)}
 
@@
 %
-Intuitively, two machine states are indistinguishable from each other if they
-only differ on secret data.
-%
-Using this notion, the particular variant of noninterference we are interested
-in this work is called \emph{single-step noninterference}
-\cite{hritcu2013testing} (SSNI).
-%
-SSNI asserts that, given two indistinguishable machine states, running a single
+Intuitively, two machine states are indistinguishable if they only differ on
+secret data.
+%
+Using this notion, the variant of noninterference we are interested in is called
+\emph{single-step noninterference} \cite{hritcu2013testing} (SSNI).
+%
+Given two indistinguishable machine states, SSNI asserts that running a single
 instruction on both machines brings them to resulting states that are also
@@
 
-\subsubsection{IFC Rules}
-
-The IFC rules enforced by this abstract machine are implemented using a rule
-table indexed by the instruction that the machine is about to perform.
-%
-This table contains: the dynamic check that the abstract machine needs to
-perform in order to enforce the IFC policy, along with the corresponding
-security label of the program counter and the instruction result after the
-operation is executed.
+\paragraph{IFC Rules}
+
+The IFC rules enforced by this machine are encoded in a table indexed by its
+different instructions.
+%
+This table contains: the dynamic check that the machine needs to perform to
+enforce the IFC policy, along with the corresponding security label of the
+program counter and the instruction result after the instruction is executed.
 %
@@
 %
-After this check, this machine overwrites the value at the destination cell
-and updates its label with the maximum sensibility of the involved labels.
+After this check, the machine overwrites the value at the destination cell and
+updates its label with the maximum sensibility of the involved labels.
 %
@@
 
+%% \vspace{-10pt}
 \begin{center}
@@
 \emph{maximum} of their sensibilities.
-
-Bugs can be systematically introduced in the IFC enforcing mechanism by removing
-or weakening the checks stored in this rule table.
+%
+Later, bugs are systematically introduced in the IFC enforcing mechanism by
+removing or weakening the checks stored in this rule table.
 
@@
 
-WebAssembly \cite{haas2017bringing} is a popular assembly-like language designed
-to be an open standard for executing low-level code in the web.%
-%
-\footnote{Although it has become increasingly popular in standalone, non-web
-  contexts as well.}
-%
-WebAssembly programs are first validated (using a simple form of type-checking)
-and later executed in a sandboxed environment (a virtual machine).
+WebAssembly \cite{haas2017bringing} is an assembly-like language designed for
+executing low-level code in the web.
+%% %
+%% \footnote{Although it has become increasingly popular in standalone, non-web
+%%   contexts as well.}
+%
+WebAssembly programs are first validated
+%
+%(using a simple form of type-checking)
+%
+and later executed in a sandboxed environment.
 %
@@
     IEEE754 floating-point numbers of either 32 or 64 bits;
-  \item values of these types are by manipulated by functions written using
+  \item values of these types are manipulated by functions written using
     sequences of stack instructions;
@@
 %% factorial function.
-
-WebAssembly semantics are fully specified, and programs are expected to be
-consistently interpreted across engines --- despite some subtle details that we
-will address soon.
-%
-Moreover, the WebAssembly standard provides a reference implementation with all
+%
+WebAssembly semantics are fully specified, and programs must be consistently
+interpreted across engines --- despite some subtle details that we will address
+soon.
+%
+For this, the WebAssembly standard provides a reference implementation with all
 the functionality expected from a compliant engine.
@@
 
-Our tool is an attractive match for testing WebAssembly engines, and
-\textit{haskell-wasm} in particular, as the space of WebAssembly programs that
-can be represented using its AST contains mostly invalid ones, and automatically
-derived random generators cannot satisfy all the invariants required to produce
-interesting test cases.
-
-In this work, we are interested in using \mutagen to test the two most complex
-subsystems of \textit{haskell-wasm}: the \emph{validator} and the
-\emph{interpreter} --- both being previously tested using a unit test suite.
-%
-We avoided spending countless hours writing an extensive property-based
-specification to mimic the reference WebAssembly specification.
-%
-Instead, we used the reference WebAssembly implementation to find discrepancies
-(that could potentially lead to bugs) via differential testing
-\cite{mckeeman1998differential}.
-%
-In this light, our testing properties assert that any result produced by
-\textit{haskell-wasm} matches that of the reference implementation.
-%
-Unsurprisingly, this engine had several latent bugs that were not caught by the
-existing unit tests and that we discovered by \mutagen while developing our test
-suite.
-%
-Moreover, \mutagen exposed two other discrepancies between \textit{haskell-wasm}
-and the reference implementation.
-%
-Not severe enough to be classified as bugs, these discrepancies trigger parts of
-the WebAssembly specification that are either not yet supported by the reference
-implementation (multi-value blocks), or that produce a well-known
-non-deterministic undefined behavior allowed by the specification (NaN
-reinterpretation) \cite{perenyi2020stack}.
-%
-These findings are briefly outlined in Table \ref{table:wasm:bugs}.
-%
-All of them were reported and later confirmed by the authors of
-\textit{haskell-wasm}.
-%
-Having sorted these issues out, we injected 10 new bugs in the validator as well
-as 5 new bugs in the interpreter of this engine (see Table
-\ref{table:wasm:injectedbugs}).
-%
-These bugs were mechanically injected by either
-%
-\begin{inparaenum}
-\item removing an existing integrity check (to weaken the WebAssembly
-  type-system/validator); or by
-\item simulating a copy-and-paste induced bug.%
-  \footnote{This kind of bugs was inspired by the real bug \#1 we found prior
-    this step.}
-  %
-  For instance, by replacing the implementation of an operation with a similar
-  compatible one (e.g., \texttt{i32.add} by \texttt{i32.sub}).
-\end{inparaenum}
-
-%% During this process, we quickly discovered 3 bugs and 2 discrepancies on this
-%% engine with respect to the reference implementation.
-%% %
-
 \begin{table}[t]
@@
   \textit{haskell-wasm}.}
-%% \vspace{-15pt}
+\vspace{-10pt}
 \end{table}
+
+Our tool is an attractive match for testing WebAssembly engines:
+%
+%% and \textit{haskell-wasm} in particular,
+%
+most of the programs that can be represented using WebAssembly's AST are
+invalid, and automatically derived random generators cannot satisfy the
+invariants required to produce interesting test cases.
+
+In this work, we are interested in using \mutagen to test the two most complex
+subsystems of \textit{haskell-wasm}: the \emph{validator} and the
+\emph{interpreter} --- both being previously tested using a unit test suite.
+
+We avoided spending countless hours writing an extensive property-based
+specification to mimic the WebAssembly specification.
+%
+Instead, we used the reference implementation to find discrepancies (that could
+potentially lead to bugs) via differential testing
+\cite{mckeeman1998differential}.
+%
+In this light, our testing properties assert that any result produced by
+\textit{haskell-wasm} matches that of the reference implementation.
+%
+Notably, this engine had several latent bugs that were not caught by the
+existing unit tests and that we discovered by \mutagen while developing our test
+suite.
+%
+Moreover, \mutagen exposed two other discrepancies between \textit{haskell-wasm}
+and the reference implementation.
+%
+These discrepancies trigger parts of the specification that are either not yet
+supported by the reference implementation (multi-value blocks), or that produce
+a well-known non-deterministic undefined behavior (NaN reinterpretation)
+\cite{perenyi2020stack}.
+%
+These findings are briefly outlined in Table \ref{table:wasm:bugs}.
+%
+All of them were reported and later confirmed by the authors of
+\textit{haskell-wasm}.
+%
+Having sorted these issues out, we injected 10 new bugs in the validator as well
+as 5 new bugs in the interpreter of this engine (see Table
+\ref{table:wasm:injectedbugs}).
+%
+These bugs were mechanically injected by either
+%
+\begin{inparaenum}
+\item removing an existing integrity check (to weaken the WebAssembly
+  type-system/validator); or by
+\item simulating a copy-and-paste induced bug,%
+  \footnote{This kind of bugs was inspired by the real bug \#1 we found prior
+    this step.}
+  %
+  replacing the implementation of an instruction with a compatible one
+  (e.g., \texttt{i32.add} by \texttt{i32.sub}).
+\end{inparaenum}
+
+%% During this process, we quickly discovered 3 bugs and 2 discrepancies on this
+%% engine with respect to the reference implementation.
+%% %
 
@@
 
-\subsubsection{Testing the WebAssembly Validator}
+\paragraph{Testing the WebAssembly Validator}
 
@@
 %
-In other words, we are testing for false negatives.
-%
-In Haskell, we write the following property:
+%% In other words, we are testing for false negatives.
+%
+In Haskell, we write the property:
 
@@
 
-Where the precondition (\texttt{isValidHaskellWasm m}) runs the input WebAssembly
-module \texttt{m} against the \textit{haskell-wasm} validator, whereas the
-postcondition (\texttt{isValidRefImpl m}) serializes \texttt{m} to a file, runs it
-against the reference implementation validator and checks that no errors are
-produced.
-
-We note that, although we only focus on finding false negatives, in a complete
-test suite, one would also want to test for false positives, i.e., when a module
-is valid and \textit{haskell-wasm} rejects it.
+\noindent The precondition (\texttt{isValidHaskellWasm m}) runs the input
+WebAssembly module \texttt{m} against \textit{haskell-wasm}'s validator, whereas
+the postcondition (\texttt{isValidRefImpl m}) serializes \texttt{m}, runs it
+against the reference implementation and checks that no errors are produced.
+
+We note that, although we only focus on finding false negatives, a comprehensive
+test suite should also test for false positives, i.e., when a module is valid
+and \textit{haskell-wasm} rejects it.
 %
@@
 %
-However, the resulting property is much slower, as every test case will always
-be serialized and run against the reference implementation.
+%% However, the resulting property is much slower, as every test case will always
+%% be serialized and run against the reference implementation.
 
@@
 
-\subsubsection{Testing the WebAssembly Interpreter}
+\paragraph{Testing the WebAssembly Interpreter}
 
 Testing the WebAssembly interpreter is substantially more complicated than
-testing the validator, since it requires running and comparing the output of
-actual programs.
-%
-To achieve this, we need the generated test cases to comply a with a common
+testing the validator, since it requires actually running and comparing the
+output of test case programs.
+%
+To achieve this, the generated test cases need to comply a with a common
 interface that can be invoked both by \textit{haskell-wasm} and the reference
-WebAssembly implementation.
-
-To simplify things here as well, we write a function \texttt{mk\_module} to
-build a stub module which initializes one memory block and exports a single
-WebAssembly function.
-%
-\texttt{mk\_module} is parameterized by the definition of the single WebAssembly
-function, along with its name and type signature.
+implementation.
+%
+To simplify things, we write a function \texttt{mk\_module} to build a stub
+module which initializes one memory block and exports a single WebAssembly
+function.
+%
+Such function is parameterized by the definition of the single function, along
+with its name and type signature.
 %
@@
 %
-This way, we can define a testing property that can be instantiated with
-randomly generated combinations of function types, function definitions and
-lists of invocation arguments:
+Then, we can use \texttt{mk\_module} to define a testing property parameterized
+by a function type, along with its definition and invocation arguments:
+
+%% define a testing property parameterized by a function type, a
+%% function body and invocation arguments:
 
@@
 
-This property instantiates the module stub using the input function and its type
-signature, and uses it to invoke the function \texttt{foo} both on the
-\textit{haskell-wasm} and reference implementation interpreter with the provided
-arguments.
+This property instantiates a module stub \texttt{m} using the input function
+(\texttt{fun}) and its type signature (\texttt{ty}).
+%
+Then, it invokes the function \texttt{foo} of the module \texttt{m} both on
+\textit{haskell-wasm} and the reference interpreter with the provided arguments
+(\texttt{args}).
 %
@@
 %
-Interestingly, equivalence in this context does not imply equality.
+Interestingly, equivalence in does not imply equality.
 %
@@
 
-This specialized property lets us generate functions using this fixed type and
-invoke them with the exact number and type of arguments required.
-%
-In our experiments (presented in the next section), we use this property when
-finding all the injected bugs into the \textit{haskell-wasm} interpreter.
-%
-As before, we remark that a complete test suite should at least include
-different variants of this property testing functions of several different
-types, as well as properties testing multiple functions simultaneously.
+\noindent This property lets us generate functions with a fixed type and invoke
+them with the exact number and type of arguments required.
+%
+%% In our experiments (presented in \S \ref{sec:evaluation}),
+%
+We use this property when finding all the injected bugs into the
+\textit{haskell-wasm} interpreter in the next section.
+%
+%% As before, we remark that a complete test suite should at least include
+%% different variants of this property testing functions of several different
+%% types, as well as properties testing multiple functions simultaneously.
 
@@
 
-We performed all experiments in a dedicated workstation with an Intel Core
-i7-8700 CPU running at 3.20GHz, and equipped with 32GB of RAM.
-%
-Each experiment was repeated 30 times except for the ones testing the
-WebAssembly interpreter, which were run 10 times due to time constraints.
+We performed all experiments in a dedicated workstation with 32GB of RAM and an
+Intel Core i7-8700 CPU running at 3.20GHz.
+%
+We repeated each experiment 30 times except for the ones testing the WebAssembly
+interpreter, which were run 10 times due to time constraints.
 %
@@
 %
-In addition, we collected the failure rate observed for each bug, i.e. the
+In addition, we collected the failure rate (FR) observed for each bug, i.e. the
 proportion of times each tool finds each bug within the one-hour testing budget.
@@
 
-\begin{figure}[t]
+\begin{figure}[b]
   \centering
@@
   \includegraphics[width=\linewidth]{tikz/ifc-fr.tikz}
-  %% \vspace{-20pt}
+  \vspace{-20pt}
   \caption{\label{fig:results:ifc} Comparison between \fuzzchick and \mutagen
     across 20 different bugs for the IFC stack machine. }
-%% \vspace{-10pt}
+\vspace{-5pt}
 \end{figure}
 
-The results of this case study are shown in Fig. \ref{fig:results:ifc} --- we
-encourage the reader to obtain a color copy of this work.
+The results of this case study are shown in Fig. \ref{fig:results:ifc}.
+%
+%% --- we encourage the reader to obtain a color copy of this work.
 %
@@
 %
-These results are somewhat pessimistic towards \fuzzchick if compared to the
-ones presented originally by \citeauthor{lampropoulos2019coverage}.
-%
-However, we remark that our experiments encompass 30 independent runs instead of
-the 5 originally used when presenting \fuzzchick
-\citeyearpar{lampropoulos2019coverage}.
+While these results are somewhat pessimistic towards \fuzzchick if compared to
+the ones presented originally by \citeauthor{lampropoulos2019coverage}, we
+remark that our experiments encompass 30 independent runs instead of the 5 used
+in the original study.
 %
 Moreover, since the MTTF simply aggregates all runs, regardless of if they found
-a bug or timed out, this metric is quite sensitive to the timeout used on each
+a bug or timed out, this metric is sensitive to the timeout used on each
 experiment, and will inflate the results as soon as the failure rate goes below
@@
 %
-For this reason, comparing the failure rate between tools give us a better
-estimation of the reliability of both tools, where \mutagen shows a clear
-improvement.
+For this reason, comparing the failure rate between tools gives a better
+estimation of their reliability, where \mutagen shows a clear improvement.
 
 %% If we only consider the succesfull runs where \fuzzchick does not time out, we
-After analyzing the results obtained using \fuzzchick, we observed a
+After carefully analyzing the results obtained using \fuzzchick, we observed a
 peculiarity.
 %
-This tool either finds bugs relatively quickly (after a few thousands tests) or
-does not find them at all within the one-hour budget, suggesting that its power
-scheduler assigns some enqueued test cases too little energy before they become
-uninteresting and get ultimately thrown away.
-%
-Under this consideration, it is fair to think that \fuzzchick might be a better
-choice if we set a shorter timeout and run it several times.
-%
-Thus, to be an improvement over \fuzzchick, our tool must be able to find bugs
-not only more reliably, but also relatively fast!
-
-Table \ref{table:ifc:tests} shows a comparison between the mean number of tests
-required by both tools to find the first failure \emph{when we only consider
-  successful runs.}
-%
-Additionally, as advised by \citeauthor{arcuri2014hitchhiker}
-\cite{arcuri2014hitchhiker} when comparing random testing tools, we computed the
-non-parametric Vargha-Delaney $A_{12}$ measure \cite{vargha2000critique}.
-%
-Intuitively, this measure encodes the probability of \mutagen to yield better
-results than \fuzzchick.
-%
-The \textit{Estimate} of this measure simply denotes the statistical difference
-between the compared distributions in a categorical manner.
-%
-Estimates in {\color{blue} blue} and {\color{red} red} indicate results
-favoring \fuzzchick and \mutagen, respectively.
-%
-As it can be observed,
-%
-%% using an exhaustive mutation approach can be not only
-%% more reliable, but also faster.
-%
-\mutagen is likely to find bugs faster than \fuzzchick in 14 of the 20 bugs
-injected into the abstract machine, while being substantially slower in only 2
-cases.
+This tool either finds bugs quickly (after a few thousands tests) or does not
+find them at all within the one-hour budget.
+%% , suggesting that its power
+%% scheduler assigns some enqueued test cases too little energy before they become
+%% uninteresting and get ultimately thrown away.
+%
+With this in mind, it is fair to think that \fuzzchick might be a better choice
+if we set a shorter timeout and run it several times.
+%
+Thus, to be an improvement over \fuzzchick, our tool must find bugs not only
+more reliably, but also fast!
 
@@
 %% \centering
-\vspace{-15pt}
+\vspace{-10pt}
 \scriptsize
@@
 
-In terms of the \mutagen heuristics, we can also arrive to some conclusions.
-%
-Firstly, this case study does not seem to be strongly affected by using a FIFO
-scheduling, where the same benchmark takes only roughly 20\% longer to complete
-when disabling this heuristic.
-%
-Upon inspection, the reason behind this is that the mutation queues remain empty
-most of the time (generation mode), and when a new interesting candidate gets
-inserted, all its mutants (and their descendants) are processed before the next
-one is enqueued.
+Table \ref{table:ifc:tests} shows a comparison between the mean number of tests
+required by both tools to find the first failure \emph{when we only consider
+  successful runs.}
+%
+Additionally, as advised by \citeauthor{arcuri2014hitchhiker}
+\cite{arcuri2014hitchhiker}, we computed the non-parametric Vargha-Delaney
+$A_{12}$ measure \cite{vargha2000critique}.
+%
+Intuitively, this measure encodes the probability of \mutagen to yield better
+results than \fuzzchick.
+%
+The \textit{Estimate} of this measure simply denotes the statistical difference
+between the compared distributions in a categorical manner.
+%
+Estimates in {\color{blue} blue} and {\color{red} red} indicate results
+favoring \fuzzchick and \mutagen, respectively.
+%
+As it can be observed,
+%
+%% using an exhaustive mutation approach can be not only
+%% more reliable, but also faster.
+%
+\mutagen is likely to find bugs faster than \fuzzchick in 14 of the 20 bugs
+injected into the abstract machine, being substantially slower in only 2 cases.
+
+In terms of the \mutagen heuristics,
+%
+%% we can also arrive to some conclusions.
+%% %
+%% Firstly,
+%
+this case study is not strongly affected by using FIFO scheduling.
+%% --- the same
+%% benchmark takes only roughly 20\% longer to complete when disabling this
+%% heuristic.
+%
+The reason behind this is that the mutation queues remain empty most of the time
+(generation mode), and when a new interesting candidate gets inserted, all its
+mutants (and their descendants) are processed before the next one is enqueued.
 %
@@
 
-On the other hand, disabling the trace saturation heuristic (\textit{no reset})
-we observed two interesting effects:
-%
-having fixed the number of random mutations to 25 adds a seemingly constant
-overhead when finding most of the (easier) bugs, suggesting that we are spending
-worthless time mutating numeric subexpressions inside the generated machine
-states and that a smaller number of random mutations could be equally effective
-to find such bugs.
+Moreover, disabling the trace saturation heuristic (\textit{no reset}) we
+observed two effects:
+%
+fixing the number of random mutations to 25 adds a constant overhead when
+finding most of the (easier) bugs, suggesting that we are spending too time
+mutating numeric subexpressions inside the generated machine states.
+%
+%% and that a
+%% smaller number of random mutations could be equally effective to find such bugs.
 %
@@
 %
-In consequence, we believe this heuristic is effective at automatically tuning
-this internal parameter of our tool, especially when the user is unsure about
-what the best value for it might be.
+In consequence, we argue that our heuristic is effective at automatically tuning
+this parameter.
+%
+%% , especially when the user is unsure about what the best value for it might be.
 
@@
 by the MTTF achieved by \mutagen.
-%
-Given that we used different properties to test the WebAssembly validator and
-interpreter, we will focus on these results independently.
+%% %
+%% Given that we used different properties to test the WebAssembly validator and
+%% interpreter, we will focus on these results independently.
 
@@
 %
-There, we can quickly conclude that \quickcheck is not well suited to find most
-of the bugs --- it merely finds a counterexample for the easier bugs \#4 and \#5
-in just 1 out of 30 runs!
-%
-The reason behind this simple: WebAssembly modules comprise several different
-interconnected components, e.g., types and functions definitions, memory blocks,
-imports, exports, etc.
-%
-All these components need to be valid and agree with each other in order for the
-module to be correctly validated (modulo the injected bugs).
-%
-Using an automatically derived generator that generates each one of these
-components independently is virtually unable to produce valid WebAssembly
-modules apart from the trivial empty one.
+There, we quickly conclude that \quickcheck is not well suited to find most of
+the bugs --- it merely finds the easier bugs \#4 and \#5 in just 1 out of 30
+runs.
+%
+The reason behind this simple:
+%
+%% WebAssembly modules comprise several different
+%% interconnected components, e.g., types and functions definitions, memory blocks,
+%% imports, exports, etc.
+%% %
+%% All these components need to be valid and agree with each other in order for the
+%% module to be correctly validated (modulo the injected bugs).
+%
+using an automatically derived generator is virtually unable to produce valid
+WebAssembly modules apart from the trivial empty one.
 %
@@
 
-In terms of the heuristics implemented in our tool, we can observe two clear
-phenomena.
+In terms of the \mutagen heuristics, we observe two clear phenomena.
 %
@@
   reset}) and using a fixed number of 25 random mutation does not affect the
-effectiveness of our tool, apart from adding again a mostly constant overhead on
-the time required to find every bug --- suggesting that the value we arbitrarily
-chose for this parameter is too large and should be left to be handled
-automatically by our tool.
+effectiveness of our tool, apart from adding again a constant time overhead.
+%% ---
+%% suggesting that the value we arbitrarily chose for this parameter is too large
+%% and should be left to be handled automatically by our tool.
 
@@
 %
-Perhaps more interestingly, we can observe a significant improvement in the
-performance of \quickcheck in terms of failure rate.
+We can also observe a significant improvement in the performance of \quickcheck
+in terms of failure rate.
 %
@@
 %
-The reason behind this is that this bug can be found using a rather small
+The reason behind this is that this bug can be found using a very small
 counterexample, and \quickcheck prefers sampling small test cases at the
@@
 Moreover, by disabling the FIFO scheduling we observe that the performance of
-our tool gets more unstable, where the bugs \#12 and \#11 take substantially
-less time to be found, whereas the bugs \#15 and \#14 cannot be found on every
-run.
+our tool becomes unstable, where bugs \#12 and \#11 take substantially less time
+to be found, whereas bugs \#15 and \#14 cannot be found on every run.
 %
@@
 
-\begin{figure}
+\begin{figure}[t]
   \centering
@@
   \includegraphics[width=\linewidth]{tikz/wasm-legend.tikz}
-  %% \vspace{-20pt}
+  \vspace{-20pt}
   \caption{\label{fig:results:wasm} Comparison between \quickcheck and \mutagen
@@
 
-Finally, this case study allows us to analyze the overhead introduced by the
-code instrumentation and internal processing used in \mutagen versus the
-stateless black-box approach of \quickcheck.
-%
-Table \ref{table:wasm:overhead} compares the total number of executed and passed
-tests per second that we observed using each tool.
-%
-It is easy to conclude that \mutagen executes tests several times slower than
-\quickcheck --- roughly 20x and 75x slower when testing \texttt{prop\_validator}
-and \texttt{prop\_interpreter\_i32}, respectively.
-%
-This slowdown is partly caused because the code instrumentation our tool
-currently uses is built upon a superficial transformation that logs traces at
-the user level.
-%
-We anticipate that this overhead could be reduced considerably by integrating
-the instrumentation directly into the Haskell runtime --- a substantial
-challenge to tackle in the future.
-%
-Despite this, our tool is still capable of running substantially more tests that
-pass the sparse preconditions than \quickcheck in the same amount of time,
-\emph{and which ultimately leads us to find bugs}.
-
-\begin{table}
+\begin{table}[t]
 \footnotesize
@@
 
+Finally, this case study allows us to analyze the overhead introduced by the
+code instrumentation and internal processing used in \mutagen versus the
+stateless black-box approach of \quickcheck.
+%
+Table \ref{table:wasm:overhead} compares the total number of executed and passed
+tests per second that we observed using each tool.
+%
+\mutagen executes tests several times slower than \quickcheck --- roughly 20x
+and 75x slower when testing \texttt{prop\_validator} and
+\texttt{prop\_interpreter\_i32}, respectively.
+%% %
+%% This slowdown is partly caused because the code instrumentation our tool
+%% currently uses is built upon a superficial transformation that logs traces at
+%% the user level.
+%% %
+%% We anticipate that this overhead could be reduced considerably by integrating
+%% the instrumentation directly into the Haskell runtime --- a substantial
+%% challenge to tackle in the future.
+%
+Despite this, our tool is still capable of running substantially more tests that
+pass the sparse preconditions than \quickcheck in the same amount of time,
+\emph{and that ultimately leads us to find bugs}.
+
 % ----------------------------------------
@@
 %
-In this section we will focus on three main topics:
+In this section we focus on three main topics:
 %
@@
 %
-exhaustive bounded testing, all of which inspired the design of \mutagen.
+exhaustive bounded testing, all of which inspired this work.
 
@@
 
-Obtaining good random generators automatically from static information (e.g.,
+Obtaining random generators automatically from static information (e.g.,
 grammars or data type definitions) has been tackled from different angles.
@@
 % DRAGEN and DRAGEN2
-In Haskell, DRAGEN \cite{DBLP:conf/haskell/MistaRH18} is a meta-programming tool
-that synthesises random generators from data types definitions, using stochastic
+DRAGEN \cite{DBLP:conf/haskell/MistaRH18} is a meta-programming tool that
+synthesises random generators from data types definitions, using stochastic
 models to predict and optimize their distribution based on a target one set by
@@
 %
-% Bolztmann-brain
-Similarly, \cite{Bendkowski2017} developed a polynomial tuning mechanism based
-on Boltzmann samplers \cite{Duchon2004} that synthesizes generators with
-approximate-size distributions for combinatorial structures like lists or trees.
-
+%% % Bolztmann-brain
+%% Similarly, \cite{Bendkowski2017} developed a tuning mechanism based on Boltzmann
+%% samplers \cite{Duchon2004} that synthesizes generators with approximate-size
+%% distributions for combinatorial structures like lists or trees.
+%
 %
@@
 \emph{QuickFuzz} \cite{GriecoCB16, grieco2017} is a fuzzer that exploits these
-ideas and uses existing Haskell libraries describing common file formats to
-synthesize random data generators that are used in tandem with existing
-off-the-shelf low-level fuzzers to find bugs in heavily used programs.
+ideas to synthesize random generators from existing Haskell libraries and that
+are used in tandem with existing low-level fuzzers to find bugs in heavily used
+programs.
 
@@
 Recently, \citeauthor{Lampropoulos2017} \citeyearpar{Lampropoulos2017} proposed
-a derivation mechanism to obtain constrained random generators directly by
-defining the target structure using inductively defined relations in Coq.
+a mechanism to obtain constrained random generators by defining their target
+structure using inductively defined relations in Coq.
 
@@
 %
-In principle, our tool is tailored to improve the performance of poor
-automatically derived generators.
-%
-However, more specialized generators can be used directly by \mutagen, and
-combining them with type-preserving mutations is an idea we aim to address in
-the future.
+While our tool is tailored to improve the performance of poor automatically
+derived generators, specialized ones can be used directly by \mutagen, and
+combining them with type-preserving mutations is part of our future work.
 
@@
 % AFLfast
-\emph{AFLFast} \cite{bohme2017coverage} extends AFL with Markov chain models to
+\emph{AFLFast} \cite{bohme2017coverage} extends AFL using Markov chain models to
 tune the power scheduler towards testing low-frequency paths.
 %
-In \mutagen, the scheduler does not account for path frequency.
+In our tool, the scheduler does not account for path frequency.
 %
@@
 % CollAFL
-Similar to our tool, \emph{CollAFL} \cite{gan2018collafl} is a variant of AFL
+Similar to \mutagen, \emph{CollAFL} \cite{gan2018collafl} is a variant of AFL
 that uses path- instead of edge-based coverage to distinguish executions more
@@
 %% to overcome this limitation.
-
+%
 All these ideas can potentially be incorporated in our tool, and we keep them as
@@
 % Korat
-On the object-oriented realm, \emph{Korat} \cite{boyapati2002korat} is a Java
-tool that uses method specification predicates to automatically generate all
-non-isomorphic test cases up to a given small size.
-
-Being exhaustive, these approaches rely on pruning mechanisms to avoid
-populating unevaluated subexpresions exhaustively before the computational cost
+\emph{Korat} \cite{boyapati2002korat} is a Java tool that uses method
+specification predicates to automatically generate all non-isomorphic test cases
+up to a given small size.
+
+These approaches rely on pruning mechanisms to avoid populating unevaluated
+subexpresions exhaustively before the computational cost of being exhaustive
 becomes too restrictive.
@@
 %
-In contrast to fully exhaustive testing tools, \mutagen relies on randomness to
+In contrast to fully-exhaustive testing tools, \mutagen relies on randomness to
 find initial interesting mutable test cases.
@@
 In terms of pruning, it is possible to instruct \mutagen to detect unused
-subexpressions (using a technique similar to that of \emph{LazySmallCheck}) to
-avoid scheduling mutations over their corresponding positions.
-%
-This could, in principle, improve the overall performance when testing
-properties that tend to short-circuit, leaving parts of the input unevaluated.
-%
-In our case studies, however, we observed that their preconditions tend to be
-quite strict when executing test cases obtained by mutating valid existing ones,
+subexpressions to avoid producing mutations over their corresponding positions.
+%
+This could improve the overall performance when testing non-strict properties
+that leave parts of their input unevaluated.
+%
+In our case studies, however, we observed that properties we tested
+%% their preconditions tend to be
+are quite strict when executing test cases obtained by mutating valid existing ones,
 fully evaluating their input before executing the postcondition.
 %
-Thus, we do not forsee considerable improvements by applying this idea in this
-particular scenario.
-%
-Gathering empirical evidence about using pruning via lazy evaluation in \mutagen
-is an effort that we keep as future work.
+Thus, we do not forsee considerable improvements by applying lazy-prunning in
+this scenario --- gathering evidence about lazy pruning in \mutagen is as part
+of our future work.
 
@@
 %
-Inspired by exhaustive bounded testing tools, our tool extends the CGPT approach
-introduced by \citeauthor{lampropoulos2019coverage} with an exhaustive mutation
-mechanism that generates every possible mutant for each interesting test case,
-and schedules it to be tested exactly once.
+%% Inspired by exhaustive bounded testing tools,
+%
+Our tool extends the CGPT approach introduced by
+\citeauthor{lampropoulos2019coverage} with an exhaustive mutation mechanism that
+generates every possible mutant for each interesting test case, and schedules it
+to be tested exactly once.
 
@@
 % ----------------------------------------
+
+\balance
 

### 2021-09-01 papers/icse22/main.tex
General pass

@@
 \title[\mutagen]{\mutagen: Coverage-Guided, Property-Based Testing using
-  Exhaustive Type-Preserving Mutations}
+  Exhaustive Structure-Preserving Mutations}
 
@@
 Coverage-Guided, Property-Based Testing (CGPT) \cite{lampropoulos2019coverage}
-is a technique that borrows ideas from the fuzzing community to generate highly
-structured values while still using automatically derived generators.
+is a technique that borrows ideas from the fuzzing community to generate
+highly-structured values while still using automatically derived generators.
 %
@@
 The data types used by the inputs of the testing properties already describe the
-structure of the random data we want to mutate in a concrete manner, and
+structure of the random data one needs to mutate in a concrete manner, and
 specialized mutators can be automatically derived directly from their definition
@@
 This case study also help us to compare the performance of our tool against a
-more traditional PBT approach that does not rely on code instrumentation.
+more traditional RPBT approach that does not rely on code instrumentation.
 %
@@
 For instance, the following testing property can be used to find bugs in the
-implementation of the \texttt{insert} function:
+implementation of a given \texttt{insert} function:
 
@@
 %% operation:
-A simple implementation of such a generator (simplified to make it more
-accessible%
+We can write a na\"ive one (simplified to make it more accessible%
 \footnote{In practice, this code will most likely be implemented using Haskell's
-  type-class mechanism, hiding the complexity behind the size limit and the payload generator.}%
-) can be written as follows:
+  type-class mechanism, hiding the complexity behind the size limit and the
+  payload generator.}%
+) as follows:
 
@@
 
+%% \begin{verbatim}
+%% genTree size = case size of
+%%   0 -> Leaf <$> genA
+%%   n -> oneof [ Leaf <$> genA
+%%              , Branch <$> genTree (n-1) <*> genA <*> genTree (n-1) ]
+%% \end{verbatim}
+
 \begin{verbatim}
-genTree 0 = Leaf <$> genA
-genTree n = oneof
-  [Leaf <$> genA, Branch <$> genTree(n-1) <*> genA <*> genTree(n-1)]
+genTree(sz) =
+  if sz == 0 then Leaf <$> genA
+  else oneof [ Leaf <$> genA
+             , Branch <$> genTree(sz-1) <*> genA <*> genTree(sz-1) ]
 \end{verbatim}
@@
 \texttt{Leaf} or a \texttt{Branch}), and then calls itself for every recursive
-subexpression needed to produce a well-typed value, carefully reducing the size
-limit by a unit on each step, ensuring termination by restricting itself to
-generating leaves when it reaches zero.
+subexpression needed to produce a well-typed value, carefully reducing the input
+size limit \texttt{sz} by a unit on each step, ensuring termination by
+restricting itself to generating leaves when it reaches zero (case \texttt{sz ==
+  0}).
 %
@@
 On each iteration, the algorithm selects the next test case by mutating the
-first value on the queue of valid test cases or, alternatively, from the queue
-of discarded test cases if the former one is empty.
-%
-If none of the queues contain test cases to mutate, CGPT generates a new value
-from scratch.
-%
-Then, the algorithm run this test case and evaluates whether it was interesting
+first value on the queue of valid test cases. If that queue is empty, it then
+selects the first one from the queue of discarded test cases.
+%
+If both queues are empty, CGPT generates a new random value from scratch.
+%
+The algorithm then run this test case and evaluates whether it was interesting
 (i.e., it exercises a new path) based on its trace information.
@@
 %
-In this light, a mutation can happen at top-level or be recursively applied to
-an immediate subexpression with approximately the same probability.
+This means that mutations can happen at the top-level or be recursively applied
+to an immediate subexpression of the input test case with approximately the same
+probability.
 %
@@
 %
-This makes deep recursive mutations very unlikely, since their probability
+This makes deep recursive mutations very unlikely, as their probability
 decreases multiplicatively with each recursive call.
@@
 Hence, these simple mutators can only be effectively used to transform shallow
-data structures, potentially excluding interesting applications that might
-require producing deeper valid values.
+data structures, excluding applications that might require producing deeper
+mutations.
 %
@@
 To keep the comparison fair, the authors replicated the same power schedule
-configuration used in AFL.
-%
-However, AFL uses a bit-level mutation approach.
+configuration used in AFL, which is tailored to work in tandem with generic
+bit-level mutators.
 %
@@
 %
-In our example, the default value of a \texttt{Tree} is \texttt{Leaf def}, where
-the inner \texttt{def} is the default value of the payload --- we use the type
-class system to abstract this complexity away in our implementation.
+In our example, the default value of a \texttt{Tree} is a leaf containing the
+smallest value of the tree payload type.
+%
+%% \texttt{Leaf def}, where
+%% the inner \texttt{def} is the default value of the payload.%
+\footnote{We use the type class system to abstract this complexity away in our
+  implementation.}
 %
 Using a small default value, as oppossed to a randomly generated one (as done by
-the original GCPT), is again inspired by exhaustive bounded testing tools, and
-avoids introducing unnecessary randomness when growing mutated expressions.
+the original CGPT), is also inspired by exhaustive bounded testing tools, and
+avoids introducing unnecessary randomness when ``growing'' mutated expressions.
 
@@
 Instead of mutating every numerical or character subexpression exhaustively, we
-sample a small number of values from a given random generator --- this amount is
-a tunable parameter of \mutagen.
+sample a small number of values from a given random generator.
+%
+This amount is a tunable parameter of \mutagen.
 %
@@
 \begin{verbatim}
-mutate n = [ RAND gen_int ]
+mutate n = [ RAND genInt ]
 \end{verbatim}
@@
 %
-To do so, we will use two functions that can be derived from the code.
+To do so, we use two utility functions that can be automatically synthesized.
 
@@
 
-\noindent Later, given the desired position to mutate within a test case, we
-define another function \texttt{inside} that finds the subexpression
-corresponding to it and applies the mutation.
+Later, we define a function \texttt{inside} that takes a desired
+position within an input test case and mutates its corresponding subexpression.
+%
+%% given the desired position to mutate within a test case, we
+%% define another function \texttt{inside} that finds the subexpression
+%% corresponding to it and mutates it.
 %
@@
 \begin{verbatim}
-inside []     x              = mutate x
+inside []     t              = mutate t
 inside (0:ps) (Leaf x)       = [ Leaf x'       | x' <- inside ps x ]
@@
 \noindent This function simply traverses the desired position, calling itself
-recursively until it reaches the desired subexpression, where the mutation can
-be applied locally at the top-level (case \texttt{inside [] x}).
+recursively until it reaches the desired subexpression, where a \texttt{mutate}
+can be applied locally at the top-level (case \texttt{inside [] t}).
 %
@@
 
-Having the exhaustive mutation mechanism in place, we can finally introduce the
-base testing loop used by \mutagen.
+Having the exhaustive mutation mechanism in place, we can now introduce the base
+testing loop used by \mutagen.
 %
@@
 
-Having selected the next test case, the main loop proceeds to execute it,
-capturing both the result (valid, discarded, or failed)
-and its execution trace over the system under test.
+Having selected the next test case, the testing loop proceeds to execute it,
+capturing both the result (valid, discarded, or failed) and its execution trace
+over the system under test.
 %
@@
 new path) based on its trace information and the one from previously executed
-test cases (represented by \textit{TLog}).
+test cases (represented by TLog).
 %
@@
 %
-Our assumption is that test cases that differ (branch) at shallower depths from
+Our assumption is that test case executions that branch at shallower depths from
 the ones already executed are more likely to discover new portions of the code
@@
 
-\begin{algorithm}[t]
-  \SetArgSty{textnormal}
-  %% \SetInd{0em}{0.75em}
-  \SetKw{KwNot}{not}
-  \SetAlgoLined
-  \DontPrintSemicolon
-  \SetKwFunction{MG}{Loop}
-  \SetKwProg{Fn}{Function}{:}{}
-  \Fn{\MG{P, N, R, gen, mut}}{
-    $\cdots$\;
-    x $\gets$ Pick(QValid, QDiscarded, gen, mut)\;
-    (result, trace) $\gets$ WithTrace(P(x))\;
-    $\cdots$\;
-    \If{Interesting(TLog, trace)}{
-      \If{\KwNot Discarded(result)}{
-        batch $\gets$ Mutate(x, mut, R)\;
-        {\color{red} prio $\gets$ BranchDepth(TLog, trace)}\; %% NEW
-        {\color{red} PushFront(QValid, prio, batch)}\;          %% NEW
-      }
-      $\cdots$\;
-    }
-  }
-  %% \vspace{5pt}
-  \SetKwFunction{Pick}{Pick}
-  \SetKwProg{Fn}{Function}{:}{}
-  \Fn{\Pick{QValid, QDiscarded, gen}}{
-    \If{\KwNot Empty(QValid)}{
-      {\color{red} (batch, prio) $\gets$ DequeMin(QValid)}\; %%NEW
-      \If{Empty(batch)}{
-        Pick(QValid, QDiscarded, gen)\;
-      } \uElse{
-        {\color{red} PushFront(QValid, prio, Rest(batch))}\; %% NEW
-        \KwRet First(batch)\;
-      }
-    }
-    $\cdots$\;
-  }
-
-\caption{\label{algo:mutagen:fifo}Priority FIFO Heuristic}
-\end{algorithm}
-
-In this light, every time we insert a new execution path into the internal trace
-log, we calculate the number of \emph{new nodes} that were executed, as well as
-the \emph{branching depth} where they got inserted.
-%
-The former is used to distinguish interesting test cases (whether or not new
-nodes were inserted), whereas the latter is used by this heuristic to schedule
-mutation candidates.
-%
-Fig. \ref{fig:mutagen:tracelog} illustrates this idea, inserting two execution
-traces (one after another) into a trace log that initially contains a single
-execution path.
-%
-The second insertion (with trace $1 \rightarrow 2 \rightarrow 6 \rightarrow 7$)
-branches at a shallower depth than the first one (2 vs. 3), and so its
-corresponding test case should be given a higher priority.
-
-Using this mechanism, we can modify \mutagen's base testing loop by replacing
-each mutation queue with a priority queue indexed by the branching depth of each
-new execution.
-%
-These changes are illustrated in Algorithm \ref{algo:mutagen:fifo}.
-%
-Statements in {\color{red} red} indicate important changes to the base
-algorithm, whereas ellipses denote parts of the code that are not relevant for
-the point being made.
-
-To pick the next test case, we retrieve the one with the highest priority (the
-smallest branching depth).
-%
-Then, when we find a new interesting test case, it gets enqueued at the
-\emph{beginning} of the queue at its corresponding priority.
-%
-This allows the testing loop to jump immediately onto mutating new interesting
-test cases as soon as they are found (even at the same priority), and to jump
-back to previous test cases as soon as their mutants become less novel.
-
-%% \caption{\label{fig:heuristics} The two main heuristics implemented in \mutagen.
-%%   Statements in {\color{red} red} indicate important changes to the base
-%%   algorithm. Ellipses denote parts of the code that are not relevant for the
-%%   point being made. }
-
-\begin{figure}[t]
+\begin{figure}[b]
 %% \vspace{-15pt}
@@
 
+\begin{algorithm}[t]
+  \SetArgSty{textnormal}
+  %% \SetInd{0em}{0.75em}
+  \SetKw{KwNot}{not}
+  \SetAlgoLined
+  \DontPrintSemicolon
+  \SetKwFunction{MG}{Loop}
+  \SetKwProg{Fn}{Function}{:}{}
+  \Fn{\MG{P, N, R, gen, mut}}{
+    $\cdots$\;
+    x $\gets$ Pick(QValid, QDiscarded, gen, mut)\;
+    (result, trace) $\gets$ WithTrace(P(x))\;
+    $\cdots$\;
+    \If{Interesting(TLog, trace)}{
+      \If{\KwNot Discarded(result)}{
+        batch $\gets$ Mutate(x, mut, R)\;
+        {\color{red} prio $\gets$ BranchDepth(TLog, trace)}\; %% NEW
+        {\color{red} PushFront(QValid, prio, batch)}\;          %% NEW
+      }
+      $\cdots$\;
+    }
+  }
+  %% \vspace{5pt}
+  \SetKwFunction{Pick}{Pick}
+  \SetKwProg{Fn}{Function}{:}{}
+  \Fn{\Pick{QValid, QDiscarded, gen}}{
+    \If{\KwNot Empty(QValid)}{
+      {\color{red} (batch, prio) $\gets$ DequeMin(QValid)}\; %%NEW
+      \If{Empty(batch)}{
+        Pick(QValid, QDiscarded, gen)\;
+      } \uElse{
+        {\color{red} PushFront(QValid, prio, Rest(batch))}\; %% NEW
+        \KwRet First(batch)\;
+      }
+    }
+    $\cdots$\;
+  }
+
+\caption{\label{algo:mutagen:fifo}Priority FIFO Heuristic}
+\end{algorithm}
+
+In this light, every time we insert a new execution path into the internal trace
+log (TLog), we calculate the number of \emph{new nodes} that were executed, as
+well as the \emph{branching depth} where they got inserted.
+%
+The former is used to distinguish interesting test cases (whether or not new
+nodes were inserted), whereas the latter is used by this heuristic to schedule
+mutation candidates.
+%
+Fig. \ref{fig:mutagen:tracelog} illustrates this idea, inserting two execution
+traces (one after another) into a trace log that initially contains a single
+execution path.
+%
+The second insertion (with trace $1 \rightarrow 2 \rightarrow 6 \rightarrow 7$)
+branches at a shallower depth than the first one (2 vs. 3), and so its
+corresponding test case is given a higher priority.
+
+Using this mechanism, we can modify \mutagen's base testing loop by replacing
+each mutation queue with a priority queue indexed by the branching depth of each
+new execution.
+%
+These changes are illustrated in Algorithm \ref{algo:mutagen:fifo}.
+%
+Statements in {\color{red} red} indicate important changes to the base
+algorithm, whereas ellipses denote parts of the code that are not relevant for
+the point being made.
+
+To pick the next test case, we retrieve the one with the highest priority (the
+smallest branching depth).
+%
+Then, when we find a new interesting test case, it gets enqueued at the
+\emph{beginning} of the queue at its corresponding priority.
+%
+This allows the testing loop to jump immediately onto mutating new interesting
+test cases as soon as they are found (even at the same priority), and to jump
+back to previous test cases as soon as their mutants become less novel.
+
+%% \caption{\label{fig:heuristics} The two main heuristics implemented in \mutagen.
+%%   Statements in {\color{red} red} indicate important changes to the base
+%%   algorithm. Ellipses denote parts of the code that are not relevant for the
+%%   point being made. }
+
 \subsection{Detecting Trace Space Saturation And Tuning Random Mutations Parameter}
@@
 %
-The first one is a simple abstract stack machine that enforces the
+The first one is a simple abstract stack machine that enforces
 \emph{noninterference} \cite{goguen1982security} using runtime checks.
@@
 %
-Such instructions are defined as:
+These instructions are defined as:
 %
@@
 
-\begin{table}[t]
+\begin{table}[b]
 \scriptsize
@@
 %
-In this light, our testing properties assert that any result produced by
+In this manner, our testing properties assert that any result produced by
 \textit{haskell-wasm} matches that of the reference implementation.
-%
+
 Notably, this engine had several latent bugs that were not caught by the
@@
 %
-Then, the property asserts whether their results are equivalent.%
+Finally, the property asserts whether their results are equivalent.%
 %
@@
 controlled by this heuristic, so we arbitrarily fixed it to 25 random mutations
-throughout the testing loop.
+throughout the execution of testing loop.
 
@@
 With this in mind, it is fair to think that \fuzzchick might be a better choice
-if we set a shorter timeout and run it several times.
+if we set a shorter timeout and run it several times in a row.
 %
@@
 \emph{QuickFuzz} \cite{GriecoCB16, grieco2017} is a fuzzer that exploits these
-ideas to synthesize random generators from existing Haskell libraries and that
-are used in tandem with existing low-level fuzzers to find bugs in heavily used
-programs.
+ideas to synthesize random generators from existing Haskell libraries. These
+generators are used in tandem with existing low-level fuzzers to find bugs in
+heavily used programs.
 
@@
 derived generators, specialized ones can be used directly by \mutagen, and
-combining them with type-preserving mutations is part of our future work.
+combining them with structure-preserving mutations is part of our future work.
 
@@
 
-Our experimental results indicate that \mutagen outperforms the simpler approach
-taken by \fuzzchick in terms of both failure rate and tests until first failure.
+Our experimental results indicate that \mutagen outperforms the simpler CGPT
+approach implemented in \fuzzchick in terms of both failure rate and tests until
+first failure.
 %
@@
 %
-combining \mutagen with specialized generators; improving the code
-instrumentation in order to target mutations towards specific parts of the code;
+combining \mutagen with specialized generators; enhancing the code
+instrumentation to be able target mutations towards specific parts of the code;
 and evaluating the effect of lazy pruning; among others.

### 2021-09-03 papers/icse22/main.tex
Abstract pass

@@
 This is coupled with heuristics that help to schedule mutants based on their
-novelty and to minimize mutating subexpressions of large enumeration types more
-than necessary.
+novelty and to minimize mutating subexpressions leading to scalability issues of
+exhaustiveness.
+%of large enumeration types more
+%than necessary.
 %

### 2021-09-03 papers/icse22/main.tex
Pass over intro

@@
 In addition, high-level, structure-preserving mutators are better suited for
-producing syntactically valid mutants, avoiding the time wasted by using generic
-low-level mutators that act over the ``serialized'' test cases and know very
-little about their structure, thus producing syntactically broken mutants most
-of the time.
-%
-This grammar-aware mutation technique has shown to be quite useful when fuzzing
-systems accepting structurally complex inputs \cite{holler2012fuzzing,
+producing syntactically valid mutants.
+%
+High-level mutators avoids wasting time with generic low-level mutators (e.g.,
+at the bit level) that know very little about data structures, thus producing
+syntactically broken mutants most of the time.
+%
+Such a grammar-aware mutation technique has shown to be quite useful when
+fuzzing systems accepting structurally complex inputs \cite{holler2012fuzzing,
   wang2019superion, xsmith}.
@@
 %
-The data types used by the inputs of the testing properties already describe the
-structure of the random data one needs to mutate in a concrete manner, and
-specialized mutators can be automatically derived directly from their definition
---- making strongly-typed programming languages an ideal match for CGPT.
+The data types of the inputs to the testing properties already provide a good source
+of information about how data can be mutated, e.g., by changing one data constructor
+by another.
+%
+% The data types used by the inputs of the testing properties already describe the
+% structure of the random data one needs to mutate in a concrete manner, and
+Specialized mutators can be automatically derived directly from data types
+definitions --- making strongly-typed programming languages an ideal match for
+CGPT.
 
@@
 \citeauthor{lampropoulos2019coverage} in their original work, and an existing
-{Web-Assembly} engine of industrial strenght written in Haskell.
+{Web-Assembly} engine of industrial strenght written in the strongly-typed
+programming language Haskell.
 %
@@
 In the IFC stack machine case study, we compare \mutagen against \fuzzchick, the
-reference CGPT implementation developed in Coq by
-\citeauthor{lampropoulos2019coverage}.
+reference CGPT implementation developed in Coq \cite{lampropoulos2019coverage}.
 %
@@
 All in all, our evaluation encompasses more than 600 hours of computing time and
-suggests that testing mutants exhaustively can be an appealing technique for
+suggests that \emph{testing mutants exhaustively together with our heuristics to
+  escape scalability issues} can be an appealing technique for
 finding bugs reliably without sacrificing speed.

### 2021-09-03 papers/icse22/main.tex
Pass Section 3 - some changes and simplifications

@@
 In this section, we briefly describe the motivation, ideas and limitations
-behind CGPT as depicted by \citeauthor{lampropoulos2019coverage}.
+behind CGPT \cite{lampropoulos2019coverage}.
 %
@@
 
-\noindent This property asserts that, given a balanced tree \texttt{t} and an
-element \texttt{x} as input, inserting \texttt{x} into \texttt{t} will produce a
+\noindent This property asserts that, given an element \texttt{x} and a balanced
+tree \texttt{t} as input, inserting \texttt{x} into \texttt{t} will produce a
 balanced tree as output.
@@
 %% operation:
-We can write a na\"ive one (simplified to make it more accessible%
+We can write a na\"ive one for trees of integers (simplified to make it more accessible%
 \footnote{In practice, this code will most likely be implemented using Haskell's
@@
 genTree(sz) =
-  if sz == 0 then Leaf <$> genA
-  else oneof [ Leaf <$> genA
-             , Branch <$> genTree(sz-1) <*> genA <*> genTree(sz-1) ]
+  if sz == 0 then do x <- genInt ; return (Leaf x)
+  else oneof [ do i <- genInt ; return (Leaf i)
+             , do l <- genTree(sz-1) ;
+                  x <- genInt        ;
+                  r <- genTree(sz-1) ;
+                  return (Branch l x r) ]
 \end{verbatim}
@@
 %
-Payloads are generated simply by delegating the task to an external random
-generator (\texttt{genA}) defined elsewhere.
+Integers are generated simply by delegating the task to an external random
+generator (\texttt{genInt}) defined elsewhere.
 %
@@
 %
-In Haskell, we define mutations as:
-
+We define mutations for a value of type \texttt{a} (written \texttt{Mutations
+  a}) as \emph{a function that takes a value of such a type and returns a list of
+mutants} (written \texttt{Mutant a}).
+%
+In Haskell, we introduce the following type synonym:
+%
 \begin{verbatim}
-type Mutation a = a -> [Mutant a]
+type Mutations a = a -> [Mutant a]
 \end{verbatim}
-
+%
 \noindent Where \texttt{Mutant}s come in two flavours, pure and random:
@@
 In our example, the default value of a \texttt{Tree} is a leaf containing the
-smallest value of the tree payload type.
+smallest value of the tree payload type, e.g., in the case of trees with
+integers \texttt{def} is defined as \texttt{Leaf 0}.
 %
@@
 %
-For instance, the positions of a \texttt{Tree} are computed as follows:
-
-\begin{verbatim}
-positions (Leaf x) =
-  node [ (0, positions x) ]
-positions (Branch l x r) =
-  node [ (0, positions l), (1, positions x), (2, positions r) ]
-\end{verbatim}
-
-\noindent where the function \texttt{node} simply builds a node of the positions
-Rose tree and propagates the accumulated prefix (the path from the root) to its
-children.
-%
-In this light, the mutable positions of the value \texttt{Branch (Leaf 1) 2
+% For instance, the positions of a \texttt{Tree} are computed as follows:
+%
+% \begin{verbatim}
+% positions (Leaf x) =
+%   node [ (0, positions x) ]
+% positions (Branch l x r) =
+%   node [ (0, positions l), (1, positions x), (2, positions r) ]
+% \end{verbatim}
+%
+% \noindent where the function \texttt{node} simply builds a node of the positions
+% Rose tree and propagates the accumulated prefix (the path from the root) to its
+% children.
+%
+For instance, the mutable positions of the value \texttt{Branch (Leaf 1) 2
   (Leaf 3)} are as follows:
@@
 %
-A slightly simplified version of this function for the \texttt{Tree} data type is
-as follows:
-
-\begin{verbatim}
-inside []     t              = mutate t
-inside (0:ps) (Leaf x)       = [ Leaf x'       | x' <- inside ps x ]
-inside (0:ps) (Branch l x r) = [ Branch l' x r | l' <- inside ps l ]
-inside (1:ps) (Branch l x r) = [ Branch l x' r | x' <- inside ps x ]
-inside (2:ps) (Branch l x r) = [ Branch l x r' | r' <- inside ps r ]
-\end{verbatim}
-
-\noindent This function simply traverses the desired position, calling itself
-recursively until it reaches the desired subexpression, where a \texttt{mutate}
-can be applied locally at the top-level (case \texttt{inside [] t}).
-%
-The rest of the function takes care of unwrapping and rewrapping the
+% A slightly simplified version of this function for the \texttt{Tree} data type is
+% as follows:
+%
+% \begin{verbatim}
+% inside []     t              = mutate t
+% inside (0:ps) (Leaf x)       = [ Leaf x'       | x' <- inside ps x ]
+% inside (0:ps) (Branch l x r) = [ Branch l' x r | l' <- inside ps l ]
+% inside (1:ps) (Branch l x r) = [ Branch l x' r | x' <- inside ps x ]
+% inside (2:ps) (Branch l x r) = [ Branch l x r' | r' <- inside ps r ]
+% \end{verbatim}
+%
+This function simply traverses the desired position, calling itself recursively
+until it reaches the desired subexpression, where a mutation encoded by \texttt{mutate} can be
+applied.
+% locally at the top-level (case \texttt{inside [] t}).
+%
+The rest of the \texttt{inside}'s functionality simply unwraps and rewraps the
 intermediate subexpressions and is not particularly relevant for the point being

### 2021-09-03 papers/icse22/main.tex
Pass over conclusion

@@
   if sz == 0 then do x <- genInt ; return (Leaf x)
-  else oneof [ do i <- genInt ; return (Leaf i)
+  else oneof [ do x <- genInt ; return (Leaf x)
              , do l <- genTree(sz-1) ;
@@
 
-We presented \mutagen, a coverage-guided, property-based testing framework
-written in Haskell.
+We presented \mutagen, a coverage-guided, property-based testing framework.
+%written in Haskell.
 %
@@
 %
-Our tool extends the CGPT approach introduced by
-\citeauthor{lampropoulos2019coverage} with an exhaustive mutation mechanism that
-generates every possible mutant for each interesting test case, and schedules it
-to be tested exactly once.
+Our tool extends the CGPT approach
+%introduced by
+%\citeauthor{lampropoulos2019coverage}
+with an exhaustive mutation mechanism that generates every possible mutant for
+each interesting test case, and schedules it to be tested exactly once.
 

### 2021-09-03 papers/icse22/main.tex
Section 5 pass

@@
 %
-Worth mentioning, we translated this case study into Haskell in order to run
-\mutagen on its test suite.
+Worth mentioning, we reimplemented this case study from scratch in Haskell in
+order to run \mutagen on its test suite.
 
@@
 targets \textit{haskell-wasm} \cite{haskellwasm}, an existing WebAssembly engine
-of industrial strength written in Haskell.
+of industrial strength.
 %

### 2021-09-03 papers/icse22/main.tex
Changed colors in evaluation and A12 estimators

@@
 \hline
-1  & 13974.9 & 3632.7 & 0.89 & {\color{red} large} \\ \hline
-2  & 23962.9 & 7122.3 & 0.90 & {\color{red} large} \\ \hline
-3  & 19678.5 & 4633.9 & 0.89 & {\color{red} large} \\ \hline
-4  & 20398.4 & 16831.2 & 0.72 & {\color{red} medium} \\ \hline
-5  & 17348.1 & 2326.6 & 0.94 & {\color{red} large} \\ \hline
-6  & 10727.1 & 2312.9 & 0.92 & {\color{red} large} \\ \hline
-7  & 5070.7 & 332.9 & 0.91 & {\color{red} large} \\ \hline
-8  & 5596.4 & 298.7 & 0.90 & {\color{red} large} \\ \hline
+1  & 13974.9 & 3632.7 & 0.89 & {\color{blue} large} \\ \hline
+2  & 23962.9 & 7122.3 & 0.90 & {\color{blue} large} \\ \hline
+3  & 19678.5 & 4633.9 & 0.89 & {\color{blue} large} \\ \hline
+4  & 20398.4 & 16831.2 & 0.72 & {\color{blue} medium} \\ \hline
+5  & 17348.1 & 2326.6 & 0.94 & {\color{blue} large} \\ \hline
+6  & 10727.1 & 2312.9 & 0.92 & {\color{blue} large} \\ \hline
+7  & 5070.7 & 332.9 & 0.91 & {\color{blue} large} \\ \hline
+8  & 5596.4 & 298.7 & 0.90 & {\color{blue} large} \\ \hline
 9  & 16402.4 & 26342.5 & 0.51 & negligible \\ \hline
 10 & 11553.0 & 25044.5 & 0.55 & negligible \\ \hline
-11 & 11304.3 & 2536.8 & 0.82 & {\color{red} large} \\ \hline
-12 & 18507.3 & 14482.7 & 0.65 & {\color{red} small} \\ \hline
-13 & 17961.5 & 13454.9 & 0.65 & {\color{red} small} \\ \hline
-14 & 10621.9 & 3928.3 & 0.78 & {\color{red} large} \\ \hline
-15 & 23866.3 & 43678.2 & 0.23 & {\color{blue} large} \\ \hline
+11 & 11304.3 & 2536.8 & 0.82 & {\color{blue} large} \\ \hline
+12 & 18507.3 & 14482.7 & 0.65 & {\color{blue} small} \\ \hline
+13 & 17961.5 & 13454.9 & 0.65 & {\color{blue} small} \\ \hline
+14 & 10621.9 & 3928.3 & 0.78 & {\color{blue} large} \\ \hline
+15 & 23866.3 & 43678.2 & 0.23 & {\color{red} large} \\ \hline
 16 & 25321.7  & 27602.0 & 0.54 & negligible \\ \hline
 17 & 28515.6 & 52344.9 & 0.47 & negligible \\ \hline
-18 & 24218.6 & 123401.3 & 0.14 & {\color{blue} large} \\ \hline
+18 & 24218.6 & 123401.3 & 0.14 & {\color{red} large} \\ \hline
 19 & 18638.9 & 30682.3 & 0.45 & negligible \\ \hline
-20 & 18883.1 & 15841.9 & 0.60 & {\color{red} small} \\ \hline
+20 & 18883.1 & 15841.9 & 0.60 & {\color{blue} small} \\ \hline
    & & Mean & 0.67 & {\color{red} medium} \\
@@
 %
-Estimates in {\color{blue} blue} and {\color{red} red} indicate results
+Estimates in {\color{red} red} and {\color{blue} blue} indicate results
 favoring \fuzzchick and \mutagen, respectively.

### 2021-09-03 papers/icse22/main.tex
Moved figures around, made it fit again, restored some heading space.

@@
 
-\titlespacing*{\section}{0pt}{5pt}{5pt}
-\titlespacing*{\subsection}{0pt}{5pt}{5pt}
-\titlespacing*{\subsubsection}{0pt}{5pt}{5pt}
+%% \titlespacing*{\section}{0pt}{5pt}{5pt}
+%% \titlespacing*{\subsection}{0pt}{5pt}{5pt}
+%% \titlespacing*{\subsubsection}{0pt}{5pt}{5pt}
 \titlespacing*{\paragraph}{10pt}{0pt}{10pt}
@@
 genTree(sz) =
-  if sz == 0 then do x <- genInt ; return (Leaf x)
-  else oneof [ do x <- genInt ; return (Leaf x)
-             , do l <- genTree(sz-1) ;
-                  x <- genInt        ;
-                  r <- genTree(sz-1) ;
-                  return (Branch l x r) ]
+  if sz == 0 then do { x <- genInt; return (Leaf x) }
+  else oneof [ do { x <- genInt; return (Leaf x) }
+             , do { l <- genTree(sz-1);
+                    x <- genInt;
+                    r <- genTree(sz-1);
+                    return (Branch l x r) } ]
 \end{verbatim}
@@
 
-\begin{figure}[b]
+\begin{figure}[t]
 \begin{verbatim}
@@
 
-\begin{algorithm}[t]
+\begin{algorithm}[b]
   \SetArgSty{textnormal}
@@
 
-Finally, the seed selection algorithm (Algorithm \ref{algo:mutagen:pick}) picks
-the next test case using the same criteria as CGPT, prioritizing valid test
-cases over discarded ones, falling back to random generation when both queues
-are empty.
-%
-Since mutations are precomputed, this function simply picks the next test case
-from the current batch, until it becomes empty and can switch to the next
-precomputed one in line.
-
-Having selected the next test case, the testing loop proceeds to execute it,
-capturing both the result (valid, discarded, or failed) and its execution trace
-over the system under test.
-%
-If the test case fails, it is reported as a bug.
-%
-If not, the algorithm evaluates whether it was interesting (i.e., it exercises a
-new path) based on its trace information and the one from previously executed
-test cases (represented by TLog).
-%
-If the test case discovers a new path, it is enqueued on its corresponding
-queue.
-% depending on whether it was valid or got discarded.
-%
-%% (i.e., it failed the precondition).
-%
-This process alternates between generation and mutation until it finds a bug or
-reaches the test limit N.
-
-Another difference with CGPT's testing loop is the criterion for enqueuing
-discarded tests.
-%
-We found that, especially for large data types, the queue of discarded
-candidates tends to grow disproportionately large during testing, making them
-hardly usable and consuming large amounts of memory.
-%
-To improve this, we resort to mutate discarded tests cases only when we have
-some evidence that they are ``almost valid.''
-%
-For this, each mutated test case remembers whether its parent (the original test
-case they derive from after being mutated) was valid.
-%
-Then, we enqueue discarded test cases only if they meet this condition.
-%
-As a result, we fill the discarded queue with lesser but much more interesting
-mutable test cases.
-%% %
-%% Moreover, this can potentially lead to \emph{2-step mutations}, where an initial
-%% mutation breaks a valid test case in a small way, it gets enqueued as discarded,
-%% and later a subsequent mutation fixes it by changing a different subexpression.
-
-\begin{algorithm}[t]
+\begin{algorithm}[b]
   \SetArgSty{textnormal}
@@
       batch $\gets$ Deque(QValid)\;
-      \If{Empty(batch)}{
-        Pick(QValid, QDiscarded, gen)\;
-      } \uElse{
+      \lIf{Empty(batch)}{ Pick(QValid, QDiscarded, gen) }
+      \uElse{
         PushFront(QValid, Rest(batch))\;
@@
     }
-    \If{\KwNot Empty(QValid)}{
-      batch $\gets$ Deque(QValid)\;
-      \If{Empty(batch)}{
-        Pick(QValid, QDiscarded, gen)\;
-      } \uElse{
-        PushFront(QValid, Rest(batch))\;
+    \If{\KwNot Empty(QDiscarded)}{
+      batch $\gets$ Deque(QDiscarded)\;
+      \lIf{Empty(batch)}{ Pick(QValid, QDiscarded, gen) }
+      \uElse{
+        PushFront(QDiscarded, Rest(batch))\;
         \KwRet First(batch)\;
@@
 \end{algorithm}
+
+Finally, the seed selection algorithm (Algorithm \ref{algo:mutagen:pick}) picks
+the next test case using the same criteria as CGPT, prioritizing valid test
+cases over discarded ones, falling back to random generation when both queues
+are empty.
+%
+Since mutations are precomputed, this function simply picks the next test case
+from the current batch, until it becomes empty and can switch to the next
+precomputed one in line.
+
+Having selected the next test case, the testing loop proceeds to execute it,
+capturing both the result (valid, discarded, or failed) and its execution trace
+over the system under test.
+%
+If the test case fails, it is reported as a bug.
+%
+If not, the algorithm evaluates whether it was interesting (i.e., it exercises a
+new path) based on its trace information and the one from previously executed
+test cases (represented by TLog).
+%
+If the test case discovers a new path, it is enqueued on its corresponding
+queue.
+% depending on whether it was valid or got discarded.
+%
+%% (i.e., it failed the precondition).
+%
+This process alternates between generation and mutation until it finds a bug or
+reaches the test limit N.
+
+Another difference with CGPT's testing loop is the criterion for enqueuing
+discarded tests.
+%
+We found that, especially for large data types, the queue of discarded
+candidates tends to grow disproportionately large during testing, making them
+hardly usable and consuming large amounts of memory.
+%
+To improve this, we resort to mutate discarded tests cases only when we have
+some evidence that they are ``almost valid.''
+%
+For this, each mutated test case remembers whether its parent (the original test
+case they derive from after being mutated) was valid.
+%
+Then, we enqueue discarded test cases only if they meet this condition.
+%
+As a result, we fill the discarded queue with lesser but much more interesting
+mutable test cases.
+%% %
+%% Moreover, this can potentially lead to \emph{2-step mutations}, where an initial
+%% mutation breaks a valid test case in a small way, it gets enqueued as discarded,
+%% and later a subsequent mutation fixes it by changing a different subexpression.
 
@@
 \end{figure}
-
 \begin{algorithm}[t]
+  \SetKwComment{Comment}{/* }{}
   \SetArgSty{textnormal}
@@
       {\color{red} (batch, prio) $\gets$ DequeMin(QValid)}\; %%NEW
-      \If{Empty(batch)}{
-        Pick(QValid, QDiscarded, gen)\;
-      } \uElse{
+      \lIf{Empty(batch)}{ Pick(QValid, QDiscarded, gen) }
+      \uElse{
         {\color{red} PushFront(QValid, prio, Rest(batch))}\; %% NEW
@@
     }
+    \If{\KwNot Empty(QDiscarded)}{\Comment {Analogous to the case above */} }
     $\cdots$\;

### 2021-09-08 papers/icst22/rebuttal.md
Created folder for ICST submission

--- 2021-06-15 papers/oopsla-21/rebuttal.md
@@
 FuzzChick and MUTAGEN, where our tool shows a noticeable improvement in terms of
-mean-time-to-failure, failure rate. We will update the introduction to reflect
+mean-time-to-failure and failure rate. We will update the introduction to reflect
 that explicitly.
@@
 without being (possibly orders of magnitude) slower on individual runs. Table 3
-shows that this in the case, where the *A12* measure indicates that MUTAGEN is
+shows that this is the case, where the *A12* measure indicates that MUTAGEN is
 not prone to need many more tests to find a bug that FuzzChick in most of the

### 2021-09-22 papers/icst22/main.tex
Created folder for ICST submission
Fine pass over paper

--- 2021-09-03 papers/icse22/main.tex
@@
 %% These commands are for a JOURNAL article.
-\acmConference[ICSE 2022]{The 44th International Conference on Software
-  Engineering}{May 21–29, 2022}{Pittsburgh, PA, USA}
+\acmConference[ICST 2022]{15th IEEE International Conference on Software
+  Testing, Verification and Validation}{Apr 4–13, 2022}{}
 
@@
 
-\author{Agust\'in Mista}
-\email{mista@chalmers.se}
-\affiliation{%
-  \institution{Chalmers University of Technology}
-  \city{Gothenburg}
-  \country{Sweden}
-}
-
-\author{Alejandro Russo}
-\email{russo@chalmers.se}
-\affiliation{%
-  \institution{Chalmers University of Technology}
-  \city{Gothenburg}
-  \country{Sweden}
-}
+%% Hidden authors in the generated PDF.
+
+%% \author{Agust\'in Mista}
+%% \email{mista@chalmers.se}
+%% \affiliation{%
+%%   \institution{Chalmers University of Technology}
+%%   \city{Gothenburg}
+%%   \country{Sweden}
+%% }
+
+%% \author{Alejandro Russo}
+%% \email{russo@chalmers.se}
+%% \affiliation{%
+%%   \institution{Chalmers University of Technology}
+%%   \city{Gothenburg}
+%%   \country{Sweden}
+%% }
 
@@
 This is coupled with heuristics that help to schedule mutants based on their
-novelty and to minimize mutating subexpressions leading to scalability issues of
+novelty and to minimize mutating certain subexpressions leading to scalability issues of
 exhaustiveness.
@@
 %
-High-level mutators avoids wasting time with generic low-level mutators (e.g.,
-at the bit level) that know very little about data structures, thus producing
+High-level mutators avoid wasting time with generic low-level mutators that work
+at the bit level and know very little about data structures, thus producing
 syntactically broken mutants most of the time.
@@
 
-Worth mentioning, this technique is not meant to replace sophisticated, manually
-written random generators, but rather to provide an adequate testing solution
+Worth mentioning, this technique is not meant to replace sophisticated manually-written
+random generators, but rather to provide an adequate testing solution
 for early stages of development.
@@
 %
-Moreover, our scheduling lets the testing loop jump back and forth between
+Moreover, our scheduling allows the testing loop to jump back and forth between
 mutable test cases as soon as a new more interesting one is enqueued for
@@
 
-In contrast to CGPT and many other popular fuzzing tools, \mutagen works by to
+In contrast to CGPT and many other popular fuzzing tools, \mutagen works by 
 mutating test cases on an exhaustive and precise manner, where
@@
 The first one is a simple abstract stack machine that enforces
-\emph{noninterference} \cite{goguen1982security} using runtime checks.
+the hyperproperty \emph{noninterference} \cite{goguen1982security} using runtime checks.
 %
@@
 \noindent Where $l_{pc}$, $l_{p}$, $l_{v}$, and $l_{v'}$ represent the labels
-of: the program counter, the memory pointer, and the old a new values stored in
+of: the program counter, the memory pointer, and the old and new values stored in
 that memory cell.
@@
 %
-In our case studies, however, we observed that properties we tested
+In our case studies, however, we observed that the properties we tested
 %% their preconditions tend to be

### 2021-09-27 papers/icst22/main.tex
Adressed Ale's comments

@@
 reliance on randomness and poor mutants scheduling, reducing its applicability
-and bug-finding capacity.
+and bug-finding capacity in existing software.
 
@@
 
-Worth mentioning, this technique is not meant to replace sophisticated manually-written
-random generators, but rather to provide an adequate testing solution
-for early stages of development.
+Worth mentioning, CGPT is not meant to replace sophisticated manually-written
+random generators, but rather to provide an adequate testing solution for early
+stages of development.
 %, and while a better test suite using manually-written generators is
@@
 %% Heuristics to improve testing loop
-Additionally, the testing loop of \mutagen incorporates two heuristics that help
-finding bugs faster and more reliably (\S \ref{sec:heuristics}).
+Additionally, the testing loop of \mutagen incorporates two novel heuristics
+that help finding bugs faster and more reliably (\S \ref{sec:heuristics}).
 %
@@
 %
-Then, if we assume existence of a predicate \texttt{balanced} of type
+Then, if we assume existence of a predicate (function) \texttt{balanced} of type
 \texttt{Tree a -> Bool} that asserts that a tree satisfies a certain notion of
@@
 \begin{verbatim}
-genTree(sz) =
-  if sz == 0 then do { x <- genInt; return (Leaf x) }
+genTree(size) =
+  if size == 0 then do { x <- genInt; return (Leaf x) }
   else oneof [ do { x <- genInt; return (Leaf x) }
-             , do { l <- genTree(sz-1);
+             , do { l <- genTree(size-1);
                     x <- genInt;
-                    r <- genTree(sz-1);
+                    r <- genTree(size-1);
                     return (Branch l x r) } ]
@@
 subexpression needed to produce a well-typed value, carefully reducing the input
-size limit \texttt{sz} by a unit on each step, ensuring termination by
-restricting itself to generating leaves when it reaches zero (case \texttt{sz ==
-  0}).
+size limit \texttt{size} by a unit on each step, ensuring termination by
+restricting itself to generating leaves when it reaches zero (case \texttt{size
+  == 0}).
 %
@@
 
-To keep the comparison fair, the authors replicated the same power schedule
-configuration used in AFL, which is tailored to work in tandem with generic
-bit-level mutators.
+To keep the comparison fair, \citeauthor{lampropoulos2019coverage} replicated
+the same power schedule configuration used in AFL, which is tailored to work in
+tandem with generic bit-level mutators.
 %
@@
 The next section introduces \mutagen, our revised CGPT that aims to tackle the
-main limitations of CGPT.
+main limitations of CGPT to make it practical in real-world scenarios.
 
@@
 
-In first place, a function \texttt{positions} traverses the input and builds a
+In first place, a function \texttt{Positions} traverses the input and builds a
 Rose tree of \emph{mutable positions}, where positions are lists of indices
@@
   \footnotesize
-  \texttt{positions}\quad
+  \texttt{Positions}\quad
   \left(
@@
 
-Later, we define a function \texttt{inside} that takes a desired
-position within an input test case and mutates its corresponding subexpression.
+Later, we define a function \texttt{MutateInside} that takes a desired position
+within an input test case and mutates its corresponding subexpression, returning
+a list of mutants.
 %
@@
 %
-The rest of the \texttt{inside}'s functionality simply unwraps and rewraps the
+The rest of the \texttt{MutateInside}'s functionality simply unwraps and rewraps the
 intermediate subexpressions and is not particularly relevant for the point being
@@
   \SetKwProg{Fn}{Function}{:}{}
-  \Fn{\MG{P, N, R, gen, mut}}{
+  \Fn{\MG{P, N, R, gen}}{
     i $\gets$ 0\;
@@
         \If{\KwNot Discarded(result)}{
-          batch $\gets$ Mutate(x, mut, R)\;
+          batch $\gets$ CreateMutationBatch(x, R)\;
           Enqueue(QValid, batch)\;
@@
         \uElseIf{\KwNot Discarded(Parent(x))}{
-          batch $\gets$ Mutate(x, mut, R)\;
+          batch $\gets$ CreateMutationBatch(x, R)\;
           Enqueue(QDiscarded, batch)\;
@@
   \DontPrintSemicolon
-  \SetKwFunction{Muts}{Mutate}
+  \SetKwFunction{Muts}{CreateMutationBatch}
   \SetKwProg{Fn}{Function}{:}{}
-  \Fn{\Muts{x, mut, R}}{
+  \Fn{\Muts{x, R}}{
     batch $\gets$ $\varnothing$\;
     \For{pos \KwIn Flatten(Positions(x))}{
-      \For{mutant \KwIn Inside(pos, mut, x)}{
+      \For{mutant \KwIn MutateInside(pos, x)}{
         \Switch{mutant}{
@@
   }
-\caption{\label{algo:mutagen:init}Mutants Initialization}
+\caption{\label{algo:mutagen:init}Mutation Batch Initialization}
 \end{algorithm}
@@
   \SetKwProg{Fn}{Function}{:}{}
-  \Fn{\MG{P, N, R, gen, mut}}{
+  \Fn{\MG{P, N, R, gen}}{
     $\cdots$\;
-    x $\gets$ Pick(QValid, QDiscarded, gen, mut)\;
+    x $\gets$ Pick(QValid, QDiscarded, gen)\;
     (result, trace) $\gets$ WithTrace(P(x))\;
@@
       \If{\KwNot Discarded(result)}{
-        batch $\gets$ Mutate(x, mut, R)\;
+        batch $\gets$ Mutate(x, R)\;
         {\color{red} prio $\gets$ BranchDepth(TLog, trace)}\; %% NEW
@@
   \SetKwProg{Fn}{Function}{:}{}
-  \Fn{\MG{P, N, gen, mut}}{
+  \Fn{\MG{P, N, gen}}{
     {\color{red} boring $\gets$ 0}; {\color{red} reset $\gets$ 1000}; {\color{red}R $\gets$ 1}\;
@@
 20 & 18883.1 & 15841.9 & 0.60 & {\color{blue} small} \\ \hline
-   & & Mean & 0.67 & {\color{red} medium} \\
+   & & Mean & 0.67 & {\color{blue} medium} \\
 \hline

### 2021-10-22 papers/icst22/main.tex
Added missing change
Added link to replication package
Replaced template and copied old version

@@
 %% The first command in your LaTeX source must be the \documentclass command.
-\documentclass[sigconf, anonymous, review]{acmart}
+\documentclass[conference, anonymous]{IEEEtran}
 
@@
 \usepackage{titlesec}
+\usepackage{amsmath}
 
@@
 
-\settopmatter{printacmref=false}
+%% \settopmatter{printacmref=false}
 
@@
 
-\linepenalty=1000
-\widowpenalty=10000
-\clubpenalty=10000
+%% \linepenalty=1000
+%% \widowpenalty=10000
+%% \clubpenalty=10000
 
@@
 \SetAlCapNameFnt{\small}
-
-% ----------------------------------------
-
-%% Rights management information. This information is sent to you when you
-%% complete the rights form. These commands have SAMPLE values in them; it is
-%% your responsibility as an author to replace the commands and values with
-%% those provided to you when you complete the rights form.
-\setcopyright{acmcopyright}
-%% \copyrightyear{2021}
-%% \acmYear{2021}
-%% \acmDOI{10.1145/1122445.1122456}
-
-%% These commands are for a JOURNAL article.
-\acmConference[ICST 2022]{15th IEEE International Conference on Software
-  Testing, Verification and Validation}{Apr 4–13, 2022}{}
-
-%% \acmJournal{JACM}
-%% \acmVolume{37}
-%% \acmNumber{4}
-%% \acmArticle{111}
-%% \acmMonth{8}
-
-%% Submission ID.
-%% Use this when submitting an article to a sponsored event. You'll
-%% receive a unique submission ID from the organizers
-%% of the event, and this ID should be used as the parameter to this command.
-%%\acmSubmissionID{123-A56-BU3}
-
-%% The majority of ACM publications use numbered citations and
-%% references.  The command \citestyle{authoryear} switches to the
-%% "author year" style.
-%%
-%% If you are preparing content for an event
-%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
-%% citations and references.
-%% Uncommenting
-%% the next command will enable that style.
-%% \citestyle{authoryear}
 
@@
 
-%%
-%% The "title" command has an optional parameter,
-%% allowing the author to define a "short title" to be used in page headers.
-\title[\mutagen]{\mutagen: Coverage-Guided, Property-Based Testing using
-  Exhaustive Structure-Preserving Mutations}
-
-%%
-%% The "author" command and its associated commands are used to define
-%% the authors and their affiliations.
-%% Of note is the shared affiliation of the first two authors, and the
-%% "authornote" and "authornotemark" commands
-%% used to denote shared contribution to the research.
-
-%% Hidden authors in the generated PDF.
-
-%% \author{Agust\'in Mista}
-%% \email{mista@chalmers.se}
-%% \affiliation{%
-%%   \institution{Chalmers University of Technology}
-%%   \city{Gothenburg}
-%%   \country{Sweden}
-%% }
-
-%% \author{Alejandro Russo}
-%% \email{russo@chalmers.se}
-%% \affiliation{%
-%%   \institution{Chalmers University of Technology}
-%%   \city{Gothenburg}
-%%   \country{Sweden}
-%% }
-
-%% By default, the full list of authors will be used in the page headers. Often,
-%% this list is too long, and will overlap other information printed in the page
-%% headers. This command allows the author to define a more concise list of
-%% authors' names for this purpose.
-
-%% \renewcommand{\shortauthors}{Mista and Russo}
+\title{
+  \mutagen: Coverage-Guided, Property-Based Testing using
+  Exhaustive Structure-Preserving Mutations
+}
+
+\author{\IEEEauthorblockN{1\textsuperscript{st} Agust\'in Mista}
+  \IEEEauthorblockA{
+    \textit{dept. name of organization (of Aff.)} \\
+    \textit{Chalmers University of Technology}\\
+    Gothenburg, Sweden \\
+    mista@chalmers.se}
+\and
+\IEEEauthorblockN{2\textsuperscript{nd} Alejandro Russo}
+\IEEEauthorblockA{
+  \textit{dept. name of organization (of Aff.)} \\
+  \textit{Chalmers University of Technology}\\
+  Gothenburg, Sweden \\
+  russo@chalmers.se}
+}
 
 % ----------------------------------------
-%% The abstract is a short summary of the work to be presented in the article.
+
+%% This command processes the author and affiliation and title information and
+%% builds the first part of the formatted document.
+\maketitle
+
+\begin{IEEEkeywords}
+  random testing, mutations, heuristics
+\end{IEEEkeywords}
 
@@
 
-%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm. Please
-%% copy and paste the code instead of the example below.
-\begin{CCSXML}
-% CCS XML here
-\end{CCSXML}
-
-%% \ccsdesc[500]{Computer systems organization~Embedded systems}
-%% \ccsdesc[300]{Computer systems organization~Redundancy}
-%% \ccsdesc{Computer systems organization~Robotics}
-%% \ccsdesc[100]{Networks~Network reliability}
-
-% ----------------------------------------
-
-%% Keywords. The author(s) should pick words that accurately describe the work
-%% being presented. Separate the keywords with commas.
-\keywords{random testing, mutations, heuristics}
-
-% ----------------------------------------
-
-%% This command processes the author and affiliation and title information and
-%% builds the first part of the formatted document.
-\maketitle
-
-% ----------------------------------------
-
 %% \vspace{-5pt}
@@
 %
-Randomly generating valid programs is a common example of this problem, where
-developers are forced to put substantial efforts in writing specialized random
-generators by hand \cite{Palka11, perenyi2020stack, yang2011finding}.
-%
-Moreover, automatically synthesized random generators offer particularly poor
-results when the testing properties are constrained by sparse preconditions.
+Randomly generating valid programs to test compilers is a common example of this
+problem \cite{chen2020asurvey}, where developers are forced to put substantial
+efforts in writing specialized random generators by hand \cite{Palka11,
+  perenyi2020stack, yang2011finding}.
+%
+Moreover, automatically synthesized random generators offer poor results when
+the testing properties are constrained by sparse preconditions.
 %
@@
 %% %
-%% Despite this limitation, \citeauthor{lampropoulos2019coverage} show that
+%% Despite this limitation, \cite{lampropoulos2019coverage} show that
 %% \fuzzchick is far better than using normal random testing coupled with na\"ive
@@
 \footnote{Some of these were already identified by
-  \citeauthor{lampropoulos2019coverage} in their original work.}
-%
-of the original CGPT approach by \citeauthor{lampropoulos2019coverage} that
+  \cite{lampropoulos2019coverage} in their original work.}
+%
+of the original CGPT approach by \cite{lampropoulos2019coverage} that
 leave considerable room for improvement (see \S \ref{sec:background}).
@@
 %% %
-%% Despite this, \citeauthor{lampropoulos2019coverage} show that \fuzzchick is far
+%% Despite this, \cite{lampropoulos2019coverage} show that \fuzzchick is far
 %% better than using normal random testing coupled with na\"ive random generators
@@
 \ref{sec:casestudies}): the Information-Flow Control (IFC) stack machine used by
-\citeauthor{lampropoulos2019coverage} in their original work, and an existing
+\cite{lampropoulos2019coverage} in their original work, and an existing
 {Web-Assembly} engine of industrial strenght written in the strongly-typed
@@
 %
-%% \citeauthor{lampropoulos2019coverage} introduced \emph{coverage-guided,
+%% \cite{lampropoulos2019coverage} introduced \emph{coverage-guided,
 %%   property-based testing} (CGPT), a technique that
@@
 Generic fuzzing tools like \emph{AFL} \cite{afl}, \emph{HonggFuzz}
-\cite{honggfuzz}, or \emph{libFuzzer} \citeyearpar{libfuzzer} as well as
+\cite{honggfuzz}, or \emph{libFuzzer} \cite{libfuzzer} as well as
 language-specific ones like Crowbar \cite{dolan2017testing} or \emph{Kelinci}
@@
 
-\citeauthor{lampropoulos2019coverage} demonstrated empirically that CGPT lies
+\cite{lampropoulos2019coverage} demonstrated empirically that CGPT lies
 comfortably in the middle ground between using pure random testing with na\"ive
@@
 %% after 30 runs (as opposed to the 5 runs used original by
-%% \citeauthor{lampropoulos2019coverage}), \emph{\fuzzchick was only able to find 5
+%% \cite{lampropoulos2019coverage}), \emph{\fuzzchick was only able to find 5
 %%   out of the 20 injected bugs with a failure rate of 1}, the hardest one being
@@
 for the sake of simplicity, the mutators proposed by
-\citeauthor{lampropoulos2019coverage} are derived to follow a top-down approach.
+\cite{lampropoulos2019coverage} are derived to follow a top-down approach.
 %% similar to that used by the automatically derived random generators.
@@
 
-To keep the comparison fair, \citeauthor{lampropoulos2019coverage} replicated
+To keep the comparison fair, \cite{lampropoulos2019coverage} replicated
 the same power schedule configuration used in AFL, which is tailored to work in
@@
     i $\gets$ 0\;
-    TLog, QValid, QDiscarded $\gets$ $\varnothing$\;
+    TLog, QValid, QDiscarded $\gets$ ${}$\;
     \While{i $<$ N}{
@@
   \Fn{\Muts{x, R}}{
-    batch $\gets$ $\varnothing$\;
+    batch $\gets$ ${}$\;
     \For{pos \KwIn Flatten(Positions(x))}{
@@
       \If{\KwNot Discarded(result)}{
-        batch $\gets$ Mutate(x, R)\;
+        batch $\gets$ CreateMutationBatch(x, R)\;
         {\color{red} prio $\gets$ BranchDepth(TLog, trace)}\; %% NEW
@@
       \If{boring $>$ reset}{
-        TLog   $\gets$ $\varnothing$\;
+        TLog   $\gets$ ${}$\;
         reset $\gets$ reset * 2\;
@@
 
-We evaluated the performance of \mutagen using two case studies.
+We evaluated the performance of \mutagen using two case studies.%
+%
+\footnote{An anonymized replication package \cite{anon_2021} is available for reviewing purposes.}
 %
@@
 %
-Its implementation was proven correct by \citeauthor{10.1145/2578855.2535839}
-\citeyearpar{10.1145/2578855.2535839} in Coq, and subsequently degraded by
+Its implementation was proven correct by \cite{10.1145/2578855.2535839}
+\cite{10.1145/2578855.2535839} in Coq, and subsequently degraded by
 systematically introducing 20 bugs on its enforcing mechanism.
 %
-\citeauthor{lampropoulos2019coverage} used this case study to compare \fuzzchick
+\cite{lampropoulos2019coverage} used this case study to compare \fuzzchick
 against RPBT using na\"ive- and hand-tuned random generators.
@@
 To preserve space, we encourage the reader to refer to the work of
-\citeauthor{hritcu2013testing} \citeyearpar{hritcu2013testing,
-  hrictcu2016testing} and \citeauthor{lampropoulos2019coverage} for more details
+\cite{hritcu2013testing} \cite{hritcu2013testing,
+  hrictcu2016testing} and \cite{lampropoulos2019coverage} for more details
 about the implementation and semantics of this case study.
@@
 %
-\citeauthor{lampropoulos2019coverage} demonstrated that generating two
+\cite{lampropoulos2019coverage} demonstrated that generating two
 independent machine states using \quickcheck has virtually no chance of
@@
 From there, we followed the approach taken by
-\citeauthor{lampropoulos2019coverage} and collected the mean-time-to-failure
+\cite{lampropoulos2019coverage} and collected the mean-time-to-failure
 (MTTF) of each bug, i.e., how quickly a bug can be found in wall clock time.
@@
 While these results are somewhat pessimistic towards \fuzzchick if compared to
-the ones presented originally by \citeauthor{lampropoulos2019coverage}, we
+the ones presented originally by \cite{lampropoulos2019coverage}, we
 remark that our experiments encompass 30 independent runs instead of the 5 used
@@
 %
-Additionally, as advised by \citeauthor{arcuri2014hitchhiker}
+Additionally, as advised by \cite{arcuri2014hitchhiker}
 \cite{arcuri2014hitchhiker}, we computed the non-parametric Vargha-Delaney
@@
 %
-\citeauthor{ClaessenDP14} \citeyearpar{ClaessenDP14} developed an algorithm for
+\cite{ClaessenDP14} \cite{ClaessenDP14} developed an algorithm for
 generating inputs constrained by boolean preconditions with almost-uniform
@@
 %
-Later, \citeauthor{LampropoulosGHH17} \citeyearpar{LampropoulosGHH17} extended
+Later, \cite{LampropoulosGHH17} \cite{LampropoulosGHH17} extended
 this approach by adding a limited form of constraint solving controllable by the
@@
 %
-Recently, \citeauthor{Lampropoulos2017} \citeyearpar{Lampropoulos2017} proposed
+Recently, \cite{Lampropoulos2017} \cite{Lampropoulos2017} proposed
 a mechanism to obtain constrained random generators by defining their target
@@
 %introduced by
-%\citeauthor{lampropoulos2019coverage}
+%\cite{lampropoulos2019coverage}
 with an exhaustive mutation mechanism that generates every possible mutant for
@@
 
-\balance
+%% \balance
 
@@
 %% bibliography file.
-\bibliographystyle{ACM-Reference-Format}
+\bibliographystyle{IEEEtran}
 %% \bibliographystyle{plainnat}

### 2021-10-22 papers/sigconf-10pages/main.tex
Replaced template and copied old version

--- 2021-09-03 papers/icse22/main.tex
@@
 
-Worth mentioning, this technique is not meant to replace sophisticated, manually
-written random generators, but rather to provide an adequate testing solution
-for early stages of development.
+Worth mentioning, CGPT is not meant to replace sophisticated manually-written
+random generators, but rather to provide an adequate testing solution for early
+stages of development.
 %, and while a better test suite using manually-written generators is
@@
 
-In contrast to CGPT and many other popular fuzzing tools, \mutagen works by to
+In contrast to CGPT and many other popular fuzzing tools, \mutagen works by
 mutating test cases on an exhaustive and precise manner, where
@@
       \If{\KwNot Discarded(result)}{
-        batch $\gets$ Mutate(x, mut, R)\;
+        batch $\gets$ CreateMutationBatch(x, R)\;
         {\color{red} prio $\gets$ BranchDepth(TLog, trace)}\; %% NEW
@@
 
-We evaluated the performance of \mutagen using two case studies.
+We evaluated the performance of \mutagen using two case studies.%
+%
+\footnote{An anonymized replication package \cite{anon_2021} is available for reviewing purposes.}
 %
 The first one is a simple abstract stack machine that enforces
-\emph{noninterference} \cite{goguen1982security} using runtime checks.
+the hyperproperty \emph{noninterference} \cite{goguen1982security} using runtime checks.
 %

### 2021-10-22 papers/sigconf-10pages/results/wasm/prop_interpreter/mutagen-base/prop_interpreter_i32.BUG_INTERPRETER_1.MUTAGEN_NO_INHERITANCE.csv.10.hs

@@ -1,0 +1,1 @@
+(1,0.39596924,Function {funcType = 0, localTypes = [F32,F32,F64], body = [Block {resultType = [], body = [Block {resultType = [], body = [F32Const 0.0,SetLocal 3,Block {resultType = [F64,I64,I64], body = [F64Const 9.687955102806763e-2,F64Const 1.025179170053963,Br 2]},Br 0]}]},I64Const 3,I64Eqz,Return]})

### 2021-10-22 papers/sigconf-10pages/results/wasm/prop_interpreter/mutagen-base/prop_interpreter_i32.BUG_INTERPRETER_1.MUTAGEN_NO_INHERITANCE.csv.5.hs

@@ -1,0 +1,1 @@
+(2,0.9365159,Function {funcType = 0, localTypes = [F64], body = [Loop {resultType = [], body = [GetLocal 0,CurrentMemory,IBinOp BS32 ISub,Return]},I32Const 0,Return]})

### 2021-10-22 papers/sigconf-10pages/results/wasm/prop_interpreter/mutagen-base/prop_interpreter_i32.BUG_INTERPRETER_3.MUTAGEN_NO_INHERITANCE.csv.9.hs

@@ -1,0 +1,1 @@
+(2,0.82717097,Function {funcType = 0, localTypes = [F32], body = [I32Const 0,GetLocal 0,IBinOp BS32 IShrU]})

### 2021-10-22 papers/sigconf-10pages/results/wasm/prop_interpreter/mutagen-no-fifo/prop_interpreter_i32.BUG_INTERPRETER_2.MUTAGEN_NO_FIFO.csv.5.hs

@@ -1,0 +1,1 @@
+(2,2.0166109,Function {funcType = 0, localTypes = [I64,F64], body = [CurrentMemory,CurrentMemory,IRelOp BS32 ILtU]})

### 2021-10-22 papers/sigconf-10pages/results/wasm/prop_interpreter/mutagen-no-fifo/prop_interpreter_i32.BUG_INTERPRETER_2.MUTAGEN_NO_FIFO.csv.8.hs

@@ -1,0 +1,1 @@
+(1,0.26555988,Function {funcType = 0, localTypes = [], body = [CurrentMemory,CurrentMemory,IRelOp BS32 ILtU]})

### 2021-10-22 papers/sigconf-10pages/results/wasm/prop_interpreter/mutagen-no-fifo/prop_interpreter_i32.BUG_INTERPRETER_5.MUTAGEN_NO_FIFO.csv.4.hs

@@ -1,0 +1,1 @@
+(1,0.64323777,Function {funcType = 0, localTypes = [F64,F64], body = [CurrentMemory,CurrentMemory,CurrentMemory,I32Store (MemArg {offset = 0, align = 0}),I32Load8S (MemArg {offset = 0, align = 0})]})

### 2021-10-22 papers/sigconf-10pages/results/wasm/prop_interpreter/mutagen-no-reset/prop_interpreter_i32.BUG_INTERPRETER_1.MUTAGEN_NO_RESET.csv.9.hs

@@ -1,0 +1,1 @@
+(2,0.7472929,Function {funcType = 0, localTypes = [], body = [GetLocal 0,GetLocal 0,IBinOp BS32 ISub]})

### 2021-10-22 papers/sigconf-10pages/results/wasm/prop_interpreter/mutagen-no-reset/prop_interpreter_i32.BUG_INTERPRETER_2.MUTAGEN_NO_RESET.csv.2.hs

@@ -1,0 +1,1 @@
+(2,-0.81162834,Function {funcType = 0, localTypes = [], body = [GetLocal 0,IUnOp BS32 ICtz,FConvertIS BS64 BS32,Drop,CurrentMemory,FReinterpretI BS32,I32Const 1,I64Load (MemArg {offset = 0, align = 2}),GetLocal 0,I64ExtendSI32,CurrentMemory,Block {resultType = [I32,I64], body = [GetLocal 0,Br 1]},I32Const 0,Return]})

### 2021-10-22 papers/sigconf-10pages/results/wasm/prop_interpreter/quickcheck/prop_interpreter_i32.BUG_INTERPRETER_3.QuickCheck.csv.9.hs

@@ -1,0 +1,1 @@
+["(0,0.0,Function {funcType = 0, localTypes = [I32,I32,I32,F32,F32,F64,F64,F32], body = [I32Const 4,CurrentMemory,IBinOp BS32 IShrU]})"]

### 2021-10-22 papers/sigconf-10pages/results/wasm/prop_interpreter/quickcheck/prop_interpreter_i32.BUG_INTERPRETER_4.QuickCheck.csv.1.hs

@@ -1,0 +1,1 @@
+["(0,0.0,Function {funcType = 0, localTypes = [F64,F32,I64,F64,F64,I32,I32,I32,I32], body = [GetLocal 9]})"]

### 2021-10-22 papers/sigconf-10pages/results/wasm/prop_validator/mutagen-base/prop_validator.BUG_VALIDATOR_10.MUTAGEN_NO_INHERITANCE.csv.7.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [I32Const 0,I32Load (MemArg {offset = 0, align = 0}),BrTable [] 0]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2021-10-22 papers/sigconf-10pages/results/wasm/prop_validator/mutagen-base/prop_validator.BUG_VALIDATOR_4.MUTAGEN_NO_INHERITANCE.csv.27.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [Br 0]}], tables = [], mems = [Memory (Limit 9 (Just 9)),Memory (Limit 5 Nothing)], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2021-10-22 papers/sigconf-10pages/results/wasm/prop_validator/mutagen-base/prop_validator.BUG_VALIDATOR_7.MUTAGEN_NO_INHERITANCE.csv.10.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [I32Const 0,IBinOp BS32 IAdd,Return]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2021-10-22 papers/sigconf-10pages/results/wasm/prop_validator/mutagen-base/prop_validator.BUG_VALIDATOR_8.MUTAGEN_NO_INHERITANCE.csv.17.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [CurrentMemory,Return]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2021-10-22 papers/sigconf-10pages/results/wasm/prop_validator/mutagen-no-fifo/prop_validator.MUTAGEN_NO_FIFO.BUG_VALIDATOR_2.csv.11.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [Nop,I32Const 2,I64Const 67,I32WrapI64,Nop]}], tables = [Table (TableType (Limit 0 Nothing) AnyFunc)], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2021-10-22 papers/sigconf-10pages/results/wasm/prop_validator/mutagen-no-fifo/prop_validator.MUTAGEN_NO_FIFO.BUG_VALIDATOR_2.csv.30.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [I32Const 0,Nop]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2021-10-22 papers/sigconf-10pages/results/wasm/prop_validator/mutagen-no-fifo/prop_validator.MUTAGEN_NO_FIFO.BUG_VALIDATOR_4.csv.9.hs

@@ -1,0 +1,1 @@
+Module {types = [], functions = [], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [Import {sourceModule = "\50121\33108W\274\SUB~\DC2", name = "r", desc = ImportMemory (Limit 5 Nothing)},Import {sourceModule = "?\f\1112898\EOT\98693\1070715", name = "\135884\SUBO\DLE8Y", desc = ImportMemory (Limit 4 Nothing)},Import {sourceModule = "\ENQ\1051971", name = "I(\v8", desc = ImportMemory (Limit 0 (Just 2))},Import {sourceModule = "(z", name = "x", desc = ImportMemory (Limit 3 (Just 3))}], exports = []}

### 2021-10-22 papers/sigconf-10pages/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_1.MUTAGEN_NO_RESET.csv.19.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [I32Const 0,If {resultType = [], true = [], false = [I64Const 30]}]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2021-10-22 papers/sigconf-10pages/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_10.MUTAGEN_NO_RESET.csv.10.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [I32Const 0,Nop,I32Load (MemArg {offset = 2, align = 3}),If {resultType = [], true = [], false = []}]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2021-10-22 papers/sigconf-10pages/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_10.MUTAGEN_NO_RESET.csv.29.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [I64Const 0,Call 0,I32WrapI64,I32Load (MemArg {offset = 0, align = 0}),F64Const (-6.025231629430419),Return]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2021-10-22 papers/sigconf-10pages/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_2.MUTAGEN_NO_RESET.csv.24.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [Nop,I32Const 2,F32Const 1.2361268,Nop]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2021-10-22 papers/sigconf-10pages/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_4.MUTAGEN_NO_RESET.csv.19.hs

@@ -1,0 +1,1 @@
+Module {types = [], functions = [], tables = [], mems = [Memory (Limit 1 (Just 7)),Memory (Limit 4 (Just 4)),Memory (Limit 5 (Just 5)),Memory (Limit 3 (Just 6)),Memory (Limit 10 Nothing)], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2021-10-22 papers/sigconf-10pages/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_7.MUTAGEN_NO_RESET.csv.16.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [I32Const 0,IBinOp BS32 IShrU,Loop {resultType = [], body = []},BrTable [] 0]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-01-13 papers/ast22/main.tex
Added AST22 version

--- 2021-10-22 papers/sigconf-10pages/main.tex
@@
 %% The first command in your LaTeX source must be the \documentclass command.
-\documentclass[sigconf, anonymous, review]{acmart}
+\documentclass[sigconf, review]{acmart}
 
@@
 
-%% \titlespacing*{\section}{0pt}{5pt}{5pt}
-%% \titlespacing*{\subsection}{0pt}{5pt}{5pt}
-%% \titlespacing*{\subsubsection}{0pt}{5pt}{5pt}
+\titlespacing*{\section}{0pt}{5pt}{2pt}
+\titlespacing*{\subsection}{0pt}{5pt}{2pt}
+\titlespacing*{\subsubsection}{0pt}{5pt}{2pt}
+
 \titlespacing*{\paragraph}{10pt}{0pt}{10pt}
@@
 %% These commands are for a JOURNAL article.
-\acmConference[ICST 2022]{15th IEEE International Conference on Software
-  Testing, Verification and Validation}{Apr 4–13, 2022}{}
+\acmConference[AST 2022]{3rd ACM/IEEE International Conference on Automation of Software Test}{May 21-22, 2022}{}
 
@@
 
-%% \author{Agust\'in Mista}
-%% \email{mista@chalmers.se}
-%% \affiliation{%
-%%   \institution{Chalmers University of Technology}
-%%   \city{Gothenburg}
-%%   \country{Sweden}
-%% }
-
-%% \author{Alejandro Russo}
-%% \email{russo@chalmers.se}
-%% \affiliation{%
-%%   \institution{Chalmers University of Technology}
-%%   \city{Gothenburg}
-%%   \country{Sweden}
-%% }
+\author{Agust\'in Mista}
+\email{mista@chalmers.se}
+\affiliation{%
+  \institution{Chalmers University of Technology}
+  \city{Gothenburg}
+  \country{Sweden}
+}
+
+\author{Alejandro Russo}
+\email{russo@chalmers.se}
+\affiliation{%
+  \institution{Chalmers University of Technology}
+  \city{Gothenburg}
+  \country{Sweden}
+}
 
@@
 
-%% \vspace{-5pt}
+\vspace{-5pt}
 
@@
 produce new execution traces --- any other test case is simply thrown away.
-
+%
 Worth mentioning, CGPT is not meant to replace sophisticated manually-written
@@
 
-In this work, we establish several aspects%
-\footnote{Some of these were already identified by
-  \citeauthor{lampropoulos2019coverage} in their original work.}
-%
-of the original CGPT approach by \citeauthor{lampropoulos2019coverage} that
-leave considerable room for improvement (see \S \ref{sec:background}).
+In this work, we establish several aspects of the original CGPT approach by
+\citeauthor{lampropoulos2019coverage} that leave considerable room for
+improvement (see \S \ref{sec:background}).
+%
 %
@@
 %
+Notably, some of these limitations were already identified
+\citeauthor{lampropoulos2019coverage} in their original work.
+%
 %% Moreover, we replicated the Information-Flow Control (IFC) stack machine case
@@
   escape scalability issues} can be an appealing technique for
-finding bugs reliably without sacrificing speed.
+finding bugs reliably without sacrificing speed.%
+%
+\footnote{A replication package \cite{anon_2021} is available for reviewing
+  purposes.}
 
@@
 % ----------------------------------------
-%% \vspace{-5pt}
+%% \vspace{-10pt}
 
@@
 %
-Then, if we assume existence of a predicate (function) \texttt{balanced} of type
+Then, if we assume existence of a function \texttt{balanced} of type
 \texttt{Tree a -> Bool} that asserts that a tree satisfies a certain notion of
@@
 (The implementation details of \texttt{balanced} and \texttt{insert} are not
-relevant for the point being made here.)
-%
-Moreover, the arrow operator (\texttt{==>}) indicates that the predicate
-\texttt{balanced t} is a precondition of this property, and so attempting to
-test it using an unbalanced tree as an input will result in the test case
-getting \emph{discarded}.
+important here.)
+%
+Moreover, the arrow operator (\texttt{==>}) indicates that \texttt{balanced t}
+is a precondition of this property, and so attempting to test it using an
+unbalanced tree as input will result in the test case getting \emph{discarded}.
 
@@
 %% operation:
-We can write a na\"ive one for trees of integers (simplified to make it more accessible%
-\footnote{In practice, this code will most likely be implemented using Haskell's
-  type-class mechanism, hiding the complexity behind the size limit and the
-  payload generator.}%
-) as follows:
+We can write a na\"ive generator for trees of integers (simplified here to make it
+more accessible) as follows:
+%% \footnote{In practice, this code will most likely be implemented using Haskell's
+%%   type-class mechanism, hiding the complexity behind the size limit and the
+%%   payload generator.}%
 
@@
 \begin{verbatim}
-genTree(size) =
-  if size == 0 then do { x <- genInt; return (Leaf x) }
+genTree(size) = if size == 0 then do { x <- genInt; return (Leaf x) }
                 else oneof [ do { x <- genInt; return (Leaf x) }
@@
 %
-Enqueued test cases are stored along with a given mutation budget, that controls
+Enqueued test cases are stored along with a given mutation budget that controls
 how many times a given test case can be mutated before being finally thrown
@@
 %
-\footnote{An anonymized replication package \cite{anon_2021} is available for reviewing purposes.}
-%
 The first one is a simple abstract stack machine that enforces
@@
   \textit{haskell-wasm}.}
-\vspace{-10pt}
+\vspace{-30pt}
 \end{table}

### 2022-01-13 papers/ast22/results/wasm/prop_interpreter/mutagen-base/prop_interpreter_i32.BUG_INTERPRETER_2.MUTAGEN_NO_INHERITANCE.csv.6.hs

@@ -1,0 +1,1 @@
+(0,2.4723327,Function {funcType = 0, localTypes = [], body = [I32Const 0,GetLocal 0,IRelOp BS32 ILtU]})

### 2022-01-13 papers/ast22/results/wasm/prop_interpreter/mutagen-base/prop_interpreter_i32.BUG_INTERPRETER_3.MUTAGEN_NO_INHERITANCE.csv.8.hs

@@ -1,0 +1,1 @@
+(1,2.8412235,Function {funcType = 0, localTypes = [I32,F64], body = [I32Const 0,GetLocal 0,IBinOp BS32 IShrU]})

### 2022-01-13 papers/ast22/results/wasm/prop_interpreter/mutagen-base/prop_interpreter_i32.BUG_INTERPRETER_4.MUTAGEN_NO_INHERITANCE.csv.10.hs

@@ -1,0 +1,1 @@
+(2,1.9046015,Function {funcType = 0, localTypes = [], body = [Block {resultType = [], body = [GetLocal 0,Nop,I64Const 1,FReinterpretI BS64,Block {resultType = [I32,F64,F32], body = [Br 1]},GetLocal 0,BrTable [] 0]},I32Const 0,Return]})

### 2022-01-13 papers/ast22/results/wasm/prop_interpreter/mutagen-no-fifo/prop_interpreter_i32.BUG_INTERPRETER_4.MUTAGEN_NO_FIFO.csv.6.hs

@@ -1,0 +1,1 @@
+(0,-2.3602166,Function {funcType = 0, localTypes = [I32,F32], body = [GetLocal 2]})

### 2022-01-13 papers/ast22/results/wasm/prop_interpreter/mutagen-no-reset/prop_interpreter_i32.BUG_INTERPRETER_1.MUTAGEN_NO_RESET.csv.7.hs

@@ -1,0 +1,1 @@
+(2,1.2629294,Function {funcType = 0, localTypes = [I32], body = [F32Const (-1.6821574),IReinterpretF BS32,SetLocal 2,GetLocal 0,I64ExtendSI32,GetLocal 0,I64Load16U (MemArg {offset = 1, align = 1}),CurrentMemory,I64Load32U (MemArg {offset = 0, align = 1}),Loop {resultType = [F32,I32], body = [CurrentMemory,Return]},I32Const 0,Br 0]})

### 2022-01-13 papers/ast22/results/wasm/prop_interpreter/mutagen-no-reset/prop_interpreter_i32.BUG_INTERPRETER_5.MUTAGEN_NO_RESET.csv.3.hs

@@ -1,0 +1,1 @@
+(2,2.9703848,Function {funcType = 0, localTypes = [I64,F32], body = [I32Const 0,Loop {resultType = [F32,I32], body = [I32Const 1,Return,F32Load (MemArg {offset = 2, align = 1})]},GetLocal 0,I64Load16S (MemArg {offset = 2, align = 0}),I32Const 0,Br 0]})

### 2022-01-13 papers/ast22/results/wasm/prop_validator/mutagen-base/prop_validator.BUG_VALIDATOR_4.MUTAGEN_NO_INHERITANCE.csv.8.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []},FuncType {params = [], results = []},FuncType {params = [], results = []},FuncType {params = [], results = []}], functions = [], tables = [], mems = [Memory (Limit 9 (Just 9)),Memory (Limit 2 (Just 3))], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-01-13 papers/ast22/results/wasm/prop_validator/mutagen-no-fifo/prop_validator.MUTAGEN_NO_FIFO.BUG_VALIDATOR_1.csv.3.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [Br 0,If {resultType = [], true = [], false = [F32Const (-1.457094)]}]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-01-13 papers/ast22/results/wasm/prop_validator/mutagen-no-fifo/prop_validator.MUTAGEN_NO_FIFO.BUG_VALIDATOR_4.csv.18.hs

@@ -1,0 +1,1 @@
+Module {types = [], functions = [], tables = [], mems = [Memory (Limit 10 Nothing),Memory (Limit 0 (Just 4))], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-01-13 papers/ast22/results/wasm/prop_validator/mutagen-no-fifo/prop_validator.MUTAGEN_NO_FIFO.BUG_VALIDATOR_4.csv.21.hs

@@ -1,0 +1,1 @@
+Module {types = [], functions = [], tables = [Table (TableType (Limit 0 Nothing) AnyFunc)], mems = [Memory (Limit 5 (Just 6)),Memory (Limit 1 (Just 2)),Memory (Limit 0 (Just 6)),Memory (Limit 6 (Just 10)),Memory (Limit 2 Nothing)], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-01-13 papers/ast22/results/wasm/prop_validator/mutagen-no-fifo/prop_validator.MUTAGEN_NO_FIFO.BUG_VALIDATOR_4.csv.23.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [], tables = [], mems = [Memory (Limit 3 (Just 4)),Memory (Limit 2 (Just 9))], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-01-13 papers/ast22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_1.MUTAGEN_NO_RESET.csv.5.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [I32Const 0,If {resultType = [], true = [], false = [F32Const 6.327547]}]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-01-13 papers/ast22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_10.MUTAGEN_NO_RESET.csv.2.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [I32Const 0,I32Load (MemArg {offset = 0, align = 0}),I32Const 14,Nop,F64Const 0.0,F32Const 6.368479,Br 0]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-01-13 papers/ast22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_10.MUTAGEN_NO_RESET.csv.25.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [F64Const 7.845268042696084,Call 0,ITruncFS BS64 BS64,I64Eqz,I32Load (MemArg {offset = 4, align = 3}),Br 0]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-01-13 papers/ast22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_10.MUTAGEN_NO_RESET.csv.3.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [I32Const 0,I32Load (MemArg {offset = 1, align = 4}),Return]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-01-13 papers/ast22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_4.MUTAGEN_NO_RESET.csv.15.hs

@@ -1,0 +1,1 @@
+Module {types = [], functions = [], tables = [], mems = [Memory (Limit 3 (Just 6)),Memory (Limit 1 (Just 5))], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-01-13 papers/ast22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_6.MUTAGEN_NO_RESET.csv.17.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [Nop,I64Eqz,Return]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-01-13 papers/ast22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_7.MUTAGEN_NO_RESET.csv.1.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [I32Const 0,IBinOp BS32 IRemU,FConvertIU BS32 BS32,I64Const 0,Return,FUnOp BS64 FAbs,F32Const (-2.873685),Nop,Return,I64ExtendUI32,Drop]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-01-13 papers/ast22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_7.MUTAGEN_NO_RESET.csv.11.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [I32Const 0,IBinOp BS32 IAdd,I64ExtendSI32,Nop,Drop]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-01-13 papers/ast22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_7.MUTAGEN_NO_RESET.csv.14.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [I32Const 0,IBinOp BS32 IAdd,BrTable [] 0]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-01-13 papers/ast22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_7.MUTAGEN_NO_RESET.csv.21.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [F64Const 0.0,ITruncFS BS32 BS64,IBinOp BS32 IAdd,Drop]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-01-13 papers/ast22/main.tex
Replaced FIFO -> FIFO. Updated replication package link.

@@
 %
-In the first place, our tool uses first-in first-out (FIFO) scheduling with
+In the first place, our tool uses last-in first-out (LIFO) scheduling with
 priority to enqueue interesting test cases for mutation.
@@
 
-\subsection{Priority FIFO Scheduling}
+\subsection{Priority LIFO Scheduling}
 
@@
 
-\caption{\label{algo:mutagen:fifo}Priority FIFO Heuristic}
+\caption{\label{algo:mutagen:lifo}Priority LIFO Heuristic}
 \end{algorithm}
@@
 %
-These changes are illustrated in Algorithm \ref{algo:mutagen:fifo}.
+These changes are illustrated in Algorithm \ref{algo:mutagen:lifo}.
 %
@@
 %
-We call these variants \textit{no FIFO} and \textit{no reset}, respectively.
+We call these variants \textit{no LIFO} and \textit{no reset}, respectively.
 %
@@
 %
-this case study is not strongly affected by using FIFO scheduling.
+this case study is not strongly affected by using LIFO scheduling.
 %% --- the same
@@
 %
-On one hand, disabling the FIFO scheduling (case \textit{no FIFO}), the
+On one hand, disabling the LIFO scheduling (case \textit{no LIFO}), the
 hardest-to-find bugs (\#7, \#10, and \#1) are no longer found on every run.
@@
 %
-Moreover, by disabling the FIFO scheduling we observe that the performance of
+Moreover, by disabling the LIFO scheduling we observe that the performance of
 our tool becomes unstable, where bugs \#12 and \#11 take substantially less time

### 2022-01-13 papers/ast22/main.tex
Rearranged figures.

@@
 
-\begin{algorithm}[b]
+Finally, the seed selection algorithm (Algorithm \ref{algo:mutagen:pick}) picks
+the next test case using the same criteria as CGPT, prioritizing valid test
+cases over discarded ones, falling back to random generation when both queues
+are empty.
+%
+Since mutations are precomputed, this function simply picks the next test case
+from the current batch, until it becomes empty and can switch to the next
+precomputed one in line.
+
+Having selected the next test case, the testing loop proceeds to execute it,
+capturing both the result (valid, discarded, or failed) and its execution trace
+over the system under test.
+%
+If the test case fails, it is reported as a bug.
+%
+If not, the algorithm evaluates whether it was interesting (i.e., it exercises a
+new path) based on its trace information and the one from previously executed
+test cases (represented by TLog).
+%
+If the test case discovers a new path, it is enqueued on its corresponding
+queue.
+% depending on whether it was valid or got discarded.
+%
+%% (i.e., it failed the precondition).
+%
+This process alternates between generation and mutation until it finds a bug or
+reaches the test limit N.
+
+\begin{algorithm}[t]
   \SetArgSty{textnormal}
@@
 \end{algorithm}
-
-Finally, the seed selection algorithm (Algorithm \ref{algo:mutagen:pick}) picks
-the next test case using the same criteria as CGPT, prioritizing valid test
-cases over discarded ones, falling back to random generation when both queues
-are empty.
-%
-Since mutations are precomputed, this function simply picks the next test case
-from the current batch, until it becomes empty and can switch to the next
-precomputed one in line.
-
-Having selected the next test case, the testing loop proceeds to execute it,
-capturing both the result (valid, discarded, or failed) and its execution trace
-over the system under test.
-%
-If the test case fails, it is reported as a bug.
-%
-If not, the algorithm evaluates whether it was interesting (i.e., it exercises a
-new path) based on its trace information and the one from previously executed
-test cases (represented by TLog).
-%
-If the test case discovers a new path, it is enqueued on its corresponding
-queue.
-% depending on whether it was valid or got discarded.
-%
-%% (i.e., it failed the precondition).
-%
-This process alternates between generation and mutation until it finds a bug or
-reaches the test limit N.
 
@@
   \right]
-  ,
+  %
   \stacked{{\color{blue} new=1\phantom{x}}}{{\color{red} depth=3}}
@@
   \right]
-  ,
+  %
   \stacked{{\color{blue} new=2\phantom{x}}}{{\color{red} depth=2}}
@@
 \caption{\label{fig:mutagen:tracelog}Inserting two new execution traces into the
-  internal trace log (represented using brackets).}
+  internal execution trace log (represented using brackets).}
 \end{figure}
+
 \begin{algorithm}[t]

### 2022-01-13 papers/ast22/main.tex
Minor changes and fixed typos

@@
 produce new execution traces --- any other test case is simply thrown away.
-%
+
 Worth mentioning, CGPT is not meant to replace sophisticated manually-written
@@
 %% Our solution
-In this work, we introduce \mutagen, a CGPT framework that tackles the
-aforementioned limitations by applying mutations \emph{exhaustively} (see
-\S \ref{sec:mutagen}).
+To tackle the aforementioned limitations, we introduce \mutagen, a CGPT
+framework that applies mutations \emph{exhaustively} (see \S \ref{sec:mutagen}).
 %
@@
 \citeauthor{lampropoulos2019coverage} in their original work, and an existing
-{Web-Assembly} engine of industrial strenght written in the strongly-typed
+{Web-Assembly} engine of industrial strength written in the strongly-typed
 programming language Haskell.
@@
 %
-On each iteration, the algorithm selects the next test case by mutating the
-first value on the queue of valid test cases. If that queue is empty, it then
-selects the first one from the queue of discarded test cases.
+Then, on each iteration the algorithm selects the next test case by mutating the
+first value on the queue of valid test cases.
+%
+If that queue is empty, it instead mutates the first test case from the queue of
+discarded test cases.
 %
@@
 For this, each mutated test case remembers whether its parent (the original test
-case they derive from after being mutated) was valid.
+case they derive from before being mutated) was valid.
 %
@@
 
-This heuristic tackles the issue of enqueuing new interesting mutation
+This heuristic first tackles the issue of enqueuing new interesting mutation
 candidates at the end of (possibly) long queues of not-so-interesting previously
@@
   \item we start the testing loop with the number of random mutations set to
-    one,
+    one, and
   %
   \item each time we find that a test is not interesting (i.e. boring), we
-    increment a counter,
+    increment a counter.
+  %
+  Then,
   \item if we have not produced any interesting test case after a certain
@@
 
-Then, if the execution of the system under test depends heavily on the values
-stored at randomly mutable subexpressions, starting with a single random
-mutation will quickly saturate the trace space, and this heuristic will
-continuously increase the random mutations parameter until that stops happening.
+Notably, this heuristic can be particularly useful when the execution of the
+system under test depends heavily on some invariant encoded by a subexpression
+of its input that get transformed using a random mutant (e.g, the number of
+pixels declared by the header of an image matching the size of its actual
+payload).
+%
+There, starting with a single random mutation will quickly saturate the trace
+space with discarded (invalid) tests, and this heuristic will continuously
+increase the random mutations parameter until the random mutant is given enough
+chances to randomly generate the value that satisfies the required invariant and
+makes the overall test case valid.
 
@@
 
-We evaluated the performance of \mutagen using two case studies.%
-%
-The first one is a simple abstract stack machine that enforces
-the hyperproperty \emph{noninterference} \cite{goguen1982security} using runtime checks.
+We evaluated the performance of \mutagen using two case studies.
+%
+The first one is a simple abstract stack machine that enforces the hyperproperty
+\emph{noninterference} \cite{goguen1982security} using runtime checks.
 %
@@
 %
-To simplify things, we write a function \texttt{mk\_module} to build a stub
-module which initializes one memory block and exports a single WebAssembly
+To simplify things, we write a helper function \texttt{mk\_module} to build a
+stub WebAssembly module that initializes one memory block and exports a single
 function.
 %
-Such function is parameterized by the definition of the single function, along
-with its name and type signature.
+This helper is parameterized by the definition of the module's single function,
+along with its type signature and name.
 %
@@
 prop_interpreter ty fun args =
-  do let m = mk_module ty "foo" fun
-     resHs   <- invokeHaskellWasm m "foo" args
-     resSpec <- invokeRefImpl     m "foo" args
+  do let m = mk_module fun ty "f"
+     resHs   <- invokeHaskellWasm m "f" args
+     resSpec <- invokeRefImpl     m "f" args
      return (equivalent resHs resSpec)
@@
 
-This property instantiates a module stub \texttt{m} using the input function
-(\texttt{fun}) and its type signature (\texttt{ty}).
-%
-Then, it invokes the function \texttt{foo} of the module \texttt{m} both on
+This testing property instantiates a module stub \texttt{m} with the input
+WebAssembly function (\texttt{fun}) and its type signature (\texttt{ty}).
+%
+Then, it invokes the function \texttt{f} of the module \texttt{m} both on
 \textit{haskell-wasm} and the reference interpreter with the provided arguments
@@
 %
-The reason behind this is that the mutation queues remain empty most of the time
-(generation mode), and when a new interesting candidate gets inserted, all its
-mutants (and their descendants) are processed before the next one is enqueued.
-%
-Hence, the small proportion of time this heuristic is active tends to be
-inconsequential.
+We noticed that when a new interesting test case gets enqueued, all its mutants
+(and their descendants) are quickly processed before new ones start piling up,
+keeping the mutation queues empty most of the time (generation mode).
+%
+Hence, the small proportion of time this heuristic is active (when the mutation
+queues are not empty) tends to be inconsequential.
 %% under this case study.
@@
 %% interpreter, we will focus on these results independently.
-
+%
 We first focus on the bugs injected in the validator (Fig.
@@
 %
-Despite this, our tool is still capable of running substantially more tests that
-pass the sparse preconditions than \quickcheck in the same amount of time,
-\emph{and that ultimately leads us to find bugs}.
+However, our tool is capable of running substantially more tests that pass the
+sparse preconditions than \quickcheck in the same amount of time, \emph{which
+  can ultimately lead us to find more bugs}.
 

### 2022-03-01 papers/fse22/results/wasm/prop_interpreter/mutagen-base/prop_interpreter_i32.BUG_INTERPRETER_2.MUTAGEN_NO_INHERITANCE.csv.3.hs
Added FSE22 initial version

@@ -1,0 +1,1 @@
+(1,0.59896225,Function {funcType = 0, localTypes = [F64], body = [I32Const 0,GrowMemory,GetLocal 0,IRelOp BS32 ILtU]})

### 2022-03-01 papers/fse22/results/wasm/prop_interpreter/mutagen-base/prop_interpreter_i32.BUG_INTERPRETER_2.MUTAGEN_NO_INHERITANCE.csv.7.hs

@@ -1,0 +1,1 @@
+(2,1.6834629,Function {funcType = 0, localTypes = [F32], body = [GetLocal 0,GetLocal 0,IRelOp BS32 ILtU]})

### 2022-03-01 papers/fse22/results/wasm/prop_interpreter/mutagen-base/prop_interpreter_i32.BUG_INTERPRETER_2.MUTAGEN_NO_INHERITANCE.csv.9.hs

@@ -1,0 +1,1 @@
+(2,2.4721553,Function {funcType = 0, localTypes = [I32], body = [I32Const 0,I32Const 0,IRelOp BS32 ILtU]})

### 2022-03-01 papers/fse22/results/wasm/prop_interpreter/mutagen-base/prop_interpreter_i32.BUG_INTERPRETER_3.MUTAGEN_NO_INHERITANCE.csv.5.hs

@@ -1,0 +1,1 @@
+(2,2.0491881,Function {funcType = 0, localTypes = [F64,I32,F32], body = [I32Const 0,Block {resultType = [], body = [Loop {resultType = [], body = []}]},I32Const 2,IBinOp BS32 IShrU,Return]})

### 2022-03-01 papers/fse22/results/wasm/prop_interpreter/mutagen-base/prop_interpreter_i32.BUG_INTERPRETER_4.MUTAGEN_NO_INHERITANCE.csv.3.hs

@@ -1,0 +1,1 @@
+(1,4.4687018e-2,Function {funcType = 0, localTypes = [I32,F64], body = [GetLocal 2]})

### 2022-03-01 papers/fse22/results/wasm/prop_interpreter/mutagen-base/prop_interpreter_i32.BUG_INTERPRETER_4.MUTAGEN_NO_INHERITANCE.csv.7.hs

@@ -1,0 +1,1 @@
+(1,1.2634474,Function {funcType = 0, localTypes = [], body = [Block {resultType = [I32,I32], body = [GetLocal 0,Br 0]},I32Const 1,Return]})

### 2022-03-01 papers/fse22/results/wasm/prop_interpreter/mutagen-no-fifo/prop_interpreter_i32.BUG_INTERPRETER_2.MUTAGEN_NO_FIFO.csv.10.hs

@@ -1,0 +1,1 @@
+(0,0.91413224,Function {funcType = 0, localTypes = [I32], body = [I32Const 0,I32Const 0,IRelOp BS32 ILtU]})

### 2022-03-01 papers/fse22/results/wasm/prop_interpreter/mutagen-no-fifo/prop_interpreter_i32.BUG_INTERPRETER_3.MUTAGEN_NO_FIFO.csv.6.hs

@@ -1,0 +1,1 @@
+(1,-1.7670708,Function {funcType = 0, localTypes = [I32,F32,F32], body = [GetLocal 2,CurrentMemory,IBinOp BS32 IShrU]})

### 2022-03-01 papers/fse22/results/wasm/prop_interpreter/mutagen-no-reset/prop_interpreter_i32.BUG_INTERPRETER_2.MUTAGEN_NO_RESET.csv.3.hs

@@ -1,0 +1,1 @@
+(1,0.6765671,Function {funcType = 0, localTypes = [I64,I32], body = [GetLocal 0,GetLocal 0,IRelOp BS32 ILtU]})

### 2022-03-01 papers/fse22/results/wasm/prop_interpreter/mutagen-no-reset/prop_interpreter_i32.BUG_INTERPRETER_2.MUTAGEN_NO_RESET.csv.4.hs

@@ -1,0 +1,1 @@
+(1,0.4831685,Function {funcType = 0, localTypes = [F64,F32], body = [F64Const 0.0,Loop {resultType = [F32,F64], body = [CurrentMemory,Return]},Block {resultType = [], body = [Block {resultType = [], body = []}]},I32Const 1,I64Load32S (MemArg {offset = 0, align = 1}),CurrentMemory,F64Load (MemArg {offset = 1, align = 3}),I32Const 0,I64ExtendSI32,GetLocal 0,Return,GrowMemory]})

### 2022-03-01 papers/fse22/results/wasm/prop_interpreter/quickcheck/prop_interpreter_i32.BUG_INTERPRETER_4.QuickCheck.csv.3.hs

@@ -1,0 +1,1 @@
+["(0,0.0,Function {funcType = 0, localTypes = [I32,I64,I32,I32,F64,F32,I64,I32], body = [GetLocal 5]})"]

### 2022-03-01 papers/fse22/results/wasm/prop_validator/mutagen-base/prop_validator.BUG_VALIDATOR_4.MUTAGEN_NO_INHERITANCE.csv.20.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []},FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = []}], tables = [], mems = [Memory (Limit 1 (Just 4)),Memory (Limit 5 (Just 7)),Memory (Limit 5 Nothing)], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-03-01 papers/fse22/results/wasm/prop_validator/mutagen-base/prop_validator.BUG_VALIDATOR_4.MUTAGEN_NO_INHERITANCE.csv.23.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []},FuncType {params = [], results = []}], functions = [], tables = [], mems = [Memory (Limit 9 Nothing),Memory (Limit 6 Nothing)], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-03-01 papers/fse22/results/wasm/prop_validator/mutagen-base/prop_validator.BUG_VALIDATOR_4.MUTAGEN_NO_INHERITANCE.csv.26.hs

@@ -1,0 +1,1 @@
+Module {types = [], functions = [], tables = [], mems = [Memory (Limit 1 (Just 5)),Memory (Limit 4 (Just 8)),Memory (Limit 4 Nothing),Memory (Limit 2 (Just 8))], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-03-01 papers/fse22/results/wasm/prop_validator/mutagen-base/prop_validator.BUG_VALIDATOR_9.MUTAGEN_NO_INHERITANCE.csv.19.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = [I32]}], functions = [Function {funcType = 0, localTypes = [], body = [Nop,I64Const 24,Return]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-03-01 papers/fse22/results/wasm/prop_validator/mutagen-no-fifo/prop_validator.MUTAGEN_NO_FIFO.BUG_VALIDATOR_10.csv.23.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [Br 0,Nop,I32Load (MemArg {offset = 0, align = 0}),Return]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-03-01 papers/fse22/results/wasm/prop_validator/mutagen-no-fifo/prop_validator.MUTAGEN_NO_FIFO.BUG_VALIDATOR_3.csv.18.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = [I32]}], functions = [Function {funcType = 0, localTypes = [], body = []}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-03-01 papers/fse22/results/wasm/prop_validator/mutagen-no-fifo/prop_validator.MUTAGEN_NO_FIFO.BUG_VALIDATOR_4.csv.28.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [], tables = [], mems = [Memory (Limit 6 Nothing),Memory (Limit 2 (Just 5))], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-03-01 papers/fse22/results/wasm/prop_validator/mutagen-no-fifo/prop_validator.MUTAGEN_NO_FIFO.BUG_VALIDATOR_8.csv.20.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [CurrentMemory,Return]},Function {funcType = 0, localTypes = [], body = []}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-03-01 papers/fse22/results/wasm/prop_validator/mutagen-no-fifo/prop_validator.MUTAGEN_NO_FIFO.BUG_VALIDATOR_9.csv.9.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []},FuncType {params = [], results = []},FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = []}], tables = [], mems = [], globals = [Global {globalType = Const I32, initializer = [I64Const 98]}], elems = [], datas = [], start = Nothing, imports = [Import {sourceModule = "", name = "", desc = ImportFunc 0}], exports = []}

### 2022-03-01 papers/fse22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_1.MUTAGEN_NO_RESET.csv.24.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [I32Const 0,If {resultType = [], true = [Nop], false = [F32Const 4.8760715]}]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-03-01 papers/fse22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_1.MUTAGEN_NO_RESET.csv.6.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [I64Const 0,FConvertIS BS32 BS64,I32Const 17,Nop,I64Const 16,Block {resultType = [I64,I32,F32,F64,I32,F64,I32], body = [Br 1]},I32Const 10,Return]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-03-01 papers/fse22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_1.MUTAGEN_NO_RESET.csv.9.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [F64Const 0.0,Loop {resultType = [F64,I64,I32,I64,F64], body = [Return]},F32Const (-0.8032148),Nop,F32Const 0.0,Return]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-03-01 papers/fse22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_10.MUTAGEN_NO_RESET.csv.12.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [I32Const 32,I32Load (MemArg {offset = 6, align = 10}),I64ExtendUI32,Drop]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-03-01 papers/fse22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_2.MUTAGEN_NO_RESET.csv.18.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [Nop,I64Const 23,Nop,Nop]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-03-01 papers/fse22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_2.MUTAGEN_NO_RESET.csv.23.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [Nop,F64Const (-0.5083633480929778),Nop]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-03-01 papers/fse22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_4.MUTAGEN_NO_RESET.csv.4.hs

@@ -1,0 +1,1 @@
+Module {types = [], functions = [], tables = [], mems = [Memory (Limit 7 (Just 9)),Memory (Limit 5 (Just 6)),Memory (Limit 1 (Just 9))], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-03-01 papers/fse22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_6.MUTAGEN_NO_RESET.csv.18.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [F32Const 0.0,F32Const 4.623046,I64Eqz,Return]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-03-01 papers/fse22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_7.MUTAGEN_NO_RESET.csv.13.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [I32Const 0,I32Eqz,IBinOp BS32 IDivU,Nop,BrTable [] 0]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-03-01 papers/fse22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_7.MUTAGEN_NO_RESET.csv.23.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [F32Const 0.0,I32Const 31,IBinOp BS32 IAdd,Loop {resultType = [], body = []},F32Const (-5.808929),Nop,Block {resultType = [], body = []},Return]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-03-01 papers/fse22/results/wasm/prop_validator/mutagen-no-reset/prop_validator.BUG_VALIDATOR_9.MUTAGEN_NO_RESET.csv.23.hs

@@ -1,0 +1,1 @@
+Module {types = [FuncType {params = [], results = []}], functions = [Function {funcType = 0, localTypes = [], body = [I64Const 0,If {resultType = [], true = [], false = []},Call 0]}], tables = [], mems = [], globals = [], elems = [], datas = [], start = Nothing, imports = [], exports = []}

### 2022-03-02 papers/fse22/main.tex
Added FSE22 initial version
Compressed abstract

--- 2022-01-13 papers/ast22/main.tex
@@
 %% The first command in your LaTeX source must be the \documentclass command.
-\documentclass[sigconf, review]{acmart}
+% \documentclass[sigconf, review]{acmart}
+\documentclass[sigconf,review,anonymous]{acmart}
 
@@
 %% These commands are for a JOURNAL article.
-\acmConference[AST 2022]{3rd ACM/IEEE International Conference on Automation of Software Test}{May 21-22, 2022}{}
-
+\acmConference[ESEC/FSE 2022]{The 30th ACM Joint European Software Engineering Conference and Symposium on the Foundations of Software Engineering}{14 - 18 November, 2022}{Singapore}
 %% \acmJournal{JACM}
@@
 %
-Such generators can be obtained using a variety of techniques that extract
-static information from the codebase in order to produce random test cases.
-%
-However, such techniques are not suitable for deriving generators producing
-random values satisfying complex invariants, which in turn complicates testing
-properties with sparse preconditions.
-
-Coverage-guided, property-based testing (CGPT) is a recent technique that
-alleviates this limitation by enhancing automatically synthesized generators
-with structure-preserving mutations guided by execution traces.
-%
-Albeit effective, the initial CGPT approach is limited by factors like large
-reliance on randomness and poor mutants scheduling, reducing its applicability
-and bug-finding capacity in existing software.
-
-In this work we present \mutagen, a CGPT framework that overcomes these
-limitations by generating mutants in an \emph{exhaustive} manner.
-%
-This is coupled with heuristics that help to schedule mutants based on their
-novelty and to minimize mutating certain subexpressions leading to scalability issues of
-exhaustiveness.
+There exists a variety of techniques that extract static information from the
+codebase to produce random test cases.
+%
+Sadly, such techniques cannot derive generators capable of producing random
+values satisfying complex invariants, often needed to test properties with
+sparse preconditions.
+
+Coverage-guided, property-based testing (CGPT) is a technique that tackles this
+limitation by enhancing synthesized generators with structure-preserving
+mutations guided by execution traces.
+%
+Albeit effective, CGPT relies largely on randomness and exhibits poor
+scheduling, preventing its applicability in real-world software.
+
+This work presents \mutagen, a CGPT framework that overcomes such limitations by
+generating mutants \emph{exhaustively}.
+%
+Our tool incorporates heuristics to help minimizing scalability issues as well
+as covering the search space in a principled manner.   
+%
+% schedule mutants based on their
+% novelty and to minimize mutating certain subexpressions leading to scalability
+% issues of exhaustiveness.
 %of large enumeration types more
@@
 %
-Our results indicate that our approach is capable of outperforming existing CGPT
-tools, as well as finding bugs in real-world scenarios.
+Our evaluation show that \mutagen not only outperforms existing CGPT tools, but
+also finds previously unknwown bugs in real-world software.
 \end{abstract}
@@
 
-\vspace{-5pt}
+% \vspace{-5pt}
 

### 2022-03-02 papers/fse22/main.tex
Compressed introduction

@@
 %
-This is achieved by using randomly generated inputs to instantiate executable
-properties encoding the expected behavior of the system under test.
-%
-%% Exemplified by frameworks like Haskell's \quickcheck \cite{ClaessenH00} and the
-%% Abundant ports of this tool in other programming languages
-%% \cite{hughes2003erlang, papadakis2011proper, bulwahn2012new,
-%%   denes2014quickchick}
-%
-%% The most common PBT approach is to instantiate the testing properties using
-%% randomly generated inputs.
-%
-However, users of \quickcheck (and RPBT in general) are well aware that one of
-the biggest challenges while using RPBT is to provide suitable random data
-generators needed to instantiate the testing properties.
-%
-In extreme cases, writing highly-tuned random generators capable of exercising
-every part of a complex system on a reasonable basis can take several thousand
-person-hours of trial and error \cite{lampropoulos2019coverage}.
-%
-To alleviate this, there exist several techniques to automatically synthesize
-random data generators by reifying the static information present in the
-codebase, e.g., data type definitions, application public interfaces (APIs),
-etc. \cite{GriecoCB16, DBLP:conf/haskell/MistaRH18, Mista2019GeneratingRS,
+% This is achieved by using randomly generated inputs to instantiate executable
+It uses randomly generated inputs to instantiate executable properties encoding
+the expected behavior of the system under test.
+%
+%% Exemplified by frameworks like Haskell's \quickcheck \cite{ClaessenH00} and
+%the % Abundant ports of this tool in other programming languages %
+%\cite{hughes2003erlang, papadakis2011proper, bulwahn2012new, %
+%denes2014quickchick}
+%
+%% The most common PBT approach is to instantiate the testing properties using %
+%randomly generated inputs.
+%
+% However, users of RPBT in general) are well aware that one of the biggest
+% challenges while using RPBT is to provide suitable random data generators
+% needed to instantiate the testing properties.
+However, a limiting factor of RPBT is having to provide the random data
+generators used to instantiate the testing properties. 
+%
+Writing highly-tuned generators capable of exercising every part of a complex
+system on a reasonable basis can take several thousand person-hours of trial and
+error \cite{lampropoulos2019coverage}.
+%
+Luckily, there exist several approaches to automatically synthesize random data
+generators by reifying the static information present in the codebase, e.g.,
+data type definitions, application public interfaces (APIs), etc.
+\cite{GriecoCB16, DBLP:conf/haskell/MistaRH18, Mista2019GeneratingRS,
 DuregardJW12, Lampropoulos2017, Bendkowski2017}.
 %
-These techniques, however, are unable to synthesize random generators capable of
+These approaches, however, are unable to synthesize generators capable of
 producing data satisfying complex invariants not easily derivable from the
@@
 %
-Randomly generating valid programs to test compilers is a common example of this
-problem \cite{chen2020asurvey}, where developers are forced to put substantial
-efforts in writing specialized random generators by hand \cite{Palka11,
+Randomly generating valid programs to test compilers is an example of this
+limitation \cite{chen2020asurvey}, where developers are forced to put
+substantial efforts in writing specialized generators by hand \cite{Palka11,
 perenyi2020stack, yang2011finding}.
 %
-Moreover, automatically synthesized random generators offer poor results when
-the testing properties are constrained by sparse preconditions.
-%
-In this scenario, most of the generated tests are simply thrown away before they
-can be used to test the main components of the system under test.
+Moreover, automatically synthesized generators behave poorly when used to test
+properties with sparse preconditions.
+%
+In this scenario, most tests are simply thrown away before reaching the main
+components of the system under test.
 
@@
 Instead of continuously generating random invalid test cases from scratch, CGPT
-works by keeping queues of \emph{interesting} previously executed test cases
-that can be mutated using high-level, structure-preserving transformations to
-produce new ones.
+keeps queues of \emph{interesting} previously executed test cases that can be
+transformed using structure-preserving mutations to produce new ones.
 %
@@
 %
-Intuitively, mutating an interesting test case in a small way (at the data
-constructor level) has a higher probability of producing a new interesting test
-case than generating a new one using a na\"ive generator.
-%
-In addition, high-level, structure-preserving mutators are better suited for
-producing syntactically valid mutants.
-%
-High-level mutators avoid wasting time with generic low-level mutators that work
-at the bit level and know very little about data structures, thus producing
-syntactically broken mutants most of the time.
-%
-Such a grammar-aware mutation technique has shown to be quite useful when
-fuzzing systems accepting structurally complex inputs \cite{holler2012fuzzing,
-  wang2019superion, xsmith}.
-%
-While most of the tools following this itechnique use existing grammars to
-tailor the mutators to the specific input structure used by the system under
-test, external grammars are not required in CGPT.
+Intuitively, mutating an existing interesting test case is more likely to
+produce a new interesting test case than generating a new one from scratch.
+
+Unlike the generic bit-level mutators often used in the fuzzing community
+\cite{}, structure-preserving mutations specified at the data type level can
+effectively produce syntactically valid mutants. 
+%
+% %
+% High-level mutators avoid wasting time with generic low-level mutators that work
+% at the bit level and know very little about data structures, thus producing
+% syntactically broken mutants most of the time.
+%
+This grammar-aware technique has shown to be effective when fuzzing systems
+accepting structurally complex inputs \cite{holler2012fuzzing, wang2019superion,
+xsmith}.
+%
+In practice, most tools use existing grammars to tailor mutations to the input
+structure used by the system under test. 
+%
+% Notably external grammars are not required in CGPT.
 %
@@
 %
-%% In CGPT, external grammars are not required.
-%
-The data types of the inputs to the testing properties already provide a good source
-of information about how data can be mutated, e.g., by changing one data constructor
-by another.
+% In CGPT, external grammars are not required.
+%
+In contast, CGPT relies on the data type information of the inputs to the
+testing properties to derive specialized structure-preserving mutators directly
+--- making strongly-typed programming languages an ideal match for this
+technique.
+%
+% already provide a good source of information about how data can be mutated,
+% e.g., by changing one data constructor by another.
+%
 %
@@
 % structure of the random data one needs to mutate in a concrete manner, and
-Specialized mutators can be automatically derived directly from data types
-definitions --- making strongly-typed programming languages an ideal match for
-CGPT.
+% Specialized mutators can be automatically derived directly from data types
+% definitions ---
+%
 
@@
 %
-Moreover, CGPT relies on execution traces to distinguish which executed test
-cases were interesting and are therefore worth mutating --- a popular technique
-known as \emph{coverage-guided fuzzing} and popularized by tools like \emph{AFL}
-\cite{afl}.
-%
-Mutated test cases are considered interesting for mutation only when they
+% Moreover, CGPT uses execution traces to distinguish which executed test cases
+Moreover, CGPT uses execution traces to distinguish interesting test cases --- a
+popular technique known as \emph{coverage-guided fuzzing} and popularized by
+tools like \emph{AFL} \cite{afl}.
+%
+% Mutated test cases are considered interesting for mutation
+Here, test cases are interesting (and therefore worth mutating) only when they
 produce new execution traces --- any other test case is simply thrown away.
 
-Worth mentioning, CGPT is not meant to replace sophisticated manually-written
-random generators, but rather to provide an adequate testing solution for early
-stages of development.
+% Worth mentioning, CGPT is not meant to replace sophisticated manually-written
+% random generators, but rather to provide an adequate testing solution for early
+% stages of development.
 %, and while a better test suite using manually-written generators is
@@
 \begin{inparaenum}
-\item structure-preserving mutations, if not done carefully, can become
-  superficial and rely heavily on randomness to produce diverse mutants,
-\item the queuing mechanism can cause delays if interesting values are enqueued
-  frequently and there is no way to prioritize them, and
-\item using an heuristic to assign a certain ``mutation budget'' to each
-  interesting test case (often referred to as a \emph{power schedule}) requires
-  fine tuning and can be hard to generalize to different kinds of testing
-  properties.
+\item automatically derived structure-preserving mutators can turn out
+"shallow", ineffective at transforming deep test cases more than superficially;
+  % superficial and rely heavily on randomness to produce diverse mutants,
+\item the queuing mechanism can cause delays if interesting test cases are
+  enqueued frequently and there is no way to prioritize them; and
+\item the heuristic used to assign a ``mutation budget'' to each interesting
+  test case (often referred to as a \emph{power schedule}) requires fine tuning
+  and can be hard to generalize.
+%to different kinds of testing properties.
 \end{inparaenum}
 %
-Notably, some of these limitations were already identified
-\citeauthor{lampropoulos2019coverage} in their original work.
+\footnote{Some of these limitations were already identified by the authors in
+their original work.}
 %
@@
 %% Our solution
-To tackle the aforementioned limitations, we introduce \mutagen, a CGPT
-framework that applies mutations \emph{exhaustively} (see \S \ref{sec:mutagen}).
+To tackle these limitations, we introduce \mutagen, a CGPT framework that
+applies mutations \emph{exhaustively} (see \S \ref{sec:mutagen}).
 %
@@
 %
-That is, given an interesting, previously executed test case, our tool
-precomputes and schedules every structure-preserving mutation that can be
-applied to it.
+That is, given an interesting test case, our tool precomputes and schedules
+every structure-preserving mutation that can be applied to it exactly once.
 %
@@
 %
-On one hand, mutations do not rely on randomness to be generated.
-%
-Instead, every subexpression of the input test case is mutated on the same
-basis, which guarantees that no interesting mutation is omitted due to
-randomness.
+On one hand,
+%  mutations do not rely on randomness to be generated.
+%
+every subexpression of the input test case is mutated on the same basis,
+guaranteeing that no interesting transformation is omitted due to randomness.
 %
@@
 On the other hand, scheduling mutations exhaustively eliminates the need for a
-power schedule to assign a given mutation budget to each input test case.
+power schedule to assign a mutation budget to every interesting test case.
 
@@
 %
-Our tool distinguishes two kinds of mutations, those that can be computed
-deterministically, yielding a single mutated value; and those that can be
-obtained non-deterministically.
-%
-On one hand, deterministic mutations encode transformations that swap data
-constructors around, as well as return or rearrange subexpressions.
-%
-Non-deterministic (random) mutations, on the other hand, are useful to represent
-transformations over large enumeration types, e.g., numbers, characters, etc.
+Internally, our tool distinguishes two kinds of mutations.
+%  those that can be computed
+% deterministically, 
+% and those that can be
+% obtained non-deterministically.
+%
+On one hand, \emph{deterministic (pure) mutations} encode transformations that
+yield a single mutated test case obtained by swapping data constructors around,
+as well as rearranging or returning subexpressions.
+%
+On the other hand, \emph{non-deterministic (random) mutations} are used to
+represent transformations over large enumeration types.
+%
+% like numbers or characters.
 %
 This mechanism let us selectively escape the scalability issues of
-exhaustiveness by randomly sampling a small random generator a reduced number of
-times.
-%
-In this way, our tool avoids, for instance, mutating a given number into every
-other number in the 32 or 64 bits range.
+exhaustiveness by yielding a random generator that replaces a specific
+subexpression of an input test case with a randomly generated one.
+%
+Such generator is then sampled a \emph{small} number of times. 
+%
+This way \mutagen avoids, for instance, mutating every numeric subexpresion of a
+test case into every other possible number.
 
 %% Heuristics to improve testing loop
-Additionally, the testing loop of \mutagen incorporates two novel heuristics
-that help finding bugs faster and more reliably (\S \ref{sec:heuristics}).
+In addition, the testing loop of \mutagen incorporates two novel heuristics that
+help finding bugs faster and more reliably (\S \ref{sec:heuristics}).
 %
 In the first place, our tool uses last-in first-out (LIFO) scheduling with
-priority to enqueue interesting test cases for mutation.
-%
-This scheduling algorithm is indexed by the novelty of each new test case with
-respect to the ones already seen.
-%
-In this light, interesting test cases that discover new parts of the code
-earlier during execution are given a higher priority.
-%
-Moreover, our scheduling allows the testing loop to jump back and forth between
-mutable test cases as soon as a new more interesting one is enqueued for
-mutation, eliminating potential delays when the mutation queues grow large.
-
-The second heuristic adjusts the number of times our tool samples random
-mutants, i.e., those corresponding to large enumeration types as described
-above.
+priority when enqueueing interesting test cases for mutation.
+%
+The priority is obtained by comparing the novelty of each new test case with
+respect to already seen ones.
+%
+This way, interesting test cases that discover new parts of the code earlier
+during execution are given a higher priority.
+%
+Moreover, this scheduling allows the testing loop to jump back and forth between
+enqueued test cases as soon as new more interesting ones become available,
+eliminating potential delays when the mutation queues grow large.
+
+The second heuristic adjusts the number of test cases generated by random
+mutations. 
+%
+% number of times our tool samples random
+% mutants, i.e., those corresponding to large enumeration types as described
+% above.
 %
@@
 %
-If this frequency suddenly stalls, \mutagen resets the testing loop increasing
-the amount of times we sample such mutants.
-%
-This automatically finds a suitable value for this external parameter on the
-fly.
+If this frequency stalls, \mutagen resets the testing loop and increases the
+effort put into sampling random mutations.
+%
+This automatically finds a suitable value for this parameter on the fly.
 
 %% Validation
-To validate our ideas, we use two case studies (described in detail in \S
+We validate our ideas using two case studies (described in detail in \S
 \ref{sec:casestudies}): the Information-Flow Control (IFC) stack machine used by
 \citeauthor{lampropoulos2019coverage} in their original work, and an existing
-{Web-Assembly} engine of industrial strength written in the strongly-typed
-programming language Haskell.
-%
-In both case studies, we additionally compare the effect of the heuristics
-implemented on top of the base testing loop of \mutagen.
+{Web-Assembly} engine of industrial strength written in Haskell.
+%
+In both case studies, we compare the effect of the heuristics implemented on top
+of the base testing loop of \mutagen.
 %
@@
 There, \mutagen is capable of reliably finding 15 planted bugs in the validator
-and interpreter, as well as 3 previously unknown bugs that flew under the radar
-of the existing unit test suite of this engine.
-%
-This case study also help us to compare the performance of our tool against a
-more traditional RPBT approach that does not rely on code instrumentation.
+and interpreter, as well as 3 previously unknown ones.
+% bugs that flew under the radar
+%of the existing unit test suite of this engine.
+%
+% This case study also help us to compare the performance of our tool against a
+% more traditional RPBT approach that does not rely on code instrumentation.
+In this case study we also compare the performance versus the overhead of our
+tool (and our custom code instrumentation mechanism) by comparing it against
+\quickcheck (black-box).  
 %
 All in all, our evaluation encompasses more than 600 hours of computing time and
-suggests that \emph{testing mutants exhaustively together with our heuristics to
-  escape scalability issues} can be an appealing technique for
-finding bugs reliably without sacrificing speed.%
+indicates that \emph{testing mutants exhaustively together with our heuristics
+to escape scalability issues} can be an appealing technique for finding bugs
+reliably without sacrificing speed.%
 %

### 2022-03-02 papers/fse22/main.tex
Compressing Background

@@
 
-In this work, we establish several aspects of the original CGPT approach by
-\citeauthor{lampropoulos2019coverage} that leave considerable room for
-improvement (see \S \ref{sec:background}).
-%
+In this work, we establish several aspects%
+%
+\footnote{Some of these limitations were already identified by the authors in
+their original work.}
+%
+of the original CGPT approach by \citeauthor{lampropoulos2019coverage} that
+leave considerable room for improvement (see \S \ref{sec:background}).
 %
@@
 %
-\footnote{Some of these limitations were already identified by the authors in
-their original work.}
-%
 %% Moreover, we replicated the Information-Flow Control (IFC) stack machine case
@@
 %% Validation
-We validate our ideas using two case studies (described in detail in \S
+We validated our ideas using two case studies (described in detail in \S
 \ref{sec:casestudies}): the Information-Flow Control (IFC) stack machine used by
@@
 
-In this section, we briefly describe the motivation, ideas and limitations
-behind CGPT \cite{lampropoulos2019coverage}.
-%
-To illustrate this technique, let us focus on a simple property defined over
-binary trees.
+This section briefly introduces the motivation, ideas and limitations behind
+CGPT \cite{lampropoulos2019coverage}.
+%
+To illustrate this technique, we focus on a simple property defined over binary
+trees.
 %
@@
 \noindent The type parameter \texttt{a} indicates that trees can be instantiated
-using any other type as payload, so the value \texttt{Leaf True} has type
-\texttt{Tree Bool}, whereas the value \texttt{Branch (Leaf 1) 2 (Leaf 3)} has
-type \texttt{Tree Int}.
+using any type as payload, so the value \texttt{Leaf True} has type \texttt{Tree
+Bool}, whereas the value \texttt{Branch (Leaf 1) 2 (Leaf 3)} has type
+\texttt{Tree Int}.
 %
@@
 %
-Then, if we assume existence of a function \texttt{balanced} of type
-\texttt{Tree a -> Bool} that asserts that a tree satisfies a certain notion of
-balancedness, we can write testing properties to assert that the operations
-defined over the \texttt{Tree} data type preserve this invariant.
-%
-For instance, the following testing property can be used to find bugs in the
-implementation of a given \texttt{insert} function:
+If we assume existence of a function \texttt{balanced} of type \texttt{Tree a ->
+Bool} that asserts that a tree satisfies a certain notion of balancedness, we
+can write testing properties to validate that the operations defined over the
+\texttt{Tree} data type preserve this invariant.
+%
+For instance, the following can be used to validate the implementation of an
+\texttt{insert} function:
 
@@
 %
-Moreover, the arrow operator (\texttt{==>}) indicates that \texttt{balanced t}
-is a precondition of this property, and so attempting to test it using an
-unbalanced tree as input will result in the test case getting \emph{discarded}.
-
-The last missing piece is a random generator of trees.
+The arrow operator (\texttt{==>}) indicates that \texttt{balanced t} is a
+precondition of this property, and so using an unbalanced tree as input will
+result in the test case getting \emph{discarded}.
+
+The only missing piece is a random generator of trees.
+%
+For this, we can define a na\"ive generator for trees of integers as:
 %
@@
 %% operation:
-We can write a na\"ive generator for trees of integers (simplified here to make it
-more accessible) as follows:
 %% \footnote{In practice, this code will most likely be implemented using Haskell's
@@
 \begin{verbatim}
-genTree(size) = if size == 0 then do { x <- genInt; return (Leaf x) }
+genTree(size) = if size == 0 
+  then do { x <- genInt; return (Leaf x) }
   else oneof [ do { x <- genInt; return (Leaf x) }
@@
 
-\noindent This definition follows a common type-directed generation approach,
-and it is a good example of what to expect from a random generator synthesized
-automatically from the codebase.
-%
-In esence, this generator randomly picks one data constructor from its
-corresponding data type definition with uniform probability (i.e, either a
-\texttt{Leaf} or a \texttt{Branch}), and then calls itself for every recursive
-subexpression needed to produce a well-typed value, carefully reducing the input
-size limit \texttt{size} by a unit on each step, ensuring termination by
-restricting itself to generating leaves when it reaches zero (case \texttt{size
-  == 0}).
+\noindent This definition (simplified to make it more accessible) follows a
+common type-directed approach, and illustrates what to expect from a generator
+synthesized automatically from the codebase.
+%
+At each step, this generator randomly picks a \texttt{Tree} data constructor
+with uniform probability (i.e, either a \texttt{Leaf} or a \texttt{Branch}), and
+calls itself for every recursive subexpression, carefully reducing the input
+size limit \texttt{size} by a unit at a time.
+%
+This enforces termination by generating only leaves when the size reaches zero
+(case \texttt{size == 0}).
 %
@@
 Readers familiar with RBPT will recognize that using \texttt{genTree} to test
-\texttt{prop\_insert\_inv} with \quickcheck directly does not work well.
+\texttt{prop\_insert\_inv} with \quickcheck directly does not work.
 %
 In practice, \quickcheck gives up if most of the generated test cases get
-discarded due to failed preconditions.
+discarded.
 %
@@
 %
-Sadly, most of the inputs produced by our na\"ive generator suffer from this
-problem, and the interesting part of the property (\texttt{balanced (insert x
-  t)}) is tested very sporadically as a result.
+Sadly, most of the test cases produced by \texttt{genTree} consist of unbalanced
+trees, thus the interesting part of the property (\texttt{balanced (insert x
+t)}) is rarely exercised.
 %% %
@@
 
-The next subsection shows how GCPT helps overcoming this limitation without
-having to write specialized generators by hand.
+% The next subsection shows how GCPT helps overcoming this limitation without
+% having to write specialized generators by hand.
 
@@
 \cite{kersten2017poster} use execution traces to recognize interesting test
-cases, e.g, those that exercise previously undiscovered parts of the target
-code.
+cases, 
+%
+% e.g, those that exercise previously undiscovered parts of the target
+% code.
 %

### 2022-03-02 papers/fse22/main.tex
Compressing Section 3

@@
 %
-Later, such tools use generic mutators to combine and produce new test cases
-from previously executed interesting ones.
+% Later, such tools use generic mutators to combine and produce new test cases
+% from previously executed interesting ones.
 %
@@
 %
-In contrast with traditional fuzzing approaches, CGPT can distinguish
-semantically valid test cases from invalid ones, i.e., those passing the testing
-property preconditions as opposed to those that are discarded early.
+In contrast with generic fuzzing approaches, CGPT can distinguish semantically
+valid test cases from invalid ones, i.e., those passing the testing property
+preconditions as opposed to those that are discarded early.
 %
@@
 
-The testing loop in CGPT uses two queues to store valid and discarded previously
+The CGPT testing loop uses two queues to store valid and discarded previously
 executed test cases.
 %
-Enqueued test cases are stored along with a given mutation budget that controls
-how many times a given test case can be mutated before being finally thrown
-away.
+Enqueued test cases are stored along with a mutation budget that controls how
+many times they can be mutated before being finally thrown away.
 %
@@
 %
-Then, on each iteration the algorithm selects the next test case by mutating the
-first value on the queue of valid test cases.
-%
-If that queue is empty, it instead mutates the first test case from the queue of
+Then, on each iteration the testing loop selects the next test case by mutating
+the first value on the queue of valid test cases.
+%
+If such queue is empty, it then mutates the first test case from the queue of
 discarded test cases.
@@
 The algorithm then run this test case and evaluates whether it was interesting
-(i.e., it exercises a new path) based on its trace information.
-%
-If the test case does in fact discover a new path, it is enqueued at the end of
-its corresponding queue, depending on whether it passed the property
-precondition or was discarded.
+based on its trace information.
+%
+If the test case was interesting, it gets enqueued into its corresponding queue
+(either valid or discarded),
+%  depending on whether it passed the property precondition or
+% was discarded.
 %
@@
 %
-This process alternates between generation and mutation until a bug is found or
-it reaches the test limit.
+This process alternates between random generation and mutation until a bug is
+found or the test limit is reached.
 
@@
 
-\citeauthor{lampropoulos2019coverage} demonstrated empirically that CGPT lies
-comfortably in the middle ground between using pure random testing with na\"ive
-automatically derived random generators and complex manually-written ones.
+\citeauthor{lampropoulos2019coverage} demonstrated that CGPT lies in the middle
+ground between using random testing with na\"ive automatically derived
+generators and complex manually-written ones.
 %
@@
 However, the authors acknowledge that certain parts of its implementation have
-room for improvement, especially when it comes to the mutator's design.
+room for improvement. 
+%, especially when it comes to the mutator's design.
 %
@@
 %
-Moreover, we replicated the IFC stack machine case study bundled with \fuzzchick
-and observed a surprising limitation:
+In addition, we observed a notable limitation after replicating the IFC stack
+machine case study bundled with \fuzzchick:
+%  and observed a notable limitation:
 %
 \emph{after repeating each experiment 30 times, \fuzzchick could only find 5 out
-  of the 20 injected bugs with a 100\% efficacy, the hardest one being found
+  of the 20 injected bugs with a 100\% effectivity, the hardest one being found
   only around 13\% of the time after an hour of testing.}
 %
-Despite this, the authors show that CGPT outperforms normal random testing
-coupled with na\"ive random generators by comparing the mean-time-to-failure
-(MTTF) against normal random property-based testing.
-%
-In turn, we believe that a more meticulous evaluation ought to take failure to find a
+Despite this, the authors show that CGPT outperforms random testing coupled with
+na\"ive generators by comparing their mean-time-to-failure (MTTF).
+%
+In turn, we argue that a meticulous evaluation ought to take failure to find a
 counterexample as an important metric when comparing PBT tools, not just mean
@@
 
-All these observations led us to consider three main aspects of CGPT that can be
-improved upon:
+In this work, we consider three aspects of CGPT to improve upon:
 %
@@
 %
-for the sake of simplicity, the mutators proposed by
-\citeauthor{lampropoulos2019coverage} are derived to follow a top-down approach.
+for simplicity, the mutators proposed by \citeauthor{lampropoulos2019coverage}
+are derived to follow a top-down approach:
 %% similar to that used by the automatically derived random generators.
 %
-This means that mutations can happen at the top-level or be recursively applied
-to an immediate subexpression of the input test case with approximately the same
-probability.
+% This means that
+mutations can happen at the top-level or be recursively applied to an immediate
+subexpression of the input test case with approximately the same probability.
 %
@@
 %
-This makes deep recursive mutations very unlikely, as their probability
+This approach makes deep recursive mutations very unlikely, as their probability
 decreases multiplicatively with each recursive call.
@@
 %
-Hence, these simple mutators can only be effectively used to transform shallow
-data structures, excluding applications that might require producing deeper
-mutations.
+Hence, these mutators can only effectively transform shallow test cases,
+excluding scenarios involving deeply nested data structures.
 %
@@
 %
-Ideally, \emph{mutations should be able to happen on every subexpression of the
-  input test case on a reasonable basis}.
+\emph{Ideally, mutations should happen on every subexpression of the input test
+  case on a reasonable basis}.
 
@@
 
-\item \emph{Enqueuing mutation candidates:}
-%
-the CGPT testing loop uses two single-ended queues for keeping valid and
-discarded interesting test cases.
+\item \emph{Scheduling:}
+%
+the CGPT testing loop uses two single-ended queues to store valid and discarded
+interesting test cases.
 %
@@
 %
-If this test case happens to have discovered a whole new portion of the target
-code, it will not be further mutated until the rest of the queue ahead of it
-gets processed.
-%
-This can limit the effectiveness of the testing loop if the queues grow more
+If this test case discovered a whole new portion of the target code, it will not
+be mutated until the rest of the queue ahead of it gets processed.
+%
+This limits the effectiveness of the testing loop whenever queues grow more
 often than they shrink, as interesting test cases can get ``buried'' at the end
-of a long queue and,
+of long queues and,
 %
@@
 %
-Ideally, \emph{we need a mechanism that prioritizes mutating test cases that
-  discover new portions of the code right away}.
+\emph{Thus, we should prioritize mutating novel test cases right away}.
 %
@@
 %
-it is not clear how the heuristic used to assign a given budget to each mutable
-test case in CGPT works in the context of high-level structure-preserving
-mutations.
-%
-If it assigns too much budget to certain not-so-interesting test cases, some
-bugs might not be discovered on a timely basis.
-%
-Conversely, assigning too little budget to interesting test cases might mean
-that some bugs cannot be discovered at all --- randomly generating the same test
-case again later does not help, as it becomes uninteresting based on historic
-trace information.
-
-To keep the comparison fair, \citeauthor{lampropoulos2019coverage} replicated
-the same power schedule configuration used in AFL, which is tailored to work in
-tandem with generic bit-level mutators.
-%
-It is unclear what the best heuristic configuration is when using a high-level
-mutation approach --- something challenging to characterize in general given the
-expressivity of the data types used to drive the mutators.
+it is unclear how well the heuristic that assigns a budget to each interesting
+test case in CGPT performs in tandem with structure-preserving mutations derived
+for an arbitrary data type.
+%
+On one hand, spending too much budget on not-so-interesting test cases might
+prevent bugs from being discovered on a timely basis.
+%
+In turn, spending too little budget on interesting test cases might prevent bugs
+from being discovered at all --- randomly generating the same initial test case
+again later does not help, as it will not be considered interesting again
+according to historic trace information.
+%
+Finding a well behaved heuristic between these two scenario can be quite
+challenging in the general case. 
+%
+\emph{Ideally, the scheduling mechanism should be as unbiased as possible.}
+
+% especially considering how expressive the data
+% types used to derive the mutators can be.  
+%
+
+% To keep the comparison fair, \citeauthor{lampropoulos2019coverage} replicated
+% the same power schedule configuration used in AFL, which is tailored to work in
+% tandem with generic bit-level mutators.
+% %
+% It is unclear what the best heuristic configuration is when using a high-level
+% mutation approach --- something challenging to characterize in general given the
+% % expressivity of the data types used to drive the mutators.
 \end{itemize}
 
-The next section introduces \mutagen, our revised CGPT that aims to tackle the
-main limitations of CGPT to make it practical in real-world scenarios.
+% We now introduce \mutagen, our revised CGPT approach designed to tackle the
+% limitations of CGPT mentioned so far.
+
+% of CGPT to make it practical in real-world scenarios.
 
@@
 In this section we describe the main ideas behind \mutagen, our CGPT tool
-written in Haskell.
-%
-Notably, these ideas extend beyond Haskell and functional programming languages
-in general.
+written in Haskell to tackle the limitations of CGPT described so far.
+%
+Notably, these ideas can be applied beyond Haskell and functional programming
+languages in general.
 %
@@
 
-In contrast to CGPT and many other popular fuzzing tools, \mutagen works by
-mutating test cases on an exhaustive and precise manner, where
+% In contrast to CGPT and many other popular fuzzing tools,
+\mutagen works by mutating test cases on an exhaustive and precise manner, where
 %
 \begin{inparaenum}
-  \item each subexpression of a mutable test case is associated with a set of
+  \item each subexpression of a test case is associated with a set of
     structure-preserving mutations, and
@@
 %
-This is inspired by exhaustive bounded testing tools like \emph{SmallCheck}
-\cite{runciman2008smallcheck} or \emph{Korat} \cite{boyapati2002korat}, that
-produce test cases exhaustively --- please refer to \S \ref{sec:related}
-for a detailed discussion.
-
-The rest of this section describes our exhaustive mutation mechanism, as well as
-the adapted testing loop used in \mutagen.
+This approach is inspired by exhaustive bounded testing tools like
+\emph{SmallCheck} \cite{runciman2008smallcheck} or \emph{Korat}
+\cite{boyapati2002korat} --- refer to \S \ref{sec:related} for a detailed
+discussion.
+%  that
+% produce test cases exhaustively ---
+
+% The rest of this section describes our exhaustive mutation mechanism, as well as
+% the adapted testing loop used in \mutagen.
 
@@
 
-Mutations in \mutagen are defined as the set of mutants that can be obtained by
-transforming the input test case \emph{at the top-level} (the root).
-%
-We define mutations for a value of type \texttt{a} (written \texttt{Mutations
-  a}) as \emph{a function that takes a value of such a type and returns a list of
-mutants} (written \texttt{Mutant a}).
-%
-In Haskell, we introduce the following type synonym:
+In \mutagen, mutations are defined as the set of mutants that can be obtained by
+transforming the input test case \emph{at the top-level} (the root node).
+%
+We define the mutations of a value of type \texttt{a} (written \texttt{Mutations
+  a}) as \emph{a function that takes a value of such a type and returns a list
+  of mutants}.
+% (written \texttt{Mutant a}).
+%
+In Haskell, we introduce the type synonym:
 %
@@
 %
-\noindent Where \texttt{Mutant}s come in two flavours, pure and random:
+\noindent As mentioned earlier, concrete mutants can be obtained either from a
+pure or a random mutation:
 
@@
 
-Pure mutants are used most of the time, and encode simple deterministic
-transformations over the outermost data constructor of the input --- recursive
-mutations will be introduced soon.
+\paragraph{Pure mutants} 
+% are used most of the time, and 
+%
+encode simple deterministic transformations over the outermost data constructor
+of the input --- recursive mutations will be introduced soon.
 %
@@
 \end{inparaenum}
-%
+
+\setlength{\intextsep}{10pt}%
+\setlength{\columnsep}{10pt}%
+\begin{wrapfigure}{r}{0.2\textwidth}
+\vspace{-10pt}
+\begin{verbatim}
+mutate t = case t of 
+  Leaf x -> [ 
+    PURE (Branch def x def)
+  ]
+  Branch l x r -> [ 
+    PURE l, PURE r, 
+    PURE (Leaf x), 
+    PURE (Branch l x l),
+    PURE (Branch r x r),
+    PURE (Branch r x l)
+  ]
+\end{verbatim}
+\vspace{-10pt}
+\caption{\label{fig:mutagen:mutator}\mutagen mutator for the \texttt{Tree} data
+  type.}
+\vspace{-10pt}
+\end{wrapfigure}
+
 To illustrate this, Fig. \ref{fig:mutagen:mutator} outlines a mutator for the
@@
 %
-Notably, this definition simply enumerates mutants that transform the outermost
+This definition simply needs to enumerate mutants that transform the outermost
 data constructor, hence no recursion is needed here.
 %
-Moreover, notice how a default value (\texttt{def}) used to fill the subtrees
-when transforming a leaf into a branch.
-%
-This value corresponds to the simplest expression we can construct for the
-mutant to be type-correct.
+Moreover, notice how a default value \texttt{def} used to fill the subtrees when
+transforming a leaf into a branch.
+%
+In practice, \texttt{def} corresponds to the simplest expression we can
+construct for the mutant to be type-correct.
 %
@@
 %% the inner \texttt{def} is the default value of the payload.%
-\footnote{We use the type class system to abstract this complexity away in our
-  implementation.}
+(Our implementation uses the type class system to abstract this complexity
+away.)
 %
@@
 the original CGPT), is also inspired by exhaustive bounded testing tools, and
-avoids introducing unnecessary randomness when ``growing'' mutated expressions.
+avoids introducing unnecessary randomness when ``growing'' a data constructor
+into another.
 
@@
 
-\begin{figure}[b]
-\begin{verbatim}
-mutate (Leaf x) =
-  [ PURE (Branch def x def) ]
-mutate (Branch l x r) =
-  [ PURE l, PURE r
-  , PURE (Leaf x)
-  , PURE (Branch l x l), PURE (Branch r x r), PURE (Branch r x l) ]
-\end{verbatim}
-\vspace{-5pt}
-\caption{\label{fig:mutagen:mutator}\mutagen mutator for the \texttt{Tree} data
-  type.}
-\vspace{-5pt}
-\end{figure}
-
 %% At this point, the reader might consider:
@@
 %
-Random mutants, on the other hand, serve as a way to \emph{selectively avoid
-  exhaustiveness} when mutating values of large enumeration types --- or any
-other type the user might want to use random mutations with.
-%
-Instead of mutating every numerical or character subexpression exhaustively, we
-sample a small number of values from a given random generator.
-%
-This amount is a tunable parameter of \mutagen.
-%
-Then, for instance, a mutator for integers becomes:
+\paragraph{Random mutants} let us \emph{selectively avoid exhaustiveness} when
+  mutating values of large enumeration types (e.g. numbers).
+  % 
+  % --- or any
+  % other type the user might want to use random mutations with.
+%
+%
+Instead of creating a new \texttt{PURE} mutant for every other numerical
+subexpression exhaustively, we condense them into a generator that can be
+repeatedly sampled to produce new random values.   
+%
+% small number of values from a
+% given random generator.
+%
+This way, a mutator for integers simply becomes:
 
@@
 
+This approach allows \mutagen to automatically control amount of effort put into
+mutating subexpressions defined using random mutations.
+%
+This can be especially useful to avoid putting unnecessary effort when the
+execution of testing property and the system under test is independent from
+their concrete values.
+
 \paragraph{Mapping top-level mutations everywhere}
@@
 %
-Now it is time to apply these mutations to every subexpression as well.
-%
-To do so, we use two utility functions that can be automatically synthesized.
-
-In first place, a function \texttt{Positions} traverses the input and builds a
-Rose tree of \emph{mutable positions}, where positions are lists of indices
+To apply these mutations to every internal subexpression we use two utility
+functions that can be automatically synthesized.
+%
+Firstly, a function \texttt{Positions} traverses the input and builds a Rose
+tree\cite{} of \emph{mutable positions}, where positions are lists of indices
 encoding the path from the root to every mutable subexpression.
@@
 
-Later, we define a function \texttt{MutateInside} that takes a desired position
+Then, we define a function \texttt{MutateInside} that takes a desired position
 within an input test case and mutates its corresponding subexpression, returning
@@
 %
-This function simply traverses the desired position, calling itself recursively
-until it reaches the desired subexpression, where a mutation encoded by \texttt{mutate} can be
-applied.
+This function traverses the desired position, calling itself recursively until
+it reaches the desired subexpression, where a mutation encoded by
+\texttt{mutate} can be applied directly.
 % locally at the top-level (case \texttt{inside [] t}).
 %
-The rest of the \texttt{MutateInside}'s functionality simply unwraps and rewraps the
-intermediate subexpressions and is not particularly relevant for the point being
-made.
+% The rest of the \texttt{MutateInside}'s functionality simply unwraps and
+% rewraps the intermediate subexpressions and is not particularly relevant for
+% the point being made.
 
@@
 
-Having the exhaustive mutation mechanism in place, we can now introduce the base
-testing loop used by \mutagen.
-%
-This is outlined in Algorithm \ref{algo:mutagen:loop}.
-%
-Like in CGPT, we use two queues, QValid and QDiscarded to keep valid and discarded mutable test cases,
-respectively.
+With the exhaustive mutation mechanism in place, we can now introduce the base
+testing loop used by \mutagen --- outlined in Algorithm \ref{algo:mutagen:loop}.
+%
+Like in CGPT, we use two queues, QValid and QDiscarded to keep valid and
+discarded mutable test cases, respectively.
 %
@@
 
-In \mutagen, we precompute all the mutations of a given mutable test case before
-enqueueing them.
-%
-These mutations are put together into lists that we call \emph{mutation batches}
---- one for each mutated test case.
+\mutagen precomputes all the mutations of a given test case before enqueueing
+them.
+%
+These mutations are put together into lists we call \emph{mutation batches} ---
+one for each mutated test case.
 %
 To initialize a mutation batch (outlined in Algorithm \ref{algo:mutagen:init}),
-we first flatten all the mutable positions of the input test case in level order
-(recall that positions are stored as a Rose tree).
-%
-Then, we iterate over all of them and retrieve all the mutants defined for each
-subexpression.
+we first flatten all the mutable positions of the input test case in level order.
+%(recall that positions are stored as a Rose tree).
+%
+Then, we iterate over all these positions, retrieving all the mutants associated
+to each corresponding subexpression.
 %
@@
   into the mutation batch directly; otherwise
-\item it is a random mutant that carries a random generator with it (e.g.,
-  corresponding to a numeric subexpression), in which case we sample and enqueue
-  \textit{R} random values using this generator, where \textit{R} is a parameter
-  set by the user.
+\item it is a random mutant that carries a random generator with it, in which
+  case we sample and enqueue \textit{R} random values using this generator,
+  where \textit{R} is a parameter set by the user.
 \end{inparaenum}
@@
 %
-Since mutations are precomputed, this function simply picks the next test case
-from the current batch, until it becomes empty and can switch to the next
-precomputed one in line.
+This function simply picks the next mutated test case from the current
+precomputed batch, jumping to the next batch in line when the current one it
+becomes empty.
 
 Having selected the next test case, the testing loop proceeds to execute it,
-capturing both the result (valid, discarded, or failed) and its execution trace
-over the system under test.
+capturing both the result (valid, discarded, or failed) and its execution trace.
 %
@@
 %
-If not, the algorithm evaluates whether it was interesting (i.e., it exercises a
-new path) based on its trace information and the one from previously executed
-test cases (represented by TLog).
-%
-If the test case discovers a new path, it is enqueued on its corresponding
-queue.
+If not, the algorithm evaluates if it was interesting based on its trace
+information and the one from previously executed test cases (represented by
+TLog).
+%
+If the test case was interesting, its mutants are precomputed and enqueued on its
+corresponding queue.
 % depending on whether it was valid or got discarded.
@@
 %
-This process alternates between generation and mutation until it finds a bug or
-reaches the test limit N.
+This process is repeated until finding a bug or reaching the test limit N.
 
@@
 
-Another difference with CGPT's testing loop is the criterion for enqueuing
+A notable difference with CGPT's testing loop is the criterion for enqueuing
 discarded tests.
@@
 candidates tends to grow disproportionately large during testing, making them
-hardly usable and consuming large amounts of memory.
+hardly usable while consuming large amounts of memory.
 %
@@
 %
-Then, we enqueue discarded test cases only if they meet this condition.
+Then, we enqueue discarded test cases only if they descend from a valid parent.
 %
@@
 
-The next section describes two heuristics we added to \mutagen's testing loop
-based on the limitations we found in CGPT.
+% The next section describes two heuristics we added to \mutagen's testing loop
+% based on the limitations we found in CGPT.
 %
@@
 In this section we introduce two heuristics implemented on top of the base
-testing loop of our tool.
+testing loop of \mutagen --- described in \S \ref{sec:mutagen}.
 %% %

### 2022-03-02 papers/fse22/main.tex
Compressing Heuristics section

@@
 
-This heuristic first tackles the issue of enqueuing new interesting mutation
-candidates at the end of (possibly) long queues of not-so-interesting previously
-executed ones.
+This first heuristic tackles the issue of enqueuing new interesting test cases
+at the end of (possibly) long queues of not-so-interesting previously executed
+ones.
 
@@
 Execution traces in \mutagen represent the specific \emph{path} in the code
-taken by the program, as opposed to just the (unordered) set of edges traversed
-in the control-flow graph (CFG) used by default by most coverage-guided fuzzers.
-%
-Using this criterion lets us gather precise information from each new execution.
+taken by the system under test, as opposed to just the unordered set of edges
+traversed in the control-flow graph used by default by most coverage-guided
+fuzzers.
+%
+Using this criterion we can gather precise information from each new execution.
 %
@@
 %
-Our assumption is that test case executions that branch at shallower depths from
-the ones already executed are more likely to discover new portions of the code
+Our assumption is that test case executions branching at shallower depths from
+already executed ones are more likely to discover new portions of the system
 under test, and hence we want to prioritize them.
@@
 
-In this light, every time we insert a new execution path into the internal trace
-log (TLog), we calculate the number of \emph{new nodes} that were executed, as
-well as the \emph{branching depth} where they got inserted.
+Every time we insert a new execution path into the internal trace log (TLog), we
+calculate the number of \emph{new nodes} that were executed, as well as the
+\emph{branching depth} where they got inserted.
 %
 The former is used to distinguish interesting test cases (whether or not new
-nodes were inserted), whereas the latter is used by this heuristic to schedule
-mutation candidates.
+nodes were inserted), whereas the latter is used by this heuristic to prioritize
+some test cases over others.
 %
@@
 
-To pick the next test case, we retrieve the one with the highest priority (the
-smallest branching depth).
+To pick the next test case, we retrieve the first one with the highest priority
+(the smallest branching depth).
 %
 Then, when we find a new interesting test case, it gets enqueued at the
-\emph{beginning} of the queue at its corresponding priority.
+\emph{beginning} of the queue of its corresponding priority.
 %
@@
 test cases as soon as they are found (even at the same priority), and to jump
-back to previous test cases as soon as their mutants become less novel.
+back to previous test cases as soon as mutants become less novel.
 
@@
 
-As introduced in \S \ref{sec:mutagen}, our tool is parameterized by the
-number of random mutations to be generated over each mutable subexpression
-defined using a random mutant, e.g., numeric values.
-%
-But, how many random mutations should we use? A single one? A few tenths? A few
-hundred?
+As introduced in \S \ref{sec:mutagen}, our tool is parameterized by the number
+of mutated test cases to be generated for each subexpression of the input test
+case associated to random mutant.
+%
+But, how many random test cases should we sample? A single one? A few tenths? A
+few hundred?
 %% %
@@
 %
-Answering this question precisely is not an easy task, and this second heuristic
-aims to tackle this issue.
-
-We found that, the smaller the number of random mutations we set, the easier it
-is for the trace log that records executions to start getting saturated, i.e.,
+Answering this question precisely can be challenging, so this second heuristic
+aims to alleviate the problem.
+
+We found that, the smaller the number of random mutations we initially set, the
+easier it is for the trace log that records executions to get saturated, i.e.,
 when interesting test cases stop getting discovered or are discovered very
@@
 %
-The process is simple:
+The process is as follows:
 %
@@
   Then,
-  \item if we have not produced any interesting test case after a certain
-    threshold (1000 tests seems to be a reasonable value in practice), we
+  \item if we have not produced any interesting test case after a certain number
+    of tests (1000 tests seems to be a reasonable threshold in practice), we
     increment the number of random mutations and the threshold by twice the
@@
     %
-    Additionally, each time this happens we also reset the trace log, so
-    interesting test cases found on a previous iteration can be found and
-    enqueued for mutation again --- this time with a higher effort dedicated to
-    producing random mutations.
+    Additionally, each time this happens we reset the trace log, so interesting
+    test cases found on a previous iteration can be found and enqueued for
+    mutation again with a higher effort dedicated to sampling mutants.
 \end{inparaenum}
 
-Notably, this heuristic can be particularly useful when the execution of the
-system under test depends heavily on some invariant encoded by a subexpression
-of its input that get transformed using a random mutant (e.g, the number of
-pixels declared by the header of an image matching the size of its actual
-payload).
+Notably, this heuristic can be useful when the execution of the system under
+test depends heavily on some invariant encoded by a subexpression of its input
+that get transformed using a random mutant (e.g, the number of pixels declared
+by the header of an image matching the size of its actual payload).
 %
@@
 increase the random mutations parameter until the random mutant is given enough
-chances to randomly generate the value that satisfies the required invariant and
+chances to randomly generate a value that satisfies the required invariant and
 makes the overall test case valid.

### 2022-03-03 papers/fse22/main.tex
Compressed Case Studies + Evaluation

@@
 %
-The first one is a simple abstract stack machine that enforces the hyperproperty
+The first one is a simple abstract stack machine that enforces
 \emph{noninterference} \cite{goguen1982security} using runtime checks.
@@
 Its implementation was proven correct by \citeauthor{10.1145/2578855.2535839}
-\citeyearpar{10.1145/2578855.2535839} in Coq, and subsequently degraded by
-systematically introducing 20 bugs on its enforcing mechanism.
+\citeyearpar{10.1145/2578855.2535839} in Coq, and later degraded by
+systematically introducing 20 bugs on its policy enforcing mechanism.
 %
@@
 %
-Here, we replicate their results and compare them against our tool.
-%
-Worth mentioning, we reimplemented this case study from scratch in Haskell in
-order to run \mutagen on its test suite.
-
-The second case study aims to evaluate \mutagen in a realistic scenario, and
-targets \textit{haskell-wasm} \cite{haskellwasm}, an existing WebAssembly engine
-of industrial strength.
-%
-%
-Sadly, comparing \mutagen with \fuzzchick on this case study has been out of the
-scope of this work, since porting \textit{haskell-wasm} into Coq, and later
-adapting our test suite to work with \fuzzchick requires a subtantial effort.
-%
-Instead, we compare \mutagen against a traditional RPBT approach, evaluating its
-effectiveness versus the relative overhead of the code instrumentation used by
-our tool.
+Here, we reproduce their results and compare them against our tool.
+%
+Worth mentioning, we mechanically translated this case study to Haskell in order
+to run \mutagen on its test suite.
+
+The second case study evaluates \mutagen in a realistic scenario, and targets
+\textit{haskell-wasm} \cite{haskellwasm}, an existing WebAssembly engine of
+industrial strength.
+%
+%
+Unfortunately, comparing \mutagen with \fuzzchick on this case study has been
+out of the scope of this work, since porting \textit{haskell-wasm} into Coq, and
+later adapting our test suite to work with \fuzzchick requires a subtantial
+effort.
+%
+Instead, we compare \mutagen against \quickcheck, evaluating its effectiveness
+versus the relative overhead of the code instrumentation used by our tool.
 
@@
 %
-The argument in the \texttt{Push} instruction represents a value to be
-inserted in the stack, whereas the argument of the \texttt{Call} instruction
-encodes the number of elements of the stack to be treated as arguments.
+The argument of the \texttt{Push} instruction represents a value to be inserted
+in the stack, whereas the argument of the \texttt{Call} instruction encodes the
+number of elements of the stack to be treated as arguments.
 %
@@
 
-This abstract machine is designed to enforce noninterference, which is based on
-the notion of \emph{indistinguishability}.
+This abstract machine enforces noninterference, a hyperproperty based on the
+notion of \emph{indistinguishability}.
 %
@@
 
-The tricky part about testing this property is to satisfy its sparse
-precondition: we need to generate two valid indistinguishable machine states in
-order to even proceed to execute the next instruction.
+The challenge about testing this property is to satisfy its sparse precondition:
+we need to generate two valid indistinguishable machine states in order to even
+proceed to execute the next instruction.
 %
@@
 %
-This table contains: the dynamic check that the machine needs to perform to
-enforce the IFC policy, along with the corresponding security label of the
-program counter and the instruction result after the instruction is executed.
+This table stores the dynamic check that the machine performs to enforce the IFC
+policy before running each instruction, along with the corresponding resulting
+security labels of the program counter and the instruction result.
 %
@@
 %
-If this condition is not met, the machine then halts as indication of a
-violation in the IFC policy.
+If this condition is not met, the machine halts to indicate a violation of the
+IFC policy.
 %
@@
 
-In this work, we are interested in using \mutagen to test the two most complex
-subsystems of \textit{haskell-wasm}: the \emph{validator} and the
-\emph{interpreter} --- both being previously tested using a unit test suite.
-
-We avoided spending countless hours writing an extensive property-based
-specification to mimic the WebAssembly specification.
-%
-Instead, we used the reference implementation to find discrepancies (that could
-potentially lead to bugs) via differential testing
+In this work, we apply \mutagen to test the two most complex subsystems of
+\textit{haskell-wasm}: the \emph{validator} and the \emph{interpreter} --- both
+being previously tested using a unit test suite.
+%
+% We avoided spending countless hours writing an extensive property-based
+% specification to mimic the WebAssembly specification.
+%
+For this, we took advantage of the reference implementation to find
+discrepancies (that could potentially lead to bugs) via differential testing
 \cite{mckeeman1998differential}.
 %
-In this manner, our testing properties assert that any result produced by
-\textit{haskell-wasm} matches that of the reference implementation.
-
-Notably, this engine had several latent bugs that were not caught by the
-existing unit tests and that we discovered by \mutagen while developing our test
-suite.
+Our testing properties assert that any result produced by \textit{haskell-wasm}
+matches that of the reference implementation.
+%
+Notably, \mutagen discovered three latent bugs 
+% in \textit{haskell-wasm} 
+that the existing test suite was unable to reveal.
+%
+% this engine had several latent bugs that were not caught by the
+% existing testsuite and that we discovered by \mutagen while developing our test
+% suite.
 %
@@
 %
-These findings are briefly outlined in Table \ref{table:wasm:bugs}.
-%
-All of them were reported and later confirmed by the authors of
-\textit{haskell-wasm}.
-%
-Having sorted these issues out, we injected 10 new bugs in the validator as well
-as 5 new bugs in the interpreter of this engine (see Table
+All these findings (outlined in Table \ref{table:wasm:bugs}) where reported and
+later confirmed by the authors of \textit{haskell-wasm}.
+%
+% All of them were reported and later confirmed by the authors of
+% \textit{haskell-wasm}.
+
+Having sorted these issues out, we mechanically injected 10 new bugs in the
+validator as well as 5 new bugs in the interpreter of this engine (see Table
 \ref{table:wasm:injectedbugs}).
 %
-These bugs were mechanically injected by either
+These bugs either
 %
 \begin{inparaenum}
-\item removing an existing integrity check (to weaken the WebAssembly
-  type-system/validator); or by
-\item simulating a copy-and-paste induced bug,%
+\item remove an existing integrity check (to weaken the WebAssembly
+  type-system/validator); or
+\item simulate a copy-paste bug \cite{CPMiner},%
   \footnote{This kind of bugs was inspired by the real bug \#1 we found prior
@@
 %% %
-To keep things simple, we assert that, whenever a randomly generated (or
+Our testing approach is to assert that, whenever a randomly generated (or
 mutated) WebAssembly module is valid according to \textit{haskell-wasm}, then
@@
 Testing the WebAssembly interpreter is substantially more complicated than
-testing the validator, since it requires actually running and comparing the
-output of test case programs.
+testing the validator, since it requires running and comparing the output of
+test case programs.
 %
@@
 %
-To simplify things, we write a helper function \texttt{mk\_module} to build a
+For simplicity, we write a helper function \texttt{mk\_module} to build a
 stub WebAssembly module that initializes one memory block and exports a single
@@
 %
-\footnote{In our implementation, we additionally set a short timeout to discard
-  potentially diverging programs with infinite loops.}
+\footnote{We additionally set a short timeout to discard potentially diverging
+  programs.}
 %
@@
 
-Using this testing property directly might not sound like a great idea, as
-randomly generated lists of input arguments will be very unlikely to match the
-type signature of randomly generated functions.
+Using this testing property directly might not sound wise, as randomly generated
+lists of input arguments will very unlikely match the type signature of randomly
+generated functions.
 %
@@
 %
-Having fixed this issue, we define a more useful specialized version of
+Having fixed this issue, we define a specialized version of
 \texttt{prop\_interpreter} that fixes the type of the generated function to take
@@
 \begin{verbatim}
-prop_interpreter_i32 fun i f = prop_interpreter
-  (FuncType { params = [I32, F32], result = [I32] }) fun [VI32 i, VF32 f]
+prop_interpreter_i32 f x y = 
+  prop_interpreter (Type {args=[I32,F32], res=[I32]}) f [VI32 x,VF32 y]
 \end{verbatim}
 
-\noindent This property lets us generate functions with a fixed type and invoke
-them with the exact number and type of arguments required.
+\noindent This property lets us generate functions of a fixed type and invoke
+them with randomly generated input of the expected types.
 %
@@
 %
-We use this property when finding all the injected bugs into the
-\textit{haskell-wasm} interpreter in the next section.
+We use this property to find all the bugs injected into \textit{haskell-wasm}'s
+interpreter in the next section.
 %
@@
 
-We performed all experiments in a dedicated workstation with 32GB of RAM and an
-Intel Core i7-8700 CPU running at 3.20GHz.
-%
-We repeated each experiment 30 times except for the ones testing the WebAssembly
-interpreter, which were run 10 times due to time constraints.
-%
-From there, we followed the approach taken by
+We performed all experiments in a workstation with 32GB of RAM and an Intel Core
+i7-8700 CPU running at 3.20GHz.
+%
+Each experiment was repeated 30 times except for the ones testing the
+WebAssembly interpreter, which were run 10 times due to time constraints.
+%
+Moreover, we followed the approach taken by
 \citeauthor{lampropoulos2019coverage} and collected the mean-time-to-failure
@@
 
-In both case studies, we show how the two \mutagen heuristics described in \S
-\ref{sec:heuristics} affect the testing performance by individually disabling
-them.
-%
-We call these variants \textit{no LIFO} and \textit{no reset}, respectively.
-%
-In the case of \textit{no reset}, the amount of random mutations is no longer
-controlled by this heuristic, so we arbitrarily fixed it to 25 random mutations
-throughout the execution of testing loop.
+In both case studies, we also show the effect of the heuristics described in \S
+\ref{sec:heuristics} by individually disabling them.
+%
+We call these variants \textit{no LIFO} and \textit{no reset}.
+%
+For \textit{no reset}, the number of times we sample random mutants (R) is no
+longer controlled by \mutagen, so we arbitrarily fixed it to R=25 throughout the
+execution of testing loop.
 
@@
 As introduced earlier, \fuzzchick can only find 5 out of the 20 bugs injected
-with a 100\% success rate within the one hour testing budget.
-%
-In turn, our tool manages to find every bug on all runs, and taking less than a
+with a 100\% failure rate within the one hour testing budget.
+%
+In turn, \mutagen manages to find every bug on all runs, taking less than a
 minute in the worst absolute case.
@@
 %
-Moreover, since the MTTF simply aggregates all runs, regardless of if they found
-a bug or timed out, this metric is sensitive to the timeout used on each
-experiment, and will inflate the results as soon as the failure rate goes below
-1.
+Moreover, since the MTTF aggregates all runs regardless of if they found a bug
+or timed out, this metric is sensitive to the timeout used on each experiment,
+and will inflate the results as soon as the failure rate goes below 1.
 %
 For this reason, comparing the failure rate between tools gives a better
-estimation of their reliability, where \mutagen shows a clear improvement.
+estimation of their effectivity, where \mutagen shows a clear improvement.
 
@@
 After carefully analyzing the results obtained using \fuzzchick, we observed a
-peculiarity.
-%
-This tool either finds bugs quickly (after a few thousands tests) or does not
+peculiarity:
+%
+this tool either finds bugs quickly (after a few thousands tests) or does not
 find them at all within the one-hour budget.
@@
 %
-With this in mind, it is fair to think that \fuzzchick might be a better choice
-if we set a shorter timeout and run it several times in a row.
-%
-Thus, to be an improvement over \fuzzchick, our tool must find bugs not only
-more reliably, but also fast!
+With this in mind, it is fair to think that \fuzzchick might still be a better
+choice if we set a shorter timeout and run it several times in a row.
+%
+Thus, to improve upon \fuzzchick, our tool must find bugs not only more
+reliably, but also fast.
 
@@
 %
-The \textit{Estimate} of this measure simply denotes the statistical difference
-between the compared distributions in a categorical manner.
+The \textit{Estimate} column denotes the statistical difference between these
+values in a categorical manner.
 %
@@
 fixed amount of random mutations (\#4, \#9, \#12, \#18, and \#10), suggesting
-that one should increase this number even further to be able to discover all
-bugs within the time budget.
-%
-In consequence, we argue that our heuristic is effective at automatically tuning
-this parameter.
+that one should increase this number even further to reliably discover all bugs
+within the time budget.
+%
+In consequence, \emph{we argue that our heuristic is effective at automatically
+tuning this parameter}.
 %
@@
 %
-The reason behind this is that this bug can be found using a very small
-counterexample, and \quickcheck prefers sampling small test cases at the
-beginning of the testing loop.
+This is because this bug can be found using a very small counterexample, and
+\quickcheck prefers sampling small test cases at the beginning of the testing
+loop.
 %
@@
 Moreover, by disabling the LIFO scheduling we observe that the performance of
-our tool becomes unstable, where bugs \#12 and \#11 take substantially less time
+\mutagen becomes unstable, where bugs \#12 and \#11 take substantially less time
 to be found, whereas bugs \#15 and \#14 cannot be found on every run. 
 %
-Finding where the best trade-off lies in cases where the heuristics cause mixed
-results like this is another interesting direction to aim our future work.
+This suggests that \emph{our LIFO scheduling heuristic can help to reliably find
+certain bugs at the cost of taking longer (on average) to find some others}.
+
+%
+% Finding where the best trade-off lies in cases where the heuristics cause mixed
+% results like this is another interesting direction to aim our future work.
 
@@
 Finally, this case study allows us to analyze the overhead introduced by the
-code instrumentation and internal processing used in \mutagen versus the
-stateless black-box approach of \quickcheck.
+custom code instrumentation and internal processing used in \mutagen versus the
+stateless black-box approach used by \quickcheck.
 %
@@
 %
-\mutagen executes tests several times slower than \quickcheck --- roughly 20x
-and 75x slower when testing \texttt{prop\_validator} and
-\texttt{prop\_interpreter\_i32}, respectively.
+Although \mutagen executes tests several times slower than \quickcheck (roughly
+20x and 75x slower when testing \texttt{prop\_validator} and
+\texttt{prop\_interpreter\_i32} respectively),
 %% %
@@
 %
-However, our tool is capable of running substantially more tests that pass the
-sparse preconditions than \quickcheck in the same amount of time, \emph{which
-  can ultimately lead us to find more bugs}.
+it runs substantially more valid tests that pass the sparse preconditions in the
+same amount of time, \emph{which can ultimately lead us to find bugs faster}.
 

### 2022-03-03 papers/fse22/main.tex
Compressed related work + conclusions

@@
 
-There exists a vast literature on fuzzing and property-based testing.
-%
-In this section we focus on three main topics:
+There exists a vast literature on fuzzing and property-based testing, so we
+focus on three main topics:
 %
@@
 
-A popular category of property-based testing tools does not rely on randomness.
-%
-Instead, all possible inputs can be enumerated and tested from smaller to larger
-up to a certain size bound.
+A different category of property-based testing tools does not rely on
+randomness.
+%
+Instead, test cases are exhaustively enumerated and tested from smaller to
+larger up to a certain size bound.
 %
@@
 %
-Later, this bijection can be traversed exhaustively or, more interestingly,
-randomly accessed.
-%
-This allows the user to easily generate random values uniformly simply by
+This bijection can be traversed exhaustively or, more interestingly, randomly
+accessed.
+%
+This allows the user to easily generate random test cases uniformly simply by
 sampling natural numbers.
 %
-However, values are enumerated based only on their type definition, so this
+However, test cases are enumerated based only on their type definition, so this
 technique is not suitable for testing properties with sparse preconditions
@@
 % SmallCheck and LazySmallCheck
-\emph{SmallCheck} \cite{runciman2008smallcheck} is a Haskell tool that follows
-this idea.
-%
-It progressively executes the testing properties against all possible input
-values up to a certain size bound.
+\emph{SmallCheck} \cite{runciman2008smallcheck} is a Haskell tool that also
+follows this approach.
+%
+It progressively executes the testing properties against all possible input test
+cases of up to a certain depth.
 %
 % Korat
-\emph{Korat} \cite{boyapati2002korat} is a Java tool that uses method
+Similarly, \emph{Korat} \cite{boyapati2002korat} is a Java tool that uses method
 specification predicates to automatically generate all non-isomorphic test cases
@@
 
-These approaches rely on pruning mechanisms to avoid populating unevaluated
-subexpresions exhaustively before the computational cost of being exhaustive
-becomes too restrictive.
+These approaches rely on pruning mechanisms to avoid generating too many
+equivalent test cases before the computational cost of their exhaustiveness
+becomes impractical.
 %
@@
 
-In this work we use exhaustiveness as a way to reliably enforce that all
-possible mutants of an interesting test case are executed.
-%
-In contrast to fully-exhaustive testing tools, \mutagen relies on randomness to
-find initial interesting mutable test cases.
-%
-In terms of pruning, it is possible to instruct \mutagen to detect unused
-subexpressions to avoid producing mutations over their corresponding positions.
-%
-This could improve the overall performance when testing non-strict properties
-that leave parts of their input unevaluated.
-%
-In our case studies, however, we observed that the properties we tested
+\mutagen uses exhaustiveness as a way to reliably enforce that all possible
+mutants of an interesting test case are scheduled.
+%
+In contrast to fully-exhaustive testing tools, our tool relies on randomly
+generated test cases as a shortcut to find initial interesting test cases
+without enumerating them exhaustively.
+%
+In terms of pruning, \mutagen can detect unused subexpressions and avoid
+producing mutations over their corresponding positions.
+%
+This can improve the overall performance when testing non-strict properties that
+leave parts of their input unevaluated.
+%
+In our case studies, however, the testing properties are quite strict:
 %% their preconditions tend to be
-are quite strict when executing test cases obtained by mutating valid existing ones,
-fully evaluating their input before executing the postcondition.
-%
-Thus, we do not forsee considerable improvements by applying lazy-prunning in
-this scenario --- gathering evidence about lazy pruning in \mutagen is as part
-of our future work.
+%are strict when executing test cases obtained by mutating valid existing ones,
+%
+their preconditions fully evaluating their inputs before executing the
+postconditions.
+%
+Thus, we did not see substantial improvements by applying lazy-prunning in this
+scenario. 
+%
+Our future work will inverstigate the effect of \mutagen's lazy-prunning against
+non-strict testing properties.
 
@@
 
-We presented \mutagen, a coverage-guided, property-based testing framework.
+We presented \mutagen, a coverage-guided, property-based testing tool that
 %written in Haskell.
@@
 %
-Our tool extends the CGPT approach
+extends the original CGPT approach
 %introduced by
@@
 with an exhaustive mutation mechanism that generates every possible mutant for
-each interesting test case, and schedules it to be tested exactly once.
+each interesting test case, scheduling them to be tested exactly once.
 
@@
 Moreover, we show how our tool can be applied in a real-world testing scenario,
-where it quickly discovers several planted and some previously unknown bugs.
-
-As indicated throughout this work, there are several directions for future work:
+where it quickly discovers 15 planted and 3 previously unknown bugs.
+
+As indicated earlier, there are several directions for future work:
 %
 combining \mutagen with specialized generators; enhancing the code
-instrumentation to be able target mutations towards specific parts of the code;
-and evaluating the effect of lazy pruning; among others.
+instrumentation to target mutations towards specific parts of the code; and
+evaluating the effect of lazy pruning; among others.
 %% %

### 2022-03-08 papers/fse22/main.tex
Added threats to validity

@@
 
-The second heuristic adjusts the number of test cases generated by random
-mutations. 
+The second heuristic adjusts the number of test cases generated from random
+mutations  
 %
@@
 %
-For this, we keep track of how often we generate interesting test cases.
+by monitoring how often we generate interesting test cases.
 %
@@
 %% mollis ultricies.
+% ----------------------------------------
+
+\section{Threats to Validity}
+\label{sec:threats}
+
+% Limited evaluation
+Our experiments cover generating test cases in two substantially different
+scenarios (valid IFC machine states and WebAssembly programs). 
+%
+Nonetheless, we cannot generalize that \mutagen will be effective at finding
+bugs in other systems or testing properties. 
+%
+However, our tool can automatically synthesize all the boilerplate code required
+to operate, becoming a drop-in replacement of \quickcheck (the most ubiquitous
+RPBT tool in Haskell). 
+%     
+This makes it an appealing alternative whenever \quickcheck is unable to
+penetrate some sparse precondition and simply gives up.
+
+% Coq -> Haskell translation
+As mentioned in \S \ref{sec:casestudies}, our evaluation required us to
+translate the IFC state machine used by \citeauthor{lampropoulos2019coverage}
+from Coq to Haskell. 
+%
+Although this translation could be done quite mechanically (this process was
+partially aided by Coq's code extraction mechanism), we cannot guarantee that
+our Haskell version of the code behaves exactly as the original one. 
+%
+% In addition, one should also consider some  
+% that the specific performance the
+% runtime systems of these two programming languages can affect, the
+% draw strong conclusions based only on time-to-failure.
 

### 2022-03-08 papers/fse22/main.tex
Added formal definition of mutations. Other fixes.

@@
 
+\newcommand\mydots{\hbox to 0.75em{.\hss.\hss.}}
+\makeatletter
+\newcommand{\vast}{\bBigg@{4}}
+\newcommand{\Vast}{\bBigg@{5}}
+\makeatother
+
 \let\oldv\verbatim
@@
 Unlike the generic bit-level mutators often used in the fuzzing community
-\cite{}, structure-preserving mutations specified at the data type level can
-effectively produce syntactically valid mutants. 
+\cite{zzuf,radamsa}, structure-preserving mutations specified at the data type
+level can effectively produce syntactically valid mutants. 
 %
@@
 %
-Such generator is then sampled a \emph{small} number of times. 
+Such generator is later sampled a \emph{relatively small} number of times. 
 %
@@
 %
-If this frequency stalls, \mutagen resets the testing loop and increases the
+Whenever this frequency stalls, \mutagen resets the testing loop and increases the
 effort put into sampling random mutations.
 %
-This automatically finds a suitable value for this parameter on the fly.
+This way, our tools automatically finds a suitable value for this parameter on
+the fly.
 
@@
 %% Last two sections
-Finally, we discuss related work in \S \ref{sec:related} and conclude
-in \S \ref{sec:conclusions}.
+We additonally present threats to validity in \S \ref{sec:threats} and discuss
+related work in \S \ref{sec:related}, to finally conclude in \S
+\ref{sec:conclusions}.
 
@@
 from being discovered at all --- randomly generating the same initial test case
-again later does not help, as it will not be considered interesting again
+again later does not help, as it will no longer be considered interesting
 according to historic trace information.
@@
 
-In \mutagen, mutations are defined as the set of mutants that can be obtained by
-transforming the input test case \emph{at the top-level} (the root node).
-%
-We define the mutations of a value of type \texttt{a} (written \texttt{Mutations
-  a}) as \emph{a function that takes a value of such a type and returns a list
-  of mutants}.
+In \mutagen, mutators are defined as the set of mutants that can be obtained by
+transforming the input test case \emph{at the top-level} (the root data
+constructor).
+%
+We define mutators of a value of type \texttt{a} (written \texttt{Mutator a}) as
+\emph{a function that takes a value of such a type and returns a list of
+mutants}.
 % (written \texttt{Mutant a}).
@@
 \begin{verbatim}
-type Mutations a = a -> [Mutant a]
+type Mutator a = a -> [Mutant a]
 \end{verbatim}
@@
 \noindent As mentioned earlier, concrete mutants can be obtained either from a
-pure or a random mutation:
+pure or a random mutation, which we define as follows:
 
@@
 
-\paragraph{Pure mutants} 
+We first focus on pure mutants, which 
 % are used most of the time, and 
 %
-encode simple deterministic transformations over the outermost data constructor
+encode deterministic transformations over the \emph{outermost} data constructor
 of the input --- recursive mutations will be introduced soon.
 %
-These transformations can:
+These transformations can either
 %
 \begin{inparaenum}
-\item return an immediate subexpression of the same type as the input;
-\item swap the outermost data constructor with any other constructor of the
-  same type, reusing existing subexpressions; and
-\item rearrange and replace fields using existing ones of the same type.
+\item return an immediate subexpression of the same type as the input, or
+\item swap the outermost data constructor with a (possibly) different one of the
+  same type, reusing the immediate subexpressions of the input in any possible
+  combination that produces a well-typed value.  
 \end{inparaenum}
@@
 \begin{wrapfigure}{r}{0.2\textwidth}
-\vspace{-10pt}
+\vspace{-20pt}
 \begin{verbatim}
@@
 
+% Defining mutations formally
+Formally, for a data type $T$ defined in terms of the data constructors $C_i$,
+each one with fields of (possibly) different types $t_j^i$:
+
+\vspace{-10pt}
+\begin{flalign*}
+T \;:=\; C_1 \; t_1^1 \; t_2^1 \; \mydots  \;|\;
+         C_2 \; t_1^2 \; t_2^2 \; \mydots  \;|\; 
+         \mydots &&
+\end{flalign*}
+\vspace{-7pt}
+
+\noindent Our tool automatically synthesizes \texttt{mutate} to follow the
+equation:
+
+\vspace{-10pt}
+\begin{flalign*}
+\texttt{mutate}(C_i \, x_1 \, x_2 \, \mydots) = 
+  \mathtt{mut_{r}}(C_i \, x_1 \, x_2 \, \mydots) \cup
+  \mathtt{mut_{s}}(C_i \, x_1 \, x_2 \, \mydots) && 
+      % \setminus
+      %  \{ (C_i \, x_1 \, x_2 \, \mydots) \} &&
+\end{flalign*}
+\vspace{-7pt}
+
+\noindent Firstly, $\mathtt{mut_r}$ computes the set of possible mutations that
+return an immediate subexpresion of the same type as the input:
+
+\vspace{-10pt}
+\begin{flalign*}
+\mathtt{mut_{r}}(C_i \, x_1 \, x_2 \, \mydots) =  
+  \{ 
+    \mathtt{PURE}\, x_k
+    \;|\;
+    x_k \!\in\! \texttt{filter}(T, \{ x_1, \, x_2, \, \mydots \})
+  \} &&
+\end{flalign*}
+\vspace{-7pt}
+
+\noindent And $\mathtt{mut_s}$ builds all the possible mutations that swap the
+root data constructor with a (possibly) different one, reusing (or defaulting)
+compatible subexpresions to instantiate the new data constructor:
+
+\vspace{-10pt}
+\begin{flalign*}
+\mathtt{mut_{s}}(C_i \, x_1 \, x_2 \, \mydots) \!=\! 
+  \Vast\{ 
+    \mathtt{PURE}(C_j \, x_1^\prime \, x_2^\prime \, \mydots)
+    \,\Vast|\,
+    \begin{aligned}
+      & C_j^T \!\in\! \{ C_1, C_2, \mydots \} \\
+      & x_k^\prime \!\in\! \overline{\mathtt{filter}}(t_k^j, \{x_1, \, x_2, \, \mydots \}) \\  
+      & C_j \, x_1^\prime \, x_2^\prime \, \mydots \;\neq\; C_i \, x_1 \, x_2 \, \mydots
+    \end{aligned}
+  \Vast\} &&
+\end{flalign*}
+\vspace{-7pt}
+
+\noindent Where the helper function $\mathtt{filter}$ simply returns the subset
+of the input values $X$ that match a target type $t$:
+
+\vspace{-10pt}
+\begin{flalign*}
+\mathtt{filter}(t, X) =  
+  \{
+    x
+    \;|\;
+    x \!\in\! X,
+    \texttt{typeof}(x) = t
+  \} &&
+\end{flalign*}
+\vspace{-7pt}
+
+\noindent And the function $\overline{\mathtt{filter}}$ is a wrapper over
+$\mathtt{filter}$ that returns the default value of the target type $t$ if there
+are no compatible subexpressions available to reuse:
+
+\vspace{-10pt}
+\begin{flalign*}
+\overline{\mathtt{filter}}(t, X) =
+\Bigg\{
+  \begin{aligned}
+    & \mathtt{filter}(t, X) 
+      &&\mathtt{if}\; \mathtt{filter}(t, X) \neq \emptyset \\
+    & \{ \mathtt{def} \}
+      &&\mathtt{otherwise}
+  \end{aligned} &&
+\end{flalign*}
+\vspace{-7pt}
+
+\noindent This ensures that a small constructor can always be grown into a
+larger one by inserting default subexpresions whenever needed.
+%
+(Our previous example defined in Fig. \ref{fig:mutagen:mutator} show this in
+practice for the case of mutating a \texttt{Leaf} into a \texttt{Branch}.)
+
+% \begin{align*}
+% \mathtt{mut_{r}}(C_i \; x_1 \; x_2 \; \mydots) = 
+%   \{ \texttt{PURE}\, x_j \, | \, 
+%   x_j \!\in\! \{ x_1, x_2, \mydots \}, \; 
+%   \texttt{typeof}(x_j) = T \}
+% \end{align*}
+
+% \begin{align*}
+% \mathtt{mut_{s}}(C_i \; x_1 \; x_2 \; \mydots) = 
+%   \{ \texttt{PURE}\, (C_j \, x_1^\prime \, x_2^\prime \, \mydots) \, | \,
+%   &C_j \!\in\! \{ C_1, C_2, \mydots \}, \, \\
+%   &x_k^\prime \!\in\! \{ x_k \, | \, x_k \!\in\! \{ x_1, x_2, \mydots \}, \mathtt{typeof}(x_k) = t_k^j \}) \}
+% \end{align*}
+
 %% \begin{figure}
@@
 %
-\paragraph{Random mutants} let us \emph{selectively avoid exhaustiveness} when
-  mutating values of large enumeration types (e.g. numbers).
+We can now focus on random mutants, which let us \emph{selectively avoid
+exhaustiveness} when mutating values of large enumeration types (e.g. numbers).
   % 
@@
 To apply these mutations to every internal subexpression we use two utility
-functions that can be automatically synthesized.
+functions.
+% that can be automatically synthesized.
 %
 Firstly, a function \texttt{Positions} traverses the input and builds a Rose
-tree\cite{} of \emph{mutable positions}, where positions are lists of indices
-encoding the path from the root to every mutable subexpression.
+tree\cite{meertens1988first} of \emph{mutable positions}, where positions are
+lists of indices encoding the path from the root to every mutable subexpression.
 %
@@
 \end{equation}
-
-Then, we define a function \texttt{MutateInside} that takes a desired position
-within an input test case and mutates its corresponding subexpression, returning
-a list of mutants.
+\vspace{-5pt}
+
+\noindent Then, we define a function \texttt{MutateInside} that takes a desired
+position within an input test case and mutates its corresponding subexpression,
+returning a list of mutants.
 %
@@
 \texttt{mutate} can be applied directly.
+%
+The implementation of these two functions consists mostly of boilerplate code
+that our tool automatically synthesizes, thus we avoid diving into details to
+preserve space.
+
 % locally at the top-level (case \texttt{inside [] t}).
@@
 
-\subsection{Detecting Trace Space Saturation And Tuning Random Mutations Parameter}
+% \subsection{Detecting Trace Space Saturation And Tuning Random Mutations Parameter}
+\subsection{Tuning Random Mutations Parameter}
 
@@
 \hline
-\textbf{Instruction}
-& \textbf{Precondition Check}
-& \textbf{Final PC Label}
-& \textbf{Final Result Label} \\
+\textbf{Instruction} & \textbf{Precondition Check} & \textbf{Final PC Label} & \textbf{Final Result Label} \\
 \hline
 \texttt{Store}
-& $l_{pc} \vee l_{p} \sqsubseteq l_{v}$
-& $l_{pc}$
-& $l_{v'} \vee l_{pc} \vee l_{p}$ \\
+& $l_{pc} \vee l_{p} \sqsubseteq l_{v}$ & $l_{pc}$ & $l_{v'} \vee l_{pc} \vee l_{p}$ \\
 \hline
@@
 removing or weakening the checks stored in this rule table.
+%
+For instance, the following constitute the buggy variations of the
+\texttt{Store} instruction:
+
+%% \vspace{-10pt}
+\begin{center}
+\scriptsize
+\begin{tabular}{|c|c|c|c|}
+\hline
+\textbf{Instruction} & \textbf{Precondition Check} & \textbf{Final PC Label} & \textbf{Final Result Label} \\
+\hline
+\texttt{Store} & $l_{pc} \sqsubseteq l_{v}$ & $l_{pc}$ & $l_{v'} \vee l_{pc} \vee l_{p}$ \\
+\texttt{Store} & $l_{p} \sqsubseteq l_{v}$ & $l_{pc}$ & $l_{v'} \vee l_{pc} \vee l_{p}$ \\
+\hline
+\texttt{Store} & $l_{pc} \vee l_{p} \sqsubseteq l_{v}$ & $\bot$ & $l_{v'} \vee l_{pc} \vee l_{p}$ \\
+\hline
+\texttt{Store} & $l_{pc} \vee l_{p} \sqsubseteq l_{v}$ & $l_{pc}$ & $l_{pc} \vee l_{p}$ \\
+\texttt{Store} & $l_{pc} \vee l_{p} \sqsubseteq l_{v}$ & $l_{pc}$ & $l_{v'} \vee l_{p}$ \\
+\texttt{Store} & $l_{pc} \vee l_{p} \sqsubseteq l_{v}$ & $l_{pc}$ & $l_{v'} \vee l_{pc}$ \\
+\hline
+\end{tabular}
+\end{center}
 
@@
 \item simulate a copy-paste bug \cite{CPMiner},%
-  \footnote{This kind of bugs was inspired by the real bug \#1 we found prior
+  \footnote{This kind of bugs was inspired by the real bug \#1 we found prior to
     this step.}
@@
 %% %
-Our testing approach is to assert that, whenever a randomly generated (or
-mutated) WebAssembly module is valid according to \textit{haskell-wasm}, then
-the reference implementation agrees upon it.
+Our approach is to assert that, whenever a randomly generated (or mutated)
+WebAssembly module is valid according to \textit{haskell-wasm}, then the
+reference implementation agrees upon it.
 %
@@
 \noindent The precondition (\texttt{isValidHaskellWasm m}) runs the input
-WebAssembly module \texttt{m} against \textit{haskell-wasm}'s validator, whereas
-the postcondition (\texttt{isValidRefImpl m}) serializes \texttt{m}, runs it
-against the reference implementation and checks that no errors are produced.
+{Web-Assembly} module \texttt{m} against \textit{haskell-wasm}'s validator,
+whereas the postcondition (\texttt{isValidRefImpl m}) serializes \texttt{m},
+runs it against the reference implementation and checks that no errors are
+produced.
 
@@
   do let m = mk_module fun ty "f"
-     resHs   <- invokeHaskellWasm m "f" args
-     resSpec <- invokeRefImpl     m "f" args
-     return (equivalent resHs resSpec)
+     resHaskellWasm <- invokeHaskellWasm m "f" args
+     resRefImpl     <- invokeRefImpl     m "f" args
+     return (resHaskellWasm === resRefImpl)
 \end{verbatim}
@@
 Thus, to improve upon \fuzzchick, our tool must find bugs not only more
-reliably, but also fast.
+reliably, but also rapidly.
+
+Table \ref{table:ifc:tests} shows a comparison between the mean number of tests
+required by both tools to find the first failure \emph{when we only consider
+  successful runs.}
+%
+Additionally, as advised by \citeauthor{arcuri2014hitchhiker}
+\cite{arcuri2014hitchhiker}, we computed the non-parametric Vargha-Delaney
+$A_{12}$ measure \cite{vargha2000critique}.
+%
+Intuitively, this measure encodes the probability of \mutagen to yield better
+results than \fuzzchick.
+%
+The \textit{Estimate} column denotes the statistical difference between these
+values in a categorical manner.
+%
+Estimates in {\color{red} red} and {\color{blue} blue} indicate results
+favoring \fuzzchick and \mutagen, respectively.
+%
+As it can be observed,
+%
+%% using an exhaustive mutation approach can be not only
+%% more reliable, but also faster.
+%
+\mutagen is likely to find bugs faster than \fuzzchick in 14 of the 20 bugs
+injected into the abstract machine, being substantially slower in only 2 cases.
+
+In terms of the \mutagen heuristics,
+%
+%% we can also arrive to some conclusions.
+%% %
+%% Firstly,
+%
+this case study is not strongly affected by using LIFO scheduling.
+%% --- the same
+%% benchmark takes only roughly 20\% longer to complete when disabling this
+%% heuristic.
+%
+We noticed that when a new interesting test case gets enqueued, all its mutants
+(and their descendants) are quickly processed before new ones start piling up,
+keeping the mutation queues empty most of the time (generation mode).
+%
+Hence, the small proportion of time this heuristic is active (when the mutation
+queues are not empty) tends to be inconsequential.
+%% under this case study.
 
@@
 %% \centering
-\vspace{-10pt}
+\vspace{-5pt}
 \scriptsize
@@
   runs.}
-\vspace{-20pt}
+\vspace{-15pt}
 %% \end{wrapfigure}
 \end{wrapfigure}
-
-Table \ref{table:ifc:tests} shows a comparison between the mean number of tests
-required by both tools to find the first failure \emph{when we only consider
-  successful runs.}
-%
-Additionally, as advised by \citeauthor{arcuri2014hitchhiker}
-\cite{arcuri2014hitchhiker}, we computed the non-parametric Vargha-Delaney
-$A_{12}$ measure \cite{vargha2000critique}.
-%
-Intuitively, this measure encodes the probability of \mutagen to yield better
-results than \fuzzchick.
-%
-The \textit{Estimate} column denotes the statistical difference between these
-values in a categorical manner.
-%
-Estimates in {\color{red} red} and {\color{blue} blue} indicate results
-favoring \fuzzchick and \mutagen, respectively.
-%
-As it can be observed,
-%
-%% using an exhaustive mutation approach can be not only
-%% more reliable, but also faster.
-%
-\mutagen is likely to find bugs faster than \fuzzchick in 14 of the 20 bugs
-injected into the abstract machine, being substantially slower in only 2 cases.
-
-In terms of the \mutagen heuristics,
-%
-%% we can also arrive to some conclusions.
-%% %
-%% Firstly,
-%
-this case study is not strongly affected by using LIFO scheduling.
-%% --- the same
-%% benchmark takes only roughly 20\% longer to complete when disabling this
-%% heuristic.
-%
-We noticed that when a new interesting test case gets enqueued, all its mutants
-(and their descendants) are quickly processed before new ones start piling up,
-keeping the mutation queues empty most of the time (generation mode).
-%
-Hence, the small proportion of time this heuristic is active (when the mutation
-queues are not empty) tends to be inconsequential.
-%% under this case study.
 
@@
 In consequence, \emph{we argue that our heuristic is effective at automatically
-tuning this parameter}.
+tuning this parameter on the fly}.
 %
@@
 Our experiments cover generating test cases in two substantially different
-scenarios (valid IFC machine states and WebAssembly programs). 
+scenarios (valid IFC machine states and WebAssembly programs), which are able to
+find several planted and real previously-unknown bugs. 
 %
@@
 %
-However, our tool can automatically synthesize all the boilerplate code required
-to operate, becoming a drop-in replacement of \quickcheck (the most ubiquitous
-RPBT tool in Haskell). 
+To compensate, our tool automatically synthesizes all the boilerplate code
+required to operate using Template Haskell \cite{SheardJ02}, becoming an almost
+drop-in replacement of \quickcheck (the most ubiquitous RPBT tool in Haskell). 
 %     
@@
 %
-Although this translation could be done quite mechanically (this process was
-partially aided by Coq's code extraction mechanism), we cannot guarantee that
-our Haskell version of the code behaves exactly as the original one. 
+Although this translation was performed mechanically (the process was partially
+aided by Coq's code extraction mechanism), we cannot formally guarantee that our
+Haskell version of the code behaves exactly as the original one. 
+%
+
+% However, we are highly confident that 
 %

### 2022-03-08 papers/fse22/main.tex
Fixed wrongly formatted citations

@@
 Generic fuzzing tools like \emph{AFL} \cite{afl}, \emph{HonggFuzz}
-\cite{honggfuzz}, or \emph{libFuzzer} \citeyearpar{libfuzzer} as well as
+\cite{honggfuzz}, or \emph{libFuzzer} \citeyear{libfuzzer} as well as
 language-specific ones like Crowbar \cite{dolan2017testing} or \emph{Kelinci}
@@
 Its implementation was proven correct by \citeauthor{10.1145/2578855.2535839}
-\citeyearpar{10.1145/2578855.2535839} in Coq, and later degraded by
+\citeyear{10.1145/2578855.2535839} in Coq, and later degraded by
 systematically introducing 20 bugs on its policy enforcing mechanism.
@@
 To preserve space, we encourage the reader to refer to the work of
-\citeauthor{hritcu2013testing} \citeyearpar{hritcu2013testing,
+\citeauthor{hritcu2013testing} \citeyear{hritcu2013testing,
   hrictcu2016testing} and \citeauthor{lampropoulos2019coverage} for more details
@@
 %
-\citeauthor{ClaessenDP14} \citeyearpar{ClaessenDP14} developed an algorithm for
+\citeauthor{ClaessenDP14} \cite{ClaessenDP14} developed an algorithm for
 generating inputs constrained by boolean preconditions with almost-uniform
@@
 %
-Later, \citeauthor{LampropoulosGHH17} \citeyearpar{LampropoulosGHH17} extended
+Later, \citeauthor{LampropoulosGHH17} \cite{LampropoulosGHH17} extended
 this approach by adding a limited form of constraint solving controllable by the
@@
 %
-Recently, \citeauthor{Lampropoulos2017} \citeyearpar{Lampropoulos2017} proposed
-a mechanism to obtain constrained random generators by defining their target
+Recently, \citeauthor{Lampropoulos2017} \cite{Lampropoulos2017} proposed a
+mechanism to obtain constrained random generators by defining their target
 structure using inductively defined relations in Coq.

### 2022-03-08 papers/fse22/main.tex
NIT

@@
 
-This approach allows \mutagen to automatically control amount of effort put into
-mutating subexpressions defined using random mutations.
-%
-This can be especially useful to avoid putting unnecessary effort when the
-execution of testing property and the system under test is independent from
-their concrete values.
+\noindent This approach allows \mutagen to selectively control the amount of
+effort put into mutating any subexpression of an input test case whose type is
+associated to a random mutation.
+%
+This can be especially useful to avoid putting unnecessary effort into mutating
+data payloads when the execution of testing property and the system under test
+is independent from their concrete values (see \S \ref{sec:heuristics}).
 

### 2022-03-09 papers/fse22/main.tex
General pass over the text

@@
 %% These commands are for a JOURNAL article.
-\acmConference[ESEC/FSE 2022]{The 30th ACM Joint European Software Engineering Conference and Symposium on the Foundations of Software Engineering}{14 - 18 November, 2022}{Singapore}
+\acmConference[ESEC/FSE 2022]{The 30th ACM Joint European Software Engineering
+Conference and Symposium on the Foundations of Software Engineering}{14 - 18
+November, 2022}{Singapore}
 %% \acmJournal{JACM}
@@
 Sadly, such techniques cannot derive generators capable of producing random
-values satisfying complex invariants, often needed to test properties with
-sparse preconditions.
+values satisfying complex invariants often needed to test properties with sparse
+preconditions.
 
@@
 %
-Our tool incorporates heuristics to help minimizing scalability issues as well
+Our tool incorporates heuristics that help to minimize scalability issues as well
 as covering the search space in a principled manner.   
@@
 %
-Our evaluation show that \mutagen not only outperforms existing CGPT tools, but
-also finds previously unknwown bugs in real-world software.
+Our evaluation shows that \mutagen not only outperforms existing CGPT tools but
+also finds previously unknown bugs in real-world software.
 \end{abstract}
@@
 % needed to instantiate the testing properties.
-However, a limiting factor of RPBT is having to provide the random data
-generators used to instantiate the testing properties. 
+However, a limiting factor of RPBT is the need for the random data generators
+used to instantiate the testing properties. 
 %
@@
 %
-Luckily, there exist several approaches to automatically synthesize random data
-generators by reifying the static information present in the codebase, e.g.,
-data type definitions, application public interfaces (APIs), etc.
+Luckily, there exist several approaches that automatically synthesize random
+data generators by reifying the static information present in the codebase,
+e.g., data type definitions, application public interfaces (APIs), etc.
 \cite{GriecoCB16, DBLP:conf/haskell/MistaRH18, Mista2019GeneratingRS,
@@
 limitation \cite{chen2020asurvey}, where developers are forced to put
-substantial efforts in writing specialized generators by hand \cite{Palka11,
+substantial efforts into writing specialized generators by hand \cite{Palka11,
 perenyi2020stack, yang2011finding}.
@@
 %
-In contast, CGPT relies on the data type information of the inputs to the
+In contrast, CGPT relies on the data type information of the inputs to the
 testing properties to derive specialized structure-preserving mutators directly
@@
 That is, given an interesting test case, our tool precomputes and schedules
-every structure-preserving mutation that can be applied to it exactly once.
+every structure-preserving mutation that can be applied to it to be evaluated
+exactly once.
 %
@@
 %
-On one hand,
+Firstly,
 %  mutations do not rely on randomness to be generated.
@@
 %
-On the other hand, scheduling mutations exhaustively eliminates the need for a
+Moreover, scheduling mutations exhaustively eliminates the need for a
 power schedule to assign a mutation budget to every interesting test case.
@@
 %
-Finding a well behaved heuristic between these two scenario can be quite
-challenging in the general case. 
+Finding a well-behaved heuristic that lies between these two scenarios can be
+quite challenging in the general case. 
 %
@@
 
-In this section we describe the main ideas behind \mutagen, our CGPT tool
-written in Haskell to tackle the limitations of CGPT described so far.
+In this section we describe the main ideas behind \mutagen, our revised CGPT
+tool written in Haskell that tackles the limitations of CGPT described so far.
 %
@@
   type.}
-\vspace{-10pt}
+\vspace{-20pt}
 \end{wrapfigure}
@@
 %
-Using a small default value, as oppossed to a randomly generated one (as done by
+Using a small default value, as opposed to a randomly generated one (as done by
 the original CGPT), is also inspired by exhaustive bounded testing tools, and
@@
 
-\noindent Our tool automatically synthesizes \texttt{mutate} to follow the
-equation:
+\noindent Our tool automatically synthesizes \texttt{mutate} by pattern matching
+on the root data constructor of the input. 
+%
+On each case, \texttt{mutate} follows:
 
@@
 
-\noindent And $\mathtt{mut_s}$ builds all the possible mutations that swap the
+\noindent Then, $\mathtt{mut_s}$ builds all the possible mutations that swap the
 root data constructor with a (possibly) different one, reusing (or defaulting)
@@
     \begin{aligned}
-      & C_j^T \!\in\! \{ C_1, C_2, \mydots \} \\
+      & C_j \!\in\! \{ C_1, C_2, \mydots \} \\
       & x_k^\prime \!\in\! \overline{\mathtt{filter}}(t_k^j, \{x_1, \, x_2, \, \mydots \}) \\  
@@
 
-\noindent Where the helper function $\mathtt{filter}$ simply returns the subset
-of the input values $X$ that match a target type $t$:
+\noindent In the previous definitions, the helper function $\mathtt{filter}$
+simply returns the subset of the input values $X$ that match a target type $t$:
 
@@
 
-\noindent And the function $\overline{\mathtt{filter}}$ is a wrapper over
+\noindent Whereas the function $\overline{\mathtt{filter}}$ is a wrapper over
 $\mathtt{filter}$ that returns the default value of the target type $t$ if there
@@
 Like in CGPT, we use two queues, QValid and QDiscarded to keep valid and
-discarded mutable test cases, respectively.
+discarded interesting test cases, respectively.
 %
@@
 
-Finally, the seed selection algorithm (Algorithm \ref{algo:mutagen:pick}) picks
+Then, the seed selection algorithm (Algorithm \ref{algo:mutagen:pick}) picks
 the next test case using the same criteria as CGPT, prioritizing valid test
@@
 
+Every time we insert a new execution path into the internal trace log (TLog), we
+calculate the number of \emph{new nodes} that were executed, as well as the
+\emph{branching depth} where they got inserted.
+%
+The former is used to distinguish interesting test cases (whether or not new
+nodes were inserted), whereas the latter is used by this heuristic to prioritize
+some test cases over others.
+%
+Fig. \ref{fig:mutagen:tracelog} illustrates this idea, inserting two execution
+traces (one after another) into a trace log that initially contains a single
+execution path.
+%
+Both executions are considered interesting, but the second one (with trace $1
+\rightarrow 2 \rightarrow 6 \rightarrow 7$) branches at a shallower depth than
+the first one (2 vs. 3), and so its corresponding test case is given a higher
+priority.
+
+Using this mechanism, we can modify \mutagen's base testing loop by replacing
+each mutation queue with a priority queue indexed by the branching depth of each
+new execution.
+%
+These changes are illustrated in Algorithm \ref{algo:mutagen:lifo}.
+%
+Statements in {\color{red} red} indicate important changes to the base
+algorithm, whereas ellipses denote unchaged parts of the code that are not
+relevant for the point being made.
+
+To pick the next test case, we retrieve the first one with the highest priority
+(the smallest branching depth).
+%
+Then, when we find a new interesting test case, it gets enqueued at the
+\emph{beginning} of the queue of its corresponding priority.
+%
+This allows the testing loop to jump immediately onto mutating new interesting
+test cases as soon as they are found (even at the same priority), and to jump
+back to previous test cases as soon as mutants become less novel.
+
+%% \caption{\label{fig:heuristics} The two main heuristics implemented in \mutagen.
+%%   Statements in {\color{red} red} indicate important changes to the base
+%%   algorithm. Ellipses denote parts of the code that are not relevant for the
+%%   point being made. }
+
 \begin{algorithm}[t]
@@
 
-Every time we insert a new execution path into the internal trace log (TLog), we
-calculate the number of \emph{new nodes} that were executed, as well as the
-\emph{branching depth} where they got inserted.
-%
-The former is used to distinguish interesting test cases (whether or not new
-nodes were inserted), whereas the latter is used by this heuristic to prioritize
-some test cases over others.
-%
-Fig. \ref{fig:mutagen:tracelog} illustrates this idea, inserting two execution
-traces (one after another) into a trace log that initially contains a single
-execution path.
-%
-The second insertion (with trace $1 \rightarrow 2 \rightarrow 6 \rightarrow 7$)
-branches at a shallower depth than the first one (2 vs. 3), and so its
-corresponding test case is given a higher priority.
-
-Using this mechanism, we can modify \mutagen's base testing loop by replacing
-each mutation queue with a priority queue indexed by the branching depth of each
-new execution.
-%
-These changes are illustrated in Algorithm \ref{algo:mutagen:lifo}.
-%
-Statements in {\color{red} red} indicate important changes to the base
-algorithm, whereas ellipses denote parts of the code that are not relevant for
-the point being made.
-
-To pick the next test case, we retrieve the first one with the highest priority
-(the smallest branching depth).
-%
-Then, when we find a new interesting test case, it gets enqueued at the
-\emph{beginning} of the queue of its corresponding priority.
-%
-This allows the testing loop to jump immediately onto mutating new interesting
-test cases as soon as they are found (even at the same priority), and to jump
-back to previous test cases as soon as mutants become less novel.
-
-%% \caption{\label{fig:heuristics} The two main heuristics implemented in \mutagen.
-%%   Statements in {\color{red} red} indicate important changes to the base
-%%   algorithm. Ellipses denote parts of the code that are not relevant for the
-%%   point being made. }
-
 % \subsection{Detecting Trace Space Saturation And Tuning Random Mutations Parameter}
@@
 of mutated test cases to be generated for each subexpression of the input test
-case associated to random mutant.
+case associated to a random mutation.
 %
@@
     test cases found on a previous iteration can be found and enqueued for
-    mutation again with a higher effort dedicated to sampling mutants.
+    mutation again with a higher effort dedicated to random mutations.
 \end{inparaenum}
@@
 test depends heavily on some invariant encoded by a subexpression of its input
-that get transformed using a random mutant (e.g, the number of pixels declared
+that get transformed using a random mutation (e.g, the number of pixels declared
 by the header of an image matching the size of its actual payload).
@@
 To preserve space, we encourage the reader to refer to the work of
-\citeauthor{hritcu2013testing} \citeyear{hritcu2013testing,
-  hrictcu2016testing} and \citeauthor{lampropoulos2019coverage} for more details
-about the implementation and semantics of this case study.
+\citeauthor{hritcu2013testing} \cite{hritcu2013testing, hrictcu2016testing} and
+\citeauthor{lampropoulos2019coverage} \cite{lampropoulos2019coverage} for more
+details about the implementation and semantics of this case study.
 
@@
 This table stores the dynamic check that the machine performs to enforce the IFC
-policy before running each instruction, along with the corresponding resulting
-security labels of the program counter and the instruction result.
+policy before running each instruction, along with the resulting security labels
+corresponding to the program counter and the instruction result.
 %
@@
   type-system/validator); or
-\item simulate a copy-paste bug \cite{CPMiner},%
+\item simulate a copy-paste bug \cite{chou2001empirical},
+  %
+  replacing the implementation of an instruction with a compatible one
+  (e.g., \texttt{i32.add} by \texttt{i32.sub}).%
   \footnote{This kind of bugs was inspired by the real bug \#1 we found prior to
     this step.}
-  %
-  replacing the implementation of an instruction with a compatible one
-  (e.g., \texttt{i32.add} by \texttt{i32.sub}).
 \end{inparaenum}
@@
 %
-Interestingly, equivalence in does not imply equality.
+Interestingly, equivalence does not imply equality.
 %
@@
 \texttt{prop\_interpreter} that fixes the type of the generated function to take
-two arguments (of type \texttt{I32} and \texttt{F32}) and return an \texttt{I32}
-as a result:
+two arguments \texttt{x} and \texttt{y} (of type \texttt{I32} and \texttt{F32},
+respectively) and return an \texttt{I32} as a result:
 
@@
 \noindent This property lets us generate functions of a fixed type and invoke
-them with randomly generated input of the expected types.
+them with randomly generated inputs of the expected types.
 %
@@
 Each experiment was repeated 30 times except for the ones testing the
-WebAssembly interpreter, which were run 10 times due to time constraints.
+{Web-Assembly} interpreter, which were run 10 times due to time constraints.
 %
@@
 %
-We used a one-hour timeout to stop the execution of both tools if they have not
+We used a one-hour timeout to stop the execution of both tools if they had not
 yet found a counterexample.
@@
 
-Moreover, disabling the trace saturation heuristic (\textit{no reset}) we
+Moreover, by disabling the trace saturation heuristic (\textit{no reset}) we
 observed two effects:
@@
 custom code instrumentation and internal processing used in \mutagen versus the
-stateless black-box approach used by \quickcheck.
+black-box approach used by \quickcheck.
 %
@@
 Our experiments cover generating test cases in two substantially different
-scenarios (valid IFC machine states and WebAssembly programs), which are able to
-find several planted and real previously-unknown bugs. 
+scenarios (valid IFC machine states and WebAssembly programs). 
+%
+There, our tool was able to find several planted and real previously-unknown
+bugs. 
 %
@@
 %     
-This makes it an appealing alternative whenever \quickcheck is unable to
-penetrate some sparse precondition and simply gives up.
+This makes it an appealing alternative to consider whenever \quickcheck is
+unable to penetrate some sparse precondition and simply gives up.
 
@@
 There exists a vast literature on fuzzing and property-based testing, so we
-focus on three main topics:
+focus on three main relevant topics:
 %

### 2022-03-11 papers/fse22/main.tex
NIT
Integrated some of Ale's feedback

@@
 %
-Sadly, such techniques cannot derive generators capable of producing random
-values satisfying complex invariants often needed to test properties with sparse
-preconditions.
+% Sadly, such techniques cannot derive generators capable of producing random
+%
+Unfortunately, such techniques suffer from severe limitations when used to
+produce random test cases satisfying complex invariants, which are often needed
+to test properties with sparse preconditions.
 
@@
 Our tool incorporates heuristics that help to minimize scalability issues as well
-as covering the search space in a principled manner.   
+as cover the search space in a principled manner.   
 %
@@
 
-% \vspace{-5pt}
+% \vspace{-10pt}
 
@@
 %
-Popularized by \quickcheck \cite{ClaessenH00}, Random Property-Based Testing
-(RPBT) is a popular technique for finding bugs
+% Popularized by \quickcheck \cite{ClaessenH00},
+%
+Random Property-Based Testing (RPBT) is a popular technique for finding bugs
 \cite{ClaessenH00,hughes2003erlang, papadakis2011proper, bulwahn2012new,
@@
 data generators by reifying the static information present in the codebase,
-e.g., data type definitions, application public interfaces (APIs), etc.
+e.g., data type definitions and application public interfaces (APIs).
 \cite{GriecoCB16, DBLP:conf/haskell/MistaRH18, Mista2019GeneratingRS,
@@
 %
-\footnote{Some of these limitations were already identified by the authors in
-their original work.}
+% \footnote{Some of these limitations were already identified by the authors in
+% their original work.}
 %
@@
 \end{inparaenum}
-%
+
 %% Moreover, we replicated the Information-Flow Control (IFC) stack machine case
@@
 %% Validation
-We validated our ideas using two case studies (described in detail in \S
-\ref{sec:casestudies}): the Information-Flow Control (IFC) stack machine used by
+We validated our ideas using two case studies which require generating
+highly-structured inputs (described in detail in \S \ref{sec:casestudies}): the
+Information-Flow Control (IFC) stack machine used by
 \citeauthor{lampropoulos2019coverage} in their original work, and an existing
-{Web-Assembly} engine of industrial strength written in Haskell.
+{Web-Assembly} engine of industrial strength written in the Haskell programming
+language.
 %
@@
 In the IFC stack machine case study, we compare \mutagen against \fuzzchick, the
-reference CGPT implementation developed in Coq \cite{lampropoulos2019coverage}.
+reference CGPT implementation developed for the Coq proof assistant
+\cite{lampropoulos2019coverage}.
 %
@@
 % more traditional RPBT approach that does not rely on code instrumentation.
-In this case study we also compare the performance versus the overhead of our
+In this case study we also evaluate the performance versus the overhead of our
 tool (and our custom code instrumentation mechanism) by comparing it against
-\quickcheck (black-box).  
+\quickcheck \cite{ClaessenH00}, the most widely used RPBT tool in Haskell.  
 %
@@
 \begin{verbatim}
-prop_insert_inv :: a -> Tree a -> Property
-prop_insert_inv x t = balanced t ==> balanced (insert x t)
+prop_insert :: a -> Tree a -> Property
+prop_insert x t = balanced t ==> balanced (insert x t)
 \end{verbatim}
@@
 
-Readers familiar with RBPT will recognize that using \texttt{genTree} to test
-\texttt{prop\_insert\_inv} with \quickcheck directly does not work.
+Readers familiar with RBPT will recognize that using \texttt{genTree} to
+generate test cases to test \texttt{prop\_insert} with \quickcheck directly does
+not work.
 %
@@
 %
-However, the authors acknowledge that certain parts of its implementation have
-room for improvement. 
+% However, the authors acknowledge that certain parts of its implementation have
+% room for improvement. 
 %, especially when it comes to the mutator's design.
@@
 %
-In addition, we observed a notable limitation after replicating the IFC stack
+% In addition, 
+However, we observed a notable limitation after replicating the IFC stack
 machine case study bundled with \fuzzchick:
@@
 counterexample as an important metric when comparing PBT tools, not just mean
-time to failure --- when one is found.
+time to failure.
+% --- when one is found.
 %
@@
 
-\section{\mutagen: No Mutant Left Behind}
+\section{\mutagen}
 \label{sec:mutagen}
@@
 %% the inner \texttt{def} is the default value of the payload.%
-(Our implementation uses the type class system to abstract this complexity
-away.)
+% (Our implementation uses the type class system to abstract this complexity
+% away.)
 %
@@
     \mathtt{PURE}(C_j \, x_1^\prime \, x_2^\prime \, \mydots)
-    \,\Vast|\,
+    \Vast|\,
     \begin{aligned}
@@
     \end{aligned}
-  \Vast\} &&
+  \!\Vast\} &&
 \end{flalign*}
@@
 \begin{figure}[b]
-%% \vspace{-15pt}
+\vspace{-5pt}
 \scriptsize
@@
 \end{align*}
-\vspace{-10pt}
+\vspace{-15pt}
 \caption{\label{fig:mutagen:tracelog}Inserting two new execution traces into the
@@
 Instead, we compare \mutagen against \quickcheck, evaluating its effectiveness
-versus the relative overhead of the code instrumentation used by our tool.
+versus the relative overhead of our custom code instrumentation.
 
@@
 These instructions are defined as:
-%
+
+\vspace{5pt}
 \begin{verbatim}
@@
 \hline
-\texttt{Store} & $l_{pc} \sqsubseteq l_{v}$ & $l_{pc}$ & $l_{v'} \vee l_{pc} \vee l_{p}$ \\
-\texttt{Store} & $l_{p} \sqsubseteq l_{v}$ & $l_{pc}$ & $l_{v'} \vee l_{pc} \vee l_{p}$ \\
-\hline
-\texttt{Store} & $l_{pc} \vee l_{p} \sqsubseteq l_{v}$ & $\bot$ & $l_{v'} \vee l_{pc} \vee l_{p}$ \\
-\hline
-\texttt{Store} & $l_{pc} \vee l_{p} \sqsubseteq l_{v}$ & $l_{pc}$ & $l_{pc} \vee l_{p}$ \\
-\texttt{Store} & $l_{pc} \vee l_{p} \sqsubseteq l_{v}$ & $l_{pc}$ & $l_{v'} \vee l_{p}$ \\
-\texttt{Store} & $l_{pc} \vee l_{p} \sqsubseteq l_{v}$ & $l_{pc}$ & $l_{v'} \vee l_{pc}$ \\
+\texttt{Store} & \color{red}{$l_{pc} \sqsubseteq l_{v}$} & $l_{pc}$ & $l_{v'} \vee l_{pc} \vee l_{p}$ \\
+\texttt{Store} & \color{red}{$l_{p} \sqsubseteq l_{v}$} & $l_{pc}$ & $l_{v'} \vee l_{pc} \vee l_{p}$ \\
+\hline
+\texttt{Store} & $l_{pc} \vee l_{p} \sqsubseteq l_{v}$ & \color{red}{$\bot$} & $l_{v'} \vee l_{pc} \vee l_{p}$ \\
+\hline
+\texttt{Store} & $l_{pc} \vee l_{p} \sqsubseteq l_{v}$ & $l_{pc}$ & \color{red}{$l_{pc} \vee l_{p}$} \\
+\texttt{Store} & $l_{pc} \vee l_{p} \sqsubseteq l_{v}$ & $l_{pc}$ & \color{red}{$l_{v'} \vee l_{p}$} \\
+\texttt{Store} & $l_{pc} \vee l_{p} \sqsubseteq l_{v}$ & $l_{pc}$ & \color{red}{$l_{v'} \vee l_{pc}$} \\
 \hline
@@
 \begin{table}[b]
+% \vspace{10pt}
 \scriptsize
@@
   \textit{haskell-wasm}.}
-\vspace{-10pt}
+% \vspace{-10pt}
 \end{table}
@@
 \begin{figure}[b]
+  \vspace{5pt}
   \centering
@@
   \includegraphics[width=\linewidth]{tikz/ifc-fr.tikz}
-  \vspace{-20pt}
+  \vspace{-15pt}
   \caption{\label{fig:results:ifc} Comparison between \fuzzchick and \mutagen
     across 20 different bugs for the IFC stack machine. }
-\vspace{-5pt}
+% \vspace{-5pt}
 \end{figure}
@@
   runs.}
-\vspace{-15pt}
+\vspace{-5pt}
 %% \end{wrapfigure}
@@
 \footnotesize
-%% \vspace{-5pt}
+\vspace{10pt}
 \begin{tabular}{|l|c|c|c|c|}
@@
 Our experiments cover generating test cases in two substantially different
-scenarios (valid IFC machine states and WebAssembly programs). 
+scenarios that require generating highly-structured inputs (valid IFC machine
+states and WebAssembly programs). 
 %
@@
 
-As indicated earlier, there are several directions for future work:
-%
-combining \mutagen with specialized generators; enhancing the code
-instrumentation to target mutations towards specific parts of the code; and
-evaluating the effect of lazy pruning; among others.
+% As indicated earlier, there are several directions for future work:
+% %
+% combining \mutagen with specialized generators; enhancing the code
+% instrumentation to target mutations towards specific parts of the code; and
+% evaluating the effect of lazy pruning; among others.
 %% %

### 2022-03-11 papers/fse22/main.tex
More small changes based on feedback

@@
 \end{inparaenum}
+%
+Overcoming these obstacles is important to make CGPT more suitable for testing
+real-world software.
 
@@
 of the input --- recursive mutations will be introduced soon.
-%
-These transformations can either
-%
-\begin{inparaenum}
-\item return an immediate subexpression of the same type as the input, or
-\item swap the outermost data constructor with a (possibly) different one of the
-  same type, reusing the immediate subexpressions of the input in any possible
-  combination that produces a well-typed value.  
-\end{inparaenum}
 
@@
 \begin{wrapfigure}{r}{0.2\textwidth}
-\vspace{-20pt}
+\vspace{-5pt}
 \begin{verbatim}
@@
   type.}
-\vspace{-20pt}
+\vspace{-10pt}
 \end{wrapfigure}
+
+%
+These transformations can either
+%
+\begin{inparaenum}
+\item return an immediate subexpression of the same type as the input, or
+\item swap the outermost data constructor with a (possibly) different one of the
+  same type, reusing the immediate subexpressions of the input in any possible
+  combination that produces a well-typed value.  
+\end{inparaenum}
 
@@
 %
-We realized that we can use this information to automatically optimize the
-number of random mutations used by our tool.
+We realized that we can use this information to dynamically adapt the number of
+random mutations used by our tool.
 %
@@
 
-\caption{\label{algo:mutagen:reset}Trace Saturation Heuristic}
+\caption{\label{algo:mutagen:reset}Adaptive Random Mutations Heuristic}
 \end{algorithm}
@@
 Later, bugs are systematically introduced in the IFC enforcing mechanism by
-removing or weakening the checks stored in this rule table.
-%
-For instance, the following constitute the buggy variations of the
-\texttt{Store} instruction:
+weakening the checks stored in this rule table.
+%
+For instance, the following are the buggy rule variations (in {\color{red} red})
+for the \texttt{Store} instruction:
 
@@
   \textit{haskell-wasm}.}
-% \vspace{-10pt}
+\vspace{-10pt}
 \end{table}
@@
 \footnotesize
-\vspace{10pt}
+% \vspace{10pt}
 \begin{tabular}{|l|c|c|c|c|}

### 2022-05-11 papers/fse22/rebuttal.md
WIP on FSE rebutal

We thank the reviewers for their quality feedback.

R1Q1: Comparing MUTAGEN with FuzzChick directly is tricky as its code is not
maintained and only provides the IFC case study shown in our paper to compare
against. Given the good results we observed after comparing it against MUTAGEN,
we focused on showing how MUTAGEN compares against QuickCheck, the reference
tool in Haskell, to (hopefully) position it as a contender in that space. Any
lack of evaluation from there is simply due to time constraints. However, we
envision MUTAGEN being effectively used for testing programs that heavily
manipulate ASTs (e.g, compilers, interpreters) -- the kind of programs Haskell
shines at.

R1Q2: These results were performed directly in Coq.

R1Q3: Yes, mutations are applied to every subexpression of the input test case.
This is implemented modularly using Haskell's type class mechanism.

R2Q1.1: To the best of our knowledge, MUTAGEN is the first CGPT tool
implementation for Haskell. Hence, we not only aim to improve against FuzzChick,
but also against QuickCheck in order to advocate its adoption in that community.

R2Q1.2: We decided not to focus on degrading our tool with random mutations to
simulate FuzzChick's behavior because this would additionally require designing
a power schedule to assign energy to test cases. As we argue in the paper, the
implementation details of such component might drastically affect the
performance of the testing loop. Moreover, MUTAGEN uses a high-level tracing
mechanism (we instrument Haskell expressions), whereas FuzzChick repurposes
OCaml's AFL bitmap-based backend to get a trace signal. The information provided
by these two tracing backends is nos directly comparable, which further
complicates porting FuzzChick's power schedule into MUTAGEN.

R2Q1.3: Refer to R1Q1.

R2Q2.1: Exhaustiveness ensures that the same mutation is evaluated exactly once,
as opposed to perhaps none or perhaps a thousand times. This can both increase
reliabilty and reduce time spent on testing the same mutated test case. Our
intuition is that random mutation might perform better when the bug area is
relatively big, whereas exhaustive mutations are better when bugs need very a
input specific pattern to get triggered.

Regarding the missing related work, we will update our discussion accordingly
with the following remarks:

* [1]:

* [2]:

* [3]:

* [4]:

* [6]:

### 2022-05-12 papers/fse22/rebuttal.md
Added more discussion to FSE rebutal

@@
-We thank the reviewers for their quality feedback.
+We thank the reviewers for their valuable feedback. We focus on addressing the
+concerns from Reviewer #1 and Reviewer #2.
 
@@
 we focused on showing how MUTAGEN compares against QuickCheck, the reference
-tool in Haskell, to (hopefully) position it as a contender in that space. Any
-lack of evaluation from there is simply due to time constraints. However, we
-envision MUTAGEN being effectively used for testing programs that heavily
-manipulate ASTs (e.g, compilers, interpreters) -- the kind of programs Haskell
-shines at.
+tool in Haskell, under a real-world scenario. Any lack of evaluation from there
+is due to time constraints. However, we argue that the WebAssembly case study is
+representative of the scenario we envision MUTAGEN being used effectively, i.e.,
+testing programs that heavily manipulate complex ASTs (e.g, compilers,
+interpreters), which are the kind of programs Haskell excels at. Other testing
+scenarios comprising non-sparse properties or with simpler input structures will
+most likely see better results using QuickCheck due to its higher test
+throughput.
 
@@
 
-R1Q3: Yes, mutations are applied to every subexpression of the input test case.
-This is implemented modularly using Haskell's type class mechanism.
+R1Q3: Yes, mutations can be applied to every subexpression of the input test
+case, regardless of its type. This is modularly implemented using Haskell's
+type-class mechanism, and every type-specific mutator can be automatically
+synthesized at compile time.
 
 R2Q1.1: To the best of our knowledge, MUTAGEN is the first CGPT tool
-implementation for Haskell. Hence, we not only aim to improve against FuzzChick,
-but also against QuickCheck in order to advocate its adoption in that community.
+implementation for Haskell. Moreover, our custom plugin-based high-level tracing
+mechanism is also novel and provides flexibility not easily achievable with
+low-level backends like `afl-gcc`, e.g., tracing only the coverage of certain
+functions/modules as opposed to the entire binary. Hence, we not only aim to
+improve against FuzzChick but also against QuickCheck to advocate for its
+adoption in the Haskell community. In addition, the comparison with QuickCheck
+gives us a baseline for MUTAGEN's overhead to improve upon.
 
@@
 a power schedule to assign energy to test cases. As we argue in the paper, the
-implementation details of such component might drastically affect the
+implementation details of such components might drastically affect the
 performance of the testing loop. Moreover, MUTAGEN uses a high-level tracing
@@
 OCaml's AFL bitmap-based backend to get a trace signal. The information provided
-by these two tracing backends is nos directly comparable, which further
+by these two tracing backends is not directly comparable, which further
 complicates porting FuzzChick's power schedule into MUTAGEN.
@@
 
-R2Q2.1: Exhaustiveness ensures that the same mutation is evaluated exactly once,
-as opposed to perhaps none or perhaps a thousand times. This can both increase
-reliabilty and reduce time spent on testing the same mutated test case. Our
+R2Q2.1: Exhaustiveness in MUTAGEN ensures that the same mutation is evaluated
+exactly once, as opposed to perhaps none or perhaps a thousand times. This can
+both increase reliability and reduce time wasted on repeated effort. Our
 intuition is that random mutation might perform better when the bug area is
@@
 
-Regarding the missing related work, we will update our discussion accordingly
-with the following remarks:
+Regarding the missing related work, we will update our discussion accordingly.
+In particular, we remark that:
 
-* [1]:
+* If we establish the similarity between grammars and values described by
+  algebraic data types, both MUTAGEN and [1] aim to cover all grammar features
+  of the input space. However, MUTAGEN uses the trace feedback to decide when to
+  grow recursive grammar nodes one step at a time, whereas [1] unfolds these
+  steps into exhaustively testing k-grams pairs of grammar constructions.
 
-* [2]:
+* In contrast to Zest, which mutates the random choices made by generators
+  producing syntactically valid inputs, MUTAGEN mutates the inputs themselves.
+  This gives us an advantage to perform certain optimizations, e.g., leveraging
+  lazy evaluation to avoid mutating the unevaluated subexpressions of a given
+  input -- this paper does not yet demonstrate this feature empirically.
 
-* [3]:
+* Similar to MUTAGEN, BeDivFuzz separates mutations into "structure-changing"
+  and "structure-preserving", which closely relate to MUTAGEN's PURE and RANDOM
+  mutant kinds, respectively. In BeDivFuzz, this distinction is used to search
+  for diverse input structures (via structure-changing mutations) and then apply
+  structure-preserving mutations to them to produce structure-equivalent
+  variants. In MUTAGEN, the distinction is used to avoid testing *every*
+  structure-equivalent variant exhaustively.
 
-* [4]:
+* While RLCheck seems to effectively avoid P4, we argue that MUTAGEN still
+  follows a valuable approach that prioritizes simplicity: instead of adding
+  reinforcement learning to dynamically tune the power schedule, our approach
+  simply removes it altogether. Simplicity is, of course, not a valid metric to
+  find bugs, but serves as a clear baseline to improve upon.
 
-* [6]:
+* The MOpt mutation scheduling optimization designed by Lye et al. is effective
+  at improving the behavior of the bit-level mutators used in traditional
+  fuzzing tools. However, we argue that bit-level mutations are still far from
+  the expressive power achieved by high-level (as well as grammar-based) mutation tools, where a single mutation can, e.g., transform an `if` statement into a `while` loop by operating directly at the AST level.

### 2022-06-07 papers/fse22/rebuttal.md
Wrapped line

@@
 maintained and only provides the IFC case study shown in our paper to compare
-against. Given the good results we observed after comparing it against MUTAGEN,
-we focused on showing how MUTAGEN compares against QuickCheck, the reference
-tool in Haskell, under a real-world scenario. Any lack of evaluation from there
-is due to time constraints. However, we argue that the WebAssembly case study is
-representative of the scenario we envision MUTAGEN being used effectively, i.e.,
-testing programs that heavily manipulate complex ASTs (e.g, compilers,
-interpreters), which are the kind of programs Haskell excels at. Other testing
-scenarios comprising non-sparse properties or with simpler input structures will
-most likely see better results using QuickCheck due to its higher test
-throughput.
+against. Given the good results we observed after comparing FuzzChick against
+MUTAGEN, we focused on showing how MUTAGEN compares against QuickCheck, the
+reference tool for property-based testing, under a real-world scenario. Any lack
+of evaluation from there is due to time constraints. However, we argue that the
+WebAssembly case study is representative of the scenario we envision MUTAGEN
+being used effectively, i.e., testing programs that heavily manipulate complex
+ASTs (e.g, compilers, interpreters), which are the kind of programs where our
+type-driven approach excels at. Other testing scenarios comprising non-sparse
+properties or with simpler input structures will most likely see better results
+using QuickCheck due to its higher test throughput.
 
@@
 R1Q3: Yes, mutations can be applied to every subexpression of the input test
-case, regardless of its type. This is modularly implemented using Haskell's
-type-class mechanism, and every type-specific mutator can be automatically
-synthesized at compile time.
+case, regardless of its type. This is modularly implemented, and every
+type-specific mutator can be automatically synthesized at compile time.
 
@@
 
-R2Q1.2: We decided not to focus on degrading our tool with random mutations to
+Ale: Here it is too nich, why would they care about the Haskell community? Can
+we rephrase it with more top-level arguments applicable to testing in general?
+
+R2Q1.2: We decided not to focus on adapting our tool with random mutations to
 simulate FuzzChick's behavior because this would additionally require designing
-a power schedule to assign energy to test cases. As we argue in the paper, the
-implementation details of such components might drastically affect the
-performance of the testing loop. Moreover, MUTAGEN uses a high-level tracing
-mechanism (we instrument Haskell expressions), whereas FuzzChick repurposes
-OCaml's AFL bitmap-based backend to get a trace signal. The information provided
-by these two tracing backends is not directly comparable, which further
-complicates porting FuzzChick's power schedule into MUTAGEN.
+a power schedule to assign energy to test cases for the comparation to be fair.
+As we argue in the paper, the implementation details of such components might
+drastically affect the performance of the testing loop. Moreover, MUTAGEN uses a
+high-level tracing mechanism (we instrument Haskell expressions), whereas
+FuzzChick repurposes OCaml's AFL bitmap-based backend to get a trace signal. The
+information provided by these two tracing backends is not directly comparable,
+which further complicates porting FuzzChick's power schedule into MUTAGEN.
 
@@
   fuzzing tools. However, we argue that bit-level mutations are still far from
-  the expressive power achieved by high-level (as well as grammar-based) mutation tools, where a single mutation can, e.g., transform an `if` statement into a `while` loop by operating directly at the AST level.
+  the expressive power achieved by high-level (as well as grammar-based)
+  mutation tools, where a single mutation can, e.g., transform an `if` statement
+  into a `while` loop by operating directly at the AST level.
